### 2022-10-04 manuscript.tex
Update on Overleaf.

% !TeX TS-program = pdfLaTeX
% !TeX encoding = UTF-8 Unicode
% !TeX spellcheck = en-US
% !BIB TS-program = bibtex
% -*- coding: UTF-8; -*-
% vim: set fenc=utf-8
% https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Readme.md
% https://www.overleaf.com/project/630dcf8012267a5bb40967dd
% https://framateam.org/spikeai/channels/ultrafastcat
% https://www.zotero.org/groups/4560566/ultrafastcat
% https://arxiv.org/abs/2205.03635
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[brainsci,review,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Review on precise spiking motifs in neurobiological and neuromorphic data}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Precise spiking motifs}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0002-9536-010X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorD}{0000-0002-3107-4788}

% Authors, for the paper (add full first names)
\Author{
Antoine Grimaldi $^{1,\dagger,\ddagger}$\orcidA{}, 
Amélie Gruel $^{2,\dagger,\ddagger}$, 
Jean Martinet $^{2,\ddagger}$ 
and Laurent Perrinet $^{1,}$\orcidD{}*}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad INT UMR 7289, Université Aix Marseille, CNRS; 27 Bd Jean Moulin, 13005 Marseille, France\\
$^{2}$ \quad SPARKS,  Côte d'Azur, CNRS, I3S; 2000 Rte des Lucioles, 06900 Sophia-Antipolis, France
}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3.} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
%A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.
Why do neurons communicate through spikes? By definition, a spike, or action potential, is a all-or-none event ---it can occur or not without further details--- at asynchronous timings, i.e.~it can occur at any continuous time - differentially to the discretized timing classically used in digital processing. In the living world, neural systems almost systematically use this so-called event-based representation, though we do not yet have a clear idea why. A better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data. It is also an emerging challenge in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal challenge in the age of the digital economy and of global warming.

The response of a biological neuron depends largely on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree. This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension. Our goal here is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines. In particular, we will focus on the hypothesis that there exists in an assembly of neurons a representation based on a set of motifs of different relative spike times. Here, we will review current litterature on the detection of such motifs in generic raster plots. It is work in progress, where anybody interested can \emph{openly} join.

This hypothesis is directly inspired by neurobiological observations in the hippocampus, and it expands the capabilities of analog representations based on the firing rate by considering a representation based on repetitions of these motifs at precise times of occurrence. A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data. We will first review some biological and theoretical evidence in neural information processing. We will then present some models for the detection of such motifs in arbitrary raster plots, synthetic, biological or artificial (notably from event-based cameras). In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree. Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}

The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions on the authors' page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX-related questions please contact latex@mdpi.com.%\endnote{This is an endnote.} % To use endnotes, please un-comment \printendnotes below (before References). Only journal Laws uses \footnote.

% The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}
\item	First bullet;
\item	Second bullet;
\item	Third bullet.
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}
\item	First item; 
\item	Second item;
\item	Third item.
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}   
\unskip

\begin{table}[H] 
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabularx}
\end{table}
\unskip

\begin{table}[H]
\caption{This is a wide table.\label{tab2}}
	\begin{adjustwidth}{-\extralength}{0cm}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\fulllength}{CCCC}
			\toprule
			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
			\midrule
			Entry 1		& Data			& Data			& Data\\
			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
			\bottomrule
		\end{tabularx}
	\end{adjustwidth}
	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
\end{table}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{linenomath}
\begin{equation}
a = 1,
\end{equation}
\end{linenomath}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

% Example of a page in landscape format (with table and table footnote).
%\startlandscape
%\begin{table}[H] %% Table in wide page
%\caption{This is a very wide table.\label{tab3}}
%	\begin{tabularx}{\textwidth}{CCCC}
%		\toprule
%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
%		\midrule
%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
%		\bottomrule
%	\end{tabularx}
%	\begin{adjustwidth}{+\extralength}{0cm}
%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
%	\end{adjustwidth}
%\end{table}
%\finishlandscape

% Example of a figure that spans the whole page width. The same concept works for tables, too.
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
\end{adjustwidth}
\caption{This is a wide figure.\label{fig2}}
\end{figure}

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}

\informedconsent{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.

Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. If the study did not report any data, you might add ``Not applicable'' here.}

\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\sampleavailability{Samples of the compounds ... are available from the authors.}

%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

\abbreviations{Abbreviations}{
The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & Linear dichroism
\end{tabular}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}
\subsection[\appendixname~\thesubsection]{}
The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.

\begin{table}[H] 
\caption{This is a table caption.\label{tab5}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabularx}
\end{table}

\section[\appendixname~\thesection]{}
All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book1}
Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
% Reference 3
\bibitem[Author3(year)]{ref-book2}
Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
% Reference 4
\bibitem[Author4(year)]{ref-unpublish}
Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
% Reference 5
\bibitem[Author5(year)]{ref-communication}
Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
% Reference 6
\bibitem[Author6(year)]{ref-proceeding}
Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of presentation. In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
% Reference 7
\bibitem[Author7(year)]{ref-thesis}
Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
% Reference 8
\bibitem[Author8(year)]{ref-url}
Title of Site. Available online: URL (accessed on Day Month Year).
\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{adjustwidth}
\end{document}

### 2022-10-04 manuscript.tex
Update on Overleaf.

@@
 % vim: set fenc=utf-8
-% https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Readme.md
-% https://www.overleaf.com/project/630dcf8012267a5bb40967dd
-% https://framateam.org/spikeai/channels/ultrafastcat
-% https://www.zotero.org/groups/4560566/ultrafastcat
-% https://arxiv.org/abs/2205.03635
+% https://www.overleaf.com/project/633bf90497b5b325a3ff692e
+% https://framateam.org/spikeai/channels/review_polychrony
+% https://www.zotero.org/groups/4562620/polychronies
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 It is also an emerging challenge in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal challenge in the age of the digital economy and of global warming.
-
+%
 The response of a biological neuron depends largely on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
@@
 It is work in progress, where anybody interested can \emph{openly} join.
-
+%
 This hypothesis is directly inspired by neurobiological observations in the hippocampus, and it expands the capabilities of analog representations based on the firing rate by considering a representation based on repetitions of these motifs at precise times of occurrence.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{section}{-1} %% Remove this when starting to work on the template.
-\section{How to Use this Template}
-
-The template details the sections that can be used in a manuscript.
-Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section).
-Please check the instructions on the authors' page of the journal to verify the correct order and names.
-For any questions, please contact the editorial office of the journal or support@mdpi.com.
-For LaTeX-related questions please contact latex@mdpi.com.%\endnote{This is an endnote.} % To use endnotes, please un-comment \printendnotes below (before References).
-Only journal Laws uses \footnote.
-
-% The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision
-
 \section{Introduction}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Patents}
-
-This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
-Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
-If the study did not report any data, you might add ``Not applicable'' here.}
+\dataavailability{
+%This works is made reproducible using the following tools.
+First the code reproducing all figures is available on \href{https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Readme.md}{GitHub}~\citep{Jeremie2022a}.
+
+%\noindent The paper is available a an \href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
+Find the associated \noindent \href{https://www.zotero.org/groups/4562620/polychronies}{zotero group} used to regroup relevant literature on the subject.
+
+}
 

### 2022-10-04 manuscript.tex
importing from manubot

@@
 %=================================================================
-\documentclass[brainsci,review,submit,pdftex,moreauthors]{Definitions/mdpi} 
+\documentclass[brainsci,review,submit,pdftex,moreauthors,draft]{Definitions/mdpi} 
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 
-The introduction should briefly place the study in a broad context and highlight why it is important.
+\hypertarget{introduction-importance-of-precise-spike-timings-in-the-brain}{%
+\section{Introduction: Importance of precise spike timings in the brain}\label{introduction-importance-of-precise-spike-timings-in-the-brain}}
+
+\hypertarget{ultra-fast-neural-codes-for-ultra-fast-vision}{%
+\subsection{Ultra fast neural codes for ultra fast vision}\label{ultra-fast-neural-codes-for-ultra-fast-vision}}
+
+Let us start our review of the state of the art on the role of dynamics in vision by presenting some surprising results that have been obtained in neuroscience.
+Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
+Its first experiment consisted in asking subjects to categorize images that do or do not contain animals {[}\protect\hyperlink{ref-G3GlupPg}{1}{]}.
+The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neuronal activity.
+These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade {[}\protect\hyperlink{ref-B9iZWoq3}{2}{]}.
+
+This fast processing also explains the surprising experiments of fast serial detection which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal.
+The performances decrease progressively as the frequency of presentation of the images increases.
+However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image {[}\protect\hyperlink{ref-EqCCVwl2}{3}{]}.
+
+This speed of the visual cortex, although surprising, is quite compatible with the latencies that are recorded at the neuro-physiological level.
+Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque {[}\protect\hyperlink{ref-Vu87UU6A}{4}{]}, see Figure {[}\ref{fig:thorpe}{]}.
+This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas {[}\protect\hyperlink{ref-15aMjRHVK}{5}{]}.
+
+\begin{figure}
+\hypertarget{fig:thorpe}{%
+\centering
+\includegraphics[width=1\textwidth,height=\textheight]{https://cdn-learn.adafruit.com/assets/assets/000/086/572/original/circuitpython_Thorpe-FabreThorpe-Neuroscience-Seeking-categories-in-the-brain-monkey-visual-latency-estimate.jpg}
+\caption{From {[}\protect\hyperlink{ref-12dWDpY61}{6}{]}}\label{fig:thorpe}
+}
+\end{figure}
+
+\hypertarget{how-timing-encodes-analogous-profile}{%
+\subsection{How timing encodes analogous profile}\label{how-timing-encodes-analogous-profile}}
+
+An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief impulses that propagate along the axons of neurons.
+These have the particularity of being essentially binary in their amplitude (that is to say, they are prototypical, all or nothing).
+An important consequence of the speed of processing is that it implies that it is carried out using only very few spikes.
+Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
+Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing {[}\protect\hyperlink{ref-zZUVkTfZ}{7}{]} but also the dependance across neurons {[}\protect\hyperlink{ref-13T2Gne9z}{8}{]}.
+
+Numerous studies demonstrated the importance of precise timing in neural population activity {[}\protect\hyperlink{ref-BNK6T5Uk}{9}{]}, efficient encoding thanks to the use of spike latencies {[}\protect\hyperlink{ref-1DpSM1kpQ}{10},\protect\hyperlink{ref-IjepoonT}{11}{]} or precise timing in the auditory system {[}\protect\hyperlink{ref-mxPMgOmW}{12},\protect\hyperlink{ref-1ECJhg2GM}{13}{]}.
+All these findings, and more {[}\protect\hyperlink{ref-WZsEyOlK}{14}{]}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
+Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons {[}\protect\hyperlink{ref-XsHCA4IZ}{15}{]}.
+
+\begin{figure}
+\hypertarget{fig:roc}{%
+\centering
+\includegraphics{images/roc.jpg}
+\caption{Latency coding.
+An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency.
+In this example, one generates at most one spike per neuron.
+From {[}\protect\hyperlink{ref-YX5wqWH4}{16}{]}}\label{fig:roc}
+}
+\end{figure}
+
+At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina {[}\protect\hyperlink{ref-IjepoonT}{11}{]}.
+Notably one can appreciate in figure 1 that the response of ganglion cells to visual gratings that are flashed onto the retina.
+The authors showed that the neuronal response could be encoded in the latency of the response and not only in the frequency of discharge as is often assumed.
+In figure 4 of the same article, these results are extended to natural images and show a qualitatively similar behavior.
+The conclusion of the authors is that the discharge latency of the neurons allows to encode spatial characteristics of the image.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account {[}\protect\hyperlink{ref-1H1Omh73g}{17}{]}.
+Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway {[}\protect\hyperlink{ref-UlbjMU0W}{18}{]}.
+
+Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area {[}\protect\hyperlink{ref-513xqzmL}{19}{]}.
+Many models have used these properties in temporal coding to build fast image categorization networks.
+These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization {[}\protect\hyperlink{ref-1DpSM1kpQ}{10}{]}.
+This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes {[}\protect\hyperlink{ref-7K80jcx9}{20},\protect\hyperlink{ref-fRudGmR1}{21}{]}.
+This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events {[}\protect\hyperlink{ref-Oet1bgdb}{22}{]}.
+This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation {[}\protect\hyperlink{ref-15cR83gqd}{23}{]}.
+Note also that timing is not entirely sensorial or internal but in {[}\protect\hyperlink{ref-1E8gkynCm}{24}{]}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information'' {[}\protect\hyperlink{ref-QK8RfvTZ}{25}{]}.
+
+\hypertarget{from-synfire-chains}{%
+\subsection{From synfire chains\ldots{}}\label{from-synfire-chains}}
+
+The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
+For instance, the firing rate is highly irregular {[}\protect\hyperlink{ref-sYHyiGsP}{26}{]}, which makes it inconsistent with temporal integration of random EPSPs.
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation {[}\protect\hyperlink{ref-XsHCA4IZ}{15}{]} (see Figure {[}\ref{fig:mainen}{]}).
+When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
+This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in {[}\protect\hyperlink{ref-1AbvCTG03}{27}{]} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
+However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
+
+\begin{figure}
+\hypertarget{fig:mainen}{%
+\centering
+\includegraphics{https://camo.githubusercontent.com/77f4a48a0fe8f98ffab3c1f21dca1e74286b14d47c17120ee9627f3351fcd5e7/687474703a2f2f692e737461636b2e696d6775722e636f6d2f69786e727a2e706e67}
+\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
+While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} nois at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
+}
+\end{figure}
+
+In his book ``Corticonics'' {[}\protect\hyperlink{ref-19MS6Bbwp}{28}{]}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
+The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models.
+While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code {[}\protect\hyperlink{ref-lT1Ato7P}{29}{]}.
+The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion.
+Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons {[}\protect\hyperlink{ref-R2E2wRFg}{32}{]}.
+
+It was shown that a simple model may allow the propagation of such synfire chains {[}\protect\hyperlink{ref-xbqYi2PD}{33}{]}.
+This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size.
+Each neuron of one group is connected by an excitatory synapse to the next.
+When a pulse is elicited in the first group, this may generate a spike in the next group.
+Depending on the weight value, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse).
+Recursively applying this to a sequence of groups generates either a synfire propagation or not.
+A simple simulation is shown in Figure {[}\ref{fig:diesman}{]}.
+A crucial aspect of this emergence is in particular the balance between excitation and inhibition {[}\protect\hyperlink{ref-DL2NpHxr}{34}{]}.
+This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex {[}\protect\hyperlink{ref-mszJ8wdv}{35}{]}.
+
+\begin{figure}
+\hypertarget{fig:diesman}{%
+\centering
+\includegraphics{https://brian2.readthedocs.io/en/stable/_images/frompapers.Diesmann_et_al_1999.1.png}
+\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
+The model consists of 10 groups of 100 neurons each.
+A pulse with a given jitter is generated in the first group which here generates a pulse after a certain processing delay in the second group with a lower jitter.
+This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
+}
+\end{figure}
+
+Attempts have been made to detect such synfire chains in neurobiological data {[}\protect\hyperlink{ref-qVu1OkoS}{36}{]}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies {[}\protect\hyperlink{ref-xBZqHOuN}{37}{]} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation {[}\protect\hyperlink{ref-p7xk5w4f}{38}{]}.
+In particular, this was implemented using the pyNN language {[}\protect\hyperlink{ref-33SRV9MB}{39}{]} both in CPU-based and neuromorphic hardware {[}\protect\hyperlink{ref-DiIieN8s}{40}{]} (see Figure {[}\ref{fig:pynn}{]}).
+
+\begin{figure}
+\hypertarget{fig:pynn}{%
+\centering
+\includegraphics{https://www.frontiersin.org/files/Articles/39056/fnins-07-00011-r2/image_m/fnins-07-00011-g004.jpg}
+\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
+The background is only utilized in the original model, where it is implemented as random Gaussian current.
+For the presented hardware implementation it has been omitted due to network size constraints.
+As compensation for missing background stimuli, the resting potential was increased to ensure a comparable excitability of the neurons.
+The state space gives the duration of synfire chains as a function of the protocol's parameters and compares (C) CPU to (D) neuromorphic implementations.}\label{fig:pynn}
+}
+\end{figure}
+
+Note also that synchronicity may explain some unintuitive results.
+Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
+However, this pathway is sufficient to drive the cortex as this input is more often synchronously active {[}\protect\hyperlink{ref-18ZU9zP9l}{41}{]}.
+
+Some experimental results show the emergence of synchrony, for instance in motor cortical function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}.
+Interestingly, authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task {[}\protect\hyperlink{ref-1HJjmmSpm}{43}{]} and showed some tuning to to movement direction and reaction time {[}\protect\hyperlink{ref-FZ9X31Qa}{44}{]}.
+It is important to note that synchronous events tended to lock to LFP beta waves {[}\protect\hyperlink{ref-Ejy1sAyh}{45}{]}, and was extended to larger assemblies {[}\protect\hyperlink{ref-Glu4NlLC}{47}{]} using statistical methods (see {[}\protect\hyperlink{ref-sec:detection}{\textbf{sec:detection?}}{]}).
+
+In {[}\protect\hyperlink{ref-3GlNNxH5}{48}{]}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
+It relies on the simple remark that a series of repeated coincidences is in itself an invariant.
+Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, the pitch of a musical note.
+
+\hypertarget{to-traveling-waves}{%
+\subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}}
+
+TODO : {[}\protect\hyperlink{ref-zNDCo9Hy}{49}{]}
+
+Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
+This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses {[}\protect\hyperlink{ref-Piao2Wzf}{50}{]}.
+This show precise spatio-temporal differences in this recurrently connected assembly.
+The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
+A similar dataset used in {[}\protect\hyperlink{ref-2AsQOOsE}{52}{]} is available from M.
+Berry's lab {[}\protect\hyperlink{ref-1FU0Y7Erc}{53}{]}.
+
+Propagating waves occur in many excitable media and were recently found in neural systems from retina to neocortex.
+While propagating waves are clearly present under anaesthesia, whether they also appear during awake and conscious states remains unclear.
+One possibility is that these waves are systematically missed in trial-averaged data, due to variability.
+A recent work {[}\protect\hyperlink{ref-hGYLURVS}{54}{]} presents a method for detecting propagating waves in noisy multichannel recordings.
+Applying this method to single-trial voltage-sensitive dye imaging data, they show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
+A network model suggests that this reliability is the hallmark of the horizontal fibre network of superficial cortical layers.
+Propagating waves with similar properties occur independently in secondary visual cortex, but maintain precise phase relations with the waves in primary visual cortex.
+These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatiotemporal frame for further neuronal interactions.
+
+\begin{figure}
+\hypertarget{fig:davis}{%
+\centering
+\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
+\caption{a Spike raster for repeated presentations (N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
+A single-trial LFP trace is plotted in gray, and the average LFP response is plotted in black.
+The relative power between baseline (−200 ms to stimulus-onset) and evoked fluctuations (stimulus-onset + 50--250 ms) significantly favored the evoked response (right panel; N = 110 trials; median = 1.89 dB, p = 0.000019, two-tailed Wilcoxon's ranked-sum test).
+b Same as in (a), but for a low contrast stimulus (\textless2\% Michelson contrast).
+The relative power between baseline and evoked LFP fluctuations was not statistically different from parity (median = 1.23 dB, p = 0.087 two-tailed Wilcoxon's ranked-sum test).
+c An example of spontaneous LFP fluctuations structured as a traveling wave recorded from a spatially distributed multielectrode array in marmoset area MT.
+d Histogram of spontaneous spike probability as a function of the generalized phase of the LFP during fixation.
+e The average evoked response to low contrast stimuli was stronger when a more excitable phase (±π rad) of a spontaneous traveling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test).
+Data for panels c--e modified from Davis et al.17 with permission.}\label{fig:davis}
+}
+\end{figure}
+
+More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex {[}\protect\hyperlink{ref-vsKvZJ82}{55}{]}.
+Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
+In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states {[}\protect\hyperlink{ref-BNK6T5Uk}{9}{]}.
+Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise.
+However, fluctuations of spontaneous activity, often organized as traveling waves, shape stimulus-evoked responses and perceptual sensitivity.
+The mechanisms underlying these waves are unknown.
+Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings.
+Here, authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
+They find that spontaneous traveling waves are a general property of these networks.
+The traveling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave.
+Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
+Further, by modulating local network state, they can shape responses to incoming inputs as observed in vivo.
+
+{[}\protect\hyperlink{ref-17EU8fPBv}{56}{]} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+
+\hypertarget{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}{%
+\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}}
+
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas {[}\protect\hyperlink{ref-yLVM9IXi}{57}{]}.
+It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
+Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials.
+Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.
+
+{[}\protect\hyperlink{ref-OAorhMRa}{58}{]}
+
+\hypertarget{todo-and-to-synfire-braids-and-polychronous-groups}{%
+\subsubsection{TODO \ldots{} and to synfire braids and polychronous groups}\label{todo-and-to-synfire-braids-and-polychronous-groups}}
+
+\begin{itemize}
+\tightlist
+\item
+  {[}\protect\hyperlink{ref-HjDb1gnq}{59}{]} : from synfire chains to Synfire braids
+\end{itemize}
+
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays {[}\protect\hyperlink{ref-KS4oxnst}{60},\protect\hyperlink{ref-ga3ZTFC7}{61}{]}.
+Notably, Izhikevich {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
+This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
+
+sparse in time and space {[}2{]} AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. {[}3{]} CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
+
+\begin{itemize}
+\item
+  recent theories of binding by synchrony : Fries 2005 trends cog neuro with spikes arriving at peak susceptibility (top of a cycle) or down, van Rullen, Laura Dugué
+\item
+  cell assembly hypothesis: neurons coordinate their activity through the formation and repetitive co-activation of groups.
+https://link-springer-com.insb.bib.cnrs.fr/article/10.1007/s10827-021-00801-9 Hebb DO.
+The organisation of behaviour: a neuropsychological theory.
+New York: Science Editions; 1949.
+\item
+  A notable exception is the polychronization model of Izhikevich {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
+\end{itemize}
+
+\hypertarget{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}{%
+\subsection{Are there precise heterosynaptic spiking motifs in the brain?}\label{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}}
+
+Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
+Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains {[}\protect\hyperlink{ref-Mf6h1KNY}{63}{]}, packets {[}\protect\hyperlink{ref-2GovoaYw}{64}{]} or hippocampal sequences {[}\protect\hyperlink{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou}{66}{]} and on the role of such precise spike timing in downstream information transfer and coding {[}\protect\hyperlink{ref-19fAM6hcv}{67},\protect\hyperlink{ref-13dQ0R4UM}{68},\protect\hyperlink{ref-MfmcMsqA}{69}{]}.
+This is for instance relevant in sensory pathways in vision {[}\protect\hyperlink{ref-1EHZgcKyL}{70}{]}, audition {[}\protect\hyperlink{ref-FlKJw2VL}{71}{]}, olfaction {[}\protect\hyperlink{ref-AhSRXBJ2}{72}{]}{]} or touch {[}\protect\hyperlink{ref-gDinei4a}{73}{]}.
+In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
+In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs (Izhikevich, 2006) which he defined as polychronous groups.
+
+Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space {[}\protect\hyperlink{ref-13dQ0R4UM}{68}{]}.
+These sequences can be internally generated {[}\protect\hyperlink{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou}{66}{]} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
+Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from {[}\protect\hyperlink{ref-qI2HUEyy}{74}{]} below).
+As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
+
+\begin{figure}
+\hypertarget{fig:haimerl}{%
+\centering
+\includegraphics{https://www.pnas.org/cms/10.1073/pnas.1718518116/asset/ab9a48f6-b28f-40c1-978e-6f6dbec39b98/assets/graphic/pnas.1718518116fig01.jpeg}
+\caption{Mixed distance and duration representation in CA1.
+(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
+Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
+The black line represents speed of the mouse.}\label{fig:haimerl}
+}
+\end{figure}
+
+TODO: read {[}\protect\hyperlink{ref-MfmcMsqA}{69}{]} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
+
+In {[}\protect\hyperlink{ref-4D8Jnhkz}{75}{]}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
+They develop a HMM model for quantifying the transitions.
+``In this study, van Kempen et al.~show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
+Top-down attention drives interareal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.''
+
+\hypertarget{cortical-songs}{%
+\subsection{Cortical songs}\label{cortical-songs}}
+
+\begin{itemize}
+\tightlist
+\item
+  Ikegaya and colleagues {[}\protect\hyperlink{ref-Mf6h1KNY}{63}{]} worked on spontaneous activity in vitro and in vivo.
+They demonstrated that in cortical activity, we can find a repetition of several motifs.
+In PSCs, but also in spike activity.
+These sequences repeat after minutes and have a precise spatio temporal structure with a ms precision.
+They can be specific of a particular layer or colomn, are synchronized with network activity oscillation and can involve several cells.
+They also demonstated that these sequences can form supersequences : the cortical songs.
+It consist of the assembly of several sequences which repeat in a specific order with a compressed timing.
+\item
+  ``We find precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons in vivo and in vitro. These patterns repeat after minutes, maintaining millisecond accuracy.''
+\item
+  Precisely repeating motifs of spontaneous synaptic activity in slices: duration around 1s +/- .5 s.
+Some events in motifs are of similar size but sometimes absent - better described by Bernouilli than SE (and covariance)
+\item
+  \emph{in vivo} spontaneous activity also reveals repeating sequences. About 3000 sequences, each involving 3-10 cells out of about 900, and last up to 3 seconds
+\item
+  topography: ``Sequences had specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster (Fig.
+4, A and B, and fig.~S3B).
+(\ldots) Therefore, repeating temporal patterns of activation (\ldots) were associated with a structured spatial organization of the neurons that formed them.''
+\item
+  ``Cortical songs: modular assemblies of repeated sequences'': hierarchical detection.
+\item
+  in cotical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
+Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
+\end{itemize}
+
+\begin{figure}
+\hypertarget{fig:Ikegaya2004}{%
+\centering
+\includegraphics{images/Ikegaya2004zse0150424620001.jpeg}
+\caption{Fig.
+1.
+from {[}\protect\hyperlink{ref-Mf6h1KNY}{63}{]} Repeated motifs of spontaneous synaptic activity in vitro and in vivo.
+(A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
+Panels show segments (red) of the same voltage-clamp recording from the same cell repeating seconds or minutes after the initial occurrence (blue).
+Arrows indicate timings of repeated PSCs.
+(i) Upper trace: low--temporal resolution display of spontaneous activity of a neuron.
+Lower traces: higher resolution display of the repeated motif at indicated regions of the trace (a to c).
+(ii) Example of a longer motif.
+(B) Three repetitions of a motif.
+The top traces show the motifs superimposed on each other (blue, green, and red), the middle traces show these same traces individually, and the bottom traces show temporally magnified regions of the motifs (a to c).
+(C) Repeated sequences of intracellular current-clamp recordings in vivo.
+Two (i) and three (ii) repetitions of motifs are shown.
+Shuffle tests were performed on traces (i), a to c, yielding significantly fewer repeats (fig.~S2, P \textless{} 0.02).
+In (i), the blue trace is shifted --2.75 mV; in (ii), the blue trace is shifted --1.58 mV, and the green +0.79 mV.}\label{fig:Ikegaya2004}
+}
+\end{figure}
+
+It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in {[}\protect\hyperlink{ref-Jzr1gwsr}{76}{]} which uses noise patterns.
+They " used random waveforms to probe the formation of new memories for arbitrary complex sounds.
+A behavioral measure was designed, based on the detection of repetitions embedded in noises up to 4 s long." The task is to detect the repetition of the same (frozen) noise within a trial.
+" Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block." they showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case.
+Note that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``Time reversal had no significant effect on the RefRN advantage'' (quite surprising).
+The Learning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).
+
+Such observations suggest that neuronal groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity {[}\protect\hyperlink{ref-UGHHptIH}{77}{]}.
+This work shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+It proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
+Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions (Fig.
+7).
+
+\hypertarget{outline}{%
+\subsection{outline}\label{outline}}
+
+A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs) {[}\protect\hyperlink{ref-DUpFdm8C}{78}{]}.
+The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient.
+This approach is quite successful, and SNNs approach in some case the performance of Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
+However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
+There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+Our approach will be to focus on the temporal representation of information directly.
+In particular, our objective is to fully exploit the capacity of spiking neurons to detect synchronous patterns.
+
+{[}
+\textbf{TODO: a paragraph on our existing work}
+Our approach would be distinct than these approaches from us and colleagues as we will directly deal with delays in the system at the presynaptic level.
+
+there are temporal delays in the nervous system, both at the neural {[}\protect\hyperlink{ref-f7gkBktf}{79}{]} and behavioral {[}\protect\hyperlink{ref-1BSbXPPDF}{80}{]} levels.
+Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
+
+may be used for motion detection and interpolation {[}\protect\hyperlink{ref-WGEaBYgy}{81}{]} {[}\protect\hyperlink{ref-7GEdzh5o}{82}{]}
+
+{]}\{.banner .lightred\}
+
+Remarkably, novel neuromorphic chips use a representation similar to that of real neurons {[}\protect\hyperlink{ref-Oet1bgdb}{22}{]}.
+For example, event-based cameras provide a stream of binary asynchronous events signaling detectable changes in luminance, and information is represented by these spike-based temporal motifs, hence their name ``silicon retinas'' (see Figure \ref{fig:silicon_retina}).
+For such devices, it is crucial to better understand the potential of using such event-based representations in order to devise novel algorithms.
+
+\begin{figure}
+\hypertarget{fig:silicon_retina}{%
+\centering
+\includegraphics{https://laurentperrinet.github.io/grant/anr-anr/event_driven_computations.png}
+\caption{A miniature, event-based ATIS sensor.
+Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).}\label{fig:silicon_retina}
+}
+\end{figure}
+
+This section has provided evidence that polychronous groups are an important apsect of information representation in biology with important application in data analysis and neuromorphic engineering.
+The rest of this review paper is organized as follows.
+
+First, we will review models for the detection of polychronous groups:
+
+\begin{itemize}
+\item
+  A crucial advantage of Spiking Neural Networks (SNNs) architectures lies in its processing of temporal information.
+Yet, most SNNs encode the temporal signal as an analog signal and try to ``cross-compile'' classical Neural Network to a spiking architecture.
+To go beyond the state-of-the-art, we will review here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized.
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous`` events, that is, volleys of spikes which are stable up to certain synaptic delays.
+These models will be compared in light of neuroscientific and computational perspectives.
+We review theoretical and computational foundations of PG detection in models.
+\item
+  Application to Image processing using sparse spiking representations: Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
+design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
+\end{itemize}
+
+Finally we will discuss future avenues for effective PG detection and learning in event streams.
+
+\hypertarget{theoretical-models-for-the-detection-of-precise-heterosynaptic-spiking-motifs}{%
+\section{Theoretical models for the detection of precise heterosynaptic spiking motifs}\label{theoretical-models-for-the-detection-of-precise-heterosynaptic-spiking-motifs}}
+
+\#sec:detection
+
+A popular model for the detection of latency patterns is the tempotron {[}\protect\hyperlink{ref-1GazUBEdT}{83}{]}.
+This model is in particular reviewed in {[}\protect\hyperlink{ref-QCyjy8wE}{84}{]}.
+The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
+This aims at extending the perceptron which does not incorporate a spike timing framework.
+The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
+It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons {[}\protect\hyperlink{ref-EF6Larp8}{85}{]}.
+The limits of this model is that it's output is only binary and that it's storage capacity are limited.
+
+Some other models of latency motifs detection using STDP learning rules.
+For instance, {[}\protect\hyperlink{ref-Dd3kSDRW}{86}{]} implements a STDP-based spiking deep convolutional neural networks for object recognition; {[}\protect\hyperlink{ref-vePjgmRh}{87}{]} develops a form of spike-based, competitive learning applied for unsupervised learning.
+In {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+
+In the following of this section, we will review theoretical models which directly tackle the problem of using precise heterosynaptic spiking motifs.
+
+\hypertarget{izhikevitchs-polychronization-model}{%
+\subsection{Izhikevitch's polychronization model}\label{izhikevitchs-polychronization-model}}
+
+Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
+However, taking axonal propagation times into account, synchronous signals may not coincide significantly at the post-synaptic neuron contrary to asynchronous or polychronous signals.
+The term polychronous was first introduced in 2006, by E.
+Izhichevitch in {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]}.
+He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
+The network is characterized by a timing-dependant learning rule for weights (STDP) and by fixed conduction delays between neurons.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
+The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
+This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike and thus to the crossing of the voltage threshold and to the discharge of a spike.
+According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and thus, constitute a polychronous group.
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
+
+{[}
+\textbf{TODO}
+
+-\textgreater{} insert fig 2 of {[}\protect\hyperlink{ref-doi}{\textbf{doi?}} 10.1162/089976606775093882{]}
+
+Thus, the learning of delays allowing this polychronous group organization may be useful to detect temporal sequences of interest.
+
+\begin{itemize}
+\item
+  Reproducing Polychronization: A Guide to Maximizing the Reproducibility of Spiking Network Models \url{https://www.frontiersin.org/articles/10.3389/fninf.2018.00046/full}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    comes with python code
+  \end{itemize}
+\item
+  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html}
+\item
+  dynamic networks \& learning delays:
+
+  \begin{itemize}
+  \item
+    Huning H., Glunder H., and Palm G.
+(1998) Synaptic delay learning in pulse-coupled neurons.
+Neural Computation, 10:555--565.
+\url{https://www.deepdyve.com/lp/mit-press/synaptic-delay-learning-in-pulse-coupled-neurons-DGMiNHxp0A}
+  \item
+    \href{https://www.researchgate.net/publication/37921636_Dynamics_of_Self-Organized_Delay_Adaptation}{Eurich C., Pawelzik K., Ernst U., Cowan J., and Milton J.
+(1999) Dynamics of self-organazed delay adaptation.
+Phys.
+Rev.~Lett., 82:1594--1597.}
+  \item
+    The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining heterosynaptic plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
+  \item
+    an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
+  \end{itemize}
+\end{itemize}
+
+{]}\{.banner .lightred\}
+
+A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
+In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
+We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.'' {[}\protect\hyperlink{ref-KS4oxnst}{60}{]}.
+Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
+
+Memory traces in dynamical systems {[}\protect\hyperlink{ref-o7rANZux}{89}{]} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
+This raises important questions about the fundamental limits of such memory traces and the properties required of dynamical systems to achieve these limits.
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+We introduce the Fisher Memory Curve (FMC) as a measure of the signal-to-noise ratio (SNR) embedded in the dynamical state relative to the input SNR.
+The integrated FMC indicates the total memory capacity.
+We apply this theory to linear neuronal networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
+A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
+The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed √𝑁.
+This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
+We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.''
+
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
+
+\hypertarget{detecting-precise-spiking-motifs-in-biological-raster-plots}{%
+\section{Detecting precise spiking motifs in biological raster plots}\label{detecting-precise-spiking-motifs-in-biological-raster-plots}}
+
+\{sec:detection\}
+
+\hypertarget{spike-distances}{%
+\subsection{spike distances}\label{spike-distances}}
+
+J. D. Victor and K. P. Purpura, ``Nature and precision of temporal coding in visual cortex: a metric-space analysis,'' J. Neurophysiol., vol.~76, pp.~1310--1326, Aug.~1996.
+
+M.
+C.
+W.
+van Rossum, ``A novel spike distance,'' Neural Comput., vol.~13, no.
+4, pp.~751--763, 2001.
+{[}21{]} D.
+Aronov and J.
+D.
+Victor, ``Non-Euclidean properties of spike train metric spaces,'' Physical Rev.~E (Statist., Nonlinear, Soft Matter Phys.), vol.~69, no.
+6, 2004.
+
+T. Kreuz, J. S. Haas, A. Morelli, H. D. I. Abarbanel, and A. Politi, ``Measuring spike train synchrony,'' J. Neurosci. Methods, vol.~165, no. 1, pp.~151--161, 2007. {[}23{]} H.
+
+Paper by {[}\protect\hyperlink{ref-LjkTSK3h}{90}{]} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
+
+Weyl's discrepency measure {[}\protect\hyperlink{ref-RYzP72lj}{91}{]} which may lead to the definition of a cross-correlation.
+
+Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~https://dx.doi.org/10.1073/pnas.1902653116
+
+A review of the methods for neuronal response latency estimation including bayesian binning {[}\protect\hyperlink{ref-Z7LAUup3}{92}{]}.
+
+\hypertarget{decoding-neural-activity}{%
+\subsection{decoding neural activity}\label{decoding-neural-activity}}
+
+In generic linear non linear lnl models, the output is assumed to be poisson.
+As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression {[}\protect\hyperlink{ref-MgINa6bU}{93}{]}.
+This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
+
+Unitary event analysis is performed by a statistical model of coincidence detection {[}\protect\hyperlink{ref-13iJohyPk}{94}{]}.
+This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]} for instance).
+
+Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity {[}\protect\hyperlink{ref-1ErNTONy0}{95}{]}.
+By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range {[}\protect\hyperlink{ref-V9qkJFZU}{96}{]}.
+
+This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents'' {[}\protect\hyperlink{ref-pExXxdMb}{97}{]} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations {[}\protect\hyperlink{ref-d2m7wQpH}{98}{]}.
+
+{[}\protect\hyperlink{ref-doi}{\textbf{doi?}}{]}
+
+TODO: summarize ``Understanding Auditory Spectro-Temporal Receptive Fields and Their Changes with Input Statistics by Efficient Coding Principles'' - Spectro-temporal receptive fields (STRFs) have been widely used as linear approximations to the signal transform from sound spectrograms to neural responses along the auditory pathway.
+Their dependence on statistical attributes of the stimuli, such as sound intensity, is usually explained by nonlinear mechanisms and models.
+Here, we apply an efficient coding principle which has been successfully used to understand receptive fields in early stages of visual processing, in order to provide a computational understanding of the STRFs.
+According to this principle, STRFs result from an optimal tradeoff between maximizing the sensory information the brain receives, and minimizing the cost of the neural activities required to represent and transmit this information.
+Both terms depend on the statistical properties of the sensory inputs and the noise that corrupts them.
+The STRFs should therefore depend on the input power spectrum and the signal-to-noise ratio, which is assumed to increase with input intensity.
+We analytically derive the optimal STRFs when signal and noise are approximated as Gaussians.
+Under the constraint that they should be spectro-temporally local, the STRFs are predicted to adapt from being band-pass to low-pass filters as the input intensity reduces, or the input correlation becomes longer range in sound frequency or time.
+These predictions qualitatively match physiological observations.
+Our prediction as to how the STRFs should be determined by the input power spectrum could readily be tested, since this spectrum depends on the stimulus ensemble.
+The potentials and limitations of the efficient coding principle are discussed.
+{[}\protect\hyperlink{ref-s63GcujC}{99}{]}
+
+\hypertarget{rastermap-decoding-large-scale-data}{%
+\subsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}}
+
+\href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
+
+\begin{itemize}
+\tightlist
+\item
+  https://www.janelia.org/lab/stringer-lab
+\item
+  described in {[}\protect\hyperlink{ref-zc16plUu}{100}{]}
+\item
+  \href{https://github.com/MouseLand/rastermap}{rastermap}
+\item
+  deconvolution strategy
+\item
+  based on a linear model
+\end{itemize}
+
+\hypertarget{stringer-et-al-2019-nature-stringer2019nature}{%
+\subsubsection{\texorpdfstring{Stringer et al 2019, Nature {[}\protect\hyperlink{ref-brg2FFsb}{101}{]}}{Stringer et al 2019, Nature {[}101{]}}}\label{stringer-et-al-2019-nature-stringer2019nature}}
+
+\begin{itemize}
+\tightlist
+\item
+  ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
+\item
+  Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
+\item
+  Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
+\end{itemize}
+
+\hypertarget{stringer-et-al-2019-science-stringer2019science}{%
+\subsubsection{\texorpdfstring{Stringer et al 2019, Science {[}\protect\hyperlink{ref-1BPHUQLTf}{102}{]}}{Stringer et al 2019, Science {[}102{]}}}\label{stringer-et-al-2019-science-stringer2019science}}
+
+\begin{itemize}
+\item
+  Stringer et al.~analyzed spontaneous neuronal firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+The variability of neuronal responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
+\item
+  see also the work showing that you can encode very precise orientation information by using many neurons: {[}\protect\hyperlink{ref-1EMs3GS0B}{103}{]}
+\end{itemize}
+
+\hypertarget{detecting-structured-temporal-patterns}{%
+\subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}}
+
+\hypertarget{paper-by-grossberger-2018-grossberger2018}{%
+\subsubsection{\texorpdfstring{Paper by Grossberger, 2018 {[}\protect\hyperlink{ref-9AK1cwY1}{104}{]}}{Paper by Grossberger, 2018 {[}104{]}}}\label{paper-by-grossberger-2018-grossberger2018}}
+
+\begin{itemize}
+\tightlist
+\item
+  Temporally ordered multi-neuron patterns likely encode information in the brain.
+We introduce an unsupervised method, SPOTDisClust (Spike Pattern Optimal Transport Dissimilarity Clustering), for their detection from high-dimensional neural ensembles.
+SPOTDisClust measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other (SPOTDis).
+\item
+  Detecting these temporal patterns represents a major methodological challenge.
+\item
+  Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
+\item
+  Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
+\item
+  Lee AK, Wilson MA.
+A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order.
+J Neurophysiol.
+2004;92(4):2555--73.
+pmid:15212425
+\item
+  Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
+\item
+  only one spike per neuron: fig 1A = ``For each pattern and each neuron, a random position was chosen for the activation pulse.''
+\item
+  t-SNE projection with HDBSCAN labels shows that our clustering method can retrieve all patterns from the data.
+\item
+  data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
+\end{itemize}
+
+\begin{figure}
+\hypertarget{fig:G2018-1}{%
+\centering
+\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
+\caption{Fig 1 of {[}\protect\hyperlink{ref-9AK1cwY1}{104}{]}: ``Simulated example illustrating the steps in SPOTDisClust.
+A) Structure of five ``ground-truth'' patterns (\ldots).
+For each pattern and each neuron, a random position was chosen for the activation pulse.
+B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
+}
+\end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
+This becomes a major problem for computing an MM dissimilarity matrix (for M time epochs) for thousands of neurons.
+(2) It relies exclusively on pairwise spike-timing relationships (i.e 2nd order correlations), because it does not solve the optimal transport problem for the entire spike pattern, but only for neuron pairs separately.
+Hence, it may not be sensitive to higher-order correlations in spiking
+  patterns.
+  Here, we develop a novel dissimilarity measure for multi-neuron spiking patterns called SpikeShip, which has linear computational complexity of O(N).
+We achieve this by (1) computing the minimum transport cost of spikes for each spike train separately, and (2) discounting a global translation term in the transport flow across neurons.''
+  https://doi.org/10.1101/2020.06.03.131573;
+\end{itemize}
+
+\hypertarget{paper-by-russo-et-al-2017-russo2017}{%
+\subsubsection{\texorpdfstring{Paper by Russo et al 2017 {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}}{Paper by Russo et al 2017 {[}88{]}}}\label{paper-by-russo-et-al-2017-russo2017}}
+
+\begin{itemize}
+\tightlist
+\item
+  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}
+\item
+  sliding window as in {[}\protect\hyperlink{ref-1XpTamcl}{105}{]} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags {[}\protect\hyperlink{ref-bc9czTxi}{46}{]}
+\item
+  based on a ``non-stationarity-corrected parametric test statistic for assessing the independence of pairs'' and ``an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies''
+\end{itemize}
+
+\hypertarget{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}{%
+\subsubsection{\texorpdfstring{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}\protect\hyperlink{ref-1300cS9RL}{106}{]}}{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}106{]}}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}}
+
+\begin{itemize}
+\tightlist
+\item
+  Stochastic sampling links perceptual uncertainty to neural response variability
+\item
+  Model accounts for independent changes in strength and variability of responses
+\item
+  Model predicts relationship between noise, signal, and spontaneous correlations
+\item
+  Stimulus statistics dependence of response statistics is explained
+\end{itemize}
+
+To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters, {[}\protect\hyperlink{ref-1365b3Wwl}{107}{]} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
+They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
+
+\hypertarget{dynamics-of-delay-coupled-excitable-neural-systems}{%
+\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}\label{dynamics-of-delay-coupled-excitable-neural-systems}}
+
+uses a FPGA
+{[}\protect\hyperlink{ref-rYLmAESf}{108}{]}
+
+\begin{verbatim}
+February 2009International Journal of Bifurcation and Chaos 19(02):745-753
+\end{verbatim}
+
+V.
+Thanasoulis, B.
+Vogginger, J.
+Partzsch and C.
+Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
+{[}\protect\hyperlink{ref-1DadetlqZ}{109}{]}
+
+\hypertarget{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}{%
+\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}}
+
+{[}\protect\hyperlink{ref-RrgzWOSk}{88}{]} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
+It is based on unitary events in multiple single-neuron spiking activity {[}\protect\hyperlink{ref-iNkKTYaI}{110}{]} and the reliable and efficient analysis of an excess or deficiency of joint-spike events {[}\protect\hyperlink{ref-14wpaupd5}{111}{]}.
+Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+
+\hypertarget{learning-to-detect-polychronous-groups}{%
+\section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}}
+
+\hypertarget{learning-weights-and-delays}{%
+\subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}}
+
+spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP) {[}\protect\hyperlink{ref-19NyVq80B}{112}{]} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
+This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach.
+This will decouple the active synapses (similarly to a logistic regression) from the values of possible synaptic delays.
+
+\begin{itemize}
+\tightlist
+\item
+  {[}\protect\hyperlink{ref-11p3NXmkg}{113}{]} : coherence detection
+\item
+  {[}\protect\hyperlink{ref-j5xS3owo}{114}{]} : STDP
+\item
+  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks {[}\protect\hyperlink{ref-15TktA0gR}{115}{]} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
+PLoS Comput Biol 9(2): e1002897.
+{[}\protect\hyperlink{ref-1FHIKpizQ}{116}{]} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+\end{itemize}
+
+Our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
+Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
+This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.
+{[}\protect\hyperlink{ref-s09KZ8Kh}{117}{]}
+
+\textbf{TODO amelie} section sur l'apprentissage des délais dans la biologie
+
+{[}\protect\hyperlink{ref-UeTLXQCp}{118}{]}
+
+{]}\{.banner .lightred\}
+
+Bio-plausible Unsupervised Delay Learning for Extracting Temporal Features in Spiking Neural Networks
+Alireza Nadafian, Mohammad Ganjtabesh
+
+\begin{verbatim}
+The plasticity of the conduction delay between neurons plays a fundamental role in learning.
+However, the exact underlying mechanisms in the brain for this modulation is still an open problem.
+Understanding the precise adjustment of synaptic delays could help us in developing effective brain-inspired computational models in providing aligned insights with the experimental evidence.
+In this paper, we propose an unsupervised biologically plausible learning rule for adjusting the synaptic delays in spiking neural networks.
+Then, we provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns.
+Furthermore, the experimental results of applying an STDP-based spiking neural network equipped with our proposed delay learning rule on Random Dot Kinematogram indicate the efficacy of the proposed delay learning rule in extracting temporal features.
+\end{verbatim}
+
+Integrating synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes, that is referred to as synaptic delay-weight plasticity {[}\protect\hyperlink{ref-ga3ZTFC7}{61}{]}.
+Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
+
+In a recent paper {[}\protect\hyperlink{ref-B493Fl4K}{119}{]}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
+In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
+
+In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity {[}\protect\hyperlink{ref-pPRQwZb}{120}{]}.
+This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
+
+\hypertarget{learning-sequences}{%
+\subsection{Learning sequences}\label{learning-sequences}}
+
+In {[}\protect\hyperlink{ref-WFwUKZb5}{121}{]}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds'' {[}\protect\hyperlink{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga}{124}{]} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
+
+\begin{itemize}
+\item
+  in {[}\protect\hyperlink{ref-Jzr1gwsr}{76}{]}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
+\item
+  In Bellec {[}\protect\hyperlink{ref-aj6TgorR}{125}{]}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+
+  \begin{itemize}
+  \tightlist
+  \item
+    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron {[}\protect\hyperlink{ref-14e3BCBii}{126}{]} with recurrent dynamics)
+  \item
+    sample and measure method to include latent / hidden neurons
+  \item
+    comes with code https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure
+  \item
+    V1-dataset : The dataset we used was collected by Smith and Kohn {[}49{]} and is publicly available at:
+    http://crcns.org/data-sets/vc/pvc-11 - it is in a sense supervised with the input being the movie and the output the spikes recorded.
+  \end{itemize}
+\end{itemize}
+
+A recent work {[}\protect\hyperlink{ref-10EU8MfxC}{127}{]} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
+The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
+The second stage takes the output of the spatiotemporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatiotemporal activity.
+For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
+Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters.
+Showing the potential of such representation
+
+\hypertarget{todo-more-bib-to-read}{%
+\subsection{TODO: more bib to read}\label{todo-more-bib-to-read}}
+
+Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
+Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021
+
+Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
+Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020
+
+Unsupervised Learning of Persistent and Sequential Activity.
+Pereira U, Brunel N.Front Comput Neurosci. 2020
+
+From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
+Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019
+
+Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
+Pang R, Fairhall AL.Elife. 2019 May {[}\protect\hyperlink{ref-12zyUFdO4}{128}{]}
+
+Emergence of spontaneous assembly activity in developing neural networks without afferent input.
+Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018
+
+Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
+Ocker GK, Doiron B.Cereb Cortex. 2019.
+{]}\{.banner .lightred\}
+
+\textbf{TODO}
+
+\hypertarget{learning-pattern-detection-on-natural-images-event-based-cameras}{%
+\subsection{learning pattern detection on natural images / event-based cameras}\label{learning-pattern-detection-on-natural-images-event-based-cameras}}
+
+\hypertarget{sparse-coding-on-spatio-temporal-data}{%
+\subsubsection{sparse coding on spatio-temporal data}\label{sparse-coding-on-spatio-temporal-data}}
+
+\hypertarget{hots}{%
+\subsubsection{HOTS}\label{hots}}
+
+\hypertarget{grimaldi-cbmi-pami}{%
+\subsubsection{Grimaldi CBMI / PAMI}\label{grimaldi-cbmi-pami}}
+
+{]}\{.banner .lightred\}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\textbf{TODO}
+
+\hypertarget{dynamical-models}{%
+\subsection{dynamical models}\label{dynamical-models}}
+
+Dumas and colleagues {[}\protect\hyperlink{ref-WxQdcMB2}{129}{]} : three levels / fourth paradigm {[}\protect\hyperlink{ref-JrZPoxk9}{130}{]} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
+
+\hypertarget{our-model}{%
+\subsection{our model}\label{our-model}}
+
+Here, we develop a model for the efficient detection of such PGs based on the inversion of a probabilistic model defining the generation of the raster plot as a combination of such groups.
+We show that such an inference can be achieved by a neural-like computation that could itself be used as a spiking neuron, as can be implemented in a neuromorphic chip for instance.
+A first result is to show the efficiency of such a scheme in detecting different PGs occurring at specific times in synthetic data.
+The representational capacity of the PGs is particularly interesting compared to traditional models of neuronal encoding using spiking frequency.
+Our second result is to propose a novel learning method for learning PGs in raster plots in a self-supervised manner.
+Finally we demonstrate the use of this algorithm to the output of an event-based camera and how this may separate independent components from the stream of events.
+This end-to-end event-based computational brick could help improve the performance of current Spiking Neural Network solution currently used in neuromorphic chips.
+
+Emergence of electronic architectures specialized in Sparse Event-Based Convolutions {[}\protect\hyperlink{ref-sJ03d2Zj}{131}{]} or of sparsity aware algorithms {[}\protect\hyperlink{ref-R5BYFgCP}{132}{]}.
+
+and oscillations? Journal Article • DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?
+
+\hypertarget{development}{%
+\section{development}\label{development}}
+
+scaffolding of neural assemblies / existence of critical periods : {[}\protect\hyperlink{ref-1HSyNOl48}{133}{]}
+{]}\{.banner .lightred\}
+
+%
+%The introduction should briefly place the study in a broad context and highlight why it is important.
 It should define the purpose of the work and its significance.
@@
 Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Materials and Methods}
+%
+%Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
 Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
@@
 New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers.
+%
+%Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers.
 If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
 They must be provided prior to publication.
-
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-\begin{quote}
-This is an example of a quote.
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-This section may be divided by subheadings.
+%
+%Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
+%\begin{quote}
+%This is an example of a quote.
+%\end{quote}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Results}
+%
+%This section may be divided by subheadings.
 It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\subsection{Subsection}
-\subsubsection{Subsubsection}
-
-Bulleted lists look like this:
-\begin{itemize}
-\item	First bullet;
-\item	Second bullet;
-\item	Third bullet.
-\end{itemize}
-
-Numbered lists can be added as follows:
-\begin{enumerate}
-\item	First item; 
-\item	Second item;
-\item	Third item.
-\end{enumerate}
-
-The text continues here.
-
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.
-
-\begin{figure}[H]
-\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
-\caption{This is a figure.
+%\section{Subsection}
+%\subsection{Subsubsection}
+%
+%Bulleted lists look like this:
+%\begin{itemize}
+%\item	First bullet;
+%\item	Second bullet;
+%\item	Third bullet.
+%\end{itemize}
+%
+%Numbered lists can be added as follows:
+%\begin{enumerate}
+%\item	First item; 
+%\item	Second item;
+%\item	Third item.
+%\end{enumerate}
+%
+%The text continues here. 
+%
+%\section{Figures, Tables and Schemes}
+%
+%All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.
+%
+%\begin{figure}[H]
+%\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
+%\caption{This is a figure.
 Schemes follow the same formatting.
@@
 A caption on a single line should be centered.\label{fig1}}
-\end{figure}   
-\unskip
-
-\begin{table}[H] 
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
-\newcolumntype{C}{>{\centering\arraybackslash}X}
-\begin{tabularx}{\textwidth}{CCC}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-Entry 1		& Data			& Data\\
-Entry 2		& Data			& Data\\
-\bottomrule
-\end{tabularx}
-\end{table}
-\unskip
-
-\begin{table}[H]
-\caption{This is a wide table.\label{tab2}}
-	\begin{adjustwidth}{-\extralength}{0cm}
-		\newcolumntype{C}{>{\centering\arraybackslash}X}
-		\begin{tabularx}{\fulllength}{CCCC}
-			\toprule
-			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
-			\midrule
-			Entry 1		& Data			& Data			& Data\\
-			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
-			\bottomrule
-		\end{tabularx}
-	\end{adjustwidth}
-	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
-\end{table}
-
-%\begin{listing}[H]
-%\caption{Title of the listing}
-%\rule{\columnwidth}{1pt}
-%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
-%\rule{\columnwidth}{1pt}
-%\end{listing}
-
-Text.
-
-Text.
-
-\subsection{Formatting of Mathematical Components}
-
-This is the example 1 of equation:
-\begin{linenomath}
-\begin{equation}
-a = 1,
-\end{equation}
-\end{linenomath}
-the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-
-This is the example 2 of equation:
-\begin{adjustwidth}{-\extralength}{0cm}
-\begin{equation}
-a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
-\end{equation}
-\end{adjustwidth}
-
-% Example of a page in landscape format (with table and table footnote).
-%\startlandscape
-%\begin{table}[H] %% Table in wide page
-%\caption{This is a very wide table.\label{tab3}}
-%	\begin{tabularx}{\textwidth}{CCCC}
+%\end{figure}   
+%\unskip
+%
+%\begin{table}[H] 
+%\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
+%\newcolumntype{C}{>{\centering\arraybackslash}X}
+%\begin{tabularx}{\textwidth}{CCC}
+%\toprule
+%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+%\midrule
+%Entry 1		& Data			& Data\\
+%Entry 2		& Data			& Data\\
+%\bottomrule
+%\end{tabularx}
+%\end{table}
+%\unskip
+%
+%\begin{table}[H]
+%\caption{This is a wide table.\label{tab2}}
+%	\begin{adjustwidth}{-\extralength}{0cm}
+%		\newcolumntype{C}{>{\centering\arraybackslash}X}
+%		\begin{tabularx}{\fulllength}{CCCC}
 %			\toprule
@@
 %			\midrule
-%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
-%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
+%			Entry 1		& Data			& Data			& Data\\
+%			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
 %			\bottomrule
 %		\end{tabularx}
-%	\begin{adjustwidth}{+\extralength}{0cm}
-%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
 %	\end{adjustwidth}
+%	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
 %\end{table}
-%\finishlandscape
-
-% Example of a figure that spans the whole page width. The same concept works for tables, too.
-\begin{figure}[H]
-\begin{adjustwidth}{-\extralength}{0cm}
-\centering
-\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
-\end{adjustwidth}
-\caption{This is a wide figure.\label{fig2}}
-\end{figure}
-
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
-
-The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
+%
+%%\begin{listing}[H]
+%%\caption{Title of the listing}
+%%\rule{\columnwidth}{1pt}
+%%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
+%%\rule{\columnwidth}{1pt}
+%%\end{listing}
+%
+%Text.
+%
+%Text.
+%
+%\section{Formatting of Mathematical Components}
+%
+%This is the example 1 of equation:
+%\begin{linenomath}
+%\begin{equation}
+%a = 1,
+%\end{equation}
+%\end{linenomath}
+%the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
+%%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
+This ensures correct linenumbering.
+The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
+%
+%This is the example 2 of equation:
+%\begin{adjustwidth}{-\extralength}{0cm}
+%\begin{equation}
+%a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
+%\end{equation}
+%\end{adjustwidth}
+%
+%% Example of a page in landscape format (with table and table footnote).
+%%\startlandscape
+%%\begin{table}[H] %% Table in wide page
+%%\caption{This is a very wide table.\label{tab3}}
+%%	\begin{tabularx}{\textwidth}{CCCC}
+%%		\toprule
+%%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
+%%		\midrule
+%%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
+%%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
+%%		\bottomrule
+%%	\end{tabularx}
+%%	\begin{adjustwidth}{+\extralength}{0cm}
+%%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
+%%	\end{adjustwidth}
+%%\end{table}
+%%\finishlandscape
+%
+%% Example of a figure that spans the whole page width. The same concept works for tables, too.
+%\begin{figure}[H]
+%\begin{adjustwidth}{-\extralength}{0cm}
+%\centering
+%\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
+%\end{adjustwidth}
+%\caption{This is a wide figure.\label{fig2}}
+%\end{figure}  
+%
+%Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
+%%% Example of a theorem:
+%\begin{Theorem}
+%Example text of a theorem.
+%\end{Theorem}
+%
+%The text continues here. Proofs must be formatted as follows:
+%
+%%% Example of a proof:
+%\begin{proof}[Proof of Theorem 1]
+%Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
+%\end{proof}
+%The text continues here.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Discussion}
+%
+%Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
 The findings and their implications should be discussed in the broadest context possible.
 Future research directions may also be highlighted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Conclusions}
+%
+%This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
 

### 2022-10-04 manuscript.tex
importing from manubot : cleanup

@@
 
-\hypertarget{introduction-importance-of-precise-spike-timings-in-the-brain}{%
-\section{Introduction: Importance of precise spike timings in the brain}\label{introduction-importance-of-precise-spike-timings-in-the-brain}}
-
-\hypertarget{ultra-fast-neural-codes-for-ultra-fast-vision}{%
-\subsection{Ultra fast neural codes for ultra fast vision}\label{ultra-fast-neural-codes-for-ultra-fast-vision}}
+\section{Introduction: Importance of precise spike timings in the brain}\label{introduction-importance-of-precise-spike-timings-in-the-brain}
+\subsection{Ultra fast neural codes for ultra fast vision}\label{ultra-fast-neural-codes-for-ultra-fast-vision}
 
@@
 \begin{figure}
-\hypertarget{fig:thorpe}{%
 \centering
@@
 \caption{From {[}\protect\hyperlink{ref-12dWDpY61}{6}{]}}\label{fig:thorpe}
-}
 \end{figure}
 
-\hypertarget{how-timing-encodes-analogous-profile}{%
-\subsection{How timing encodes analogous profile}\label{how-timing-encodes-analogous-profile}}
-
+\subsection{How timing encodes analogous profile}\label{how-timing-encodes-analogous-profile}
 An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief impulses that propagate along the axons of neurons.
@@
 \begin{figure}
-\hypertarget{fig:roc}{%
 \centering
@@
 From {[}\protect\hyperlink{ref-YX5wqWH4}{16}{]}}\label{fig:roc}
-}
 \end{figure}
@@
 
-\hypertarget{from-synfire-chains}{%
-\subsection{From synfire chains\ldots{}}\label{from-synfire-chains}}
-
+\subsection{From synfire chains\ldots{}}\label{from-synfire-chains}
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
@@
 \begin{figure}
-\hypertarget{fig:mainen}{%
 \centering
@@
 See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
-}
 \end{figure}
@@
 \begin{figure}
-\hypertarget{fig:diesman}{%
 \centering
@@
 This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
-}
 \end{figure}
@@
 \begin{figure}
-\hypertarget{fig:pynn}{%
 \centering
@@
 The state space gives the duration of synfire chains as a function of the protocol's parameters and compares (C) CPU to (D) neuromorphic implementations.}\label{fig:pynn}
-}
 \end{figure}
@@
 
-\hypertarget{to-traveling-waves}{%
-\subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}}
-
+\subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}
 TODO : {[}\protect\hyperlink{ref-zNDCo9Hy}{49}{]}
@@
 \begin{figure}
-\hypertarget{fig:davis}{%
 \centering
@@
 Data for panels c--e modified from Davis et al.17 with permission.}\label{fig:davis}
-}
 \end{figure}
@@
 
-\hypertarget{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}{%
-\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}}
-
+\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
 It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas {[}\protect\hyperlink{ref-yLVM9IXi}{57}{]}.
@@
 
-\hypertarget{todo-and-to-synfire-braids-and-polychronous-groups}{%
-\subsubsection{TODO \ldots{} and to synfire braids and polychronous groups}\label{todo-and-to-synfire-braids-and-polychronous-groups}}
-
+\subsubsection{TODO \ldots{} and to synfire braids and polychronous groups}\label{todo-and-to-synfire-braids-and-polychronous-groups}
 \begin{itemize}
@@
 
-\hypertarget{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}{%
-\subsection{Are there precise heterosynaptic spiking motifs in the brain?}\label{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}}
-
+\subsection{Are there precise heterosynaptic spiking motifs in the brain?}\label{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}
 Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
@@
 \begin{figure}
-\hypertarget{fig:haimerl}{%
 \centering
@@
 The black line represents speed of the mouse.}\label{fig:haimerl}
-}
 \end{figure}
@@
 
-\hypertarget{cortical-songs}{%
-\subsection{Cortical songs}\label{cortical-songs}}
-
+\subsection{Cortical songs}\label{cortical-songs}
 \begin{itemize}
@@
 \begin{figure}
-\hypertarget{fig:Ikegaya2004}{%
 \centering
@@
 In (i), the blue trace is shifted --2.75 mV; in (ii), the blue trace is shifted --1.58 mV, and the green +0.79 mV.}\label{fig:Ikegaya2004}
-}
 \end{figure}
@@
 
-\hypertarget{outline}{%
-\subsection{outline}\label{outline}}
-
+\subsection{outline}\label{outline}
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs) {[}\protect\hyperlink{ref-DUpFdm8C}{78}{]}.
@@
 
-{]}\{.banner .lightred\}
-
 Remarkably, novel neuromorphic chips use a representation similar to that of real neurons {[}\protect\hyperlink{ref-Oet1bgdb}{22}{]}.
@@
 \begin{figure}
-\hypertarget{fig:silicon_retina}{%
 \centering
@@
 These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).}\label{fig:silicon_retina}
-}
 \end{figure}
@@
 
-\hypertarget{theoretical-models-for-the-detection-of-precise-heterosynaptic-spiking-motifs}{%
-\section{Theoretical models for the detection of precise heterosynaptic spiking motifs}\label{theoretical-models-for-the-detection-of-precise-heterosynaptic-spiking-motifs}}
-
+\section{Theoretical models for the detection of precise heterosynaptic spiking motifs}\label{theoretical-models-for-the-detection-of-precise-heterosynaptic-spiking-motifs}
 \#sec:detection
@@
 
-\hypertarget{izhikevitchs-polychronization-model}{%
-\subsection{Izhikevitch's polychronization model}\label{izhikevitchs-polychronization-model}}
-
+\subsection{Izhikevitch's polychronization model}\label{izhikevitchs-polychronization-model}
 Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
@@
 
-{]}\{.banner .lightred\}
-
 A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
@@
 
-\hypertarget{detecting-precise-spiking-motifs-in-biological-raster-plots}{%
-\section{Detecting precise spiking motifs in biological raster plots}\label{detecting-precise-spiking-motifs-in-biological-raster-plots}}
-
+\section{Detecting precise spiking motifs in biological raster plots}\label{detecting-precise-spiking-motifs-in-biological-raster-plots}
 \{sec:detection\}
 
-\hypertarget{spike-distances}{%
-\subsection{spike distances}\label{spike-distances}}
-
+\subsection{spike distances}\label{spike-distances}
 J. D. Victor and K. P. Purpura, ``Nature and precision of temporal coding in visual cortex: a metric-space analysis,'' J. Neurophysiol., vol.~76, pp.~1310--1326, Aug.~1996.
@@
 
-\hypertarget{decoding-neural-activity}{%
-\subsection{decoding neural activity}\label{decoding-neural-activity}}
-
+\subsection{decoding neural activity}\label{decoding-neural-activity}
 In generic linear non linear lnl models, the output is assumed to be poisson.
@@
 
-\hypertarget{rastermap-decoding-large-scale-data}{%
-\subsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}}
-
+\subsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}
 \href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
@@
 
-\hypertarget{stringer-et-al-2019-nature-stringer2019nature}{%
-\subsubsection{\texorpdfstring{Stringer et al 2019, Nature {[}\protect\hyperlink{ref-brg2FFsb}{101}{]}}{Stringer et al 2019, Nature {[}101{]}}}\label{stringer-et-al-2019-nature-stringer2019nature}}
-
+\subsubsection{\texorpdfstring{Stringer et al 2019, Nature {[}\protect\hyperlink{ref-brg2FFsb}{101}{]}}{Stringer et al 2019, Nature {[}101{]}}}\label{stringer-et-al-2019-nature-stringer2019nature}
 \begin{itemize}
@@
 
-\hypertarget{stringer-et-al-2019-science-stringer2019science}{%
-\subsubsection{\texorpdfstring{Stringer et al 2019, Science {[}\protect\hyperlink{ref-1BPHUQLTf}{102}{]}}{Stringer et al 2019, Science {[}102{]}}}\label{stringer-et-al-2019-science-stringer2019science}}
-
+\subsubsection{\texorpdfstring{Stringer et al 2019, Science {[}\protect\hyperlink{ref-1BPHUQLTf}{102}{]}}{Stringer et al 2019, Science {[}102{]}}}\label{stringer-et-al-2019-science-stringer2019science}
 \begin{itemize}
@@
 
-\hypertarget{detecting-structured-temporal-patterns}{%
-\subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}}
-
-\hypertarget{paper-by-grossberger-2018-grossberger2018}{%
-\subsubsection{\texorpdfstring{Paper by Grossberger, 2018 {[}\protect\hyperlink{ref-9AK1cwY1}{104}{]}}{Paper by Grossberger, 2018 {[}104{]}}}\label{paper-by-grossberger-2018-grossberger2018}}
-
+\subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
+\subsubsection{\texorpdfstring{Paper by Grossberger, 2018 {[}\protect\hyperlink{ref-9AK1cwY1}{104}{]}}{Paper by Grossberger, 2018 {[}104{]}}}\label{paper-by-grossberger-2018-grossberger2018}
 \begin{itemize}
@@
 \begin{figure}
-\hypertarget{fig:G2018-1}{%
 \centering
@@
 B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
-}
 \end{figure}
@@
 
-\hypertarget{paper-by-russo-et-al-2017-russo2017}{%
-\subsubsection{\texorpdfstring{Paper by Russo et al 2017 {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}}{Paper by Russo et al 2017 {[}88{]}}}\label{paper-by-russo-et-al-2017-russo2017}}
-
+\subsubsection{\texorpdfstring{Paper by Russo et al 2017 {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}}{Paper by Russo et al 2017 {[}88{]}}}\label{paper-by-russo-et-al-2017-russo2017}
 \begin{itemize}
@@
 
-\hypertarget{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}{%
-\subsubsection{\texorpdfstring{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}\protect\hyperlink{ref-1300cS9RL}{106}{]}}{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}106{]}}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}}
-
+\subsubsection{\texorpdfstring{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}\protect\hyperlink{ref-1300cS9RL}{106}{]}}{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}106{]}}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
 \begin{itemize}
@@
 
-\hypertarget{dynamics-of-delay-coupled-excitable-neural-systems}{%
-\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}\label{dynamics-of-delay-coupled-excitable-neural-systems}}
-
+\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}\label{dynamics-of-delay-coupled-excitable-neural-systems}
 uses a FPGA
@@
 
-\hypertarget{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}{%
-\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}}
-
+\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
 {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
@@
 
-\hypertarget{learning-to-detect-polychronous-groups}{%
-\section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}}
-
-\hypertarget{learning-weights-and-delays}{%
-\subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}}
-
+\section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}
+\subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
 spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP) {[}\protect\hyperlink{ref-19NyVq80B}{112}{]} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
@@
 {[}\protect\hyperlink{ref-UeTLXQCp}{118}{]}
-
-{]}\{.banner .lightred\}
 
@@
 
-\hypertarget{learning-sequences}{%
-\subsection{Learning sequences}\label{learning-sequences}}
-
+\subsection{Learning sequences}\label{learning-sequences}
 In {[}\protect\hyperlink{ref-WFwUKZb5}{121}{]}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds'' {[}\protect\hyperlink{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga}{124}{]} and this model shows how one could vary HVC activity using something similar to dropout in ML.
@@
 
-\hypertarget{todo-more-bib-to-read}{%
-\subsection{TODO: more bib to read}\label{todo-more-bib-to-read}}
-
+\subsection{TODO: more bib to read}\label{todo-more-bib-to-read}
 Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
@@
 Ocker GK, Doiron B.Cereb Cortex. 2019.
-{]}\{.banner .lightred\}
 
@@
 
-\hypertarget{learning-pattern-detection-on-natural-images-event-based-cameras}{%
-\subsection{learning pattern detection on natural images / event-based cameras}\label{learning-pattern-detection-on-natural-images-event-based-cameras}}
-
-\hypertarget{sparse-coding-on-spatio-temporal-data}{%
-\subsubsection{sparse coding on spatio-temporal data}\label{sparse-coding-on-spatio-temporal-data}}
-
-\hypertarget{hots}{%
-\subsubsection{HOTS}\label{hots}}
-
-\hypertarget{grimaldi-cbmi-pami}{%
-\subsubsection{Grimaldi CBMI / PAMI}\label{grimaldi-cbmi-pami}}
-
-{]}\{.banner .lightred\}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
+\subsection{learning pattern detection on natural images / event-based cameras}\label{learning-pattern-detection-on-natural-images-event-based-cameras}
+\subsubsection{sparse coding on spatio-temporal data}\label{sparse-coding-on-spatio-temporal-data}
+\subsubsection{HOTS}\label{hots}
+\subsubsection{Grimaldi CBMI / PAMI}\label{grimaldi-cbmi-pami}
+\section{Discussion}\label{discussion}
 \textbf{TODO}
 
-\hypertarget{dynamical-models}{%
-\subsection{dynamical models}\label{dynamical-models}}
-
+\subsection{dynamical models}\label{dynamical-models}
 Dumas and colleagues {[}\protect\hyperlink{ref-WxQdcMB2}{129}{]} : three levels / fourth paradigm {[}\protect\hyperlink{ref-JrZPoxk9}{130}{]} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
 
-\hypertarget{our-model}{%
-\subsection{our model}\label{our-model}}
-
+\subsection{our model}\label{our-model}
 Here, we develop a model for the efficient detection of such PGs based on the inversion of a probabilistic model defining the generation of the raster plot as a combination of such groups.
@@
 
-\hypertarget{development}{%
-\section{development}\label{development}}
-
+\section{development}\label{development}
 scaffolding of neural assemblies / existence of critical periods : {[}\protect\hyperlink{ref-1HSyNOl48}{133}{]}
-{]}\{.banner .lightred\}
 

### 2022-10-04 manuscript.tex
importing from manubot : figures

@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{https://cdn-learn.adafruit.com/assets/assets/000/086/572/original/circuitpython_Thorpe-FabreThorpe-Neuroscience-Seeking-categories-in-the-brain-monkey-visual-latency-estimate.jpg}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/visual-latency-estimate.jpg}
 \caption{From {[}\protect\hyperlink{ref-12dWDpY61}{6}{]}}\label{fig:thorpe}
@@
 \centering
-\includegraphics{https://camo.githubusercontent.com/77f4a48a0fe8f98ffab3c1f21dca1e74286b14d47c17120ee9627f3351fcd5e7/687474703a2f2f692e737461636b2e696d6775722e636f6d2f69786e727a2e706e67}
+\includegraphics{figures/MainenSejnowski1995.png} % TODO replace by the simulation
 \caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
@@
 \centering
-\includegraphics{https://brian2.readthedocs.io/en/stable/_images/frompapers.Diesmann_et_al_1999.1.png}
+\includegraphics{figures/Diesmann_et_al_1999.png}
 \caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
@@
 \centering
-\includegraphics{https://www.frontiersin.org/files/Articles/39056/fnins-07-00011-r2/image_m/fnins-07-00011-g004.jpg}
+\includegraphics{figures/fnins-07-00011-g004.jpg}
 \caption{A pyNN implementation of a Synfire chain with feedforward inhibition.

### 2022-10-04 manuscript.tex
importing from manubot : figures

@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
+%\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
 \caption{a Spike raster for repeated presentations (N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
@@
 \centering
-\includegraphics{https://www.pnas.org/cms/10.1073/pnas.1718518116/asset/ab9a48f6-b28f-40c1-978e-6f6dbec39b98/assets/graphic/pnas.1718518116fig01.jpeg}
+%\includegraphics{https://www.pnas.org/cms/10.1073/pnas.1718518116/asset/ab9a48f6-b28f-40c1-978e-6f6dbec39b98/assets/graphic/pnas.1718518116fig01.jpeg}
 \caption{Mixed distance and duration representation in CA1.
@@
 \centering
-\includegraphics{https://laurentperrinet.github.io/grant/anr-anr/event_driven_computations.png}
+\includegraphics{figures/event_driven_computations.png}
 \caption{A miniature, event-based ATIS sensor.

### 2022-10-04 build/build.sh
ezxporting

@@ -19,7 +19,7 @@ DOCKER_RUNNING="$(docker info &> /dev/null && echo "true" || (true && echo "fals
 CI="${CI:-false}"
 BUILD_PDF="${BUILD_PDF:-true}"
 BUILD_DOCX="${BUILD_DOCX:-false}"
-BUILD_LATEX="${BUILD_LATEX:-false}"
+BUILD_LATEX="${BUILD_LATEX:-true}"
 SPELLCHECK="${SPELLCHECK:-false}"
 MANUBOT_USE_DOCKER="${MANUBOT_USE_DOCKER:-$DOCKER_RUNNING}"
 # Pandoc's configuration is specified via files of option defaults

### 2022-10-04 manuscript.tex
Update manuscript.tex

@@
 Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
-Its first experiment consisted in asking subjects to categorize images that do or do not contain animals {[}\protect\hyperlink{ref-G3GlupPg}{1}{]}.
+Its first experiment consisted in asking subjects to categorize images that do or do not contain animals~\citep{ref-G3GlupPg}.
 The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neuronal activity.
-These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade {[}\protect\hyperlink{ref-B9iZWoq3}{2}{]}.
+These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade~\citep{ref-B9iZWoq3}.
 
@@
 The performances decrease progressively as the frequency of presentation of the images increases.
-However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image {[}\protect\hyperlink{ref-EqCCVwl2}{3}{]}.
+However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image~\citep{ref-EqCCVwl2}.
 
 This speed of the visual cortex, although surprising, is quite compatible with the latencies that are recorded at the neuro-physiological level.
-Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque {[}\protect\hyperlink{ref-Vu87UU6A}{4}{]}, see Figure {[}\ref{fig:thorpe}{]}.
-This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas {[}\protect\hyperlink{ref-15aMjRHVK}{5}{]}.
+Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque~\citep{ref-Vu87UU6A}, see Figure {[}\ref{fig:thorpe}{]}.
+This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{ref-15aMjRHVK}.
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/visual-latency-estimate.jpg}
-\caption{From {[}\protect\hyperlink{ref-12dWDpY61}{6}{]}}\label{fig:thorpe}
+\caption{From~\citep{ref-12dWDpY61}}\label{fig:thorpe}
 \end{figure}
@@
 Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
-Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing {[}\protect\hyperlink{ref-zZUVkTfZ}{7}{]} but also the dependance across neurons {[}\protect\hyperlink{ref-13T2Gne9z}{8}{]}.
-
-Numerous studies demonstrated the importance of precise timing in neural population activity {[}\protect\hyperlink{ref-BNK6T5Uk}{9}{]}, efficient encoding thanks to the use of spike latencies {[}\protect\hyperlink{ref-1DpSM1kpQ}{10},\protect\hyperlink{ref-IjepoonT}{11}{]} or precise timing in the auditory system {[}\protect\hyperlink{ref-mxPMgOmW}{12},\protect\hyperlink{ref-1ECJhg2GM}{13}{]}.
-All these findings, and more {[}\protect\hyperlink{ref-WZsEyOlK}{14}{]}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
-Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons {[}\protect\hyperlink{ref-XsHCA4IZ}{15}{]}.
+Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{ref-zZUVkTfZ} but also the dependance across neurons~\citep{ref-13T2Gne9z}.
+
+Numerous studies demonstrated the importance of precise timing in neural population activity~\citep{ref-BNK6T5Uk}, efficient encoding thanks to the use of spike latencies~\citep{ref-1DpSM1kpQ}{10},\protect\hyperlink{ref-IjepoonT} or precise timing in the auditory system~\citep{ref-mxPMgOmW}{12},\protect\hyperlink{ref-1ECJhg2GM}.
+All these findings, and more~\citep{ref-WZsEyOlK}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
+Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{ref-XsHCA4IZ}.
 
@@
 In this example, one generates at most one spike per neuron.
-From {[}\protect\hyperlink{ref-YX5wqWH4}{16}{]}}\label{fig:roc}
+From~\citep{ref-YX5wqWH4}}\label{fig:roc}
 \end{figure}
 
-At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina {[}\protect\hyperlink{ref-IjepoonT}{11}{]}.
+At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{ref-IjepoonT}.
 Notably one can appreciate in figure 1 that the response of ganglion cells to visual gratings that are flashed onto the retina.
@@
 The conclusion of the authors is that the discharge latency of the neurons allows to encode spatial characteristics of the image.
-First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account {[}\protect\hyperlink{ref-1H1Omh73g}{17}{]}.
-Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway {[}\protect\hyperlink{ref-UlbjMU0W}{18}{]}.
-
-Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area {[}\protect\hyperlink{ref-513xqzmL}{19}{]}.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{ref-1H1Omh73g}.
+Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{ref-UlbjMU0W}.
+
+Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{ref-513xqzmL}.
 Many models have used these properties in temporal coding to build fast image categorization networks.
-These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization {[}\protect\hyperlink{ref-1DpSM1kpQ}{10}{]}.
-This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes {[}\protect\hyperlink{ref-7K80jcx9}{20},\protect\hyperlink{ref-fRudGmR1}{21}{]}.
-This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events {[}\protect\hyperlink{ref-Oet1bgdb}{22}{]}.
-This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation {[}\protect\hyperlink{ref-15cR83gqd}{23}{]}.
-Note also that timing is not entirely sensorial or internal but in {[}\protect\hyperlink{ref-1E8gkynCm}{24}{]}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
-Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information'' {[}\protect\hyperlink{ref-QK8RfvTZ}{25}{]}.
+These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{ref-1DpSM1kpQ}.
+This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{ref-7K80jcx9}{20},\protect\hyperlink{ref-fRudGmR1}.
+This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events~\citep{ref-Oet1bgdb}.
+This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{ref-15cR83gqd}.
+Note also that timing is not entirely sensorial or internal but in~\citep{ref-1E8gkynCm}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{ref-QK8RfvTZ}.
 
@@
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
-For instance, the firing rate is highly irregular {[}\protect\hyperlink{ref-sYHyiGsP}{26}{]}, which makes it inconsistent with temporal integration of random EPSPs.
-It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation {[}\protect\hyperlink{ref-XsHCA4IZ}{15}{]} (see Figure {[}\ref{fig:mainen}{]}).
+For instance, the firing rate is highly irregular~\citep{ref-sYHyiGsP}, which makes it inconsistent with temporal integration of random EPSPs.
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{ref-XsHCA4IZ} (see Figure {[}\ref{fig:mainen}{]}).
 When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
-This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in {[}\protect\hyperlink{ref-1AbvCTG03}{27}{]} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
+This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in~\citep{ref-1AbvCTG03} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
 However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
@@
 
-In his book ``Corticonics'' {[}\protect\hyperlink{ref-19MS6Bbwp}{28}{]}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
+In his book ``Corticonics''~\citep{ref-19MS6Bbwp}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
 The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models.
-While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code {[}\protect\hyperlink{ref-lT1Ato7P}{29}{]}.
+While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{ref-lT1Ato7P}.
 The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion.
-Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons {[}\protect\hyperlink{ref-R2E2wRFg}{32}{]}.
-
-It was shown that a simple model may allow the propagation of such synfire chains {[}\protect\hyperlink{ref-xbqYi2PD}{33}{]}.
+Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{ref-R2E2wRFg}.
+
+It was shown that a simple model may allow the propagation of such synfire chains~\citep{ref-xbqYi2PD}.
 This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size.
@@
 A simple simulation is shown in Figure {[}\ref{fig:diesman}{]}.
-A crucial aspect of this emergence is in particular the balance between excitation and inhibition {[}\protect\hyperlink{ref-DL2NpHxr}{34}{]}.
-This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex {[}\protect\hyperlink{ref-mszJ8wdv}{35}{]}.
+A crucial aspect of this emergence is in particular the balance between excitation and inhibition~\citep{ref-DL2NpHxr}.
+This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{ref-mszJ8wdv}.
 
@@
 
-Attempts have been made to detect such synfire chains in neurobiological data {[}\protect\hyperlink{ref-qVu1OkoS}{36}{]}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies {[}\protect\hyperlink{ref-xBZqHOuN}{37}{]} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation {[}\protect\hyperlink{ref-p7xk5w4f}{38}{]}.
-In particular, this was implemented using the pyNN language {[}\protect\hyperlink{ref-33SRV9MB}{39}{]} both in CPU-based and neuromorphic hardware {[}\protect\hyperlink{ref-DiIieN8s}{40}{]} (see Figure {[}\ref{fig:pynn}{]}).
+Attempts have been made to detect such synfire chains in neurobiological data~\citep{ref-qVu1OkoS}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{ref-xBZqHOuN} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{ref-p7xk5w4f}.
+In particular, this was implemented using the pyNN language~\citep{ref-33SRV9MB} both in CPU-based and neuromorphic hardware~\citep{ref-DiIieN8s} (see Figure {[}\ref{fig:pynn}{]}).
 
@@
 Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
-However, this pathway is sufficient to drive the cortex as this input is more often synchronously active {[}\protect\hyperlink{ref-18ZU9zP9l}{41}{]}.
-
-Some experimental results show the emergence of synchrony, for instance in motor cortical function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}.
+However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{ref-18ZU9zP9l}.
+
+Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{ref-wVZpQvSk}.
 Interestingly, authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
 Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
-These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task {[}\protect\hyperlink{ref-1HJjmmSpm}{43}{]} and showed some tuning to to movement direction and reaction time {[}\protect\hyperlink{ref-FZ9X31Qa}{44}{]}.
-It is important to note that synchronous events tended to lock to LFP beta waves {[}\protect\hyperlink{ref-Ejy1sAyh}{45}{]}, and was extended to larger assemblies {[}\protect\hyperlink{ref-Glu4NlLC}{47}{]} using statistical methods (see {[}\protect\hyperlink{ref-sec:detection}{\textbf{sec:detection?}}{]}).
-
-In {[}\protect\hyperlink{ref-3GlNNxH5}{48}{]}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{ref-1HJjmmSpm} and showed some tuning to to movement direction and reaction time~\citep{ref-FZ9X31Qa}.
+It is important to note that synchronous events tended to lock to LFP beta waves~\citep{ref-Ejy1sAyh}, and was extended to larger assemblies~\citep{ref-Glu4NlLC} using statistical methods (see~\citep{ref-sec:detection}{\textbf{sec:detection?}}{]}).
+
+In~\citep{ref-3GlNNxH5}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
 It relies on the simple remark that a series of repeated coincidences is in itself an invariant.
@@
 \subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}
-TODO : {[}\protect\hyperlink{ref-zNDCo9Hy}{49}{]}
+TODO :~\citep{ref-zNDCo9Hy}
 
 Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses {[}\protect\hyperlink{ref-Piao2Wzf}{50}{]}.
+This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{ref-Piao2Wzf}.
 This show precise spatio-temporal differences in this recurrently connected assembly.
 The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
-A similar dataset used in {[}\protect\hyperlink{ref-2AsQOOsE}{52}{]} is available from M.
-Berry's lab {[}\protect\hyperlink{ref-1FU0Y7Erc}{53}{]}.
+A similar dataset used in~\citep{ref-2AsQOOsE} is available from M.
+Berry's lab~\citep{ref-1FU0Y7Erc}.
 
@@
 One possibility is that these waves are systematically missed in trial-averaged data, due to variability.
-A recent work {[}\protect\hyperlink{ref-hGYLURVS}{54}{]} presents a method for detecting propagating waves in noisy multichannel recordings.
+A recent work~\citep{ref-hGYLURVS} presents a method for detecting propagating waves in noisy multichannel recordings.
 Applying this method to single-trial voltage-sensitive dye imaging data, they show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
@@
 
-More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex {[}\protect\hyperlink{ref-vsKvZJ82}{55}{]}.
+More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex~\citep{ref-vsKvZJ82}.
 Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
-In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states {[}\protect\hyperlink{ref-BNK6T5Uk}{9}{]}.
+In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{ref-BNK6T5Uk}.
 Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise.
@@
 
-{[}\protect\hyperlink{ref-17EU8fPBv}{56}{]} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
-
-\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas {[}\protect\hyperlink{ref-yLVM9IXi}{57}{]}.
+~\citep{ref-17EU8fPBv} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+
+\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{ref-yLVM9IXi}.
 It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
@@
 
-{[}\protect\hyperlink{ref-OAorhMRa}{58}{]}
+~\citep{ref-OAorhMRa}
 
@@
 \item
-  {[}\protect\hyperlink{ref-HjDb1gnq}{59}{]} : from synfire chains to Synfire braids
-\end{itemize}
-
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays {[}\protect\hyperlink{ref-KS4oxnst}{60},\protect\hyperlink{ref-ga3ZTFC7}{61}{]}.
-Notably, Izhikevich {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
+ ~\citep{ref-HjDb1gnq} : from synfire chains to Synfire braids
+\end{itemize}
+
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{ref-KS4oxnst}{60},\protect\hyperlink{ref-ga3ZTFC7}.
+Notably, Izhikevich~\citep{ref-SM9G0xBK} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
 This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
@@
 \item
-  A notable exception is the polychronization model of Izhikevich {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+  A notable exception is the polychronization model of Izhikevich~\citep{ref-SM9G0xBK}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
 In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
@@
 Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
-Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains {[}\protect\hyperlink{ref-Mf6h1KNY}{63}{]}, packets {[}\protect\hyperlink{ref-2GovoaYw}{64}{]} or hippocampal sequences {[}\protect\hyperlink{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou}{66}{]} and on the role of such precise spike timing in downstream information transfer and coding {[}\protect\hyperlink{ref-19fAM6hcv}{67},\protect\hyperlink{ref-13dQ0R4UM}{68},\protect\hyperlink{ref-MfmcMsqA}{69}{]}.
-This is for instance relevant in sensory pathways in vision {[}\protect\hyperlink{ref-1EHZgcKyL}{70}{]}, audition {[}\protect\hyperlink{ref-FlKJw2VL}{71}{]}, olfaction {[}\protect\hyperlink{ref-AhSRXBJ2}{72}{]}{]} or touch {[}\protect\hyperlink{ref-gDinei4a}{73}{]}.
+Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ref-Mf6h1KNY}, packets~\citep{ref-2GovoaYw} or hippocampal sequences~\citep{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou} and on the role of such precise spike timing in downstream information transfer and coding~\citep{ref-19fAM6hcv}{67},\protect\hyperlink{ref-13dQ0R4UM}{68},\protect\hyperlink{ref-MfmcMsqA}.
+This is for instance relevant in sensory pathways in vision~\citep{ref-1EHZgcKyL}, audition~\citep{ref-FlKJw2VL}, olfaction~\citep{ref-AhSRXBJ2}{]} or touch~\citep{ref-gDinei4a}.
 In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
@@
 
-Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space {[}\protect\hyperlink{ref-13dQ0R4UM}{68}{]}.
-These sequences can be internally generated {[}\protect\hyperlink{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou}{66}{]} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
+Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{ref-13dQ0R4UM}.
+These sequences can be internally generated~\citep{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
 Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from {[}\protect\hyperlink{ref-qI2HUEyy}{74}{]} below).
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from~\citep{ref-qI2HUEyy} below).
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
@@
 
-TODO: read {[}\protect\hyperlink{ref-MfmcMsqA}{69}{]} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
-
-In {[}\protect\hyperlink{ref-4D8Jnhkz}{75}{]}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
+TODO: read~\citep{ref-MfmcMsqA} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
+
+In~\citep{ref-4D8Jnhkz}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
 They develop a HMM model for quantifying the transitions.
@@
 \item
-  Ikegaya and colleagues {[}\protect\hyperlink{ref-Mf6h1KNY}{63}{]} worked on spontaneous activity in vitro and in vivo.
+  Ikegaya and colleagues~\citep{ref-Mf6h1KNY} worked on spontaneous activity in vitro and in vivo.
 They demonstrated that in cortical activity, we can find a repetition of several motifs.
@@
 1.
-from {[}\protect\hyperlink{ref-Mf6h1KNY}{63}{]} Repeated motifs of spontaneous synaptic activity in vitro and in vivo.
+from~\citep{ref-Mf6h1KNY} Repeated motifs of spontaneous synaptic activity in vitro and in vivo.
 (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
@@
 
-It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in {[}\protect\hyperlink{ref-Jzr1gwsr}{76}{]} which uses noise patterns.
+It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{ref-Jzr1gwsr} which uses noise patterns.
 They " used random waveforms to probe the formation of new memories for arbitrary complex sounds.
@@
 
-Such observations suggest that neuronal groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity {[}\protect\hyperlink{ref-UGHHptIH}{77}{]}.
+Such observations suggest that neuronal groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{ref-UGHHptIH}.
 This work shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
@@
 \subsection{outline}\label{outline}
-A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs) {[}\protect\hyperlink{ref-DUpFdm8C}{78}{]}.
+A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{ref-DUpFdm8C}.
 The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
@@
 
-there are temporal delays in the nervous system, both at the neural {[}\protect\hyperlink{ref-f7gkBktf}{79}{]} and behavioral {[}\protect\hyperlink{ref-1BSbXPPDF}{80}{]} levels.
+there are temporal delays in the nervous system, both at the neural~\citep{ref-f7gkBktf} and behavioral~\citep{ref-1BSbXPPDF} levels.
 Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
 
-may be used for motion detection and interpolation {[}\protect\hyperlink{ref-WGEaBYgy}{81}{]} {[}\protect\hyperlink{ref-7GEdzh5o}{82}{]}
-
-Remarkably, novel neuromorphic chips use a representation similar to that of real neurons {[}\protect\hyperlink{ref-Oet1bgdb}{22}{]}.
+may be used for motion detection and interpolation~\citep{ref-WGEaBYgy}~\citep{ref-7GEdzh5o}
+
+Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{ref-Oet1bgdb}.
 For example, event-based cameras provide a stream of binary asynchronous events signaling detectable changes in luminance, and information is represented by these spike-based temporal motifs, hence their name ``silicon retinas'' (see Figure \ref{fig:silicon_retina}).
@@
 
-A popular model for the detection of latency patterns is the tempotron {[}\protect\hyperlink{ref-1GazUBEdT}{83}{]}.
-This model is in particular reviewed in {[}\protect\hyperlink{ref-QCyjy8wE}{84}{]}.
+A popular model for the detection of latency patterns is the tempotron~\citep{ref-1GazUBEdT}.
+This model is in particular reviewed in~\citep{ref-QCyjy8wE}.
 The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
@@
 The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
-It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons {[}\protect\hyperlink{ref-EF6Larp8}{85}{]}.
+It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{ref-EF6Larp8}.
 The limits of this model is that it's output is only binary and that it's storage capacity are limited.
@@
 Some other models of latency motifs detection using STDP learning rules.
-For instance, {[}\protect\hyperlink{ref-Dd3kSDRW}{86}{]} implements a STDP-based spiking deep convolutional neural networks for object recognition; {[}\protect\hyperlink{ref-vePjgmRh}{87}{]} develops a form of spike-based, competitive learning applied for unsupervised learning.
-In {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+For instance,~\citep{ref-Dd3kSDRW} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{ref-vePjgmRh} develops a form of spike-based, competitive learning applied for unsupervised learning.
+In~\citep{ref-RrgzWOSk}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
 
@@
 The term polychronous was first introduced in 2006, by E.
-Izhichevitch in {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]}.
+Izhichevitch in~\citep{ref-SM9G0xBK}.
 He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
@@
 
--\textgreater{} insert fig 2 of {[}\protect\hyperlink{ref-doi}{\textbf{doi?}} 10.1162/089976606775093882{]}
+-\textgreater{} insert fig 2 of~\citep{ref-doi}{\textbf{doi?}} 10.1162/089976606775093882{]}
 
@@
 In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
-We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.'' {[}\protect\hyperlink{ref-KS4oxnst}{60}{]}.
+We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{ref-KS4oxnst}.
 Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
 
-Memory traces in dynamical systems {[}\protect\hyperlink{ref-o7rANZux}{89}{]} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+Memory traces in dynamical systems~\citep{ref-o7rANZux} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
 It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
@@
 
-Paper by {[}\protect\hyperlink{ref-LjkTSK3h}{90}{]} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
-
-Weyl's discrepency measure {[}\protect\hyperlink{ref-RYzP72lj}{91}{]} which may lead to the definition of a cross-correlation.
+Paper by~\citep{ref-LjkTSK3h} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
+
+Weyl's discrepency measure~\citep{ref-RYzP72lj} which may lead to the definition of a cross-correlation.
 
@@
 
-A review of the methods for neuronal response latency estimation including bayesian binning {[}\protect\hyperlink{ref-Z7LAUup3}{92}{]}.
+A review of the methods for neuronal response latency estimation including bayesian binning~\citep{ref-Z7LAUup3}.
 
@@
 In generic linear non linear lnl models, the output is assumed to be poisson.
-As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression {[}\protect\hyperlink{ref-MgINa6bU}{93}{]}.
+As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{ref-MgINa6bU}.
 This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
 
-Unitary event analysis is performed by a statistical model of coincidence detection {[}\protect\hyperlink{ref-13iJohyPk}{94}{]}.
-This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]} for instance).
-
-Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity {[}\protect\hyperlink{ref-1ErNTONy0}{95}{]}.
-By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range {[}\protect\hyperlink{ref-V9qkJFZU}{96}{]}.
-
-This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents'' {[}\protect\hyperlink{ref-pExXxdMb}{97}{]} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations {[}\protect\hyperlink{ref-d2m7wQpH}{98}{]}.
-
-{[}\protect\hyperlink{ref-doi}{\textbf{doi?}}{]}
+Unitary event analysis is performed by a statistical model of coincidence detection~\citep{ref-13iJohyPk}.
+This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{ref-wVZpQvSk} for instance).
+
+Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{ref-1ErNTONy0}.
+By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range~\citep{ref-V9qkJFZU}.
+
+This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{ref-pExXxdMb} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{ref-d2m7wQpH}.
+
+~\citep{ref-doi}{\textbf{doi?}}{]}
 
@@
 The potentials and limitations of the efficient coding principle are discussed.
-{[}\protect\hyperlink{ref-s63GcujC}{99}{]}
+~\citep{ref-s63GcujC}
 
@@
 \item
-  described in {[}\protect\hyperlink{ref-zc16plUu}{100}{]}
+  described in~\citep{ref-zc16plUu}{100}{]}
 \item
@@
 
-\subsubsection{\texorpdfstring{Stringer et al 2019, Nature {[}\protect\hyperlink{ref-brg2FFsb}{101}{]}}{Stringer et al 2019, Nature {[}101{]}}}\label{stringer-et-al-2019-nature-stringer2019nature}
+\subsubsection{\texorpdfstring{Stringer et al 2019, Nature~\citep{ref-brg2FFsb}{101}{]}}{Stringer et al 2019, Nature {[}101{]}}}\label{stringer-et-al-2019-nature-stringer2019nature}
 \begin{itemize}
@@
 
-\subsubsection{\texorpdfstring{Stringer et al 2019, Science {[}\protect\hyperlink{ref-1BPHUQLTf}{102}{]}}{Stringer et al 2019, Science {[}102{]}}}\label{stringer-et-al-2019-science-stringer2019science}
+\subsubsection{\texorpdfstring{Stringer et al 2019, Science~\citep{ref-1BPHUQLTf}{102}{]}}{Stringer et al 2019, Science {[}102{]}}}\label{stringer-et-al-2019-science-stringer2019science}
 \begin{itemize}
@@
 \item
-  see also the work showing that you can encode very precise orientation information by using many neurons: {[}\protect\hyperlink{ref-1EMs3GS0B}{103}{]}
+  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{ref-1EMs3GS0B}{103}{]}
 \end{itemize}
@@
 \subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
-\subsubsection{\texorpdfstring{Paper by Grossberger, 2018 {[}\protect\hyperlink{ref-9AK1cwY1}{104}{]}}{Paper by Grossberger, 2018 {[}104{]}}}\label{paper-by-grossberger-2018-grossberger2018}
+\subsubsection{\texorpdfstring{Paper by Grossberger, 2018~\citep{ref-9AK1cwY1}{104}{]}}{Paper by Grossberger, 2018 {[}104{]}}}\label{paper-by-grossberger-2018-grossberger2018}
 \begin{itemize}
@@
 \includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
-\caption{Fig 1 of {[}\protect\hyperlink{ref-9AK1cwY1}{104}{]}: ``Simulated example illustrating the steps in SPOTDisClust.
+\caption{Fig 1 of~\citep{ref-9AK1cwY1}{104}{]}: ``Simulated example illustrating the steps in SPOTDisClust.
 A) Structure of five ``ground-truth'' patterns (\ldots).
@@
 
-\subsubsection{\texorpdfstring{Paper by Russo et al 2017 {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}}{Paper by Russo et al 2017 {[}88{]}}}\label{paper-by-russo-et-al-2017-russo2017}
+\subsubsection{\texorpdfstring{Paper by Russo et al 2017~\citep{ref-RrgzWOSk}}{Paper by Russo et al 2017 {[}88{]}}}\label{paper-by-russo-et-al-2017-russo2017}
 \begin{itemize}
@@
 \item
-  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by {[}\protect\hyperlink{ref-RrgzWOSk}{88}{]}
-\item
-  sliding window as in {[}\protect\hyperlink{ref-1XpTamcl}{105}{]} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags {[}\protect\hyperlink{ref-bc9czTxi}{46}{]}
+  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{ref-RrgzWOSk}
+\item
+  sliding window as in~\citep{ref-1XpTamcl}{105}{]} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{ref-bc9czTxi}
 \item
@@
 
-\subsubsection{\texorpdfstring{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}\protect\hyperlink{ref-1300cS9RL}{106}{]}}{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}106{]}}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
+\subsubsection{\texorpdfstring{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{ref-1300cS9RL}{106}{]}}{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}106{]}}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
 \begin{itemize}
@@
 
-To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters, {[}\protect\hyperlink{ref-1365b3Wwl}{107}{]} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
+To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{ref-1365b3Wwl}{107}{]} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
 They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
@@
 uses a FPGA
-{[}\protect\hyperlink{ref-rYLmAESf}{108}{]}
+~\citep{ref-rYLmAESf}{108}{]}
 
@@
 Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
-{[}\protect\hyperlink{ref-1DadetlqZ}{109}{]}
+~\citep{ref-1DadetlqZ}{109}{]}
 
 \subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
-{[}\protect\hyperlink{ref-RrgzWOSk}{88}{]} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
-It is based on unitary events in multiple single-neuron spiking activity {[}\protect\hyperlink{ref-iNkKTYaI}{110}{]} and the reliable and efficient analysis of an excess or deficiency of joint-spike events {[}\protect\hyperlink{ref-14wpaupd5}{111}{]}.
+~\citep{ref-RrgzWOSk} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
+It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI}{110}{]} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}{111}{]}.
 Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
@@
 \subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
-spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP) {[}\protect\hyperlink{ref-19NyVq80B}{112}{]} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
+spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{ref-19NyVq80B}{112}{]} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
 This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach.
@@
 \item
-  {[}\protect\hyperlink{ref-11p3NXmkg}{113}{]} : coherence detection
-\item
-  {[}\protect\hyperlink{ref-j5xS3owo}{114}{]} : STDP
-\item
-  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks {[}\protect\hyperlink{ref-15TktA0gR}{115}{]} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
-PLoS Comput Biol 9(2): e1002897.
-{[}\protect\hyperlink{ref-1FHIKpizQ}{116}{]} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+ ~\citep{ref-11p3NXmkg}{113}{]} : coherence detection
+\item
+ ~\citep{ref-j5xS3owo}{114}{]} : STDP
+\item
+  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR}{115}{]} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
+PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ}{116}{]} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 \end{itemize}
@@
 Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
-This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.
-{[}\protect\hyperlink{ref-s09KZ8Kh}{117}{]}
+This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{ref-s09KZ8Kh}{117}{]}
 
@@
 
-{[}\protect\hyperlink{ref-UeTLXQCp}{118}{]}
+~\citep{ref-UeTLXQCp}{118}{]}
 
@@
 
-Integrating synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes, that is referred to as synaptic delay-weight plasticity {[}\protect\hyperlink{ref-ga3ZTFC7}{61}{]}.
+Integrating synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes, that is referred to as synaptic delay-weight plasticity~\citep{ref-ga3ZTFC7}.
 Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
 
-In a recent paper {[}\protect\hyperlink{ref-B493Fl4K}{119}{]}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
+In a recent paper~\citep{ref-B493Fl4K}{119}{]}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
 In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
 
-In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity {[}\protect\hyperlink{ref-pPRQwZb}{120}{]}.
+In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{ref-pPRQwZb}{120}{]}.
 This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
@@
 \subsection{Learning sequences}\label{learning-sequences}
-In {[}\protect\hyperlink{ref-WFwUKZb5}{121}{]}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds'' {[}\protect\hyperlink{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga}{124}{]} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+In~\citep{ref-WFwUKZb5}{121}{]}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga}{124}{]} and this model shows how one could vary HVC activity using something similar to dropout in ML.
 Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
@@
 \item
-  in {[}\protect\hyperlink{ref-Jzr1gwsr}{76}{]}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
-\item
-  In Bellec {[}\protect\hyperlink{ref-aj6TgorR}{125}{]}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+  in~\citep{ref-Jzr1gwsr}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
+\item
+  In Bellec~\citep{ref-aj6TgorR}{125}{]}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
 
@@
   \item
-    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron {[}\protect\hyperlink{ref-14e3BCBii}{126}{]} with recurrent dynamics)
+    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{ref-14e3BCBii}{126}{]} with recurrent dynamics)
   \item
@@
 
-A recent work {[}\protect\hyperlink{ref-10EU8MfxC}{127}{]} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
+A recent work~\citep{ref-10EU8MfxC}{127}{]} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
 The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
@@
 Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
-Pang R, Fairhall AL.Elife. 2019 May {[}\protect\hyperlink{ref-12zyUFdO4}{128}{]}
+Pang R, Fairhall AL.Elife. 2019 May~\citep{ref-12zyUFdO4}{128}{]}
 
@@
 \subsection{dynamical models}\label{dynamical-models}
-Dumas and colleagues {[}\protect\hyperlink{ref-WxQdcMB2}{129}{]} : three levels / fourth paradigm {[}\protect\hyperlink{ref-JrZPoxk9}{130}{]} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
+Dumas and colleagues~\citep{ref-WxQdcMB2}{129}{]} : three levels / fourth paradigm~\citep{ref-JrZPoxk9}{130}{]} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
 
@@
 
-Emergence of electronic architectures specialized in Sparse Event-Based Convolutions {[}\protect\hyperlink{ref-sJ03d2Zj}{131}{]} or of sparsity aware algorithms {[}\protect\hyperlink{ref-R5BYFgCP}{132}{]}.
+Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj}{131}{]} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}{132}{]}.
 
@@
 \section{development}\label{development}
-scaffolding of neural assemblies / existence of critical periods : {[}\protect\hyperlink{ref-1HSyNOl48}{133}{]}
+scaffolding of neural assemblies / existence of critical periods :~\citep{ref-1HSyNOl48}{133}{]}
 

### 2022-10-04 manuscript.tex
Update manuscript.tex
importing from manubot : compiles ! :-O

@@
 %\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
-\caption{a Spike raster for repeated presentations (N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
+\caption{a Spike raster for repeated presentations (%N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
 A single-trial LFP trace is plotted in gray, and the average LFP response is plotted in black.
 The relative power between baseline (−200 ms to stimulus-onset) and evoked fluctuations (stimulus-onset + 50--250 ms) significantly favored the evoked response (right panel; N = 110 trials; median = 1.89 dB, p = 0.000019, two-tailed Wilcoxon's ranked-sum test).
-b Same as in (a), but for a low contrast stimulus (\textless2\% Michelson contrast).
+b Same as in (a), but for a low contrast stimulus (2\% Michelson contrast).
 The relative power between baseline and evoked LFP fluctuations was not statistically different from parity (median = 1.23 dB, p = 0.087 two-tailed Wilcoxon's ranked-sum test).
@@
 d Histogram of spontaneous spike probability as a function of the generalized phase of the LFP during fixation.
-e The average evoked response to low contrast stimuli was stronger when a more excitable phase (±π rad) of a spontaneous traveling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test).
-Data for panels c--e modified from Davis et al.17 with permission.}\label{fig:davis}
+e The average evoked response to low contrast stimuli was stronger when a more excitable phase (+ / - \pi rad) of a spontaneous traveling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test).
+Data for panels c--e modified from Davis et al.17 with permission.
+}\label{fig:davis}
 \end{figure}
@@
 \begin{itemize}
-\tightlist
  \item
@@
 Two (i) and three (ii) repetitions of motifs are shown.
-Shuffle tests were performed on traces (i), a to c, yielding significantly fewer repeats (fig.~S2, P \textless{} 0.02).
+Shuffle tests were performed on traces (i), a to c, yielding significantly fewer repeats (fig.~S2, P  0.02).
 In (i), the blue trace is shifted --2.75 mV; in (ii), the blue trace is shifted --1.58 mV, and the green +0.79 mV.}\label{fig:Ikegaya2004}
@@
   \begin{itemize}
-  \tightlist
      \item
@@
 A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
-The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed √𝑁.
+The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed $\sqrt N$.
 This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
@@
 \begin{itemize}
-\tightlist
  \item
   ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
-This becomes a major problem for computing an MM dissimilarity matrix (for M time epochs) for thousands of neurons.
+This becomes a major problem for computing an M dissimilarity matrix (for M time epochs) for thousands of neurons.
 (2) It relies exclusively on pairwise spike-timing relationships (i.e 2nd order correlations), because it does not solve the optimal transport problem for the entire spike pattern, but only for neuron pairs separately.

### 2022-10-04 manuscript.tex
Update manuscript.tex

@@
 
-\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
+\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
 It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{ref-yLVM9IXi}.
@@
 \item
-  described in~\citep{ref-zc16plUu}{100}{]}
+  described in~\citep{ref-zc16plUu}
 \item
@@
 
-\subsubsection{\texorpdfstring{Stringer et al 2019, Nature~\citep{ref-brg2FFsb}{101}{]}}{Stringer et al 2019, Nature {[}101{]}}}\label{stringer-et-al-2019-nature-stringer2019nature}
+\subsubsection{Stringer et al 2019, Nature~\citep{ref-brg2FFsb}}\label{stringer-et-al-2019-nature-stringer2019nature}
 \begin{itemize}
@@
 
-\subsubsection{\texorpdfstring{Stringer et al 2019, Science~\citep{ref-1BPHUQLTf}{102}{]}}{Stringer et al 2019, Science {[}102{]}}}\label{stringer-et-al-2019-science-stringer2019science}
+\subsubsection{Stringer et al 2019, Science~\citep{ref-1BPHUQLTf}}\label{stringer-et-al-2019-science-stringer2019science}
 \begin{itemize}
@@
 \item
-  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{ref-1EMs3GS0B}{103}{]}
+  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{ref-1EMs3GS0B}
 \end{itemize}
@@
 \subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
-\subsubsection{\texorpdfstring{Paper by Grossberger, 2018~\citep{ref-9AK1cwY1}{104}{]}}{Paper by Grossberger, 2018 {[}104{]}}}\label{paper-by-grossberger-2018-grossberger2018}
+\subsubsection{Paper by Grossberger, 2018~\citep{ref-9AK1cwY1}}\label{paper-by-grossberger-2018-grossberger2018}
 \begin{itemize}
@@
 \includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
-\caption{Fig 1 of~\citep{ref-9AK1cwY1}{104}{]}: ``Simulated example illustrating the steps in SPOTDisClust.
+\caption{Fig 1 of~\citep{ref-9AK1cwY1}: ``Simulated example illustrating the steps in SPOTDisClust.
 A) Structure of five ``ground-truth'' patterns (\ldots).
@@
 
-\subsubsection{\texorpdfstring{Paper by Russo et al 2017~\citep{ref-RrgzWOSk}}{Paper by Russo et al 2017 {[}88{]}}}\label{paper-by-russo-et-al-2017-russo2017}
+\subsubsection{Paper by Russo et al 2017~\citep{ref-RrgzWOSk}}\label{paper-by-russo-et-al-2017-russo2017}
 \begin{itemize}
@@
 \item
-  sliding window as in~\citep{ref-1XpTamcl}{105}{]} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{ref-bc9czTxi}
+  sliding window as in~\citep{ref-1XpTamcl} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{ref-bc9czTxi}
 \item
@@
 
-\subsubsection{\texorpdfstring{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{ref-1300cS9RL}{106}{]}}{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}106{]}}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
+\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{ref-1300cS9RL}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
 \begin{itemize}
@@
 
-To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{ref-1365b3Wwl}{107}{]} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
+To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{ref-1365b3Wwl} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
 They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
@@
 uses a FPGA
-~\citep{ref-rYLmAESf}{108}{]}
+~\citep{ref-rYLmAESf}
 
@@
 Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
-~\citep{ref-1DadetlqZ}{109}{]}
+~\citep{ref-1DadetlqZ}
 
@@
 ~\citep{ref-RrgzWOSk} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
-It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI}{110}{]} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}{111}{]}.
+It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}.
 Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
@@
 \subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
-spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{ref-19NyVq80B}{112}{]} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
+spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{ref-19NyVq80B} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
 This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach.
@@
  \item
- ~\citep{ref-11p3NXmkg}{113}{]} : coherence detection
-\item
- ~\citep{ref-j5xS3owo}{114}{]} : STDP
-\item
-  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR}{115}{]} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
-PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ}{116}{]} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+ ~\citep{ref-11p3NXmkg} : coherence detection
+\item
+ ~\citep{ref-j5xS3owo} : STDP
+\item
+  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
+PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 \end{itemize}
@@
 Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
-This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{ref-s09KZ8Kh}{117}{]}
+This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{ref-s09KZ8Kh}
 
@@
 
-~\citep{ref-UeTLXQCp}{118}{]}
+~\citep{ref-UeTLXQCp}
 
@@
 
-In a recent paper~\citep{ref-B493Fl4K}{119}{]}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
+In a recent paper~\citep{ref-B493Fl4K}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
 In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
 
-In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{ref-pPRQwZb}{120}{]}.
+In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{ref-pPRQwZb}.
 This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
@@
 \subsection{Learning sequences}\label{learning-sequences}
-In~\citep{ref-WFwUKZb5}{121}{]}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga}{124}{]} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+In~\citep{ref-WFwUKZb5}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga} and this model shows how one could vary HVC activity using something similar to dropout in ML.
 Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
@@
 \item
-  In Bellec~\citep{ref-aj6TgorR}{125}{]}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
-
-  \begin{itemize}
-     \item
-    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{ref-14e3BCBii}{126}{]} with recurrent dynamics)
+  In Bellec~\citep{ref-aj6TgorR}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+
+  \begin{itemize}
+     \item
+    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{ref-14e3BCBii} with recurrent dynamics)
   \item
@@
 
-A recent work~\citep{ref-10EU8MfxC}{127}{]} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
+A recent work~\citep{ref-10EU8MfxC} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
 The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
@@
 Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
-Pang R, Fairhall AL.Elife. 2019 May~\citep{ref-12zyUFdO4}{128}{]}
+Pang R, Fairhall AL.Elife. 2019 May~\citep{ref-12zyUFdO4}
 
@@
 \subsection{dynamical models}\label{dynamical-models}
-Dumas and colleagues~\citep{ref-WxQdcMB2}{129}{]} : three levels / fourth paradigm~\citep{ref-JrZPoxk9}{130}{]} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
+Dumas and colleagues~\citep{ref-WxQdcMB2} : three levels / fourth paradigm~\citep{ref-JrZPoxk9} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
 
@@
 
-Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj}{131}{]} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}{132}{]}.
+Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}.
 
@@
 \section{development}\label{development}
-scaffolding of neural assemblies / existence of critical periods :~\citep{ref-1HSyNOl48}{133}{]}
+scaffolding of neural assemblies / existence of critical periods :~\citep{ref-1HSyNOl48}
 

### 2022-10-04 manuscript.tex
Update on Overleaf.

@@
 %=================================================================
-\documentclass[brainsci,review,submit,pdftex,moreauthors,draft]{Definitions/mdpi} 
+\documentclass[brainsci,review,submit,pdftex,moreauthors]{Definitions/mdpi} 
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
-Its first experiment consisted in asking subjects to categorize images that do or do not contain animals~\citep{ref-G3GlupPg}.
+Its first experiment consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
 The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neuronal activity.
-These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade~\citep{ref-B9iZWoq3}.
+These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
 
@@
 The performances decrease progressively as the frequency of presentation of the images increases.
-However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image~\citep{ref-EqCCVwl2}.
+However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image~\citep{keysers_speed_2001}.
 
 This speed of the visual cortex, although surprising, is quite compatible with the latencies that are recorded at the neuro-physiological level.
-Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque~\citep{ref-Vu87UU6A}, see Figure {[}\ref{fig:thorpe}{]}.
-This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{ref-15aMjRHVK}.
+Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque~\citep{nowak_timing_1997}, see Figure {[}\ref{fig:thorpe}{]}.
+This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/visual-latency-estimate.jpg}
-\caption{From~\citep{ref-12dWDpY61}}\label{fig:thorpe}
+\caption{From~\citep{thorpe_seeking_2001}}\label{fig:thorpe}
 \end{figure}
@@
 Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
-Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{ref-zZUVkTfZ} but also the dependance across neurons~\citep{ref-13T2Gne9z}.
-
-Numerous studies demonstrated the importance of precise timing in neural population activity~\citep{ref-BNK6T5Uk}, efficient encoding thanks to the use of spike latencies~\citep{ref-1DpSM1kpQ}{10},\protect\hyperlink{ref-IjepoonT} or precise timing in the auditory system~\citep{ref-mxPMgOmW}{12},\protect\hyperlink{ref-1ECJhg2GM}.
-All these findings, and more~\citep{ref-WZsEyOlK}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
-Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{ref-XsHCA4IZ}.
+Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependance across neurons~\citep{perkel_neuronal_1967-1}.
+
+Numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
+Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
 
@@
 In this example, one generates at most one spike per neuron.
-From~\citep{ref-YX5wqWH4}}\label{fig:roc}
+%From~\citep{ref-YX5wqWH4}
+}\label{fig:roc}
 \end{figure}
 
-At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{ref-IjepoonT}.
+At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{gollisch_rapid_2008}.
 Notably one can appreciate in figure 1 that the response of ganglion cells to visual gratings that are flashed onto the retina.
@@
 The conclusion of the authors is that the discharge latency of the neurons allows to encode spatial characteristics of the image.
-First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{ref-1H1Omh73g}.
-Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{ref-UlbjMU0W}.
-
-Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{ref-513xqzmL}.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{chase_first-spike_2007}.
+Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+
+Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
 Many models have used these properties in temporal coding to build fast image categorization networks.
-These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{ref-1DpSM1kpQ}.
-This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{ref-7K80jcx9}{20},\protect\hyperlink{ref-fRudGmR1}.
-This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events~\citep{ref-Oet1bgdb}.
-This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{ref-15cR83gqd}.
-Note also that timing is not entirely sensorial or internal but in~\citep{ref-1E8gkynCm}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
-Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{ref-QK8RfvTZ}.
+These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{thorpe_speed_1996}.
+This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
+This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events~\citep{rasetto_challenges_2022}.
+This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
+Note also that timing is not entirely sensorial or internal but in~\citep{safaie_turning_2020}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
 
@@
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
-For instance, the firing rate is highly irregular~\citep{ref-sYHyiGsP}, which makes it inconsistent with temporal integration of random EPSPs.
-It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{ref-XsHCA4IZ} (see Figure {[}\ref{fig:mainen}{]}).
+For instance, the firing rate is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of random EPSPs.
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure {[}\ref{fig:mainen}{]}).
 When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
-This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in~\citep{ref-1AbvCTG03} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
+This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
 However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
@@
 
-In his book ``Corticonics''~\citep{ref-19MS6Bbwp}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
+In his book ``Corticonics''~\citep{abeles_corticonics_1991}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
 The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models.
-While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{ref-lT1Ato7P}.
+While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{paugam-moisy_computing_2012}.
 The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion.
-Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{ref-R2E2wRFg}.
-
-It was shown that a simple model may allow the propagation of such synfire chains~\citep{ref-xbqYi2PD}.
+Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
+
+It was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_1970}.
 This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size.
@@
 A simple simulation is shown in Figure {[}\ref{fig:diesman}{]}.
-A crucial aspect of this emergence is in particular the balance between excitation and inhibition~\citep{ref-DL2NpHxr}.
-This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{ref-mszJ8wdv}.
+A crucial aspect of this emergence is in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
+This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
 
@@
 
-Attempts have been made to detect such synfire chains in neurobiological data~\citep{ref-qVu1OkoS}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{ref-xBZqHOuN} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{ref-p7xk5w4f}.
-In particular, this was implemented using the pyNN language~\citep{ref-33SRV9MB} both in CPU-based and neuromorphic hardware~\citep{ref-DiIieN8s} (see Figure {[}\ref{fig:pynn}{]}).
+Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
+In particular, this was implemented using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure {[}\ref{fig:pynn}{]}).
 
@@
 Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
-However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{ref-18ZU9zP9l}.
-
-Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{ref-wVZpQvSk}.
+However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{bruno_cortex_2006}.
+
+Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}.
 Interestingly, authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
 Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
-These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{ref-1HJjmmSpm} and showed some tuning to to movement direction and reaction time~\citep{ref-FZ9X31Qa}.
-It is important to note that synchronous events tended to lock to LFP beta waves~\citep{ref-Ejy1sAyh}, and was extended to larger assemblies~\citep{ref-Glu4NlLC} using statistical methods (see~\citep{ref-sec:detection}{\textbf{sec:detection?}}{]}).
-
-In~\citep{ref-3GlNNxH5}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to to movement direction and reaction time~\citep{grammont_spike_2003}.
+It is important to note that synchronous events tended to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see~\citep{ref-sec:detection}{\textbf{sec:detection?}}{]}).
+
+In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
 It relies on the simple remark that a series of repeated coincidences is in itself an invariant.
@@
 \subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}
-TODO :~\citep{ref-zNDCo9Hy}
+TODO :~\citep{bringuier_horizontal_1999} latency basin
 
 Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{ref-Piao2Wzf}.
+This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}.
 This show precise spatio-temporal differences in this recurrently connected assembly.
 The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
-A similar dataset used in~\citep{ref-2AsQOOsE} is available from M.
-Berry's lab~\citep{ref-1FU0Y7Erc}.
+A similar dataset used in~\citep{schneidman_weak_2006} is available from M.
+Berry's lab~\citep{berry_spike_2022}.
 
@@
 One possibility is that these waves are systematically missed in trial-averaged data, due to variability.
-A recent work~\citep{ref-hGYLURVS} presents a method for detecting propagating waves in noisy multichannel recordings.
+A recent work~\citep{muller_stimulus-evoked_2014} presents a method for detecting propagating waves in noisy multichannel recordings.
 Applying this method to single-trial voltage-sensitive dye imaging data, they show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
@@
 
-More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex~\citep{ref-vsKvZJ82}.
+More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex~\citep{muller_cortical_2018}.
 Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
-In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{ref-BNK6T5Uk}.
+In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
 Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise.
@@
 
-~\citep{ref-17EU8fPBv} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
-
-\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{ref-yLVM9IXi}.
+~\citep{linden_movement_2021} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+
+\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{riehle_spike_1997}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{grammont_precise_1999}.
 It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
@@
 
-~\citep{ref-OAorhMRa}
+~\citep{chemla_suppressive_2019}
 
@@
  \item
- ~\citep{ref-HjDb1gnq} : from synfire chains to Synfire braids
-\end{itemize}
-
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{ref-KS4oxnst}{60},\protect\hyperlink{ref-ga3ZTFC7}.
-Notably, Izhikevich~\citep{ref-SM9G0xBK} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
+ ~\citep{bienenstock_model_1995} : from synfire chains to Synfire braids
+\end{itemize}
+
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
+Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
 This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
@@
 \item
-  A notable exception is the polychronization model of Izhikevich~\citep{ref-SM9G0xBK}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+  A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
 In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
@@
 Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
-Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ref-Mf6h1KNY}, packets~\citep{ref-2GovoaYw} or hippocampal sequences~\citep{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou} and on the role of such precise spike timing in downstream information transfer and coding~\citep{ref-19fAM6hcv}{67},\protect\hyperlink{ref-13dQ0R4UM}{68},\protect\hyperlink{ref-MfmcMsqA}.
-This is for instance relevant in sensory pathways in vision~\citep{ref-1EHZgcKyL}, audition~\citep{ref-FlKJw2VL}, olfaction~\citep{ref-AhSRXBJ2}{]} or touch~\citep{ref-gDinei4a}.
+Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015} and on the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
+This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996}{]} or touch~\citep{johansson_first_2004}.
 In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
@@
 
-Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{ref-13dQ0R4UM}.
-These sequences can be internally generated~\citep{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
+Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
+These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
 Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from~\citep{ref-qI2HUEyy} below).
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from~\citep{haimerl_internal_2019} below).
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
@@
 
-TODO: read~\citep{ref-MfmcMsqA} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
-
-In~\citep{ref-4D8Jnhkz}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
+TODO: read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
+
+In~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
 They develop a HMM model for quantifying the transitions.
@@
  \item
-  Ikegaya and colleagues~\citep{ref-Mf6h1KNY} worked on spontaneous activity in vitro and in vivo.
+  Ikegaya and colleagues~\citep{ikegaya_synfire_2004} worked on spontaneous activity in vitro and in vivo.
 They demonstrated that in cortical activity, we can find a repetition of several motifs.
@@
 1.
-from~\citep{ref-Mf6h1KNY} Repeated motifs of spontaneous synaptic activity in vitro and in vivo.
+from~\citep{ikegaya_synfire_2004} Repeated motifs of spontaneous synaptic activity in vitro and in vivo.
 (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
@@
 
-It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{ref-Jzr1gwsr} which uses noise patterns.
+It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns.
 They " used random waveforms to probe the formation of new memories for arbitrary complex sounds.
@@
 
-Such observations suggest that neuronal groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{ref-UGHHptIH}.
+Such observations suggest that neuronal groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{miller_visual_2014}.
 This work shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
@@
 \subsection{outline}\label{outline}
-A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{ref-DUpFdm8C}.
+A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
 The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
@@
 
-there are temporal delays in the nervous system, both at the neural~\citep{ref-f7gkBktf} and behavioral~\citep{ref-1BSbXPPDF} levels.
+there are temporal delays in the nervous system, both at the neural~\citep{perrinet_active_2014} and behavioral~\citep{khoei_flash-lag_2017} levels.
 Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
 
-may be used for motion detection and interpolation~\citep{ref-WGEaBYgy}~\citep{ref-7GEdzh5o}
-
-Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{ref-Oet1bgdb}.
+may be used for motion detection and interpolation~\citep{kaplan_anisotropic_2013}~\citep{khoei_flash-lag_2017}
+
+Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}.
 For example, event-based cameras provide a stream of binary asynchronous events signaling detectable changes in luminance, and information is represented by these spike-based temporal motifs, hence their name ``silicon retinas'' (see Figure \ref{fig:silicon_retina}).
@@
 
-A popular model for the detection of latency patterns is the tempotron~\citep{ref-1GazUBEdT}.
-This model is in particular reviewed in~\citep{ref-QCyjy8wE}.
+A popular model for the detection of latency patterns is the tempotron~\citep{gutig_tempotron_2006}.
+This model is in particular reviewed in~\citep{gutig_spike_2014}.
 The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
@@
 The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
-It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{ref-EF6Larp8}.
+It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
 The limits of this model is that it's output is only binary and that it's storage capacity are limited.
@@
 Some other models of latency motifs detection using STDP learning rules.
-For instance,~\citep{ref-Dd3kSDRW} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{ref-vePjgmRh} develops a form of spike-based, competitive learning applied for unsupervised learning.
-In~\citep{ref-RrgzWOSk}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
+In~\citep{russo_cell_2017}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
 
@@
 The term polychronous was first introduced in 2006, by E.
-Izhichevitch in~\citep{ref-SM9G0xBK}.
+Izhichevitch in~\citep{izhikevich_polychronization_2006}.
 He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
@@
 
--\textgreater{} insert fig 2 of~\citep{ref-doi}{\textbf{doi?}} 10.1162/089976606775093882{]}
+-\textgreater{} insert fig 2 of~\citep{izhikevich_polychronization_2006}
 
@@
 In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
-We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{ref-KS4oxnst}.
+We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{guise_bayesian_2014}.
 Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
 
-Memory traces in dynamical systems~\citep{ref-o7rANZux} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
 It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
@@
 
-Paper by~\citep{ref-LjkTSK3h} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
-
-Weyl's discrepency measure~\citep{ref-RYzP72lj} which may lead to the definition of a cross-correlation.
+Paper by~\citep{moser_stability_2014} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
+
+Weyl's discrepency measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation.
 
@@
 
-A review of the methods for neuronal response latency estimation including bayesian binning~\citep{ref-Z7LAUup3}.
+A review of the methods for neuronal response latency estimation including bayesian binning~\citep{levakova_review_2015}.
 
@@
 In generic linear non linear lnl models, the output is assumed to be poisson.
-As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{ref-MgINa6bU}.
+As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
 This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
 
-Unitary event analysis is performed by a statistical model of coincidence detection~\citep{ref-13iJohyPk}.
-This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{ref-wVZpQvSk} for instance).
-
-Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{ref-1ErNTONy0}.
-By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range~\citep{ref-V9qkJFZU}.
-
-This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{ref-pExXxdMb} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{ref-d2m7wQpH}.
-
-~\citep{ref-doi}{\textbf{doi?}}{]}
+Unitary event analysis is performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
+This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance).
+
+Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{torre_statistical_2013}.
+By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range~\citep{quaglio_methods_2018}.
+
+This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
 
@@
 The potentials and limitations of the efficient coding principle are discussed.
-~\citep{ref-s63GcujC}
+~\citep{zhao_understanding_2011}
 
@@
 \item
-  described in~\citep{ref-zc16plUu}
+  described in~\citep{pachitariu_robustness_2018}
 \item
@@
 
-\subsubsection{Stringer et al 2019, Nature~\citep{ref-brg2FFsb}}\label{stringer-et-al-2019-nature-stringer2019nature}
+\subsubsection{Stringer et al 2019, Nature~\citep{stringer_high-dimensional_2019}}\label{stringer-et-al-2019-nature-stringer2019nature}
 \begin{itemize}
@@
 
-\subsubsection{Stringer et al 2019, Science~\citep{ref-1BPHUQLTf}}\label{stringer-et-al-2019-science-stringer2019science}
+\subsubsection{Stringer et al 2019, Science~\citep{stringer_spontaneous_2019}}\label{stringer-et-al-2019-science-stringer2019science}
 \begin{itemize}
@@
 \item
-  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{ref-1EMs3GS0B}
+  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{stringer_high-precision_2021}
 \end{itemize}
@@
 \subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
-\subsubsection{Paper by Grossberger, 2018~\citep{ref-9AK1cwY1}}\label{paper-by-grossberger-2018-grossberger2018}
+\subsubsection{Paper by Grossberger, 2018~\citep{grossberger_unsupervised_2018}}\label{paper-by-grossberger-2018-grossberger2018}
 \begin{itemize}
@@
 \includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
-\caption{Fig 1 of~\citep{ref-9AK1cwY1}: ``Simulated example illustrating the steps in SPOTDisClust.
+\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}: ``Simulated example illustrating the steps in SPOTDisClust.
 A) Structure of five ``ground-truth'' patterns (\ldots).
@@
 
-\subsubsection{Paper by Russo et al 2017~\citep{ref-RrgzWOSk}}\label{paper-by-russo-et-al-2017-russo2017}
-\begin{itemize}
- \item
-  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{ref-RrgzWOSk}
-\item
-  sliding window as in~\citep{ref-1XpTamcl} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{ref-bc9czTxi}
+\subsubsection{Paper by Russo et al 2017~\citep{russo_cell_2017}}\label{paper-by-russo-et-al-2017-russo2017}
+\begin{itemize}
+ \item
+  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{russo_cell_2017}
+\item
+  sliding window as in~\citep{grun_unitary_2002} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{torre_synchronous_2016}
 \item
@@
 
-\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{ref-1300cS9RL}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
+\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
 \begin{itemize}
@@
 
-To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{ref-1365b3Wwl} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
+To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
 They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
@@
 uses a FPGA
-~\citep{ref-rYLmAESf}
+~\citep{dahlem_dynamics_2009}
 
@@
 Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
-~\citep{ref-1DadetlqZ}
+~\citep{thanasoulis_delay-based_2021}
 
 \subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
-~\citep{ref-RrgzWOSk} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
-It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}.
+~\citep{russo_cell_2017} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
+It is based on unitary events in multiple single-neuron spiking activity~\citep{grun_unitary_2002-1} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
 Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
@@
  \item
- ~\citep{ref-11p3NXmkg} : coherence detection
-\item
- ~\citep{ref-j5xS3owo} : STDP
-\item
-  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
-PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+ ~\citep{perrinet_coherence_2002} : coherence detection
+\item
+ ~\citep{perrinet_networks_2001} : STDP
+\item
+  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{gilson_stdp_2010} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
+PLoS Comput Biol 9(2): e1002897.~\citep{kerr_delay_2013} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 \end{itemize}
@@
 Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
-This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{ref-s09KZ8Kh}
+This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{burkitt_predictive_2021}
 
@@
 
-~\citep{ref-UeTLXQCp}
+~\citep{nadafian_bio-plausible_2020}
 
@@
 
-Integrating synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes, that is referred to as synaptic delay-weight plasticity~\citep{ref-ga3ZTFC7}.
+Integrating synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes, that is referred to as synaptic delay-weight plasticity~\citep{zhang_supervised_2020}.
 Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
@@
 
-In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{ref-pPRQwZb}.
+In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{sun_learning_2016}.
 This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
@@
 \subsection{Learning sequences}\label{learning-sequences}
-In~\citep{ref-WFwUKZb5}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
 Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
@@
 \item
-  in~\citep{ref-Jzr1gwsr}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
-\item
-  In Bellec~\citep{ref-aj6TgorR}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
-
-  \begin{itemize}
-     \item
-    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{ref-14e3BCBii} with recurrent dynamics)
+  in~\citep{agus_rapid_2010}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
+\item
+  In Bellec~\citep{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+
+  \begin{itemize}
+     \item
+    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics)
   \item
@@
 
-A recent work~\citep{ref-10EU8MfxC} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
+A recent work~\citep{ghosh_spatiotemporal_2019} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
 The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
@@
 Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
-Pang R, Fairhall AL.Elife. 2019 May~\citep{ref-12zyUFdO4}
+Pang R, Fairhall AL.Elife. 2019 May~\citep{pang_fast_2019}
 
@@
 \subsection{dynamical models}\label{dynamical-models}
-Dumas and colleagues~\citep{ref-WxQdcMB2} : three levels / fourth paradigm~\citep{ref-JrZPoxk9} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
+Dumas and colleagues~\citep{panahi_generative_2021} : three levels / fourth paradigm~\citep{tolle_fourth_2011} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
 
@@
 
-Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}.
+Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{di_mauro_alfio_sne_2022} or of sparsity aware algorithms~\citep{yin_sata_2022}.
 
@@
 \section{development}\label{development}
-scaffolding of neural assemblies / existence of critical periods :~\citep{ref-1HSyNOl48}
+scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2021}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{adjustwidth}{-\extralength}{0cm}
@@
 
+%%%%%%%%% REFERENCES
 \reftitle{References}
@@
 
-%=====================================
-% References, variant A: external bibliography
-%=====================================
-%\bibliography{your_external_BibTeX_file}
-
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-\begin{thebibliography}{999}
-% Reference 1
-\bibitem[Author1(year)]{ref-journal}
-Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-% Reference 2
-\bibitem[Author2(year)]{ref-book1}
-Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-% Reference 3
-\bibitem[Author3(year)]{ref-book2}
-Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
-% Reference 4
-\bibitem[Author4(year)]{ref-unpublish}
-Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
-% Reference 5
-\bibitem[Author5(year)]{ref-communication}
-Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
-% Reference 6
-\bibitem[Author6(year)]{ref-proceeding}
-Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
-Title of presentation.
-In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
-% Reference 7
-\bibitem[Author7(year)]{ref-thesis}
-Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
-% Reference 8
-\bibitem[Author8(year)]{ref-url}
-Title of Site. Available online: URL (accessed on Day Month Year).
-\end{thebibliography}
-
-% If authors have biography, please use the format below
-%\section*{Short Biography of Authors}
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
-%{\textbf{Firstname Lastname} Biography of first author}
-%
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
-%{\textbf{Firstname Lastname} Biography of second author}
+\bibliography{references}
+
+%\begin{thebibliography}{00}
+%\printbibliography[heading=none]
+%\end{thebibliography}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{adjustwidth}

### 2022-10-04 manuscript.tex
Update on Overleaf.
Update manuscript.tex

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% a simpler annotation system:
+\usepackage{soulutf8}
+\usepackage{color}
+\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+%\newcommand{\note}[1]{}
+%
+
 %=================================================================
@@
 % Full title of the paper (Capitalized)
-\Title{Review on precise spiking motifs in neurobiological and neuromorphic data}
+\Title{Precise spiking motifs in neurobiological and neuromorphic data}
 
@@
 Amélie Gruel $^{2,\dagger,\ddagger}$, 
-Jean Martinet $^{2,\ddagger}$ 
+Jean Martinet $^{2,}$ 
 and Laurent Perrinet $^{1,}$\orcidD{}*}
@@
 By definition, a spike, or action potential, is a all-or-none event ---it can occur or not without further details--- at asynchronous timings, i.e.~it can occur at any continuous time - differentially to the discretized timing classically used in digital processing.
-In the living world, neural systems almost systematically use this so-called event-based representation, though we do not yet have a clear idea why.
+In the living world, neural systems almost systematically use this so-called event-based representation.
 A better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data.
 It is also an emerging challenge in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal challenge in the age of the digital economy and of global warming.
-%
-The response of a biological neuron depends largely on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
-This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
-Our goal here is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
-In particular, we will focus on the hypothesis that there exists in an assembly of neurons a representation based on a set of motifs of different relative spike times.
-Here, we will review current litterature on the detection of such motifs in generic raster plots.
-It is work in progress, where anybody interested can \emph{openly} join.
-%
-This hypothesis is directly inspired by neurobiological observations in the hippocampus, and it expands the capabilities of analog representations based on the firing rate by considering a representation based on repetitions of these motifs at precise times of occurrence.
-A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
-We will first review some biological and theoretical evidence in neural information processing.
-We will then present some models for the detection of such motifs in arbitrary raster plots, synthetic, biological or artificial (notably from event-based cameras).
-In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
-Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
 
 \section{Introduction: Importance of precise spike timings in the brain}\label{introduction-importance-of-precise-spike-timings-in-the-brain}
+
+%
+The response of a biological neuron depends largely on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
+This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
+Our goal here is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
+In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
+%Here, we will review current literature on the detection of such motifs in generic raster plots.
+It is work in progress, where anybody interested can \emph{openly} join.
+%
+This hypothesis is directly inspired by neurobiological observations in the hippocampus, and it expands the capabilities of analog representations based on the firing rate by considering a representation based on repetitions of these motifs at precise times of occurrence.
+A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+We will first review some biological and theoretical evidence in neural information processing.
+We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
+In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
+Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
+
 \subsection{Ultra fast neural codes for ultra fast vision}\label{ultra-fast-neural-codes-for-ultra-fast-vision}
@@
 This speed of the visual cortex, although surprising, is quite compatible with the latencies that are recorded at the neuro-physiological level.
-Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque~\citep{nowak_timing_1997}, see Figure {[}\ref{fig:thorpe}{]}.
+Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
 This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/visual-latency-estimate.jpg}
-\caption{From~\citep{thorpe_seeking_2001}}\label{fig:thorpe}
+\includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}
+\caption{From~\citep{thorpe_seeking_2001} \note{get the rights or redo a schematics}}\label{fig:thorpe}
 \end{figure}
@@
 \centering
-\includegraphics{images/roc.jpg}
-\caption{Latency coding.
-An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency.
-In this example, one generates at most one spike per neuron.
-%From~\citep{ref-YX5wqWH4}
+\note{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
+\caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
 }\label{fig:roc}
@@
 
+\section{Precise spike timings in neural assemblies}
+
 \subsection{From synfire chains\ldots{}}\label{from-synfire-chains}
@@
 For instance, the firing rate is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of random EPSPs.
-It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure {[}\ref{fig:mainen}{]}).
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figur
+~\ref{fig:mainen}).
 When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
@@
 \centering
-\includegraphics{figures/MainenSejnowski1995.png} % TODO replace by the simulation
+\includegraphics[width=.7\textwidth]{figures/MainenSejnowski1995.png} % TODO replace by the simulation
 \caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
-While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} nois at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
 See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
@@
 Recursively applying this to a sequence of groups generates either a synfire propagation or not.
-A simple simulation is shown in Figure {[}\ref{fig:diesman}{]}.
+A simple simulation is shown in Figure
+~\ref{fig:diesman}.
 A crucial aspect of this emergence is in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
@@
 Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
-In particular, this was implemented using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure {[}\ref{fig:pynn}{]}).
+In particular, this was implemented using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
 
@@
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tended to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see~\citep{ref-sec:detection}{\textbf{sec:detection?}}{]}).
+It is important to note that synchronous events tended to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see~\ref{sec:detection}).
 
@@
 
-sparse in time and space {[}2{]} AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. {[}3{]} CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
+sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. 
+CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
 
@@
 Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015} and on the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
-This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996}{]} or touch~\citep{johansson_first_2004}.
+This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996} or touch~\citep{johansson_first_2004}.
 In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
@@
 Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from~\citep{haimerl_internal_2019} below).
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure
+~\ref{fig:haimerl} from~\citep{haimerl_internal_2019} below).
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
@@
 \centering
-\includegraphics{images/Ikegaya2004zse0150424620001.jpeg}
+%\includegraphics{figures/Ikegaya2004zse0150424620001.jpeg}
 \caption{Fig.
@@
 
-{[}
 \textbf{TODO: a paragraph on our existing work}
@@
 
-\section{Theoretical models for the detection of precise heterosynaptic spiking motifs}\label{theoretical-models-for-the-detection-of-precise-heterosynaptic-spiking-motifs}
-\#sec:detection
+\section{Models for the detection of precise heterosynaptic spiking motifs}\label{sec:detection}
 
@@
 
-{[}
-\textbf{TODO}
-
--\textgreater{} insert fig 2 of~\citep{izhikevich_polychronization_2006}
+This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance~\citep{wang}.
+
+\begin{figure}
+\centering
+\includegraphics{figures/event_driven_computations.png}
+\caption{\note{insert fig 2 of~\citep{izhikevich_polychronization_2006}}.}\label{fig:PG}
+\end{figure}
 
@@
 
-\section{Detecting precise spiking motifs in biological raster plots}\label{detecting-precise-spiking-motifs-in-biological-raster-plots}
-\{sec:detection\}
-
 \subsection{spike distances}\label{spike-distances}
@@
 
-M.
-C.
-W.
-van Rossum, ``A novel spike distance,'' Neural Comput., vol.~13, no.
-4, pp.~751--763, 2001.
-{[}21{]} D.
-Aronov and J.
-D.
-Victor, ``Non-Euclidean properties of spike train metric spaces,'' Physical Rev.~E (Statist., Nonlinear, Soft Matter Phys.), vol.~69, no.
-6, 2004.
-
-T. Kreuz, J. S. Haas, A. Morelli, H. D. I. Abarbanel, and A. Politi, ``Measuring spike train synchrony,'' J. Neurosci. Methods, vol.~165, no. 1, pp.~151--161, 2007. {[}23{]} H.
+M. C. W. van Rossum, ``A novel spike distance,'' Neural Comput., vol.~13, no. 4, pp.~751--763, 2001.
+
+D. Aronov and J. D. Victor, ``Non-Euclidean properties of spike train metric spaces,'' Physical Rev.~E (Statist., Nonlinear, Soft Matter Phys.), vol.~69, no. 6, 2004.
+
+T. Kreuz, J. S. Haas, A. Morelli, H. D. I. Abarbanel, and A. Politi, ``Measuring spike train synchrony,'' J. Neurosci. Methods, vol.~165, no. 1, pp.~151--161, 2007. 23 H.
 
@@
 \centering
-\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
+%\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
 \caption{Fig 1 of~\citep{grossberger_unsupervised_2018}: ``Simulated example illustrating the steps in SPOTDisClust.
@@
   \item
-    V1-dataset : The dataset we used was collected by Smith and Kohn {[}49{]} and is publicly available at:
+    V1-dataset : The dataset we used was collected by Smith and Kohn 49{and is publicly available at:
     http://crcns.org/data-sets/vc/pvc-11 - it is in a sense supervised with the input being the movie and the output the spikes recorded.

### 2022-10-04 manuscript.tex
Update on Overleaf.
Update manuscript.tex

@@
 \abstract{
-%A single paragraph of about 200 words maximum.
-For research articles, abstracts should give a pertinent overview of the work.
-We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations.
-The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.
+
 Why do neurons communicate through spikes?
@@
 
-\section{Introduction: Importance of precise spike timings in the brain}\label{introduction-importance-of-precise-spike-timings-in-the-brain}
-
-%
-The response of a biological neuron depends largely on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
+\section{Introduction: Importance of precise spike timings in the brain}\label{sec:time}
+%
+The response of a biological neuron largely depends on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
 This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
@@
 
-\subsection{Ultra fast neural codes for ultra fast vision}\label{ultra-fast-neural-codes-for-ultra-fast-vision}
-
+\subsection{Ultra fast neural codes for efficient vision}
+%
 Let us start our review of the state of the art on the role of dynamics in vision by presenting some surprising results that have been obtained in neuroscience.

### 2022-10-04 manuscript.tex
Update manuscript.tex

@@
 
-%--------------------
-% Class Options:
-%--------------------
-%----------
-% journal
-%----------
-% Choose between the following MDPI journals:
-% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
-
-%---------
-% article
-%---------
-% The default type of manuscript is "article", but can be replaced by: 
-% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
-% supfile = supplementary materials
-
 %----------
@@
 Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
-
-%------------------
-% moreauthors
-%------------------
-% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
-
-%---------
-% pdftex
-%---------
-% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
-
 %=================================================================
@@
 %
-
 %=================================================================
@@
 \abstract{
-
 Why do neurons communicate through spikes?
@@
 It is work in progress, where anybody interested can \emph{openly} join.
-%
-This hypothesis is directly inspired by neurobiological observations in the hippocampus, and it expands the capabilities of analog representations based on the firing rate by considering a representation based on repetitions of these motifs at precise times of occurrence.
+
+This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of these motifs at precise times of occurrence.
 A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
@@
 Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
-
+%
 \subsection{Ultra fast neural codes for efficient vision}
@@
 \centering
+\note{get the rights or redo a schematics}
 \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}
-\caption{From~\citep{thorpe_seeking_2001} \note{get the rights or redo a schematics}}\label{fig:thorpe}
+\caption{From~\citep{thorpe_seeking_2001} }\label{fig:thorpe}
 \end{figure}
 
-\subsection{How timing encodes analogous profile}\label{how-timing-encodes-analogous-profile}
+\subsection{How timing encodes analogous profile}
+
 An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief impulses that propagate along the axons of neurons.
@@
 \centering
+\note{insert fig 2 of izhikevich 2006}
 \includegraphics{figures/event_driven_computations.png}
-\caption{\note{insert fig 2 of~\citep{izhikevich_polychronization_2006}}.}\label{fig:PG}
+\caption{blah.}\label{fig:PG}
 \end{figure}
@@
   \item
-    comes with code https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure
-  \item
-    V1-dataset : The dataset we used was collected by Smith and Kohn 49{and is publicly available at:
-    http://crcns.org/data-sets/vc/pvc-11 - it is in a sense supervised with the input being the movie and the output the spikes recorded.
+    comes with code \url{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}
+  \item
+    V1-dataset : The dataset we used was collected by Smith and Kohn 49 and is publicly available at:
+     \url{http://crcns.org/data-sets/vc/pvc-11} - it is in a sense supervised with the input being the movie and the output the spikes recorded.
   \end{itemize}

### 2022-10-05 manuscript.tex
Update manuscript.tex

@@
 
-\subsection{Are there precise heterosynaptic spiking motifs in the brain?}\label{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}
+\subsection{Are there precise heterosynaptic spiking motifs in the brain?}
+%
 Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
@@
 
-\subsection{Izhikevitch's polychronization model}\label{izhikevitchs-polychronization-model}
+\subsection{Izhikevitch's polychronization model}
 Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.

### 2022-10-06 manuscript.tex
Update on Overleaf.

@@
 \begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction: Importance of precise spike timings in the brain}\label{sec:time}
-%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Spiking neurons: integrators our synchrony detectors?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\note{TODO: a paragraph saying that we think of neurons of integrators, but that they may also switch to a mode fro synchrony detection}~\citep{paugam-moisy_computing_2012}
+
 The response of a biological neuron largely depends on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
 This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
-Our goal here is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
+Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
 In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
@@
 We will first review some biological and theoretical evidence in neural information processing.
+The following section will focus on precise spike time motifs.
 We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
@@
 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Ultra fast neural codes for efficient vision}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{How timing encodes analogous profile}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
-Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependance across neurons~\citep{perkel_neuronal_1967-1}.
+Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
 
@@
 
+%----------------------------%
 \begin{figure}
@@
 \end{figure}
-
+%----------------------------%
 At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{gollisch_rapid_2008}.
@@
 First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{chase_first-spike_2007}.
-Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+Interestingly, such fine latency effect can be at the origin of some visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
 
@@
 This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
-Note also that timing is not entirely sensorial or internal but in~\citep{safaie_turning_2020}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Note also that timing is not entirely sensory or internal but in~\citep{safaie_turning_2020}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
 Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Precise spike timings in neural assemblies}
-
-\subsection{From synfire chains\ldots{}}\label{from-synfire-chains}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{From synfire chains\ldots{}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
@@
 However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
-
+%----------------------------%
 \begin{figure}
@@
 \end{figure}
+%----------------------------%
 
@@
 \end{figure}
-
+%----------------------------%
 Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
@@
 \end{figure}
-
+%----------------------------%
 Note also that synchronicity may explain some unintuitive results.
@@
 
-\subsection{Are there precise heterosynaptic spiking motifs in the brain?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{\ldots to precise spiking motifs in the brain?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 
-\subsection{Cortical songs}\label{cortical-songs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Cortical songs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{itemize}
@@
 7).
-
-\subsection{outline}\label{outline}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Precisely timing motifs in neuromorphic engineering}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
@@
 
+Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
+The task consists in localizing a target using only the relative spike timing of eight spatially-separated vibration sensors.
+Haessig2020a
+
 Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}.
@@
 \end{figure}
-
+%----------------------------%
 This section has provided evidence that polychronous groups are an important apsect of information representation in biology with important application in data analysis and neuromorphic engineering.
@@
 
-\section{Models for the detection of precise heterosynaptic spiking motifs}\label{sec:detection}
+%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Models for the detection of precise spiking motifs}\label{sec:detection}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In the following of this section, we will review theoretical models which directly tackle the problem of using precise spiking motifs.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{In computational neuroscience: Izhikevitch's polychronization model}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% dynamic networks \& learning delays:
+
+    Huning H., Glunder H., and Palm G.
+(1998) Synaptic delay learning in pulse-coupled neurons.
+Neural Computation, 10:555--565.
+\url{https://www.deepdyve.com/lp/mit-press/synaptic-delay-learning-in-pulse-coupled-neurons-DGMiNHxp0A}
+    \href{https://www.researchgate.net/publication/37921636_Dynamics_of_Self-Organized_Delay_Adaptation}{Eurich C., Pawelzik K., Ernst U., Cowan J., and Milton J.
+(1999) Dynamics of self-organazed delay adaptation.
+Phys.
+Rev.~Lett., 82:1594--1597.}
 
@@
 
+A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
+In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
+We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{guise_bayesian_2014}.
+Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
+
+% STDP based
+Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
+This raises important questions about the fundamental limits of such memory traces and the properties required of dynamical systems to achieve these limits.
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+We introduce the Fisher Memory Curve (FMC) as a measure of the signal-to-noise ratio (SNR) embedded in the dynamical state relative to the input SNR.
+The integrated FMC indicates the total memory capacity.
+We apply this theory to linear neuronal networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
+A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
+The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed $\sqrt N$.
+This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
+We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.''
+
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
+
 Some other models of latency motifs detection using STDP learning rules.
@@
 
-In the following of this section, we will review theoretical models which directly tackle the problem of using precise heterosynaptic spiking motifs.
-
-\subsection{Izhikevitch's polychronization model}
+% polychronization
+%  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html} Basic idea: different temporal activation of neurons can activate different populations based on conduction delays.
+This allows for a much higher dimensional representation space, and a lot more memory capacity.
+Firing is not synchronous, but it is time-locked = polychrony.
+STDP automatically associates the neurons in the network into polychronous groups.
+There can be far more polychronous groups than neurons.
+He lets the network just settle for 24 hours - and these groups emerge spontaneously from the random connectivity initial conditions.
+The dendritic tree can be used to select the polychronous inputs more precisely.
+Can act as a full feed-forward neural network and make non-linear classifications.
+Two dendritic trees on pyramidal cells - one for the feed-forward input, and one for feed-back/recurrent.
+These can interact such that the feed-back connections are predicting the feed-forward inputs.
+
 Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
@@
 \end{figure}
+%----------------------------%
+
+% polychronization
 
@@
 
-\begin{itemize}
-\item
-  Reproducing Polychronization: A Guide to Maximizing the Reproducibility of Spiking Network Models \url{https://www.frontiersin.org/articles/10.3389/fninf.2018.00046/full}
-
-  \begin{itemize}
-     \item
-    comes with python code
-  \end{itemize}
-\item
-  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html}
-\item
-  dynamic networks \& learning delays:
-
-  \begin{itemize}
-  \item
-    Huning H., Glunder H., and Palm G.
-(1998) Synaptic delay learning in pulse-coupled neurons.
-Neural Computation, 10:555--565.
-\url{https://www.deepdyve.com/lp/mit-press/synaptic-delay-learning-in-pulse-coupled-neurons-DGMiNHxp0A}
-  \item
-    \href{https://www.researchgate.net/publication/37921636_Dynamics_of_Self-Organized_Delay_Adaptation}{Eurich C., Pawelzik K., Ernst U., Cowan J., and Milton J.
-(1999) Dynamics of self-organazed delay adaptation.
-Phys.
-Rev.~Lett., 82:1594--1597.}
-  \item
-    The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining heterosynaptic plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
-  \item
+Wang et al.
+(2013) presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
+
+Reproducing Polychronization: A Guide to Maximizing the Reproducibility of Spiking Network Models \url{https://www.frontiersin.org/articles/10.3389/fninf.2018.00046/full}     comes with python code
+
+    The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
     an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
-  \end{itemize}
-\end{itemize}
-
-A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
-In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
-We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{guise_bayesian_2014}.
-Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
-
-Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
-It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
-This raises important questions about the fundamental limits of such memory traces and the properties required of dynamical systems to achieve these limits.
-We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
-We introduce the Fisher Memory Curve (FMC) as a measure of the signal-to-noise ratio (SNR) embedded in the dynamical state relative to the input SNR.
-The integrated FMC indicates the total memory capacity.
-We apply this theory to linear neuronal networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
-A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
-The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed $\sqrt N$.
-This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
-We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.''
-
-We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
-Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
-
-\subsection{spike distances}\label{spike-distances}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Neurobiology spike distances}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{decoding neural activity/ spike distances}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 J. D. Victor and K. P. Purpura, ``Nature and precision of temporal coding in visual cortex: a metric-space analysis,'' J. Neurophysiol., vol.~76, pp.~1310--1326, Aug.~1996.
@@
 
-\subsection{decoding neural activity}\label{decoding-neural-activity}
 In generic linear non linear lnl models, the output is assumed to be poisson.
@@
 
-\subsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}
+\subsubsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}
 \href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
@@
 
-\subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
-\subsubsection{Paper by Grossberger, 2018~\citep{grossberger_unsupervised_2018}}\label{paper-by-grossberger-2018-grossberger2018}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{detecting structured temporal patterns}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Paper by Grossberger, 2018~\citep{grossberger_unsupervised_2018}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{itemize}
@@
 
-\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
+\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
 \begin{itemize}
@@
 
-\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}\label{dynamics-of-delay-coupled-excitable-neural-systems}
+\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
+~\citep{russo_cell_2017} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
+It is based on unitary events in multiple single-neuron spiking activity~\citep{grun_unitary_2002-1} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
+Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+
+\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}
 uses a FPGA
@@
 
-\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
-~\citep{russo_cell_2017} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
-It is based on unitary events in multiple single-neuron spiking activity~\citep{grun_unitary_2002-1} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
-Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
-spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{ref-19NyVq80B} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{markram_regulation_1997} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
 This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach.
@@
 
-\textbf{TODO amelie} section sur l'apprentissage des délais dans la biologie
+\note{\textbf{TODO amelie} section sur l'apprentissage / la plasticité des délais dans la biologie}
 
@@
 
-In a recent paper~\citep{ref-B493Fl4K}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
+In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
 In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Learning sequences}\label{learning-sequences}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
@@
 
-\subsection{TODO: more bib to read}\label{todo-more-bib-to-read}
 Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
@@
 Ocker GK, Doiron B.Cereb Cortex. 2019.
-
-\textbf{TODO}
 
@@
 \subsubsection{Grimaldi CBMI / PAMI}\label{grimaldi-cbmi-pami}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}\label{discussion}
-\textbf{TODO}
-
-\subsection{dynamical models}\label{dynamical-models}
-Dumas and colleagues~\citep{panahi_generative_2021} : three levels / fourth paradigm~\citep{tolle_fourth_2011} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Spiking neural models with heterogeneous delays}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{our model}\label{our-model}
@@
 
+\subsection{Limits}
+
+Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks in vivo~\citep{Isbister} They discovered (in vivo) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
+% dynamical models
+Dumas and colleagues~\citep{panahi_generative_2021} : three levels / fourth paradigm~\citep{tolle_fourth_2011} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
 Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{di_mauro_alfio_sne_2022} or of sparsity aware algorithms~\citep{yin_sata_2022}.
 
-and oscillations? Journal Article • DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?
-
-\section{development}\label{development}
+and oscillations? Journal Article DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?
+
+% development
 scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2021}
+
+\subsection{Perspectives}
 

### 2022-10-10 manuscript.tex
Update on Overleaf.
1.1

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% In recognition of your well-respected work, we are inviting you to submit one feature contribution (i.e., a long review or research paper) to an MDPI journal free of charge.
+If you wish to take advantage of this opportunity, please provide your information and the name of the journal you would like to submit to: https://www.surveymonkey.com/r/XQZ5FP9 (If you cannot find the journal through the link, please feel free to reply to us about your choice).
+
+% The submission deadline for this special offer is 31 October 2022.
+The journal editorial office will contact you soon once you choose the journal.
+Please contact the revelant journal’s Editorial Office if you require an extension.
+
 %  LaTeX support: latex@mdpi.com 
@@
 %=================================================================
-\documentclass[brainsci,review,submit,pdftex,moreauthors]{Definitions/mdpi} 
+\documentclass[brainsci, %arts,
+               review,submit,pdftex,moreauthors]{Definitions/mdpi} 
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 %
-\note{TODO: a paragraph saying that we think of neurons of integrators, but that they may also switch to a mode fro synchrony detection}~\citep{paugam-moisy_computing_2012}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Is there a neural code?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Neural activity is directly influenced by our neighboring environment, and also by internal states, which in turn allow to generate actions for survival, the ultimate determinant in the light of natural selection.
+A central question in neuroscience is to better understand this relationship.
+This is sometimes coined as ``decoding the neural code'', though this would imply that there would be a code, that is an explicit representation within the neural activity.
+Still, we will use that terminology in all generality.
+In that regard, it is known that some nervous activity is linked to a specific variable, for instance the stretch of a frog's limb (Galvani).
+An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief electrochemical impulses that propagate along the axons of neurons.
+These have the particularity of being essentially binary in their amplitude (that is to say, they are prototypical, all or nothing).
+The frequency of those spikes has been shown to be proportional to the stretch in the frog's limb (Adrian) , and this variable (measured in spikes per seconds, or Hertz) is believed to be the major contributor to the neural code.
+However, computational neuroscience models have hinted that the precise timing may play a crucial role~\citep{paugam-moisy_computing_2012}, an hypothesis which we will review in this manuscript.
 
@@
 
-An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief impulses that propagate along the axons of neurons.
-These have the particularity of being essentially binary in their amplitude (that is to say, they are prototypical, all or nothing).
 An important consequence of the speed of processing is that it implies that it is carried out using only very few spikes.
@@
 \centering
-\includegraphics{figures/event_driven_computations.png}
+%\includegraphics{figures/event_driven_computations.png}
 \caption{A miniature, event-based ATIS sensor.
@@
 \note{insert fig 2 of izhikevich 2006}
-\includegraphics{figures/event_driven_computations.png}
+%includegraphics{figures/event_driven_computations.png}
 \caption{blah.}\label{fig:PG}
@@
 % Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
-
+%\bibliographystyle{Definitions/chicago2}
 \bibliography{references}

### 2022-10-10 manuscript.tex
1.1

@@
 %
-Neural activity is directly influenced by our neighboring environment, and also by internal states, which in turn allow to generate actions for survival, the ultimate determinant in the light of natural selection.
+Neuronal activity is directly influenced by our immediate environment and by internal states, which then generate actions.
+These actions are essential for survival, which is determinant in the light of natural selection.
 A central question in neuroscience is to better understand this relationship.
-This is sometimes coined as ``decoding the neural code'', though this would imply that there would be a code, that is an explicit representation within the neural activity.
-Still, we will use that terminology in all generality.
-In that regard, it is known that some nervous activity is linked to a specific variable, for instance the stretch of a frog's limb (Galvani).
-An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief electrochemical impulses that propagate along the axons of neurons.
-These have the particularity of being essentially binary in their amplitude (that is to say, they are prototypical, all or nothing).
-The frequency of those spikes has been shown to be proportional to the stretch in the frog's limb (Adrian) , and this variable (measured in spikes per seconds, or Hertz) is believed to be the major contributor to the neural code.
-However, computational neuroscience models have hinted that the precise timing may play a crucial role~\citep{paugam-moisy_computing_2012}, an hypothesis which we will review in this manuscript.
+We sometimes use the expression "decoding the neuronal code", although this implies the existence of a code, i.e.
+an explicit representation within the neuronal activity.
+Nevertheless, we will use this terminology in all generality.
+In this respect, it is known that this neural activity is related to a specific variable, for example the stretching of a frog's limb (Galvani).
+A very important characteristic of neural activity is the fact that it consists mainly of action potentials, or spikes, which are brief (about one millisecond) electrochemical impulses that propagate along the axons of neurons.
+These have the particularity of being essentially binary in their amplitude (i.e.
+they are prototypical, all or nothing).
+The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb (Adrian), and this variable (measured in spikes per second, or Hertz) is considered to be the main constituent of the neuronal code.
+However, computational neuroscience models have suggested that precise timing may play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this manuscript.
 

### 2022-10-10 manuscript.tex
1.2

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Spiking neurons: integrators our synchrony detectors?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 A central question in neuroscience is to better understand this relationship.
-We sometimes use the expression "decoding the neuronal code", although this implies the existence of a code, i.e.
+We sometimes use the expression ``decoding the neuronal code'', although this implies the existence of a code, i.e.
 an explicit representation within the neuronal activity.
 Nevertheless, we will use this terminology in all generality.
-In this respect, it is known that this neural activity is related to a specific variable, for example the stretching of a frog's limb (Galvani).
+In this respect, it is known that this neural activity is related to a specific variable, and for example that electrical activity can cause the stretching of a frog's limb~\citep{piccolino_luigi_1997}.
 A very important characteristic of neural activity is the fact that it consists mainly of action potentials, or spikes, which are brief (about one millisecond) electrochemical impulses that propagate along the axons of neurons.
@@
 they are prototypical, all or nothing).
-The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb (Adrian), and this variable (measured in spikes per second, or Hertz) is considered to be the main constituent of the neuronal code.
+The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is considered to be the main constituent of the neuronal code.
 However, computational neuroscience models have suggested that precise timing may play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this manuscript.
 
-The response of a biological neuron largely depends on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
+Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
 This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
+This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of these motifs at precise times of occurrence.
+A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+
 Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
@@
 It is work in progress, where anybody interested can \emph{openly} join.
-
-This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of these motifs at precise times of occurrence.
-A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
-We will first review some biological and theoretical evidence in neural information processing.
-The following section will focus on precise spike time motifs.
+We will first review some biological and theoretical evidence in neural information processing in the particular case of vision.
+The next section will focus on precise spike time motifs.
 We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
@@
 %
-Let us start our review of the state of the art on the role of dynamics in vision by presenting some surprising results that have been obtained in neuroscience.
+Let us start our review of the state of the art on the role of dynamics in vision.
+Vision is the set of processes which allow to make sense of the luminous world, and it is an intensively studied field in neuroscience with respect to the question of the neural code.
+In most mammals, light enters the eye to induce a neural activity on the retina, while keeping a certain homology between visual space and retinotopy.
+The origins of visual neuroscience may be found in Pierre Flouren's early experiments showing the dependence of visual perception upon cerebral cortex (Fite).
+In a series of seminal studies, hubel_receptive_1968 have shown that this activity could be selective to different features, such as orientation or motion .
+Remarkably, there exists a monotonous relationship between the visual contrast of the features and the firing rate of neurons.
+However, there is no consensus as to explain the multiple non-linear mechanism which transform the visual scene into a map of vernal activity, while they seem to be an essential piece to that puzzle (carandini_normalization_2012).
+
+In particular, there has been some remarkable findings when studying the dynamics of vision.
 Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.

### 2022-10-10 manuscript.tex
1.2

--- 2022-10-10 manuscript.tex
@@
 A central question in neuroscience is to better understand this relationship.
-We sometimes use the expression "decoding the neuronal code", although this implies the existence of a code, i.e.
+We sometimes use the expression ``decoding the neuronal code'', although this implies the existence of a code, i.e.
 an explicit representation within the neuronal activity.
 Nevertheless, we will use this terminology in all generality.
-In this respect, it is known that this neural activity is related to a specific variable, for example the stretching of a frog's limb (Galvani).
+In this respect, it is known that this neural activity is related to a specific variable, and since Galvani's experiments, we know that electrical activity can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
 A very important characteristic of neural activity is the fact that it consists mainly of action potentials, or spikes, which are brief (about one millisecond) electrochemical impulses that propagate along the axons of neurons.
@@
 they are prototypical, all or nothing).
-The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb (Adrian), and this variable (measured in spikes per second, or Hertz) is considered to be the main constituent of the neuronal code.
-However, computational neuroscience models have suggested that precise timing may play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this manuscript.
-
-The response of a biological neuron largely depends on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
+The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is today considered to be the main constituent of the neuronal code.
+However, computational neuroscience models have suggested that precise timing may also play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this manuscript.
+
+Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes as they reach its soma.
 This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
+This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
+A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+Importantly, validating this hypothesis would also be crucial in our understanding of neuroscientific processes.
+
 Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
@@
 %
-Let us start our review of the state of the art on the role of dynamics in vision by presenting some surprising results that have been obtained in neuroscience.
+Let us start with a review of the state of the art on the role of dynamics in vision.
+Vision is the set of processes that allow us to make sense of the luminous world, and is an area intensively studied in neuroscience with respect to the question of neural code.
+In most mammals, light enters the eye to induce neuronal activity on the retina, while maintaining a certain similitude between visual space and retinotopy.
+The origins of this  neuroscientific question can be found in the first experiments of Pierre Flouren showing the relationship between sensations, perception and activity in the cerebral cortex using lesions in animals~\citep{pearce_marie-jean-pierre_2009}.
+In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as orientation or motion.
+Remarkably, there is a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
+However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into a neural activity map, even though these seem to be an essential piece to this puzzle~\citep{carandini_normalization_2012}.
+
+In particular, there has been some remarkable findings when studying the dynamics of vision.
 Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.

### 2022-10-10 manuscript.tex
1.3

@@
 This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
+Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
+Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
 A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
@@
 These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
-
 This fast processing also explains the surprising experiments of fast serial detection which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal.
@@
 This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
-
+%
+An important consequence of the speed of processing of vision is that it implies that it is carried out using only very few spikes.
+Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
+Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
+
+%----------------------------%
 \begin{figure}
@@
 \end{figure}
-
+%----------------------------%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-An important consequence of the speed of processing is that it implies that it is carried out using only very few spikes.
-Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
-Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
-
-Numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
-Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
-
+%
 %----------------------------%
@@
 %----------------------------%
-At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{gollisch_rapid_2008}.
-Notably one can appreciate in figure 1 that the response of ganglion cells to visual gratings that are flashed onto the retina.
+%
+At the level of dynamic processing of visual information, it has been shown in the retina that an encoding of the values of luminance in the image in the timing of the spikes rather than the frequency~\citep{gollisch_rapid_2008}.
+Notably one can appreciate in these results that the response of ganglion cells to visual gratings that are flashed onto the retina.
 The authors showed that the neuronal response could be encoded in the latency of the response and not only in the frequency of discharge as is often assumed.
-In figure 4 of the same article, these results are extended to natural images and show a qualitatively similar behavior.
-The conclusion of the authors is that the discharge latency of the neurons allows to encode spatial characteristics of the image.
+These results were extended to natural images and show a qualitatively similar behavior.
+The conclusion of the authors is that the spiking latency of the neurons allows to encode spatial characteristics of the image.
 First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{chase_first-spike_2007}.
-Interestingly, such fine latency effect can be at the origin of some visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+Interestingly, such fine latency effect can be at the origin of some visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based  on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
 
@@
 Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-10-10 manuscript.tex
1.3

@@
 %
-At the level of dynamic processing of visual information, it has been shown in the retina that an encoding of the values of luminance in the image in the timing of the spikes rather than the frequency~\citep{gollisch_rapid_2008}.
-Notably one can appreciate in these results that the response of ganglion cells to visual gratings that are flashed onto the retina.
-The authors showed that the neuronal response could be encoded in the latency of the response and not only in the frequency of discharge as is often assumed.
-These results were extended to natural images and show a qualitatively similar behavior.
-The conclusion of the authors is that the spiking latency of the neurons allows to encode spatial characteristics of the image.
-First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{chase_first-spike_2007}.
-Interestingly, such fine latency effect can be at the origin of some visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based  on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+At the level of dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~{gollisch_rapid_2008}.
+In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge as it is often assumed.
+These results have been extended to natural images and show a qualitatively similar behavior.
+The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
 

### 2022-10-10 manuscript.tex
1.3

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Ultra fast neural codes for efficient vision}
+\subsection{Role of timing in the neural code for efficient vision}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
+Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
 
@@
 %
-At the level of dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~{gollisch_rapid_2008}.
+At the level of dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
 In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge as it is often assumed.
@@
 This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
-Note also that timing is not entirely sensory or internal but in~\citep{safaie_turning_2020}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
-Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Note also that timing is not entirely sensory or internal but can be used as a general coding principle.
+In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Timing, at least in animals, may thus be fundamentally embodied and situated.''
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Precise spike timings in neural assemblies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{From synfire chains\ldots{}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.

### 2022-10-10 manuscript.tex
1.3

@@
 %
+The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
+For instance, the firing rate is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of random EPSPs.
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
+When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
+This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
+However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=.7\textwidth]{figures/MainenSejnowski1995.png} % TODO replace by the simulation
+\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
+While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
+\end{figure}
+%----------------------------%
+
+%
+At the level of dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
+In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge as it is often assumed.
+These results have been extended to natural images and show a qualitatively similar behavior.
+The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
+
 %----------------------------%
@@
 %----------------------------%
-%
-At the level of dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
-In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge as it is often assumed.
-These results have been extended to natural images and show a qualitatively similar behavior.
-The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
-First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
-
 Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
-For instance, the firing rate is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of random EPSPs.
-It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figur
-~\ref{fig:mainen}).
-When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
-This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
-However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=.7\textwidth]{figures/MainenSejnowski1995.png} % TODO replace by the simulation
-\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
-While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
-See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
-\end{figure}
-%----------------------------%
-
+%
 In his book ``Corticonics''~\citep{abeles_corticonics_1991}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.

### 2022-10-10 manuscript.tex
1.3

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{From synfire chains\ldots{}}
+\subsection{A first hypothesis: synchronous firing in cell assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-In his book ``Corticonics''~\citep{abeles_corticonics_1991}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
+In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
 The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models.
-While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{paugam-moisy_computing_2012}.
+While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{paugam-moisy_computing_2012}.
 The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion.
 Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
-
-It was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_1970}.
+%
+In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_1970}.
 This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size.
@@
 When a pulse is elicited in the first group, this may generate a spike in the next group.
-Depending on the weight value, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse).
+Depending on the concentration of synaptic weight values, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse).
 Recursively applying this to a sequence of groups generates either a synfire propagation or not.
-A simple simulation is shown in Figure
-~\ref{fig:diesman}.
+A simple simulation is shown in Figure~\ref{fig:diesman}.
 A crucial aspect of this emergence is in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
@@
 %----------------------------%
-Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
+Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}, which envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003} while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
 In particular, this was implemented using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
-
+%
 %----------------------------%
@@
 \caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
-The background is only utilized in the original model, where it is implemented as random Gaussian current.
-For the presented hardware implementation it has been omitted due to network size constraints.
-As compensation for missing background stimuli, the resting potential was increased to ensure a comparable excitability of the neurons.
-The state space gives the duration of synfire chains as a function of the protocol's parameters and compares (C) CPU to (D) neuromorphic implementations.}\label{fig:pynn}
+The background noise  is implemented as random Gaussian current.
+The state space gives the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
 \end{figure}
 %----------------------------%
-Note also that synchronicity may explain some unintuitive results.
-Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
-However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{bruno_cortex_2006}.
-
+%
 Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}.
@@
 It is important to note that synchronous events tended to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see~\ref{sec:detection}).
-
+Note also that synchronicity may explain some unintuitive results.
+Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
+However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{bruno_cortex_2006}.
+
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{riehle_spike_1997,grammont_precise_1999}.
+It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
+Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials.
+Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.
 In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
@@
 
-\subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}
+\subsubsection{Extension to traveling waves}
+%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
 TODO :~\citep{bringuier_horizontal_1999} latency basin
@@
 ~\citep{linden_movement_2021} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
-
-\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{riehle_spike_1997}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{grammont_precise_1999}.
-It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
-Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials.
-Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
-Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
-These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.
 

### 2022-10-10 manuscript.tex
1

@@
 %
-Neuronal activity is directly influenced by our immediate environment and by internal states, which then generate actions.
-These actions are essential for survival, which is determinant in the light of natural selection.
-A central question in neuroscience is to better understand this relationship.
-We sometimes use the expression ``decoding the neuronal code'', although this implies the existence of a code, i.e.
-an explicit representation within the neuronal activity.
+ Neural activity is directly influenced by our immediate environment and by internal states in order to generate actions.
+These are essential for survival, which is the sole determinant in the light of natural selection.
+A central question of modern neuroscience is to better understand the essence of  Neural activity.
+We sometimes use the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
+an explicit representation within the neural activity.
 Nevertheless, we will use this terminology in all generality.
@@
 they are prototypical, all or nothing).
-The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is today considered to be the main constituent of the neuronal code.
+The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
 However, computational neuroscience models have suggested that precise timing may also play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this manuscript.
@@
 Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes as they reach its soma.
-This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
+This \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this minute temporal dimension.
 This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
@@
 Vision is the set of processes that allow us to make sense of the luminous world, and is an area intensively studied in neuroscience with respect to the question of neural code.
-In most mammals, light enters the eye to induce neuronal activity on the retina, while maintaining a certain similitude between visual space and retinotopy.
+In most mammals, light enters the eye to induce neural activity on the retina, while maintaining a certain similitude between visual space and retinotopy.
 The origins of this  neuroscientific question can be found in the first experiments of Pierre Flouren showing the relationship between sensations, perception and activity in the cerebral cortex using lesions in animals~\citep{pearce_marie-jean-pierre_2009}.
@@
 Its first experiment consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
-The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neuronal activity.
+The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neural activity.
 These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
@@
 
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{riehle_spike_1997,grammont_precise_1999}.
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas~\citep{riehle_spike_1997,grammont_precise_1999}.
 It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
@@
 Propagating waves with similar properties occur independently in secondary visual cortex, but maintain precise phase relations with the waves in primary visual cortex.
-These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatiotemporal frame for further neuronal interactions.
+These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatiotemporal frame for further neural interactions.
 
@@
 
-More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex~\citep{muller_cortical_2018}.
+More recently Advanced recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
 Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
@@
 
-Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
+Stereotyped sequences of neural activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
 These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
@@
 
-Such observations suggest that neuronal groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{miller_visual_2014}.
+Such observations suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{miller_visual_2014}.
 This work shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
@@
 The integrated FMC indicates the total memory capacity.
-We apply this theory to linear neuronal networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
+We apply this theory to linear neural networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
 A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
@@
 \item
-  Stringer et al.~analyzed spontaneous neuronal firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
-The variability of neuronal responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
+  Stringer et al.~analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+The variability of neural responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
 \item
@@
 
-In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{sun_learning_2016}.
+In a computational model, authors show that the frequently activated polychronous neural groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{sun_learning_2016}.
 This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.

### 2022-10-10 manuscript.tex
1

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
+
+\newcommand{\ms}{\si{\milli\second}}%
+
 % a simpler annotation system:
@@
 The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
-However, computational neuroscience models have suggested that precise timing may also play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this manuscript.
-
-Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes as they reach its soma.
-This \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this minute temporal dimension.
+Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
+However, computational neuroscience models have suggested that precise timing may also play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this review.
+
+Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma.
+In comparison tyo a classical analogical vector of inputs, this \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this minute temporal dimension.
 This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
 Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
-Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and also suggest the existence of repeated spatio-temporal spiking motifs in the spike train reaching neurons.
 A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
 Importantly, validating this hypothesis would also be crucial in our understanding of neuroscientific processes.
-
-Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
-In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
-%Here, we will review current literature on the detection of such motifs in generic raster plots.
-It is work in progress, where anybody interested can \emph{openly} join.
-We will first review some biological and theoretical evidence in neural information processing in the particular case of vision.
-The next section will focus on precise spike time motifs.
-We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
-In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
-Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Role of timing in the neural code for efficient vision}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Precise timing in the neural code enables efficient vision}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Remarkably, there is a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
-However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into a neural activity map, even though these seem to be an essential piece to this puzzle~\citep{carandini_normalization_2012}.
+However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitue essential pieces to this puzzle~\citep{carandini_normalization_2012}.
 
@@
 The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neural activity.
-These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
+These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of $120~\ms$ when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
 This fast processing also explains the surprising experiments of fast serial detection which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal.
 The performances decrease progressively as the frequency of presentation of the images increases.
-However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image~\citep{keysers_speed_2001}.
-
-This speed of the visual cortex, although surprising, is quite compatible with the latencies that are recorded at the neuro-physiological level.
-Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
+However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only $14~\ms$ per image~\citep{keysers_speed_2001}.
+
+This speed of the visual cortex in primates, although surprising, is  compatible with the latencies that are recorded at the neuro-physiological level.
+Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
 This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
@@
 An important consequence of the speed of processing of vision is that it implies that it is carried out using only very few spikes.
-Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
-Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
+Indeed, if we consider that a behavioral response in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
 Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{How timing encodes analogous profile}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
+\subsection{How timing may encode vectors of real values}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Sensory data is most often represented by continuous values, such as the luminance in the different pixels of a digital image.
+How may such information be encoded in neuronal activity?
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
-For instance, the firing rate is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of random EPSPs.
+For instance, the firing rate is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of firing rate.
+%random EPSPs.
+% Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
 It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
 When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
-This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
+This is consistent with the differential role of different stimulus frequencies (for instance the gamma range around $80~\si{\Hz}$) on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
 However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
@@
 %
-At the level of dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
+At the level of the dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
 In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge as it is often assumed.
@@
 The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
-First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
+This evidence found in the retina  can be extended to the visual cortex.
 
@@
 %----------------------------%
-Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
+In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
 Many models have used these properties in temporal coding to build fast image categorization networks.
-These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{thorpe_speed_1996}.
+These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
 This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
@@
 In PSCs, but also in spike activity.
-These sequences repeat after minutes and have a precise spatio temporal structure with a ms precision.
+These sequences repeat after minutes and have a precise spatio temporal structure with a $~\ms$ precision.
 They can be specific of a particular layer or colomn, are synchronized with network activity oscillation and can involve several cells.
@@
 
+Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
+In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
+%Here, we will review current literature on the detection of such motifs in generic raster plots.
+It is work in progress, where anybody interested can \emph{openly} join.
+%We will first review some biological and theoretical evidence in neural information processing in the particular case of vision. 
+%The next section will focus on precise spike time motifs. 
+We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
+In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
+Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
+
 \begin{itemize}
@@
 Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{torre_statistical_2013}.
-By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range~\citep{quaglio_methods_2018}.
+By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
 

### 2022-10-10 manuscript.tex
2.2

@@
 
-\subsubsection{Extension to traveling waves}
+\subsection{Extension to traveling waves}
 %: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
-TODO :~\citep{bringuier_horizontal_1999} latency basin
+To further investigate the role of precise timing , let us also focus on the role of differential timings across nervous.
+As we have seen, a visual feature will induce the firing of different cells at different latencies~\citep{celebrini_dynamics_1993}.
+Using intra-cellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is a graded onset of the neural response~\citep{bringuier_horizontal_1999}.
+In particular, it was shown that the network of horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within the area.
+For generic visual scenes, this processes would generate an interplay between the dynamics of the sensory signal and the spatio- temporal determinants of these interactions.
 

### 2022-10-10 manuscript.tex
2.2

--- 2022-10-10 manuscript.tex
@@
 \documentclass[brainsci, %arts,
-               review,submit,pdftex,moreauthors]{Definitions/mdpi} 
+               review,submit,pdftex,moreauthors%,draft
+               ]{Definitions/mdpi} 
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
-However, computational neuroscience models have suggested that precise timing may also play a crucial role~\citep{paugam-moisy_computing_2012}, a hypothesis that we will investigate in this review.
+However, computational neuroscience models have suggested that precise timing may also play a crucial role, and that neurons may be as well integrators than synchrony detectors~\citep{paugam-moisy_computing_2012}.
+We will investigate this very hypothesis in this review.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A first hypothesis: synchronous firing in cell assemblies}
+\subsection{One first hypothesis: synchronous firing in cell assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \centering
-\includegraphics{figures/Diesmann_et_al_1999.png}
+\includegraphics[width=.4\textwidth]{figures/Diesmann_et_al_1999.png}
 \caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
@@
 Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}, which envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003} while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
-In particular, this was implemented using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
+In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
 %
@@
 \centering
-\includegraphics{figures/fnins-07-00011-g004.jpg}
+\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
 \caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
@@
 Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
-These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tended to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see~\ref{sec:detection}).
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
+It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
@@
 Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials.
-Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Accurate spike synchronization occurs in relation to external events (stimuli, movements) and is commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
 Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
@@
 
-\subsubsection{Extension to traveling waves}
+\subsection{A further hypothesis: traveling waves}
 %: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
-TODO :~\citep{bringuier_horizontal_1999} latency basin
+To further investigate the role of precise timing, let us also focus on the role of differential timings in an assembly of neurons.
+As we have seen, a visual feature will induce the firing of different cells at different latencies~\citep{celebrini_dynamics_1993}.
+Using intra-cellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
+In particular, it was shown that the network of horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within the area.
+For generic visual scenes, this processes would generate a complex interplay between the dynamics of the sensory signal and the spatio- temporal determinants of these interactions and may underly anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
 

### 2022-10-13 manuscript.tex
2.2
Update on Overleaf.

@@
 % https://www.zotero.org/groups/4562620/polychronies
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % In recognition of your well-respected work, we are inviting you to submit one feature contribution (i.e., a long review or research paper) to an MDPI journal free of charge.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
@@
 Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma.
-In comparison tyo a classical analogical vector of inputs, this \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this minute temporal dimension.
+In comparison to a classical analogical vector of inputs, this \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this minute temporal dimension.
 This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\ldots to precise spiking motifs in the brain?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{A novel hypothesis: precise spiking motifs in cell assemblies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %\subsection{Cortical songs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{itemize}
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Precisely timing motifs in neuromorphic engineering}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Models for the detection of precise spiking motifs}\label{sec:detection}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In the following of this section, we will review theoretical models which directly tackle the problem of using precise spiking motifs.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{In computational neuroscience: Izhikevitch's polychronization model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Neurobiology spike distances}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{decoding neural activity/ spike distances}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{detecting structured temporal patterns}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Paper by Grossberger, 2018~\citep{grossberger_unsupervised_2018}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{markram_regulation_1997} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Learning sequences}\label{learning-sequences}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
@@
 \subsubsection{Grimaldi CBMI / PAMI}\label{grimaldi-cbmi-pami}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}\label{discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Spiking neural models with heterogeneous delays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{our model}\label{our-model}

### 2022-10-17 manuscript.tex
2.2

@@
 
+%: rhythms as a 
+The idea of using the synchrony of coactivation in a cell assembly is an hypothesis that was formalized by~\citet{hebb_organization_1949}. 
+This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}.
+The synchronicity will generate rhythms at different range of frequencies with spikes arriving at peak susceptibility (top of a cycle) or down.
+Such a thory has surprising ly been validated in EEG recordings for example to explain the continuous wagon wheel illusion which corresponds to the pervceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
+More generallly, it can be shown that the phase of alpha (about $10$ Hz) oscillations  is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
+%
 \subsection{A further hypothesis: traveling waves}
@@
 For generic visual scenes, this processes would generate a complex interplay between the dynamics of the sensory signal and the spatio- temporal determinants of these interactions and may underly anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
-
+%
 Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
@@
 
-More recently Advanced recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
+More recently, advanced recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
 Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
@@
 Further, by modulating local network state, they can shape responses to incoming inputs as observed in vivo.
-
-~\citep{linden_movement_2021} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
-
-~\citep{chemla_suppressive_2019}
-
-\subsubsection{TODO \ldots{} and to synfire braids and polychronous groups}\label{todo-and-to-synfire-braids-and-polychronous-groups}
-\begin{itemize}
- \item
+Such waves also occur in motor areas and~\citep{linden_movement_2021} have presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+
+Interestingly, it can be shown that these travelling waves could have a measurable role on theactivity of the visual cortex.
+This has been illustrated in a recent study studying the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
+The lrAM is the simple effect of perceiving a smooth motion when showing two dots in a temporal sequence and in close visual proximity.
+The lrAM is the building block of the use of sequences of images to induce the perception of smooth, realistic visual scenes which is at the base of movies seen in cinema theaters.
+In this study, authors used voltage-sensitive dye imaging to record the activty of the primary visual cortex of macaque monkeys to the presentation of the dots alone or in conjunction.
+A probabilistic modelling has shown that the activity of the joint presentation induced a suppressive wave in the direction opposed to the perceived direction, shaping the formation of a wave of propagation travelling at a speed compatible with the perceived motion.
+A computational model validated the hypothesis that this process could be mediated by diffusion in the horizontal layers connecting the differnt locations within this cortical area.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{A novel hypothesis: precise spiking motifs in cell assemblies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 ~\citep{bienenstock_model_1995} : from synfire chains to Synfire braids
-\end{itemize}
+
+A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
 
@@
 
-\begin{itemize}
-\item
-  recent theories of binding by synchrony : Fries 2005 trends cog neuro with spikes arriving at peak susceptibility (top of a cycle) or down, van Rullen, Laura Dugué
-\item
-  cell assembly hypothesis: neurons coordinate their activity through the formation and repetitive co-activation of groups.
-https://link-springer-com.insb.bib.cnrs.fr/article/10.1007/s10827-021-00801-9 Hebb DO.
-The organisation of behaviour: a neuropsychological theory.
-New York: Science Editions; 1949.
-\item
-  A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
-In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A novel hypothesis: precise spiking motifs in cell assemblies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %

### 2022-10-17 manuscript.tex
2.3

@@
 \subsection{A further hypothesis: traveling waves}
-%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
 To further investigate the role of precise timing, let us also focus on the role of differential timings in an assembly of neurons.
@@
 %
-Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}.
-This show precise spatio-temporal differences in this recurrently connected assembly.
-The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
-A similar dataset used in~\citep{schneidman_weak_2006} is available from M.
-Berry's lab~\citep{berry_spike_2022}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-~\citep{bienenstock_model_1995} : from synfire chains to Synfire braids
-
-A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
-In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
-
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
-This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
-
-sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. 
-CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
-
-%
-Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
+Travelling waves indicate that spatio-temporal correlations could play an important role in shaping neural activity.
+For instance, statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
+In particular, this was put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}.
+This shows precise spatio-temporal differences in this recurrently connected assembly.
+The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
+Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from M.
+Berry's lab~\citep{berry_spike_2022} and allows to perform and validate thiese hypothesis.
+However a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationaryn, which is incomptaible with the limits (in space and time) of a neural system.
+
+Moreover, there is a consensus for rate-based coding models in computational and biological neuroscience.
 Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015} and on the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
 This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996} or touch~\citep{johansson_first_2004}.
-In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs (Izhikevich, 2006) which he defined as polychronous groups.
+In particular, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
+In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which he defined as \emph{polychronous groups}.
 
 Stereotyped sequences of neural activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
-These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
+These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
 Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
@@
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
+The rest of this review will be devoted to present evidence for the use of precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
+
+%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
+
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Precisely timing motifs in computational neuroscience}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In the following section, we will review theoretical models which directly tackle the problem of using precise spiking motifs. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
+The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient and try to ``cross-compile'' classical Neural Network to a spiking architecture.
+This approach is quite successful, and SNNs approach in some case the performance of Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
+However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
+There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+To go beyond the state-of-the-art, we will review here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized.
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous`` events, that is, volleys of spikes which are stable up to certain synaptic delays.
+These models will be compared in light of neuroscientific and computational perspectives.
+We review theoretical and computational foundations of PG detection in models.
+In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
+
+Finally we will discuss future avenues for effective PG detection and learning in event streams.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Izhikevitch's polychronization model}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+First, we will review models for the detection of polychronous groups:
+
+Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
+In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
+%Here, we will review current literature on the detection of such motifs in generic raster plots.
+It is work in progress, where anybody interested can \emph{openly} join.
+%We will first review some biological and theoretical evidence in neural information processing in the particular case of vision. 
+%The next section will focus on precise spike time motifs. 
+We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
+In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
+Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
+
+% dynamic networks \& learning delays:
+
+    Huning H., Glunder H., and Palm G.
+(1998) Synaptic delay learning in pulse-coupled neurons.
+Neural Computation, 10:555--565.
+\url{https://www.deepdyve.com/lp/mit-press/synaptic-delay-learning-in-pulse-coupled-neurons-DGMiNHxp0A}
+    \href{https://www.researchgate.net/publication/37921636_Dynamics_of_Self-Organized_Delay_Adaptation}{Eurich C., Pawelzik K., Ernst U., Cowan J., and Milton J.
+(1999) Dynamics of self-organazed delay adaptation.
+Phys.
+Rev.~Lett., 82:1594--1597.}
+
+A popular model for the detection of latency patterns is the tempotron~\citep{gutig_tempotron_2006}.
+This model is in particular reviewed in~\citep{gutig_spike_2014}.
+The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
+This aims at extending the perceptron which does not incorporate a spike timing framework.
+The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
+It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
+The limits of this model is that it's output is only binary and that it's storage capacity are limited.
+
+A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
+In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
+We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{guise_bayesian_2014}.
+Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
+
+% STDP based
+Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
+This raises important questions about the fundamental limits of such memory traces and the properties required of dynamical systems to achieve these limits.
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+We introduce the Fisher Memory Curve (FMC) as a measure of the signal-to-noise ratio (SNR) embedded in the dynamical state relative to the input SNR.
+The integrated FMC indicates the total memory capacity.
+We apply this theory to linear neural networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
+A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
+The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed $\sqrt N$.
+This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
+We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.''
+
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
+
+Some other models of latency motifs detection using STDP learning rules.
+For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
+In~\citep{russo_cell_2017}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+
+% polychronization
+%  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html} Basic idea: different temporal activation of neurons can activate different populations based on conduction delays.
+This allows for a much higher dimensional representation space, and a lot more memory capacity.
+Firing is not synchronous, but it is time-locked = polychrony.
+STDP automatically associates the neurons in the network into polychronous groups.
+There can be far more polychronous groups than neurons.
+He lets the network just settle for 24 hours - and these groups emerge spontaneously from the random connectivity initial conditions.
+The dendritic tree can be used to select the polychronous inputs more precisely.
+Can act as a full feed-forward neural network and make non-linear classifications.
+Two dendritic trees on pyramidal cells - one for the feed-forward input, and one for feed-back/recurrent.
+These can interact such that the feed-back connections are predicting the feed-forward inputs.
+
+Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
+However, taking axonal propagation times into account, synchronous signals may not coincide significantly at the post-synaptic neuron contrary to asynchronous or polychronous signals.
+The term polychronous was first introduced in 2006, by E.
+Izhichevitch in~\citep{izhikevich_polychronization_2006}.
+He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
+The network is characterized by a timing-dependant learning rule for weights (STDP) and by fixed conduction delays between neurons.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
+The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
+This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike and thus to the crossing of the voltage threshold and to the discharge of a spike.
+According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and thus, constitute a polychronous group.
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
+
+This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance~\citep{wang}.
+
+%----------------------------%
+\begin{figure}
+\centering
+\note{insert fig 2 of izhikevich 2006}
+%includegraphics{figures/event_driven_computations.png}
+\caption{blah.}\label{fig:PG}
+\end{figure}
+%----------------------------%
+
+% polychronization
+
+Thus, the learning of delays allowing this polychronous group organization may be useful to detect temporal sequences of interest.
+
+Wang et al.
+(2013) presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
+
+Reproducing Polychronization: A Guide to Maximizing the Reproducibility of Spiking Network Models \url{https://www.frontiersin.org/articles/10.3389/fninf.2018.00046/full}     comes with python code
+
+    The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
+    an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Application to Image processing}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Application to Image processing using sparse spiking representations: Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
+design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
+
+there are temporal delays in the nervous system, both at the neural~\citep{perrinet_active_2014} and behavioral~\citep{khoei_flash-lag_2017} levels.
+Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
+
+may be used for motion detection and interpolation~\citep{kaplan_anisotropic_2013}~\citep{khoei_flash-lag_2017}
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ \subsection{Precise spiking motifs in neurobiology}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{decoding neural activity/ spike distances}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+J. D. Victor and K. P. Purpura, ``Nature and precision of temporal coding in visual cortex: a metric-space analysis,'' J. Neurophysiol., vol.~76, pp.~1310--1326, Aug.~1996.
+
+M. C. W. van Rossum, ``A novel spike distance,'' Neural Comput., vol.~13, no. 4, pp.~751--763, 2001.
+
+D. Aronov and J. D. Victor, ``Non-Euclidean properties of spike train metric spaces,'' Physical Rev.~E (Statist., Nonlinear, Soft Matter Phys.), vol.~69, no. 6, 2004.
+
+T. Kreuz, J. S. Haas, A. Morelli, H. D. I. Abarbanel, and A. Politi, ``Measuring spike train synchrony,'' J. Neurosci. Methods, vol.~165, no. 1, pp.~151--161, 2007. 23 H.
+
+Paper by~\citep{moser_stability_2014} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
+
+Weyl's discrepency measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation.
+
+Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~https://dx.doi.org/10.1073/pnas.1902653116
+
+A review of the methods for neuronal response latency estimation including bayesian binning~\citep{levakova_review_2015}.
+
+In generic linear non linear lnl models, the output is assumed to be poisson.
+As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
+This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
+
+Unitary event analysis is performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
+This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance).
+
+Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{torre_statistical_2013}.
+By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
+
+This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
+
+TODO: summarize ``Understanding Auditory Spectro-Temporal Receptive Fields and Their Changes with Input Statistics by Efficient Coding Principles'' - Spectro-temporal receptive fields (STRFs) have been widely used as linear approximations to the signal transform from sound spectrograms to neural responses along the auditory pathway.
+Their dependence on statistical attributes of the stimuli, such as sound intensity, is usually explained by nonlinear mechanisms and models.
+Here, we apply an efficient coding principle which has been successfully used to understand receptive fields in early stages of visual processing, in order to provide a computational understanding of the STRFs.
+According to this principle, STRFs result from an optimal tradeoff between maximizing the sensory information the brain receives, and minimizing the cost of the neural activities required to represent and transmit this information.
+Both terms depend on the statistical properties of the sensory inputs and the noise that corrupts them.
+The STRFs should therefore depend on the input power spectrum and the signal-to-noise ratio, which is assumed to increase with input intensity.
+We analytically derive the optimal STRFs when signal and noise are approximated as Gaussians.
+Under the constraint that they should be spectro-temporally local, the STRFs are predicted to adapt from being band-pass to low-pass filters as the input intensity reduces, or the input correlation becomes longer range in sound frequency or time.
+These predictions qualitatively match physiological observations.
+Our prediction as to how the STRFs should be determined by the input power spectrum could readily be tested, since this spectrum depends on the stimulus ensemble.
+The potentials and limitations of the efficient coding principle are discussed.
+~\citep{zhao_understanding_2011}
+
+\subsubsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}
+\href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
+
+\begin{itemize}
+ \item
+  https://www.janelia.org/lab/stringer-lab
+\item
+  described in~\citep{pachitariu_robustness_2018}
+\item
+  \href{https://github.com/MouseLand/rastermap}{rastermap}
+\item
+  deconvolution strategy
+\item
+  based on a linear model
+\end{itemize}
+
+\subsubsection{Stringer et al 2019, Nature~\citep{stringer_high-dimensional_2019}}\label{stringer-et-al-2019-nature-stringer2019nature}
+\begin{itemize}
+ \item
+  ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
+\item
+  Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
+\item
+  Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
+\end{itemize}
+
+\subsubsection{Stringer et al 2019, Science~\citep{stringer_spontaneous_2019}}\label{stringer-et-al-2019-science-stringer2019science}
+\begin{itemize}
+\item
+  Stringer et al.~analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+The variability of neural responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
+\item
+  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{stringer_high-precision_2021}
+\end{itemize}
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{detecting structured temporal patterns}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Paper by Grossberger, 2018~\citep{grossberger_unsupervised_2018}}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{itemize}
+ \item
+  Temporally ordered multi-neuron patterns likely encode information in the brain.
+We introduce an unsupervised method, SPOTDisClust (Spike Pattern Optimal Transport Dissimilarity Clustering), for their detection from high-dimensional neural ensembles.
+SPOTDisClust measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other (SPOTDis).
+\item
+  Detecting these temporal patterns represents a major methodological challenge.
+\item
+  Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
+\item
+  Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
+\item
+  Lee AK, Wilson MA.
+A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order.
+J Neurophysiol.
+2004;92(4):2555--73.
+pmid:15212425
+\item
+  Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
+\item
+  only one spike per neuron: fig 1A = ``For each pattern and each neuron, a random position was chosen for the activation pulse.''
+\item
+  t-SNE projection with HDBSCAN labels shows that our clustering method can retrieve all patterns from the data.
+\item
+  data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
+\end{itemize}
+
+%----------------------------%
+\begin{figure}
+\centering
+%\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
+\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}: ``Simulated example illustrating the steps in SPOTDisClust.
+A) Structure of five ``ground-truth'' patterns (\ldots).
+For each pattern and each neuron, a random position was chosen for the activation pulse.
+B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
+\end{figure}
+%----------------------------%
+
+\begin{itemize}
+ \item
+  ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
+This becomes a major problem for computing an M dissimilarity matrix (for M time epochs) for thousands of neurons.
+(2) It relies exclusively on pairwise spike-timing relationships (i.e 2nd order correlations), because it does not solve the optimal transport problem for the entire spike pattern, but only for neuron pairs separately.
+Hence, it may not be sensitive to higher-order correlations in spiking
+  patterns.
+  Here, we develop a novel dissimilarity measure for multi-neuron spiking patterns called SpikeShip, which has linear computational complexity of O(N).
+We achieve this by (1) computing the minimum transport cost of spikes for each spike train separately, and (2) discounting a global translation term in the transport flow across neurons.''
+  https://doi.org/10.1101/2020.06.03.131573;
+\end{itemize}
+
+\subsubsection{Paper by Russo et al 2017~\citep{russo_cell_2017}}\label{paper-by-russo-et-al-2017-russo2017}
+\begin{itemize}
+ \item
+  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{russo_cell_2017}
+\item
+  sliding window as in~\citep{grun_unitary_2002} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{torre_synchronous_2016}
+\item
+  based on a ``non-stationarity-corrected parametric test statistic for assessing the independence of pairs'' and ``an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies''
+\end{itemize}
+
+\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
+\begin{itemize}
+ \item
+  Stochastic sampling links perceptual uncertainty to neural response variability
+\item
+  Model accounts for independent changes in strength and variability of responses
+\item
+  Model predicts relationship between noise, signal, and spontaneous correlations
+\item
+  Stimulus statistics dependence of response statistics is explained
+\end{itemize}
+
+To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
+They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
+
+\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
+~\citep{russo_cell_2017} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
+It is based on unitary events in multiple single-neuron spiking activity~\citep{grun_unitary_2002-1} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
+Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+
+\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}
+uses a FPGA
+~\citep{dahlem_dynamics_2009}
+
+\begin{verbatim}
+February 2009International Journal of Bifurcation and Chaos 19(02):745-753
+\end{verbatim}
+
+V.
+Thanasoulis, B.
+Vogginger, J.
+Partzsch and C.
+Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
+~\citep{thanasoulis_delay-based_2021}
+
+A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
+Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
+This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
 
@@
 \end{figure}
+
+sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. 
+CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
 
@@
 7).
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Precisely timing motifs in neuromorphic engineering}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
-The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient.
-This approach is quite successful, and SNNs approach in some case the performance of Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
-However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
-There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-Our approach will be to focus on the temporal representation of information directly.
-In particular, our objective is to fully exploit the capacity of spiking neurons to detect synchronous patterns.
-
-\textbf{TODO: a paragraph on our existing work}
-Our approach would be distinct than these approaches from us and colleagues as we will directly deal with delays in the system at the presynaptic level.
-
-there are temporal delays in the nervous system, both at the neural~\citep{perrinet_active_2014} and behavioral~\citep{khoei_flash-lag_2017} levels.
-Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
-
-may be used for motion detection and interpolation~\citep{kaplan_anisotropic_2013}~\citep{khoei_flash-lag_2017}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Precise spiking motifs in  neuromorphic engineering}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-First, we will review models for the detection of polychronous groups:
-
-Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
-In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
-%Here, we will review current literature on the detection of such motifs in generic raster plots.
-It is work in progress, where anybody interested can \emph{openly} join.
-%We will first review some biological and theoretical evidence in neural information processing in the particular case of vision. 
-%The next section will focus on precise spike time motifs. 
-We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
-In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
-Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
-
-\begin{itemize}
-\item
-  A crucial advantage of Spiking Neural Networks (SNNs) architectures lies in its processing of temporal information.
-Yet, most SNNs encode the temporal signal as an analog signal and try to ``cross-compile'' classical Neural Network to a spiking architecture.
-To go beyond the state-of-the-art, we will review here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized.
-To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous`` events, that is, volleys of spikes which are stable up to certain synaptic delays.
-These models will be compared in light of neuroscientific and computational perspectives.
-We review theoretical and computational foundations of PG detection in models.
-\item
-  Application to Image processing using sparse spiking representations: Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
-design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
-\end{itemize}
-
-Finally we will discuss future avenues for effective PG detection and learning in event streams.
-
-%%
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Models for the detection of precise spiking motifs}\label{sec:detection}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the following of this section, we will review theoretical models which directly tackle the problem of using precise spiking motifs.
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{In computational neuroscience: Izhikevitch's polychronization model}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% dynamic networks \& learning delays:
-
-    Huning H., Glunder H., and Palm G.
-(1998) Synaptic delay learning in pulse-coupled neurons.
-Neural Computation, 10:555--565.
-\url{https://www.deepdyve.com/lp/mit-press/synaptic-delay-learning-in-pulse-coupled-neurons-DGMiNHxp0A}
-    \href{https://www.researchgate.net/publication/37921636_Dynamics_of_Self-Organized_Delay_Adaptation}{Eurich C., Pawelzik K., Ernst U., Cowan J., and Milton J.
-(1999) Dynamics of self-organazed delay adaptation.
-Phys.
-Rev.~Lett., 82:1594--1597.}
-
-A popular model for the detection of latency patterns is the tempotron~\citep{gutig_tempotron_2006}.
-This model is in particular reviewed in~\citep{gutig_spike_2014}.
-The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
-This aims at extending the perceptron which does not incorporate a spike timing framework.
-The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
-It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
-The limits of this model is that it's output is only binary and that it's storage capacity are limited.
-
-A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
-In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
-We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{guise_bayesian_2014}.
-Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
-
-% STDP based
-Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
-It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
-This raises important questions about the fundamental limits of such memory traces and the properties required of dynamical systems to achieve these limits.
-We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
-We introduce the Fisher Memory Curve (FMC) as a measure of the signal-to-noise ratio (SNR) embedded in the dynamical state relative to the input SNR.
-The integrated FMC indicates the total memory capacity.
-We apply this theory to linear neural networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
-A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
-The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed $\sqrt N$.
-This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
-We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.''
-
-We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
-Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
-
-Some other models of latency motifs detection using STDP learning rules.
-For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-In~\citep{russo_cell_2017}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-
-% polychronization
-%  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html} Basic idea: different temporal activation of neurons can activate different populations based on conduction delays.
-This allows for a much higher dimensional representation space, and a lot more memory capacity.
-Firing is not synchronous, but it is time-locked = polychrony.
-STDP automatically associates the neurons in the network into polychronous groups.
-There can be far more polychronous groups than neurons.
-He lets the network just settle for 24 hours - and these groups emerge spontaneously from the random connectivity initial conditions.
-The dendritic tree can be used to select the polychronous inputs more precisely.
-Can act as a full feed-forward neural network and make non-linear classifications.
-Two dendritic trees on pyramidal cells - one for the feed-forward input, and one for feed-back/recurrent.
-These can interact such that the feed-back connections are predicting the feed-forward inputs.
-
-Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
-However, taking axonal propagation times into account, synchronous signals may not coincide significantly at the post-synaptic neuron contrary to asynchronous or polychronous signals.
-The term polychronous was first introduced in 2006, by E.
-Izhichevitch in~\citep{izhikevich_polychronization_2006}.
-He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
-The network is characterized by a timing-dependant learning rule for weights (STDP) and by fixed conduction delays between neurons.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
-The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike and thus to the crossing of the voltage threshold and to the discharge of a spike.
-According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and thus, constitute a polychronous group.
-Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
-
-This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance~\citep{wang}.
-
-%----------------------------%
-\begin{figure}
-\centering
-\note{insert fig 2 of izhikevich 2006}
-%includegraphics{figures/event_driven_computations.png}
-\caption{blah.}\label{fig:PG}
-\end{figure}
-%----------------------------%
-
-% polychronization
-
-Thus, the learning of delays allowing this polychronous group organization may be useful to detect temporal sequences of interest.
-
-Wang et al.
-(2013) presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
-
-Reproducing Polychronization: A Guide to Maximizing the Reproducibility of Spiking Network Models \url{https://www.frontiersin.org/articles/10.3389/fninf.2018.00046/full}     comes with python code
-
-    The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
-    an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Neurobiology spike distances}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{decoding neural activity/ spike distances}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-J. D. Victor and K. P. Purpura, ``Nature and precision of temporal coding in visual cortex: a metric-space analysis,'' J. Neurophysiol., vol.~76, pp.~1310--1326, Aug.~1996.
-
-M. C. W. van Rossum, ``A novel spike distance,'' Neural Comput., vol.~13, no. 4, pp.~751--763, 2001.
-
-D. Aronov and J. D. Victor, ``Non-Euclidean properties of spike train metric spaces,'' Physical Rev.~E (Statist., Nonlinear, Soft Matter Phys.), vol.~69, no. 6, 2004.
-
-T. Kreuz, J. S. Haas, A. Morelli, H. D. I. Abarbanel, and A. Politi, ``Measuring spike train synchrony,'' J. Neurosci. Methods, vol.~165, no. 1, pp.~151--161, 2007. 23 H.
-
-Paper by~\citep{moser_stability_2014} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
-
-Weyl's discrepency measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation.
-
-Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~https://dx.doi.org/10.1073/pnas.1902653116
-
-A review of the methods for neuronal response latency estimation including bayesian binning~\citep{levakova_review_2015}.
-
-In generic linear non linear lnl models, the output is assumed to be poisson.
-As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
-This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
-
-Unitary event analysis is performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
-This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance).
-
-Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{torre_statistical_2013}.
-By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
-
-This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
-
-TODO: summarize ``Understanding Auditory Spectro-Temporal Receptive Fields and Their Changes with Input Statistics by Efficient Coding Principles'' - Spectro-temporal receptive fields (STRFs) have been widely used as linear approximations to the signal transform from sound spectrograms to neural responses along the auditory pathway.
-Their dependence on statistical attributes of the stimuli, such as sound intensity, is usually explained by nonlinear mechanisms and models.
-Here, we apply an efficient coding principle which has been successfully used to understand receptive fields in early stages of visual processing, in order to provide a computational understanding of the STRFs.
-According to this principle, STRFs result from an optimal tradeoff between maximizing the sensory information the brain receives, and minimizing the cost of the neural activities required to represent and transmit this information.
-Both terms depend on the statistical properties of the sensory inputs and the noise that corrupts them.
-The STRFs should therefore depend on the input power spectrum and the signal-to-noise ratio, which is assumed to increase with input intensity.
-We analytically derive the optimal STRFs when signal and noise are approximated as Gaussians.
-Under the constraint that they should be spectro-temporally local, the STRFs are predicted to adapt from being band-pass to low-pass filters as the input intensity reduces, or the input correlation becomes longer range in sound frequency or time.
-These predictions qualitatively match physiological observations.
-Our prediction as to how the STRFs should be determined by the input power spectrum could readily be tested, since this spectrum depends on the stimulus ensemble.
-The potentials and limitations of the efficient coding principle are discussed.
-~\citep{zhao_understanding_2011}
-
-\subsubsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}
-\href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
-
-\begin{itemize}
- \item
-  https://www.janelia.org/lab/stringer-lab
-\item
-  described in~\citep{pachitariu_robustness_2018}
-\item
-  \href{https://github.com/MouseLand/rastermap}{rastermap}
-\item
-  deconvolution strategy
-\item
-  based on a linear model
-\end{itemize}
-
-\subsubsection{Stringer et al 2019, Nature~\citep{stringer_high-dimensional_2019}}\label{stringer-et-al-2019-nature-stringer2019nature}
-\begin{itemize}
- \item
-  ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
-Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
-\item
-  Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
-\item
-  Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
-\end{itemize}
-
-\subsubsection{Stringer et al 2019, Science~\citep{stringer_spontaneous_2019}}\label{stringer-et-al-2019-science-stringer2019science}
-\begin{itemize}
-\item
-  Stringer et al.~analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
-The variability of neural responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
-\item
-  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{stringer_high-precision_2021}
-\end{itemize}
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{detecting structured temporal patterns}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Paper by Grossberger, 2018~\citep{grossberger_unsupervised_2018}}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{itemize}
- \item
-  Temporally ordered multi-neuron patterns likely encode information in the brain.
-We introduce an unsupervised method, SPOTDisClust (Spike Pattern Optimal Transport Dissimilarity Clustering), for their detection from high-dimensional neural ensembles.
-SPOTDisClust measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other (SPOTDis).
-\item
-  Detecting these temporal patterns represents a major methodological challenge.
-\item
-  Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
-\item
-  Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
-\item
-  Lee AK, Wilson MA.
-A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order.
-J Neurophysiol.
-2004;92(4):2555--73.
-pmid:15212425
-\item
-  Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
-\item
-  only one spike per neuron: fig 1A = ``For each pattern and each neuron, a random position was chosen for the activation pulse.''
-\item
-  t-SNE projection with HDBSCAN labels shows that our clustering method can retrieve all patterns from the data.
-\item
-  data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
-\end{itemize}
-
-%----------------------------%
-\begin{figure}
-\centering
-%\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
-\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}: ``Simulated example illustrating the steps in SPOTDisClust.
-A) Structure of five ``ground-truth'' patterns (\ldots).
-For each pattern and each neuron, a random position was chosen for the activation pulse.
-B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
-\end{figure}
-%----------------------------%
-
-\begin{itemize}
- \item
-  ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
-This becomes a major problem for computing an M dissimilarity matrix (for M time epochs) for thousands of neurons.
-(2) It relies exclusively on pairwise spike-timing relationships (i.e 2nd order correlations), because it does not solve the optimal transport problem for the entire spike pattern, but only for neuron pairs separately.
-Hence, it may not be sensitive to higher-order correlations in spiking
-  patterns.
-  Here, we develop a novel dissimilarity measure for multi-neuron spiking patterns called SpikeShip, which has linear computational complexity of O(N).
-We achieve this by (1) computing the minimum transport cost of spikes for each spike train separately, and (2) discounting a global translation term in the transport flow across neurons.''
-  https://doi.org/10.1101/2020.06.03.131573;
-\end{itemize}
-
-\subsubsection{Paper by Russo et al 2017~\citep{russo_cell_2017}}\label{paper-by-russo-et-al-2017-russo2017}
-\begin{itemize}
- \item
-  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{russo_cell_2017}
-\item
-  sliding window as in~\citep{grun_unitary_2002} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{torre_synchronous_2016}
-\item
-  based on a ``non-stationarity-corrected parametric test statistic for assessing the independence of pairs'' and ``an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies''
-\end{itemize}
-
-\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
-\begin{itemize}
- \item
-  Stochastic sampling links perceptual uncertainty to neural response variability
-\item
-  Model accounts for independent changes in strength and variability of responses
-\item
-  Model predicts relationship between noise, signal, and spontaneous correlations
-\item
-  Stimulus statistics dependence of response statistics is explained
-\end{itemize}
-
-To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
-They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
-
-\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
-~\citep{russo_cell_2017} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
-It is based on unitary events in multiple single-neuron spiking activity~\citep{grun_unitary_2002-1} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
-Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
-
-\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}
-uses a FPGA
-~\citep{dahlem_dynamics_2009}
-
-\begin{verbatim}
-February 2009International Journal of Bifurcation and Chaos 19(02):745-753
-\end{verbatim}
-
-V.
-Thanasoulis, B.
-Vogginger, J.
-Partzsch and C.
-Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
-~\citep{thanasoulis_delay-based_2021}
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{burkitt_predictive_2021}
-
-\note{\textbf{TODO amelie} section sur l'apprentissage / la plasticité des délais dans la biologie}
 
@@
 scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2021}
+
+\note{\textbf{TODO amelie} section sur l'apprentissage / la plasticité des délais dans la biologie}
 

### 2022-10-17 manuscript.tex
3.1

@@
 %: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
-
 %
@@
 These models will be compared in light of neuroscientific and computational perspectives.
-We review theoretical and computational foundations of PG detection in models.
+We review in this section theoretical and computational foundations of PG detection in models.
 In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
-
-Finally we will discuss future avenues for effective PG detection and learning in event streams.
+%Finally we will discuss future avenues for effective PG detection and learning in event streams.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-First, we will review models for the detection of polychronous groups:
-
-Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
+First, we will review models for the detection of polychronous groups.
+% Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines. 
 In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
@@
 Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
+%
+% polychronization
+%  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html} Basic idea: different temporal activation of neurons can activate different populations based on conduction delays.
+This allows for a much higher dimensional representation space, and a lot more memory capacity.
+Firing is not synchronous, but it is time-locked = polychrony.
+STDP automatically associates the neurons in the network into polychronous groups.
+There can be far more polychronous groups than neurons.
+He lets the network just settle for 24 hours - and these groups emerge spontaneously from the random connectivity initial conditions.
+The dendritic tree can be used to select the polychronous inputs more precisely.
+Can act as a full feed-forward neural network and make non-linear classifications.
+Two dendritic trees on pyramidal cells - one for the feed-forward input, and one for feed-back/recurrent.
+These can interact such that the feed-back connections are predicting the feed-forward inputs.
+%
+Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
+However, taking axonal propagation times into account, synchronous signals may not coincide significantly at the post-synaptic neuron contrary to asynchronous or polychronous signals.
+The term polychronous was first introduced by~\citet{izhikevich_polychronization_2006}.
+He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
+The network is characterized by a timing-dependant learning rule for weights (STDP) and by fixed conduction delays between neurons.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
+The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
+This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike and thus to the crossing of the voltage threshold and to the discharge of a spike.
+According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and thus, constitute a polychronous group.
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
+
+%----------------------------%
+\begin{figure}
+\centering
+\note{insert fig 2 of izhikevich 2006}
+%includegraphics{figures/event_driven_computations.png}
+\caption{blah.}\label{fig:PG}
+\end{figure}
+%----------------------------%
+
+% polychronization
+Interestingly the paper by~\citet{izhikevich_polychronization_2006} stirred a debate in the field of computational neuroscience, with a general postive acceptance, but quite few works extended this seminal paper.
 
@@
 The limits of this model is that it's output is only binary and that it's storage capacity are limited.
+The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
+an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Extensions of Izhikevitch's polychronization model}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A notable exception is an extensive work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
+Indeed, the original paper contained material within the text to reproduce the whole model.
+This more recent work details how this code was slighlty corrected.
+It comes with python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
 
@@
 In~\citep{russo_cell_2017}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-
-% polychronization
-%  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html} Basic idea: different temporal activation of neurons can activate different populations based on conduction delays.
-This allows for a much higher dimensional representation space, and a lot more memory capacity.
-Firing is not synchronous, but it is time-locked = polychrony.
-STDP automatically associates the neurons in the network into polychronous groups.
-There can be far more polychronous groups than neurons.
-He lets the network just settle for 24 hours - and these groups emerge spontaneously from the random connectivity initial conditions.
-The dendritic tree can be used to select the polychronous inputs more precisely.
-Can act as a full feed-forward neural network and make non-linear classifications.
-Two dendritic trees on pyramidal cells - one for the feed-forward input, and one for feed-back/recurrent.
-These can interact such that the feed-back connections are predicting the feed-forward inputs.
-
-Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
-However, taking axonal propagation times into account, synchronous signals may not coincide significantly at the post-synaptic neuron contrary to asynchronous or polychronous signals.
-The term polychronous was first introduced in 2006, by E.
-Izhichevitch in~\citep{izhikevich_polychronization_2006}.
-He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
-The network is characterized by a timing-dependant learning rule for weights (STDP) and by fixed conduction delays between neurons.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
-The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike and thus to the crossing of the voltage threshold and to the discharge of a spike.
-According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and thus, constitute a polychronous group.
-Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
-
-This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance~\citep{wang}.
-
-%----------------------------%
-\begin{figure}
-\centering
-\note{insert fig 2 of izhikevich 2006}
-%includegraphics{figures/event_driven_computations.png}
-\caption{blah.}\label{fig:PG}
-\end{figure}
-%----------------------------%
-
-% polychronization
-
-Thus, the learning of delays allowing this polychronous group organization may be useful to detect temporal sequences of interest.
-
-Wang et al.
-(2013) presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
-
-Reproducing Polychronization: A Guide to Maximizing the Reproducibility of Spiking Network Models \url{https://www.frontiersin.org/articles/10.3389/fninf.2018.00046/full}     comes with python code
-
-    The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
-    an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
 
@@
 
+ ~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
+This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
+
 Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.

### 2022-10-17 manuscript.tex
3.2 & 3.3

@@
 He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
-The network is characterized by a timing-dependant learning rule for weights (STDP) and by fixed conduction delays between neurons.
+The network is characterized by a timing-dependant learning rule for weights (STDP)~\citep{markram_regulation_1997} and by fixed conduction delays between neurons.
+It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
 Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
@@
 % polychronization
-Interestingly the paper by~\citet{izhikevich_polychronization_2006} stirred a debate in the field of computational neuroscience, with a general postive acceptance, but quite few works extended this seminal paper.
-
-% dynamic networks \& learning delays:
-
-    Huning H., Glunder H., and Palm G.
-(1998) Synaptic delay learning in pulse-coupled neurons.
-Neural Computation, 10:555--565.
-\url{https://www.deepdyve.com/lp/mit-press/synaptic-delay-learning-in-pulse-coupled-neurons-DGMiNHxp0A}
-    \href{https://www.researchgate.net/publication/37921636_Dynamics_of_Self-Organized_Delay_Adaptation}{Eurich C., Pawelzik K., Ernst U., Cowan J., and Milton J.
-(1999) Dynamics of self-organazed delay adaptation.
-Phys.
-Rev.~Lett., 82:1594--1597.}
-
-A popular model for the detection of latency patterns is the tempotron~\citep{gutig_tempotron_2006}.
+Interestingly the paper by~\citet{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general postive acceptance, but quite few works extended this seminal paper.
+Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the deterction of spiking motifs.
+A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}.
 This model is in particular reviewed in~\citep{gutig_spike_2014}.
@@
 The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
-It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
 The limits of this model is that it's output is only binary and that it's storage capacity are limited.
-The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
-an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
+The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
+The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
+See also the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronized_2022}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A notable exception is an extensive work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
-Indeed, the original paper contained material within the text to reproduce the whole model.
+%
+A notable extension of~\citep{izhikevich_polychronization_2006} is an extensive work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
+Indeed, while the original paper contained material within the text to reproduce the whole model, it was not complete for the repoduction of all results presented in that manuscript.
 This more recent work details how this code was slighlty corrected.
@@
 
-A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
-In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
-We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{guise_bayesian_2014}.
-Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
+Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
+Based on the fact that previous methods for studying polychronous neural groups (PNG) activation response to stimuli have been limited by the template-based methods used to identify PNG activation, authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
+They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.
+Stimulations of the trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
 
 % STDP based
-Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+%Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
 It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
@@
 This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
-We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.''
-
-We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.'' We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
 Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
 
-Some other models of latency motifs detection using STDP learning rules.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Learning synaptic delays}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A whole range of models of latency motifs detection are using STDP-based learning rules.
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
 In~\citep{russo_cell_2017}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Application to Image processing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Application to Image processing using sparse spiking representations: Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
+%
+Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific recetpive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection in recurrent networks of spiking neurons receiving oscillatory inputs\citep{kerr_delay_2013} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
+Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
+This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{burkitt_predictive_2021}
+
+Specifically, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
+Authors provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns.
+Furthermore, the experimental results of applying an STDP-based spiking neural network was applied on Random Dot Kinematogram.
+Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
+% Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
+In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
+In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
+In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity.
+This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
+
+%In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in birdsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
+In~\citep{agus_rapid_2010}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
+
+% Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
+% Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021
+
+% Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
+% Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020
+
+% Unsupervised Learning of Persistent and Sequential Activity.
+% Pereira U, Brunel N.Front Comput Neurosci. 2020
+
+% From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
+% Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019
+
+% Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
+% Pang R, Fairhall AL.Elife. 2019 May~\citep{pang_fast_2019}
+
+% Emergence of spontaneous assembly activity in developing neural networks without afferent input.
+% Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018
+
+% Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
+% Ocker GK, Doiron B.Cereb Cortex. 2019.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Real-world applications}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Most of these theoretical models are trying to reproduce neurobiological observations and some others also propose applications to machine learning mothods such as image processing.
+For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
+The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
+The second stage takes the output of the spatiotemporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatiotemporal activity.
+For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
+Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters.
+Showing the potential of such representation
+
+There are more applications to image processing using sparse spiking representations.
+Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
 design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
 
-there are temporal delays in the nervous system, both at the neural~\citep{perrinet_active_2014} and behavioral~\citep{khoei_flash-lag_2017} levels.
-Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
-
-may be used for motion detection and interpolation~\citep{kaplan_anisotropic_2013}~\citep{khoei_flash-lag_2017}
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \subsection{Precise spiking motifs in neurobiology}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+there are temporal delays in the nervous system, both at the neural~\citep{perrinet_active_2014}
+
+and behavioral~\citep{khoei_flash-lag_2017} levels. Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
+
+may be used for motion detection and Motion-based prediction explains the role of tracking in motion extrapolation interpolation~\citep{kaplan_anisotropic_2013}~\citep{khoei_flash-lag_2017}
+
+We will review some more applications of using precise spiking motifs in neuromorphic engineering, but will first review current evidence in neurobiology, that is, from actual recordings in biological tissues.
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ \section{Precise spiking motifs in neurobiology}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+ In Bellec~\citep{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
+The model uses samples and measure method sto include latent / hidden neurons.
+In particular, it    comes with  \href{code}{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}.
+It uses the publicly available \href{V1-dataset}{http://crcns.org/data-sets/vc/pvc-11}, which allows to supervise the model with the input being the movie and the output the spikes recorded.
 
@@
 Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
-
-\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}
-uses a FPGA
-~\citep{dahlem_dynamics_2009}
-
-\begin{verbatim}
-February 2009International Journal of Bifurcation and Chaos 19(02):745-753
-\end{verbatim}
-
-V.
-Thanasoulis, B.
-Vogginger, J.
-Partzsch and C.
-Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
-~\citep{thanasoulis_delay-based_2021}
-
-A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
-In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
-This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
 
@@
 
+\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}
+uses a FPGA
+~\citep{dahlem_dynamics_2009}
+
+\begin{verbatim}
+February 2009International Journal of Bifurcation and Chaos 19(02):745-753
+\end{verbatim}
+
+V.
+Thanasoulis, B.
+Vogginger, J.
+Partzsch and C.
+Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
+~\citep{thanasoulis_delay-based_2021}
+
+A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
+Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
+This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
+
  ~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
@@
 The rest of this review paper is organized as follows.
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{markram_regulation_1997} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
-This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach.
-This will decouple the active synapses (similarly to a logistic regression) from the values of possible synaptic delays.
-
-\begin{itemize}
- \item
- ~\citep{perrinet_coherence_2002} : coherence detection
-\item
- ~\citep{perrinet_networks_2001} : STDP
-\item
-  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{gilson_stdp_2010} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
-PLoS Comput Biol 9(2): e1002897.~\citep{kerr_delay_2013} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
-\end{itemize}
-
-Our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
-Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
-This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{burkitt_predictive_2021}
-
-~\citep{nadafian_bio-plausible_2020}
-
-Bio-plausible Unsupervised Delay Learning for Extracting Temporal Features in Spiking Neural Networks
-Alireza Nadafian, Mohammad Ganjtabesh
-
-\begin{verbatim}
-The plasticity of the conduction delay between neurons plays a fundamental role in learning.
-However, the exact underlying mechanisms in the brain for this modulation is still an open problem.
-Understanding the precise adjustment of synaptic delays could help us in developing effective brain-inspired computational models in providing aligned insights with the experimental evidence.
-In this paper, we propose an unsupervised biologically plausible learning rule for adjusting the synaptic delays in spiking neural networks.
-Then, we provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns.
-Furthermore, the experimental results of applying an STDP-based spiking neural network equipped with our proposed delay learning rule on Random Dot Kinematogram indicate the efficacy of the proposed delay learning rule in extracting temporal features.
-\end{verbatim}
-
-Integrating synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes, that is referred to as synaptic delay-weight plasticity~\citep{zhang_supervised_2020}.
-Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
-
-In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
-In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
-
-In a computational model, authors show that the frequently activated polychronous neural groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{sun_learning_2016}.
-This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Learning sequences}\label{learning-sequences}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
-Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
-
-\begin{itemize}
-\item
-  in~\citep{agus_rapid_2010}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
-\item
-  In Bellec~\citep{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
-
-  \begin{itemize}
-     \item
-    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics)
-  \item
-    sample and measure method to include latent / hidden neurons
-  \item
-    comes with code \url{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}
-  \item
-    V1-dataset : The dataset we used was collected by Smith and Kohn 49 and is publicly available at:
-     \url{http://crcns.org/data-sets/vc/pvc-11} - it is in a sense supervised with the input being the movie and the output the spikes recorded.
-  \end{itemize}
-\end{itemize}
-
-A recent work~\citep{ghosh_spatiotemporal_2019} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
-The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
-The second stage takes the output of the spatiotemporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatiotemporal activity.
-For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
-Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters.
-Showing the potential of such representation
-
-Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
-Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021
-
-Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
-Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020
-
-Unsupervised Learning of Persistent and Sequential Activity.
-Pereira U, Brunel N.Front Comput Neurosci. 2020
-
-From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
-Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019
-
-Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
-Pang R, Fairhall AL.Elife. 2019 May~\citep{pang_fast_2019}
-
-Emergence of spontaneous assembly activity in developing neural networks without afferent input.
-Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018
-
-Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
-Ocker GK, Doiron B.Cereb Cortex. 2019.
 

### 2022-10-17 manuscript.tex
3.4

@@
 %
-A notable extension of~\citep{izhikevich_polychronization_2006} is an extensive work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
+A notable extension of~\citep{izhikevich_polychronization_2006} is a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
 Indeed, while the original paper contained material within the text to reproduce the whole model, it was not complete for the repoduction of all results presented in that manuscript.
@@
 
-There are more applications to image processing using sparse spiking representations.
+%There are more applications to image processing using sparse spiking representations.
 Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
 design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
-
-there are temporal delays in the nervous system, both at the neural~\citep{perrinet_active_2014}
-
-and behavioral~\citep{khoei_flash-lag_2017} levels. Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
-
-may be used for motion detection and Motion-based prediction explains the role of tracking in motion extrapolation interpolation~\citep{kaplan_anisotropic_2013}~\citep{khoei_flash-lag_2017}
-
+A predictive model defined in the framework of the free-energy principle was able to account for temporal delays in the nervous system, both at the sensory and motor levels~\citep{perrinet_active_2014}.
+%
+%and behavioral~\citep{khoei_flash-lag_2017} levels. Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
+Such a model was implemented at the network level by showing that motion-based prediction explains the role of tracking in motion extrapolation interpolation~\citep{khoei_motion-based_2013}.
+Note that this motion-based prediction is implemented in the weights of the neural network by defining connections between neurons with specific delays.
+Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
+This was aso implemented in a spiking neural network and reproduced the observation that neural activty was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
+This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the flash-lag illusion~\citep{khoei_flash-lag_2017}.
 We will review some more applications of using precise spiking motifs in neuromorphic engineering, but will first review current evidence in neurobiology, that is, from actual recordings in biological tissues.
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
  In Bellec~\citep{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.

### 2022-10-17 manuscript.tex
4.1

@@
 
-%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
+% Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~https://dx.doi.org/10.1073/pnas.1902653116
 %
@@
 %
- In Bellec~\citep{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+% In Bellec~\citep{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
 The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{decoding neural activity/ spike distances}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-J. D. Victor and K. P. Purpura, ``Nature and precision of temporal coding in visual cortex: a metric-space analysis,'' J. Neurophysiol., vol.~76, pp.~1310--1326, Aug.~1996.
-
-M. C. W. van Rossum, ``A novel spike distance,'' Neural Comput., vol.~13, no. 4, pp.~751--763, 2001.
-
-D. Aronov and J. D. Victor, ``Non-Euclidean properties of spike train metric spaces,'' Physical Rev.~E (Statist., Nonlinear, Soft Matter Phys.), vol.~69, no. 6, 2004.
-
-T. Kreuz, J. S. Haas, A. Morelli, H. D. I. Abarbanel, and A. Politi, ``Measuring spike train synchrony,'' J. Neurosci. Methods, vol.~165, no. 1, pp.~151--161, 2007. 23 H.
-
-Paper by~\citep{moser_stability_2014} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
-
-Weyl's discrepency measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation.
-
-Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~https://dx.doi.org/10.1073/pnas.1902653116
-
-A review of the methods for neuronal response latency estimation including bayesian binning~\citep{levakova_review_2015}.
-
-In generic linear non linear lnl models, the output is assumed to be poisson.
+\subsection{Decoding neural activity using spike distances}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In most generic computational models the neural activity is assumed to be encoded in the firing rate.
+For instance, linear non linear L-NL models, the output is assumed to be distributed according to a Poisson law.
 As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
@@
 
-Unitary event analysis is performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
+There are different solution to provide with a distance between two given spike trains.
+A known measure is the Victor-Pupura distance which showed that their findings are inconsistent with Poisson models of spike trains~\citep{victor_nature_1996}.
+In a later study,~\citet{van_rossum_novel_2001} have defined a distance that has a time constant as a parameter.
+Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
+Such distances were extended to non-Euclidean metrics~\citep{} and use morphological manipulmations to compute spike train dissimilarity~\citet{kruz}.
+Mathelmatically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepency measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
+These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures.
+
+Let's now focus on distances defined based on a generative model formed by using precise spiking motifs.
+A review of the methods for neuronal response latency estimation exists that includes bayesian binning~\citep{levakova_review_2015}.
+More braodly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
 This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance).
-
-Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{torre_statistical_2013}.
+A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}.
 By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
-
 This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
@@
 
-TODO: summarize ``Understanding Auditory Spectro-Temporal Receptive Fields and Their Changes with Input Statistics by Efficient Coding Principles'' - Spectro-temporal receptive fields (STRFs) have been widely used as linear approximations to the signal transform from sound spectrograms to neural responses along the auditory pathway.
+% TODO: summarize ``Understanding Auditory Spectro-Temporal Receptive Fields and Their Changes with Input Statistics by Efficient Coding Principles'' - Spectro-temporal receptive fields (STRFs) have been widely used as linear approximations to the signal transform from sound spectrograms to neural responses along the auditory pathway.
 Their dependence on statistical attributes of the stimuli, such as sound intensity, is usually explained by nonlinear mechanisms and models.
@@
 
-\subsubsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}
+%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
+\subsection{Scaling up to very large scale data}
+
 \href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
@@
 \item
-  in cotical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
+  in cortical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
@@
 7).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-10-17 manuscript.tex
4.2

@@
 \subsection{Scaling up to very large scale data}
-
-\href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
-
-\begin{itemize}
- \item
-  https://www.janelia.org/lab/stringer-lab
-\item
-  described in~\citep{pachitariu_robustness_2018}
-\item
-  \href{https://github.com/MouseLand/rastermap}{rastermap}
-\item
-  deconvolution strategy
-\item
-  based on a linear model
-\end{itemize}
-
-\subsubsection{Stringer et al 2019, Nature~\citep{stringer_high-dimensional_2019}}\label{stringer-et-al-2019-nature-stringer2019nature}
-\begin{itemize}
- \item
-  ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
+Bridging across different spatial and temporal scales, combination of in vivo two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
+The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurophysics to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
+
+One algorithm capable of achieveing such daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
+Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
+The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
+In~\citet{stringer_high-dimensional_2019}, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+Then, in~\citep{stringer_high-precision_2021}, authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
+% \href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
+
+% \begin{itemize}
+%  \item
+%   https://www.janelia.org/lab/stringer-lab
+% \item
+%   described in
+% \item
+%   \href{https://github.com/MouseLand/rastermap}{rastermap}
+% \item
+%   deconvolution strategy
+% \item
+%   based on a linear model
+% \end{itemize}
+% \subsubsection{Stringer et al 2019, Nature~\citep{stringer_high-dimensional_2019}}\label{stringer-et-al-2019-nature-stringer2019nature}
+% \begin{itemize}
+%  \item
+%   ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
 Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
-\item
-  Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
-\item
-  Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
-\end{itemize}
-
-\subsubsection{Stringer et al 2019, Science~\citep{stringer_spontaneous_2019}}\label{stringer-et-al-2019-science-stringer2019science}
-\begin{itemize}
-\item
-  Stringer et al.~analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+% \item
+%   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
+% \item
+%   Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
+% \end{itemize}
+
+% \subsubsection{Stringer et al 2019, Science~\citep{stringer_spontaneous_2019}}\label{stringer-et-al-2019-science-stringer2019science}
+% \begin{itemize}
+% \item
+%   Stringer et al.~analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
 The variability of neural responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
-\item
-  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{stringer_high-precision_2021}
-\end{itemize}
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{detecting structured temporal patterns}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Paper by Grossberger, 2018~\citep{grossberger_unsupervised_2018}}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{itemize}
- \item
-  Temporally ordered multi-neuron patterns likely encode information in the brain.
-We introduce an unsupervised method, SPOTDisClust (Spike Pattern Optimal Transport Dissimilarity Clustering), for their detection from high-dimensional neural ensembles.
-SPOTDisClust measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other (SPOTDis).
-\item
-  Detecting these temporal patterns represents a major methodological challenge.
-\item
-  Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
-\item
-  Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
-\item
-  Lee AK, Wilson MA.
+% \item
+%   see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{stringer_high-precision_2021}
+% \end{itemize}
+
+Another important algorithm is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}.
+They introduced an unsupervised method based for their detection from high-dimensional neural ensembles.
+The algorithm measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other.
+Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
+% \item
+%   Detecting these temporal patterns represents a major methodological challenge.
+% \item
+%   Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
+% \item
+%   Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
+% \item
+%   Lee AK, Wilson MA.
 A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order.
@@
 pmid:15212425
-\item
-  Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
-\item
-  only one spike per neuron: fig 1A = ``For each pattern and each neuron, a random position was chosen for the activation pulse.''
-\item
-  t-SNE projection with HDBSCAN labels shows that our clustering method can retrieve all patterns from the data.
-\item
-  data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
-\end{itemize}
+% \item
+%   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
+% \item
+In that method, the learning is unsupervised. It uses the prior that there is only one spike per pattern. 
+% \item
+Using a t-SNE projection with knwon labels validated that this clustering method can retrieve all patterns from the data.
+% \item
+%   data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
+% \end{itemize}
+The limits of this method is that it is block-based and stricly specialized for the task at hand.
+To overcome these diculties a novel method was recently developped~\citep{sotomayor-gomez_spikeship_2021}.
 
@@
 
-\begin{itemize}
- \item
-  ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
+% \begin{itemize}
+%  \item
+%   ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
 This becomes a major problem for computing an M dissimilarity matrix (for M time epochs) for thousands of neurons.
@@
 Hence, it may not be sensitive to higher-order correlations in spiking
-  patterns.
-  Here, we develop a novel dissimilarity measure for multi-neuron spiking patterns called SpikeShip, which has linear computational complexity of O(N).
+%   patterns.
+%   Here, we develop a novel dissimilarity measure for multi-neuron spiking patterns called SpikeShip, which has linear computational complexity of O(N).
 We achieve this by (1) computing the minimum transport cost of spikes for each spike train separately, and (2) discounting a global translation term in the transport flow across neurons.''
-  https://doi.org/10.1101/2020.06.03.131573;
-\end{itemize}
-
+%   https://doi.org/10.1101/2020.06.03.131573;
+% \end{itemize}
+
+\subsection{Cell assemblies at multiple time scales with arbitrary lag constellations}
 \subsubsection{Paper by Russo et al 2017~\citep{russo_cell_2017}}\label{paper-by-russo-et-al-2017-russo2017}
@@
 
-\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
 ~\citep{russo_cell_2017} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
@@
 
-sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. 
+% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. 
 CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
@@
 In~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
-They develop a HMM model for quantifying the transitions.
-``In this study, van Kempen et al.~show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
-Top-down attention drives interareal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.''
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Cortical songs}
+They develop a Hidden Markov Model for quantifying the transitions.
+In particular, they show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
+Top-down attention drives interareal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Cortical songs}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\subsection{learning pattern detection on natural images / event-based cameras}
+
 Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}.
@@
 
-\subsection{learning pattern detection on natural images / event-based cameras}\label{learning-pattern-detection-on-natural-images-event-based-cameras}
 \subsubsection{sparse coding on spatio-temporal data}\label{sparse-coding-on-spatio-temporal-data}

### 2022-10-17 manuscript.tex
4.3

@@
 
-\subsection{Cell assemblies at multiple time scales with arbitrary lag constellations}
-\subsubsection{Paper by Russo et al 2017~\citep{russo_cell_2017}}\label{paper-by-russo-et-al-2017-russo2017}
-\begin{itemize}
- \item
-  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{russo_cell_2017}
-\item
-  sliding window as in~\citep{grun_unitary_2002} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{torre_synchronous_2016}
-\item
-  based on a ``non-stationarity-corrected parametric test statistic for assessing the independence of pairs'' and ``an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies''
-\end{itemize}
-
-\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
-\begin{itemize}
- \item
-  Stochastic sampling links perceptual uncertainty to neural response variability
-\item
-  Model accounts for independent changes in strength and variability of responses
-\item
-  Model predicts relationship between noise, signal, and spontaneous correlations
-\item
-  Stimulus statistics dependence of response statistics is explained
-\end{itemize}
+\subsection{Extending the decoding at multiple time scales}
+
+In the paper by~\citet{russo_cell_2017}, authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.
+It uses  sliding window as in~\citep{grun_unitary_2002} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
+They extend the measure to multiple lags~\citep{torre_synchronous_2016}.
+The core measure is based on a non-stationarity-corrected parametric test statistic for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+
+% \subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
+% \begin{itemize}
+%  \item
+%   Stochastic sampling links perceptual uncertainty to neural response variability
+% \item
+%   Model accounts for independent changes in strength and variability of responses
+% \item
+%   Model predicts relationship between noise, signal, and spontaneous correlations
+% \item
+%   Stimulus statistics dependence of response statistics is explained
+% \end{itemize}
 
@@
 They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
-
-~\citep{russo_cell_2017} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
-It is based on unitary events in multiple single-neuron spiking activity~\citep{grun_unitary_2002-1} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
-Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
 
@@
 
-% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. 
-CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
-
-TODO: read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
+% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
 
@@
 
+TODO: read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
+
 Such observations suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{miller_visual_2014}.

### 2022-10-17 manuscript.tex
4.3
abstract

@@
 \Author{
-Antoine Grimaldi $^{1,\dagger,\ddagger}$\orcidA{}, 
-Amélie Gruel $^{2,\dagger,\ddagger}$, 
-Jean Martinet $^{2,}$ 
-and Laurent Perrinet $^{1,}$\orcidD{}*}
+Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
+\orcidA{}, 
+Amélie Gruel $^{2}$, % $^{2,\dagger,\ddagger}$, 
+Jean Martinet $^{2}$,  
+and Laurent Perrinet $^{1,}$\orcidD{},*}
 
@@
 Why do neurons communicate through spikes?
-By definition, a spike, or action potential, is a all-or-none event ---it can occur or not without further details--- at asynchronous timings, i.e.~it can occur at any continuous time - differentially to the discretized timing classically used in digital processing.
+By definition, a spike, or action potential, is a all-or-none event ---it can occur or not without further details--- at timings asynchronous within a cell assembly, i.e.~it can occur at any continuous time.
+This stands in stark contrast to the discretized timing classically used in digital processing at the base of modern-day neural networks.
 In the living world, neural systems almost systematically use this so-called event-based representation.
 A better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data.
-It is also an emerging challenge in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal challenge in the age of the digital economy and of global warming.
+It is also an emerging topic in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal mission in the age of the digital economy and of global warming.
 }
@@
 It uses the publicly available \href{V1-dataset}{http://crcns.org/data-sets/vc/pvc-11}, which allows to supervise the model with the input being the movie and the output the spikes recorded.
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding neural activity using spike distances}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In most generic computational models the neural activity is assumed to be encoded in the firing rate.
@@
 This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
-
+As such, all these models are dependent on  a core definition of spike measures and we will review here how precise spiking motifs are taken into account in various neurobiological results.
+%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decoding neural activity using spike distances}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 There are different solution to provide with a distance between two given spike trains.

### 2022-10-18 manuscript.tex
3.1

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A novel hypothesis: precise spiking motifs in cell assemblies}
+\subsection{A rediscovered  hypothesis: precise spiking motifs in cell assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 These can interact such that the feed-back connections are predicting the feed-forward inputs.
-%
+
+A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
+Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
+This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
+
+% REDONDANCE
 Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
@@
 %
-A notable extension of~\citep{izhikevich_polychronization_2006} is a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
+An extension of~\citep{izhikevich_polychronization_2006} is a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
 Indeed, while the original paper contained material within the text to reproduce the whole model, it was not complete for the repoduction of all results presented in that manuscript.
@@
 %
-Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific recetpive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection in recurrent networks of spiking neurons receiving oscillatory inputs\citep{kerr_delay_2013} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific recetpive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection in recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
 Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
-This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{burkitt_predictive_2021}
+This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
 
@@
 \subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}
-uses a FPGA
-~\citep{dahlem_dynamics_2009}
-
-\begin{verbatim}
-February 2009International Journal of Bifurcation and Chaos 19(02):745-753
-\end{verbatim}
+
+uses a FPGA~\citep{dahlem_dynamics_2009}
 
@@
 ~\citep{thanasoulis_delay-based_2021}
-
-A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
-In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
-This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
 

### 2022-10-18 manuscript.tex
plan de 5

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}
+We have seen that precise spiking moifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobuological data. 
+A third class of models coming from the field of neuromorphic engineering, a priori different from that used in neuroscience, are of interest in this review as they often take advantage of minute temporal information.
+
+\subsubsection{The emergence of neuromorphic engineering}
+% intro 
+the field of neuromorphic engineering...
 
@@
 
+% un exemple de polychronization
  ~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
@@
 
+% d'autres applications
 Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
@@
 
-\subsection{learning pattern detection on natural images / event-based cameras}
-
+\subsection{Temporal pattern detection in event-based cameras}
+% le cas spécifique des event-based cameras
 Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}.
@@
 
-\subsubsection{sparse coding on spatio-temporal data}\label{sparse-coding-on-spatio-temporal-data}
-\subsubsection{HOTS}\label{hots}
-\subsubsection{Grimaldi CBMI / PAMI}\label{grimaldi-cbmi-pami}
+% solutions usuellement utilisées - surrogate gradient
+
+% autres méthodes
+
+\subsubsection{Methods using precise timings}
+
+% HOTS
+
+% Grimaldi CBMI + PAMI
+
+% Grimaldi ICIP
+
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-10-18 manuscript.tex
4.4

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{itemize}
- \item
-  Ikegaya and colleagues~\citep{ikegaya_synfire_2004} worked on spontaneous activity in vitro and in vivo.
-They demonstrated that in cortical activity, we can find a repetition of several motifs.
-In PSCs, but also in spike activity.
-These sequences repeat after minutes and have a precise spatio temporal structure with a $~\ms$ precision.
-They can be specific of a particular layer or colomn, are synchronized with network activity oscillation and can involve several cells.
-They also demonstated that these sequences can form supersequences : the cortical songs.
+
+Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons in vivo and in vitro.
+These patterns repeat after minutes, maintaining millisecond accuracy.
+Indeed,~\citet{ikegaya_synfire_2004} demonstrated that in cortical activity, one can find a repetition of several motifs in spike activity (duration around 1s +/- .5 s, some events in motifs are of similar size but sometimes absent).
+These sequences can be specific of a particular layer or colomn, can be synchronized with network activity oscillation and can involve several cells.
+They also demonstated that these sequences can form supersequences, so-called cortical songs.
 It consist of the assembly of several sequences which repeat in a specific order with a compressed timing.
-\item
-  ``We find precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons in vivo and in vitro. These patterns repeat after minutes, maintaining millisecond accuracy.''
-\item
-  Precisely repeating motifs of spontaneous synaptic activity in slices: duration around 1s +/- .5 s.
-Some events in motifs are of similar size but sometimes absent - better described by Bernouilli than SE (and covariance)
-\item
-  \emph{in vivo} spontaneous activity also reveals repeating sequences. About 3000 sequences, each involving 3-10 cells out of about 900, and last up to 3 seconds
-\item
-  topography: ``Sequences had specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster (Fig.
-4, A and B, and fig.~S3B).
-(\ldots) Therefore, repeating temporal patterns of activation (\ldots) were associated with a structured spatial organization of the neurons that formed them.''
-\item
-  ``Cortical songs: modular assemblies of repeated sequences'': hierarchical detection.
-\item
-  in cortical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
+\emph{In vivo} spontaneous activity also reveals repeating sequences.
+About 3000 sequences, each involving 3-10 cells out of about 900, and last up to 3 seconds.
+Sequences had specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster and were associated with a structured spatial organization of the neurons that formed them.
+%In cortical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
-\end{itemize}
 
@@
 
-TODO: read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
+TODO: re-read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
 
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{our model}\label{our-model}
 Here, we develop a model for the efficient detection of such PGs based on the inversion of a probabilistic model defining the generation of the raster plot as a combination of such groups.

### 2022-10-18 manuscript.tex
4.4

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% what we say in section 2.3: Stereotyped sequences of neural activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
+These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
+Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019} below).
+We have seen previously that neurobiologists havbe described Stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019} below).
 Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons in vivo and in vitro.

### 2022-10-18 manuscript.tex
Update on Overleaf.
Update manuscript.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Precisely timing motifs in computational neuroscience}
+\section{Precisely-timed spiking motifs in computational neuroscience}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Extending the decoding at multiple time scales}
+https://www.overleaf.com/project/633bf90497b5b325a3ff692e
+\subsection{Extending the decoding athttps://www.overleaf.com/project/633bf90497b5b325a3ff692e multiple time scales}
 
@@
 
+% Currently, few interdisciplinary research projects take full advantage of event-based representations in computational and biological neuroscience.
+Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains (Ikegaya et al., 2004), packets (Luczak et al., 2007) or hippocampal sequences (Pastalkova et al., 2008; Villette et al., 2015)  and on the role of such precise spike timing in downstream information transfer and coding (Branco et al., 2010; Buzsáki & Tingley, 2018; Luczak et al., 2015).
+In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
+In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs (Izhikevich, 2006) which he defined as polychronous groups.
+
+%Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space (Buzsáki & Tingley, 2018).
+These sequences can be internally generated (Pastalkova et al., 2008; Villette et al., 2015) and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
+Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure 2 of (Haimerl et al., 2019)).
+As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
 \begin{figure}
 \centering
-%\includegraphics{https://www.pnas.org/cms/10.1073/pnas.1718518116/asset/ab9a48f6-b28f-40c1-978e-6f6dbec39b98/assets/graphic/pnas.1718518116fig01.jpeg}
-\caption{Mixed distance and duration representation in CA1.
+\includegraphics{figures/haimerl2019.jpg}
+\caption{
+%  Figure 2: An example of a polychronous group seen in cortical slices.
+In this study by (Haimerl et al., 2019), an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+Mixed distance and duration representation in CA1.
 (A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.

### 2022-10-19 manuscript.tex
Update on Overleaf.

@@
 Why do neurons communicate through spikes?
-By definition, a spike, or action potential, is a all-or-none event ---it can occur or not without further details--- at timings asynchronous within a cell assembly, i.e.~it can occur at any continuous time.
+By definition, a spike, or action potential, is an all-or-none event ---it can occur or not without further details--- at asynchronous timings within a cell assembly, i.e.~it can occur at any continuous time.
 This stands in stark contrast to the discretized timing classically used in digital processing at the base of modern-day neural networks.
@@
 A central question of modern neuroscience is to better understand the essence of  Neural activity.
-We sometimes use the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
+One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
 an explicit representation within the neural activity.
@@
 This speed of the visual cortex in primates, although surprising, is  compatible with the latencies that are recorded at the neuro-physiological level.
-Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
+Indeed, when an image is presented to the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
 This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.

### 2022-10-26 manuscript.tex
Update on Overleaf.

@@
 Why do neurons communicate through spikes?
-By definition, a spike, or action potential, is an all-or-none event ---it can occur or not without further details--- at asynchronous timings within a cell assembly, i.e.~it can occur at any continuous time.
+By definition, a spike, or action potential, is an all-or-none event which occurs at asynchronous times.
+In other words, spikes on one side occur or not without further details and second can occur at any continuous time.
 This stands in stark contrast to the discretized timing classically used in digital processing at the base of modern-day neural networks.
@@
 These are essential for survival, which is the sole determinant in the light of natural selection.
-A central question of modern neuroscience is to better understand the essence of  Neural activity.
+A central question of modern neuroscience is to better understand the essence of neural activity.
 One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
@@
 In this respect, it is known that this neural activity is related to a specific variable, and since Galvani's experiments, we know that electrical activity can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-A very important characteristic of neural activity is the fact that it consists mainly of action potentials, or spikes, which are brief (about one millisecond) electrochemical impulses that propagate along the axons of neurons.
+Among various phenomenon, a well studied way of communication between neurons are action potentials, or spikes.
+In the scope of this article we focus mainly on spikes, which are brief (about one millisecond) electrochemical impulses that propagate along the axons of neurons.
 These have the particularity of being essentially binary in their amplitude (i.e.
@@
 Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma.
-In comparison to a classical analogical vector of inputs, this \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this minute temporal dimension.
+In comparison to a classical analogical vector of inputs, this \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this fine-scaled temporal dimension.
 This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
@@
 Vision is the set of processes that allow us to make sense of the luminous world, and is an area intensively studied in neuroscience with respect to the question of neural code.
-In most mammals, light enters the eye to induce neural activity on the retina, while maintaining a certain similitude between visual space and retinotopy.
+In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between external visual space and its representation on the retina.
 The origins of this  neuroscientific question can be found in the first experiments of Pierre Flouren showing the relationship between sensations, perception and activity in the cerebral cortex using lesions in animals~\citep{pearce_marie-jean-pierre_2009}.
@@
 Remarkably, there is a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
-However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitue essential pieces to this puzzle~\citep{carandini_normalization_2012}.
-
-In particular, there has been some remarkable findings when studying the dynamics of vision.
+However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitute essential pieces to this puzzle~\citep{carandini_normalization_2012}.
+
+In particular, there have been some remarkable findings when studying the dynamics of vision.
 Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
@@
 These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of $120~\ms$ when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
-This fast processing also explains the surprising experiments of fast serial detection which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal.
+This fast processing also explains the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
 The performances decrease progressively as the frequency of presentation of the images increases.
-However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only $14~\ms$ per image~\citep{keysers_speed_2001}.
+However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only $14~\ms$ per image.
 
 This speed of the visual cortex in primates, although surprising, is  compatible with the latencies that are recorded at the neuro-physiological level.
-Indeed, when an image is presented to the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
+Indeed, when an image is presented to the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex, and it takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
 This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
@@
 Indeed, if we consider that a behavioral response in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
-Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
+Altogether, these results show that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
 
@@
 % Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
-It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned, but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
 When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
@@
 At the level of the dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
-In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge as it is often assumed.
+In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
 These results have been extended to natural images and show a qualitatively similar behavior.
@@
 These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}.
-First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
-This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
+First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
+This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
 This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events~\citep{rasetto_challenges_2022}.
-This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
-Note also that timing is not entirely sensory or internal but can be used as a general coding principle.
+This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
+Note also that timing is not entirely sensory or internal, but can be used as a general coding principle.
 In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
@@
 %----------------------------%
-Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}, which envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003} while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
+Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}, which envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
 In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
@@
 Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}.
-Interestingly, authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Interestingly, the authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
 Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
-Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
-However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{bruno_cortex_2006}.
+Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intra-cortical activity.
+However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active~\citep{bruno_cortex_2006}.
 
@@
 %: rhythms as a 
-The idea of using the synchrony of coactivation in a cell assembly is an hypothesis that was formalized by~\citet{hebb_organization_1949}. 
+The idea of using the synchrony of co-activation in a cell assembly is a hypothesis that was formalized by~\citet{hebb_organization_1949}. 
 This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}.
-The synchronicity will generate rhythms at different range of frequencies with spikes arriving at peak susceptibility (top of a cycle) or down.
-Such a thory has surprising ly been validated in EEG recordings for example to explain the continuous wagon wheel illusion which corresponds to the pervceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
-More generallly, it can be shown that the phase of alpha (about $10$ Hz) oscillations  is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
+The synchronicity will generate rhythms at different range of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down.
+Such a theory has surprisingly been validated in EEG recordings, for example, to explain the continuous wagon-wheel illusion which corresponds to the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
+More generally, it can be shown that the phase of alpha (about $10$ Hz) oscillations  is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
 %
@@
 As we have seen, a visual feature will induce the firing of different cells at different latencies~\citep{celebrini_dynamics_1993}.
-Using intra-cellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
+Using intracellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is, a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
 In particular, it was shown that the network of horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within the area.
-For generic visual scenes, this processes would generate a complex interplay between the dynamics of the sensory signal and the spatio- temporal determinants of these interactions and may underly anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
+For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions and may underlay anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
 %
@@
 Propagating waves occur in many excitable media and were recently found in neural systems from retina to neocortex.
-While propagating waves are clearly present under anaesthesia, whether they also appear during awake and conscious states remains unclear.
+While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remains unclear.
 One possibility is that these waves are systematically missed in trial-averaged data, due to variability.
@@
 Applying this method to single-trial voltage-sensitive dye imaging data, they show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
-A network model suggests that this reliability is the hallmark of the horizontal fibre network of superficial cortical layers.
+A network model suggests that this reliability is the hallmark of the horizontal fiber network of superficial cortical layers.
 Propagating waves with similar properties occur independently in secondary visual cortex, but maintain precise phase relations with the waves in primary visual cortex.
-These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatiotemporal frame for further neural interactions.
+These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatio-temporal frame for further neural interactions.
 
@@
 More recently, advanced recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
-Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
+Authors review these findings, consider the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function.
 In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
@@
 Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings.
-Here, authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
+Here, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
 They find that spontaneous traveling waves are a general property of these networks.
@@
 
-Interestingly, it can be shown that these travelling waves could have a measurable role on theactivity of the visual cortex.
+Interestingly, it can be shown that these travelling waves could have a measurable role on the activity of the visual cortex.
 This has been illustrated in a recent study studying the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
 The lrAM is the simple effect of perceiving a smooth motion when showing two dots in a temporal sequence and in close visual proximity.
-The lrAM is the building block of the use of sequences of images to induce the perception of smooth, realistic visual scenes which is at the base of movies seen in cinema theaters.
-In this study, authors used voltage-sensitive dye imaging to record the activty of the primary visual cortex of macaque monkeys to the presentation of the dots alone or in conjunction.
+The lrAM is the building block of the use of sequences of images to induce the perception of smooth, realistic visual scenes, which is at the base of movies seen in cinema theaters.
+In this study, the authors used voltage-sensitive dye imaging to record the activity of the primary visual cortex of macaque monkeys to the presentation of the dots alone or in conjunction.
 A probabilistic modelling has shown that the activity of the joint presentation induced a suppressive wave in the direction opposed to the perceived direction, shaping the formation of a wave of propagation travelling at a speed compatible with the perceived motion.
-A computational model validated the hypothesis that this process could be mediated by diffusion in the horizontal layers connecting the differnt locations within this cortical area.
+A computational model validated the hypothesis that this process could be mediated by diffusion in the horizontal layers connecting the different locations within this cortical area.
 %
@@
 For instance, statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-In particular, this was put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}.
+In particular, this was put in evidence by analyzing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}.
 This shows precise spatio-temporal differences in this recurrently connected assembly.
-The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
+The different aspects of information in the data are evaluated by a decoding strategy, highlighting the role of correlations.
 Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from M.
-Berry's lab~\citep{berry_spike_2022} and allows to perform and validate thiese hypothesis.
-However a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationaryn, which is incomptaible with the limits (in space and time) of a neural system.
+Berry's lab~\citep{berry_spike_2022} and allows performing and validate these hypothesis.
+However, a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationary, which is incompatible with the limits (in space and time) of a neural system.
 
@@
 He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
-The network is characterized by a timing-dependant learning rule for weights (STDP)~\citep{markram_regulation_1997} and by fixed conduction delays between neurons.
-It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
+The network is characterized by a timing-dependent learning rule for weights (STDP)~\citep{markram_regulation_1997} and by fixed conduction delays between neurons.
+It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
 Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
 The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike and thus to the crossing of the voltage threshold and to the discharge of a spike.
-According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and thus, constitute a polychronous group.
+This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike.
+According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and, thus, constitute a polychronous group.
 Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
@@
 % polychronization
-Interestingly the paper by~\citet{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general postive acceptance, but quite few works extended this seminal paper.
-Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the deterction of spiking motifs.
+Interestingly, the paper by~\citet{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but quite few works extended this seminal paper.
+Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs.
 A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}.
@@
 The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
-The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
 See also the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronized_2022}.
@@
 An extension of~\citep{izhikevich_polychronization_2006} is a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
-Indeed, while the original paper contained material within the text to reproduce the whole model, it was not complete for the repoduction of all results presented in that manuscript.
-This more recent work details how this code was slighlty corrected.
+Indeed, while the original paper contained material within the text to reproduce the whole model, it was not complete for the reproduction of all results presented in that manuscript.
+This more recent work details how this code was slightly corrected.
 It comes with python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
@@
 Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-Based on the fact that previous methods for studying polychronous neural groups (PNG) activation response to stimuli have been limited by the template-based methods used to identify PNG activation, authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
+Based on the fact that previous methods for studying polychronous neural groups (PNG) activation response to stimuli have been limited by the template-based methods used to identify PNG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
 They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.
-Stimulations of the trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
+Stimulation of the trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
 
@@
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-In~\citep{russo_cell_2017}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-%
-Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific recetpive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection in recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+%
+Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection in recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
@@
 Specifically, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
-Authors provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns.
+The authors provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns.
 Furthermore, the experimental results of applying an STDP-based spiking neural network was applied on Random Dot Kinematogram.
@@
 % Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
-In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
+In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron.
 In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
 In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity.
-This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
+This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show an efficient learning of polychronous groups.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Most of these theoretical models are trying to reproduce neurobiological observations and some others also propose applications to machine learning mothods such as image processing.
-For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
-The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
-The second stage takes the output of the spatiotemporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatiotemporal activity.
+Most of these theoretical models are trying to reproduce neurobiological observations, and some others also propose applications to machine learning methods such as image processing.
+For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
+The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern.
+The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity.
 For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
 Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters.
-Showing the potential of such representation
+%Showing the potential of such representation
 
@@
 Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
-This was aso implemented in a spiking neural network and reproduced the observation that neural activty was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
+This was aso implemented in a spiking neural network and reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
 This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the flash-lag illusion~\citep{khoei_flash-lag_2017}.
@@
  \section{Precise spiking motifs in neurobiology}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+also
 %
@@
 In most generic computational models the neural activity is assumed to be encoded in the firing rate.
-For instance, linear non linear L-NL models, the output is assumed to be distributed according to a Poisson law.
-As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
-This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
-As such, all these models are dependent on  a core definition of spike measures and we will review here how precise spiking motifs are taken into account in various neurobiological results.
+For instance, linear non-linear (L-NL) models, the output is assumed to be distributed according to a Poisson law.
+As such, a simple decoding strategy is to assume it is to be inferred for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
+This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
+As such, all these models are dependent on  a core definition of spike measures, and we will review here how precise spiking motifs are taken into account in various neurobiological results.
 %
@@
 There are different solution to provide with a distance between two given spike trains.
-A known measure is the Victor-Pupura distance which showed that their findings are inconsistent with Poisson models of spike trains~\citep{victor_nature_1996}.
+A known measure is the Victor-Purpura distance, which showed that their findings are inconsistent with Poisson models of spike trains~\citep{victor_nature_1996}.
 In a later study,~\citet{van_rossum_novel_2001} have defined a distance that has a time constant as a parameter.
 Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
-Such distances were extended to non-Euclidean metrics~\citep{} and use morphological manipulmations to compute spike train dissimilarity~\citet{kruz}.
-Mathelmatically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepency measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
+Such distances were extended to non-Euclidean metrics~\citep{} and use morphological manipulations to compute spike train dissimilarity~\citet{kruz}.
+Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepancy measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
 These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures.
@@
 Let's now focus on distances defined based on a generative model formed by using precise spiking motifs.
-A review of the methods for neuronal response latency estimation exists that includes bayesian binning~\citep{levakova_review_2015}.
-More braodly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
+A review of the methods for neuronal response latency estimation exists that includes Bayesian binning~\citep{levakova_review_2015}.
+More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
 This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance).
 A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}.
-By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
+By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
 This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
@@
 
-One algorithm capable of achieveing such daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
+One algorithm capable of achieving such daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
 Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
@@
 In~\citet{stringer_high-dimensional_2019}, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
-Then, in~\citep{stringer_high-precision_2021}, authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
 Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
@@
 % \item
-Using a t-SNE projection with knwon labels validated that this clustering method can retrieve all patterns from the data.
+Using a t-SNE projection with known labels validated that this clustering method can retrieve all patterns from the data.
 % \item
@@
 % \end{itemize}
-The limits of this method is that it is block-based and stricly specialized for the task at hand.
-To overcome these diculties a novel method was recently developped~\citep{sotomayor-gomez_spikeship_2021}.
+The limits of this method is that it is block-based and strictly specialized for the task at hand.
+To overcome these difficulties a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
 
@@
 
-In the paper by~\citet{russo_cell_2017}, authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.
+In the paper by~\citet{russo_cell_2017}, the authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.
 It uses  sliding window as in~\citep{grun_unitary_2002} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
 They extend the measure to multiple lags~\citep{torre_synchronous_2016}.
-The core measure is based on a non-stationarity-corrected parametric test statistic for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+The core measure is based on a non-stationarity corrected parametric test statistic for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
 
@@
 
-To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
-They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
+To overcome the limits of models which require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
+They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
 
@@
 Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents speed of the mouse.}\label{fig:haimerl}
+The black line represents the speed of the mouse.}\label{fig:haimerl}
 \end{figure}
@@
 They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019} below).
-We have seen previously that neurobiologists havbe described Stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019} below).
+We have seen previously that neurobiologists have described Stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019} below).
 Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons in vivo and in vitro.
@@
 Indeed,~\citet{ikegaya_synfire_2004} demonstrated that in cortical activity, one can find a repetition of several motifs in spike activity (duration around 1s +/- .5 s, some events in motifs are of similar size but sometimes absent).
-These sequences can be specific of a particular layer or colomn, can be synchronized with network activity oscillation and can involve several cells.
-They also demonstated that these sequences can form supersequences, so-called cortical songs.
-It consist of the assembly of several sequences which repeat in a specific order with a compressed timing.
+These sequences can be specific of a particular layer or column, can be synchronized with network activity oscillation, and can involve several cells.
+They also demonstrated that these sequences can form super sequences, so-called cortical songs.
+It consists of the assembly of several sequences which repeat in a specific order with a compressed timing.
 \emph{In vivo} spontaneous activity also reveals repeating sequences.
 About 3000 sequences, each involving 3-10 cells out of about 900, and last up to 3 seconds.
-Sequences had specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster and were associated with a structured spatial organization of the neurons that formed them.
+Sequences had specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and were associated with a structured spatial organization of the neurons that formed them.
 %In cortical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
@@
 1.
-from~\citep{ikegaya_synfire_2004} Repeated motifs of spontaneous synaptic activity in vitro and in vivo.
+from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity in vitro and in vivo.
 (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
-Panels show segments (red) of the same voltage-clamp recording from the same cell repeating seconds or minutes after the initial occurrence (blue).
+Panels show segments (red) of the same voltage-clamp recording from the same cell, repeating seconds or minutes after the initial occurrence (blue).
 Arrows indicate timings of repeated PSCs.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-We have seen that precise spiking moifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobuological data. 
+We have seen that precise spiking motifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobiological data. 
 A third class of models coming from the field of neuromorphic engineering, a priori different from that used in neuroscience, are of interest in this review as they often take advantage of minute temporal information.
@@
 %----------------------------%
-This section has provided evidence that polychronous groups are an important apsect of information representation in biology with important application in data analysis and neuromorphic engineering.
+This section has provided evidence that polychronous groups are an important aspect of information representation in biology, with important application in data analysis and neuromorphic engineering.
 The rest of this review paper is organized as follows.
@@
 Our second result is to propose a novel learning method for learning PGs in raster plots in a self-supervised manner.
-Finally we demonstrate the use of this algorithm to the output of an event-based camera and how this may separate independent components from the stream of events.
+Finally, we demonstrate the use of this algorithm to the output of an event-based camera and how this may separate independent components from the stream of events.
 This end-to-end event-based computational brick could help improve the performance of current Spiking Neural Network solution currently used in neuromorphic chips.
@@
 \note{\textbf{TODO amelie} section sur l'apprentissage / la plasticité des délais dans la biologie}
+
+% (mainly based on JOBIM poster)
 

### 2022-10-26 manuscript.tex
abstract

@@
 Why do neurons communicate through spikes?
-By definition, a spike, or action potential, is an all-or-none event which occurs at asynchronous times.
-In other words, spikes on one side occur or not without further details and second can occur at any continuous time.
+By definition, a spike is an all-or-none event which occurs at continuous times.
+In other words, spikes are on one side binary, they occur or not without further details,  and second can occur at any asynchronous time, without a centralized clock.
 This stands in stark contrast to the discretized timing classically used in digital processing at the base of modern-day neural networks.
@@
 It is also an emerging topic in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal mission in the age of the digital economy and of global warming.
+In this review paper, we provide evidence from biology and engineering that the precise timing of spikes plays a key role for our understanding of efficient neural networks.
 }

### 2022-10-26 manuscript.tex
introduction

@@
 It is also an emerging topic in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal mission in the age of the digital economy and of global warming.
-In this review paper, we provide evidence from biology and engineering that the precise timing of spikes plays a key role for our understanding of efficient neural networks.
+In this review paper, we provide evidence from biology and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
 }
@@
 %
- Neural activity is directly influenced by our immediate environment and by internal states in order to generate actions.
-These are essential for survival, which is the sole determinant in the light of natural selection.
-A central question of modern neuroscience is to better understand the essence of neural activity.
+ Neural activity is directly influenced by our immediate environment and by internal states, and is shaped in order to generate actions.
+The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection.
+A central question of modern neuroscience is to better understand the essence of neural activity, as one can observe it in neurobilogical experiments.
 One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
@@
 Nevertheless, we will use this terminology in all generality.
-In this respect, it is known that this neural activity is related to a specific variable, and since Galvani's experiments, we know that electrical activity can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-Among various phenomenon, a well studied way of communication between neurons are action potentials, or spikes.
-In the scope of this article we focus mainly on spikes, which are brief (about one millisecond) electrochemical impulses that propagate along the axons of neurons.
-These have the particularity of being essentially binary in their amplitude (i.e.
+In this respect, it is known that neural activity may be related to a specific measurable variable, and since Galvani's experiments, we know that an electrical activity applied on musuclar nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
+Among various phenomenon, a well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes.
+Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and eventually further propagated along its "ouput" along the axon to reach efferent neurons.
+In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
+First, these have the particular property of being essentially binary in their amplitude (i.e.
 they are prototypical, all or nothing).

### 2022-10-26 manuscript.tex
introduction

@@
 Let us start with a review of the state of the art on the role of dynamics in vision.
-Vision is the set of processes that allow us to make sense of the luminous world, and is an area intensively studied in neuroscience with respect to the question of neural code.
+Vision is the set of processes that allow us to make sense of the luminous world, and is an area intensively studied in neuroscience, in particular with respect to deciphering the neural code.
 In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between external visual space and its representation on the retina.
@@
 In particular, there have been some remarkable findings when studying the dynamics of vision.
-Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
+For instance, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
 Its first experiment consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
-The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neural activity.
-These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of $120~\ms$ when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
+The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
+These results have been extended to several species including primates, and was also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of $120~\ms$ when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
 This fast processing also explains the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
@@
 An important consequence of the speed of processing of vision is that it implies that it is carried out using only very few spikes.
-Indeed, if we consider that a behavioral response in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
+Indeed, if we consider that a behavioral response in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system.
+This was was indeed modeled in feed-forward models trained with back-propagation~\cite{serre_feedforward_2007,jeremie_ultrafast_2022}.
+This imposes that the processing in a single area is performed with a reduced number of spikes.
 Altogether, these results show that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.

### 2022-10-26 manuscript.tex
introduction

@@
 Indeed, if we consider that a behavioral response in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system.
-This was was indeed modeled in feed-forward models trained with back-propagation~\cite{serre_feedforward_2007,jeremie_ultrafast_2022}.
+This was was indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
 This imposes that the processing in a single area is performed with a reduced number of spikes.
@@
 % Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
-It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned, but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
-When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
+Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned, but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
+However, if that same neuron is now driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
 This is consistent with the differential role of different stimulus frequencies (for instance the gamma range around $80~\si{\Hz}$) on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
@@
 In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
-Many models have used these properties in temporal coding to build fast image categorization networks.
+Many models have used these properties in temporal coding to build fast image categorization networks~\citep{{gautrais_rate_1998,delorme_spikenet_1999,perrinet_coding_2004}.
 These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}.

### 2022-10-26 manuscript.tex
Precise spike timing

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Precise spike timings in neural assemblies}
+\section{Precise spike timing in neural assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{paugam-moisy_computing_2012}.
-The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion.
+The book then demonstrates that this could take the form of ``synfire chains,'' that is, the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
 Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
 %
-In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_1970}.
+In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001}.
 This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size.

### 2022-10-26 manuscript.tex
Precise spike timing

@@
 In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
-Many models have used these properties in temporal coding to build fast image categorization networks~\citep{{gautrais_rate_1998,delorme_spikenet_1999,perrinet_coding_2004}.
+Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}.
 These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Precise spike timing in neural assemblies}
+\section{Role of precise spike timing in neural assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding neural activity using spike distances}
+\subsection{Decoding neural activity using spike distances}\label{sec:discussion}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-https://www.overleaf.com/project/633bf90497b5b325a3ff692e
-\subsection{Extending the decoding athttps://www.overleaf.com/project/633bf90497b5b325a3ff692e multiple time scales}
+\subsection{Extending the decoding at multiple time scales}
 
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}\label{discussion}
+\section{Discussion}\label{sec:discussion}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-10-26 manuscript.tex
Precise spike timing

@@
 
-Propagating waves occur in many excitable media and were recently found in neural systems from retina to neocortex.
-While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remains unclear.
-One possibility is that these waves are systematically missed in trial-averaged data, due to variability.
+Propagating waves occur in many excitable media and were recently found in neural systems, for instance in the retina or the neocortex.
+While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries.
+One possibility is that these waves were systematically missed in trial-averaged data, due to variability.
 A recent work~\citep{muller_stimulus-evoked_2014} presents a method for detecting propagating waves in noisy multichannel recordings.
@@
 This has been illustrated in a recent study studying the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
-The lrAM is the simple effect of perceiving a smooth motion when showing two dots in a temporal sequence and in close visual proximity.
+The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in close visual proximity.
 The lrAM is the building block of the use of sequences of images to induce the perception of smooth, realistic visual scenes, which is at the base of movies seen in cinema theaters.
@@
 In particular, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which he defined as \emph{polychronous groups}.
+In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
 
@@
 Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure
-~\ref{fig:haimerl} from~\citep{haimerl_internal_2019} below).
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019} reproduced below).
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding neural activity using spike distances}\label{sec:discussion}
+\subsection{Decoding neural activity using spike distances}\label{sec:detection}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-10-26 manuscript.tex
Precise spike timing

@@
 The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient and try to ``cross-compile'' classical Neural Network to a spiking architecture.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture.
 This approach is quite successful, and SNNs approach in some case the performance of Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
@@
 There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will review here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized.
+To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized.
 To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous`` events, that is, volleys of spikes which are stable up to certain synaptic delays.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-First, we will review models for the detection of polychronous groups.
+% First, we will review models for the detection of polychronous groups.
 % Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines. 
-In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
+% In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
 %Here, we will review current literature on the detection of such motifs in generic raster plots.
@@
 %The next section will focus on precise spike time motifs. 
-We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
+% We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
 In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
@@
 
-A notable exception is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+A notable example of a SNN using precise spike timings is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
 In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.

### 2022-10-26 manuscript.tex
Precise spike timing

@@
 The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
 This approach is quite successful, and SNNs approach in some case the performance of Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
@@
 Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
-Such distances were extended to non-Euclidean metrics~\citep{} and use morphological manipulations to compute spike train dissimilarity~\citet{kruz}.
+Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citet{kreuz_measuring_2007}.
 Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepancy measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
 These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures.
+% ~\citep{}
 

### 2022-10-27 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Training delays ~\citep{state_training_2019}
 A whole range of models of latency motifs detection are using STDP-based learning rules.
@@
 %
-Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection in recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}, in particular for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.

### 2022-10-27 manuscript.tex
section 3

@@
 One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-This approach is quite successful, and SNNs approach in some case the performance of Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
+This approach is quite successful, and SNNs reach in some case a similar performance as Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
 However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
 There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized.
-To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous`` events, that is, volleys of spikes which are stable up to certain synaptic delays.
+To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{paugam-moisy_computing_2012}.
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
 These models will be compared in light of neuroscientific and computational perspectives.
@@
 These can interact such that the feed-back connections are predicting the feed-forward inputs.
-
-A notable example of a SNN using precise spike timings is the polychronization model of Izhikevich~\citep{izhikevich_polychronization_2006}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
-In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-Notably, Izhikevich~\citep{izhikevich_polychronization_2006} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
-This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
-
-% REDONDANCE
-Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
-However, taking axonal propagation times into account, synchronous signals may not coincide significantly at the post-synaptic neuron contrary to asynchronous or polychronous signals.
-The term polychronous was first introduced by~\citet{izhikevich_polychronization_2006}.
-He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
-The network is characterized by a timing-dependent learning rule for weights (STDP)~\citep{markram_regulation_1997} and by fixed conduction delays between neurons.
+As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on a firing-rate based encoding of information.
+A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}.
+This model is based on a random recurrent model of spiking neurons including such synaptic delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
+It is worth mentioning that the set of delays is set randomly at initialization in a range of biologically realistic dealys (from $0$ to $20~\ms$) and is then frozen for the rest of the simulation, only the weights are changed using the STDP rule.
 It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
@@
 Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
+In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
+This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
 
@@
 \centering
-\note{insert fig 2 of izhikevich 2006}
+\note{insert fig 2 of izhikevich 2006 - or an open-source reproduction ?}
 %includegraphics{figures/event_driven_computations.png}
@@
 % polychronization
-Interestingly, the paper by~\citet{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but quite few works extended this seminal paper.
+Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper.
 Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs.
@@
 The limits of this model is that it's output is only binary and that it's storage capacity are limited.
-The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
+The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
 See also the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronized_2022}.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-An extension of~\citep{izhikevich_polychronization_2006} is a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
-Indeed, while the original paper contained material within the text to reproduce the whole model, it was not complete for the reproduction of all results presented in that manuscript.
+An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
+Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript.
 This more recent work details how this code was slightly corrected.

### 2022-10-27 manuscript.tex
section 3

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Extensions of Izhikevitch's polychronization model}
+% \subsection{Extensions of Izhikevitch's polychronization model}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 It comes with python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
-
+%
 Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
@@
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-%
-Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}, in particular for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+%
+Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
+In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
@@
 Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
+This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
 % Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
@@
 \subsection{Scaling up to very large scale data}
+
+In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+
 Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
@@
 % autres méthodes
+spatitemporal kernels: \citep{yu_stsc-snn_2022}
 

### 2022-10-27 analyze_bibs.py
analyzing citations

@@ -1,0 +1,53 @@
+import re
+
+if False:
+    latex = r"""
+    blablabla
+    Author et. al \cite{author92} bla bla. % should match
+    \citep{author93} % should match
+    \nocite{author94} % should match
+    100\%\nocite{author95} % should match
+    100\% \nocite{author95} % should match
+    %\nocite{author96} % should not match
+    \cite{author97, author98, author99} % should match
+    \nocite{*} % should not match
+    """
+else:
+    with open('manuscript.tex') as f:
+        latexlines = f.readlines()
+
+    # latex = ''.join(line.strip() for line in latexlines)
+    latex = ''.join(line for line in latexlines)
+    # latex = latex.replace('\\', '\')
+
+rx = re.compile(r'''(?<!\\)%.+|(\\(?:no)?citep?\{((?!\*)[^{}]+)\})''')
+# rx = re.compile(r'''^(?!(%\\(?:no)?cite\w*\{([^}]*?)\}))[^*\n]*$''')
+# rx = re.compile(r'''^(?!(%\\(?:no)?cite\w*\{([^}]*?)\}))[^*\n]*$''')
+
+citekeys = [m.group(2) for m in rx.finditer(latex) if m.group(2)]
+# print(citekeys)
+
+citekey_list = []
+for citekey in citekeys:
+    for citekey_ in citekey.split(', '):
+        citekey_list.append(citekey_)
+import numpy as np
+citekey_list = np.unique(citekey_list)
+# print(citekey_list)
+
+
+with open('references.bib') as f:
+        bibtexlines = f.readlines()
+
+# latex = ''.join(line.strip() for line in latexlines)
+bibtex = ''.join(line for line in bibtexlines)
+
+rx = re.compile(r'''@\w+{([\w:-]+)''')
+bibkeys = [m.group(1) for m in rx.finditer(bibtex) if m.group(1)]
+bibkeys = np.unique(bibkeys)
+# print(bibkeys)
+    
+for bibkey in bibkeys:
+    if not(bibkey in citekey_list):
+        print('This key in not used in the paper', bibkey)
+

### 2022-10-27 analyze_bibs.py
analyze biblio

@@ -20,7 +20,7 @@ else:
     latex = ''.join(line for line in latexlines)
     # latex = latex.replace('\\', '\')
 
-rx = re.compile(r'''(?<!\\)%.+|(\\(?:no)?citep?\{((?!\*)[^{}]+)\})''')
+rx = re.compile(r'''(?<!\\)%.+|(\\(?:no)?cite[pt]?\{((?!\*)[^{}]+)\})''')
 # rx = re.compile(r'''^(?!(%\\(?:no)?cite\w*\{([^}]*?)\}))[^*\n]*$''')
 # rx = re.compile(r'''^(?!(%\\(?:no)?cite\w*\{([^}]*?)\}))[^*\n]*$''')
 
@@ -29,8 +29,8 @@ citekeys = [m.group(2) for m in rx.finditer(latex) if m.group(2)]
 
 citekey_list = []
 for citekey in citekeys:
-    for citekey_ in citekey.split(', '):
-        citekey_list.append(citekey_)
+    for citekey_ in citekey.split(','):
+        citekey_list.append(citekey_.strip(' '))
 import numpy as np
 citekey_list = np.unique(citekey_list)
 # print(citekey_list)
@@ -46,8 +46,11 @@ rx = re.compile(r'''@\w+{([\w:-]+)''')
 bibkeys = [m.group(1) for m in rx.finditer(bibtex) if m.group(1)]
 bibkeys = np.unique(bibkeys)
 # print(bibkeys)
+
+for citekey in citekey_list:
+    if not(citekey in bibkeys):
+        print('This key in not present in the bibliography', citekey)
     
 for bibkey in bibkeys:
     if not(bibkey in citekey_list):
         print('This key in not used in the paper', bibkey)
-

### 2022-10-27 manuscript.tex
fix biblio

@@
  \section{Precise spiking motifs in neurobiology}
-also
 %
@@
 Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citet{kreuz_measuring_2007}.
-Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepancy measure~\citep{weyl_ber_1916} which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
+Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepancy measure~\citep{weyl_ueber_1916} which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
 These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures.
@@
 
-Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks in vivo~\citep{Isbister} They discovered (in vivo) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
+Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks in vivo~\citep{isbister_clustering_2021} They discovered (in vivo) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
 % dynamical models
@@
 % development
-scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2021}
+scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2022}
 

### 2022-11-04 manuscript.tex
section 3
Update on Overleaf.

@@
 %%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% cite or die
+% ❯ ipython analyze_bibs.py
+
 % In recognition of your well-respected work, we are inviting you to submit one feature contribution (i.e., a long review or research paper) to an MDPI journal free of charge.
@@
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0003-3916-0514}
+\newcommand{\orcidauthorC}{0000-0001-8821-5556}
 \newcommand{\orcidauthorD}{0000-0002-3107-4788}
@@
 \orcidA{}, 
-Amélie Gruel $^{2}$, % $^{2,\dagger,\ddagger}$, 
-Jean Martinet $^{2}$,  
+Amélie Gruel $^{2}$\orcidB{}, % $^{2,\dagger,\ddagger}$, 
+Jean Martinet $^{2}$\orcidC{},  
 and Laurent Perrinet $^{1,}$\orcidD{},*}
@@
 By definition, a spike is an all-or-none event which occurs at continuous times.
-In other words, spikes are on one side binary, they occur or not without further details,  and second can occur at any asynchronous time, without a centralized clock.
+In other words, spikes are on one side binary, arising or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock.
 This stands in stark contrast to the discretized timing classically used in digital processing at the base of modern-day neural networks.
-In the living world, neural systems almost systematically use this so-called event-based representation.
-A better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data.
+As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data.
 It is also an emerging topic in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal mission in the age of the digital economy and of global warming.
@@
 The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection.
-A central question of modern neuroscience is to better understand the essence of neural activity, as one can observe it in neurobilogical experiments.
+A central question of modern neuroscience is to better understand the essence of neural activity, as one can observe it in neurobiological experiments.
 One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
@@
 Nevertheless, we will use this terminology in all generality.
-In this respect, it is known that neural activity may be related to a specific measurable variable, and since Galvani's experiments, we know that an electrical activity applied on musuclar nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
+In this respect, it is known that neural activity may be related to a specific measurable variable, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
 Among various phenomenon, a well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes.
-Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and eventually further propagated along its "ouput" along the axon to reach efferent neurons.
+Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and eventually further propagated along its "output" axon to reach efferent neurons.
 In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
-First, these have the particular property of being essentially binary in their amplitude (i.e.
-they are prototypical, all or nothing).
+First, these have the particular property of being essentially binary in their amplitude, i.e.
+they are prototypical, "all or nothing".
 The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
 Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
-However, computational neuroscience models have suggested that precise timing may also play a crucial role, and that neurons may be as well integrators than synchrony detectors~\citep{paugam-moisy_computing_2012}.
+However, computational neuroscience models have suggested that precise timing may also play a crucial role, and that neurons may be as well integrators than synchrony detectors~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
 We will investigate this very hypothesis in this review.
@@
 Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma.
-In comparison to a classical analogical vector of inputs, this \emph{event-based representation} present in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this fine-scaled temporal dimension.
+In comparison to a classical analogical vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this fine-scaled temporal dimension.
 This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
@@
 In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between external visual space and its representation on the retina.
-The origins of this  neuroscientific question can be found in the first experiments of Pierre Flouren showing the relationship between sensations, perception and activity in the cerebral cortex using lesions in animals~\citep{pearce_marie-jean-pierre_2009}.
+The origins of this  neuroscientific question can be found in the first experiments of Pierre Flourens showing the relationship between sensations, perception and activity in the cerebral cortex using lesions in animals~\citep{pearce_marie-jean-pierre_2009}.
 In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as orientation or motion.
-Remarkably, there is a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
+There is remarkably a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
 However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitute essential pieces to this puzzle~\citep{carandini_normalization_2012}.
@@
 The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
-These results have been extended to several species including primates, and was also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of $120~\ms$ when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
+These results have been extended to several species including primates.
+Different experimental protocols have shown for example that the response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
 This fast processing also explains the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
-The performances decrease progressively as the frequency of presentation of the images increases.
+As expected, the performances decrease progressively as the frequency of presentation of the images increases.
 However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only $14~\ms$ per image.
 
-This speed of the visual cortex in primates, although surprising, is  compatible with the latencies that are recorded at the neuro-physiological level.
-Indeed, when an image is presented to the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex, and it takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}, see Figure~\ref{fig:thorpe}.
+Although surprising, this speed of the visual cortex in primates is  compatible with the latencies that are recorded at the neuro-physiological level.
+The rapid propagation of the visual information in the thalamus then in the primary visual cortex (see Figure~\ref{fig:thorpe}) takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}.
 This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
@@
 Indeed, if we consider that a behavioral response in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system.
-This was was indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
-This imposes that the processing in a single area is performed with a reduced number of spikes.
+This was indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
+This imposes that the processing in a single area is performed with a reduced number of spikes, as it has been demonstrated that one spike requires a significant amount of time ($X~\ms$) to traverse the synapse of a neuron.
 Altogether, these results show that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
@@
 %
-At the level of the dynamic processing of visual information, it has been shown in the retina that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
-In particular, these results show that using  the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
+At the level of the dynamic processing of visual information, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
+In particular, these results show that the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
 These results have been extended to natural images and show a qualitatively similar behavior.
-The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
+Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
 This evidence found in the retina  can be extended to the visual cortex.
@@
 First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
+One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods.
+The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
 This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
-This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events~\citep{rasetto_challenges_2022}.
+This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas which represents the visual information as events, similarly to the representations described above and instead of using a basic frame-based representation~\citep{rasetto_challenges_2022}.
 This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
@@
 %
-In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
+In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
 The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models.
-While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{paugam-moisy_computing_2012}.
+While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
 The book then demonstrates that this could take the form of ``synfire chains,'' that is, the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
 Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
-%
-In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001}.
-This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size.
+One possible function of this synchronization may serve the binding of information distributed in the brain~\citep{singer_visual_1995, roelfsema_visuomotor_1997}.
+%
+In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire neurons in different groups of similar size.
 Each neuron of one group is connected by an excitatory synapse to the next.
@@
 Recursively applying this to a sequence of groups generates either a synfire propagation or not.
-A simple simulation is shown in Figure~\ref{fig:diesman}.
-A crucial aspect of this emergence is in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
-This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
+A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
+A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model÷\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
+This was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
 
@@
 %----------------------------%
-Attempts have been made to detect such synfire chains in neurobiological data~\citep{schrader_detecting_2008}, which envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
+Attempts have been made to detect such synfire chains in neurobiological data.
+\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
 In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
@@
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016,torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
@@
 The synchronicity will generate rhythms at different range of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down.
-Such a theory has surprisingly been validated in EEG recordings, for example, to explain the continuous wagon-wheel illusion which corresponds to the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
-More generally, it can be shown that the phase of alpha (about $10$ Hz) oscillations  is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
+Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion , i.e.
+to the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
+More generally, it can be shown that the phase of alpha oscillations (about $10$ Hz) is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
 %
@@
 
-Propagating waves occur in many excitable media and were recently found in neural systems, for instance in the retina or the neocortex.
+Propagating waves in the neuronal response occur in many excitable media and were recently found in neural systems, for instance in the retina or the neocortex.
 While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries.
@@
 Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
-Further, by modulating local network state, they can shape responses to incoming inputs as observed in vivo.
-Such waves also occur in motor areas and~\citep{linden_movement_2021} have presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+Further, by modulating local network state, they can shape responses to incoming inputs as observed \textit{in vivo}.
+Such waves also occur in motor areas and~\citet{linden_movement_2021} have presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
 
 Interestingly, it can be shown that these travelling waves could have a measurable role on the activity of the visual cortex.
-This has been illustrated in a recent study studying the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
+This has been illustrated in a recent study investigating the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
 The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in close visual proximity.
@@
 For instance, statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-In particular, this was put in evidence by analyzing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}.
-This shows precise spatio-temporal differences in this recurrently connected assembly.
+In particular, this was put in evidence by analyzing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}, which shows precise spatio-temporal differences in this recurrently connected assembly.
 The different aspects of information in the data are evaluated by a decoding strategy, highlighting the role of correlations.
@@
 Berry's lab~\citep{berry_spike_2022} and allows performing and validate these hypothesis.
-However, a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationary, which is incompatible with the limits (in space and time) of a neural system.
+However, a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
 
 Moreover, there is a consensus for rate-based coding models in computational and biological neuroscience.
-Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015} and on the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
+Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015}.
+Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
 This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996} or touch~\citep{johansson_first_2004}.
 In particular, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
+Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
 
@@
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
-The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
+The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
 One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
@@
 There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{paugam-moisy_computing_2012}.
+To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
 To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
@@
 This model is based on a random recurrent model of spiking neurons including such synaptic delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
-It is worth mentioning that the set of delays is set randomly at initialization in a range of biologically realistic dealys (from $0$ to $20~\ms$) and is then frozen for the rest of the simulation, only the weights are changed using the STDP rule.
+It is worth mentioning that the set of delays is set randomly at initialization in a range of biologically realistic delays (from $0$ to $20~\ms$) and is then frozen for the rest of the simulation, only the weights are changed using the STDP rule.
 It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
@@
 This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike.
-According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and, thus, constitute a polychronous group.
+According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, constitute a polychronous group.
 Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
 In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
+This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
 
@@
 Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs.
-A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}.
-This model is in particular reviewed in~\citep{gutig_spike_2014}.
-The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
-This aims at extending the perceptron which does not incorporate a spike timing framework.
-The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
-The limits of this model is that it's output is only binary and that it's storage capacity are limited.
+A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}.
+The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron which does not incorporate a spike timing framework.
+The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a STDP rule.
+The limits of this model is that its output is only binary and that its storage capacity are limited.
 The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
-See also the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronized_2022}.
+Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
 %
@@
 This more recent work details how this code was slightly corrected.
-It comes with python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
+It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
 %
@@
 They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.
-Stimulation of the trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
+Stimulation of the trained network produces the activation of a PNG, i.e.
+the propagation of firing activity through multiple layers due to convergent patterns of firing.
 
@@
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-%
-Modified STPD rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001} and were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
+To address  how transmission delays and STDP can jointly determine the emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
+%
+Modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
+They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
 In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
@@
 This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
+Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
 % Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
@@
 In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
-In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity.
-This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show an efficient learning of polychronous groups.
+In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Most of these theoretical models are trying to reproduce neurobiological observations, and some others also propose applications to machine learning methods such as image processing.
+Most of these theoretical models are trying to reproduce neurobiological observations, while others propose applications to machine learning methods such as image processing.
 For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
@@
 For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
-Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters.
-%Showing the potential of such representation
+Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters thus showing the potential of such representation.
 
@@
 Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
-This was aso implemented in a spiking neural network and reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
+This was also implemented in a spiking neural network and reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
 This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-We will review some more applications of using precise spiking motifs in neuromorphic engineering, but will first review current evidence in neurobiology, that is, from actual recordings in biological tissues.
+In the following section, we will review some more applications of using precise spiking motifs in neuromorphic engineering, but will first study current evidence in neurobiology, that is, from actual recordings in biological tissues.
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-% In Bellec~\citep{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
+For instance, the output of linear non-linear (L-NL) models is assumed to be distributed according to a Poisson law.
+As such, a simple decoding strategy is to assume it is to be inferred for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
+This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
+An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
+In Bellec~\citet{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
 The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
@@
 It uses the publicly available \href{V1-dataset}{http://crcns.org/data-sets/vc/pvc-11}, which allows to supervise the model with the input being the movie and the output the spikes recorded.
-In most generic computational models the neural activity is assumed to be encoded in the firing rate.
-For instance, linear non-linear (L-NL) models, the output is assumed to be distributed according to a Poisson law.
-As such, a simple decoding strategy is to assume it is to be inferred for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
-This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
+These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
+Using the right metric, as implemented in that paper by the models themselves, is essential in better understanding neural data.
 As such, all these models are dependent on  a core definition of spike measures, and we will review here how precise spiking motifs are taken into account in various neurobiological results.
@@
 A known measure is the Victor-Purpura distance, which showed that their findings are inconsistent with Poisson models of spike trains~\citep{victor_nature_1996}.
-In a later study,~\citet{van_rossum_novel_2001} have defined a distance that has a time constant as a parameter.
+In a later study,~\citet{van_rossum_novel_2001} have defined a distance that uses a time constant as a parameter.
 Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
 Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citet{kreuz_measuring_2007}.
-Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepancy measure~\citep{weyl_ueber_1916} which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
-These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures.
+Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
+These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
 % ~\citep{}
@@
 Let's now focus on distances defined based on a generative model formed by using precise spiking motifs.
-A review of the methods for neuronal response latency estimation exists that includes Bayesian binning~\citep{levakova_review_2015}.
-More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2010}.
-This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance).
+A review of the methods for neuronal response latency estimation exists and includes Bayesian binning~\citep{levakova_review_2015}.
+More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
+This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}.
 A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}.
 By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
-This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+This was recently extended in~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
 The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
@@
 Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
-Bridging across different spatial and temporal scales, combination of in vivo two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
+Bridging across different spatial and temporal scales, combination of \textit{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
 The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurophysics to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
 
-One algorithm capable of achieving such daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
+One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
 Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
 The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
-In~\citet{stringer_high-dimensional_2019}, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+In~\citet{stringer_spontaneous_2019,stringer_high-dimensional_2019}, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
 Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
@@
 The algorithm measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other.
-Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
+Many approaches to this problem are supervised.
+In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
 % \item
@@
 The limits of this method is that it is block-based and strictly specialized for the task at hand.
-To overcome these difficulties a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
+To overcome these difficulties, a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
 
@@
 In particular, they show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
-Top-down attention drives interareal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
+Top-down attention drives inter-areal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
 
@@
 We have seen previously that neurobiologists have described Stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019} below).
-Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons in vivo and in vitro.
+Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons \textit{in vivo} and \textit{in vitro}.
 These patterns repeat after minutes, maintaining millisecond accuracy.
@@
 It consists of the assembly of several sequences which repeat in a specific order with a compressed timing.
-\emph{In vivo} spontaneous activity also reveals repeating sequences.
-About 3000 sequences, each involving 3-10 cells out of about 900, and last up to 3 seconds.
-Sequences had specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and were associated with a structured spatial organization of the neurons that formed them.
+\emph{In vivo} spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3-10 cells out of about 900, and lasting up to 3 seconds.
+Sequences have specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and are associated with a structured spatial organization of the neurons that formed them.
 %In cortical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
@@
 1.
-from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity in vitro and in vivo.
+from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity \textit{in vitro} and \textit{in vivo}.
 (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
@@
 The top traces show the motifs superimposed on each other (blue, green, and red), the middle traces show these same traces individually, and the bottom traces show temporally magnified regions of the motifs (a to c).
-(C) Repeated sequences of intracellular current-clamp recordings in vivo.
+(C) Repeated sequences of intracellular current-clamp recordings \textit{in vivo}.
 Two (i) and three (ii) repetitions of motifs are shown.
@@
 It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns.
-They " used random waveforms to probe the formation of new memories for arbitrary complex sounds.
-A behavioral measure was designed, based on the detection of repetitions embedded in noises up to 4 s long." The task is to detect the repetition of the same (frozen) noise within a trial.
-" Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block." they showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case.
-Note that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``Time reversal had no significant effect on the RefRN advantage'' (quite surprising).
-The Learning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).
+They ``used random waveforms to probe the formation of new memories for arbitrary complex sounds.
+A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \s$ long''.
+The task is to detect the repetition of the same (frozen) noise within a trial.
+" Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block." They showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case.
+Note that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``time reversal had no significant effect on the RefRN advantage'' (which is quite surprising).
+The mearning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).
 
@@
 We have seen that precise spiking motifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobiological data.
+Interestingly some of these computational neuroscience models aim at solving the spike sorting method using STDP rules~\citep{bernert_fully_2017,bernert_attention-based_2018}
 A third class of models coming from the field of neuromorphic engineering, a priori different from that used in neuroscience, are of interest in this review as they often take advantage of minute temporal information.
@@
 Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
-The task consists in localizing a target using only the relative spike timing of eight spatially-separated vibration sensors.
-Haessig2020a
+The task consists in localizing a target using only the relative spike timing of eight spatially-separated vibration sensors~\citep{haessig_event-based_2020}.
 
@@
 % autres méthodes
-spatitemporal kernels: \citep{yu_stsc-snn_2022}
-
-\subsubsection{Methods using precise timings}
+Spatio-temporal kernels: \citep{yu_stsc-snn_2022}
+
+\subsection{Methods using precise timings}
 
@@
 
+ ~\citep{lagorce_hots_2017}
+
+ ~\citep{benosman_asynchronous_2012}
+
+ ~\citep{benosman_event-based_2014}
+
+  ~\citep{dardelet_event-by-event_2021}
+
 % Grimaldi CBMI + PAMI
+~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}
 
@@
 
-Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks in vivo~\citep{isbister_clustering_2021} They discovered (in vivo) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
+Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks \textit{in vivo}~\citep{isbister_clustering_2021} They discovered (\textit{in vivo}) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
 % dynamical models
@@
 
-\note{\textbf{TODO amelie} section sur l'apprentissage / la plasticité des délais dans la biologie}
+\subsection{What biological mechanism enables delay learning?}
+
+Despite the extensive computer evidence of the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
+The discovery of the existence of an equivalent biological mechanism in Vertebrates as well as the demonstration of the importance of its role in various developed behaviours allow us to put these doubts to rest.
+In the following paragraphs we will successively describe the first biological observations of biological delay learning, identify myelination as an important actor and finally study this phenomenon at the molecular level.
+
+% First biological observations
+
+One of the first significant evidence of any neuronal delay in the information propagation within the animal brain came from Young's study of the squid giant axon \cite{young_1938}:
+
+The interaural time difference, allowing for the azimutal localization of sound by barn owls.
+According to Gerstner in 1996, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower” \cite{gerstner_1996}.
+
+Various biological experiments have thus revealed the existence of a biological axonal delay precisely adjusted according to variations of parameters in the brainstem.
+
+% The importance of myelination
+
+Over and over, myelin has been identified as one of the parameters mentioned above.
+Indeed, this multilaminar coating formed by the glial cells in the Vertebrates' nervous system facilitates both the neural circuit function and the behavioural performance \cite{fields_2015}.
+Experiments on mammals show that myelination is directly related to learning and memory consolidation,
+%especially motor training and in enriched environments, and 
+both at an early age and in older animals, due to its involvement in coupling the activity of distant neuron populations. 
+%It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
+% Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.
+
+%\note{\textbf{TODO amelie} section sur l'apprentissage / la plasticité des délais dans la biologie}
 
@@
 \subsection{Perspectives}
+
+todo: implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
+it may be difficult but solutions exist  ~\citep{foss_multistability_2000,perrinet_active_2014}
+
+hierarchies
+
+developpment
 

### 2022-11-07 manuscript.tex
Update on Overleaf.

@@
 Nevertheless, we will use this terminology in all generality.
-In this respect, it is known that neural activity may be related to a specific measurable variable, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
+In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
 Among various phenomenon, a well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes.
@@
 A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
-Importantly, validating this hypothesis would also be crucial in our understanding of neuroscientific processes.
+Importantly, validating this hypothesis would also be crucial in our understanding of neural processes.
 %
@@
 Let us start with a review of the state of the art on the role of dynamics in vision.
-Vision is the set of processes that allow us to make sense of the luminous world, and is an area intensively studied in neuroscience, in particular with respect to deciphering the neural code.
+Vision is the set of processes that allow us to make sense of the luminous world, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
 In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between external visual space and its representation on the retina.
@@
 An important consequence of the speed of processing of vision is that it implies that it is carried out using only very few spikes.
-Indeed, if we consider that a behavioral response in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system.
+Indeed, if we consider that a behavioral response occurring in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system.
 This was indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
@@
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
-For instance, the firing rate is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of firing rate.
+For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of firing rate.
 %random EPSPs.
@@
 
-\subsubsection{The emergence of neuromorphic engineering}
+\subsection{The emergence of neuromorphic engineering}
 % intro 
-the field of neuromorphic engineering...
+The field of neuromorphic engineering has emerged with the need for dedicated hardware to simulate SNNs.
 
@@
 
-The interaural time difference, allowing for the azimutal localization of sound by barn owls.
+However, it was not before The interaural time difference, allowing for the azimutal localization of sound by barn owls.
 According to Gerstner in 1996, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower” \cite{gerstner_1996}.

### 2022-11-07 manuscript.tex
Update on Overleaf.

@@
 Amélie Gruel $^{2}$\orcidB{}, % $^{2,\dagger,\ddagger}$,
+Camille Besnainou $^{1}$,
 Jean Martinet $^{2}$\orcidC{},  
-and Laurent Perrinet $^{1,}$\orcidD{},*}
+and Laurent Perrinet $^{1,}$\orcidD{}*}
 
@@
 As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data.
-It is also an emerging topic in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal mission in the age of the digital economy and of global warming.
+With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption --- a major societal issue in the era of the digital economy and global warming.
 In this review paper, we provide evidence from biology and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.

### 2022-11-07 manuscript.tex
Update on Overleaf.

@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0002-9536-010X} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorAnG}{0000-0002-9536-010X} % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0003-3916-0514}
-\newcommand{\orcidauthorC}{0000-0001-8821-5556}
-\newcommand{\orcidauthorD}{0000-0002-3107-4788}
+\newcommand{\orcidauthorAmG}{0000-0003-3916-0514}
+\newcommand{\orcidauthorJN}{0000-0001-8821-5556}
+\newcommand{\orcidauthorLP}{0000-0002-3107-4788}
 
@@
 Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
-\orcidA{}, 
-Amélie Gruel $^{2}$\orcidB{}, % $^{2,\dagger,\ddagger}$,
+\orcidAnG{}, 
+Amélie Gruel $^{2}$\orcidAmG{}, % $^{2,\dagger,\ddagger}$,
 Camille Besnainou $^{1}$,
-Jean Martinet $^{2}$\orcidC{},  
-and Laurent Perrinet $^{1,}$\orcidD{}*}
+Jean Martinet $^{2}$\orcidJN{},  
+and Laurent Perrinet $^{1,}$\orcidLP{}*}
 
@@
 %----------------------------%
-Attempts have been made to detect such synfire chains in neurobiological data.
-\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
+Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
 In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
@@
 % autres méthodes
-Spatio-temporal kernels: \citep{yu_stsc-snn_2022}
+Spatio-temporal kernels:~\citep{yu_stsc-snn_2022}
 
@@
 
-One of the first significant evidence of any neuronal delay in the information propagation within the animal brain came from Young's study of the squid giant axon \cite{young_1938}:
+One of the first significant evidence of any neuronal delay in the information propagation within the animal brain came from Young's study of the squid giant axon~\cite{young_1938}:
 
 However, it was not before The interaural time difference, allowing for the azimutal localization of sound by barn owls.
-According to Gerstner in 1996, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower” \cite{gerstner_1996}.
+According to Gerstner in 1996, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”~\cite{gerstner_1996}.
 
@@
 Over and over, myelin has been identified as one of the parameters mentioned above.
-Indeed, this multilaminar coating formed by the glial cells in the Vertebrates' nervous system facilitates both the neural circuit function and the behavioural performance \cite{fields_2015}.
+Indeed, this multilaminar coating formed by the glial cells in the Vertebrates' nervous system facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
 Experiments on mammals show that myelination is directly related to learning and memory consolidation,
@@
 As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}.
-Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
+Citing a journal paper~\cite{ref-journal}.
+Now citing a book reference~\cite{ref-book1,ref-book2} or other reference types~\cite{ref-unpublish,ref-communication,ref-proceeding}.
+Please use the command~\citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
+% To cite two works by the same author:~\citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a},~\citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages:~\citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328;~\citeyear{ref-journal-3b}, p.475).
 This produces: Wong (1999, p.

### 2022-11-07 manuscript.tex
Update on Overleaf.

@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean Martinet and Laurent Perrinet}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
+\AuthorCitation{Grimaldi, A.; Gruel, A.; Besnainou, C.; Martinet, J.; and Perrinet, L.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 % Contact information of the corresponding author
-\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
+\corres{Correspondence: laurent.perrinet@univ-amu.fr}
 
 % Current address and/or shared authorship
-\firstnote{Current address: Affiliation 3.} 
-\secondnote{These authors contributed equally to this work.}
+% \firstnote{Current address: Affiliation 3.} 
+% \secondnote{These authors contributed equally to this work.}
 % The commands \thirdnote{} till \eighthnote{} are available for further notes
@@
 an explicit representation within the neural activity.
-Nevertheless, we will use this terminology in all generality.
+Nevertheless, we will use this terminology in all generality to denote the existence of a structure in neural activity.
 In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-Among various phenomenon, a well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes.
-Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and eventually further propagated along its "output" axon to reach efferent neurons.
+Among various phenomenon, a well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}.
 In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
 First, these have the particular property of being essentially binary in their amplitude, i.e.
-they are prototypical, "all or nothing".
-The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
+they are prototypical, "\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}".
+Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and eventually further propagated along its "output" axon to reach efferent neurons The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
 Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
@@
 
-However, it was not before The interaural time difference, allowing for the azimutal localization of sound by barn owls.
-According to Gerstner in 1996, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”~\cite{gerstner_1996}.
+However, it was not until the study of the interaural time difference (ITD) by Gerstner in 1996~\citep{gerstner_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
+This ITD is a biological mechanism which allows for the azimutal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
+According to Gerstner, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.
+\citet{pena_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
+A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
+the temporal disparity between left and right cochlear loci in the owl's brain.
+However it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variations were perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
+The authors support a second hypothesis, that of variating axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
+This second theory
 

### 2022-11-07 manuscript.tex
neuromorphic intro

@@
 Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
-\orcidAnG{}, 
-Amélie Gruel $^{2}$\orcidAmG{}, % $^{2,\dagger,\ddagger}$,
+\orcidauthorAnG{}, 
+Amélie Gruel $^{2}$\orcidauthorAmG{}, % $^{2,\dagger,\ddagger}$,
 Camille Besnainou $^{1}$,
-Jean Martinet $^{2}$\orcidJN{},  
-and Laurent Perrinet $^{1,}$\orcidLP{}*}
+Jean Martinet $^{2}$\orcidauthorJN{},  
+and Laurent Perrinet $^{1,}$\orcidauthorLP{}*}
 
@@
 They ``used random waveforms to probe the formation of new memories for arbitrary complex sounds.
-A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \s$ long''.
+A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \si{\second}$ long''.
 The task is to detect the repetition of the same (frozen) noise within a trial.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Artificial intelligence has made huge advances in the past decades, now deep learning algorithms outperform humans at complex tasks like natural image recognition or abstract strategy board games.
+Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems (ref de revue?), and to develop tools improving the capacities of current artificial intelligence.
+Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make computations.
+This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
+These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation for the recording of a visual scene (ref).
+Another important distinction with classical Von Neumann architectures is the localized memory of this new type of chips.
+It is materialized by the physical connections between the processing units that can store information, an example directly inspired by synaptic plasticity of such connection is the memristor (ref).
+Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
+Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges (ref).
+Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware is described as event streams by an Address-Event Representation (AER).
+Here we extend the review of the use of precise spike timing to this emerging field.
+
+\subsection{On the importance of spatio-temporal information in event-based recordings}
+
+With their AER and their submillisecond temporal precision, DVS bring a new approach to the storage and use of visual information. This neuromorphic sensor is mostly used in robotics
+
 We have seen that precise spiking motifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobiological data.

### 2022-11-07 manuscript.tex
Update on Overleaf.

@@
 However it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variations were perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
-The authors support a second hypothesis, that of variating axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
-This second theory
-
-Various biological experiments have thus revealed the existence of a biological axonal delay precisely adjusted according to variations of parameters in the brainstem.
+The authors support a second hypothesis, that of different axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
+\citet{seidl_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that "variations of parameters such as axon diameter, internode distance, and others in the mammalian brainstem might be responsible for precise adjustments of physiological delays, thereby creating the framework and adjustments of the ITD detection circuit".
+
+The experiments conducted by \citet{seidl_2010} also confirmed ... axon diameter (def \cite{gasser_1939}) + conduction velocity depending on internode distance \cite{brills_1977}
 
@@
 Indeed, this multilaminar coating formed by the glial cells in the Vertebrates' nervous system facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
-Experiments on mammals show that myelination is directly related to learning and memory consolidation,
-%especially motor training and in enriched environments, and 
-both at an early age and in older animals, due to its involvement in coupling the activity of distant neuron populations. 
-%It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
-% Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.
-
-%\note{\textbf{TODO amelie} section sur l'apprentissage / la plasticité des délais dans la biologie}
-
-% (mainly based on JOBIM poster)
+Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments, and both at an early age and in older animals, due to its involvement in coupling the activity of distant neuron populations.
+It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.
 

### 2022-11-07 manuscript.tex
review polychrony on events

@@
 %
-In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire neurons in different groups of similar size.
+In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
 Each neuron of one group is connected by an excitatory synapse to the next.
@@
 Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems (ref de revue?), and to develop tools improving the capacities of current artificial intelligence.
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\cite{roy2018toward}).
 Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make computations.
@@
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges (ref).
+Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges (zenke2021visualizing).
 Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware is described as event streams by an Address-Event Representation (AER).
-Here we extend the review of the use of precise spike timing to this emerging field.
+We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.
+
+%On one hand, neuromorphic engineering is inspired by biology to develop efficient algorithms and technology.
+For instance, DVS cameras are inspired by the retina, memristors are used to implement neuroplasticity and neuromorphic chips are designed to emulate neural elements in order to perform sparse and asynchronous computations as observed in the biological brain.
+Therefore, the neuromorphic field is interested in understanding neural circuitry to be able to reproduce such mechanisms.
+On the other hand, computational neuroscience uses mathematical models to describe and explain a wide range of cognitive processes.
+It focuses on biological plausibility to validate its models and in comparison with experimental data.
+As a result, both fields have a common interest in neural information processing and how these may implement the computations that happen in the brain (for a review, see~\cite{zenke2021visualizing}).
 
@@
 
-With their AER and their submillisecond temporal precision, DVS bring a new approach to the storage and use of visual information. This neuromorphic sensor is mostly used in robotics
+%A time surface represents therefore a combination of an object and its motion, but both elements are not explicitly dissociated.
+This entangled information may need to be separated in order to reach optimum object categorization.
+
+With the AER specification and their submillisecond temporal precision, DVS bring a new approach to the storage and use of visual information.
+From their generative model, these sparse events are markers of the dynamics of visual scene captured by the DVS.
+The dynamics of the event streams have to be used to make sense of the recorded information and new algorithms are developed to solve classical computer vision tasks.
+Different methods represent this spatio-temporal information in order to use it efficiently.
+In~\cite{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+An exponential decay is applied on this delay to obtain the analog values of the time surfaces, it gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
+Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs on the input signals.
+Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022}.
+Three dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022}.
+With their additional temporal dimension, they can capture multiple events on the same pixel address as long as they belong to the local temporal window.
+This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
+Other methods make direct use of the precise timing of events captured by the DVS to solve complicated computer vision tasks.
+
+such as estimating optical flow and time-to-contact~\cite{benosman2013event, tschechne2014bio, bardow2016simultaneous}, inferring 3D reconstruction~\cite{hidalgo2020learning, osswald2017spiking, zhu2018realtime} or solving the simultaneous localization and mapping problem~\cite{gallego2017event, kim2016real} %à voir si je reformule cette phrase empruntée du pami
 
@@
 % Grimaldi CBMI + PAMI
-~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}
+~\citep{lagorce_hots_2017, grimaldi_homeostatic_2021,grimaldi_robust_2022}
 

### 2022-11-07 manuscript.tex
formattage

@@
 
-We have seen that precise spiking motifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobiological data.
+%We have seen that precise spiking motifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobiological data.
 Interestingly some of these computational neuroscience models aim at solving the spike sorting method using STDP rules~\citep{bernert_fully_2017,bernert_attention-based_2018}
-A third class of models coming from the field of neuromorphic engineering, a priori different from that used in neuroscience, are of interest in this review as they often take advantage of minute temporal information.
-
-\subsection{The emergence of neuromorphic engineering}
+%A third class of models coming from the field of neuromorphic engineering, a priori different from that used in neuroscience, are of interest in this review as they often take advantage of minute temporal information.
+
+%\subsection{The emergence of neuromorphic engineering}
 % intro 
-The field of neuromorphic engineering has emerged with the need for dedicated hardware to simulate SNNs.
-
-uses a FPGA~\citep{dahlem_dynamics_2009}
-
-V.
+%The field of neuromorphic engineering has emerged with the need for dedicated hardware to simulate SNNs.
+
+%uses a FPGA~\citep{dahlem_dynamics_2009}
+
+%V.
 Thanasoulis, B.
@@
 Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
-~\citep{thanasoulis_delay-based_2021}
+%~\citep{thanasoulis_delay-based_2021}
 
 % un exemple de polychronization
- ~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
+% ~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
 This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
@@
 % d'autres applications
-Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
+%Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
 The task consists in localizing a target using only the relative spike timing of eight spatially-separated vibration sensors~\citep{haessig_event-based_2020}.
 
-\subsection{Temporal pattern detection in event-based cameras}
+%\subsection{Temporal pattern detection in event-based cameras}
 % le cas spécifique des event-based cameras
-Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}.
+%Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}.
 For example, event-based cameras provide a stream of binary asynchronous events signaling detectable changes in luminance, and information is represented by these spike-based temporal motifs, hence their name ``silicon retinas'' (see Figure \ref{fig:silicon_retina}).
@@
 
-\subsection{Methods using precise timings}
+\subsection{Precise timings on neuromorphic hardware}
 
@@
 
- ~\citep{lagorce_hots_2017}
-
- ~\citep{benosman_asynchronous_2012}
-
- ~\citep{benosman_event-based_2014}
-
-  ~\citep{dardelet_event-by-event_2021}
+ %~\citep{lagorce_hots_2017}
+
+ %~\citep{benosman_asynchronous_2012}
+
+ %~\citep{benosman_event-based_2014}
+
+  %~\citep{dardelet_event-by-event_2021}
 
 % Grimaldi CBMI + PAMI
-~\citep{lagorce_hots_2017, grimaldi_homeostatic_2021,grimaldi_robust_2022}
+%~\citep{lagorce_hots_2017, grimaldi_homeostatic_2021,grimaldi_robust_2022}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 \documentclass[brainsci, %arts,
-               review,submit,pdftex,moreauthors%,draft
+               review,submit,pdftex,moreauthors,%
+               draft
                ]{Definitions/mdpi} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
- Neural activity is directly influenced by our immediate environment and by internal states, and is shaped in order to generate actions.
+% CONTEXT
+Neural activity is directly influenced by our immediate environment and by internal states, and is structured in order to generate motor actions.
 The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection.
-A central question of modern neuroscience is to better understand the essence of neural activity, as one can observe it in neurobiological experiments.
+A central question of modern neuroscience is to better understand the essence of neural activity, as the recordings that can be observed in neurobiological experiments.
 One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
@@
 In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-Among various phenomenon, a well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}.
+A central and well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}.
 In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
@@
 they are prototypical, "\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}".
-Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and eventually further propagated along its "output" axon to reach efferent neurons The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
-Neurophysiologists typically use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
-However, computational neuroscience models have suggested that precise timing may also play a crucial role, and that neurons may be as well integrators than synchrony detectors~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
+Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and possibly further propagated along its "output" axon to reach efferent neurons The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
+Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance as computedf in small time windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
+This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
+
+% GAP
+However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role, and that neurons may be as well integrators than synchrony detectors~\citep{abeles_role_1982}.
 We will investigate this very hypothesis in this review.
-
-Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma.
-In comparison to a classical analogical vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing and yet, most neural models do not take advantage of this fine-scaled temporal dimension.
+Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
+In comparison to a classical analogical vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
+For instance, it expands the capabilities of analog representations (which rely on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
 This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
@@
 All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and also suggest the existence of repeated spatio-temporal spiking motifs in the spike train reaching neurons.
-A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
 Importantly, validating this hypothesis would also be crucial in our understanding of neural processes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Precise timing in the neural code enables efficient vision}
+\subsection{Dynamics of vision and consequences on the neural code}%Precise timing in the neural code enables efficient vision}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Let us start with a review of the state of the art on the role of dynamics in vision.
-Vision is the set of processes that allow us to make sense of the luminous world, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
-In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between external visual space and its representation on the retina.
-The origins of this  neuroscientific question can be found in the first experiments of Pierre Flourens showing the relationship between sensations, perception and activity in the cerebral cortex using lesions in animals~\citep{pearce_marie-jean-pierre_2009}.
+Broadly speaking, vision is the set of processes that allow us to make sense of the luminous world, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
+In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between the topology of external visual space and its representation on the retina, called retinotopy.
+The origins of this neuroscientific question can be found in the first experiments from Pierre Flourens showing the relationship between visual sensations and activity in the cerebral cortex using lesions in animals~\citep{flourens_recherches_1842,pearce_marie-jean-pierre_2009}.
+This was also observed by Lord Adrian when recording the activity the frog's visual system~\citep{adrian_impulses_1926}.
 In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as orientation or motion.
-There is remarkably a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
+For a large proportion of neurons, there is remarkably a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
 However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitute essential pieces to this puzzle~\citep{carandini_normalization_2012}.
@@
 The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
-These results have been extended to several species including primates.
+These results have been extended to several species, including primates.
 Different experimental protocols have shown for example that the response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
@@
 Although surprising, this speed of the visual cortex in primates is compatible with the latencies that are recorded at the neuro-physiological level.
-The rapid propagation of the visual information in the thalamus then in the primary visual cortex (see Figure~\ref{fig:thorpe}) takes about $55~\ms$ in the macaque~\citep{nowak_timing_1997}.
-This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
-%
+The rapid propagation of the visual information in the thalamus, then in the primary visual cortex (see Figure~\ref{fig:thorpe}) takes about $45~\ms$ in the macaque~\citep{schmolesky_signal_1998} and about $60~\ms$ in humans~\citep{vanni_coinciding_2001}.
+This functioning of visual processing as a forward pass is most prominent in fast processing, but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
 An important consequence of the speed of processing of vision is that it implies that it is carried out using only very few spikes.
-Indeed, if we consider that a behavioral response occurring in only $120~\ms$ consists of about ten processing stages following the ``forward'' pathways of the visual system.
-This was indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
-This imposes that the processing in a single area is performed with a reduced number of spikes, as it has been demonstrated that one spike requires a significant amount of time ($X~\ms$) to traverse the synapse of a neuron.
-Altogether, these results show that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
-
+Indeed, we may consider that a behavioral response occurring in only $120~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
+Such models were indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
+At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
+As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
 %----------------------------%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{How timing may encode vectors of real values}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Sensory data is most often represented by continuous values, such as the luminance in the different pixels of a digital image.
+\subsection{How precise spike timing may encode vectors of real values}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% reliability of spike timings
+Sensory data is most often represented by continuous values, such as the energy produced by a flow of photons which hits the different photoreceptors of the retina.
 How may such information be encoded in neuronal activity?
@@
 For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of firing rate.
-%random EPSPs.
-% Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
 Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned, but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
-However, if that same neuron is now driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
+However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see\citep{ermentrout_reliability_2008}).
 This is consistent with the differential role of different stimulus frequencies (for instance the gamma range around $80~\si{\Hz}$) on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
@@
 
-%
+% retina
 At the level of the dynamic processing of visual information, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
-In particular, these results show that the response of the ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
+In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
 These results have been extended to natural images and show a qualitatively similar behavior.
@@
 In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
-Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}.
+First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered, for instance for sound localization~\citep{chase_first-spike_2007}.
+Note also that timing is not entirely sensory or internal, but can be used as a general coding principle.
+In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Timing, at least in animals, may thus be fundamentally embodied and situated.''  Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}.
 These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}.
-First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered~\citep{chase_first-spike_2007}.
-One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron by using the rank of neurons' activation~\citep{delorme_spikenet_1999,bonilla_analyzing_2022}.
 Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
@@
 The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
+This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
 However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
-This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
-This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas which represents the visual information as events, similarly to the representations described above and instead of using a basic frame-based representation~\citep{rasetto_challenges_2022}.
-This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
-Note also that timing is not entirely sensory or internal, but can be used as a general coding principle.
-In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
-Timing, at least in animals, may thus be fundamentally embodied and situated.''
+Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas which represents the visual information as events, similarly to the representations described above and instead of using a basic frame-based representation~\citep{lichtsteiner_128-128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}).
+% This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
+To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.
 %
@@
 Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from M.
-Berry's lab~\citep{berry_spike_2022} and allows performing and validate these hypothesis.
+Berry's lab~\citep{berry_spike_2022} and allows performing and validate these hypotheses.
 However, a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
@@
 \section{Precise spiking motifs in neuromorphic engineering}
+\label{sec:neuromorphic}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The authors support a second hypothesis, that of different axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
-\citet{seidl_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that "variations of parameters such as axon diameter, internode distance, and others in the mammalian brainstem might be responsible for precise adjustments of physiological delays, thereby creating the framework and adjustments of the ITD detection circuit".
-
-The experiments conducted by \citet{seidl_2010} also confirmed ... axon diameter (def \cite{gasser_1939}) + conduction velocity depending on internode distance \cite{brills_1977}
+\citet{seidl_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
+These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously stated \cite{??}.
 
@@
 
-Over and over, myelin has been identified as one of the parameters mentioned above.
-Indeed, this multilaminar coating formed by the glial cells in the Vertebrates' nervous system facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
+\cite{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
+the amount of myelin wrapped around the axon.
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, wrap \cite{???}
+
+Over and over, myelin thus has been identified as a main actor in the delay regulation in many biological behaviors.
+
+facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
 Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments, and both at an early age and in older animals, due to its involvement in coupling the activity of distant neuron populations.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
-The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models.
+The book gradually leads the reader from the macroscopic cortical anatomy and standard electro-physiological properties of single neurons to neural network models.
 While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{abeles_role_1982,paugam-moisy_computing_2012}.

### 2022-11-08 manuscript.tex
draft
Update on Overleaf.

@@
 \documentclass[brainsci, %arts,
-               review,submit,pdftex,moreauthors,%
-               draft
+               review,submit,pdftex,moreauthors
+               %,  draft
                ]{Definitions/mdpi} 
@@
 One possible function of this synchronization may serve the binding of information distributed in the brain~\citep{singer_visual_1995, roelfsema_visuomotor_1997}.
-%
-In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
-Each neuron of one group is connected by an excitatory synapse to the next.
-When a pulse is elicited in the first group, this may generate a spike in the next group.
-Depending on the concentration of synaptic weight values, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse).
-Recursively applying this to a sequence of groups generates either a synfire propagation or not.
-A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
-A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model÷\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
-This was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
-
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=.4\textwidth]{figures/Diesmann_et_al_1999.png}
-\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
-The model consists of 10 groups of 100 neurons each.
-A pulse with a given jitter is generated in the first group which here generates a pulse after a certain processing delay in the second group with a lower jitter.
-This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
-\end{figure}
-%----------------------------%
-Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
-In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013} (see Figure~\ref{fig:pynn}).
-%
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
-\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
-The background noise  is implemented as random Gaussian current.
-The state space gives the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
-\end{figure}
-%----------------------------%
-%
+
 Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}.
@@
 
+In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
+Each neuron of one group is connected by an excitatory synapse to the next.
+When a pulse is elicited in the first group, this may generate a spike in the next group.
+Depending on the concentration of synaptic weight values, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse).
+Recursively applying this to a sequence of groups generates either a synfire propagation or not.
+A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
+A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model÷\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
+This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
+
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=.4\textwidth]{figures/Diesmann_et_al_1999.png}
+\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
+The model consists of 10 groups of 100 neurons each.
+A pulse with a given jitter is generated in the first group, which here generates a pulse after a certain processing delay in the second group with a lower jitter.
+This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
+\end{figure}
+%----------------------------%
+Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
+In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013}.
+% (see Figure~\ref{fig:pynn}).
+%
+%----------------------------%
+%\begin{figure}
+%\centering
+%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
+%\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
+The background noise  is implemented as random Gaussian current.
+The state space gives %the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
+%\end{figure}
+%----------------------------%
+
 It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas~\citep{riehle_spike_1997,grammont_precise_1999}.
@@
 Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials.
-Accurate spike synchronization occurs in relation to external events (stimuli, movements) and is commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
-Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
-These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.
 In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
@@
 
-%: rhythms as a 
 The idea of using the synchrony of co-activation in a cell assembly is a hypothesis that was formalized by~\citet{hebb_organization_1949}. 
@@
 The synchronicity will generate rhythms at different range of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down.
-Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion , i.e.
-to the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
+Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion, i.e.
+the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
 More generally, it can be shown that the phase of alpha oscillations (about $10$ Hz) is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.

### 2022-11-08 manuscript.tex
2.1

@@
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016,torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
@@
 A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
-A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model÷\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
+A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
 This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
-
+Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
+In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013}.
+% (see Figure~\ref{fig:pynn}).
+%
+%----------------------------%
+%\begin{figure}
+%\centering
+%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
+%\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
+The background noise  is implemented as random Gaussian current.
+The state space gives %the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
+%\end{figure}
+%----------------------------%
 %----------------------------%
@@
 %----------------------------%
-Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
-In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013}.
-% (see Figure~\ref{fig:pynn}).
-%
-%----------------------------%
-%\begin{figure}
-%\centering
-%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
-%\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
-The background noise  is implemented as random Gaussian current.
-The state space gives %the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
-%\end{figure}
-%----------------------------%
-
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas~\citep{riehle_spike_1997,grammont_precise_1999}.
+
+Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas~\citep{riehle_spike_1997,grammont_precise_1999}.
 It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
@@
 
-The idea of using the synchrony of co-activation in a cell assembly is a hypothesis that was formalized by~\citet{hebb_organization_1949}. 
+The idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}. 
 This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}.
@@
 
-\begin{figure}
-\centering
-%\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
-\caption{a Spike raster for repeated presentations (%N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
+% \begin{figure}
+% \centering
+% %\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
+% \caption{a Spike raster for repeated presentations (%N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
 A single-trial LFP trace is plotted in gray, and the average LFP response is plotted in black.
@@
 Data for panels c--e modified from Davis et al.17 with permission.
-}\label{fig:davis}
-\end{figure}
+% }\label{fig:davis}
+% \end{figure}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 
-Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas~\citep{riehle_spike_1997,grammont_precise_1999}.
-It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
-Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials.
+Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Indeed, simultaneously recorded activities of neurons in the primary motor cortex of monkeys exhibited context-dependent, rapid changes in the patterns of coincident action potentials~\citep{riehle_spike_1997,grammont_precise_1999}.
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas yet it is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
 In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
@@
 
-The idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}. 
+More generally, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}. 
 This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 \subsection{A further hypothesis: traveling waves}
+%
 To further investigate the role of precise timing, let us also focus on the role of differential timings in an assembly of neurons.
@@
 Using intracellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is, a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
-In particular, it was shown that the network of horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within the area.
+In particular, it was shown that the network of the so-called horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within that area.
 For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions and may underlay anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
@@
 
-More recently, advanced recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
+More recently, novel multi-unit recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
 Authors review these findings, consider the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 More recently, novel multi-unit recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
-Authors review these findings, consider the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function.
+Authors review these findings by considering the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function.
 In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
@@
 Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings.
-Here, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
+In that paper, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
 They find that spontaneous traveling waves are a general property of these networks.
@@
 Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
-Further, by modulating local network state, they can shape responses to incoming inputs as observed \textit{in vivo}.
+Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \textit{in vivo}.
 Such waves also occur in motor areas and~\citet{linden_movement_2021} have presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 
-However, it was not until the study of the interaural time difference (ITD) by Gerstner in 1996~\citep{gerstner_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
+However, it was not until the study of the interaural time difference (ITD) by Gerstner in 1996~\citep{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
 This ITD is a biological mechanism which allows for the azimutal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
@@
 The authors support a second hypothesis, that of different axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
-\citet{seidl_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
+\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
 These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously stated \cite{??}.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 \citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously stated \cite{??}.
+These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously thought \cite{??}.
 
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, wrap \cite{???}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce  \cite{???}
 

### 2022-11-08 manuscript.tex
2.3

--- 2022-11-08 manuscript.tex
@@
 
-Moreover, there is a consensus for rate-based coding models in computational and biological neuroscience.
-Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015}.
+From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015}.
 Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
 This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996} or touch~\citep{johansson_first_2004}.
-In particular, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
-
-Stereotyped sequences of neural activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
+
+In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
 These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
@@
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
+
+Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
+Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
 The rest of this review will be devoted to present evidence for the use of precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce  \cite{???}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce  \citep{reynolds_1928}
 
@@
 facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
-Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments, and both at an early age and in older animals, due to its involvement in coupling the activity of distant neuron populations.
+Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments.
+This biological phenomenom both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
 It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce  \citep{reynolds_1928}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin proteic sheets wrapped concentircallu about the  \citep{schmidt_1939}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce  \citep{reynolds_1928}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin proteic sheets wrapped concentircall the axon \citep{schmidt_1939}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin proteic sheets wrapped concentircall the axon \citep{schmidt_1939}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets wrapped concentrically around the axon \citep{schmidt_1939}
 

### 2022-11-08 manuscript.tex

--- 2022-11-08 manuscript.tex
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin proteic sheets wrapped concentircallu about the  \citep{schmidt_1939}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets wrapped concentrically around the axon \citep{schmidt_1939}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin proteic sheets wrapped concentircallu about the  \citep{schmidt_1939}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets interspwrapped concentrically around the axon \citep{schmidt_1939}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets wrapped concentrically around the axon \citep{schmidt_1939}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets intewrapped concentrically around the axon \citep{schmidt_1939}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 %
+% Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~ https://dx.doi.org/10.1073/pnas.1902653116
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The rest of this review will be devoted to present evidence for the use of precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
-
-% Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~https://dx.doi.org/10.1073/pnas.1902653116
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the following section, we will review theoretical models which directly tackle the problem of using precise spiking motifs. 
+In the following section, we will review theoretical models which directly take advantage of using precise spiking motifs. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-This approach is quite successful, and SNNs reach in some case a similar performance as Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
+This approach is quite successful, and SNNs reach in some case a similar performance as Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
 However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
@@
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
 These models will be compared in light of neuroscientific and computational perspectives.
@@
 In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
-%Finally we will discuss future avenues for effective PG detection and learning in event streams.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% First, we will review models for the detection of polychronous groups.
-% Our goal in this review is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines. 
-% In particular, we will focus on the hypothesis that there exists at the scale of an assembly of neurons a representation based on a set of motifs defined by different relative spike times.
-%Here, we will review current literature on the detection of such motifs in generic raster plots.
-It is work in progress, where anybody interested can \emph{openly} join.
-%We will first review some biological and theoretical evidence in neural information processing in the particular case of vision. 
-%The next section will focus on precise spike time motifs. 
-% We will then present some models for the detection of such motifs in arbitrary raster plots, biological or artificial (notably from event-based cameras).
-In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
-Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
-%
 % polychronization
-%  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html} Basic idea: different temporal activation of neurons can activate different populations based on conduction delays.
-This allows for a much higher dimensional representation space, and a lot more memory capacity.
-Firing is not synchronous, but it is time-locked = polychrony.
-STDP automatically associates the neurons in the network into polychronous groups.
-There can be far more polychronous groups than neurons.
-He lets the network just settle for 24 hours - and these groups emerge spontaneously from the random connectivity initial conditions.
-The dendritic tree can be used to select the polychronous inputs more precisely.
-Can act as a full feed-forward neural network and make non-linear classifications.
-Two dendritic trees on pyramidal cells - one for the feed-forward input, and one for feed-back/recurrent.
-These can interact such that the feed-back connections are predicting the feed-forward inputs.
-As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on a firing-rate based encoding of information.
+As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on an encoding of information based on a continuously varying firing-rate.
 A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}.
-This model is based on a random recurrent model of spiking neurons including such synaptic delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
-It is worth mentioning that the set of delays is set randomly at initialization in a range of biologically realistic delays (from $0$ to $20~\ms$) and is then frozen for the rest of the simulation, only the weights are changed using the STDP rule.
+This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
 It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
+It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation.
 Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
-The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike.
+The neurons composing a group discharge at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
+This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:PG}).
 According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, constitute a polychronous group.
@@
 This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
-
 %----------------------------%
@@
 \centering
-\note{insert fig 2 of izhikevich 2006 - or an open-source reproduction ?}
+\note{insert fig 2 of izhikevich 2006 - or an open-source reproduction like that from Camille ? }
 %includegraphics{figures/event_driven_computations.png}
@@
 
-% polychronization
+% follow-ups
 Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper.
@@
 Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
+As such, these models do not explicitly take advantage of the representation capacity offered by polychronous groups.
 %
@@
 This ITD is a biological mechanism which allows for the azimutal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-According to Gerstner, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.
+According to \citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.
 \citet{pena_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
@@
 \citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously thought \cite{??}.
+These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously thought.
 
@@
 
-\cite{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
+\citet{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets wrapped concentrically around the axon \citep{schmidt_1939}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon \citep{schmidt_1939}
 
@@
 It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
+
+%
+
 Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-Based on the fact that previous methods for studying polychronous neural groups (PNG) activation response to stimuli have been limited by the template-based methods used to identify PNG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
+Based on the fact that previous methods for studying polychronous neural groups (PNG) PG activation response to stimuli have been limited by the template-based methods used to identify PNG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
 They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.
@@
 
+% role of myelination and neuronal delay
+
 facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
@@
 
-%
+% How is the neuronal activity detected ?
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-Based on the fact that previous methods for studying polychronous neural groups (PNG) PG activation response to stimuli have been limited by the template-based methods used to identify PNG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
+Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PNG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
 They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon \citep{schmidt_1939}
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon~\citep{schmidt_1939}
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PNG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
+Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
 They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
-As such, these models do not explicitly take advantage of the representation capacity offered by polychronous groups.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsection{Extensions of Izhikevitch's polychronization model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
+As such, these models do not explicitly take full advantage of the representation capacity offered by polychronous groups.
+
 An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
@@
 It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
-%
 Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation.
-They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.
-Stimulation of the trained network produces the activation of a PNG, i.e.
+Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation.
+They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system.
+Stimulation of the trained network produces the activation of a PG, i.e.
 the propagation of firing activity through multiple layers due to convergent patterns of firing.
-
-% STDP based
-%Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
-It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
-This raises important questions about the fundamental limits of such memory traces and the properties required of dynamical systems to achieve these limits.
-We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
-We introduce the Fisher Memory Curve (FMC) as a measure of the signal-to-noise ratio (SNR) embedded in the dynamical state relative to the input SNR.
-The integrated FMC indicates the total memory capacity.
-We apply this theory to linear neural networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
-A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
-The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed $\sqrt N$.
-This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
-We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.'' We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
-Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 % Training delays ~\citep{state_training_2019}
@@
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-To address  how transmission delays and STDP can jointly determine the emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
-%
-Modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
+In particular, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
+Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
 They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
-However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
+For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition or~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
+
+However, the performance of SNNs is still lagging compared to that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
 Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas which represents the visual information as events, similarly to the representations described above and instead of using a basic frame-based representation~\citep{lichtsteiner_128-128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}).
@@
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
@@
 A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
-This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
+This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling so-called push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
 Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
@@
 It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation.
+Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.
 Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
@@
 the propagation of firing activity through multiple layers due to convergent patterns of firing.
+While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
 %
@@
 % Training delays ~\citep{state_training_2019}
-A whole range of models of latency motifs detection are using STDP-based learning rules.
-For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-In particular, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
+First, the original model use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
+For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
 Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
@@
 More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
-Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
-This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
-
-Specifically, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
-The authors provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns.
-Furthermore, the experimental results of applying an STDP-based spiking neural network was applied on Random Dot Kinematogram.
+Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
+
+Furthermore, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
+The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns.
+Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task.
 Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
@@
 Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
-% Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
 In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Most of these theoretical models are trying to reproduce neurobiological observations, while others propose applications to machine learning methods such as image processing.
+A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios.
+Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility.
 For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
@@
 For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
-Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters thus showing the potential of such representation.
-
-%There are more applications to image processing using sparse spiking representations.
-Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
-design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
-A predictive model defined in the framework of the free-energy principle was able to account for temporal delays in the nervous system, both at the sensory and motor levels~\citep{perrinet_active_2014}.
-%
-%and behavioral~\citep{khoei_flash-lag_2017} levels. Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
-Such a model was implemented at the network level by showing that motion-based prediction explains the role of tracking in motion extrapolation interpolation~\citep{khoei_motion-based_2013}.
-Note that this motion-based prediction is implemented in the weights of the neural network by defining connections between neurons with specific delays.
+Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.
+
+There are more applications to image processing using spiking neural networks based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections.
+Using this core computational unit defined, and extending to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation.
+This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks.
 Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
-This was also implemented in a spiking neural network and reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
-This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-In the following section, we will review some more applications of using precise spiking motifs in neuromorphic engineering, but will first study current evidence in neurobiology, that is, from actual recordings in biological tissues.
+This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
+This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash-lag illusion~\citep{khoei_flash-lag_2017}.
+However, these latter models used parametric rules for defining the wrights.
+Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore results from neurobiology to get better insights on that aspect.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Precise spiking motifs in neurobiology}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In the following section, we will first review current evidence of taking advantage of spiking motifs in neurobiology, that is, from actual recordings in biological tissues.
+% some more applications of using precise spiking motifs in neuromorphic engineering, but will first study
 %

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the following section, we will first review current evidence of taking advantage of spiking motifs in neurobiology, that is, from actual recordings in biological tissues.
-% some more applications of using precise spiking motifs in neuromorphic engineering, but will first study
-%
+In this section, we will review current evidence of taking advantage of spiking motifs in neurobiology, that is, from actual recordings in biological tissues.
 In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
-For instance, the output of linear non-linear (L-NL) models is assumed to be distributed according to a Poisson law.
-As such, a simple decoding strategy is to assume it is to be inferred for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
+For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process.
+As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.
 This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
 An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
-In Bellec~\citet{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
-The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
-The model uses samples and measure method sto include latent / hidden neurons.
+In~\citet{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
 In particular, it comes with  \href{code}{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}.
-It uses the publicly available \href{V1-dataset}{http://crcns.org/data-sets/vc/pvc-11}, which allows to supervise the model with the input being the movie and the output the spikes recorded.
+It uses the publicly available \href{V1-dataset}{http://crcns.org/data-sets/vc/pvc-11}, which allows supervising the model with the input being the movie and the output the spikes recorded.
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
-Using the right metric, as implemented in that paper by the models themselves, is essential in better understanding neural data.
-As such, all these models are dependent on  a core definition of spike measures, and we will review here how precise spiking motifs are taken into account in various neurobiological results.
+Using the right metric, as implemented in that paper by the chosen neural models, is essential in better understanding neural data.
+As such, all these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
 %
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-There are different solution to provide with a distance between two given spike trains.
-A known measure is the Victor-Purpura distance, which showed that their findings are inconsistent with Poisson models of spike trains~\citep{victor_nature_1996}.
+There are different solutions to provide with a distance between two given spike trains.
+A known measure is the Victor-Purpura distance, which showed that their findings are overcoming inconsistencies experienced with Poisson models of spike trains~\citep{victor_nature_1996}.
 In a later study,~\citet{van_rossum_novel_2001} have defined a distance that uses a time constant as a parameter.
@@
 Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citet{kreuz_measuring_2007}.
-Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
+Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
 These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.

### 2022-11-08 manuscript.tex
4

@@
 The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
-In particular, it comes with  \href{code}{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}.
-It uses the publicly available \href{V1-dataset}{http://crcns.org/data-sets/vc/pvc-11}, which allows supervising the model with the input being the movie and the output the spikes recorded.
+In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code}.
+It uses the publicly available \href{http://crcns.org/data-sets/vc/pvc-11}{V1-dataset} which allows supervising the model with the input being the movie and the output the spikes recorded.
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.

### 2022-11-08 manuscript.tex
4

@@
 An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
-In~\citet{bellec_fitting_2021}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+In that paper, authors fit summary statistics of neural data with a differentiable spiking network model.
 The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
-In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code}.
-It uses the publicly available \href{http://crcns.org/data-sets/vc/pvc-11}{V1-dataset} which allows supervising the model with the input being the movie and the output the spikes recorded.
+In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the publicly available \href{http://crcns.org/data-sets/vc/pvc-11}{V1-dataset} which allows supervising the model with the input being the movie and the output the spikes recorded.
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 % \item
-Using a t-SNE projection with known labels validated that this clustering method can retrieve all patterns from the data.
+Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data.
 % \item

### 2022-11-08 manuscript.tex
Update on Overleaf.

--- 2022-11-08 manuscript.tex
@@
 
-From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008,villette_internally_2015}.
+From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008}.
 Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
@@
 Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019} reproduced below).
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}).
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
@@
 The rest of this review will be devoted to present evidence for the use of precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
+%
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
+\caption{An example of a polychronous group seen in cortical slices.
+In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
+Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
+The black line represents the speed of the mouse.}\label{fig:haimerl}
+\end{figure}
+%----------------------------%
 %
@@
 By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
-This was recently extended in~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
 The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
 
-% TODO: summarize ``Understanding Auditory Spectro-Temporal Receptive Fields and Their Changes with Input Statistics by Efficient Coding Principles'' - Spectro-temporal receptive fields (STRFs) have been widely used as linear approximations to the signal transform from sound spectrograms to neural responses along the auditory pathway.
-Their dependence on statistical attributes of the stimuli, such as sound intensity, is usually explained by nonlinear mechanisms and models.
-Here, we apply an efficient coding principle which has been successfully used to understand receptive fields in early stages of visual processing, in order to provide a computational understanding of the STRFs.
-According to this principle, STRFs result from an optimal tradeoff between maximizing the sensory information the brain receives, and minimizing the cost of the neural activities required to represent and transmit this information.
-Both terms depend on the statistical properties of the sensory inputs and the noise that corrupts them.
-The STRFs should therefore depend on the input power spectrum and the signal-to-noise ratio, which is assumed to increase with input intensity.
-We analytically derive the optimal STRFs when signal and noise are approximated as Gaussians.
-Under the constraint that they should be spectro-temporally local, the STRFs are predicted to adapt from being band-pass to low-pass filters as the input intensity reduces, or the input correlation becomes longer range in sound frequency or time.
-These predictions qualitatively match physiological observations.
-Our prediction as to how the STRFs should be determined by the input power spectrum could readily be tested, since this spectrum depends on the stimulus ensemble.
-The potentials and limitations of the efficient coding principle are discussed.
-~\citep{zhao_understanding_2011}
-
-%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
-\subsection{Scaling up to very large scale data}
-
-In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-
-Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
-Bridging across different spatial and temporal scales, combination of \textit{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
-The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurophysics to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
-
-One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
-Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
-The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
-In~\citet{stringer_spontaneous_2019,stringer_high-dimensional_2019}, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
-Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
-Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
-% \href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
-
-% \begin{itemize}
-%  \item
-%   https://www.janelia.org/lab/stringer-lab
-% \item
-%   described in
-% \item
-%   \href{https://github.com/MouseLand/rastermap}{rastermap}
-% \item
-%   deconvolution strategy
-% \item
-%   based on a linear model
-% \end{itemize}
-% \subsubsection{Stringer et al 2019, Nature~\citep{stringer_high-dimensional_2019}}\label{stringer-et-al-2019-nature-stringer2019nature}
-% \begin{itemize}
-%  \item
-%   ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
-Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
-% \item
-%   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
-% \item
-%   Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
-% \end{itemize}
-
-% \subsubsection{Stringer et al 2019, Science~\citep{stringer_spontaneous_2019}}\label{stringer-et-al-2019-science-stringer2019science}
-% \begin{itemize}
-% \item
-%   Stringer et al.~analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
-The variability of neural responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
-% \item
-%   see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{stringer_high-precision_2021}
-% \end{itemize}
-
-Another important algorithm is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}.
+Another important algorithm, called SPOTDisClust, is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}.
 They introduced an unsupervised method based for their detection from high-dimensional neural ensembles.
@@
 In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
-% \item
-%   Detecting these temporal patterns represents a major methodological challenge.
-% \item
-%   Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
-% \item
-%   Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
-% \item
-%   Lee AK, Wilson MA.
-A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order.
-J Neurophysiol.
-2004;92(4):2555--73.
-pmid:15212425
-% \item
+In SPOTDisClust, the learning is unsupervised.
+It uses the prior that there is only one spike per pattern.
+Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data (see Figure~\ref{fig:SPOTDisClust}).
+The limits of this method are that it is computationally complex, block-based and strictly specialized for the task at hand.
+To overcome these difficulties, a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
 %   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
-% \item
-In that method, the learning is unsupervised. It uses the prior that there is only one spike per pattern. 
-% \item
-Using a t-SNE projection with known labels validated that this clustering method can retrieve all patterns from the data.
-% \item
-%   data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
-% \end{itemize}
-The limits of this method is that it is block-based and strictly specialized for the task at hand.
-To overcome these difficulties, a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
 
@@
 \centering
-%\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
-\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}: ``Simulated example illustrating the steps in SPOTDisClust.
+\includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
+\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in SPOTDisClust.
 A) Structure of five ``ground-truth'' patterns (\ldots).
 For each pattern and each neuron, a random position was chosen for the activation pulse.
-B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
+B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
+(© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
 \end{figure}
 %----------------------------%
-
-% \begin{itemize}
-%  \item
-%   ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
-This becomes a major problem for computing an M dissimilarity matrix (for M time epochs) for thousands of neurons.
-(2) It relies exclusively on pairwise spike-timing relationships (i.e 2nd order correlations), because it does not solve the optimal transport problem for the entire spike pattern, but only for neuron pairs separately.
-Hence, it may not be sensitive to higher-order correlations in spiking
-%   patterns.
-%   Here, we develop a novel dissimilarity measure for multi-neuron spiking patterns called SpikeShip, which has linear computational complexity of O(N).
-We achieve this by (1) computing the minimum transport cost of spikes for each spike train separately, and (2) discounting a global translation term in the transport flow across neurons.''
-%   https://doi.org/10.1101/2020.06.03.131573;
-% \end{itemize}
-
-\subsection{Extending the decoding at multiple time scales}
-
+%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Scaling up to very large scale data}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
+Bridging across different spatial and temporal scales, combination of \textit{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
+The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurobiological knowledge to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
+As a consequence, there is a growing need to scale these methods to larger amounts of data.
+
+There are multiple solutions which aim at tackling this problem.
+One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
+Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
+The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
+In~\citet{stringer_spontaneous_2019,stringer_high-dimensional_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
+%   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
+
+% \subsection{Extending the decoding at multiple time scales}
 In the paper by~\citet{russo_cell_2017}, the authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.
@@
 The core measure is based on a non-stationarity corrected parametric test statistic for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
-
-% \subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
-% \begin{itemize}
-%  \item
-%   Stochastic sampling links perceptual uncertainty to neural response variability
-% \item
-%   Model accounts for independent changes in strength and variability of responses
-% \item
-%   Model predicts relationship between noise, signal, and spontaneous correlations
-% \item
-%   Stimulus statistics dependence of response statistics is explained
-% \end{itemize}
-
-To overcome the limits of models which require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
+To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
 They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
-
-% Currently, few interdisciplinary research projects take full advantage of event-based representations in computational and biological neuroscience.
-Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains (Ikegaya et al., 2004), packets (Luczak et al., 2007) or hippocampal sequences (Pastalkova et al., 2008; Villette et al., 2015)  and on the role of such precise spike timing in downstream information transfer and coding (Branco et al., 2010; Buzsáki & Tingley, 2018; Luczak et al., 2015).
-In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs (Izhikevich, 2006) which he defined as polychronous groups.
-
-%Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space (Buzsáki & Tingley, 2018).
-These sequences can be internally generated (Pastalkova et al., 2008; Villette et al., 2015) and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
-Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure 2 of (Haimerl et al., 2019)).
-As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
-\begin{figure}
-\centering
-\includegraphics{figures/haimerl2019.jpg}
-\caption{
-%  Figure 2: An example of a polychronous group seen in cortical slices.
-In this study by (Haimerl et al., 2019), an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
-Mixed distance and duration representation in CA1.
-(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
-Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse.}\label{fig:haimerl}
-\end{figure}
-
-% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
-
-In~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
+At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
 They develop a Hidden Markov Model for quantifying the transitions.
@@
 Top-down attention drives inter-areal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Cortical songs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% what we say in section 2.3: Stereotyped sequences of neural activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
-These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
-Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019} below).
-We have seen previously that neurobiologists have described Stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019} below).
+Building such models for predicting changes of timings based on context, such as using a so-called change point model for blocked-based experimental protocols~\citep{pasturel_humans_2020}, could therefore provide useful prior information to enhance the decoding from neural activity.
+
+% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
+%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Explictly decoding spiking motifs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
 Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons \textit{in vivo} and \textit{in vitro}.
@@
 These sequences can be specific of a particular layer or column, can be synchronized with network activity oscillation, and can involve several cells.
-They also demonstrated that these sequences can form super sequences, so-called cortical songs.
+They also demonstrated that these sequences can form super sequences, so-called \emph{cortical songs}.
 It consists of the assembly of several sequences which repeat in a specific order with a compressed timing.
@@
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
+Such precise structural information should therefore be included in decoding algorithm for a better understanding of neural activity.
 
@@
 
+% \subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
+% \begin{itemize}
+%  \item
+%   Stochastic sampling links perceptual uncertainty to neural response variability
+% \item
+%   Model accounts for independent changes in strength and variability of responses
+% \item
+%   Model predicts relationship between noise, signal, and spontaneous correlations
+% \item
+%   Stimulus statistics dependence of response statistics is explained
+% \end{itemize}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 % sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
-%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
-Such precise structural information should therefore be included in decoding algorithm for a better understanding of neural activity.
-
+
+%----------------------------%
 \begin{figure}
 \centering
-%\includegraphics{figures/Ikegaya2004zse0150424620001.jpeg}
+\includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
 \caption{Fig.
@@
 \end{figure}
-
-It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns.
-They ``used random waveforms to probe the formation of new memories for arbitrary complex sounds.
-A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \si{\second}$ long''.
+%----------------------------%
+
+That results suggest that polychronous groups are not necessarily grouped on a topography and can be forming apparently randomly arranged connections.
+It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some (frozen) noise patterns.
+In paerticular, they used random waveforms to probe the formation of new memories for arbitrary complex sounds.
+A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \si{\second}$ long.
 The task is to detect the repetition of the same (frozen) noise within a trial.
-" Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block." They showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case.
-Note that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``time reversal had no significant effect on the RefRN advantage'' (which is quite surprising).
-The mearning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).
-
-TODO: re-read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
-
-Such observations suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{miller_visual_2014}.
-This work shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block.
+They showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case.
+Note also that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``time reversal had no significant effect on the detection accuracy (which is quite surprising).
+The learning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).
+
+Such observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
+~\citep{miller_visual_2014} shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
 It proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
-Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions (Fig.
+Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions.
+% (Fig.
 7).
-
+As a summary, it seems imperative to include such precise structural information should therefore be included in decoding algorithm for a better understanding of neural activity.
+
+% TODO: re-read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
 % \subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
+
 Here, we develop a model for the efficient detection of such PGs based on the inversion of a probabilistic model defining the generation of the raster plot as a combination of such groups.
@@
 developpment
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``Conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work~reported.}
+
+\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+
+\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
+You might choose to exclude this statement if the study did not require ethical approval.
+Please note that the Editorial Office might ask you for further information.
+Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
+OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
+OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
+
+\informedconsent{Any research article describing a study involving humans should contain this statement.
+Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
+You might also choose to exclude this statement if the study did not involve humans.
+
+Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
+Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
+
+\dataavailability{
+%This works is made reproducible using the following tools.
+First the code reproducing all figures is available on \href{https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Readme.md}{GitHub}~\citep{Jeremie2022a}.
+
+%\noindent The paper is available a an \href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
+Find the associated \noindent \href{https://www.zotero.org/groups/4562620/polychronies}{zotero group} used to regroup relevant literature on the subject.
+
+}
+
+\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results''.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\sampleavailability{Samples of the compounds ... are available from the authors.}
+
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+\abbreviations{Abbreviations}{
+The following abbreviations are used in this manuscript:\\
+
+\noindent 
+\begin{tabular}{@{}ll}
+MDPI & Multidisciplinary Digital Publishing Institute\\
+DOAJ & Directory of open access journals\\
+TLA & Three letter acronym\\
+LD & Linear dichroism
+\end{tabular}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
+\appendix
+\section[\appendixname~\thesection]{}
+\subsection[\appendixname~\thesubsection]{}
+The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+\begin{table}[H] 
+\caption{This is a table caption.\label{tab5}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCC}
+\toprule
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+\midrule
+Entry 1		& Data			& Data\\
+Entry 2		& Data			& Data\\
+\bottomrule
+\end{tabularx}
+\end{table}
+
+\section[\appendixname~\thesection]{}
+All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{adjustwidth}{-\extralength}{0cm}
+%\printendnotes[custom] % Un-comment to print a list of endnotes
+
+%%%%%%%%% REFERENCES
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
+%\bibliographystyle{Definitions/chicago2}
+\bibliography{references}
+
+%\begin{thebibliography}{00}
+%\printbibliography[heading=none]
+%\end{thebibliography}
+
+% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author:~\citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a},~\citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages:~\citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328;~\citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+\end{adjustwidth}
+\end{document}
 
@@
 %This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
-A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work~reported.}
-
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
-You might choose to exclude this statement if the study did not require ethical approval.
-Please note that the Editorial Office might ask you for further information.
-Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
-OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
-OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
-
-\informedconsent{Any research article describing a study involving humans should contain this statement.
-Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
-You might also choose to exclude this statement if the study did not involve humans.
-
-Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
-Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
-
-\dataavailability{
-%This works is made reproducible using the following tools.
-First the code reproducing all figures is available on \href{https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Readme.md}{GitHub}~\citep{Jeremie2022a}.
-
-%\noindent The paper is available a an \href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
-Find the associated \noindent \href{https://www.zotero.org/groups/4562620/polychronies}{zotero group} used to regroup relevant literature on the subject.
-
-}
-
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
-
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results''.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-\sampleavailability{Samples of the compounds ... are available from the authors.}
-
-%% Only for journal Encyclopedia
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
-
-\abbreviations{Abbreviations}{
-The following abbreviations are used in this manuscript:\\
-
-\noindent 
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & Linear dichroism
-\end{tabular}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendixstart
-\appendix
-\section[\appendixname~\thesection]{}
-\subsection[\appendixname~\thesubsection]{}
-The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
-Mathematical proofs of results not central to the paper can be added as an appendix.
-
-\begin{table}[H] 
-\caption{This is a table caption.\label{tab5}}
-\newcolumntype{C}{>{\centering\arraybackslash}X}
-\begin{tabularx}{\textwidth}{CCC}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-Entry 1		& Data			& Data\\
-Entry 2		& Data			& Data\\
-\bottomrule
-\end{tabularx}
-\end{table}
-
-\section[\appendixname~\thesection]{}
-All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{adjustwidth}{-\extralength}{0cm}
-%\printendnotes[custom] % Un-comment to print a list of endnotes
-
-%%%%%%%%% REFERENCES
-\reftitle{References}
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
-%\bibliographystyle{Definitions/chicago2}
-\bibliography{references}
-
-%\begin{thebibliography}{00}
-%\printbibliography[heading=none]
-%\end{thebibliography}
-
-% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author:~\citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a},~\citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages:~\citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328;~\citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-\end{adjustwidth}
-\end{document}

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 However, a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
+Recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
+~\citep{miller_visual_2014} shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+It proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
+Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions.
+% (Fig.
+7).
 
 From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008}.
-Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
-This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996} or touch~\citep{johansson_first_2004}.
-
-In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
-These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
-Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}).
-As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
-
-Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
-The rest of this review will be devoted to present evidence for the use of precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
-%
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
-\caption{An example of a polychronous group seen in cortical slices.
-In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
-(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
-Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse.}\label{fig:haimerl}
-\end{figure}
-%----------------------------%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Precisely-timed spiking motifs in computational neuroscience}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the following section, we will review theoretical models which directly take advantage of using precise spiking motifs. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
-The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-This approach is quite successful, and SNNs reach in some case a similar performance as Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
-However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
-There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
-These models will be compared in light of neuroscientific and computational perspectives.
-We review in this section theoretical and computational foundations of PG detection in models.
-In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Izhikevitch's polychronization model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% polychronization
-As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}.
-This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
-It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
-It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation.
-Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
-The neurons composing a group discharge at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:PG}).
-According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, constitute a polychronous group.
-Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
-In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
-%----------------------------%
-\begin{figure}
-\centering
-\note{insert fig 2 of izhikevich 2006 - or an open-source reproduction like that from Camille ? }
-%includegraphics{figures/event_driven_computations.png}
-\caption{blah.}\label{fig:PG}
-\end{figure}
-%----------------------------%
-
-% follow-ups
-Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper.
-Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs.
-A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}.
-The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron which does not incorporate a spike timing framework.
-The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a STDP rule.
-The limits of this model is that its output is only binary and that its storage capacity are limited.
-The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
-Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
-As such, these models do not explicitly take full advantage of the representation capacity offered by polychronous groups.
-
-An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
-Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript.
-This more recent work details how this code was slightly corrected.
-It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
-Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation.
-They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system.
-Stimulation of the trained network produces the activation of a PG, i.e.
-the propagation of firing activity through multiple layers due to convergent patterns of firing.
-While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Learning synaptic delays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Training delays ~\citep{state_training_2019}
-First, the original model use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
-For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
-Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
-They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
-In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
-More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
-Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
-
-Furthermore, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
-The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns.
-Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task.
-Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
-This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
-In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron.
-In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
-In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.
-
-%In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in birdsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
-Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
-In~\citep{agus_rapid_2010}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
-
-% Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
-% Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021
-
-% Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
-% Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020
-
-% Unsupervised Learning of Persistent and Sequential Activity.
-% Pereira U, Brunel N.Front Comput Neurosci. 2020
-
-% From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
-% Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019
-
-% Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
-% Pang R, Fairhall AL.Elife. 2019 May~\citep{pang_fast_2019}
-
-% Emergence of spontaneous assembly activity in developing neural networks without afferent input.
-% Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018
-
-% Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
-% Ocker GK, Doiron B.Cereb Cortex. 2019.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Real-world applications}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios.
-Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility.
-For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
-The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern.
-The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity.
-For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
-Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.
-
-There are more applications to image processing using spiking neural networks based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections.
-Using this core computational unit defined, and extending to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation.
-This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks.
-Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
-This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
-This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash-lag illusion~\citep{khoei_flash-lag_2017}.
-However, these latter models used parametric rules for defining the wrights.
-Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore results from neurobiology to get better insights on that aspect.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Precise spiking motifs in neurobiology}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we will review current evidence of taking advantage of spiking motifs in neurobiology, that is, from actual recordings in biological tissues.
-In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
-For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process.
-As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.
-This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
-An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
-In that paper, authors fit summary statistics of neural data with a differentiable spiking network model.
-The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
-In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the publicly available \href{http://crcns.org/data-sets/vc/pvc-11}{V1-dataset} which allows supervising the model with the input being the movie and the output the spikes recorded.
-These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
-Using the right metric, as implemented in that paper by the chosen neural models, is essential in better understanding neural data.
-As such, all these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
-%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding neural activity using spike distances}\label{sec:detection}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-There are different solutions to provide with a distance between two given spike trains.
-A known measure is the Victor-Purpura distance, which showed that their findings are overcoming inconsistencies experienced with Poisson models of spike trains~\citep{victor_nature_1996}.
-In a later study,~\citet{van_rossum_novel_2001} have defined a distance that uses a time constant as a parameter.
-Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
-Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citet{kreuz_measuring_2007}.
-Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
-These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
-% ~\citep{}
-
-Let's now focus on distances defined based on a generative model formed by using precise spiking motifs.
-A review of the methods for neuronal response latency estimation exists and includes Bayesian binning~\citep{levakova_review_2015}.
-More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
-This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}.
-A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}.
-By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
-This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
-
-Another important algorithm, called SPOTDisClust, is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}.
-They introduced an unsupervised method based for their detection from high-dimensional neural ensembles.
-The algorithm measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other.
-Many approaches to this problem are supervised.
-In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
-In SPOTDisClust, the learning is unsupervised.
-It uses the prior that there is only one spike per pattern.
-Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data (see Figure~\ref{fig:SPOTDisClust}).
-The limits of this method are that it is computationally complex, block-based and strictly specialized for the task at hand.
-To overcome these difficulties, a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
-%   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
-
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
-\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in SPOTDisClust.
-A) Structure of five ``ground-truth'' patterns (\ldots).
-For each pattern and each neuron, a random position was chosen for the activation pulse.
-B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
-(© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
-\end{figure}
-%----------------------------%
-%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Scaling up to very large scale data}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
-Bridging across different spatial and temporal scales, combination of \textit{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
-The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurobiological knowledge to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
-As a consequence, there is a growing need to scale these methods to larger amounts of data.
-
-There are multiple solutions which aim at tackling this problem.
-One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
-Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
-The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
-In~\citet{stringer_spontaneous_2019,stringer_high-dimensional_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
-Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
-In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
-%   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
-
-% \subsection{Extending the decoding at multiple time scales}
-In the paper by~\citet{russo_cell_2017}, the authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.
-It uses  sliding window as in~\citep{grun_unitary_2002} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
-They extend the measure to multiple lags~\citep{torre_synchronous_2016}.
-The core measure is based on a non-stationarity corrected parametric test statistic for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
-To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
-They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
-At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
-They develop a Hidden Markov Model for quantifying the transitions.
-In particular, they show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
-Top-down attention drives inter-areal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
-Building such models for predicting changes of timings based on context, such as using a so-called change point model for blocked-based experimental protocols~\citep{pasturel_humans_2020}, could therefore provide useful prior information to enhance the decoding from neural activity.
-
-% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Explictly decoding spiking motifs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Explictly decoding spiking motifs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
 Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons \textit{in vivo} and \textit{in vitro}.
@@
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
-
 %----------------------------%
@@
 
-That results suggest that polychronous groups are not necessarily grouped on a topography and can be forming apparently randomly arranged connections.
+Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
+This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996} or touch~\citep{johansson_first_2004}.
+In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
+These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
+Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}).
+As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
+
+%
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
+\caption{An example of a polychronous group seen in cortical slices.
+In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
+Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
+The black line represents the speed of the mouse.}\label{fig:haimerl}
+\end{figure}
+%----------------------------%
+
 It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some (frozen) noise patterns.
-In paerticular, they used random waveforms to probe the formation of new memories for arbitrary complex sounds.
+In particular, they used random waveforms to probe the formation of new memories for arbitrary complex sounds.
 A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \si{\second}$ long.
@@
 
-Such observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
-~\citep{miller_visual_2014} shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
-It proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
-Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions.
-% (Fig.
-7).
+That results suggest that polychronous groups are not necessarily grouped on a topography and can be forming apparently randomly arranged connections.
+Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
+Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
+The rest of this review will be devoted to present evidence for the use of precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
 As a summary, it seems imperative to include such precise structural information should therefore be included in decoding algorithm for a better understanding of neural activity.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Precise spiking motifs in neuromorphic engineering}
+\section{Understanding precise spiking motifs in neurobiology}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In this section, we will review current evidence of taking advantage of spiking motifs in neurobiology, that is, from actual recordings in biological tissues.
+In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
+For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process.
+As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.
+This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
+An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
+In that paper, authors fit summary statistics of neural data with a differentiable spiking network model.
+The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
+In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the publicly available \href{http://crcns.org/data-sets/vc/pvc-11}{V1-dataset} which allows supervising the model with the input being the movie and the output the spikes recorded.
+These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
+Using the right metric, as implemented in that paper by the chosen neural models, is essential in better understanding neural data.
+As such, all these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
+%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decoding neural activity using spike distances}\label{sec:detection}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+There are different solutions to provide with a distance between two given spike trains.
+A known measure is the Victor-Purpura distance, which showed that their findings are overcoming inconsistencies experienced with Poisson models of spike trains~\citep{victor_nature_1996}.
+In a later study,~\citet{van_rossum_novel_2001} have defined a distance that uses a time constant as a parameter.
+Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
+Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citet{kreuz_measuring_2007}.
+Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
+These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
+% ~\citep{}
+
+Let's now focus on distances defined based on a generative model formed by using precise spiking motifs.
+A review of the methods for neuronal response latency estimation exists and includes Bayesian binning~\citep{levakova_review_2015}.
+More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
+This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}.
+A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}.
+By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
+This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
+
+Another important algorithm, called SPOTDisClust, is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}.
+They introduced an unsupervised method based for their detection from high-dimensional neural ensembles.
+The algorithm measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other.
+Many approaches to this problem are supervised.
+In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
+In SPOTDisClust, the learning is unsupervised.
+It uses the prior that there is only one spike per pattern.
+Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data (see Figure~\ref{fig:SPOTDisClust}).
+The limits of this method are that it is computationally complex, block-based and strictly specialized for the task at hand.
+To overcome these difficulties, a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
+%   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
+
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
+\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in SPOTDisClust.
+A) Structure of five ``ground-truth'' patterns (\ldots).
+For each pattern and each neuron, a random position was chosen for the activation pulse.
+B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
+(© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
+\end{figure}
+%----------------------------%
+%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Scaling up to very large scale data}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
+Bridging across different spatial and temporal scales, combination of \textit{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
+The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurobiological knowledge to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
+As a consequence, there is a growing need to scale these methods to larger amounts of data.
+
+There are multiple solutions which aim at tackling this problem.
+One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
+Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
+The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
+In~\citet{stringer_spontaneous_2019,stringer_high-dimensional_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
+%   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
+
+% \subsection{Extending the decoding at multiple time scales}
+In the paper by~\citet{russo_cell_2017}, the authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.
+It uses  sliding window as in~\citep{grun_unitary_2002} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
+They extend the measure to multiple lags~\citep{torre_synchronous_2016}.
+The core measure is based on a non-stationarity corrected parametric test statistic for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
+They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
+At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
+They develop a Hidden Markov Model for quantifying the transitions.
+In particular, they show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
+Top-down attention drives inter-areal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
+Building such models for predicting changes of timings based on context, such as using a so-called change point model for blocked-based experimental protocols~\citep{pasturel_humans_2020}, could therefore provide useful prior information to enhance the decoding from neural activity.
+
+% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
+
+\subsection{What biological mechanism enables delay learning?}
+
+Despite the extensive computer evidence of the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
+The discovery of the existence of an equivalent biological mechanism in Vertebrates as well as the demonstration of the importance of its role in various developed behaviours allow us to put these doubts to rest.
+In the following paragraphs we will successively describe the first biological observations of biological delay learning, identify myelination as an important actor and finally study this phenomenon at the molecular level.
+
+% First biological observations
+
+One of the first significant evidence of any neuronal delay in the information propagation within the animal brain came from Young's study of the squid giant axon~\cite{young_1938}:
+
+However, it was not until the study of the interaural time difference (ITD) by Gerstner in 1996~\citep{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
+This ITD is a biological mechanism which allows for the azimutal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
+According to \citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.
+\citet{pena_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
+A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
+the temporal disparity between left and right cochlear loci in the owl's brain.
+However it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variations were perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
+The authors support a second hypothesis, that of different axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
+\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
+These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously thought.
+
+% The importance of myelination
+
+\citet{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
+the amount of myelin wrapped around the axon.
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon~\citep{schmidt_1939}
+
+Over and over, myelin thus has been identified as a main actor in the delay regulation in many biological behaviors.
+
+% role of myelination and neuronal delay
+
+facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
+Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments.
+This biological phenomenom both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
+It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
+
+% How is the neuronal activity detected ?
+
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Modeling precisely-timed spiking motifs in computational neuroscience}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In the following section, we will review theoretical models which directly take advantage of using precise spiking motifs. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
+The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
+This approach is quite successful, and SNNs reach in some case a similar performance as Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
+However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
+There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
+These models will be compared in light of neuroscientific and computational perspectives.
+We review in this section theoretical and computational foundations of PG detection in models.
+In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Izhikevitch's polychronization model}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% polychronization
+As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on an encoding of information based on a continuously varying firing-rate.
+A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}.
+This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
+It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
+It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation.
+Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
+The neurons composing a group discharge at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
+This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:PG}).
+According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, constitute a polychronous group.
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
+In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
+This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
+%----------------------------%
+\begin{figure}
+\centering
+\note{insert fig 2 of izhikevich 2006 - or an open-source reproduction like that from Camille ? }
+%includegraphics{figures/event_driven_computations.png}
+\caption{blah.}\label{fig:PG}
+\end{figure}
+%----------------------------%
+
+% follow-ups
+Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper.
+Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs.
+A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}.
+The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron which does not incorporate a spike timing framework.
+The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a STDP rule.
+The limits of this model is that its output is only binary and that its storage capacity are limited.
+The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
+Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
+As such, these models do not explicitly take full advantage of the representation capacity offered by polychronous groups.
+
+An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
+Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript.
+This more recent work details how this code was slightly corrected.
+It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
+Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
+Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation.
+They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system.
+Stimulation of the trained network produces the activation of a PG, i.e.
+the propagation of firing activity through multiple layers due to convergent patterns of firing.
+While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Learning synaptic delays}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Training delays ~\citep{state_training_2019}
+First, the original model use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
+For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
+Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
+They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
+In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
+Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
+
+Furthermore, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
+The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns.
+Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task.
+Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
+This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
+Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
+In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron.
+In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
+In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.
+
+%In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in birdsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
+In~\citep{agus_rapid_2010}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
+
+% Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
+% Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021
+
+% Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
+% Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020
+
+% Unsupervised Learning of Persistent and Sequential Activity.
+% Pereira U, Brunel N.Front Comput Neurosci. 2020
+
+% From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
+% Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019
+
+% Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
+% Pang R, Fairhall AL.Elife. 2019 May~\citep{pang_fast_2019}
+
+% Emergence of spontaneous assembly activity in developing neural networks without afferent input.
+% Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018
+
+% Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
+% Ocker GK, Doiron B.Cereb Cortex. 2019.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Real-world applications}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios.
+Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility.
+For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
+The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern.
+The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity.
+For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
+Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.
+
+There are more applications to image processing using spiking neural networks based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections.
+Using this core computational unit defined, and extending to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation.
+This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks.
+Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
+This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
+This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash-lag illusion~\citep{khoei_flash-lag_2017}.
+However, these latter models used parametric rules for defining the wrights.
+Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore results from neurobiology to get better insights on that aspect.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Applications of precise spiking motifs in neuromorphic engineering}
 \label{sec:neuromorphic}
@@
 
-%\subsection{Temporal pattern detection in event-based cameras}
+\subsection{Temporal pattern detection in event-based cameras}
 % le cas spécifique des event-based cameras
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Spiking neural models with heterogeneous delays}
+\subsection{Summary}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Here, we develop a model for the efficient detection of such PGs based on the inversion of a probabilistic model defining the generation of the raster plot as a combination of such groups.
-We show that such an inference can be achieved by a neural-like computation that could itself be used as a spiking neuron, as can be implemented in a neuromorphic chip for instance.
-A first result is to show the efficiency of such a scheme in detecting different PGs occurring at specific times in synthetic data.
-The representational capacity of the PGs is particularly interesting compared to traditional models of neuronal encoding using spiking frequency.
-Our second result is to propose a novel learning method for learning PGs in raster plots in a self-supervised manner.
-Finally, we demonstrate the use of this algorithm to the output of an event-based camera and how this may separate independent components from the stream of events.
-This end-to-end event-based computational brick could help improve the performance of current Spiking Neural Network solution currently used in neuromorphic chips.
+In this review paper, we have presented
+recent evidence for the role of precise
+spiking motifs in neuroscience. In
+particular, we have shown that those
+motifs may play a crucial role in
+neurobiology, that they may be understood
+at a theoretical and computational levels
+and that they may have numerous applications
+in neuromorphic engineering.
+In particular, we have shown that:
+- the efficiency of neural systems, and
+in particular the visual system, imposes
+strong constraints on the structure of
+neural activity which highlight the
+importance of precise spike times,
+- that growing evidence from neurobiology
+proves that neural systems are we ne thar
+integrators and may use synchrony
+detection in different forms: synfne chains,
+traveling waves on arbitrary spiking motif,
+- that many theoretical models exit
+taking into account the specificity of
+spiking motifs, notably by using
+heterogeneous delays,
+- ultimately that using precise puking
+motifs could be a key ingredient in
+men nomorphic systems to reach similar
+efficiencies as biological neural systems.
+Ova, our reviewing effort has shown
+that a growing community is focusing
+on that aspect. This community is
+based on solid and validated evidence
+which is getting novel grounds thanks h
+current advances in neurobiological
+technical advances. Moreover, we have also
+shown that this community as highly
+diverse, ranging from in vivo, he in
+computo and in silico. As a consequence
+the global effort is still largely scattered,
+which limits its larger acceptance in
+neuroscience.
 
@@
 
-Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks \textit{in vivo}~\citep{isbister_clustering_2021} They discovered (\textit{in vivo}) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
+Les différents modèles de detection et d’apprentissage de motifs de Spike que nous avons présenté aux différents niveaux (neurobiologiques, théoriques, neuromorphic ) présentent individuellement des limites qui empêchent leur généralisation en neurosciences.
+
+Tout d’abord, de nombreux modèles sont basé sur une discretisation du temps.
+cette hypothèse est importante pour permettre une représentation correcte de l'information neuronale afin d’etre traité dans des ordinateurs de façon computationnelle.
+Ce traitement revient à transformer les Rasster de Spike sous forme matricielle pour leur traitement dans des algorithmes et des formalisme classique d’apprentissage machine classique.
+cette hypothèse implique un usage inefficace de la mémoire, car cette représentation transforme la représentation parcimonieuse d’une séquence de Spike en des matrices denses.
+D’autre part, cette représentation peut induire des erreurs dû à la discréditassions et à l'échelle de l’échantillonnage temporel.
+Enfin, cette représentation encourage l'utilisation de méthodes classiques,qui ne sont pas adaptées à des applications disruptives telles que les représentations event-driven.
+
+D’autre part, l’apprentissage des motifs est souvent effectué de façon supervisé.
+En effet, le problème de détecter des groupes polichrone implique d’inférer à la fois l’adresse et le temps précis de occurrence de ces motifs.
+la plupart des modèles que nous avons présenté sont basé sur l’hypothèse qu'au moins une de ses variables est connu : soit le pattern, soit son identité, soit son temps d’occurrence.
+Cette contrainte est à mettre en parallèle du fonctionnement d’un système nerveux biologique dans lequel ses apprentissages sont effectués de façon autonome, c’est-à-dire sans supervision.
+On peut noter toutefois que certains modèles peuvent effectuer un tel apprentissage, mais seulement, dans le cas de données pour lequel les motifs sont aisément séparables.
+
+Une dernière limite des modèles que nous avons présenté et le fait que ceci sont constitués d’une seule couche de traitement qui relit une entrée à une sortie.
+Nous avons vu que toutefois le système neurobiologiques utilise des boucles de traitement.
+
+%Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks \textit{in vivo}~\citep{isbister_clustering_2021} They discovered (\textit{in vivo}) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
 % dynamical models
-Dumas and colleagues~\citep{panahi_generative_2021} : three levels / fourth paradigm~\citep{tolle_fourth_2011} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
-Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{di_mauro_alfio_sne_2022} or of sparsity aware algorithms~\citep{yin_sata_2022}.
-
-and oscillations? Journal Article DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?
+%Dumas and colleagues~\citep{panahi_generative_2021} : three levels / fourth paradigm~\citep{tolle_fourth_2011} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
+%Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{di_mauro_alfio_sne_2022} or of sparsity aware algorithms~\citep{yin_sata_2022}.
+
+%and oscillations? Journal Article DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?
 
 % development
-scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2022}
-
-\subsection{What biological mechanism enables delay learning?}
-
-Despite the extensive computer evidence of the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
-The discovery of the existence of an equivalent biological mechanism in Vertebrates as well as the demonstration of the importance of its role in various developed behaviours allow us to put these doubts to rest.
-In the following paragraphs we will successively describe the first biological observations of biological delay learning, identify myelination as an important actor and finally study this phenomenon at the molecular level.
-
-% First biological observations
-
-One of the first significant evidence of any neuronal delay in the information propagation within the animal brain came from Young's study of the squid giant axon~\cite{young_1938}:
-
-However, it was not until the study of the interaural time difference (ITD) by Gerstner in 1996~\citep{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
-This ITD is a biological mechanism which allows for the azimutal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-According to \citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.
-\citet{pena_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
-A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
-the temporal disparity between left and right cochlear loci in the owl's brain.
-However it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variations were perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
-The authors support a second hypothesis, that of different axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
-\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously thought.
-
-% The importance of myelination
-
-\citet{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
-the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon~\citep{schmidt_1939}
-
-Over and over, myelin thus has been identified as a main actor in the delay regulation in many biological behaviors.
-
-% role of myelination and neuronal delay
-
-facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
-Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments.
-This biological phenomenom both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
-
-% How is the neuronal activity detected ?
-
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.
+%scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2022}
+
+% todo: implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
+it may be difficult but solutions exist  ~\citep{foss_multistability_2000,perrinet_active_2014}
+
+%hierarchies
+
+%developpment
 
@@
 
-todo: implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
-it may be difficult but solutions exist  ~\citep{foss_multistability_2000,perrinet_active_2014}
-
-hierarchies
-
-developpment
+Mt#the limits that we have
+presented can be treated individually
+as evidenced by individual efforts
+which try he orecome them. However,
+to propose a real breakthrough, we
+believe that future venues should
+provide with a unified, inter-disciplinary
+approach, with applications to deal-
+word, ecological scenarios and
+with open and reproducible methods.
+First, as we already noted, the effort
+is still largely scattered -this is
+in part due to the fact that interaction
+between neurobiology, theoretical
+and computational neuroscience
+and meromorphic engineering
+are still scarce as of today.
+It was largely demonstrated that
+close, bi-directional interactions
+are essential to foster breakthrough
+For instance, the design of model-
+driven protocols as proven to be
+essential in modern neuroscience.
+Also, if neural networks were
+essential in shaping modern-day
+machine learning, e-g. computer
+vision using deep-learning,
+spiking neural networks should
+prove essential in future
+emerging technologies.
+
+In that perspective, it is
+essential that such models
+are tested on ecologically-
+relevant, real-world scenarios.
+Indeed, classic convolutional
+neural networks emerged
+as optimal solutions for
+instance he classify static
+images into categories, yet
+thay are not well adapted
+for processing dynamic,
+multimodal sensory flows.
+The emerging necessity to be
+able to process more complex
+flows, such as the multiples
+flows of information in a
+car designed for autonomous
+driving, necessitates to
+modify modelling paradigms.
+and in particular to take into
+account that the generated actions
+may modify the sensory inputs.
+Notably, the protocols used
+as well in neurobiology,
+theory on engineering should
+take into account this hovel
+complexity.
+Ultimately, the community should
+encourage the adoption of open,
+reproducible science. Indeed,
+the different models that we have
+displayed often come with the
+tools necessary he reproduce the
+results obtained. This is true
+in neurobiology (Belles), in
+theoretical neuroscience (Ishi)
+or in engineering (PAMI).
+This aspect is essential to
+foster the emergence of inter-
+disciplinary projects, such
+as model-driven neurobiological
+experiments or biologically-
+inspired vennomorphic
+engineering.
+As such, this review aimed
+at paving the way to openly
+share the variety of
+resources and to offer an
+unified view on the role of
+precise spiking motifs in
+neuroscience.
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 
-Les différents modèles de detection et d’apprentissage de motifs de Spike que nous avons présenté aux différents niveaux (neurobiologiques, théoriques, neuromorphic ) présentent individuellement des limites qui empêchent leur généralisation en neurosciences.
-
-Tout d’abord, de nombreux modèles sont basé sur une discretisation du temps.
-cette hypothèse est importante pour permettre une représentation correcte de l'information neuronale afin d’etre traité dans des ordinateurs de façon computationnelle.
-Ce traitement revient à transformer les Rasster de Spike sous forme matricielle pour leur traitement dans des algorithmes et des formalisme classique d’apprentissage machine classique.
-cette hypothèse implique un usage inefficace de la mémoire, car cette représentation transforme la représentation parcimonieuse d’une séquence de Spike en des matrices denses.
-D’autre part, cette représentation peut induire des erreurs dû à la discréditassions et à l'échelle de l’échantillonnage temporel.
-Enfin, cette représentation encourage l'utilisation de méthodes classiques,qui ne sont pas adaptées à des applications disruptives telles que les représentations event-driven.
-
-D’autre part, l’apprentissage des motifs est souvent effectué de façon supervisé.
-En effet, le problème de détecter des groupes polichrone implique d’inférer à la fois l’adresse et le temps précis de occurrence de ces motifs.
-la plupart des modèles que nous avons présenté sont basé sur l’hypothèse qu'au moins une de ses variables est connu : soit le pattern, soit son identité, soit son temps d’occurrence.
-Cette contrainte est à mettre en parallèle du fonctionnement d’un système nerveux biologique dans lequel ses apprentissages sont effectués de façon autonome, c’est-à-dire sans supervision.
-On peut noter toutefois que certains modèles peuvent effectuer un tel apprentissage, mais seulement, dans le cas de données pour lequel les motifs sont aisément séparables.
-
-Une dernière limite des modèles que nous avons présenté et le fait que ceci sont constitués d’une seule couche de traitement qui relit une entrée à une sortie.
-Nous avons vu que toutefois le système neurobiologiques utilise des boucles de traitement.
+The different models of spike motif detection and learning that we have presented at different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
+
+First of all, many models are based on a discretization of time.
+This assumption is important to allow a correct representation of the neuronal information in order to be processed in computers.
+This treatment amounts to transforming the Spike Raster into a matrix form for processing in classical machine learning algorithms.
+This assumption implies an ineffective use of the memory because this representation transforms the sparse representation of a spike sequence into dense matrices.
+In addition, this representation can induce errors due to the discretization and the scale of temporal sampling.
+Finally, this representation encourages the use of classical methods, which are not adapted to disruptive applications such as event-driven representations.
+
+Moreover, the learning of patterns is often done in a supervised way.
+Indeed, the problem of detecting polychronous groups implies to infer both the address and the precise time of occurrence of these motifs.
+Most of the models we have presented are based on the assumption that at least one of its variables is known: either the pattern, its identity, or its time of occurrence.
+This constraint is to be put in parallel with the way a biological nervous system works in which learning is done autonomously, i.e.
+without supervision.
+However, we can note that some models can perform such learning, but only in the case of data for which the motifs are easily separable.
+
+A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output.
+However, we have seen that the neurobiological system uses processing loops.
+
+Translated with www.DeepL.com/Translator (free version)
 

### 2022-11-08 manuscript.tex
Update on Overleaf.

@@
 
-Translated with www.DeepL.com/Translator (free version)
-
 %Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks \textit{in vivo}~\citep{isbister_clustering_2021} They discovered (\textit{in vivo}) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
@@
 
-Mt#the limits that we have
+The limits that we have
 presented can be treated individually

### 2022-11-08 manuscript.tex
Discussion
Update on Overleaf.

@@
 
-However, it was not until the study of the interaural time difference (ITD) by Gerstner in 1996~\citep{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
+However, it was not until the study of the interaural time difference (ITD) by~\citet{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
 This ITD is a biological mechanism which allows for the azimutal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
@@
 
-In this review paper, we have presented
-recent evidence for the role of precise
-spiking motifs in neuroscience. In
-particular, we have shown that those
-motifs may play a crucial role in
-neurobiology, that they may be understood
-at a theoretical and computational levels
-and that they may have numerous applications
-in neuromorphic engineering.
+In this review paper, we have presented recent evidence for the role of precise spiking motifs in neuroscience.
+In particular, we have shown that those motifs may play a crucial role in neurobiology, that they may be understood at a theoretical and computational levels and that they may have numerous applications in neuromorphic engineering.
 In particular, we have shown that:
-- the efficiency of neural systems, and
-in particular the visual system, imposes
-strong constraints on the structure of
-neural activity which highlight the
-importance of precise spike times,
-- that growing evidence from neurobiology
-proves that neural systems are we ne thar
-integrators and may use synchrony
-detection in different forms: synfne chains,
-traveling waves on arbitrary spiking motif,
-- that many theoretical models exit
-taking into account the specificity of
-spiking motifs, notably by using
-heterogeneous delays,
-- ultimately that using precise puking
-motifs could be a key ingredient in
-men nomorphic systems to reach similar
-efficiencies as biological neural systems.
-Ova, our reviewing effort has shown
-that a growing community is focusing
-on that aspect. This community is
-based on solid and validated evidence
-which is getting novel grounds thanks h
-current advances in neurobiological
-technical advances. Moreover, we have also
-shown that this community as highly
-diverse, ranging from in vivo, he in
-computo and in silico. As a consequence
-the global effort is still largely scattered,
-which limits its larger acceptance in
-neuroscience.
+- the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
+- that growing evidence from neurobiology proves that neural systems are we ne thar integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motif,
+- that many theoretical models exit taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
+- ultimately that using precise puking motifs could be a key ingredient in men nomorphic systems to reach similar efficiencies as biological neural systems.
+Ova, our reviewing effort has shown that a growing community is focusing on that aspect.
+This community is based on solid and validated evidence which is getting novel grounds thanks h current advances in neurobiological technical advances.
+Moreover, we have also shown that this community as highly diverse, ranging from in vivo, he in computo and in silico.
+As a consequence the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
 
@@
 The limits that we have
-presented can be treated individually
-as evidenced by individual efforts
-which try he orecome them. However,
-to propose a real breakthrough, we
-believe that future venues should
-provide with a unified, inter-disciplinary
-approach, with applications to deal-
-word, ecological scenarios and
-with open and reproducible methods.
-First, as we already noted, the effort
-is still largely scattered -this is
-in part due to the fact that interaction
-between neurobiology, theoretical
-and computational neuroscience
-and meromorphic engineering
-are still scarce as of today.
-It was largely demonstrated that
-close, bi-directional interactions
-are essential to foster breakthrough
-For instance, the design of model-
-driven protocols as proven to be
-essential in modern neuroscience.
-Also, if neural networks were
-essential in shaping modern-day
-machine learning, e-g. computer
-vision using deep-learning,
-spiking neural networks should
-prove essential in future
-emerging technologies.
-
-In that perspective, it is
-essential that such models
-are tested on ecologically-
-relevant, real-world scenarios.
-Indeed, classic convolutional
-neural networks emerged
-as optimal solutions for
-instance he classify static
-images into categories, yet
-thay are not well adapted
-for processing dynamic,
-multimodal sensory flows.
-The emerging necessity to be
-able to process more complex
-flows, such as the multiples
-flows of information in a
-car designed for autonomous
-driving, necessitates to
-modify modelling paradigms.
-and in particular to take into
-account that the generated actions
-may modify the sensory inputs.
-Notably, the protocols used
-as well in neurobiology,
-theory on engineering should
-take into account this hovel
-complexity.
-Ultimately, the community should
-encourage the adoption of open,
-reproducible science. Indeed,
-the different models that we have
-displayed often come with the
-tools necessary he reproduce the
-results obtained. This is true
-in neurobiology (Belles), in
-theoretical neuroscience (Ishi)
-or in engineering (PAMI).
-This aspect is essential to
-foster the emergence of inter-
-disciplinary projects, such
-as model-driven neurobiological
-experiments or biologically-
-inspired vennomorphic
-engineering.
-As such, this review aimed
-at paving the way to openly
-share the variety of
-resources and to offer an
-unified view on the role of
-precise spiking motifs in
-neuroscience.
+presented can be treated individually as evidenced by individual efforts which try he orecome them.
+However, to propose a real breakthrough, we believe that future venues should provide with a unified, inter-disciplinary approach, with applications to deal- word, ecological scenarios and with open and reproducible methods.
+First, as we already noted, the effort is still largely scattered -this is in part due to the fact that interaction between neurobiology, theoretical and computational neuroscience and meromorphic engineering are still scarce as of today.
+It was largely demonstrated that close, bi-directional interactions are essential to foster breakthrough For instance, the design of model- driven protocols as proven to be essential in modern neuroscience.
+Also, if neural networks were essential in shaping modern-day machine learning, e-g.
+computer vision using deep-learning, spiking neural networks should prove essential in future emerging technologies.
+
+In that perspective, it is essential that such models are tested on ecologically- relevant, real-world scenarios.
+Indeed, classic convolutional neural networks emerged as optimal solutions for instance he classify static images into categories, yet thay are not well adapted for processing dynamic, multimodal sensory flows.
+The emerging necessity to be able to process more complex flows, such as the multiples flows of information in a car designed for autonomous driving, necessitates to modify modelling paradigms.
+and in particular to take into account that the generated actions may modify the sensory inputs.
+Notably, the protocols used as well in neurobiology, theory on engineering should take into account this hovel complexity.
+
+Ultimately, the community should encourage the adoption of open, reproducible science.
+Indeed, the different models that we have displayed often come with the tools necessary he reproduce the results obtained.
+This is true in neurobiology (Belles), in theoretical neuroscience (Ishi) or in engineering (PAMI).
+This aspect is essential to foster the emergence of inter- disciplinary projects, such as model-driven neurobiological experiments or biologically- inspired vennomorphic engineering.
+As such, this review aimed at paving the way to openly share the variety of resources and to offer an unified view on the role of precise spiking motifs in neuroscience.
 

### 2022-11-08 manuscript.tex
abstract

@@
 Why do neurons communicate through spikes?
-By definition, a spike is an all-or-none event which occurs at continuous times.
-In other words, spikes are on one side binary, arising or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock.
-This stands in stark contrast to the discretized timing classically used in digital processing at the base of modern-day neural networks.
-As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data.
+By definition, a spike is an all-or-none neural event which occurs at continuous times.
+In other words, spikes are on one side binary, existing or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock.
+This stands in stark contrast to the analog represnetation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks.
+As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the profusion of recorded data.
 With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption --- a major societal issue in the era of the digital economy and global warming.

### 2022-11-08 manuscript.tex
1

@@
 The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection.
-A central question of modern neuroscience is to better understand the essence of neural activity, as the recordings that can be observed in neurobiological experiments.
+A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings that can be observed in neurobiological experiments.
 One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
@@
 Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and possibly further propagated along its "output" axon to reach efferent neurons The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
-Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance as computedf in small time windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
+Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance as computed as the average occurrence of spikes in small time windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
 This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.

### 2022-11-08 manuscript.tex
Update on Overleaf.
1

@@
 For instance, it expands the capabilities of analog representations (which rely on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
-This hypothesis is directly inspired by neurobiological observations, particularly in the hippocampus, and it expands the capabilities of analog representations (which are based on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
+This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.
 Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and also suggest the existence of repeated spatio-temporal spiking motifs in the spike train reaching neurons.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons.
 A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
-Importantly, validating this hypothesis would also be crucial in our understanding of neural processes.
+Crucially, validating this hypothesis would also be crucial in our understanding of neural processes.
+%
 %
@@
 %
-Let us start with a review of the state of the art on the role of dynamics in vision.
+Let us start with a focus of the state-of-the-art on the role of dynamics in vision.
 Broadly speaking, vision is the set of processes that allow us to make sense of the luminous world, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
 In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between the topology of external visual space and its representation on the retina, called retinotopy.
-The origins of this neuroscientific question can be found in the first experiments from Pierre Flourens showing the relationship between visual sensations and activity in the cerebral cortex using lesions in animals~\citep{flourens_recherches_1842,pearce_marie-jean-pierre_2009}.
+The origins of this neuroscientific question can be found in the first experiments from Pierre Flourens which, using lesions in animals, demonstrated the relationship between visual sensations and activity in the cerebral cortex~\citep{flourens_recherches_1842,pearce_marie-jean-pierre_2009}.
 This was also observed by Lord Adrian when recording the activity the frog's visual system~\citep{adrian_impulses_1926}.
-In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as orientation or motion.
+In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as the visual orientation or motion of elementary contours.
 For a large proportion of neurons, there is remarkably a monotonic relationship between the contrast of visual features and the firing frequency of neurons.

### 2022-11-08 manuscript.tex
1

@@
 For instance, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
-Its first experiment consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
+His experiments consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
 The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
 These results have been extended to several species, including primates.
-Different experimental protocols have shown for example that the response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
-This fast processing also explains the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
+Different experimental protocols have shown for example that the motor response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
+This fast processing correlates with the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
 As expected, the performances decrease progressively as the frequency of presentation of the images increases.
@@
 This functioning of visual processing as a forward pass is most prominent in fast processing, but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
-An important consequence of the speed of processing of vision is that it implies that it is carried out using only very few spikes.
+An important consequence of this speed of processing of vision is that it implies that procesing is carried out using only very few spikes per layer.
 Indeed, we may consider that a behavioral response occurring in only $120~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
@@
 
-%----------------------------%
-\begin{figure}
-\centering
-\note{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
-\caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
-}\label{fig:roc}
-\end{figure}
-%----------------------------%
+% %----------------------------%
+% \begin{figure}
+% \centering
+% \note{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
+% \caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
+% }\label{fig:roc}
+% \end{figure}
+% %----------------------------%
 In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.

### 2022-11-08 manuscript.tex
1

@@
 The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
-For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it inconsistent with temporal integration of firing rate.
-Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned, but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
-However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see\citep{ermentrout_reliability_2008}).
+For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it at first sight inconsistent with the temporal integration of firing rate.
+Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: whike the first spike are aligned, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
+However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see~\citep{ermentrout_reliability_2008}).
 This is consistent with the differential role of different stimulus frequencies (for instance the gamma range around $80~\si{\Hz}$) on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
@@
 % retina
-At the level of the dynamic processing of visual information, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work in the retina~\citep{gollisch_rapid_2008}.
+How may information about the sensory world be encoded in the spiking pattern?
+At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
 In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
@@
 Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
-This evidence found in the retina  can be extended to the visual cortex.
+This evidence found in the retina can be extended to other areas such as the visual cortex.
 
@@
 One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron by using the rank of neurons' activation~\citep{delorme_spikenet_1999,bonilla_analyzing_2022}.
-Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
+Another class of artificial SNNs use precise spike timing as a metric in order to determine the structure of the network in order to minimize a cost function.
 This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods.
-The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
-This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
+The subsequent surrogate gradient method is now widely used in methods that attempt to transfer performance from analog (CNNs) to spike-based (SNNs) architectures~\citep{zenke_remarkable_2021}.
+This type of modeling often uses the classical task of categorizing images developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition or~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
@@
 However, the performance of SNNs is still lagging compared to that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
-Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas which represents the visual information as events, similarly to the representations described above and instead of using a basic frame-based representation~\citep{lichtsteiner_128-128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}).
+Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas.
+These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation~\citep{lichtsteiner_128-128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}).
 % This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.

### 2022-11-08 manuscript.tex
1

@@
 Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas.
-These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation~\citep{lichtsteiner_128-128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}).
+These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}).
 % This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
@@
 The book gradually leads the reader from the macroscopic cortical anatomy and standard electro-physiological properties of single neurons to neural network models.
-While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
+While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
 The book then demonstrates that this could take the form of ``synfire chains,'' that is, the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
+More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}.
 Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
@@
 Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}.
-Interestingly, the authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Interestingly, these authors showed that ``accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
 Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018}, and were also extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
@@
 
-More generally, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}. 
 This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}.

### 2022-11-08 manuscript.tex
2

@@
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018}, and were also extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018}, and were also extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
 Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intra-cortical activity.
-However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active~\citep{bruno_cortex_2006}.
-
-In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
+However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons~\citep{bruno_cortex_2006}.
+
+Theoretically, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
 Each neuron of one group is connected by an excitatory synapse to the next.
 When a pulse is elicited in the first group, this may generate a spike in the next group.
-Depending on the concentration of synaptic weight values, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse).
-Recursively applying this to a sequence of groups generates either a synfire propagation or not.
+Depending on the concentration of synaptic weight values, this new activity may become more or less synchronized with respect to that of the previous pulse (as measured by the standard deviation of spike times within the pulse).
+Recursively applying this to a sequence of groups within a chain generates either a synfire propagation or not.
 A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
@@
 It relies on the simple remark that a series of repeated coincidences is in itself an invariant.
-Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, the pitch of a musical note.
+Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, or the pitch of a musical note.
 
@@
 More generally, it can be shown that the phase of alpha oscillations (about $10$ Hz) is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
+The questions still remains open as to know if this is an epiphenomenom or a working mechanism of the neural code.
 %
@@
 Using intracellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is, a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
-In particular, it was shown that the network of the so-called horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within that area.
+In particular, it was shown that the network of the so-called horizontal connections (that is, connections within a cortical area) are typically not myelinated and that this latency basin would be determined by the propagation speed within that area.
 For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions and may underlay anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
-%
+The underlying process could be the emergence of propagating waves on the surface of the cortex.
 
@@
 A recent work~\citep{muller_stimulus-evoked_2014} presents a method for detecting propagating waves in noisy multichannel recordings.
-Applying this method to single-trial voltage-sensitive dye imaging data, they show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
+Applying this method to single-trial voltage-sensitive dye imaging data, the authors show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
 A network model suggests that this reliability is the hallmark of the horizontal fiber network of superficial cortical layers.
@@
 Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \textit{in vivo}.
-Such waves also occur in motor areas and~\citet{linden_movement_2021} have presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+Such waves also occur in motor areas and~\citet{linden_movement_2021} have recently presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
 

### 2022-11-08 manuscript.tex
2

@@
 However, a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
-Recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
+Moreover, recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
 ~\citep{miller_visual_2014} shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
@@
 %We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
-Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons \textit{in vivo} and \textit{in vitro}.
+Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \textit{in vivo} and \textit{in vitro}.
 These patterns repeat after minutes, maintaining millisecond accuracy.
@@
 It consists of the assembly of several sequences which repeat in a specific order with a compressed timing.
-\emph{In vivo} spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3-10 cells out of about 900, and lasting up to 3 seconds.
+\emph{In vivo} spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3-10 cells out (of about 900), and lasting up to 3 seconds.
 Sequences have specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and are associated with a structured spatial organization of the neurons that formed them.
@@
 In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
-These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
-Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
+These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and may be formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
+In that protocol, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
 They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}).
@@
 
-That results suggest that polychronous groups are not necessarily grouped on a topography and can be forming apparently randomly arranged connections.
+That results suggest that precise spiking motifs are not necessarily grouped on a topography and can be forming apparently randomly arranged connections.
 Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
 Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
-The rest of this review will be devoted to present evidence for the use of precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
-As a summary, it seems imperative to include such precise structural information should therefore be included in decoding algorithm for a better understanding of neural activity.
+The rest of this review will be devoted to present evidence for the use of such precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
+As a summary, it seems that such precise structural information is essential to the neural code and that it seems imperative to include this information in decoding algorithm for a better understanding of neural activity.
 

### 2022-11-09 manuscript.tex
3

@@
 An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
-In that paper, authors fit summary statistics of neural data with a differentiable spiking network model.
-The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995} with recurrent dynamics).
+In this particular paper, authors fit summary statistics of neural data with a (differentiable) spiking network model.
+The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995}) and embedded with recurrent dynamics.
 In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the publicly available \href{http://crcns.org/data-sets/vc/pvc-11}{V1-dataset} which allows supervising the model with the input being the movie and the output the spikes recorded.
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
-Using the right metric, as implemented in that paper by the chosen neural models, is essential in better understanding neural data.
-As such, all these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
+Using the right metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.
+Importantly, these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
 %
@@
 There are different solutions to provide with a distance between two given spike trains.
-A known measure is the Victor-Purpura distance, which showed that their findings are overcoming inconsistencies experienced with Poisson models of spike trains~\citep{victor_nature_1996}.
+A known measure is the Victor-Purpura distance, which  overcomes inconsistencies experienced with a simple firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
 In a later study,~\citet{van_rossum_novel_2001} have defined a distance that uses a time constant as a parameter.
 Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
-Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citet{kreuz_measuring_2007}.
+Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.
 Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
@@
 Many approaches to this problem are supervised.
-In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
+In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for determining hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
 In SPOTDisClust, the learning is unsupervised.
@@
 To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
-They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
+They also introduce learnable time-warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
 At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
@@
 
-\subsection{What biological mechanism enables delay learning?}
+\subsection{What biological mechanism could allow learning spiking motifs?}
 
 Despite the extensive computer evidence of the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
-The discovery of the existence of an equivalent biological mechanism in Vertebrates as well as the demonstration of the importance of its role in various developed behaviours allow us to put these doubts to rest.
+The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviours allow us to put these doubts to rest.
 In the following paragraphs we will successively describe the first biological observations of biological delay learning, identify myelination as an important actor and finally study this phenomenon at the molecular level.

### 2022-11-09 manuscript.tex
4

@@
 Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns.
 The neurons composing a group discharge at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
@@
 Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation.
+In that work, based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation.
 They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system.
@@
 
-Furthermore, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
+At the implementation level, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
 The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns.
@@
 Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility.
-For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
+For instance, in a recent work,~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
 The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern.
@@
 This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash-lag illusion~\citep{khoei_flash-lag_2017}.
-However, these latter models used parametric rules for defining the wrights.
-Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore results from neurobiology to get better insights on that aspect.
+However, these latter models used parametric rules for defining the weights.
+Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore results from neuromorphic engineering to get better insights on that aspect.
 %

### 2022-11-09 manuscript.tex
6

@@
 %: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
-
 In this review paper, we have presented recent evidence for the role of precise spiking motifs in neuroscience.
-In particular, we have shown that those motifs may play a crucial role in neurobiology, that they may be understood at a theoretical and computational levels and that they may have numerous applications in neuromorphic engineering.
+In particular, we have shown that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels and that they may have numerous applications in neuromorphic engineering.
 In particular, we have shown that:
-- the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
-- that growing evidence from neurobiology proves that neural systems are we ne thar integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motif,
-- that many theoretical models exit taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
-- ultimately that using precise puking motifs could be a key ingredient in men nomorphic systems to reach similar efficiencies as biological neural systems.
-Ova, our reviewing effort has shown that a growing community is focusing on that aspect.
-This community is based on solid and validated evidence which is getting novel grounds thanks h current advances in neurobiological technical advances.
-Moreover, we have also shown that this community as highly diverse, ranging from in vivo, he in computo and in silico.
-As a consequence the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
+\begin{itemize}
+  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
+  \item  that growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motifs,
+  \item  that many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
+  \item  ultimately that using precise puking motifs could be a key ingredient in men nomorphic systems to reach similar efficiencies as biological neural systems.
+\end{itemize}
+ Overall, our reviewing effort has shown that a growing community is focusing on that aspect.
+This community is based on solid and validated evidence which is getting novel grounds thanks to current advances in technical advances.
+Moreover, we have also shown that this community is highly diverse, ranging from \emph{in vivo}, \emph{in computo} and \emph{in silico}.
+As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
 
 \subsection{Limits}
-
 The different models of spike motif detection and learning that we have presented at different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
@@
 First of all, many models are based on a discretization of time.
-This assumption is important to allow a correct representation of the neuronal information in order to be processed in computers.
-This treatment amounts to transforming the Spike Raster into a matrix form for processing in classical machine learning algorithms.
-This assumption implies an ineffective use of the memory because this representation transforms the sparse representation of a spike sequence into dense matrices.
+This assumption is important to allow for a useful representation of neuronal information in order to be processed in computers.
+This treatment amounts to transforming spike rasters into a matrix form for processing in classical machine learning algorithms.
+This assumption therefore implies an ineffective use of the memory as this representation transforms the sparse representation of a spike sequence into dense matrices.
 In addition, this representation can induce errors due to the discretization and the scale of temporal sampling.
@@
 A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output.
-However, we have seen that the neurobiological system uses processing loops.
+However, we have seen that the neurobiological system uses processing loops within hierarchical graphs.
 

### 2022-11-09 manuscript.tex
6

@@
 
-The limits that we have
-presented can be treated individually as evidenced by individual efforts which try he orecome them.
-However, to propose a real breakthrough, we believe that future venues should provide with a unified, inter-disciplinary approach, with applications to deal- word, ecological scenarios and with open and reproducible methods.
-First, as we already noted, the effort is still largely scattered -this is in part due to the fact that interaction between neurobiology, theoretical and computational neuroscience and meromorphic engineering are still scarce as of today.
-It was largely demonstrated that close, bi-directional interactions are essential to foster breakthrough For instance, the design of model- driven protocols as proven to be essential in modern neuroscience.
+The limits that we have presented can be treated individually in each model as evidenced by individual efforts which try he orecome them.
+However, to propose a real breakthrough, we believe that future venues should provide with a unified, inter-disciplinary approach, with applications to real-word, ecological scenarios and with open and reproducible methods.
+
+First, as we already noted, the effort is still largely scattered.
+This is in part due to the fact that interaction between neurobiology, theoretical and computational neuroscience and neuromorphic engineering are still scarce as of today.
+It was largely demonstrated that close, bi-directional interactions are essential to foster breakthroughs.
+For instance, the design of model-driven protocols has proven to be essential in modern neuroscience.
 Also, if neural networks were essential in shaping modern-day machine learning, e-g.
@@
 
-In that perspective, it is essential that such models are tested on ecologically- relevant, real-world scenarios.
-Indeed, classic convolutional neural networks emerged as optimal solutions for instance he classify static images into categories, yet thay are not well adapted for processing dynamic, multimodal sensory flows.
-The emerging necessity to be able to process more complex flows, such as the multiples flows of information in a car designed for autonomous driving, necessitates to modify modelling paradigms.
-and in particular to take into account that the generated actions may modify the sensory inputs.
-Notably, the protocols used as well in neurobiology, theory on engineering should take into account this hovel complexity.
+In that perspective, it is essential that such models are tested on ecologically-relevant, real-world scenarios.
+Indeed, classic convolutional neural networks emerged as optimal solutions for instance to classify static images into categories, yet thay are not well adapted for processing dynamic, multimodal sensory flows.
+The emerging necessity to be able to process more complex flows, such as the multiples flows of information in a car designed for autonomous driving, necessitates to modify such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs.
+Notably, the protocols used as well in neurobiology, theory on engineering should take into account this hovel levels of complexity.
 
@@
 Indeed, the different models that we have displayed often come with the tools necessary he reproduce the results obtained.
-This is true in neurobiology (Belles), in theoretical neuroscience (Ishi) or in engineering (PAMI).
-This aspect is essential to foster the emergence of inter- disciplinary projects, such as model-driven neurobiological experiments or biologically- inspired vennomorphic engineering.
+This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}.
+This aspect is essential to foster the emergence of inter- disciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering.
 As such, this review aimed at paving the way to openly share the variety of resources and to offer an unified view on the role of precise spiking motifs in neuroscience.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-11-09 manuscript.tex
6

@@
   \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
-  \item  that growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motifs,
-  \item  that many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
-  \item  ultimately that using precise puking motifs could be a key ingredient in men nomorphic systems to reach similar efficiencies as biological neural systems.
+  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motifs,
+  \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
+  \item  ultimately that using precise puking motifs could be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
 \end{itemize}
  Overall, our reviewing effort has shown that a growing community is focusing on that aspect.
-This community is based on solid and validated evidence which is getting novel grounds thanks to current advances in technical advances.
-Moreover, we have also shown that this community is highly diverse, ranging from \emph{in vivo}, \emph{in computo} and \emph{in silico}.
+This community is based on solid and validated evidence which is getting novel grounds thanks to current technical advances.
+Moreover, we have also shown that this community is highly diverse, operating on \emph{in vivo}, \emph{in computo} and \emph{in silico} systems.
 As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
@@
 \subsection{Limits}
-The different models of spike motif detection and learning that we have presented at different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
+Additionnally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
 
@@
 However, we can note that some models can perform such learning, but only in the case of data for which the motifs are easily separable.
+More generally, to reproduce the efficiency of biological systems, one should account for the different temporal scales of adaptation, from seconds to years.
+For instance, the scaffolding of neural assemblies seem to follow critical periods during development~\citep{dard_rapid_2022}.
 
@@
 However, we have seen that the neurobiological system uses processing loops within hierarchical graphs.
-
-%Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks \textit{in vivo}~\citep{isbister_clustering_2021} They discovered (\textit{in vivo}) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
-% dynamical models
-%Dumas and colleagues~\citep{panahi_generative_2021} : three levels / fourth paradigm~\citep{tolle_fourth_2011} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
-%Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{di_mauro_alfio_sne_2022} or of sparsity aware algorithms~\citep{yin_sata_2022}.
-
-%and oscillations? Journal Article DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?
-
-% development
-%scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2022}
-
-% todo: implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
-it may be difficult but solutions exist  ~\citep{foss_multistability_2000,perrinet_active_2014}
-
-%hierarchies
-
-%developpment
-
+In general, these systems are bidirectionnally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}.
+The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
+%
 \subsection{Perspectives}
-
+%
 The limits that we have presented can be treated individually in each model as evidenced by individual efforts which try he orecome them.
@@
 This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}.
-This aspect is essential to foster the emergence of inter- disciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering.
+This aspect is essential to foster the emergence of inter-disciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering.
+Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011} i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms.
 As such, this review aimed at paving the way to openly share the variety of resources and to offer an unified view on the role of precise spiking motifs in neuroscience.
@@
 \vspace{6pt} 
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-11-10 manuscript.tex
Update on Overleaf.

@@
 
-Let's now focus on distances defined based on a generative model formed by using precise spiking motifs.
+Let us now focus on distances defined based on a generative model formed by using precise spiking motifs.
 A review of the methods for neuronal response latency estimation exists and includes Bayesian binning~\citep{levakova_review_2015}.
@@
 They extend the measure to multiple lags~\citep{torre_synchronous_2016}.
-The core measure is based on a non-stationarity corrected parametric test statistic for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+The core measure is based on a non-stationarity corrected parametric statistical test for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
 To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
@@
 
-One of the first significant evidence of any neuronal delay in the information propagation within the animal brain came from Young's study of the squid giant axon~\cite{young_1938}:
+One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmotz's study of a frog's sciatic nerve in 1850 \citep{vonHelmotz_1850,peyrard_2020}, and was later confirmed with Young's study of the squid giant axon~\cite{young_1938}.
+Extensive measures of nerve conduction velocities were done in different animals.
+% unfinished sendence:, such as Stetson'
 
@@
 the temporal disparity between left and right cochlear loci in the owl's brain.
-However it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variations were perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
+However, it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
 The authors support a second hypothesis, that of different axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
@@
 
-Over and over, myelin thus has been identified as a main actor in the delay regulation in many biological behaviors.
+Thus in several occasions, myelin thus has been identified as a main actor in the delay regulation in many biological behaviors.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the following section, we will review theoretical models which directly take advantage of using precise spiking motifs. 
+In the following section, we review theoretical models which directly take advantage of using precise spiking motifs. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
-The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient, and try to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-This approach is quite successful, and SNNs reach in some case a similar performance as Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
+The approach which is currently most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs.
+This algorithms is the most widely used algorithm to tune the weights of a classical (non spiking) Neural Network.
+It iteratively updates the weights according to the direction and magnitude of the error function gradient.
+Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient [cite Neftci 2019] to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
+This approach is quite successful, and SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
+The main reason for adopting this approach is to benefit from dedicated neuromorphic low-energy hardware.
 However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
@@
 % polychronization
-As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on an encoding of information based on a continuously varying firing-rate.
+As we saw, most SNN and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
 A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}.
@@
 Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns.
-The neurons composing a group discharge at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
+The neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
 This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:PG}).
-According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, constitute a polychronous group.
+According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, form a polychronous group.
 Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
@@
 A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}.
-The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron which does not incorporate a spike timing framework.
-The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a STDP rule.
-The limits of this model is that its output is only binary and that its storage capacity are limited.
+The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron, which does not incorporate a spike timing framework.
+The \emph{tempotron} learning rule is derived by an optimization process and takes the form of an STDP rule.
+The limits of this model is that its output is only binary and that its storage capacities are limited.
 The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
@@
 %
-The limits that we have presented can be treated individually in each model as evidenced by individual efforts which try he orecome them.
+The limits that we have presented can be treated individually in each model as evidenced by individual efforts which try to overcome them.
 However, to propose a real breakthrough, we believe that future venues should provide with a unified, inter-disciplinary approach, with applications to real-word, ecological scenarios and with open and reproducible methods.

### 2022-11-10 manuscript.tex
Update on Overleaf.

@@
 Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges (zenke2021visualizing).
-Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware is described as event streams by an Address-Event Representation (AER).
+Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
 We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.
@@
 With the AER specification and their submillisecond temporal precision, DVS bring a new approach to the storage and use of visual information.
-From their generative model, these sparse events are markers of the dynamics of visual scene captured by the DVS.
+From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the DVS.
 The dynamics of the event streams have to be used to make sense of the recorded information and new algorithms are developed to solve classical computer vision tasks.

### 2022-11-10 manuscript.tex
Update on Overleaf.

@@
 One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmotz's study of a frog's sciatic nerve in 1850 \citep{vonHelmotz_1850,peyrard_2020}, and was later confirmed with Young's study of the squid giant axon~\cite{young_1938}.
-Extensive measures of nerve conduction velocities were done in different animals.
-% unfinished sendence:, such as Stetson'
+Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_1992}.
 
@@
 According to \citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.
-\citet{pena_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
+\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
 A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
@@
 \citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_1939} and the internode distance between Ranvier nodes \cite{brills_1977} --- instead of the axon's length as was previously thought.
+These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_axon_1939} and the internode distance between Ranvier nodes \cite{brill_conduction_1977} --- instead of solely the axon's length as was previously thought.
 
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon~\citep{schmidt_1939}
-
-Thus in several occasions, myelin thus has been identified as a main actor in the delay regulation in many biological behaviors.
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez_cerda_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}.
+Multiple sheaths of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def).
+Each sheath \citep{simons_2016}.
+
+Thus in several occasions, myelin thus has been identified as a main actor in the regulation of conduction velocity in neurons, i.e. delay regulation.
 

### 2022-11-10 manuscript.tex
advances in part.2" klmds Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022 "

@@
 In other words, spikes are on one side binary, existing or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock.
-This stands in stark contrast to the analog represnetation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks.
+This stands in stark contrast to the analog representation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks.
 As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the profusion of recorded data.
@@
 Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\cite{roy2018toward}).
-Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make computations.
+Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
 This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
@@
 Another important distinction with classical Von Neumann architectures is the localized memory of this new type of chips.
-It is materialized by the physical connections between the processing units that can store information, an example directly inspired by synaptic plasticity of such connection is the memristor (ref).
+It is materialized by capacity of the physical connections between the processing units to store information.
+An example of such connection, directly inspired by synaptic plasticity, is the memristor (ref).
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
@@
 With the AER specification and their submillisecond temporal precision, DVS bring a new approach to the storage and use of visual information.
-From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the DVS.
-The dynamics of the event streams have to be used to make sense of the recorded information and new algorithms are developed to solve classical computer vision tasks.
-Different methods represent this spatio-temporal information in order to use it efficiently.
-In~\cite{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
+The dynamics of the event streams have to be used to make sense of the recorded information and new algorithms are developed to solve efficiently classical computer vision tasks.
+In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
 An exponential decay is applied on this delay to obtain the analog values of the time surfaces, it gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
@@
 This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
-Other methods make direct use of the precise timing of events captured by the DVS to solve complicated computer vision tasks.
-
-such as estimating optical flow and time-to-contact~\cite{benosman2013event, tschechne2014bio, bardow2016simultaneous}, inferring 3D reconstruction~\cite{hidalgo2020learning, osswald2017spiking, zhu2018realtime} or solving the simultaneous localization and mapping problem~\cite{gallego2017event, kim2016real} %à voir si je reformule cette phrase empruntée du pami
-
-%We have seen that precise spiking motifs may be used in theoretical neuroscience models, themselves being used to derive novel methods for the understanding of neurobiological data.
-Interestingly some of these computational neuroscience models aim at solving the spike sorting method using STDP rules~\citep{bernert_fully_2017,bernert_attention-based_2018}
-%A third class of models coming from the field of neuromorphic engineering, a priori different from that used in neuroscience, are of interest in this review as they often take advantage of minute temporal information.
-
-%\subsection{The emergence of neuromorphic engineering}
-% intro 
-%The field of neuromorphic engineering has emerged with the need for dedicated hardware to simulate SNNs.
-
-%uses a FPGA~\citep{dahlem_dynamics_2009}
-
-%V.
-Thanasoulis, B.
-Vogginger, J.
-Partzsch and C.
-Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
-%~\citep{thanasoulis_delay-based_2021}
+Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\cite{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
+This non-exhaustive list of complex task solving are not directly linked to biological processes but show the potential of using the precise temporal resolution of retinomorphic cameras.
+By essence, these sensors offer a novel view of visual information processing and the use of spatio-temporal motifs of embedded in the event streams are essential to solve high level visual tasks.
+
+\subsection{Computations with delays in neuromorphic hardware}
+
+Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}
 
 % un exemple de polychronization
-% ~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
+~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
 This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
@@
 % d'autres applications
-%Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
+Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
 The task consists in localizing a target using only the relative spike timing of eight spatially-separated vibration sensors~\citep{haessig_event-based_2020}.
-
-\subsection{Temporal pattern detection in event-based cameras}
-% le cas spécifique des event-based cameras
-%Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}.
-For example, event-based cameras provide a stream of binary asynchronous events signaling detectable changes in luminance, and information is represented by these spike-based temporal motifs, hence their name ``silicon retinas'' (see Figure \ref{fig:silicon_retina}).
-For such devices, it is crucial to better understand the potential of using such event-based representations in order to devise novel algorithms.
 
@@
 %----------------------------%
-This section has provided evidence that polychronous groups are an important aspect of information representation in biology, with important application in data analysis and neuromorphic engineering.
-The rest of this review paper is organized as follows.
 
@@
 Spatio-temporal kernels:~\citep{yu_stsc-snn_2022}
-
-\subsection{Precise timings on neuromorphic hardware}
-
-% HOTS
-
- %~\citep{lagorce_hots_2017}
-
- %~\citep{benosman_asynchronous_2012}
-
- %~\citep{benosman_event-based_2014}
-
-  %~\citep{dardelet_event-by-event_2021}
-
-% Grimaldi CBMI + PAMI
-%~\citep{lagorce_hots_2017, grimaldi_homeostatic_2021,grimaldi_robust_2022}
-
-% Grimaldi ICIP
 

### 2022-11-10 manuscript.tex
starting delay on chip
Update on Overleaf.

@@
 Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez_cerda_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}.
-Multiple sheaths of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def).
-Each sheath \citep{simons_2016}.
-
-Thus in several occasions, myelin thus has been identified as a main actor in the regulation of conduction velocity in neurons, i.e. delay regulation.
+Myelination consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\cite{simons_2016}.
+In average, each oligodendrocyte produce 20 to 60 myelinating processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_2016}.
+Multiple sheaths of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def); this demonstrate an additional impact of myelination on the conduction velocity as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
+Thus in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e.
+axonal delay regulation.
 
@@
 
-facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}.
-Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments.
-This biological phenomenom both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
+\citet{fields_2015} states that myeline facilitates both the neural circuit function and the behavioural performance: experiments on mammals show that myelination is activity-dependant and directly related to learning and memory consolidation, especially motor training and in enriched environments.
+This biological phenomenom takes place both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
 It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
@@
 
+However, one question remains: how do oligodendrocytes detect neuronal activity and regulate the myelination accordingly?
+To answer this, we must first study the myelination process.
+The oligodendrocyte precursor cells (OPCs) first proliferate in the white matter \textit{via} a self-repursive process, thus allowing for an evenly spaced network, and identify target axons.
+Most OPCs then differentiate into oligodendrocytes and immediatly initiate myelination, with no further migration.
+
+axo-glial signaling + how do they identify the target axon
+
+Extensive studies of the glialo axo-glial synaptic communication
 Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.
+
+% myelination for axonal delay, but what about dendritic delay ? 
+The previous paragraphs present the biological mechanisms behind axonal delay.
+However dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes.
+cite all info from \citep{MaladiAsl_2018} + \citep{branco_dendritic_2010}
 
@@
 It is materialized by capacity of the physical connections between the processing units to store information.
-An example of such connection, directly inspired by synaptic plasticity, is the memristor (ref).
+An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{markovic_physics_2020}.
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
@@
 The task consists in localizing a target using only the relative spike timing of eight spatially-separated vibration sensors~\citep{haessig_event-based_2020}.
+
+\citep{renner_sparse_2022}
+
+\citep{sandamirskaya_neuromorphic_2022}
+
+\citep{hussain_deltron_2012}
+\citep{sandamirskaya_neuromorphic_2022}
+\citep{sandamirskaya_neuromorphic_2022}
+
+%https://ieeexplore.ieee.org/abstract/document/8123631
+
+%https://ieeexplore.ieee.org/abstract/document/9622153
+
+%Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160
+
+%Sparse Vector Binding on Spiking Neuromorphic Hardware Using Synaptic Delays
 

### 2022-11-11 manuscript.tex
chip review

@@
 Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\cite{roy2018toward}).
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\cite{roy_towards_2019}).
 Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
@@
 These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation for the recording of a visual scene (ref).
-Another important distinction with classical Von Neumann architectures is the localized memory of this new type of chips.
-It is materialized by capacity of the physical connections between the processing units to store information.
-An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{markovic_physics_2020}.
+Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips.
+It can be materialized by capacity of the physical connections between the processing units to store information~\citep{markovic_physics_2020}.
+An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted.
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges (zenke2021visualizing).
+This field of research is inspired by neuroscientific advances and computational formalisms to design innovative architectures and by reproducing such mechanisms it is interesting to study neural circuitry.
+Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\cite{zenke_visualizing_2021}.
 Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
 We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.
-
-%On one hand, neuromorphic engineering is inspired by biology to develop efficient algorithms and technology.
-For instance, DVS cameras are inspired by the retina, memristors are used to implement neuroplasticity and neuromorphic chips are designed to emulate neural elements in order to perform sparse and asynchronous computations as observed in the biological brain.
-Therefore, the neuromorphic field is interested in understanding neural circuitry to be able to reproduce such mechanisms.
-On the other hand, computational neuroscience uses mathematical models to describe and explain a wide range of cognitive processes.
-It focuses on biological plausibility to validate its models and in comparison with experimental data.
-As a result, both fields have a common interest in neural information processing and how these may implement the computations that happen in the brain (for a review, see~\cite{zenke2021visualizing}).
-
-\subsection{On the importance of spatio-temporal information in event-based recordings}
-
-%A time surface represents therefore a combination of an object and its motion, but both elements are not explicitly dissociated.
-This entangled information may need to be separated in order to reach optimum object categorization.
-
+We report different types of event-based sensing devices for sound~\cite{chan_aer_2007} and touch~\cite{haessig_event-based_2020} but first focus on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
+Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
+%
+\subsection{On the importance of spatio-temporal information in silicon retinas}
+%
 With the AER specification and their submillisecond temporal precision, DVS bring a new approach to the storage and use of visual information.
 From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
-The dynamics of the event streams have to be used to make sense of the recorded information and new algorithms are developed to solve efficiently classical computer vision tasks.
-In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+The dynamics of the event streams have to be used to make sense of the recorded information and new algorithms are needed to solve efficiently classical computer vision tasks.
+In~\cite{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
 An exponential decay is applied on this delay to obtain the analog values of the time surfaces, it gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
-Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs on the input signals.
-Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022}.
-Three dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022}.
+Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
+Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022}.
+Three dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
 With their additional temporal dimension, they can capture multiple events on the same pixel address as long as they belong to the local temporal window.
 This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
-Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\cite{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
-This non-exhaustive list of complex task solving are not directly linked to biological processes but show the potential of using the precise temporal resolution of retinomorphic cameras.
-By essence, these sensors offer a novel view of visual information processing and the use of spatio-temporal motifs of embedded in the event streams are essential to solve high level visual tasks.
-
+Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
+This non-exhaustive list of complex task solving is not directly linked to biological processes but show the potential of using the precise temporal resolution of retinomorphic cameras.
+By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
+With this type of signal, the use of spatio-temporal motifs embedded in the event streams are essential to solve high level visual tasks.
+%
 \subsection{Computations with delays in neuromorphic hardware}
-
-Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{rasetto_challenges_2022}
+%
+SNNs are considered the third generation of ANNs and, also for the asynchronous nature of computations, neuromorphic engineering is promising for the future of artificial intelligence.
+Both fields of research are characterized by local asynchronous computing units and the absence of a central clock and this is why the use of the temporal dimension in these types of architectures seems crucial.
+The previous section demonstrates the importance of using spatio-temporal motifs in event-based recordings, here we focus on the implementation of such event-based algorithms in neuromorphic hardware.
+Even if some useful SNNs simulators run on GPUs (brian, etc..
+ref), such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
+
+To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\cite{mead_pulsed_1989}.
+It was implemented with transistors operating in the subthreshold domain and responded to asynchronous events but was not capable of discriminating two different spiking sequences with the same firing rate.
+Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decreased in time like it can be observed in biological neurons.
+Then, electronics circuits became more and more bio-realistics and~\cite{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the dynamics of the biological neurons.
+The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector".
+Today many devices are good candidates for implementing event-based algorithms and use the AER, they can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
+For a more complete review the reader can refer to~\cite{schuman_survey_2017}, here we site the most popular ones.
+SpiNNaker, Loihi and TrueNorth chips (ref) are widely used examples of digital hardware implementations.
+Compared to TrueNorth which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning.
+This flexibility in the implementation comes at the cost of an increased energy consumption.
+Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop (ref).
+They are mostly used by computational neuroscientist to model brain activity with different levels of abstraction.
+BrainScaleS (FPGA, etc...
+voir la revue faite pour agileneurobot).
+
+\citep{hussain_deltron_2012} reported an implementation of a delay-learning algorithm on an analog chip.
+Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\cite{gutig_tempotron_2006}.
+It
+
+\note{parler un peu des chips qui existent? This poten-
+tial advantage of spiking has motivated a number of neural
+computing architectures such as IBM’s TrueNorth [16] and
+University of Manchester’s SpiNNaker processor [17]. }
+Oui commencer par décrire le développement de ces chips avec un historique, dire qu'elles sont
+
+they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is
+well-suited for areas such as pattern recognition and natural
+language processing. In this paper, we formally describe the
+STPU, implement the STPU on a field programmable gate array,
+and show measured performance data.\cite{hill_spike-timing_2017}
 
 % un exemple de polychronization
-~\citet{wang_neuromorphic_2015} presented a hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
+~\citet{wang_neuromorphic_2015} presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
 This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
@@
 
+However, the binding of spatially sparse representations is computationally expensive because it is not local to corresponding pairs of neurons as in VSAs with dense vectors.
+Here, we present the first implementation of a sparse VSA on spiking neuromorphic hardware, specifically Intel’s neuromorphic research chip Loihi.
+To reduce the cost of binding, a delay line and coincidence detection are used, trading off space with time.
+We show as proof of principle that our network on Loihi can perform the binding operation of a classical analogical reasoning task and discuss the cost of different sparse binding operations.
+The proposed binding mechanism can be used as a building block for VSA-based architectures on neuromorphic hardware.
 \citep{renner_sparse_2022}
 
-\citep{sandamirskaya_neuromorphic_2022}
-
-\citep{hussain_deltron_2012}
-\citep{sandamirskaya_neuromorphic_2022}
-\citep{sandamirskaya_neuromorphic_2022}
-
-%https://ieeexplore.ieee.org/abstract/document/8123631
-
-%https://ieeexplore.ieee.org/abstract/document/9622153
+review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)
+
+deltron is an implementation of delay learning based on membrane potential optimization/supervised learning to detect spatio-temporal patterns in spike train.
+Only delays no weights and supervision with timing of the patterns (not sure because they take Vmax and then compare with Vthres)\citep{hussain_deltron_2012}
+
+aer ear a silicon cochlea with AER \citep{chan_aer_2007}
+
+new analog VLSI synapse circuit (the DPI of section 2.6)
+useful for implementing postsynaptic currents in neuromorphic VLSI net-
+works of spiking neurons with biologically realistic temporal dynamics.
+\citep{bartolozzi_synaptic_2007} Several subthreshold synaptic circuit designs have been proposed (Mead, 1989; Lazzaro, 1994; Boahen, 1998; Fusi et al., 2000; Chicca, Indiveri,  Douglas, 2003; Shi  Horiuchi, 2004a; Gordon, Farquhar,  Hasler, 2004; Hynna  Boahen, 2006) covering a range of trade-offs between functionality and complexity of temporal dynamics versus circuit and layout size.
+these properties in our VLSI devices because there is evidence that they play an important role in detecting coincidence between the presynaptic activity and postsynaptic depolarization for inducing long-term-potentiation (LTP) (Morris, Davis,  Butcher, 1990)
+
+We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method.
+Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware\cite{liang_neuromorphic_2021}
+
+they implement on a neuromorphic chip the detection of spikes delay up to a 50 ns precision, simulation of the auditory cortex of the barn owl\cite{pfeil_neuromorphic_2013}
 
 %Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160
-
-%Sparse Vector Binding on Spiking Neuromorphic Hardware Using Synaptic Delays
 
@@
 % solutions usuellement utilisées - surrogate gradient
-
-% autres méthodes
-Spatio-temporal kernels:~\citep{yu_stsc-snn_2022}
+for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)
 

### 2022-11-11 manuscript.tex
Update on Overleaf.

@@
 Why do neurons communicate through spikes?
-By definition, a spike is an all-or-none neural event which occurs at continuous times.
+By definition, spikes are all-or-none neural events which occur at continuous times.
 In other words, spikes are on one side binary, existing or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock.
@@
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+\keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; polychronization.)}
 
@@
 Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and possibly further propagated along its "output" axon to reach efferent neurons The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
-Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance as computed as the average occurrence of spikes in small time windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
+Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
 This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
@@
 For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it at first sight inconsistent with the temporal integration of firing rate.
-Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: whike the first spike are aligned, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
+Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: while the first spike is aligned to stimulus' time, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
 However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see~\citep{ermentrout_reliability_2008}).
@@
 Building such models for predicting changes of timings based on context, such as using a so-called change point model for blocked-based experimental protocols~\citep{pasturel_humans_2020}, could therefore provide useful prior information to enhance the decoding from neural activity.
-
 % sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{What biological mechanism could allow learning spiking motifs?}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 Despite the extensive computer evidence of the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
-The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviours allow us to put these doubts to rest.
-In the following paragraphs we will successively describe the first biological observations of biological delay learning, identify myelination as an important actor and finally study this phenomenon at the molecular level.
+The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviors allow us to put these doubts to rest.
+In the following paragraphs we will successively describe the first biological observations of biological delay learning, identify myelinization as an important actor and finally study this phenomenon at the molecular level.
 
@@
 
-One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmotz's study of a frog's sciatic nerve in 1850 \citep{vonHelmotz_1850,peyrard_2020}, and was later confirmed with Young's study of the squid giant axon~\cite{young_1938}.
+One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{vonHelmotz_1850,peyrard_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1938}.
 Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_1992}.
@@
 However, it was not until the study of the interaural time difference (ITD) by~\citet{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
-This ITD is a biological mechanism which allows for the azimutal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-According to \citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviourally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.
-\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
+This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
+According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.~\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
 A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
@@
 However, it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
-The authors support a second hypothesis, that of different axonal delays in the ipsi and contralateral cochlear nucleus magnocellularis.
-\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter \cite{gasser_axon_1939} and the internode distance between Ranvier nodes \cite{brill_conduction_1977} --- instead of solely the axon's length as was previously thought.
+The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
+These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977} --- instead of solely the axon's length as was previously thought.
 
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez_cerda_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}.
-Myelination consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\cite{simons_2016}.
-In average, each oligodendrocyte produce 20 to 60 myelinating processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_2016}.
-Multiple sheaths of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def); this demonstrate an additional impact of myelination on the conduction velocity as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
-Thus in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e.
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}.
+Myelinization consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}.
+In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
+Multiple Myelinization of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def); this demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
+Thus, in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e.
 axonal delay regulation.
@@
 
-\citet{fields_2015} states that myeline facilitates both the neural circuit function and the behavioural performance: experiments on mammals show that myelination is activity-dependant and directly related to learning and memory consolidation, especially motor training and in enriched environments.
-This biological phenomenom takes place both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
+\citet{fields_2015} states that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially motor training and in enriched environments.
+This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
+It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
 
@@
 
-However, one question remains: how do oligodendrocytes detect neuronal activity and regulate the myelination accordingly?
-To answer this, we must first study the myelination process.
-The oligodendrocyte precursor cells (OPCs) first proliferate in the white matter \textit{via} a self-repursive process, thus allowing for an evenly spaced network, and identify target axons.
-Most OPCs then differentiate into oligodendrocytes and immediatly initiate myelination, with no further migration.
+However, one question remains: how do oligodendrocytes detect neuronal activity and regulate the myelinization accordingly?
+To answer this, we must first study the myelinization process.
+The oligodendrocyte precursor cells (OPCs) first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
+Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration.
 
@@
 The previous paragraphs present the biological mechanisms behind axonal delay.
-However dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes.
-cite all info from \citep{MaladiAsl_2018} + \citep{branco_dendritic_2010}
+However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes.
+cite all info from~\citep{madadi_asl_delay-dependent_2022}% MaladiAsl_2018}
++~\citep{branco_dendritic_2010}
 
@@
 The \emph{tempotron} learning rule is derived by an optimization process and takes the form of an STDP rule.
-The limits of this model is that its output is only binary and that its storage capacities are limited.
-The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}.
+The limits of this model are that its output is only binary and that its storage capacities are limited.
+The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the modulation of spike latency emission~\citep{susi_nmnsd-spiking_2021}.
 Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
-As such, these models do not explicitly take full advantage of the representation capacity offered by polychronous groups.
+As such, these models use the neural dynamics to handle input signals with different delays but do not explicitly take full advantage of the representation capacity offered by heterogeneous delays.
 
@@
 The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns.
-Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task.
+Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task, but were prone to convergence issues.
 Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
@@
 Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\cite{roy_towards_2019}).
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\citep{roy_towards_2019}).
 Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
@@
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-This field of research is inspired by neuroscientific advances and computational formalisms to design innovative architectures and by reproducing such mechanisms it is interesting to study neural circuitry.
-Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\cite{zenke_visualizing_2021}.
+This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and by reproducing such mechanisms it is interesting to study neural circuitry.
+Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
 Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
 We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.
-We report different types of event-based sensing devices for sound~\cite{chan_aer_2007} and touch~\cite{haessig_event-based_2020} but first focus on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
+We report different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but first focus on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
 Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
@@
 %
-With the AER specification and their submillisecond temporal precision, DVS bring a new approach to the storage and use of visual information.
+With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and use of visual information.
 From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
-The dynamics of the event streams have to be used to make sense of the recorded information and new algorithms are needed to solve efficiently classical computer vision tasks.
-In~\cite{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision tasks.
+In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
 An exponential decay is applied on this delay to obtain the analog values of the time surfaces, it gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
@@
 Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022}.
-Three dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
-With their additional temporal dimension, they can capture multiple events on the same pixel address as long as they belong to the local temporal window.
+Three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
+With their additional temporal dimension, they can capture multiple events on the same pixel address, as long as they belong to the local temporal window.
 This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
-Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
-This non-exhaustive list of complex task solving is not directly linked to biological processes but show the potential of using the precise temporal resolution of retinomorphic cameras.
+Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
+This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of using the precise temporal resolution of neuromorphic retina-like cameras.
 By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
@@
 SNNs are considered the third generation of ANNs and, also for the asynchronous nature of computations, neuromorphic engineering is promising for the future of artificial intelligence.
-Both fields of research are characterized by local asynchronous computing units and the absence of a central clock and this is why the use of the temporal dimension in these types of architectures seems crucial.
+Both fields of research are characterized by local asynchronous computing units and the absence of a central clock, and this is why the use of the temporal dimension in these types of architectures seems crucial.
 The previous section demonstrates the importance of using spatio-temporal motifs in event-based recordings, here we focus on the implementation of such event-based algorithms in neuromorphic hardware.
@@
 
-To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\cite{mead_pulsed_1989}.
-It was implemented with transistors operating in the subthreshold domain and responded to asynchronous events but was not capable of discriminating two different spiking sequences with the same firing rate.
-Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decreased in time like it can be observed in biological neurons.
-Then, electronics circuits became more and more bio-realistics and~\cite{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the dynamics of the biological neurons.
+To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_pulsed_1989}.
+It was implemented with transistors operating in the sub-threshold domain and responded to asynchronous events, but was not capable of discriminating two different spiking sequences with the same firing rate.
+Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time like it can be observed in biological neurons.
+Then, electronics circuits became more and more bio-realistic and~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the dynamics of the biological neurons.
 The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector".
 Today many devices are good candidates for implementing event-based algorithms and use the AER, they can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
-For a more complete review the reader can refer to~\cite{schuman_survey_2017}, here we site the most popular ones.
+For a more complete review the reader can refer to~\citep{schuman_survey_2017}, here we site the most popular ones.
 SpiNNaker, Loihi and TrueNorth chips (ref) are widely used examples of digital hardware implementations.
@@
 \citep{hussain_deltron_2012} reported an implementation of a delay-learning algorithm on an analog chip.
-Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\cite{gutig_tempotron_2006}.
+Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}.
 It
@@
 STPU, implement the STPU on a field programmable gate array,
-and show measured performance data.\cite{hill_spike-timing_2017}
+and show measured performance data.\citep{hill_spike-timing_2017}
 
@@
 We show as proof of principle that our network on Loihi can perform the binding operation of a classical analogical reasoning task and discuss the cost of different sparse binding operations.
-The proposed binding mechanism can be used as a building block for VSA-based architectures on neuromorphic hardware.
-\citep{renner_sparse_2022}
+The proposed binding mechanism can be used as a building block for VSA-based architectures on neuromorphic hardware.~\citep{renner_sparse_2022}
 
@@
 
-aer ear a silicon cochlea with AER \citep{chan_aer_2007}
+aer ear a silicon cochlea with AER~\citep{chan_aer_2007}
 
@@
 useful for implementing postsynaptic currents in neuromorphic VLSI net-
-works of spiking neurons with biologically realistic temporal dynamics.
-\citep{bartolozzi_synaptic_2007} Several subthreshold synaptic circuit designs have been proposed (Mead, 1989; Lazzaro, 1994; Boahen, 1998; Fusi et al., 2000; Chicca, Indiveri,  Douglas, 2003; Shi  Horiuchi, 2004a; Gordon, Farquhar,  Hasler, 2004; Hynna  Boahen, 2006) covering a range of trade-offs between functionality and complexity of temporal dynamics versus circuit and layout size.
+works of spiking neurons with biologically realistic temporal dynamics.~\citep{bartolozzi_synaptic_2007} Several subthreshold synaptic circuit designs have been proposed (Mead, 1989; Lazzaro, 1994; Boahen, 1998; Fusi et al., 2000; Chicca, Indiveri,  Douglas, 2003; Shi  Horiuchi, 2004a; Gordon, Farquhar,  Hasler, 2004; Hynna  Boahen, 2006) covering a range of trade-offs between functionality and complexity of temporal dynamics versus circuit and layout size.
 these properties in our VLSI devices because there is evidence that they play an important role in detecting coincidence between the presynaptic activity and postsynaptic depolarization for inducing long-term-potentiation (LTP) (Morris, Davis,  Butcher, 1990)
@@
 We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method.
-Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware\cite{liang_neuromorphic_2021}
-
-they implement on a neuromorphic chip the detection of spikes delay up to a 50 ns precision, simulation of the auditory cortex of the barn owl\cite{pfeil_neuromorphic_2013}
+Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware~\citep{liang_neuromorphic_2022}
+
+they implement on a neuromorphic chip the detection of spikes delay up to a 50 ns precision, simulation of the auditory cortex of the barn owl\citep{pfeil_neuromorphic_2013}
 
@@
   \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
-  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motifs,
+  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, traveling waves on arbitrary spiking motifs,
   \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
@@
 \subsection{Limits}
-Additionnally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
-
-First of all, many models are based on a discretization of time.
+Additionally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
+
+First, many models are based on a discretization of time.
 This assumption is important to allow for a useful representation of neuronal information in order to be processed in computers.
 This treatment amounts to transforming spike rasters into a matrix form for processing in classical machine learning algorithms.
-This assumption therefore implies an ineffective use of the memory as this representation transforms the sparse representation of a spike sequence into dense matrices.
+This assumption therefore implies an ineffective use of the memory, as this representation transforms the sparse representation of a spike sequence into dense matrices.
 In addition, this representation can induce errors due to the discretization and the scale of temporal sampling.
@@
 However, we have seen that the neurobiological system uses processing loops within hierarchical graphs.
-In general, these systems are bidirectionnally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}.
+In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}.
 The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
@@
 %
-The limits that we have presented can be treated individually in each model as evidenced by individual efforts which try to overcome them.
-However, to propose a real breakthrough, we believe that future venues should provide with a unified, inter-disciplinary approach, with applications to real-word, ecological scenarios and with open and reproducible methods.
+The limits that we have presented can be treated individually in each model, as evidenced by individual efforts which try to overcome them.
+However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.
 
@@
 This is in part due to the fact that interaction between neurobiology, theoretical and computational neuroscience and neuromorphic engineering are still scarce as of today.
-It was largely demonstrated that close, bi-directional interactions are essential to foster breakthroughs.
+It was largely demonstrated that close, bidirectional interactions are essential to foster breakthroughs.
 For instance, the design of model-driven protocols has proven to be essential in modern neuroscience.
@@
 In that perspective, it is essential that such models are tested on ecologically-relevant, real-world scenarios.
-Indeed, classic convolutional neural networks emerged as optimal solutions for instance to classify static images into categories, yet thay are not well adapted for processing dynamic, multimodal sensory flows.
+Indeed, classic convolutional neural networks emerged as optimal solutions for instance to classify static images into categories, yet they are not well adapted for processing dynamic, multimodal sensory flows.
 The emerging necessity to be able to process more complex flows, such as the multiples flows of information in a car designed for autonomous driving, necessitates to modify such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs.
@@
 Ultimately, the community should encourage the adoption of open, reproducible science.
-Indeed, the different models that we have displayed often come with the tools necessary he reproduce the results obtained.
+Indeed, the different models that we have displayed often come with the tools necessary he reproduces the results obtained.
 This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}.
-This aspect is essential to foster the emergence of inter-disciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering.
+This aspect is essential to foster the emergence of interdisciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering.
 Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011} i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms.
-As such, this review aimed at paving the way to openly share the variety of resources and to offer an unified view on the role of precise spiking motifs in neuroscience.
+As such, this review aimed at paving the way to openly share the variety of resources and to offer a unified view on the role of precise spiking motifs in neuroscience.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\dataavailability{
-%This works is made reproducible using the following tools.
-First the code reproducing all figures is available on \href{https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Readme.md}{GitHub}~\citep{Jeremie2022a}.
-
-%\noindent The paper is available a an \href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
-Find the associated \noindent \href{https://www.zotero.org/groups/4562620/polychronies}{zotero group} used to regroup relevant literature on the subject.
-
-}
+\dataavailability{This works is made reproducible using the following tools.
+First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/XXX}{GitHub}~\citep{Grimaldi2022review}.
+The paper is available as an \href{https://arxiv.org/abs/YYYY.XXX}{arXiv preprint} with links to previous versions.
+Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{zotero group} which was used to regroup relevant literature on the subject.}
 
@@
 As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper~\cite{ref-journal}.
-Now citing a book reference~\cite{ref-book1,ref-book2} or other reference types~\cite{ref-unpublish,ref-communication,ref-proceeding}.
+Citing a journal paper~\citep{ref-journal}.
+Now citing a book reference~\citep{ref-book1,ref-book2} or other reference types~\citep{ref-unpublish,ref-communication,ref-proceeding}.
 Please use the command~\citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.

### 2022-11-11 manuscript.tex
appendix

@@
 
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results''.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-\sampleavailability{Samples of the compounds ... are available from the authors.}
-
-%% Only for journal Encyclopedia
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
-
-\abbreviations{Abbreviations}{
-The following abbreviations are used in this manuscript:\\
-
-\noindent 
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & Linear dichroism
-\end{tabular}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendixstart
-\appendix
-\section[\appendixname~\thesection]{}
-\subsection[\appendixname~\thesubsection]{}
-The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
+\conflictsofinterest{The authors declare no conflict of interest.
+The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% %% Optional
+% \sampleavailability{Samples of the compounds ... are available from the authors.}
+
+% %% Only for journal Encyclopedia
+% %\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+% \abbreviations{Abbreviations}{
+% The following abbreviations are used in this manuscript:\\
+
+% \noindent 
+% \begin{tabular}{@{}ll}
+% MDPI & Multidisciplinary Digital Publishing Institute\\
+% DOAJ & Directory of open access journals\\
+% TLA & Three letter acronym\\
+% LD & Linear dichroism
+% \end{tabular}
+% }
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% %% Optional
+% \appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+% \appendixstart
+% \appendix
+% \section[\appendixname~\thesection]{}
+% \subsection[\appendixname~\thesubsection]{}
+% The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
 Mathematical proofs of results not central to the paper can be added as an appendix.
 
-\begin{table}[H] 
-\caption{This is a table caption.\label{tab5}}
-\newcolumntype{C}{>{\centering\arraybackslash}X}
-\begin{tabularx}{\textwidth}{CCC}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-Entry 1		& Data			& Data\\
-Entry 2		& Data			& Data\\
-\bottomrule
-\end{tabularx}
-\end{table}
-
-\section[\appendixname~\thesection]{}
-All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
+% \begin{table}[H] 
+% \caption{This is a table caption.\label{tab5}}
+% \newcolumntype{C}{>{\centering\arraybackslash}X}
+% \begin{tabularx}{\textwidth}{CCC}
+% \toprule
+% \textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+% \midrule
+% Entry 1		& Data			& Data\\
+% Entry 2		& Data			& Data\\
+% \bottomrule
+% \end{tabularx}
+% \end{table}
+
+% \section[\appendixname~\thesection]{}
+% All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
 

### 2022-11-11 manuscript.tex
Update on Overleaf.

@@
 
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+\funding{
+This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}'').
+% Agence Nationale de la Recherche %HL and
+LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
+% LP received funding 
+and LP from the Amidex project % grant number N° ANR-22-XXX-XXX 
+``\href{PolyChronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
+Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
 Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

### 2022-11-11 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+\authorcontributions{Conceptualization, LP; writing---original draft preparation, LP.; writing---review and editing, all authors; supervision, LP; funding acquisition, Y.Y.
 All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
@@
 % LP received funding 
-and LP from the Amidex project % grant number N° ANR-22-XXX-XXX 
-``\href{PolyChronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
-Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+and LP from A*Midex grant number AMX-21-RID-025
+``\href{PolyChronies}{https://laurentperrinet.github.io/grant/polychronies/}''.}
 

### 2022-11-11 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization, LP; writing---original draft preparation, LP.; writing---review and editing, all authors; supervision, LP; funding acquisition, Y.Y.
+\authorcontributions{Conceptualization, LP; writing---original draft preparation, LP.; writing---review and editing, all authors; supervision, LP; funding acquisition, LP.
 All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.

### 2022-11-11 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization, LP; writing---original draft preparation, LP.; writing---review and editing, all authors; supervision, LP; funding acquisition, LP.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work~reported.}
+\authorcontributions{Conceptualization, LP; writing---original draft preparation, LP.; writing---review and editing, all authors; supervision, LP; funding acquisition, LP and JM.
+All authors have read and agreed to the published version of the manuscript.}
 
@@
 
-\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
+%\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
 You might choose to exclude this statement if the study did not require ethical approval.
@@
 
-\informedconsent{Any research article describing a study involving humans should contain this statement.
+%\informedconsent{Any research article describing a study involving humans should contain this statement.
 Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
@@
 
-Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
+%Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
 Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

### 2022-11-11 manuscript.tex
figure from https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles.git

@@
 \centering
-\includegraphics[width=.7\textwidth]{figures/MainenSejnowski1995.png} % TODO replace by the simulation
+\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
 \caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
+The stimulus is shown in the top row, while repetitions of a simulated Leaky Integrate-and-Fire neuron are shown in the second row (with membrane potential represented by  blue color when low hot yellow colors when depolarized).
 While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.

### 2022-11-11 manuscript.tex
synfire chain

--- 2022-11-11 manuscript.tex
@@
 \centering
-\includegraphics[width=.7\textwidth]{figures/MainenSejnowski1995.png} % TODO replace by the simulation
+\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
 \caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
+The stimulus is shown in the top row, while trial repetitions of a simulated Leaky Integrate-and-Fire neuron are shown in the second row (with membrane potential represented by  dark blue color when light yellow colors when depolarized).
 While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
@@
 \centering
-\includegraphics[width=.4\textwidth]{figures/Diesmann_et_al_1999.png}
+\includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
 \caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.

### 2022-11-11 figures/Diesmann_et_al_1999.py
synfire chain

@@ -1,0 +1,75 @@
+#!/usr/bin/env python
+"""
+Synfire chains
+--------------
+M. Diesmann et al. (1999). Stable propagation of synchronous spiking in cortical
+neural networks. Nature 402, 529-533.
+"""
+from brian2 import *
+
+duration = 100*ms
+
+# Neuron model parameters
+Vr = -70*mV
+Vt = -55*mV
+taum = 10*ms
+taupsp = 0.325*ms
+# weight = 4.86*mV
+weight = 4.10*mV
+# Neuron model
+eqs = Equations('''
+dV/dt = (-(V-Vr)+x)*(1./taum) : volt
+dx/dt = (-x+y)*(1./taupsp) : volt
+dy/dt = -y*(1./taupsp)+25.27*mV/ms+
+        (39.24*mV/ms**0.5)*xi : volt
+''')
+
+# Neuron groups
+n_groups = 10
+group_size = 100
+P = NeuronGroup(N=n_groups*group_size, model=eqs,
+                threshold='V>Vt', reset='V=Vr', refractory=1*ms,
+                method='euler')
+
+N_inputs = 100
+N_pulses = 5
+spike_times = np.zeros(0)
+np.random.seed(42)
+for i_pulse in range(N_pulses):
+        pulse = .5 * (N_pulses-i_pulse) * np.random.randn(N_inputs) + 10 + i_pulse*20
+        spike_times = np.concatenate((spike_times, pulse))
+
+spike_times = np.mod(spike_times, 100)
+
+Pinput = SpikeGeneratorGroup(N_inputs*N_pulses, 
+                            np.arange(N_inputs*N_pulses), 
+                            spike_times*1*ms)
+                             
+                             
+# The network structure
+S = Synapses(P, P, on_pre='y+=weight')
+S.connect(j='k for k in range((int(i/group_size)+1)*group_size, (int(i/group_size)+2)*group_size) '
+            'if i<N_pre-group_size')
+Sinput = Synapses(Pinput, P[:group_size], on_pre='y+=weight')
+Sinput.connect()
+
+# Record the spikes
+Mgp = SpikeMonitor(P)
+Minput = SpikeMonitor(Pinput)
+# Setup the network, and run it
+P.V = 'Vr + rand() * (Vt - Vr)'
+run(duration)
+
+# figure(figsize=(13, 8))
+figure(figsize=(13, 5))
[…]

### 2022-11-11 manuscript.tex
synfire chain

@@
 \caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
-The model consists of 10 groups of 100 neurons each.
-A pulse with a given jitter is generated in the first group, which here generates a pulse after a certain processing delay in the second group with a lower jitter.
+The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
+Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreaseses from $5$ to $1$~\ms).
+This generates a pulse after a certain processing delay in the second group with a diffrent jitter.
+While the first two pulses progressively vanish, from the third input pulse, it is propagated to the following groups.
 This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}

### 2022-11-11 manuscript.tex
figure pathways

@@
 \centering
-\note{get the rights or redo a schematics}
-\includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}
-\caption{From~\citep{thorpe_seeking_2001} }\label{fig:thorpe}
+\includegraphics[width=.7\textwidth]{figures/visual-latency.pdf}
+% thorpe
+% \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
+\caption{Latency of the different processing steps along the visual pathway in order of activation from a visual stimulus.
+Different areas are schematically represented by ellipses, along with their activation latency.
+In order, Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger mescles, $180-260~\ms$.Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
+The latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger mescles, $180-260~\ms$.}\label{fig:thorpe}
 \end{figure}

### 2022-11-13 manuscript.tex
Update on Overleaf.

@@
 Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}.
-Myelinization consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}.
+Myelinisation consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}.
+Multiple myelin regions can appear on one neuron axon and form the following subdomains: "the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the ∼1 μm gap between adjacent myelin internodes [allowing for] the saltatory conduction)"~\citep{duncan_2021}.
 In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
-Multiple Myelinization of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def); this demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
+This demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
 Thus, in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e.
@@
 \citet{fields_2015} states that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially motor training and in enriched environments.
-This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.
+This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
+Myelination helps memory consolidation by coupling the activity of distant neuron populations, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_2020}, contextual fear conditioning~\citep{Pan_2020} or OPCs knock-out~\citep{wang_2020}.
+Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_2015}.
+
+It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_2015}.
 
@@
 The oligodendrocyte precursor cells (OPCs) first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
-Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration.
-
-axo-glial signaling + how do they identify the target axon
-
-Extensive studies of the glialo axo-glial synaptic communication
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.
+Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
+Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_2019}, such as $Ca^{2+}$ activity~\citep{baraban_2018} or the neuregulin 1 growth factor~\citep{nave_2006}.
+The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity.
+Indeed, \citet{cullen_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
+They also confirmed that the delay correlates with the level of skill acquisition.
+\citet{gibson_2014} suggests that neuronaly activity does not solely promote adaptive mylination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_2015}.
+
+A potential scenario for selective myelination on electrically active axons using non synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
+This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins.
 
@@
 Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\citep{roy_towards_2019}).
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence \citep{roy_towards_2019}.
 Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
@@
 For a more complete review the reader can refer to~\citep{schuman_survey_2017}, here we site the most popular ones.
-SpiNNaker, Loihi and TrueNorth chips (ref) are widely used examples of digital hardware implementations.
+SpiNNaker~\citep{furber_overview_2013, furber_spinnaker_2020}, Loihi~\citep{davies_loihi_2018} and TrueNorth~\citep{merolla_million_2014} chips are widely used examples of digital hardware implementations.
 Compared to TrueNorth which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning.

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 
-One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{vonHelmotz_1850,peyrard_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1938}.
-Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_1992}.
+One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1938}.
+Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{aslmadadi_2018}.
+Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_effects_1992}.
 
@@
 
-\citet{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
+\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}.
-Myelinisation consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}.
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
+Myelinization consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}.
 Multiple myelin regions can appear on one neuron axon and form the following subdomains: "the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the ∼1 μm gap between adjacent myelin internodes [allowing for] the saltatory conduction)"~\citep{duncan_2021}.
@@
 
-\citet{fields_2015} states that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially motor training and in enriched environments.
+\citet{fields_2015,fields_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments.
 This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-Myelination helps memory consolidation by coupling the activity of distant neuron populations, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_2020}, contextual fear conditioning~\citep{Pan_2020} or OPCs knock-out~\citep{wang_2020}.
-Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_2015}.
-
-It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_2015}.
+Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{MaladiAsl_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_2020}, contextual fear conditioning~\citep{Pan_2020} or OPCs knock-out~\citep{wan_2020}.
+Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_2021}.
+
+It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{MaladiAsl_2018}.
+Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_2021}.
+~\citet{duncan_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".
 
@@
 They also confirmed that the delay correlates with the level of skill acquisition.
-\citet{gibson_2014} suggests that neuronaly activity does not solely promote adaptive mylination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
+\citet{gibson_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
 Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_2015}.
@@
 A potential scenario for selective myelination on electrically active axons using non synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
-This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins.
+This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_2015}.
 
 % myelination for axonal delay, but what about dendritic delay ? 
-The previous paragraphs present the biological mechanisms behind axonal delay.
-However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes.
-cite all info from~\citep{madadi_asl_delay-dependent_2022}% MaladiAsl_2018}
-+~\citep{branco_dendritic_2010}
+The previous paragraphs present the biological mechanisms behind axonal delay, regulated by myelination.
+However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{MadadiAsl_2018}.
+Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_2018}.
+\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output".
+The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.
+
+All in all, learning spike motifs requires significantly complex pathways and biological mechanisms whether in the dendrites or the axon of the neuron.
+More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data can only help in the development of computer-based learning mechanisms of precise spiking motifs in neuromorphic data.
 
@@
 It iteratively updates the weights according to the direction and magnitude of the error function gradient.
-Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient [cite Neftci 2019] to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
+Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
 This approach is quite successful, and SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.

### 2022-11-14 manuscript.tex
THC figures

--- 2022-11-11 manuscript.tex
@@
 %----------------------------%
-\begin{figure}
+%
+\begin{figure}%[t!]
   \centering
-\note{insert fig 2 of izhikevich 2006 - or an open-source reproduction like that from Camille ? }
-%includegraphics{figures/event_driven_computations.png}
-\caption{blah.}\label{fig:PG}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
+    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
+    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
+    {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
+    {\bf (c)}~We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
+    Note that each afferent may be connected with multiple weights at different delays.
+The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
+%These PG motifs were efficiently learned by a logistic regression.
+    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
+    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
+     }
+  \label{fig:model}
 \end{figure}
+% 
 %----------------------------%

### 2022-11-14 manuscript.tex
THC figures

@@
 \funding{
-This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}'').
+This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
 % Agence Nationale de la Recherche %HL and
-LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
+LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}''
 % LP received funding 
 and LP from A*Midex grant number AMX-21-RID-025
-``\href{PolyChronies}{https://laurentperrinet.github.io/grant/polychronies/}''.}
+``\href{https://laurentperrinet.github.io/grant/polychronies/}{Polychronies}''.}
 

### 2022-11-14 manuscript.tex
correcting compilation mistakes

@@
 Myelinization consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}.
-Multiple myelin regions can appear on one neuron axon and form the following subdomains: "the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the ∼1 μm gap between adjacent myelin internodes [allowing for] the saltatory conduction)"~\citep{duncan_2021}.
+Multiple myelin regions can appear on one neuron axon and form the following subdomains: "the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the $\backsim$1 $\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)"~\citep{duncan_2021}.
 In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.

### 2022-11-14 manuscript.tex
ref update

@@
 This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
-These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation for the recording of a visual scene (ref).
+These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene.
 Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips.
-It can be materialized by capacity of the physical connections between the processing units to store information~\citep{markovic_physics_2020}.
+It can be materialized by the capacity of the physical connections between the processing units to store information~\citep{markovic_physics_2020}.
 An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted.
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and by reproducing such mechanisms it is interesting to study neural circuitry.
+This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by reproducing such mechanisms, it is interesting to study neural circuitry.
 Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
@@
 We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.
-We report different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but first focus on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
+There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but first we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
 Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
@@
 In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
-An exponential decay is applied on this delay to obtain the analog values of the time surfaces, it gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
+An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
+It gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
 Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
@@
 Three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
-With their additional temporal dimension, they can capture multiple events on the same pixel address, as long as they belong to the local temporal window.
+With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window.
 This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
 Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
-This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of using the precise temporal resolution of neuromorphic retina-like cameras.
+This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of the precise temporal resolution of neuromorphic retina-like cameras.
 By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
@@
 The previous section demonstrates the importance of using spatio-temporal motifs in event-based recordings, here we focus on the implementation of such event-based algorithms in neuromorphic hardware.
-Even if some useful SNNs simulators run on GPUs (brian, etc..
-ref), such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
-
-To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_pulsed_1989}.
+Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2001, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
+
+To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_analog_1989}.
 It was implemented with transistors operating in the sub-threshold domain and responded to asynchronous events, but was not capable of discriminating two different spiking sequences with the same firing rate.
-Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time like it can be observed in biological neurons.
-Then, electronics circuits became more and more bio-realistic and~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the dynamics of the biological neurons.
+Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, like it can be observed in biological neurons.
+Then, electronics circuits became more and more bio-realistic and, two decades later, ~\cite{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons.
 The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector".

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=.61803\textwidth]{figures/visual-latency.pdf}
+% thorpe
+% \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
+\caption{{\bf Latency of the different processing steps along the visual pathway.} Though the visual system is a highly inter-connected and complex system, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
+Different areas are schematically represented by ellipses and arrows denotes the fastest feed-forward activation, ordered with respect to their activation latency.
+In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-80~\ms$).
+Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $100-120~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to to finger muscles, with latencies about $200-350~\ms$.
+Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
+The latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger mescles, $180-260~\ms$.}\label{fig:thorpe}
+\end{figure}
+%----------------------------%
 In particular, there have been some remarkable findings when studying the dynamics of vision.
@@
 As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=.7\textwidth]{figures/visual-latency.pdf}
-% thorpe
-% \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
-\caption{Latency of the different processing steps along the visual pathway in order of activation from a visual stimulus.
-Different areas are schematically represented by ellipses, along with their activation latency.
-In order, Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger mescles, $180-260~\ms$.Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
-The latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger mescles, $180-260~\ms$.}\label{fig:thorpe}
-\end{figure}
-%----------------------------%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-11-14 manuscript.tex
finish hardware

--- 2022-11-14 manuscript.tex
@@
 This flexibility in the implementation comes at the cost of an increased energy consumption.
-Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop (ref).
+Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2018}.
 They are mostly used by computational neuroscientist to model brain activity with different levels of abstraction.
-BrainScaleS (FPGA, etc...
-voir la revue faite pour agileneurobot).
+BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introduing_2011}.
+It is a wafer-scale neuromorphic hardware with analog components.
+Analog arrays (i.e.
+field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices.
+The pulsed current source synapse and the DPI are examples of such devices, we can also mention the field programmable neural array~\cite{farquhar_field_2006} and the NeuroFPAA~\cite{liu_fpaa_2009} specifically designed for neuromorphic systems.
+Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.
 

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 \centering
-\includegraphics[width=.61803\textwidth]{figures/visual-latency.pdf}
+\includegraphics[width=.75\textwidth]{figures/visual-latency.pdf}
 % thorpe
 % \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
-\caption{{\bf Latency of the different processing steps along the visual pathway.} Though the visual system is a highly inter-connected and complex system, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
+\caption{{\bf Latency of the different processing steps along the visual pathway.} Though this system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
 Different areas are schematically represented by ellipses and arrows denotes the fastest feed-forward activation, ordered with respect to their activation latency.
 In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-80~\ms$).
-Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $100-120~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to to finger muscles, with latencies about $200-350~\ms$.
-Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
-The latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger mescles, $180-260~\ms$.}\label{fig:thorpe}
+Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $100-120~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to finger muscles, with latencies about $200-350~\ms$.}\label{fig:thorpe}
 \end{figure}
@@
 At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
+Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
+The latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.
 As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 \newcommand{\orcidauthorLP}{0000-0002-3107-4788}
+\newcommand{\orcidauthorCB}{0000-0002-0054-2729}
 
@@
 Amélie Gruel $^{2}$\orcidauthorAmG{}, % $^{2,\dagger,\ddagger}$,
-Camille Besnainou $^{1}$,
+Camille Besnainou $^{1}$\orcidauthorCB{},
 Jean Martinet $^{2}$\orcidauthorJN{},  
@@
 
-One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1938}.
-Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{aslmadadi_2018}.
+One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1939}.
+Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{madadi_asl_dendritic_2018}.
 Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_effects_1992}.
@@
 
-\citet{fields_2015,fields_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments.
+\citet{fields_new_2015, fields_myelin_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments.
 This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{MaladiAsl_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_2020}, contextual fear conditioning~\citep{Pan_2020} or OPCs knock-out~\citep{wan_2020}.
-Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_2021}.
-
-It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{MaladiAsl_2018}.
+Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{madadi_asl_dendritic_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_disruption_2020}, contextual fear conditioning~\citep{pan_preservation_2020} or OPCs knock-out~\citep{wan_impaired_2020}.
+Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_new_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_neuron-oligodendrocyte_2021}.
+
+It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_new_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{madadi_asl_dendritic_2018}.
 Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_2021}.
-~\citet{duncan_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".
+~\citet{duncan_neuron-oligodendrocyte_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".
 

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_new_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{madadi_asl_dendritic_2018}.
-Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_2021}.
+Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_demyelination_2021}.
 ~\citet{duncan_neuron-oligodendrocyte_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".
@@
 Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
-Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_2019}, such as $Ca^{2+}$ activity~\citep{baraban_2018} or the neuregulin 1 growth factor~\citep{nave_2006}.
+Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as $Ca^{2+}$ activity~\citep{baraban_ca2_2018} or the neuregulin 1 growth factor~\citep{nave_axonal_2006}.
 The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity.
-Indeed, \citet{cullen_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
+Indeed, \citet{cullen_periaxonal_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
 They also confirmed that the delay correlates with the level of skill acquisition.
-\citet{gibson_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_2015}.
+\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
 
 A potential scenario for selective myelination on electrically active axons using non synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
-This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_2015}.
+This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_new_2015}.
 
@@
 The previous paragraphs present the biological mechanisms behind axonal delay, regulated by myelination.
-However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{MadadiAsl_2018}.
-Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_2018}.
+However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
+Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.
 \citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output".

### 2022-11-14 manuscript.tex
Figure 1

@@
 In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-80~\ms$).
-Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $100-120~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to finger muscles, with latencies about $200-350~\ms$.}\label{fig:thorpe}
+Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $120-140~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to trigger finger muscles, with latencies about $200-350~\ms$.}\label{fig:thorpe}
 \end{figure}

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorAnG}{0000-0002-9536-010X} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0002-9536-010X} % AmG % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorAmG}{0000-0003-3916-0514}
-\newcommand{\orcidauthorJN}{0000-0001-8821-5556}
-\newcommand{\orcidauthorLP}{0000-0002-3107-4788}
-\newcommand{\orcidauthorCB}{0000-0002-0054-2729}
+\newcommand{\orcidauthorB}{0000-0003-3916-0514} % AmG
+\newcommand{\orcidauthorC}{0000-0002-0054-2729} % CB
+\newcommand{\orcidauthorD}{0000-0001-8821-5556} % JM
+\newcommand{\orcidauthorE}{0000-0002-3107-4788} % LP
 
@@
 Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
-\orcidauthorAnG{}, 
-Amélie Gruel $^{2}$\orcidauthorAmG{}, % $^{2,\dagger,\ddagger}$,
-Camille Besnainou $^{1}$\orcidauthorCB{},
-Jean Martinet $^{2}$\orcidauthorJN{},  
-and Laurent Perrinet $^{1,}$\orcidauthorLP{}*}
+\orcidA{}, 
+Amélie Gruel $^{2}$\orcidB{}, % $^{2,\dagger,\ddagger}$,
+Camille Besnainou $^{1}$\orcidC{},
+Jean Martinet $^{2}$\orcidD{},
+and Laurent Perrinet $^{1,}$\orcidE{}*}
 
@@
 % Keywords
-\keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; polychronization.)}
+\keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; spiking motifs; polychronization.)}
 
@@
 \caption{{\bf Latency of the different processing steps along the visual pathway.} Though this system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
-Different areas are schematically represented by ellipses and arrows denotes the fastest feed-forward activation, ordered with respect to their activation latency.
+Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency.
 In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-80~\ms$).

### 2022-11-14 manuscript.tex
Figure replicating_MainenSejnowski1995

--- 2022-11-14 manuscript.tex
@@
 %----------------------------%
+% C_MainenSejnowski1995_Perrinet.ipynb
 \begin{figure}
@@
 \caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
-The stimulus is shown in the top row, while trial repetitions of a simulated Leaky Integrate-and-Fire neuron are shown in the second row (with membrane potential represented by  dark blue color when light yellow colors when depolarized).
+The stimulus is shown in the top row, while trial repetitions of a simulated Leaky Integrate-and-Fire neuron are shown in the second row (with membrane potential represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row).
 While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
-See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
+See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
 \end{figure}

### 2022-11-14 figures/Diesmann_et_al_1999.py

@@ -66,10 +66,12 @@ plot(Mgp.t/ms, 1.0*Mgp.i/group_size, '.')
 plot([0, duration/ms], np.arange(n_groups).repeat(2).reshape(-1, 2).T, 'k-')
 ylabel('group number')
 yticks(np.arange(n_groups)+.5, labels=range(1, n_groups+1))
-xlabel('time (ms)')
-# show()
 
-fname = 'Diesmann_et_al_1999'
-import matplotlib.pyplot as plt
-plt.savefig(fname + '.pdf', bbox_inches='tight', pad_inches=0.0, backend='pdf')
-#plt.savefig(fname + '.png', bbox_inches='tight', pad_inches=0.0, dpi=150, backend='agg')
+xlabel('time (ms)')
+if True:
+        show()
+else:
+        fname = 'Diesmann_et_al_1999'
+        import matplotlib.pyplot as plt
+        plt.savefig(fname + '.pdf', bbox_inches='tight', pad_inches=0.0, backend='pdf')
+        #plt.savefig(fname + '.png', bbox_inches='tight', pad_inches=0.0, dpi=150, backend='agg')

### 2022-11-14 manuscript.tex
Figure Haimerl 2019

@@
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
-\caption{Fig.
+% %----------------------------%
+% \begin{figure}
+% \centering
+% \includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
+% \caption{Fig.
 1.
@@
 In (i), the blue trace is shifted --2.75 mV; in (ii), the blue trace is shifted --1.58 mV, and the green +0.79 mV.}\label{fig:Ikegaya2004}
-\end{figure}
-%----------------------------%
+% \end{figure}
+% %----------------------------%
 
@@
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
-
 %
@@
 Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse.}\label{fig:haimerl}
+The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}.).}\label{fig:haimerl}
 \end{figure}

### 2022-11-14 figures/Diesmann_et_al_1999.py

@@ -68,7 +68,7 @@ ylabel('group number')
 yticks(np.arange(n_groups)+.5, labels=range(1, n_groups+1))
 
 xlabel('time (ms)')
-if True:
+if False:
         show()
 else:
         fname = 'Diesmann_et_al_1999'

### 2022-11-14 manuscript.tex
Figure Haimerl 2019

@@
 Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}.).}\label{fig:haimerl}
+The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}
 \end{figure}

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 Although surprising, this speed of the visual cortex in primates is compatible with the latencies that are recorded at the neuro-physiological level.
-The rapid propagation of the visual information in the thalamus, then in the primary visual cortex (see Figure~\ref{fig:thorpe}) takes about $45~\ms$ in the macaque~\citep{schmolesky_signal_1998} and about $60~\ms$ in humans~\citep{vanni_coinciding_2001}.
-This functioning of visual processing as a forward pass is most prominent in fast processing, but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
-An important consequence of this speed of processing of vision is that it implies that procesing is carried out using only very few spikes per layer.
-Indeed, we may consider that a behavioral response occurring in only $120~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
+The rapid propagation of the visual information in the thalamus, then in the primary visual cortex takes about $45~\ms$ in the macaque~\citep{schmolesky_signal_1998} and about $60~\ms$ in humans~\citep{vanni_coinciding_2001}.
+This functioning of visual processing as a forward pass is most prominent in fast processing (see Figure~\ref{fig:thorpe}), and can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
+An important consequence of this speed of processing of vision is that it implies that processing is carried out using only very few spikes per layer.
+As a comparison, the latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.
+Indeed, we may consider that a behavioral response occurring in only $200~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
 Such models were indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
@@
 Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
-The latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.
 As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
@@
 However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
+
+% retina
+How may information about the sensory world be encoded in the spiking pattern?
+At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
+In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
+These results have been extended to natural images and show a qualitatively similar behavior.
+The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
+Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+This evidence found in the retina can be extended to other areas such as the visual cortex.
 %----------------------------%
@@
 %----------------------------%
-
-% retina
-How may information about the sensory world be encoded in the spiking pattern?
-At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
-In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
-These results have been extended to natural images and show a qualitatively similar behavior.
-The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
-Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
-This evidence found in the retina can be extended to other areas such as the visual cortex.
 

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 As a summary, it seems that such precise structural information is essential to the neural code and that it seems imperative to include this information in decoding algorithm for a better understanding of neural activity.
-
+%
 % TODO: re-read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
@@
 The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity.
-Indeed, \citet{cullen_periaxonal_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
-They also confirmed that the delay correlates with the level of skill acquisition.
-\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
+Indeed,~\citet{cullen_periaxonal_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
+They also confirmed that the delay correlates with the level of skill acquisition.~\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
 Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
@@
 However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
-Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.
-\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output".
+Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output".
 The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.
@@
 All in all, learning spike motifs requires significantly complex pathways and biological mechanisms whether in the dendrites or the axon of the neuron.
-More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data can only help in the development of computer-based learning mechanisms of precise spiking motifs in neuromorphic data.
-
+More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data and to help develop artificial neural networks, whose learning rules would be more neuromorphic.
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, form a polychronous group.
-Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:model}).
 In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
@@
     \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
+    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
     {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
-    {\bf (c)}~We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
+    {\bf (c)}~By defining a layer of neurons that detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
     Note that each afferent may be connected with multiple weights at different delays.
@@
 Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence \citep{roy_towards_2019}.
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019}.
 Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
@@
 We show as proof of principle that our network on Loihi can perform the binding operation of a classical analogical reasoning task and discuss the cost of different sparse binding operations.
-The proposed binding mechanism can be used as a building block for VSA-based architectures on neuromorphic hardware.~\citep{renner_sparse_2022}
+The proposed binding mechanism can be used as a building block for VSA-based architectures on neuromorphic hardware~\citep{renner_sparse_2022}.
 
@@
 This assumption is important to allow for a useful representation of neuronal information in order to be processed in computers.
-This treatment amounts to transforming spike rasters into a matrix form for processing in classical machine learning algorithms.
+This treatment amounts to transforming spike trains into a matrix form for processing in classical machine learning algorithms.
 This assumption therefore implies an ineffective use of the memory, as this representation transforms the sparse representation of a spike sequence into dense matrices.
@@
 
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+\acknowledgments{We would like to thank Jean-Nicolas Jérémie, Hugo LAdret for useful discussions while writing this review.}
 

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019}.
-Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
+Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{ref}, neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
 This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.

### 2022-11-14 manuscript.tex
Section 5

--- 2022-11-14 manuscript.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Artificial intelligence has made huge advances in the past decades, now deep learning algorithms outperform humans at complex tasks like natural image recognition or abstract strategy board games.
-Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
+Artificial intelligence has made huge advances in the past decades and deep learning algorithms nowadays outperform humans at complex tasks like natural image recognition or abstract strategy board games.
+Yet, machine learning algorithms suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
 Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019}.

### 2022-11-14 manuscript.tex
end of section 5, feel free to change the end!

--- 2022-11-14 manuscript.tex
@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0002-9536-010X} % AmG % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorAnG}{0000-0002-9536-010X} % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0003-3916-0514} % AmG
-\newcommand{\orcidauthorC}{0000-0002-0054-2729} % CB
-\newcommand{\orcidauthorD}{0000-0001-8821-5556} % JM
-\newcommand{\orcidauthorE}{0000-0002-3107-4788} % LP
+\newcommand{\orcidauthorAmG}{0000-0003-3916-0514}
+\newcommand{\orcidauthorJN}{0000-0001-8821-5556}
+\newcommand{\orcidauthorLP}{0000-0002-3107-4788}
 
@@
 Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
-\orcidA{}, 
-Amélie Gruel $^{2}$\orcidB{}, % $^{2,\dagger,\ddagger}$,
-Camille Besnainou $^{1}$\orcidC{},
-Jean Martinet $^{2}$\orcidD{},
-and Laurent Perrinet $^{1,}$\orcidE{}*}
+\orcidauthorAnG{}, 
+Amélie Gruel $^{2}$\orcidauthorAmG{}, % $^{2,\dagger,\ddagger}$,
+Camille Besnainou $^{1}$,
+Jean Martinet $^{2}$\orcidauthorJN{},  
+and Laurent Perrinet $^{1,}$\orcidauthorLP{}*}
 
@@
 % Keywords
-\keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; spiking motifs; polychronization.)}
+\keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; polychronization.)}
 
@@
 \centering
-\includegraphics[width=.75\textwidth]{figures/visual-latency.pdf}
+\includegraphics[width=.61803\textwidth]{figures/visual-latency.pdf}
 % thorpe
 % \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
-\caption{{\bf Latency of the different processing steps along the visual pathway.} Though this system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
-Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency.
+\caption{{\bf Latency of the different processing steps along the visual pathway.} Though the visual system is a highly inter-connected and complex system, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
+Different areas are schematically represented by ellipses and arrows denotes the fastest feed-forward activation, ordered with respect to their activation latency.
 In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-80~\ms$).
-Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $120-140~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to trigger finger muscles, with latencies about $200-350~\ms$.}\label{fig:thorpe}
+Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $100-120~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to to finger muscles, with latencies about $200-350~\ms$.
+Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
+The latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger mescles, $180-260~\ms$.}\label{fig:thorpe}
 \end{figure}
@@
 Although surprising, this speed of the visual cortex in primates is compatible with the latencies that are recorded at the neuro-physiological level.
-The rapid propagation of the visual information in the thalamus, then in the primary visual cortex takes about $45~\ms$ in the macaque~\citep{schmolesky_signal_1998} and about $60~\ms$ in humans~\citep{vanni_coinciding_2001}.
-This functioning of visual processing as a forward pass is most prominent in fast processing (see Figure~\ref{fig:thorpe}), and can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
-An important consequence of this speed of processing of vision is that it implies that processing is carried out using only very few spikes per layer.
-As a comparison, the latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.
-Indeed, we may consider that a behavioral response occurring in only $200~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
+The rapid propagation of the visual information in the thalamus, then in the primary visual cortex (see Figure~\ref{fig:thorpe}) takes about $45~\ms$ in the macaque~\citep{schmolesky_signal_1998} and about $60~\ms$ in humans~\citep{vanni_coinciding_2001}.
+This functioning of visual processing as a forward pass is most prominent in fast processing, but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
+An important consequence of this speed of processing of vision is that it implies that procesing is carried out using only very few spikes per layer.
+Indeed, we may consider that a behavioral response occurring in only $120~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
 Such models were indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
 At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
-Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
 As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
@@
 However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
+\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
+The stimulus is shown in the top row, while trial repetitions of a simulated Leaky Integrate-and-Fire neuron are shown in the second row (with membrane potential represented by  dark blue color when light yellow colors when depolarized).
+While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
+\end{figure}
+%----------------------------%
 
@@
 This evidence found in the retina can be extended to other areas such as the visual cortex.
-%----------------------------%
-% C_MainenSejnowski1995_Perrinet.ipynb
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
-\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
-The stimulus is shown in the top row, while trial repetitions of a simulated Leaky Integrate-and-Fire neuron are shown in the second row (with membrane potential represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row).
-While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
-See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
-\end{figure}
-%----------------------------%
 
@@
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
-% %----------------------------%
-% \begin{figure}
-% \centering
-% \includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
-% \caption{Fig.
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
+\caption{Fig.
 1.
@@
 In (i), the blue trace is shifted --2.75 mV; in (ii), the blue trace is shifted --1.58 mV, and the green +0.79 mV.}\label{fig:Ikegaya2004}
-% \end{figure}
-% %----------------------------%
+\end{figure}
+%----------------------------%
 
@@
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
+
 %
@@
 Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}
+The black line represents the speed of the mouse.}\label{fig:haimerl}
 \end{figure}
@@
 As a summary, it seems that such precise structural information is essential to the neural code and that it seems imperative to include this information in decoding algorithm for a better understanding of neural activity.
-%
+
 % TODO: re-read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
@@
 
-One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1939}.
-Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{madadi_asl_dendritic_2018}.
+One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1938}.
+Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{aslmadadi_2018}.
 Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_effects_1992}.
@@
 
-\citet{fields_new_2015, fields_myelin_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments.
+\citet{fields_2015,fields_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments.
 This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{madadi_asl_dendritic_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_disruption_2020}, contextual fear conditioning~\citep{pan_preservation_2020} or OPCs knock-out~\citep{wan_impaired_2020}.
-Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_new_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_neuron-oligodendrocyte_2021}.
-
-It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_new_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{madadi_asl_dendritic_2018}.
-Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_demyelination_2021}.
-~\citet{duncan_neuron-oligodendrocyte_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".
+Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{MaladiAsl_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_2020}, contextual fear conditioning~\citep{Pan_2020} or OPCs knock-out~\citep{wan_2020}.
+Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_2021}.
+
+It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{MaladiAsl_2018}.
+Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_2021}.
+~\citet{duncan_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".
 
@@
 Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
-Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as $Ca^{2+}$ activity~\citep{baraban_ca2_2018} or the neuregulin 1 growth factor~\citep{nave_axonal_2006}.
+Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_2019}, such as $Ca^{2+}$ activity~\citep{baraban_2018} or the neuregulin 1 growth factor~\citep{nave_2006}.
 The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity.
-Indeed,~\citet{cullen_periaxonal_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
-They also confirmed that the delay correlates with the level of skill acquisition.~\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
+Indeed, \citet{cullen_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
+They also confirmed that the delay correlates with the level of skill acquisition.
+\citet{gibson_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_2015}.
 
 A potential scenario for selective myelination on electrically active axons using non synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
-This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_new_2015}.
+This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_2015}.
 
@@
 The previous paragraphs present the biological mechanisms behind axonal delay, regulated by myelination.
-However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
-Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output".
+However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{MadadiAsl_2018}.
+Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_2018}.
+\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output".
 The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.
@@
 All in all, learning spike motifs requires significantly complex pathways and biological mechanisms whether in the dendrites or the axon of the neuron.
-More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data and to help develop artificial neural networks, whose learning rules would be more neuromorphic.
-%
+More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data can only help in the development of computer-based learning mechanisms of precise spiking motifs in neuromorphic data.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, form a polychronous group.
-Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:model}).
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
 In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
@@
     \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
+    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
     {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
-    {\bf (c)}~By defining a layer of neurons that detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
+    {\bf (c)}~We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
     Note that each afferent may be connected with multiple weights at different delays.
@@
 Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019}.
-Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{ref}, neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence \citep{roy_towards_2019}.
+Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
 This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
@@
 \subsection{On the importance of spatio-temporal information in silicon retinas}
+%
+%----------------------------%
+\begin{figure}
+\centering
+%\includegraphics{figures/event_driven_computations.png}
+\caption{A miniature, event-based ATIS sensor.
+Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).}\label{fig:silicon_retina}
+\end{figure}
+%----------------------------%
 %
@@
 
-\citep{hussain_deltron_2012} reported an implementation of a delay-learning algorithm on an analog chip.
+For the rest of this section, we report examples of implementations of event-based algorithms using precise spatio-temporal motifs on neuromorphic hardware. 
+\citep{hussain_deltron_2012} implemented a delay-learning algorithm on an analog chip.
 Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}.
-It
-
-\note{parler un peu des chips qui existent? This poten-
-tial advantage of spiking has motivated a number of neural
-computing architectures such as IBM’s TrueNorth [16] and
-University of Manchester’s SpiNNaker processor [17]. }
-Oui commencer par décrire le développement de ces chips avec un historique, dire qu'elles sont
-
-they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is
-well-suited for areas such as pattern recognition and natural
-language processing. In this paper, we formally describe the
-STPU, implement the STPU on a field programmable gate array,
-and show measured performance data.\citep{hill_spike-timing_2017}
-
-% un exemple de polychronization
-~\citet{wang_neuromorphic_2015} presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron.
-This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-
-% d'autres applications
-Event-Based Computation for Touch Localization Based on Precise Spike Timing : Here we propose a neuromorphic model inspired by the sand scorpion to explore the benefits of temporal coding, and validate it in an event-based sensory-processing task.
-The task consists in localizing a target using only the relative spike timing of eight spatially-separated vibration sensors~\citep{haessig_event-based_2020}.
-
-However, the binding of spatially sparse representations is computationally expensive because it is not local to corresponding pairs of neurons as in VSAs with dense vectors.
-Here, we present the first implementation of a sparse VSA on spiking neuromorphic hardware, specifically Intel’s neuromorphic research chip Loihi.
-To reduce the cost of binding, a delay line and coincidence detection are used, trading off space with time.
-We show as proof of principle that our network on Loihi can perform the binding operation of a classical analogical reasoning task and discuss the cost of different sparse binding operations.
-The proposed binding mechanism can be used as a building block for VSA-based architectures on neuromorphic hardware~\citep{renner_sparse_2022}.
-
-review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)
-
-deltron is an implementation of delay learning based on membrane potential optimization/supervised learning to detect spatio-temporal patterns in spike train.
-Only delays no weights and supervision with timing of the patterns (not sure because they take Vmax and then compare with Vthres)\citep{hussain_deltron_2012}
-
-aer ear a silicon cochlea with AER~\citep{chan_aer_2007}
-
-new analog VLSI synapse circuit (the DPI of section 2.6)
-useful for implementing postsynaptic currents in neuromorphic VLSI net-
-works of spiking neurons with biologically realistic temporal dynamics.~\citep{bartolozzi_synaptic_2007} Several subthreshold synaptic circuit designs have been proposed (Mead, 1989; Lazzaro, 1994; Boahen, 1998; Fusi et al., 2000; Chicca, Indiveri,  Douglas, 2003; Shi  Horiuchi, 2004a; Gordon, Farquhar,  Hasler, 2004; Hynna  Boahen, 2006) covering a range of trade-offs between functionality and complexity of temporal dynamics versus circuit and layout size.
-these properties in our VLSI devices because there is evidence that they play an important role in detecting coincidence between the presynaptic activity and postsynaptic depolarization for inducing long-term-potentiation (LTP) (Morris, Davis,  Butcher, 1990)
-
-We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method.
+In addition to delay learning, a group, in the University of West Sydney, developed a neurmorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}.
+They show that STDP and Spike Timing Dependant Delay Plasticity rules could be implemented in both a digital and an analog chip.
+Physical limitations to store the network on-chip and, from the parameters of the network and the current technology, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
+The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citet{wang_fpga_2013} .
+%This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
+\cite{pfeil_neuromorphic_2013} implemented STDP on a mixed-analog-digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
+Coherence detection on input spikes coming from two sources was obtained with a 50 $ns$ precision.
+They claim that this unsupervised learning denoise the input and compensates for variations between neural components.
+The variability of response of the analog components can be compensated by population coding for responses robust to noise and this phenomenon is also observed in biological neural networks~\citep{boerlin_spike-based_2011}.
+A recent work makes the first implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to com-
+pute the binding operation~\citep{renner_sparse_2022}.
+They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equilavent to the maximal delay.
+Note that this problem is also due the algorithm used in this paper and that the analog chips must not suffer from this type of limitations.
+
+Online on-chip learning and computations with delays are still emerging in neuromorphic engineering.
+The technical challenges linked to the development of this type of implementations and the growing interest for delay learning makes advances in this field interesting for the future of computations with precise spatio-temporal motifs.
+While improvements are still to be made, neuromorphic chips seem to be good candidate to efficiently make use of these particular features.
+
+%they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is well-suited for areas such as pattern recognition and natural language processing.
+In this paper, we formally describe the STPU, implement the STPU on a field programmable gate array, and show measured performance data.\citep{hill_spike-timing_2017} not accessible
+
+%review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)
+
+%We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method.
 Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware~\citep{liang_neuromorphic_2022}
 
-they implement on a neuromorphic chip the detection of spikes delay up to a 50 ns precision, simulation of the auditory cortex of the barn owl\citep{pfeil_neuromorphic_2013}
-
 %Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160
 
-%----------------------------%
-\begin{figure}
-\centering
-%\includegraphics{figures/event_driven_computations.png}
-\caption{A miniature, event-based ATIS sensor.
-Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
-These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).}\label{fig:silicon_retina}
-\end{figure}
-%----------------------------%
-
 % solutions usuellement utilisées - surrogate gradient
-for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)
+%for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)
 
@@
 This assumption is important to allow for a useful representation of neuronal information in order to be processed in computers.
-This treatment amounts to transforming spike trains into a matrix form for processing in classical machine learning algorithms.
+This treatment amounts to transforming spike rasters into a matrix form for processing in classical machine learning algorithms.
 This assumption therefore implies an ineffective use of the memory, as this representation transforms the sparse representation of a spike sequence into dense matrices.
@@
 
-\acknowledgments{We would like to thank Jean-Nicolas Jérémie, Hugo LAdret for useful discussions while writing this review.}
+\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
 

### 2022-11-14 manuscript.tex

--- 2022-11-14 manuscript.tex
@@
 %
+%----------------------------%
+\begin{figure}
+\centering
+%\includegraphics{figures/event_driven_computations.png}
+\caption{A miniature, event-based ATIS sensor.
+Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).}\label{fig:silicon_retina}
+\end{figure}
+%----------------------------%
+%
 With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and use of visual information.

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
   \centering
+  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}
   \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\subsection{Neuromorphic engineering and the emergence of novel computational architectures}
+%
 Artificial intelligence has made huge advances in the past decades and deep learning algorithms nowadays outperform humans at complex tasks like natural image recognition or abstract strategy board games.
@@
 Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{ref}, neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
+%
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
+\caption{A miniature, event-based ATIS sensor.
+Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
+Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
+\end{figure}
+%----------------------------%
+%
 This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
@@
 An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted.
+
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
@@
 Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
-We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.
-There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but first we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
+In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
+There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
 Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
@@
 %
-%----------------------------%
-\begin{figure}
-\centering
-%\includegraphics{figures/event_driven_computations.png}
-\caption{A miniature, event-based ATIS sensor.
-Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
-These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).}\label{fig:silicon_retina}
-\end{figure}
-%----------------------------%
-%
-With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and use of visual information.
+With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and processing of visual information.
 From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
@@
 Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
-Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022}.
-Three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
+Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022} and show that this method could be efficiently applied to state-of-the-art benchmarks.
+
+More generally, three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
 With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window.

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 The previous section demonstrates the importance of using spatio-temporal motifs in event-based recordings, here we focus on the implementation of such event-based algorithms in neuromorphic hardware.
-Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2001, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
+Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2003, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
 
@@
 This flexibility in the implementation comes at the cost of an increased energy consumption.
-Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2018}.
+Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}.
 They are mostly used by computational neuroscientist to model brain activity with different levels of abstraction.
-BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introduing_2011}.
+BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}.
 It is a wafer-scale neuromorphic hardware with analog components.

### 2022-11-14 manuscript.tex
Update on Overleaf.

@@
 Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, like it can be observed in biological neurons.
-Then, electronics circuits became more and more bio-realistic and, two decades later, ~\cite{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons.
+Then, electronics circuits became more and more bio-realistic and, two decades later, ~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons.
 The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector".
@@
 field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices.
-The pulsed current source synapse and the DPI are examples of such devices, we can also mention the field programmable neural array~\cite{farquhar_field_2006} and the NeuroFPAA~\cite{liu_fpaa_2009} specifically designed for neuromorphic systems.
+The pulsed current source synapse and the DPI are examples of such devices, we can also mention the field programmable neural array~\citep{farquhar_field_2006} and the NeuroFPAA~\citep{cheng_fpaa_2009} specifically designed for neuromorphic systems.
 Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.
@@
 Physical limitations to store the network on-chip and, from the parameters of the network and the current technology, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
-The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citet{wang_fpga_2013} .
+The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citet{wang_fpga_2013}.
 %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-\cite{pfeil_neuromorphic_2013} implemented STDP on a mixed-analog-digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
+\citep{pfeil_neuromorphic_2013} implemented STDP on a mixed-analog-digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
 Coherence detection on input spikes coming from two sources was obtained with a 50 $ns$ precision.

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption --- a major societal issue in the era of the digital economy and global warming.
-In this review paper, we provide evidence from biology and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
+In this review paper, we provide evidence from biology, theory and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
 }
@@
 First, these have the particular property of being essentially binary in their amplitude, i.e.
-they are prototypical, "\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}".
-Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and possibly further propagated along its "output" axon to reach efferent neurons The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
+they are prototypical,
+``\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}''.
+Typically, these are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.
+The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
 Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
@@
 However, it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
-The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
+The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
 These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977} --- instead of solely the axon's length as was previously thought.
@@
 Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
-Myelinization consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}.
-Multiple myelin regions can appear on one neuron axon and form the following subdomains: "the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the $\backsim$1 $\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)"~\citep{duncan_2021}.
+Myelinization consists in ``two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}.
+Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the $\backsim$1 $\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_2021}.
 In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
@@
 Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_demyelination_2021}.
-~\citet{duncan_neuron-oligodendrocyte_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".
+~\citet{duncan_neuron-oligodendrocyte_2021} states that ``the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration''.
 
@@
 However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
-Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output".
+Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
 The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.
@@
 %
+%----------------------------%
+%
+\begin{figure}%[t!]
+  \centering
+  \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
+    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
+    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
+    {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
+    {\bf (c)}~By defining a layer of neurons that detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
+    Note that each afferent may be connected with multiple weights at different delays.
+The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
+%These PG motifs were efficiently learned by a logistic regression.
+    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
+    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
+     }
+  \label{fig:model}
+\end{figure}
+% 
+%----------------------------%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------%
+%
+\begin{figure}%[t!]
+  \centering
+  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}
+    \caption{Detecting motion in an event stream with heterogeneous delays. 
+     }
+  \label{fig:model}
+\end{figure}
+% 
+%----------------------------%
+
 A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios.
@@
 Then, electronics circuits became more and more bio-realistic and, two decades later, ~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons.
-The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector".
+The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector''.
 Today many devices are good candidates for implementing event-based algorithms and use the AER, they can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
@@
 While improvements are still to be made, neuromorphic chips seem to be good candidate to efficiently make use of these particular features.
-
+%
 %they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is well-suited for areas such as pattern recognition and natural language processing.
 In this paper, we formally describe the STPU, implement the STPU on a field programmable gate array, and show measured performance data.\citep{hill_spike-timing_2017} not accessible
-
+%
 %review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)
-
+%
 %We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method.
 Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware~\citep{liang_neuromorphic_2022}
-
+%
 %Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160
-
+%
 % solutions usuellement utilisées - surrogate gradient
 %for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\acknowledgments{We would like to thank Jean-Nicolas Jérémie, Hugo LAdret for useful discussions while writing this review.}
+\acknowledgments{We would like to thank Jean-Nicolas Jérémie, Hugo Ladret for useful discussions while writing this review.}
 

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns.
-The neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:PG}).
+On core ingrredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
+This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevitch}).
 According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, form a polychronous group.
@@
   \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}
-  \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
-  \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
-  \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
-  \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
-    {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
-    {\bf (c)}~By defining a layer of neurons that detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
-    Note that each afferent may be connected with multiple weights at different delays.
-The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
-%These PG motifs were efficiently learned by a logistic regression.
-    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
-    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
+    \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In that example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
+\textit{(Middle)}~if a synchronous pulse is emitted in presynaptic neurons, this will generate post-synaptic potentials at these different delays, and these may not be sufficient to generate a spike in either neuron.
+\textit{(Right)}~If a sequence of activation is generated in the presynaptic neurons such that, taking into account the heterogeneous delays, these converge at the same time on neuron \textit{a} (here, at $t=10~\ms$) then a spike is emitted, while none is emitted on neuron \textit{e}.
      }
-  \label{fig:model}
+  \label{fig:izhikevitch}
 \end{figure}
@@
     {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
+grimaldi_learning_2022
      }
-  \label{fig:model}
+  \label{fig:THC} % TODO: citer biol cyb
 \end{figure}
@@
      }
-  \label{fig:model}
+  \label{fig:icip}
 \end{figure}
@@
 More generally, three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
-With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window.
+With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window (see Figure~\ref{fig:icip}).
 This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, form a polychronous group.
-Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:model}).
-In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
 %----------------------------%
@@
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In that example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~if a synchronous pulse is emitted in presynaptic neurons, this will generate post-synaptic potentials at these different delays, and these may not be sufficient to generate a spike in either neuron.
-\textit{(Right)}~If a sequence of activation is generated in the presynaptic neurons such that, taking into account the heterogeneous delays, these converge at the same time on neuron \textit{a} (here, at $t=10~\ms$) then a spike is emitted, while none is emitted on neuron \textit{e}.
+\textit{(Middle)}~if a synchronous pulse is emitted in all presynaptic neurons, this will generate post-synaptic potentials at these different delays, and these may not be sufficient to generate a spike in either neuron.
+\textit{(Right)}~If a sequence of activation is generated in the presynaptic neurons such that, taking into account the heterogeneous delays, these converge at the same time on neuron \textit{a} (here, at $t=10~\ms$) then a spike is emitted (here in red color), while none is emitted on neuron \textit{e}.
      }
@@
 The limits of this model are that its output is only binary and that its storage capacities are limited.
-The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the modulation of spike latency emission~\citep{susi_nmnsd-spiking_2021}.
-Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
-As such, these models use the neural dynamics to handle input signals with different delays but do not explicitly take full advantage of the representation capacity offered by heterogeneous delays.
-
 An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
@@
 While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
+
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:THC}).
+In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
+The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the modulation of spike latency emission~\citep{susi_nmnsd-spiking_2021}.
+Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
+This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
+Recently, by using a logistic regression model coupled with a temporal convolution implemented a model with heterogeneous delays was implemented to test the detection of the spiking motifs embedded in the event stream~\citep{grimaldi_learning_2022}.
+This allowed to detect the detection of a high number of superposed polychronous motifs in synthetic data, illustrating the computational benefit of such representations compared to that with a unique delay (see Figure~\ref{fig:THC}).
+As such, these models use the neural dynamics to handle input signals with different delays but do not explicitly take full advantage of the representation capacity offered by heterogeneous delays.
 %
@@
     {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
-    {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
-    {\bf (c)}~By defining a layer of neurons that detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
-    Note that each afferent may be connected with multiple weights at different delays.
+    {\bf (b)}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay.
+    {\bf (c)}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevitch}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}.
+% such that the addresses are the synapses of these neurons (which we therefore call here synapses).
+    %Note that each afferent may be connected with multiple weights at different delays.
 The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
 %These PG motifs were efficiently learned by a logistic regression.
-    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
+    %{\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
     {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
-grimaldi_learning_2022
      }
-  \label{fig:THC} % TODO: citer biol cyb
+  \label{fig:THC} % TODO: citer biol cyb fig:THC
 \end{figure}
@@
 %----------------------------%
-%
+% TODO
 \begin{figure}%[t!]
   \centering
-  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}
+  \includegraphics[width=0.980\linewidth]{figures/2022-11-10_MotionDetection_kernel.png}
     \caption{Detecting motion in an event stream with heterogeneous delays. 

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 %
-Let us start with a focus of the state-of-the-art on the role of dynamics in vision.
+Let us start with a focus on the state-of-the-art of the role of dynamics in vision.
 Broadly speaking, vision is the set of processes that allow us to make sense of the luminous world, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
@@
 \includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
-\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
-The stimulus is shown in the top row, while trial repetitions of a simulated Leaky Integrate-and-Fire neuron are shown in the second row (with membrane potential represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row).
+\caption{The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
+The stimulus current value over time for a step stimulus (top left) and for a noisy one (top right).
+Trial repetitions of a Leaky Integrate-and-Fire neuron stimulated by the stimulus on the upper row (middle row).
+Membrane potential represented by dark blue color when light yellow colors when depolarized) and quantified by .
+The average firing rate across trials (lower row).
 While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
@@
   \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}
-    \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In that example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~if a synchronous pulse is emitted in all presynaptic neurons, this will generate post-synaptic potentials at these different delays, and these may not be sufficient to generate a spike in either neuron.
-\textit{(Right)}~If a sequence of activation is generated in the presynaptic neurons such that, taking into account the heterogeneous delays, these converge at the same time on neuron \textit{a} (here, at $t=10~\ms$) then a spike is emitted (here in red color), while none is emitted on neuron \textit{e}.
+    \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted.
+    %at these different delays, and these may not be sufficient to generate a spike in either neuron.
+    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, these converge at the same time on neuron \textit{a} (here, at $t=10~\ms$) then a spike is emitted (here in red color), while none is emitted on neuron \textit{e}.
      }

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
     \caption{Detecting motion in an event stream with heterogeneous delays.
+Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
+The models parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
+When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push-pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}.
      }

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
-A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
+A crucial aspect of this emergence is explained by the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
 This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling so-called push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
@@
   \centering
-  \includegraphics[width=0.980\linewidth]{figures/2022-11-10_MotionDetection_kernel.png}
+  \includegraphics[width=0.980\linewidth]{figures/2022-09-27_MotionDetection_kernel.png}
+  %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
     \caption{Detecting motion in an event stream with heterogeneous delays.
@@
 When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push-pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}.
+Global weight is globally decreasing from the lowest delay (right) to less recent information (left).
      }

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
-Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreaseses from $5$ to $1$~\ms).
-This generates a pulse after a certain processing delay in the second group with a diffrent jitter.
+Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreases from $5$ to $1$~\ms).
+This generates a pulse after a certain processing delay in the second group with a different jitter.
 While the first two pulses progressively vanish, from the third input pulse, it is propagated to the following groups.
@@
 It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas yet it is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
-In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
-It relies on the simple remark that a series of repeated coincidences is in itself an invariant.
+In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical artificial neural networks.
+It relies on the simple remark that a serie of repeated coincidences is in itself an invariant.
 Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, or the pitch of a musical note.
@@
 %
-\subsection{A further hypothesis: traveling waves}
+\subsection{A further hypothesis: travelling waves}
 %
@@
 The relative power between baseline and evoked LFP fluctuations was not statistically different from parity (median = 1.23 dB, p = 0.087 two-tailed Wilcoxon's ranked-sum test).
-c An example of spontaneous LFP fluctuations structured as a traveling wave recorded from a spatially distributed multielectrode array in marmoset area MT.
+c An example of spontaneous LFP fluctuations structured as a travelling wave recorded from a spatially distributed multielectrode array in marmoset area MT.
 d Histogram of spontaneous spike probability as a function of the generalized phase of the LFP during fixation.
-e The average evoked response to low contrast stimuli was stronger when a more excitable phase (+ / - \pi rad) of a spontaneous traveling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test).
+e The average evoked response to low contrast stimuli was stronger when a more excitable phase (+ / - \pi rad) of a spontaneous travelling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test).
 Data for panels c--e modified from Davis et al.17 with permission.
@@
 Authors review these findings by considering the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function.
-In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
+In particular, spontaneous travelling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
 Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise.
-However, fluctuations of spontaneous activity, often organized as traveling waves, shape stimulus-evoked responses and perceptual sensitivity.
+However, fluctuations of spontaneous activity, often organized as travelling waves, shape stimulus-evoked responses and perceptual sensitivity.
 The mechanisms underlying these waves are unknown.
@@
 In that paper, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
-They find that spontaneous traveling waves are a general property of these networks.
-The traveling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave.
+They find that spontaneous travelling waves are a general property of these networks.
+The travelling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave.
 Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
@@
 The different aspects of information in the data are evaluated by a decoding strategy, highlighting the role of correlations.
-Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from M.
-Berry's lab~\citep{berry_spike_2022} and allows performing and validate these hypotheses.
-However, a limit of explaining such data with traveling waves is that in theory a traveling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
+Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from Michael Berry's lab~\citep{berry_spike_2022} and allows performing and validate these hypotheses.
+However, a limit of explaining such data with travelling waves is that in theory a travelling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
 Moreover, recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
-~\citep{miller_visual_2014} shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+~\cite{miller_visual_2014} shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
 It proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
@@
 \includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
-\caption{An example of a polychronous group seen in cortical slices.
+\caption{An example of a precise temporal motifs seen in cortical slices.
 In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
@@
 
-One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_1939}.
+One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_functioning_1938}.
 Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{madadi_asl_dendritic_2018}.
@@
 All in all, learning spike motifs requires significantly complex pathways and biological mechanisms whether in the dendrites or the axon of the neuron.
-More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data and to help develop artificial neural networks, whose learning rules would be more neuromorphic.
+More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop artificial neural networks, whose learning rules would be more neuromorphic.
 %
@@
   \centering
-  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}
+  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
@@
 \begin{itemize}
-  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
-  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, traveling waves on arbitrary spiking motifs,
+  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlights the importance of precise spike times,
+  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, travelling waves on arbitrary spiking motifs, etc,
   \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
-  \item  ultimately that using precise puking motifs could be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
+  \item  using precise puking motifs could ultimately be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
 \end{itemize}
@@
 More generally, to reproduce the efficiency of biological systems, one should account for the different temporal scales of adaptation, from seconds to years.
-For instance, the scaffolding of neural assemblies seem to follow critical periods during development~\citep{dard_rapid_2022}.
+For instance, the scaffolding of neural assemblies seems to follow critical periods during development~\citep{dard_rapid_2022}.
 
@@
 Indeed, classic convolutional neural networks emerged as optimal solutions for instance to classify static images into categories, yet they are not well adapted for processing dynamic, multimodal sensory flows.
-The emerging necessity to be able to process more complex flows, such as the multiples flows of information in a car designed for autonomous driving, necessitates to modify such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs.
-Notably, the protocols used as well in neurobiology, theory on engineering should take into account this hovel levels of complexity.
+The emerging necessity to be able to process more complex flows, such as the multiple flows of information in a car designed for autonomous driving, necessitates to modify such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs.
+Notably, the protocols used as well in neurobiology, theory or engineering should take into account this novel levels of complexity.
 
 Ultimately, the community should encourage the adoption of open, reproducible science.
-Indeed, the different models that we have displayed often come with the tools necessary he reproduces the results obtained.
+Indeed, the different models that we have displayed often come with the tools necessary to reproduce the results obtained.
 This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}.
@@
 Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011} i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms.
-As such, this review aimed at paving the way to openly share the variety of resources and to offer a unified view on the role of precise spiking motifs in neuroscience.
+As such, this review aims at paving the way to openly share the variety of resources and to offer a unified view on the role of precise spiking motifs in neuroscience.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 \includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
-\caption{The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
+\caption{\textbf{Reproducibility of the spiking response of a neuron.
+}The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
 The stimulus current value over time for a step stimulus (top left) and for a noisy one (top right).
@@
 \includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
-\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
-The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
+\caption{\textbf{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.} The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
 Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreases from $5$ to $1$~\ms).
@@
 \includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
-\caption{An example of a precise temporal motifs seen in cortical slices.
-In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+\caption{\textbf{An example of a precise temporal motifs seen in cortical slices.} In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
 (A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
@@
 \includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
-\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in SPOTDisClust.
-A) Structure of five ``ground-truth'' patterns (\ldots).
+\caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm.
+\textbf{(A)} Structure of five ``ground-truth'' patterns (\ldots).
 For each pattern and each neuron, a random position was chosen for the activation pulse.
-B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
+\textbf{(B)} Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
 (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
@@
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
-The approach which is currently most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs.
-This algorithms is the most widely used algorithm to tune the weights of a classical (non spiking) Neural Network.
-It iteratively updates the weights according to the direction and magnitude of the error function gradient.
+The approach which is at present most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.
+In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error function gradient.
 Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
 This approach is quite successful, and SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
-The main reason for adopting this approach is to benefit from dedicated neuromorphic low-energy hardware.
+The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
 However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
@@
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain onset time (see Figure~\ref{fig:THC}).
 These models will be compared in light of neuroscientific and computational perspectives.
-We review in this section theoretical and computational foundations of PG detection in models.
-In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
+We review in this section theoretical and computational foundations of spiking motif detection.
+%In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As we saw, most SNN and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}.
+A notable exception of a SNN using precise spike timings is the \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
 This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
@@
 Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns.
-On core ingrredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
+On core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
 This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevitch}).
@@
   \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
+    \caption{\textbf{Detecting event-based motifs using spiking neurons with heterogeneous delays.} 
     {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
@@
   %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
-    \caption{Detecting motion in an event stream with heterogeneous delays.
-Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
+    \caption{\textbf{Detecting visual motion in an event stream with heterogeneous delays.} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
 The models parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
@@
 \includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
-\caption{A miniature, event-based ATIS sensor.
-Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+\caption{\textbf{A miniature, event-based ATIS sensor.} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
 These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
-It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation.
 Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns.
-On core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
+It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that delays are then ``frozen'' for the rest of the simulation.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs).
+One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
 This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevitch}).
-According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, form a polychronous group.
+According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group.
 %----------------------------%
@@
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
     %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, these converge at the same time on neuron \textit{a} (here, at $t=10~\ms$) then a spike is emitted (here in red color), while none is emitted on neuron \textit{e}.
+    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$), then a spike is emitted (red color), while none is emitted from post-synaptic neuron \textit{e}.
      }
@@
 An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
-Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript.
-This more recent work details how this code was slightly corrected.
+Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete such as to allow for the reproduction of all results presented in that manuscript.
+This more recent work details how this code could be slightly corrected.
 It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
@@
 
-Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:THC}).
+Strikingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:THC}).
 In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
@@
 Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task, but were prone to convergence issues.
-Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
+Another model of synaptic delay-weight plasticity integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes~\citep{zhang_supervised_2020}.
 This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 
-There are more applications to image processing using spiking neural networks based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections.
-Using this core computational unit defined, and extending to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation.
+There are more applications to image processing using spiking neural networks.
+For instance, a set of models are based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections.
+Using this core computational unit, and extending it to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation.
 This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks.
-Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
+In essence, two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
 This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
@@
 However, these latter models used parametric rules for defining the weights.
-Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore results from neuromorphic engineering to get better insights on that aspect.
+Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore some exemplar results from neuromorphic engineering to get better insights on that aspect.
 %

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 We will investigate this very hypothesis in this review.
-Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
+%Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
 In comparison to a classical analogical vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
-For instance, it expands the capabilities of analog representations (which rely on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
+For instance, it expands the capabilities of such representations (which rely on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
 This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.
@@
 At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
-Figured inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
+This figure is inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
 As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
@@
 The average firing rate across trials (lower row).
-While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
 See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}

### 2022-11-15 manuscript.tex
5 reformulation

--- 2022-11-15 manuscript.tex
@@
 \includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
-\caption{\textbf{Reproducibility of the spiking response of a neuron.
-}The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
+\caption{The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
 The stimulus current value over time for a step stimulus (top left) and for a noisy one (top right).
@@
 \includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
-\caption{\textbf{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.} The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
+\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
+The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
 Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreases from $5$ to $1$~\ms).
@@
 \includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
-\caption{\textbf{An example of a precise temporal motifs seen in cortical slices.} In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+\caption{An example of a precise temporal motifs seen in cortical slices.
+In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
 (A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
@@
 \includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
-\caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm.
-\textbf{(A)} Structure of five ``ground-truth'' patterns (\ldots).
+\caption{Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in SPOTDisClust.
+A) Structure of five ``ground-truth'' patterns (\ldots).
 For each pattern and each neuron, a random position was chosen for the activation pulse.
-\textbf{(B)} Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
+B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
 (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
@@
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
-The approach which is at present most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.
-In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error function gradient.
+The approach which is currently most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs.
+This algorithms is the most widely used algorithm to tune the weights of a classical (non spiking) Neural Network.
+It iteratively updates the weights according to the direction and magnitude of the error function gradient.
 Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
 This approach is quite successful, and SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
-The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
+The main reason for adopting this approach is to benefit from dedicated neuromorphic low-energy hardware.
 However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
@@
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain onset time (see Figure~\ref{fig:THC}).
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain synaptic delays.
 These models will be compared in light of neuroscientific and computational perspectives.
-We review in this section theoretical and computational foundations of spiking motif detection.
-%In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
+We review in this section theoretical and computational foundations of PG detection in models.
+In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As we saw, most SNN and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-A notable exception of a SNN using precise spike timings is the \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
+A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}.
 This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
 It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
+It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation.
 Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.
-It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that delays are then ``frozen'' for the rest of the simulation.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs).
-One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns.
+On core ingrredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
 This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevitch}).
-According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group.
+According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, form a polychronous group.
 %----------------------------%
@@
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted.
     %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$), then a spike is emitted (red color), while none is emitted from post-synaptic neuron \textit{e}.
+    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, these converge at the same time on neuron \textit{a} (here, at $t=10~\ms$) then a spike is emitted (here in red color), while none is emitted on neuron \textit{e}.
      }
@@
 An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
-Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete such as to allow for the reproduction of all results presented in that manuscript.
-This more recent work details how this code could be slightly corrected.
+Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript.
+This more recent work details how this code was slightly corrected.
 It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
@@
 
-Strikingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:THC}).
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:THC}).
 In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
@@
   \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-    \caption{\textbf{Detecting event-based motifs using spiking neurons with heterogeneous delays.} 
+    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
     {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
@@
 Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task, but were prone to convergence issues.
-Another model of synaptic delay-weight plasticity integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes~\citep{zhang_supervised_2020}.
+Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes.
 This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
@@
   %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
-    \caption{\textbf{Detecting visual motion in an event stream with heterogeneous delays.} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
+    \caption{Detecting motion in an event stream with heterogeneous delays.
+Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
 The models parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
@@
 %
-\subsection{Neuromorphic engineering and the emergence of novel computational architectures}
-%
-Artificial intelligence has made huge advances in the past decades and deep learning algorithms nowadays outperform humans at complex tasks like natural image recognition or abstract strategy board games.
+Artificial intelligence has made huge advances in the past decades and deep learning algorithms, nowadays, outperform humans at complex tasks like natural image recognition or abstract strategy board games.
 Yet, machine learning algorithms suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
 Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019}.
-Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{ref}, neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
-%
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
-\caption{\textbf{A miniature, event-based ATIS sensor.} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
-These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
-Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
-\end{figure}
-%----------------------------%
-%
-This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
-These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene.
+Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
 Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips.
@@
 An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted.
-
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by reproducing such mechanisms, it is interesting to study neural circuitry.
+Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2003, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
+
+This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by artificially reproducing such mechanisms, it is interesting to study neural circuitry.
 Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
-Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
-In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
-There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
-Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
-%
-\subsection{On the importance of spatio-temporal information in silicon retinas}
-%
-With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and processing of visual information.
-From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
-The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision tasks.
-In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
-An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
-It gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
-Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
-Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022} and show that this method could be efficiently applied to state-of-the-art benchmarks.
-
-More generally, three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
-With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window (see Figure~\ref{fig:icip}).
-This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
-Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
-This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of the precise temporal resolution of neuromorphic retina-like cameras.
-By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
-With this type of signal, the use of spatio-temporal motifs embedded in the event streams are essential to solve high level visual tasks.
-%
-\subsection{Computations with delays in neuromorphic hardware}
-%
-SNNs are considered the third generation of ANNs and, also for the asynchronous nature of computations, neuromorphic engineering is promising for the future of artificial intelligence.
-Both fields of research are characterized by local asynchronous computing units and the absence of a central clock, and this is why the use of the temporal dimension in these types of architectures seems crucial.
-The previous section demonstrates the importance of using spatio-temporal motifs in event-based recordings, here we focus on the implementation of such event-based algorithms in neuromorphic hardware.
-Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2003, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
-
+In this section, we make a description of the different neuromorphic hardware that have been developed and see how they can be used to deal with precise temporal motifs.
+%
+\subsection{The emergence of novel computational architectures}
+%
 To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_analog_1989}.
@@
 The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector''.
-Today many devices are good candidates for implementing event-based algorithms and use the AER, they can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
+Today, many devices are good candidates for implementing event-based algorithms and use the AER.
+They can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
 For a more complete review the reader can refer to~\citep{schuman_survey_2017}, here we site the most popular ones.
@@
 Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.
-
+%
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
+\caption{A miniature, event-based ATIS sensor.
+Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
+Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
+\end{figure}
+%----------------------------%
+%
+Neuromorphic sensors are also developed with the idea to capture external stimuli more efficiently and closer to biological systems.
+A widely used example is the Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
+These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene.
+We also report other event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but will focus on DVS for the next subsection about the use of dynamics embedded in event-based signals.
+
+%Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
+This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by reproducing such mechanisms, it is interesting to study neural circuitry.
+Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
+Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
+In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
+There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
+Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
+%
+\subsection{On the importance of spatio-temporal information in silicon retinas}
+%
+With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and processing of visual information.
+From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
+The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision tasks.
+
+In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
+It gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
+Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
+Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022} and show that this method could be efficiently applied to state-of-the-art benchmarks.
+
+More generally, three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
+With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window (see Figure~\ref{fig:icip}).
+This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
+Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
+This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of the precise temporal resolution of neuromorphic retina-like cameras.
+By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
+With this type of signal, the use of spatio-temporal motifs embedded in the event streams are essential to solve high level visual tasks.
+%
+\subsection{Computations with delays in neuromorphic hardware}
+%
 For the rest of this section, we report examples of implementations of event-based algorithms using precise spatio-temporal motifs on neuromorphic hardware. 
@@
 They show that STDP and Spike Timing Dependant Delay Plasticity rules could be implemented in both a digital and an analog chip.
-Physical limitations to store the network on-chip and, from the parameters of the network and the current technology, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
-The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citet{wang_fpga_2013}.
+From the network parameters and the physical limitations to store it on-chip, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
+The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2013}.
 %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
@@
 Coherence detection on input spikes coming from two sources was obtained with a 50 $ns$ precision.
-They claim that this unsupervised learning denoise the input and compensates for variations between neural components.
+They claim that this unsupervised learning denoises the input and compensates for variations between neural components.
 The variability of response of the analog components can be compensated by population coding for responses robust to noise and this phenomenon is also observed in biological neural networks~\citep{boerlin_spike-based_2011}.
-A recent work makes the first implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to com-
-pute the binding operation~\citep{renner_sparse_2022}.
+A recent work makes an implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}.
 They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equilavent to the maximal delay.

### 2022-11-15 manuscript.tex

--- 2022-11-15 manuscript.tex
@@
 Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.
-
+%
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
+\caption{\textbf{A miniature, event-based ATIS sensor.} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
+Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
+\end{figure}
+%----------------------------%
+%
+Neuromorphic sensors are also developed with the idea to capture external stimuli more efficiently and closer to biological systems.
+A widely used example is the Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
+These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene.
+We also report other event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but will focus on DVS for the next subsection about the use of dynamics embedded in event-based signals.
+
+%Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
+This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by reproducing such mechanisms, it is interesting to study neural circuitry.
+Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
+Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
+In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
+There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
+Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
+%
+\subsection{On the importance of spatio-temporal information in silicon retinas}
+%
+With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and processing of visual information.
+From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
+The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision tasks.
+
+In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
+It gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
+Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
+Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022} and show that this method could be efficiently applied to state-of-the-art benchmarks.
+
+More generally, three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
+With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window (see Figure~\ref{fig:icip}).
+This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
+Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
+This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of the precise temporal resolution of neuromorphic retina-like cameras.
+By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
+With this type of signal, the use of spatio-temporal motifs embedded in the event streams are essential to solve high level visual tasks.
+%
+\subsection{Computations with delays in neuromorphic hardware}
+%
 For the rest of this section, we report examples of implementations of event-based algorithms using precise spatio-temporal motifs on neuromorphic hardware. 

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-The book then demonstrates that this could take the form of ``synfire chains,'' that is, the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
+The book then demonstrates that neural assemblies could form so-called ``synfire chains,'' that is, showing the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
 More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}.
@@
 
-It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some (frozen) noise patterns.
+It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some frozen noise patterns.
 In particular, they used random waveforms to probe the formation of new memories for arbitrary complex sounds.
 A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \si{\second}$ long.
-The task is to detect the repetition of the same (frozen) noise within a trial.
+%The task is to detect the repetition of the same (frozen) noise within a trial.
 Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block.
@@
     %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$), then a spike is emitted (red color), while none is emitted from post-synaptic neuron \textit{e}.
+    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
      }

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we will review current evidence of taking advantage of spiking motifs in neurobiology, that is, from actual recordings in biological tissues.
+In this section, we will review current evidence on the role of spiking motifs in neurobiology, that is, from our understanding arising from actual recordings in biological tissues.
 In most generic computational models, the neural activity is assumed to be encoded in the firing rate.

### 2022-11-15 manuscript.tex
Update on Overleaf.

@@
 Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
-These observations lead to the intuition that each distance may be related as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
+These observations lead to the intuition that each distance may be as well  as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
 % ~\citep{}

### 2022-11-15 manuscript.tex
section 3

@@
 % https://www.zotero.org/groups/4562620/polychronies
-%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
@@
 While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
-See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
+See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of the results from~\citep{mainen_reliability_1995}  using a simple LIF model.}\label{fig:mainen}
 \end{figure}
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{A rediscovered hypothesis: precise spiking motifs in cell assemblies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Travelling waves indicate that spatio-temporal correlations could play an important role in shaping neural activity.
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %\subsection{Explictly decoding spiking motifs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
@@
 % \end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Understanding precise spiking motifs in neurobiology}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we will review current evidence on the role of spiking motifs in neurobiology, that is, from our understanding arising from actual recordings in biological tissues.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In this section, we will review current evidence on how we may take advantage of spiking motifs in neurobiology, that is, in actual recordings from biological neural tissues.
 In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
@@
 %
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Decoding neural activity using spike distances}\label{sec:detection}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 There are different solutions to provide with a distance between two given spike trains.
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Scaling up to very large scale data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{What biological mechanism could allow learning spiking motifs?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Modeling precisely-timed spiking motifs in computational neuroscience}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In the following section, we review theoretical models which directly take advantage of using precise spiking motifs. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
@@
 %In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Izhikevitch's polychronization model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % polychronization
@@
 %----------------------------%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Learning synaptic delays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Training delays ~\citep{state_training_2019}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Real-world applications}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %----------------------------%
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Applications of precise spiking motifs in neuromorphic engineering}
 \label{sec:neuromorphic}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}\label{sec:discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Summary}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %: >>>>>>>> Laurent is here <<<<<<<<<<<<<<

### 2022-11-15 manuscript.tex
section 1 & 2

@@
 This generates a pulse after a certain processing delay in the second group with a different jitter.
-While the first two pulses progressively vanish, from the third input pulse, it is propagated to the following groups.
+While the first two pulses progressively vanish in the following groups, starting from the third input pulse (with a jitter of $3~\ms$), it is propagated to the following groups.
 This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we will review current evidence on how we may take advantage of spiking motifs in neurobiology, that is, in actual recordings from biological neural tissues.
+In this section, we will review current evidence on how we may take advantage of spiking motifs in neurobiology, that is, in an effort to understand actual recordings from biological neural tissues.
 In most generic computational models, the neural activity is assumed to be encoded in the firing rate.

### 2022-11-15 manuscript.tex
new section :-0

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Understanding precise spiking motifs in neurobiology}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Understanding precise spiking motifs in neurobiology}\label{sec:detection}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decoding neural activity from firing rates}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding neural activity using spike distances}\label{sec:detection}
+\subsection{Decoding neural activity using spike distances}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{What biological mechanism could allow learning spiking motifs?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{What biological mechanism could allow learning spiking motifs?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Biological observations of delay adaptation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % First biological observations
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The importance of myelination}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % The importance of myelination
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Interplay of delay adaptation and neural activity}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % How is the neuronal activity detected ?

### 2022-11-16 manuscript.tex
new section :-0
Update on Overleaf.

@@
 This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.~\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
+According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.~\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
 A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
@@
 % The importance of myelination
-
 \citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
@@
 Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
-Myelinization consists in ``two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}.
-Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the $\backsim$1 $\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_2021}.
+Myelinization consists in ``two motions: the wrapping of the leading edge of the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}.
+Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the approx $1~\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_neuron-oligodendrocyte_2021}.
 In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
@@
 % How is the neuronal activity detected ?
-
 However, one question remains: how do oligodendrocytes detect neuronal activity and regulate the myelinization accordingly?
@@
 Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
-Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as $Ca^{2+}$ activity~\citep{baraban_ca2_2018} or the neuregulin 1 growth factor~\citep{nave_axonal_2006}.
+Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as $Ca^{2+}$ activity~\citep{baraban_ca2_2018} of the neuregulin 1 growth factor~\citep{nave_axonal_2006}.
 The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity.
@@
 
-A potential scenario for selective myelination on electrically active axons using non synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
+A potential scenario for selective myelination on electrically active axons using non-synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
 This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_new_2015}.
@@
 However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
-Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
+Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
 The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.
 
-All in all, learning spike motifs requires significantly complex pathways and biological mechanisms whether in the dendrites or the axon of the neuron.
+All in all, learning spike motifs requires significantly complex pathways and biological mechanisms, whether in the dendrites or the axon of the neuron.
 More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop artificial neural networks, whose learning rules would be more neuromorphic.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Modeling precisely-timed spiking motifs in computational neuroscience}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the following section, we review theoretical models which directly take advantage of using precise spiking motifs. 
+\section{Modeling precise spiking motifs in theoretical \& computational neuroscience}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Now that we have reviewed biological foundations forthe role of delays in neural computations, we review in the following section theoretical models which directly take advantage of using precise spiking motifs.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.
-In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error function gradient.
-Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
+In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error's gradient.
+Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
 This approach is quite successful, and SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
 The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
-However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
+However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
 There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-To overcome the diversity of input presynaptic patterns, we will explore different existing architectures which are able to learn to detect stable ``polychronous'' events, that is, volleys of spikes which are stable up to certain onset time (see Figure~\ref{fig:THC}).
+In particular, we will explore different existing architectures which are able to overcome the diversity of input presynaptic patterns and learn to detect stable spiking motifs, that is, volleys of spikes which are stable up to a certain onset time (see Figure~\ref{fig:THC}).
 These models will be compared in light of neuroscientific and computational perspectives.
-We review in this section theoretical and computational foundations of spiking motif detection.
+%We review in this section theoretical and computational foundations of spiking motif detection.
 %In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.

### 2022-11-16 manuscript.tex
Figure visual-latency.svg

@@
 In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-80~\ms$).
-Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $120-140~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal chord to trigger finger muscles, with latencies about $200-350~\ms$.}\label{fig:thorpe}
+Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $120-140~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal cord to trigger finger muscles, with latencies about $200-350~\ms$.}\label{fig:thorpe}
 \end{figure}

### 2022-11-16 manuscript.tex
section 4

@@
 First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered, for instance for sound localization~\citep{chase_first-spike_2007}.
-Note also that timing is not entirely sensory or internal, but can be used as a general coding principle.
+Note also that timing is not entirely sensory or internal, but can be used as a general neural coding principle.
 In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
@@
 
-Interestingly, it can be shown that these travelling waves could have a measurable role on the activity of the visual cortex.
+Interestingly, it can be shown that these travelling waves could have a measurable impact on the activity of the visual cortex.
 This has been illustrated in a recent study investigating the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
 The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in close visual proximity.
-The lrAM is the building block of the use of sequences of images to induce the perception of smooth, realistic visual scenes, which is at the base of movies seen in cinema theaters.
-In this study, the authors used voltage-sensitive dye imaging to record the activity of the primary visual cortex of macaque monkeys to the presentation of the dots alone or in conjunction.
+The lrAM is the core building block underlying the use of sequences of images to induce the perception of smooth, realistic visual scenes, which is at the base of movies seen in cinema theaters.
+In this study, the authors used voltage-sensitive dye imaging to record the activity of the primary visual cortex of macaque monkeys to the presentation of the pair of dots presented independently or in conjunction.
 A probabilistic modelling has shown that the activity of the joint presentation induced a suppressive wave in the direction opposed to the perceived direction, shaping the formation of a wave of propagation travelling at a speed compatible with the perceived motion.
 A computational model validated the hypothesis that this process could be mediated by diffusion in the horizontal layers connecting the different locations within this cortical area.
+In summary, the study by~\citet{chemla_suppressive_2019} gave a multi-disciplinary account to demonstrate the effect of travelling waves in the visual cortex.
 %
@@
 For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process.
-As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.
+As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves, that is, its selectivity to a range of features~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.
 This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
@@
 
-Let us now focus on distances defined based on a generative model formed by using precise spiking motifs.
-A review of the methods for neuronal response latency estimation exists and includes Bayesian binning~\citep{levakova_review_2015}.
+%Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
+Concerning spike timings,~\citet{levakova_review_2015} has reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning.
 More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
@@
 According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.~\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
+In the mouse somatosensory cortex,~\citep{camon_timing_2019}  found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.
 A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.

### 2022-11-16 manuscript.tex
Update on Overleaf.

@@
 %
-Despite the extensive computer evidence of the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
+Despite the evidence for the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
 The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviors allow us to put these doubts to rest.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Now that we have reviewed biological foundations forthe role of delays in neural computations, we review in the following section theoretical models which directly take advantage of using precise spiking motifs.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{maass_networks_1997}.
-The approach which is at present most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
+Now that we have reviewed biological foundations for the role of delays in neural computations, we review, in the following section, theoretical models which directly take advantage of using precise spiking motifs.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Spiking Neural Networks (SNNs)~\citep{maass_networks_1997} are natural candidates to use these precise temporal patterns in the brain.
+The approach which is currently most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
 One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.

### 2022-11-16 manuscript.tex
figures/izhikevich-repro.svg

--- 2022-11-16 manuscript.tex
@@
 This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{madadi_asl_dendritic_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_disruption_2020}, contextual fear conditioning~\citep{pan_preservation_2020} or OPCs knock-out~\citep{wan_impaired_2020}.
+Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{madadi_asl_dendritic_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_disruption_2020}, contextual fear conditioning~\citep{pan_preservation_2020} or oligodendrocyte precursor cells (OPCs) knock-out~\citep{wan_impaired_2020}.
 Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_new_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_neuron-oligodendrocyte_2021}.
@@
 To answer this, we must first study the myelinization process.
-The oligodendrocyte precursor cells (OPCs) first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
+The OPCs first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
 Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
@@
 From the network parameters and the physical limitations to store it on-chip, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
-The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2013}.
+The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2014}.
 %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.

### 2022-11-16 manuscript.tex
figures/izhikevich-repro.svg

@@
   \centering
-  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
+  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.

### 2022-11-16 manuscript.tex
figures/izhikevich-repro.svg
Update on Overleaf.

--- 2022-11-16 manuscript.tex
@@
 % polychronization
-As we saw, most SNN and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
+As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
 A notable exception of a SNN using precise spike timings is the \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.

### 2022-11-16 manuscript.tex

--- 2022-11-16 manuscript.tex
@@
   \centering
-  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
+  \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.

### 2022-11-16 manuscript.tex
Update on Overleaf.

--- 2022-11-16 manuscript.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Izhikevitch's polychronization model}
+\subsection{Izhikevich's polychronization model}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevitch}).
+This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevich}).
 According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group.
@@
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no output spike will be emitted.
     %at these different delays, and these may not be sufficient to generate a spike in either neuron.
@@
      }
-  \label{fig:izhikevitch}
+  \label{fig:izhikevich}
 \end{figure}
@@
 The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron, which does not incorporate a spike timing framework.
-The \emph{tempotron} learning rule is derived by an optimization process and takes the form of an STDP rule.
+The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a supervised STDP rule.
 The limits of this model are that its output is only binary and that its storage capacities are limited.
@@
 This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
-Recently, by using a logistic regression model coupled with a temporal convolution implemented a model with heterogeneous delays was implemented to test the detection of the spiking motifs embedded in the event stream~\citep{grimaldi_learning_2022}.
-This allowed to detect the detection of a high number of superposed polychronous motifs in synthetic data, illustrating the computational benefit of such representations compared to that with a unique delay (see Figure~\ref{fig:THC}).
+Recently, by using a logistic regression model coupled with a temporal convolution, a model with heterogeneous delays was implemented to test the detection of the spiking motifs embedded in an event stream~\citep{grimaldi_learning_2022}.
+This allowed to detect a high number of superposed polychronous motifs in synthetic data, illustrating the computational benefit of such representations compared to that with a unique delay (see Figure~\ref{fig:THC}).
 As such, these models use the neural dynamics to handle input signals with different delays but do not explicitly take full advantage of the representation capacity offered by heterogeneous delays.
@@
     {\bf (b)}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay.
-    {\bf (c)}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevitch}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}.
+    {\bf (c)}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}.
 % such that the addresses are the synapses of these neurons (which we therefore call here synapses).
@@
 % Training delays ~\citep{state_training_2019}
-First, the original model use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
+First, the original model by Izhikevich use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
 For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.

### 2022-11-16 manuscript.tex
figures/izhikevich-repro.svg

--- 2022-11-16 manuscript.tex
@@
 Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}.
-In addition to delay learning, a group, in the University of West Sydney, developed a neurmorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}.
-They show that STDP and Spike Timing Dependant Delay Plasticity rules could be implemented in both a digital and an analog chip.
+In addition to delay learning, a group at the University of West Sydney developed a neuromorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}.
+They show that STDP and Spike Timing Dependent Delay Plasticity rules could be implemented in both a digital and an analog chip.
 From the network parameters and the physical limitations to store it on-chip, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
@@
 A recent work makes an implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}.
-They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equilavent to the maximal delay.
+They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay.
 Note that this problem is also due the algorithm used in this paper and that the analog chips must not suffer from this type of limitations.
@@
 The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
+Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts).
 %

### 2022-11-16 manuscript.tex
figures/izhikevich-repro.svg

@@
 \textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+% Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no output spike will be emitted.
     %at these different delays, and these may not be sufficient to generate a spike in either neuron.

### 2022-11-16 manuscript.tex
figures/izhikevich-repro.svg

@@
     \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
-% Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no output spike will be emitted.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
     %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
+    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
      }

### 2022-11-16 manuscript.tex
discussion

@@
 The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
-Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts).
+Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts)~\citep{coull_distinction_2022}.
 %

### 2022-11-16 manuscript.tex
discussion

@@
 \dataavailability{This works is made reproducible using the following tools.
-First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/XXX}{GitHub}~\citep{Grimaldi2022review}.
+First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/XXX}{GitHub}.
+%~\citep{Grimaldi2023review}.
 The paper is available as an \href{https://arxiv.org/abs/YYYY.XXX}{arXiv preprint} with links to previous versions.
-Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{zotero group} which was used to regroup relevant literature on the subject.}
-
-\acknowledgments{We would like to thank Jean-Nicolas Jérémie, Hugo Ladret for useful discussions while writing this review.}
+Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{Zotero group} which was used to regroup relevant literature on the subject.}
+
+\acknowledgments{We would like to thank Jean-Nicolas Jérémie, George Abitbol and Hugo Ladret for useful discussions while writing this review.}
 

### 2022-11-16 manuscript.tex
Update on Overleaf.

@@
 Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \textit{in vivo}.
-Such waves also occur in motor areas and~\citet{linden_movement_2021} have recently presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+Such waves also occur in motor areas and~\citet{linden_movement_2022} have recently presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
 
@@
   \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
-  \item  using precise puking motifs could ultimately be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
+  \item  using precise spiking motifs could ultimately be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
 \end{itemize}

### 2022-11-16 2022-11-16_coverletter.tex
covzer letter

%!TeX TS-program = lualatex
%!TeX encoding = UTF-8 Unicode
%!TeX spellcheck = fr-FR
%!BIB TS-program = biber
% -*- coding: UTF-8; -*-
% vim: set fenc=utf-8
%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\documentclass[10pt,english]{article}

\usepackage{charter,graphicx}
\usepackage[margin=1in]{geometry}

\usepackage{hyperref,tabularx}
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{.4pt}% Default header rule
\renewcommand{\footrulewidth}{0pt}% No footer rule
\fancyhf{}% Clear header/footer
\fancypagestyle{plain}{
  \renewcommand{\headrulewidth}{0pt}% No header rule
  \renewcommand{\footrulewidth}{.4pt}% Default footer rule
  \fancyhf{}% Clear header/footer
}
\AtBeginDocument{\thispagestyle{plain}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{.5\baselineskip plus 1pt minus 1pt}
\usepackage[utf8]{inputenc}%
\usepackage{babel}%
\usepackage{etoolbox}
\makeatletter
\patchcmd{\@zfancyhead}{\fancy@reset}{\f@nch@reset}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyorh}{\f@nch@orh}{}{}
\makeatother
%\usepackage{kpfonts}
%\usepackage{babel}
%\usepackage{csquotes}
\usepackage{url}
\usepackage{charter} % Use the Charter font for the document text
%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------
%Jean-Nicolas Jérémie\\
%Institute of Neuroscience of la Timone\\
%CNRS - Aix Marseille University\\
%27, boulevard Jean Moulin\\
%13005 Marseille, France
\newcommand{\LastName}{Perrinet}%{Grimaldi}%
\newcommand{\FirstName}{Laurent U}%{Antoine}%
\newcommand{\Institute}{Institut de Neurosciences de la Timone, CNRS / Aix-Marseille Universit\'e}%
\newcommand{\Address}{27, Bd. Jean Moulin, 13385 Marseille Cedex 5, France}%
\newcommand{\Website}{\url{https://laurentperrinet.github.io/}}%
\newcommand{\Email}{\url{laurent.perrinet@univ-amu.fr}}%
%----------------------------------------------------------------------------------------
%	YOUR DOCUMENT
%----------------------------------------------------------------------------------------
\usepackage[document]{ragged2e}
\begin{document}
%\centerline{
\includegraphics[width=.4\textwidth]{~/metagit/blog/perrinet_curriculum-vitae_tex/troislogos.png}
%}
\\
\vspace{.1\baselineskip}
\hrulefill
\vspace{.1\baselineskip}

\begin{flushright}
	\FirstName\  \LastName\  \\
	\Institute\\[6pt]
	\Address\\%[6pt]
	\Website \\
	  Email: \Email \\[6pt]
\end{flushright}
\justifying
\vspace{1\baselineskip}
%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------
%À qui de droit,\\
Marseille, \today\\[12pt] % Date

Dear editor,

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

Following your invitation to submit an article to Brain Sciences, please find enclosed a manuscript entitled ``Precise spiking motifs in neurobiological and neuromorphic data'' for your consideration as a review paper in \emph{Brain Sciences}. All the authors have been involved with the work, have approved the manuscript and agreed to its submission.

This is an invited Feature Paper, with a full waiver on the APC. In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source

Many thanks for your consideration.
\vspace{.2\baselineskip}
%----------------------------------------------------------------------------------------
Sincerely yours,

\vspace{.1cm}

\FirstName\ \LastName \\

\includegraphics[width=.3\linewidth]{~/quantic/RTC/signature_LuP.jpg}

\end{document}

### 2022-11-16 manuscript.tex
covzer letter

--- 2022-11-16 manuscript.tex
@@
 \address{%
-$^{1}$ \quad INT UMR 7289, Université Aix Marseille, CNRS; 27 Bd Jean Moulin, 13005 Marseille, France\\
+$^{1}$ \quad INT UMR 7289, Aix Marseille Univ, CNRS; 27 Bd Jean Moulin, 13005 Marseille, France\\
 $^{2}$ \quad SPARKS,  Côte d'Azur, CNRS, I3S; 2000 Rte des Lucioles, 06900 Sophia-Antipolis, France

### 2022-12-12 manuscript.tex
Update on Overleaf.

@@
 % CONTEXT
-Neural activity is directly influenced by our immediate environment and by internal states, and is structured in order to generate motor actions.
+Neural activity is directly influenced by our immediate environment and by internal states and is structured in order to generate motor actions.
 The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection.
-A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings that can be observed in neurobiological experiments.
+A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings observed in neurobiological experiments.
 One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
@@
 In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-A central and well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}.
+A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}.
 In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
@@
 The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
-Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
+Indeed, most neurophysiologists use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
 This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
@@
 % GAP
-However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role, and that neurons may be as well integrators than synchrony detectors~\citep{abeles_role_1982}.
+However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role and that neurons may be synchrony detectors as well as integrators~\citep{abeles_role_1982}.
 We will investigate this very hypothesis in this review.
 %Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
-In comparison to a classical analogical vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
-For instance, it expands the capabilities of such representations (which rely on the firing rate) by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
+In comparison to a classical analog vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
+For instance, it expands the capabilities of representations, that rely on the firing rate, by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
 This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.
@@
 Let us start with a focus on the state-of-the-art of the role of dynamics in vision.
-Broadly speaking, vision is the set of processes that allow us to make sense of the luminous world, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
+Broadly speaking, vision is the set of processes that allow us to make sense of the world through luminous signals, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
 In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between the topology of external visual space and its representation on the retina, called retinotopy.

### 2022-12-13 manuscript.tex
Update on Overleaf.

@@
 \newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+\usepackage[margins]{trackchanges}
+
+\addeditor{Antoine}
+\addeditor{Amélie}
+\addeditor{Camille}
+\addeditor{Jean}
+\addeditor{JN}
+\addeditor{Laurent}
+
 %\newcommand{\note}[1]{}
@@
 ``\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}''.
-Typically, these are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.
+\change[Antoine]{Typically, these are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
+Once integrated by the dendritic tree and the soma of the neuron, these spikes result in a modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.}
 The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.

### 2022-12-13 manuscript.tex
Update on Overleaf.

@@
 \newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
-\usepackage[margins]{trackchanges}
+\usepackage{trackchanges}
 
@@
 In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}.
+A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered at the beginning of the $XX^{th}$ century~\citep{adrian_impulses_1926}.
 In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
@@
 \change[Antoine]{Typically, these are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
-Once integrated by the dendritic tree and the soma of the neuron, these spikes result in a modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.}
-The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
-Indeed, most neurophysiologists use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
+The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.}
+The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code \note{change that and add a more wide intro about other possible constituent of the neural code}.
+\change[Antoine]{Indded, neurophysiologists typically}{Most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
 This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
@@
 % GAP
-However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role and that neurons may be synchrony detectors as well as integrators~\citep{abeles_role_1982}.
+However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role and that neurons \change[Antoine]{may be as well integrators than synchrony detectors}{may be synchrony detectors as well as integrators}~\citep{abeles_role_1982}.
 We will investigate this very hypothesis in this review.
 %Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
-In comparison to a classical analog vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
-For instance, it expands the capabilities of representations, that rely on the firing rate, by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
+In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
+For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{that rely on the firing rate} by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
 This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.
 Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons.
+All these findings, and more~\citep{bohte_evidence_2004,dilorenzo_spike_2013}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons.
 A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
@@
 Let us start with a focus on the state-of-the-art of the role of dynamics in vision.
-Broadly speaking, vision is the set of processes that allow us to make sense of the world through luminous signals, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
+Broadly speaking, vision is the set of processes that allow us \change[Antoine]{to make sense of the luminous world}{to make sense of the world through luminous signals}, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
 In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between the topology of external visual space and its representation on the retina, called retinotopy.
 The origins of this neuroscientific question can be found in the first experiments from Pierre Flourens which, using lesions in animals, demonstrated the relationship between visual sensations and activity in the cerebral cortex~\citep{flourens_recherches_1842,pearce_marie-jean-pierre_2009}.
-This was also observed by Lord Adrian when recording the activity the frog's visual system~\citep{adrian_impulses_1926}.
+This was also observed \remove[Antoine]{by Lord Adrian} when recording the activity of the frog's visual system~\citep{adrian_impulses_1926}.
 In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as the visual orientation or motion of elementary contours.
-For a large proportion of neurons, there is remarkably a monotonic relationship between the contrast of visual features and the firing frequency of neurons.
+For a large proportion of neurons, there is \change[Antoine]{remarkably a monotonic}{a remarkably monotonic} relationship between the contrast of visual features and the firing frequency of neurons.
 However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitute essential pieces to this puzzle~\citep{carandini_normalization_2012}.
@@
 \end{figure}
+\note{Je changerai cette figure car nous parlons du système visuel des primates non humains}
 %----------------------------%
@@
 For instance, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
-His experiments consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
+\change[Antoine]{His}{Their} experiments consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
 The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
@@
 Different experimental protocols have shown for example that the motor response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
-This fast processing correlates with the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
+This fast processing correlates with the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and decoding via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
 As expected, the performances decrease progressively as the frequency of presentation of the images increases.
@@
 An important consequence of this speed of processing of vision is that it implies that processing is carried out using only very few spikes per layer.
-As a comparison, the latencies in macaque monkey are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.
+As a comparison, the latencies in macaque monkeys are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.
 Indeed, we may consider that a behavioral response occurring in only $200~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
@@
 At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
-This figure is inspired by a similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
+This figure is inspired by similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
 As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
@@
 % reliability of spike timings
-Sensory data is most often represented by continuous values, such as the energy produced by a flow of photons which hits the different photoreceptors of the retina.
+Sensory data is most often represented by continuous values, such as the energy produced by a flow of photons that hits the different photoreceptors of the retina.
 How may such information be encoded in neuronal activity?
@@
 For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it at first sight inconsistent with the temporal integration of firing rate.
-Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: while the first spike is aligned to stimulus' time, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
+Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: while the first spike is aligned to stimulus' time, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{bryant_spike_1976,mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
 However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see~\citep{ermentrout_reliability_2008}).
@@
 At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
-In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
+\note{here we have to make the choice of studying only vertebrates, otherwise there is a huge review to add on insects.
+Gollisch and Meister is the fly retina?
+Maybe find another ref} In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
 These results have been extended to natural images and show a qualitatively similar behavior.
@@
 Trial repetitions of a Leaky Integrate-and-Fire neuron stimulated by the stimulus on the upper row (middle row).
-Membrane potential represented by dark blue color when light yellow colors when depolarized) and quantified by .
+Membrane potential is represented by dark blue color when light yellow colors when depolarized) and quantified by .
 The average firing rate across trials (lower row).
-While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{frozen} noise at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{frozen} noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
 See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of the results from~\citep{mainen_reliability_1995}  using a simple LIF model.}\label{fig:mainen}

### 2022-12-13 manuscript.tex
Update on Overleaf.

@@
 One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron by using the rank of neurons' activation~\citep{delorme_spikenet_1999,bonilla_analyzing_2022}.
-Another class of artificial SNNs use precise spike timing as a metric in order to determine the structure of the network in order to minimize a cost function.
+Another class of artificial SNNs uses precise spike timing as a metric in order to determine the structure of the network in order to minimize a cost function.
 This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods.
@@
 However, the performance of SNNs is still lagging compared to that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
-Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas.
-These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}).
-% This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
+\change[Antoine]{Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of event-based cameras}{Improvements in this new generation of Artificial Neural networks (ANN) would bring major advances in terms of efficient computations in machine learning.
+They would benefit in particular to a new generation of cameras called Silicon Retinas}~\citep{gallego_event-based_2022} (see Section~\ref{sec:neuromorphic}).
+%These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developed in Section~\ref{sec:neuromorphic}).% This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
 To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.
@@
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018}, and were also extended to larger assemblies~\citep{torre_synchronous_2016} using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
 Note also that synchronicity may explain some unintuitive results.
@@
 However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons~\citep{bruno_cortex_2006}.
+\noet{neuron-centric view from the reviewer comment, make some changes to broaden the proposed vision.
+Also change the very specific example given after, or put in context (works by Sophie Dénève or Ballard and Jehee)}
 
@@
 Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Indeed, simultaneously recorded activities of neurons in the primary motor cortex of monkeys exhibited context-dependent, rapid changes in the patterns of coincident action potentials~\citep{riehle_spike_1997,grammont_precise_1999}.
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas yet it is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
-In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical artificial neural networks.
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas yet it is less clear \remove[Antoine]{, though,} how these populations organize dynamically to cope with the momentary computational demands.
+In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical ANNs.
 It relies on the simple remark that a serie of repeated coincidences is in itself an invariant.
@@
 This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}.
-The synchronicity will generate rhythms at different range of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down.
+The synchronicity will generate rhythms at different ranges of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down.
 Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion, i.e.
@@
 More generally, it can be shown that the phase of alpha oscillations (about $10$ Hz) is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
-The questions still remains open as to know if this is an epiphenomenom or a working mechanism of the neural code.
+The question still remains open as to know if this is an epiphenomenon or a working mechanism of the neural code.
 %
@@
 All in all, learning spike motifs requires significantly complex pathways and biological mechanisms, whether in the dendrites or the axon of the neuron.
-More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop artificial neural networks, whose learning rules would be more neuromorphic.
+More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop ANNs, whose learning rules would be more neuromorphic.
 %
@@
 % LP received funding 
-and LP from A*Midex grant number AMX-21-RID-025
-``\href{https://laurentperrinet.github.io/grant/polychronies/}{Polychronies}''.}
+LP received support from the french government under the France 2030 investment plan, as part of the Initiative d’Excellence d’Aix-Marseille Université – A*MIDEX ” AMX-21-RID-025 ``\href{https://laurentperrinet.github.io/grant/polychronies/}{Polychronies}''.}
 

### 2022-12-14 manuscript.tex
Update on Overleaf.

@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0002-9536-010X} % AmG % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0002-3107-4788} % AmG % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
@@
 \newcommand{\orcidauthorC}{0000-0002-0054-2729} % CB
-\newcommand{\orcidauthorD}{0000-0001-8821-5556} % JM
-\newcommand{\orcidauthorE}{0000-0002-3107-4788} % LP
+\newcommand{\orcidauthorD}{0000-0002-9238-6654} % JN
+\newcommand{\orcidauthorE}{0000-0001-8821-5556} % JM
+\newcommand{\orcidauthorF}{0000-0002-9536-010X} % LP
 
@@
 Camille Besnainou $^{1}$\orcidC{},
-Jean Martinet $^{2}$\orcidD{},
-and Laurent Perrinet $^{1,}$\orcidE{}*}
+Jean-Nicolas Jérémie $^{1}$\orcidD{},
+Jean Martinet $^{2}$\orcidE{},
+and Laurent Perrinet $^{1,}$\orcidF{}*}
 
@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean Martinet and Laurent Perrinet}
+\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent Perrinet}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Grimaldi, A.; Gruel, A.; Besnainou, C.; Martinet, J.; and Perrinet, L.}
+\AuthorCitation{Grimaldi, A.; Gruel, A.; Besnainou, C.; Jérémie, J.-N.;Martinet, J.; and Perrinet, L.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
-an explicit representation within the neural activity.
+an explicit representation \add[Antoine]{of cognitive processes} within the neural activity.
 Nevertheless, we will use this terminology in all generality to denote the existence of a structure in neural activity.
-In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered at the beginning of the $XX^{th}$ century~\citep{adrian_impulses_1926}.
+In this respect, it is \change[Antoine]{known}{reasonable to declare} that neural activity may be related to specific measurable variables\change[Antoine]{, and since Galvani's}{.
+Since Galvani's} experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
+A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered \change[Antoine]{by Lord Adrian}{at the beginning of the $XX^{th}$ century}~\citep{adrian_impulses_1926}.
 In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
@@
 The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.}
-The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code \note{change that and add a more wide intro about other possible constituent of the neural code}.
+The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
 \change[Antoine]{Indded, neurophysiologists typically}{Most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
@@
 \end{figure}
-\note{Je changerai cette figure car nous parlons du système visuel des primates non humains}
 %----------------------------%
@@
 At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
-\note{here we have to make the choice of studying only vertebrates, otherwise there is a huge review to add on insects.
-Gollisch and Meister is the fly retina?
-Maybe find another ref} In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
+In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
 These results have been extended to natural images and show a qualitatively similar behavior.
@@
 A recent work~\citep{muller_stimulus-evoked_2014} presents a method for detecting propagating waves in noisy multichannel recordings.
-Applying this method to single-trial voltage-sensitive dye imaging data, the authors show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
+Applying this method to single-trial voltage-sensitive dye imaging data, the authors show that the stimulus-evoked population response in the primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
 A network model suggests that this reliability is the hallmark of the horizontal fiber network of superficial cortical layers.
-Propagating waves with similar properties occur independently in secondary visual cortex, but maintain precise phase relations with the waves in primary visual cortex.
+Propagating waves with similar properties occur independently in the secondary visual cortex, but maintain precise phase relations with the waves in the primary visual cortex.
 These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatio-temporal frame for further neural interactions.
@@
 The different aspects of information in the data are evaluated by a decoding strategy, highlighting the role of correlations.
-Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from Michael Berry's lab~\citep{berry_spike_2022} and allows performing and validate these hypotheses.
-However, a limit of explaining such data with travelling waves is that in theory a travelling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
+Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from Michael Berry's lab~\citep{berry_spike_2022} and allows \change[Antoine]{performing and validate}{testing in order to validate or falsify} these hypotheses.
+However, a limit of explaining such data with travelling waves is that in theory a \add[Antoine]{cortical} travelling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
 Moreover, recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
@@
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
-Using the right metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.
-Importantly, these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
+\change[Antoine]{Using the right metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.}{Carefully picking the more appropriate metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.} Importantly, these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
 %
@@
 %Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
-Concerning spike timings,~\citet{levakova_review_2015} has reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning.
+Concerning spike timings,~\citet{levakova_review_2015} \change[Antoine]{has}{have} reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning.
 More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
@@
 
-There are multiple solutions which aim at tackling this problem.
+There are multiple \change[Antoine]{solutions}{approaches} which aim at tackling this problem.
 One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
-Basically, it re-arranges neurons in the raster plot based on similarity of activity and applies a deconvolution strategy based on a linear model.
+Basically, it rearranges neurons in the raster plot based on the similarity of their activity and applies a deconvolution strategy based on a linear model.
 The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
@@
 the amount of myelin wrapped around the axon.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
+Indeed, the oligodendrocytes, one of the many glial cells present in the \change[Antoine]{Vertebrates'}{vertebrates'} nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
 Myelinization consists in ``two motions: the wrapping of the leading edge of the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}.
 Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the approx $1~\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_neuron-oligodendrocyte_2021}.
-In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
+\change[Antoine]{In}{On} average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
 This demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
-Thus, in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e.
+Thus, in several occasions, myelin has been identified as \change[Antoine]{a main}{an important} actor in the regulation of conduction velocity in neurons, i.e.
 axonal delay regulation.
@@
 Artificial intelligence has made huge advances in the past decades and deep learning algorithms, nowadays, outperform humans at complex tasks like natural image recognition or abstract strategy board games.
-Yet, machine learning algorithms suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
+Yet, machine learning algorithms suffer from adversarial attacks or a lack of generalization \change[Antoine]{capacity but}{capacity.
+But} their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
 Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019}.
-Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to make energy efficient computations.
+Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to \change[Antoine]{make}{perform} energy efficient computations.
 Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips.
@@
 It was implemented with transistors operating in the sub-threshold domain and responded to asynchronous events, but was not capable of discriminating two different spiking sequences with the same firing rate.
-Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, like it can be observed in biological neurons.
+Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, as it can be observed in biological neurons.
 Then, electronics circuits became more and more bio-realistic and, two decades later, ~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons.
 The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector''.
-Today, many devices are good candidates for implementing event-based algorithms and use the AER.
+Today, many devices are good candidates for implementing event-based algorithms and use the \change[Antoine]{AER}{address event representation (AER)}.
 They can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
@@
 The variability of response of the analog components can be compensated by population coding for responses robust to noise and this phenomenon is also observed in biological neural networks~\citep{boerlin_spike-based_2011}.
-A recent work makes an implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}.
+A recent work makes implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}.
 They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay.
-Note that this problem is also due the algorithm used in this paper and that the analog chips must not suffer from this type of limitations.
+Note that this problem is also due to the algorithm used in this paper and that the analog chips must not suffer from this type of limitation.
 
 Online on-chip learning and computations with delays are still emerging in neuromorphic engineering.
-The technical challenges linked to the development of this type of implementations and the growing interest for delay learning makes advances in this field interesting for the future of computations with precise spatio-temporal motifs.
-While improvements are still to be made, neuromorphic chips seem to be good candidate to efficiently make use of these particular features.
+The technical challenges linked to the development of this type of implementation and the growing interest in delay learning make advances in this field interesting for the future of computations with precise spatio-temporal motifs.
+While improvements are still to be made, neuromorphic chips seem to be a good candidate to efficiently make use of these particular features.
 %

### 2022-12-14 manuscript.tex
Update on Overleaf.

@@
 A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered \change[Antoine]{by Lord Adrian}{at the beginning of the $XX^{th}$ century}~\citep{adrian_impulses_1926}.
-In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons.
+In this study, the frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb.
+In the scope of this article, we focus mainly on these spikes\change[Antoine]{which are brief (about one millisecond) impulses that propagate along the axons of neurons.
 First, these have the particular property of being essentially binary in their amplitude, i.e.
-they are prototypical,
-``\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}''.
-\change[Antoine]{Typically, these are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
-The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.}
-The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.
-\change[Antoine]{Indded, neurophysiologists typically}{Most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
+they are prototypical, ``all-or-none''}{in vertebrate systems.
+They can be described as brief (about one millisecond) and prototypical, i.e.
+``all-or-none'', impulses that propagate along the axons of neurons.}
+ \change[Antoine]{Typically, action potentials are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
+The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except in the retina where neurons communicate with graded potentials}~\citep{goudas_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
+\remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
 This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
@@
 In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
-For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{that rely on the firing rate} by considering a representation based on repetitions of spiking motifs at precise times of occurrence.
-This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.
-Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
+For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{of the rate coding hypothesis that relies only on the firing rate} by considering \change[Antoine]{a representation based on repetitions of spiking motifs at precise times of occurrence.}{representations based on the precise timing of single spikes.} \remove[Antoine]{This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.} Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
 All these findings, and more~\citep{bohte_evidence_2004,dilorenzo_spike_2013}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons.
-A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+\add[Antoine]{This review } A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
 Crucially, validating this hypothesis would also be crucial in our understanding of neural processes.
@@
 Indeed, we may consider that a behavioral response occurring in only $200~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
-Such models were indeed efficiently modeled in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
+Such processes were indeed efficiently reproduced in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
 At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
@@
 % retina
-How may information about the sensory world be encoded in the spiking pattern?
-At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
+\remove[Antoine]{How may information about the sensory world be encoded in the spiking pattern?} At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
 In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
@@
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition or~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-
 However, the performance of SNNs is still lagging compared to that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
@@
 %These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developed in Section~\ref{sec:neuromorphic}).% This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
-To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.
+
+\change[Antoine]{To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.}{Even if technology lags far behind biology, this introduction demonstrates the importance of timing in neural processes and we will further review the role of precise spike timing in neural assemblies.
+We start by reviewing the different hypotheses that aim at deciphering the neural code with spatio-temporal spiking motifs.
+After listing some biological evidence for the use of precise spike timing we review some computational models and neuromorphic technics that add this temporal dimensions to their computations.}
 %

### 2022-12-15 manuscript.tex
Update on Overleaf.

@@
 We start by reviewing the different hypotheses that aim at deciphering the neural code with spatio-temporal spiking motifs.
-After listing some biological evidence for the use of precise spike timing we review some computational models and neuromorphic technics that add this temporal dimensions to their computations.}
+After listing some biological evidence for the use of precise spike timing we review some computational models and neuromorphic technics that add this temporal dimension to their computations.}
 %

### 2022-12-15 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
+Experimentally, there are multiple pieces of evidence that neurons use different strategies to communicate.
+Population coding \citep{Maunsell_Functional_1983}, time-to-first-spike (described above), phase-of-firing code~\cipte{montemurro_phase-of-firing_2008}, correlation coding~\cite{decharms_primary_1996}, sparse coding~\citep{vinje_sparse_2000}, }
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-12-15 manuscript.tex
Update on Overleaf.

--- 2022-12-15 manuscript.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
+There are multiple experimental evidences that neurons use different strategies to communicate.}~\cite{Maunsell_Functional_1983} \add[Antoine]{demonstrate that neurons of visual area MT in macaque monkeys encode motion direction through population activity, offering robustness to single neuron activity fluctuations.
+For rapid image processing, it has been observed that neurons in the salamander retina represent the spatial structure of an image with the relative timing of their first spikes}~\citep{gollisch_rapid_2008}\add[Antoine]{.
+In the primary visual area of macaque monkeys, the phase of firing convey information about naturalistic stimuli that cannot be described by firing rates alone}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{.
+Among a variety of strategies to communicate, so far we have also evidence that neurons can use correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Note also that synchronicity may explain some unintuitive results.
-Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intra-cortical activity.
+Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity.
 However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons~\citep{bruno_cortex_2006}.

### 2022-12-15 manuscript.tex
Update on Overleaf.

--- 2022-12-15 manuscript.tex
@@
 All these findings, and more~\citep{bohte_evidence_2004,dilorenzo_spike_2013}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons.
-\add[Antoine]{This review } A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
 Crucially, validating this hypothesis would also be crucial in our understanding of neural processes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
+There are multiple experimental evidences that neurons use different strategies to communicate.}~\cite{Maunsell_Functional_1983} \add[Antoine]{demonstrate that neurons of visual area MT in macaque monkeys encode motion direction through population activity, offering robustness to single neuron activity fluctuations.
+For rapid image processing, it has been observed that neurons in the salamander retina represent the spatial structure of an image with the relative timing of their first spikes}~\citep{gollisch_rapid_2008}\add[Antoine]{.
+In the primary visual area of macaque monkeys, the phase of firing convey information about naturalistic stimuli that cannot be described by firing rates alone}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{.
+Among a variety of strategies to communicate, so far we have also evidence that neurons can use correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}.
+\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
+Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\cite{Maunsell_Functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
+In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space.
+In all generality, this representation can encompass all the previous ones if the precision in time .
+}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-12-16 manuscript.tex
Update on Overleaf.

@@
 In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space.
-In all generality, this representation can encompass all the previous ones if the precision in time .
-}
+In all generality, this representation can encompass all the previous ones except for rate coding which is not defined locally in time.
+In this section, we choose to described different hypotheses making use of spatio-temporal patterns of spikes that can be propagated among neural assemblies.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The book then demonstrates that neural assemblies could form so-called ``synfire chains,'' that is, showing the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
-More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}.
+More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}: \add[Antoine]{"cells that fire together wire together"}.
 Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
@@
 However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons~\citep{bruno_cortex_2006}.
-\noet{neuron-centric view from the reviewer comment, make some changes to broaden the proposed vision.
-Also change the very specific example given after, or put in context (works by Sophie Dénève or Ballard and Jehee)}
 
@@
 
-Propagating waves in the neuronal response occur in many excitable media and were recently found in neural systems, for instance in the retina or the neocortex.
+Propagating waves in the neuronal response occur in many excitable media and were recently found in neural systems, for instance in the retina~\citep{feller_dynamic_1997} or the neocortex.
 While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries.
@@
 Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}.
-They are mostly used by computational neuroscientist to model brain activity with different levels of abstraction.
+They are mostly used by computational neuroscientists to model brain activity with different levels of abstraction.
 BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}.

### 2022-12-16 src/Izhikevich_2006_Polychronization_longer.py
Create Izhikevich_2006_Polychronization_longer.py

@@ -1,0 +1,221 @@
+'''
+Polychronization: Computation with Spikes
+Eugene M. Izhikevich
+Neural Computation, 2006
+
+See Izhikevich_2006_Polychronization.py for an explanation of this code.
+
+In this version, we use various tricks to make the simulation much faster,
+including inline C++ code. We also save the progress as the simulation runs,
+and analyse it at the end for polychronous groups.
+
+Note that you will need to create two folders data and imgs to run this
+script.
+'''
+import brian_no_units_no_warnings
+from brian import *
+import random
+from scipy import weave
+import cPickle as pickle
+
+try:
+    imported_data = pickle.load(open('data/izhikevich.pickle', 'rb'))
+    defaultclock.t = imported_data['t']
+    print 'Starting from saved progress at time', defaultclock.t
+except IOError:
+    imported_data = {}
+
+dt = defaultclock.dt = 0.1 * ms
+
+M = 100                # number of synapses per neuron
+D = 20 * ms              # maximal conduction delay
+D_min = 1 * ms           # minimal conduction delay
+D_i = 1 * ms             # inhibitory neuron conduction delay
+
+# excitatory neurons
+Ne = 800
+a_e = 0.02 / ms
+d_e = 8 * mV / ms
+s_e = 6 * mV / dt # ms
+# inhibitory neurons
+Ni = 200
+a_i = 0.1 / ms
+d_i = 2 * mV / ms
+s_i = -5 * mV / dt # ms
+# all neurons
+b = 0.2 / ms
+c = -65 * mV
+sm = 10 * mV / dt # ms             # maximal synaptic strength
+N = Ne + Ni
+
+thalamic_input_weight = 20 * mV / dt # ms
+
+eqs = Equations('''
+dv/dt = (0.04/ms/mV)*v**2+(5/ms)*v+140*mV/ms-u+I : volt
+du/dt = a*(b*v-u)                                : volt/second
+a                                                : 1/second
+d                                                : volt/second
+I                                                : volt/second
+''')
+reset = '''
+v = c
+u += d
+'''
+threshold = 30 * mV
+
[…]

### 2022-12-16 manuscript.tex
Update on Overleaf.

@@
 It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
-Note also that synchronicity may explain some unintuitive results.
+\change[Antoine]{}{Synchronicity is also an interesting proposition to explain the relatively weak thalamo-cortical synapses able to drive the cortical neurons.
+Among other explanations }Note also that synchronicity may explain some unintuitive results.
 Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity.

### 2022-12-16 manuscript.tex
Update on Overleaf.

--- 2022-12-16 manuscript.tex
@@
 It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
-Note also that synchronicity may explain some unintuitive results.
+\change[Antoine]{}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons.
+Among different explanations concerning neuromodulation and other mechanisms, a synchronous activity at the synaptic level }Note also that synchronicity may explain some unintuitive results.
 Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity.

### 2022-12-16 src/Diesmann_et_al_1999.py
figures Diesmann_et_al_1999

@@ -66,12 +66,10 @@ plot(Mgp.t/ms, 1.0*Mgp.i/group_size, '.')
 plot([0, duration/ms], np.arange(n_groups).repeat(2).reshape(-1, 2).T, 'k-')
 ylabel('group number')
 yticks(np.arange(n_groups)+.5, labels=range(1, n_groups+1))
-
 xlabel('time (ms)')
+
 if False:
         show()
 else:
-        fname = 'Diesmann_et_al_1999'
         import matplotlib.pyplot as plt
-        plt.savefig(fname + '.pdf', bbox_inches='tight', pad_inches=0.0, backend='pdf')
-        #plt.savefig(fname + '.png', bbox_inches='tight', pad_inches=0.0, dpi=150, backend='agg')
+        plt.savefig('../figures/Diesmann_et_al_1999.pdf', bbox_inches='tight', pad_inches=0.0, backend='pdf')

### 2022-12-16 manuscript.tex

@@
 \change[Antoine]{}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons.
-Among different explanations concerning neuromodulation and other mechanisms, a synchronous activity at the synaptic level }Note also that synchronicity may explain some unintuitive results.
+Among different explanations concerning neuromodulation and other mechanisms, a synchronous activity at the synaptic level may be sufficient to elicit }Note also that synchronicity may explain some unintuitive results.
 Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity.

### 2022-12-16 manuscript.tex
Update on Overleaf.

--- 2022-12-16 manuscript.tex
@@
 \change[Antoine]{}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons.
-Among different explanations concerning neuromodulation and other mechanisms, a synchronous activity at the synaptic level }Note also that synchronicity may explain some unintuitive results.
+Among different explanations concerning neuromodulation and other mechanisms, a synchronous activity at the synaptic level may be sufficient to elicit activity in the cortex.} Note also that synchronicity may explain some unintuitive results.
 Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity.

### 2022-12-16 src/Diesmann_et_al_1999.py
Update Diesmann_et_al_1999.py

@@ -36,8 +36,8 @@ N_pulses = 5
 spike_times = np.zeros(0)
 np.random.seed(42)
 for i_pulse in range(N_pulses):
-        pulse = .5 * (N_pulses-i_pulse) * np.random.randn(N_inputs) + 10 + i_pulse*20
-        spike_times = np.concatenate((spike_times, pulse))
+    pulse = .5 * (N_pulses-i_pulse) * np.random.randn(N_inputs) + 10 + i_pulse*20
+    spike_times = np.concatenate((spike_times, pulse))
 
 spike_times = np.mod(spike_times, 100)
 
@@ -69,7 +69,7 @@ yticks(np.arange(n_groups)+.5, labels=range(1, n_groups+1))
 xlabel('time (ms)')
 
 if False:
-        show()
+    show()
 else:
-        import matplotlib.pyplot as plt
-        plt.savefig('../figures/Diesmann_et_al_1999.pdf', bbox_inches='tight', pad_inches=0.0, backend='pdf')
+    import matplotlib.pyplot as plt
+    plt.savefig('../figures/Diesmann_et_al_1999.pdf', bbox_inches='tight', pad_inches=0.0, backend='pdf')

### 2022-12-16 manuscript.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newcommand{\ResponsReviewerTwo}{
+The authors present a relatively thorough review on neural coding based on spike patterns in populations of neurons, from biological, theoretical and neuromorphic engineering perspective.
+I find the review mostly fine, with some omissions that, when corrected, should enhance it.
+
+My main concern is the overall reliance on spike motifs as a coding mechanism.
+The authors should list alternatives and compare to spike motifs, at least cursory.
+A minor concern is the focus on mylenation-based delay learning, ignoring the obvious length-based delay adjustments.
+
+I have commented heavily within the manuscript, and am attaching it for the authors to use.
+I tried to delineate wording suggestions from more significant critiques.
+I am also listing the bigger concerns below.
+
+Specific issues:
+
+- not referencing much of insect work, with a few exceptions that are put in mammalian context.
+OK to do just timing in vertebrate systems (as birds are used), but should be stated clearly, and cited work should comport to that constraint (e.g., cite rat olfactory, not locust olfactory).
+Should the authors wish to include invertebrate work, there is a LOT more to review.
+- some references to original research are missing, while often much older ones are used (and not reviews).
+- missing important lines of work on spike timing. In particular, Aurel Lazar's TEM work. 
+- missing important lines of work on dendritic processing: Mel, Spruston.
+- discounting axonal length as delay mechanism, focus on myelin effects too high.
+- singular focus on synfire chain/polychronicity-like mechanisms.
+Alternative options, like Bayesian spiking (Deneve), probabilistic ensemble motifs (Ballard) and probabilistic population coding (Pouget and Latham), which do not rely on exact motifs of spikes, are not discussed.
+- there are also some discrepancies - are SNN-s differentiable or not.
+}
 

### 2022-12-18 manuscript.tex
Update on Overleaf.

@@
 
-\newcommand{\ResponsReviewerTwo}{
+\newcommand{\ResponseReviewerTwo}{
 The authors present a relatively thorough review on neural coding based on spike patterns in populations of neurons, from biological, theoretical and neuromorphic engineering perspective.
@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0002-3107-4788} % AmG % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0002-3107-4788} % AnG % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
@@
 It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
-\change[Antoine]{}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons.
-Among different explanations concerning neuromodulation and other mechanisms, a synchronous activity at the synaptic level may be sufficient to elicit activity in the cortex.} Note also that synchronicity may explain some unintuitive results.
+\change[Antoine]{Note also that synchronicity may explain some unintuitive results.
 Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity.
-However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons~\citep{bruno_cortex_2006}.
-
-Theoretically, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
+However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons.
+Among different explanations including travelling waves}~\citep{ben-yishai_traveling_1997}\add[Antoine]{, a synchronous activity at the synaptic level may be sufficient to elicit activity in the cortex.}~\citep{bruno_cortex_2006}.
+
+Theoretically\add[Antoine]{, even if vertebrate's neural system is not likely to be modeled only by such deterministic connectivity}~\citep{deneve_bayesian_2004, ballard_dual_2011}, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
 Each neuron of one group is connected by an excitatory synapse to the next.
@@
 Using intracellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is, a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
-In particular, it was shown that the network of the so-called horizontal connections (that is, connections within a cortical area) are typically not myelinated and that this latency basin would be determined by the propagation speed within that area.
-For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions and may underlay anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
+In particular, it was shown that the network of the so-called horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within that area.
+For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions.
+Such interactions may underlay anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
 The underlying process could be the emergence of propagating waves on the surface of the cortex.
 
-Propagating waves in the neuronal response occur in many excitable media and were recently found in neural systems, for instance in the retina~\citep{feller_dynamic_1997} or the neocortex.
+Propagating waves in the neuronal response occur in many excitable media and were found in neural systems, for instance in the retina~\citep{feller_dynamic_1997} or the neocortex~\citep{bienenstock_model_1995}.
 While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries.
@@
 Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from Michael Berry's lab~\citep{berry_spike_2022} and allows \change[Antoine]{performing and validate}{testing in order to validate or falsify} these hypotheses.
-However, a limit of explaining such data with travelling waves is that in theory a \add[Antoine]{cortical} travelling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
-Moreover, recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
-~\cite{miller_visual_2014} shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
-It proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
+However, \remove[Antoine]{a limit of explaining such data with travelling waves is that} in theory, a \add[Antoine]{cortical} travelling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
+Recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
+~\cite{miller_visual_2014} show that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+Experiments are performed using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity.
+The study proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
 Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions.
@@
 %We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
-Going further, researchers have found that  precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \textit{in vivo} and \textit{in vitro}.
+Going further, researchers have found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \textit{in vivo} and \textit{in vitro}.
 These patterns repeat after minutes, maintaining millisecond accuracy.
@@
 Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
-This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996} or touch~\citep{johansson_first_2004}.
+This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{cleland_construction_2014, kashiwadani_synchronized_1999, rinberg_speed-accuracy_2006} or touch~\citep{johansson_first_2004}.
 In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
@@
 In that protocol, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}).
+They are stereotyped and robust since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}).
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
@@
 There are different solutions to provide with a distance between two given spike trains.
-A known measure is the Victor-Purpura distance, which  overcomes inconsistencies experienced with a simple firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
-In a later study,~\citet{van_rossum_novel_2001} have defined a distance that uses a time constant as a parameter.
-Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.
-Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.
-Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure which is adapted to the event-based nature of spiking signals.
+A known measure is the Victor-Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
+\change[JN]{In a later study,van_rossum_novel_2001 have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
+\change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.
+Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure \change[JN]{which}{.
+An interesting conclusion since cross-correlation measure is} which is adapted to the event-based nature of spiking signals.
 These observations lead to the intuition that each distance may be as well  as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
@@
 %Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
-Concerning spike timings,~\citet{levakova_review_2015} \change[Antoine]{has}{have} reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning.
-More broadly, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
+Concerning spike timings,~\citet{levakova_review_2015} \change[Antoine]{has}{have} \change[JN]{reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning}{reviewed existing methods for estimating the latency of neural responses that include Bayesian binning.}.
+\change[JN]{More broadly }{Alternatively }, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
 This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}.
@@
 %   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
+\add[JN]{Whether it is the distance between two given spike trains or the spike timings the complexity and the diversity of the methods used to measure them are witnesses of the growing interest of the integration of these measures in the understanding of the neural code.
+One of the steps to test their potential usefulness is then to scale these methods to larger amounts of data.}
 
@@
 
-However, it was not until the study of the interaural time difference (ITD) by~\citet{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
+However, it was not until the study of the interaural time difference (ITD) by \remove[Amélie]{gerstnerneuronal1996}~\citet{carr_circuit_1990} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
 This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.~\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
-In the mouse somatosensory cortex,~\citep{camon_timing_2019}  found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.
-A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
+\add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\cite{jeffress}.
+\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\cite{konishi}.
+According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.
+\add[Amélie]{This assertion confused the mean interspike interval, i.e.
+how often a neuron can fire, and the specific spike time, i.e.
+how precisely a spike can be generated.
+However, it has nevertheless contributed to the recognition of the importance of time in various biological mechanisms.} \remove[Amélie]{penacochlear2001 conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
+In the mouse somatosensory cortex, camontiming2019 found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
 the temporal disparity between left and right cochlear loci in the owl's brain.
@@
 The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977} --- instead of solely the axon's length as was previously thought.
+\add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implement a ``pure delay line''.} \add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
+However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\citep{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.
+This mechanism adds to the axonal length delay, which was previously thought sole influence on the conduction velocity due to its anatomical soundness --- as}~\citet{seidl_mechanisms_2010}\add[Amélie]{ experimentally demonstrated, this mechanism by itself is not sufficient to explain the biological functionality but should be added to the one of axonal length delay.}
 
@@
 % The importance of myelination
-\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e.
-the amount of myelin wrapped around the axon.
+\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to \add[Amélie]{both the axon diameter by itself and the diameter of the axon wrapped in myelin}{the axon diameter, i.e.
+the amount of myelin wrapped around the axon}.
 Indeed, the oligodendrocytes, one of the many glial cells present in the \change[Antoine]{Vertebrates'}{vertebrates'} nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
@@
 % myelination for axonal delay, but what about dendritic delay ? 
-The previous paragraphs present the biological mechanisms behind axonal delay, regulated by myelination.
-However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
-Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
-The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.
+The previous paragraphs present the biological mechanisms behind the axonal delay, regulated by myelination.
+However, the dendritic delay \add[Amélie]{as well as the axonal length delay (see }~\cite{seidl_mechanisms_2010}) seem to play an equally important role in the precise timing within a sequence of spikes.
+Dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
+Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}.
+\cite{mel_dendrite} \add[Amélie]{highlights the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
+Dendritic spiking involvement in synaptic potentiation following active backpropagation into dendrites was experimentally uncovered using calcium imaging to highlight dendritic calcium entry allowing for long-tem potentiation}~\citep{golding}.
+~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
+The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree and relies on the dendritic calcium influx moderation by NMDA receptors.
 

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 %   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
-\add[JN]{Whether it is the distance between two given spike trains or the spike timings the complexity and the diversity of the methods used to measure them are witnesses of the growing interest of the integration of these measures in the understanding of the neural code.
+\add[JN]{Whether it is the distance between two given spike trains or a comparison of the spike timings, the complexity and the diversity of the methods used to measure them are witnesses of the growing interest of the integration of these measures in the understanding of the neural code.
 One of the steps to test their potential usefulness is then to scale these methods to larger amounts of data.}
@@
 The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviors allow us to put these doubts to rest.
-In the following paragraphs we will successively describe the first biological observations of biological delay learning, identify myelinization as an important actor and finally study this phenomenon at the molecular level.
+In the following paragraphs we will successively describe the first biological observations of delay learning, identify myelinization as an important actor and finally study this phenomenon at the molecular level.
 
@@
 This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-\add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\cite{jeffress}.
-\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\cite{konishi}.
+\add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\cite{jeffress_place_1948}.
+\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\cite{konishi_coding_2003}.
 According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.
@@
 The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-\add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implement a ``pure delay line''.} \add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
-However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\citep{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.
+\add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.} \add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
+However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\cite{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.
 This mechanism adds to the axonal length delay, which was previously thought sole influence on the conduction velocity due to its anatomical soundness --- as}~\citet{seidl_mechanisms_2010}\add[Amélie]{ experimentally demonstrated, this mechanism by itself is not sufficient to explain the biological functionality but should be added to the one of axonal length delay.}
@@
 % The importance of myelination
-\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to \add[Amélie]{both the axon diameter by itself and the diameter of the axon wrapped in myelin}{the axon diameter, i.e.
+\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to \change[Amélie]{both the axon diameter by itself and the diameter of the axon wrapped in myelin}{the axon diameter, i.e.
 the amount of myelin wrapped around the axon}.
@@
 Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}.
-\cite{mel_dendrite} \add[Amélie]{highlights the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
-Dendritic spiking involvement in synaptic potentiation following active backpropagation into dendrites was experimentally uncovered using calcium imaging to highlight dendritic calcium entry allowing for long-tem potentiation}~\citep{golding}.
+\cite{mel_synaptic_2017} \add[Amélie]{highlights the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
+Dendritic spiking involvement in synaptic potentiation following active backpropagation into dendrites was experimentally uncovered using calcium imaging to highlight dendritic calcium entry allowing for long-tem potentiation}~\citep{golding_dendritic_2002}.
 ~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
@@
 Spiking Neural Networks (SNNs)~\citep{maass_networks_1997} are natural candidates to use these precise temporal patterns in the brain.
-The approach which is currently most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.
+The approach which is currently most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021} One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.
 In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error's gradient.
-Since the activation function of a spiking neuron is not differentiable, a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-This approach is quite successful, and SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
-The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
-However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
+\change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient}~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
+SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
+\remove[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.}{Even if improvements in terms of efficiency} However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
 There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
+To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
 In particular, we will explore different existing architectures which are able to overcome the diversity of input presynaptic patterns and learn to detect stable spiking motifs, that is, volleys of spikes which are stable up to a certain onset time (see Figure~\ref{fig:THC}).

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
-\remove[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.}{Even if improvements in terms of efficiency} However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
+\remove[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.} However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
 There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
-\remove[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.} However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
-There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+\change[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
+However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 \change[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
-However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{So far, this approach does not outperform classical architectures both in term of training efficiency and performances~\citep{davies_loihi_2018}.
+However, they remain the best candidates to reproduce } There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{So far, this approach does not outperform classical architectures both in term of training efficiency and performances~\citep{davies_loihi_2018}.
-However, they remain the best candidates to reproduce } There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+However, they remain the best candidates to reproduce biological neural systems and their efficiency both in terms of accuracy, speed and energy consumption.} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.

### 2022-12-19 manuscript.tex
Update on Overleaf.

--- 2022-12-19 manuscript.tex
@@
 \change[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
-However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{So far, this approach does not outperform classical architectures both in term of training efficiency and performances}~\citep{davies_loihi_2018}\add[Antoine]{.
+However, they remain the best candidates to reproduce biological neural systems and their capacities both in terms of accuracy, speed and energy consumption.} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 \change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient}~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the N-MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
+SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the \remove[Antoine]{N-}MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
 \change[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
@@
 However, they remain the best candidates to reproduce biological neural systems and their capacities both in terms of accuracy, speed and energy consumption.} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
+To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
 In particular, we will explore different existing architectures which are able to overcome the diversity of input presynaptic patterns and learn to detect stable spiking motifs, that is, volleys of spikes which are stable up to a certain onset time (see Figure~\ref{fig:THC}).

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
-It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}.
+\change[Antoine]{It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons}{It was shown that spike timing (STDP) has an impact on the development of synaptic efficacy for many kinds of neurons}~\citep{caporale_spike_2008}.
 Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.

### 2022-12-19 manuscript.tex
in sync

@@
 This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \change[LP]{}{such as the synthetic data for synfire chains (see Figure~\ref{fig:Diesmann_et_al_1999_spade}),} such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
 
@@
 It uses the prior that there is only one spike per pattern.
-Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data (see Figure~\ref{fig:SPOTDisClust}).
+Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data.
+% (see Figure~\ref{fig:SPOTDisClust}).
 The limits of this method are that it is computationally complex, block-based and strictly specialized for the task at hand.
@@
 %----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
+\caption{\textbf{Detecting motifs using SPADE.} We use the SPADE algorithm~\citep{stella_3d-spade_2019} on the data generated in  Figure~\ref{fig:diesman} to label different precise spike motifs which are denoted by different colors.
+Spikes are labelled by the same color when they belong to the same motif.
+As the synfire chain gets more precise, the detected patterns grow in size.}\label{fig:Diesmann_et_al_1999_spade}
+\end{figure}
+%----------------------------%
+
+%%----------------------------%
+\remove[LP]{
 \begin{figure}
@@
 %----------------------------%
+}
 %

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 A known measure is the Victor-Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
-\change[JN]{In a later study,van_rossum_novel_2001 have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
+\change[JN]{In a later study, van_rossum_novel_2001 have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
 \change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.
@@
 This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \change[LP]{}{such as the synthetic data for synfire chains (see Figure~\ref{fig:Diesmann_et_al_1999_spade}),} such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[LP]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
 
@@
 However, they remain the best candidates to reproduce biological neural systems and their capacities both in terms of accuracy, speed and energy consumption.} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
+To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) \change[Antoine]{to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized}{to a detector of precise spiking motifs}~\citep{paugam-moisy_computing_2012}.
 In particular, we will explore different existing architectures which are able to overcome the diversity of input presynaptic patterns and learn to detect stable spiking motifs, that is, volleys of spikes which are stable up to a certain onset time (see Figure~\ref{fig:THC}).
@@
 As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-A notable exception of a SNN using precise spike timings is the \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
-This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
+Notable exceptions of a SNN using precise spike timings are the \add[Antoine]{time encoding machine by}~\cite{lazar_time_2004} and \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
+\add[Antoine]{In this section we focus on the polychronization model} based on a random recurrent model of spiking neurons including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
 \change[Antoine]{It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons}{It was shown that spike timing (STDP) has an impact on the development of synaptic efficacy for many kinds of neurons}~\citep{caporale_spike_2008}.

### 2022-12-19 manuscript.tex
in sync

--- 2022-12-19 manuscript.tex
@@
 
-%%----------------------------%
-\remove[LP]{
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
-\caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm.
+% %%----------------------------%
+% \remove[LP]{
+% \begin{figure}
+% \centering
+% \includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
+% \caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm.
 \textbf{(A)} Structure of five ``ground-truth'' patterns (\ldots).
@@
 (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
-\end{figure}
-%----------------------------%
-}
+% \end{figure}
+% %----------------------------%
+% }
 %

### 2022-12-19 manuscript.tex
in sync

@@
 \usepackage{color}
-\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+\newcommand{\mynote}[1]{{\sethlcolor{yellow}\hl{#1}}}
 \usepackage{trackchanges}
@@
 % \centering
-% \note{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
+% \mynote{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
 % \caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
@@
 A known measure is the Victor-Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
-\change[JN]{In a later study,van_rossum_novel_2001 have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
+\change[JN]{In a later study, Van Rossum (2001) have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
 \change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.

### 2022-12-19 manuscript.tex
in sync

@@
 A known measure is the Victor-Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
-\change[JN]{In a later study, van_rossum_novel_2001 have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
+\change[JN]{In a later study, Van Rossum (2001) have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
 \change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.

### 2022-12-19 manuscript.tex
in sync

@@
 In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error's gradient.
-\change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient}~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
+\change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
 SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the \remove[Antoine]{N-}MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.

### 2022-12-19 manuscript.tex
in sync: update numbers in fig:thorpe

@@
 \usepackage{trackchanges}
-
+%\usepackage[finalnew]{trackchanges}
 \addeditor{Antoine}
@@
 \Title{Precise spiking motifs in neurobiological and neuromorphic data}
-
+%
 % MDPI internal command: Title for citation in the left column
 \TitleCitation{Precise spiking motifs}
-
+%
 % Author Orchid ID: enter ID or remove command
@@
 \newcommand{\orcidauthorF}{0000-0002-9536-010X} % LP
-
+%
 % Authors, for the paper (add full first names)
@@
 and Laurent Perrinet $^{1,}$\orcidF{}*}
-
+%
 %\longauthorlist{yes}
-
+%
 % MDPI internal command: Authors, for metadata in PDF
 \AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent Perrinet}
-
+%
 % MDPI internal command: Authors, for citation in the left column
@@
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
-
+%
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
@@
 }
-
+%
 % Contact information of the corresponding author
 \corres{Correspondence: laurent.perrinet@univ-amu.fr}
-
+%
 % Current address and/or shared authorship
@@
 % The commands \thirdnote{} till \eighthnote{} are available for further notes
-
+%
 %\simplesumm{} % Simple summary
-
+%
 %\conference{} % An extended version of a conference paper
-
+%
 % Abstract (Do not insert blank lines, i.e. \\) 
@@
 }
-
+%
 % Keywords
 \keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; spiking motifs; polychronization.)} 
-
+%
 % The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
@@
 %\JEL{}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \caption{{\bf Latency of the different processing steps along the visual pathway.} Though this system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
-Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency.
-In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-80~\ms$).
-Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $120-140~\ms$), and then the prefrontal cortex (PFC) which modulates decision making and the Motor cortex (MC, $140-190~\ms$) which mediates a motor action which is then relayed through the spinal cord to trigger finger muscles, with latencies about $200-350~\ms$.}\label{fig:thorpe}
+Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
+In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-90~\ms$).
+Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $150~\ms$), and then the prefrontal cortex (PFC, $180~\ms$) which modulates decision making and the Motor cortex (MC, $220~\ms$) which mediates a motor action which is then relayed through the spinal cord to trigger finger muscles, with latencies about $280-400~\ms$.}\label{fig:thorpe}
 \end{figure}

### 2022-12-19 manuscript.tex
in sync: citep and citet

@@
 \add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
-Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\cite{Maunsell_Functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
+Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\citep{Maunsell_Functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
 In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space.
@@
 Recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
-~\cite{miller_visual_2014} show that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+~\citet{miller_visual_2014} show that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
 Experiments are performed using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity.
@@
 This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-\add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\cite{jeffress_place_1948}.
-\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\cite{konishi_coding_2003}.
+\add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\citep{jeffress_place_1948}.
+\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
 According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.
@@
 \add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.} \add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
-However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\cite{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.
+However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\citep{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.
 This mechanism adds to the axonal length delay, which was previously thought sole influence on the conduction velocity due to its anatomical soundness --- as}~\citet{seidl_mechanisms_2010}\add[Amélie]{ experimentally demonstrated, this mechanism by itself is not sufficient to explain the biological functionality but should be added to the one of axonal length delay.}
@@
 The previous paragraphs present the biological mechanisms behind the axonal delay, regulated by myelination.
-However, the dendritic delay \add[Amélie]{as well as the axonal length delay (see }~\cite{seidl_mechanisms_2010}) seem to play an equally important role in the precise timing within a sequence of spikes.
+However, the dendritic delay \add[Amélie]{as well as the axonal length delay (see }~\citep{seidl_mechanisms_2010}) seem to play an equally important role in the precise timing within a sequence of spikes.
 Dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
 Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}.
-\cite{mel_synaptic_2017} \add[Amélie]{highlights the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
+\citet{mel_synaptic_2017} \add[Amélie]{highlights the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
 Dendritic spiking involvement in synaptic potentiation following active backpropagation into dendrites was experimentally uncovered using calcium imaging to highlight dendritic calcium entry allowing for long-tem potentiation}~\citep{golding_dendritic_2002}.
@@
 As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-Notable exceptions of a SNN using precise spike timings are the \add[Antoine]{time encoding machine by}~\cite{lazar_time_2004} and \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
+Notable exceptions of a SNN using precise spike timings are the \add[Antoine]{time encoding machine by}~\citet{lazar_time_2004} and \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
 \add[Antoine]{In this section we focus on the polychronization model} based on a random recurrent model of spiking neurons including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.

### 2022-12-19 manuscript.tex
in sync: \ResponseReviewerTwo

@@
 %%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \newcommand{\ResponseReviewerTwo}{
-The authors present a relatively thorough review on neural coding based on spike patterns in populations of neurons, from biological, theoretical and neuromorphic engineering perspective.
+We would like to thank reviewer 2 for his very detailed response which were essential in providing this new revision.
+
+> The authors present a relatively thorough review on neural coding based on spike patterns in populations of neurons, from biological, theoretical and neuromorphic engineering perspective.
 I find the review mostly fine, with some omissions that, when corrected, should enhance it.
 
-My main concern is the overall reliance on spike motifs as a coding mechanism.
+Thanks for the encouraging comment, we hope that in its present form, the result is even more useful to the community.
+
+> My main concern is the overall reliance on spike motifs as a coding mechanism.
 The authors should list alternatives and compare to spike motifs, at least cursory.
@@
 
-I have commented heavily within the manuscript, and am attaching it for the authors to use.
+We have more progressively commented different alternatives for coding mechanisms in the beginning of section 2, highlighting their relative benefits.
+
+> I have commented heavily within the manuscript, and am attaching it for the authors to use.
 I tried to delineate wording suggestions from more significant critiques.
@@
 
-Specific issues:
-
-- not referencing much of insect work, with a few exceptions that are put in mammalian context.
+Many thanks for providing the PDF of the manuscript with annotations, this was immensely useful to annotate and improve our revision.
+To respond to all these points, we have constructed a tracked changes manuscript in which we respond (by a change, removal or addition) to the reviewer's comments.
+
+> Specific issues:
+> - not referencing much of insect work, with a few exceptions that are put in mammalian context.
 OK to do just timing in vertebrate systems (as birds are used), but should be stated clearly, and cited work should comport to that constraint (e.g., cite rat olfactory, not locust olfactory).
 Should the authors wish to include invertebrate work, there is a LOT more to review.
-- some references to original research are missing, while often much older ones are used (and not reviews).
-- missing important lines of work on spike timing. In particular, Aurel Lazar's TEM work. 
-- missing important lines of work on dendritic processing: Mel, Spruston.
-- discounting axonal length as delay mechanism, focus on myelin effects too high.
-- singular focus on synfire chain/polychronicity-like mechanisms.
+
+Indeed, our focus was implicitly on mammals.
+worse, we were sometimes using evidence form insects without highlighting the differences.
+We believe that the coding principle we put forward is general enough to cover both orders.
+We also tried to keep the review at a reasonable length to keep the balance between computational and biological evidence.
+Therefore, while we keep our focus on mammals, we have been careful now to mention the order of animals.
+
+> - some references to original research are missing, while often much older ones are used (and not reviews).
+> - missing important lines of work on spike timing. In particular, Aurel Lazar's TEM work. 
+> - missing important lines of work on dendritic processing: Mel, Spruston.
+
+These critical references were indeed missing and we have added them now to the manuscript. Thanks for pointing out these omissions.
+
+> - discounting axonal length as delay mechanism, focus on myelin effects too high.
+
+This was corrected in section 4, providing a more balanced review.
+
+> - singular focus on synfire chain/polychronicity-like mechanisms.
 Alternative options, like Bayesian spiking (Deneve), probabilistic ensemble motifs (Ballard) and probabilistic population coding (Pouget and Latham), which do not rely on exact motifs of spikes, are not discussed.
-- there are also some discrepancies - are SNN-s differentiable or not.
+
+Indeed, we had made that choice to
+
+> - there are also some discrepancies - are SNN-s differentiable or not.
+
+We have corrected these discrepancies. Thanks again for all these useful comments.
+
 }

### 2022-12-19 manuscript.tex
Update on Overleaf.
in sync: spade

@@
 This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[LP]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[Laurent]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
 
@@
 \includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
-\caption{\textbf{Detecting motifs using SPADE.} We use the SPADE algorithm~\citep{stella_3d-spade_2019} on the data generated in  Figure~\ref{fig:diesman} to label different precise spike motifs which are denoted by different colors.
-Spikes are labelled by the same color when they belong to the same motif.
-As the synfire chain gets more precise, the detected patterns grow in size.}\label{fig:Diesmann_et_al_1999_spade}
+\caption{\textbf{Detecting motifs using SPADE.} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}.
+This allowed to label different precise spike motifs which are denoted by different colors.
+Spikes belonging to the same motif have the same color.
+%As the synfire chain gets more precise, the detected patterns grow in size.
+}\label{fig:Diesmann_et_al_1999_spade}
 \end{figure}

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 
-Indeed, our focus was implicitly on mammals.
-worse, we were sometimes using evidence form insects without highlighting the differences.
+Indeed, our focus was implicitly on mammals while we were sometimes using evidence from insects without highlighting the differences.
 We believe that the coding principle we put forward is general enough to cover both orders.
@@
 
-Indeed, we had made that choice to
+Indeed, we had made that choice to study precise temporal motifs.
+In the new manuscript, we mention these alternative options briefly even if we keep the focus on deterministic spiking motifs for this review.
 

### 2022-12-19 manuscript.tex
in sync: \acknowledgments

@@
 
-Indeed, we had made that choice to study precise temporal motifs.
-In the new manuscript, we mention these alternative options briefly even if we keep the focus on deterministic spiking motifs for this review.
+Indeed, we had made that choice to
 
@@
 
-\acknowledgments{We would like to thank Jean-Nicolas Jérémie, George Abitbol and Hugo Ladret for useful discussions while writing this review.}
+\acknowledgments{We would like to thank George Abitbol and Hugo Ladret for useful feedbacks while writing this review.
+Many thanks to Simon Thorpe for the useful discussions and references on the latency of the visual system in human and non-human primates.}
 

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
  \change[Antoine]{Typically, action potentials are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
-The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except in the retina where neurons communicate with graded potentials}~\citep{goudas_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
+The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960as_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
 \remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.

### 2022-12-19 manuscript.tex
in syncl

@@
 % LP received funding 
-LP received support from the french government under the France 2030 investment plan, as part of the Initiative d’Excellence d’Aix-Marseille Université – A*MIDEX ” AMX-21-RID-025 ``\href{https://laurentperrinet.github.io/grant/polychronies/}{Polychronies}''.}
+LP received support from the french government under the France 2030 investment plan, as part of the Initiative d’Excellence d’Aix-Marseille Université – A*MIDEX number AMX-21-RID-025 ``\href{https://laurentperrinet.github.io/grant/polychronies/}{Polychronies}''.}
 

### 2022-12-19 manuscript.tex
Update manuscript.tex

@@
 
-Propagating waves in the neuronal response occur in many excitable media and were found in neural systems, for instance in the retina~\citep{feller_dynamic_1997} or the neocortex~\citep{bienenstock_model_1995}.
+Propagating waves in the neuronal response occur in many excitable media and were found in neural systems, for instance in the retina\note[Antoine]{added ref to Feller}~\citep{feller_dynamic_1997} or the neocortex~\citep{bienenstock_model_1995}.
 While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries.
@@
 In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
-For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process.
+For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process\note[Laurent]{adding reference}~\citep{simoncelli_characterization_2003}.
 As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves, that is, its selectivity to a range of features~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.

### 2022-12-19 manuscript.tex
Update manuscript.tex

@@
 The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995}) and embedded with recurrent dynamics.
-In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the publicly available \href{http://crcns.org/data-sets/vc/pvc-11}{V1-dataset} which allows supervising the model with the input being the movie and the output the spikes recorded.
+In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the \note[Laurent]{now citing the URL}publicly available V1 dataset~\citep{kohn_utah_2016} which allows supervising the model with the input being the movie and the output the spikes recorded.
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
@@
 A known measure is the Victor-Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
-\change[JN]{In a later study, Van Rossum (2001) have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citet{van_rossum_novel_2001}.
+\change[JN]{In a later study, Van Rossum (2001) have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citep{van_rossum_novel_2001}.
 \change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.
@@
 Basically, it rearranges neurons in the raster plot based on the similarity of their activity and applies a deconvolution strategy based on a linear model.
-The model is \href{https://github.com/MouseLand/rastermap}{openly accessible} and has led to important discoveries.
+\add[Laurent]{Yet this method was mainly tested on calcium imaging data which is known to add some imprecision to the timing of the original spiking activity.} The model is \note[Laurent]{now citing the URL}openly accessible~\citep{stringer_mouselandrastermap_2020} and has led to important discoveries.
 In~\citet{stringer_spontaneous_2019,stringer_high-dimensional_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.

### 2022-12-19 manuscript.tex
Update manuscript.tex

@@
 To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
-They also introduce learnable time-warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
+\change[Laurent]{They also introduce}{As originally introduced by}~\citep{kass_statistical_2005}, they use learnable time-warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
 At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 
-Indeed, our focus was implicitly on mammals while we were sometimes using evidence from insects without highlighting the differences.
+Indeed, our focus was implicitly on vertebrates but we cited some works about insect.
+As mentioned by the reviewer, adding researches done on insects would increase too much the size of the article.
+We decide to focus on vertebrates and mention this choice in the new manuscript.
+%Indeed, our focus was implicitly on mammals while we were sometimes using evidence from insects without highlighting the differences.
 We believe that the coding principle we put forward is general enough to cover both orders.
@@
 \add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
-Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\citep{Maunsell_Functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
+Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\citep{Maunsell_Functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase--firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
 In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization, LP; writing---original draft preparation, LP.; writing---review and editing, all authors; supervision, LP; funding acquisition, LP and JM.
+\authorcontributions{Conceptualization, LP; writing---original draft preparation, all authors; writing---review and editing, all authors; supervision, LP; funding acquisition, LP and JM.
 All authors have read and agreed to the published version of the manuscript.}
@@
 
-\acknowledgments{We would like to thank George Abitbol and Hugo Ladret for useful feedbacks while writing this review.
+\acknowledgments{We would like to thank Hugo Ladret and Frdéric Chavane for valuable feedback while writing this review.
 Many thanks to Simon Thorpe for the useful discussions and references on the latency of the visual system in human and non-human primates.}

### 2022-12-19 2022-12-19_coverletter.tex
Update on Overleaf.

--- 2022-11-16 2022-11-16_coverletter.tex
@@
 %----------------------------------------------------------------------------------------
+Many thanks for handling our manuscript entitled ``Precise spiking motifs in neurobiological and neuromorphic data'' for your consideration as a review paper in \emph{Brain Sciences}.
+We are very grateful to the reviewers for their very insightful remarks and we have now responded to all their remarks.
+Note that in agreement with the editorial office, we have added Jean-Nicolas Jérémie as an author, as his contributions were substantial during the process of writing this reviews.
 
-Following your invitation to submit an article to Brain Sciences, please find enclosed a manuscript entitled ``Precise spiking motifs in neurobiological and neuromorphic data'' for your consideration as a review paper in \emph{Brain Sciences}.
 All the authors have been involved with the work, have approved the manuscript and agreed to its submission.
 
-This is an invited Feature Paper, with a full waiver on the APC. In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source
+This is an invited Feature Paper, with a full waiver on the APC. %In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source
 

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
  \change[Antoine]{Typically, action potentials are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
-The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960as_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
+The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
 \remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
@@
 They would benefit in particular to a new generation of cameras called Silicon Retinas}~\citep{gallego_event-based_2022} (see Section~\ref{sec:neuromorphic}).
-%These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developed in Section~\ref{sec:neuromorphic}).% This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
+\remove{These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation}%~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developed in Section~\ref{sec:neuromorphic}).This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
 
@@
 A recent work~\citep{muller_stimulus-evoked_2014} presents a method for detecting propagating waves in noisy multichannel recordings.
-Applying this method to single-trial voltage-sensitive dye imaging data, the authors show that the stimulus-evoked population response in the primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
+Applying this method to single-trial voltage-sensitive dye imaging data\note{The reviewer is right to point that calcium imaging may not be the best tool to capture precise spiking motif as both spatial and temporal resolution are not that great.
+We argue that few imaging techniques can capture the precise spike location, even good spike sorting algorithms.
+Also, here we study a propagation of activity defining a spatio-temporal pattern that may not have the same precision as some spiking motifs we describe later.}, the authors show that the stimulus-evoked population response in the primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
 A network model suggests that this reliability is the hallmark of the horizontal fiber network of superficial cortical layers.
@@
 the amount of myelin wrapped around the axon}.
-Indeed, the oligodendrocytes, one of the many glial cells present in the \change[Antoine]{Vertebrates'}{vertebrates'} nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
 Myelinization consists in ``two motions: the wrapping of the leading edge of the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}.
@@
 The previous paragraphs present the biological mechanisms behind the axonal delay, regulated by myelination.
-However, the dendritic delay \add[Amélie]{as well as the axonal length delay (see }~\citep{seidl_mechanisms_2010}) seem to play an equally important role in the precise timing within a sequence of spikes.
+However, the dendritic delay \add[Amélie]{as well as the axonal delay (see }~\citep{seidl_mechanisms_2010}) seem to play an equally important role in the precise timing within a sequence of spikes.
 Dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
@@
 As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-Notable exceptions of a SNN using precise spike timings are the \add[Antoine]{time encoding machine by}~\citet{lazar_time_2004} and \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
+Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time encoding machine by}~\citet{lazar_time_2004} and \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
 \add[Antoine]{In this section we focus on the polychronization model} based on a random recurrent model of spiking neurons including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
@@
 But} their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019}.
-Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to \change[Antoine]{make}{perform} energy efficient computations.
+Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019, javanshir_advancements_2022}.
+Because the reduced energy consumption of biological networks can be explained \add{in part} by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to \change{make}{perform} energy efficient computations.
 Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
  \change[Antoine]{Typically, action potentials are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
-The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
+The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except notably in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
 \remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role and that neurons \change[Antoine]{may be as well integrators than synchrony detectors}{may be synchrony detectors as well as integrators}~\citep{abeles_role_1982}.
-We will investigate this very hypothesis in this review.
+\add[Laurent]{In particular, it is possible that the minute arrangement of temporal delays between neurons may provide a computational advantage.} We will investigate this very hypothesis in this review.
 %Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
@@
 % \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
-\caption{{\bf Latency of the different processing steps along the visual pathway.} Though this system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
+\caption{{\bf Latency of the different processing steps along the human visual pathway.} Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
 Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
-In order, the retina is first activated ($20-40~\ms$) then the thalamus and the primary visual cortex (V1, $60-90~\ms$).
-Visual information used for object recognition follows the temporal lobe to reach the infero-temporal area (IT, $150~\ms$), and then the prefrontal cortex (PFC, $180~\ms$) which modulates decision making and the Motor cortex (MC, $220~\ms$) which mediates a motor action which is then relayed through the spinal cord to trigger finger muscles, with latencies about $280-400~\ms$.}\label{fig:thorpe}
+In order, the retina is first activated ($20-40~\ms$), then the thalamus and the primary visual cortex (V1, $60-90~\ms$).
+This visual information  projects to the temporal lobe to reach the infero-temporal area (IT, $150~\ms$) for object recognition.
+It then reaches the prefrontal cortex (PFC, $180~\ms$) which modulates decision making to evoke the Motor Cortex (MC, $220~\ms$) which may mediate an action.
+This is eventually relayed through the spinal cord to trigger finger muscles, with latencies about $280-400~\ms$.}\label{fig:thorpe}
 \end{figure}

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 An important consequence of this speed of processing of vision is that it implies that processing is carried out using only very few spikes per layer.
-As a comparison, the latencies in macaque monkeys are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.
-Indeed, we may consider that a behavioral response occurring in only $200~\ms$ would involve about ten processing stages along the ``forward'' pathways of the visual system.
+As a comparison, the latencies in macaque monkeys are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.\add[Laurent]{Note that, since maximal conduction speeds are roughly constant, theses latencies are comparable to that found in humans, with a ratio given by the physical size of the whole system.} It follows that if we consider that a behavioral response occurra in only $200~\ms$, it would involve about ten processing stages along the ``forward'' pathways of the visual system.
 Such processes were indeed efficiently reproduced in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\add[Laurent]{Let's now focus on one processing step along a cortical pathway. }
 % reliability of spike timings

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 This has been illustrated in a recent study investigating the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
-The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in close visual proximity.
+The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in relatively close visual proximity.
 The lrAM is the core building block underlying the use of sequences of images to induce the perception of smooth, realistic visual scenes, which is at the base of movies seen in cinema theaters.
@@
 \add[Laurent]{Yet this method was mainly tested on calcium imaging data which is known to add some imprecision to the timing of the original spiking activity.} The model is \note[Laurent]{now citing the URL}openly accessible~\citep{stringer_mouselandrastermap_2020} and has led to important discoveries.
-In~\citet{stringer_spontaneous_2019,stringer_high-dimensional_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+In~\citep{stringer_spontaneous_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
 Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.

### 2022-12-19 manuscript.tex
Update on Overleaf.

@@
 %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-\citep{pfeil_neuromorphic_2013} implemented STDP on a mixed-analog-digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
+\citet{pfeil_neuromorphic_2013} implemented STDP on a mixed-analog-digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
 Coherence detection on input spikes coming from two sources was obtained with a 50 $ns$ precision.
@@
 A recent work makes implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}.
-They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay.
+They highlight the fact that using delays can be expensive \add[Laurent]{notably in memory bandwidth} because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay.
 Note that this problem is also due to the algorithm used in this paper and that the analog chips must not suffer from this type of limitation.
@@
   \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlights the importance of precise spike times,
-  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, travelling waves on arbitrary spiking motifs, etc,
+  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, travelling waves on arbitrary spiking motifs, \add[Laurent]{and notably that a an encoding based on precise spiking motifs may provide huge computational benefits},
   \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization, LP; writing---original draft preparation, all authors; writing---review and editing, all authors; supervision, LP; funding acquisition, LP and JM.
+% \authorcontributions{Conceptualization, LP; writing---original draft preparation, all authors; writing---review and editing, all authors; supervision, LP; funding acquisition, LP and JM.
 All authors have read and agreed to the published version of the manuscript.}

### 2022-12-19 manuscript.tex
maunsell_Functional_1983

@@
 We decide to focus on vertebrates and mention this choice in the new manuscript.
-%Indeed, our focus was implicitly on mammals while we were sometimes using evidence from insects without highlighting the differences.
-We believe that the coding principle we put forward is general enough to cover both orders.
-We also tried to keep the review at a reasonable length to keep the balance between computational and biological evidence.
-Therefore, while we keep our focus on mammals, we have been careful now to mention the order of animals.
 
@@
 %\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
-There are multiple experimental evidences that neurons use different strategies to communicate.}~\cite{Maunsell_Functional_1983} \add[Antoine]{demonstrate that neurons of visual area MT in macaque monkeys encode motion direction through population activity, offering robustness to single neuron activity fluctuations.
+There are multiple experimental evidences that neurons use different strategies to communicate.}~\cite{maunsell_functional_1983} \add[Antoine]{demonstrate that neurons of visual area MT in macaque monkeys encode motion direction through population activity, offering robustness to single neuron activity fluctuations.
 For rapid image processing, it has been observed that neurons in the salamander retina represent the spatial structure of an image with the relative timing of their first spikes}~\citep{gollisch_rapid_2008}\add[Antoine]{.
@@
 \add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
-Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\citep{Maunsell_Functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase--firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
+Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\citep{maunsell_functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase--firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
 In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space.

### 2022-12-19 manuscript.tex
Update manuscript.tex

@@
 \newcommand{\mynote}[1]{{\sethlcolor{yellow}\hl{#1}}}
-\usepackage{trackchanges}
-%\usepackage[finalnew]{trackchanges}
+% \usepackage{trackchanges}
+\usepackage[finalnew]{trackchanges}
 \addeditor{Antoine}
@@
 Trial repetitions of a Leaky Integrate-and-Fire neuron stimulated by the stimulus on the upper row (middle row).
-Membrane potential is represented by dark blue color when light yellow colors when depolarized) and quantified by .
-The average firing rate across trials (lower row).
+Membrane potential is represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row).
 While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{frozen} noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
@@
 The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-\add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.} \add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
+\add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.}
+
+\add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
 However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\citep{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.

### 2022-12-20 manuscript.tex
Update manuscript.tex
push REV1
Update on Overleaf.

@@
 % vim: set fenc=utf-8
+%
+% REV1
+%
 % https://www.overleaf.com/project/633bf90497b5b325a3ff692e
@@
 % role of myelination and neuronal delay
-
 \citet{fields_new_2015, fields_myelin_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments.
@@
 
-\dataavailability{This works is made reproducible using the following tools.
-First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/XXX}{GitHub}.
+\dataavailability{This work is made reproducible using the following tools.
+First the manuscript's source code, and the code reproducing all figures are available on \href{https://github.com/SpikeAI/XXX}{GitHub}.
 %~\citep{Grimaldi2023review}.
@@
 
-\acknowledgments{We would like to thank Hugo Ladret and Frdéric Chavane for valuable feedback while writing this review.
+\acknowledgments{We would like to thank Hugo Ladret and Frédéric Chavane for valuable feedback while writing this review.
 Many thanks to Simon Thorpe for the useful discussions and references on the latency of the visual system in human and non-human primates.}

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -1,0 +1,253 @@
+# 2022-12-23_video-abstract
+"""
+ Using the template @ https://laurentperrinet.github.io/sciblog/posts/2019-09-11_video-abstract-vision.html
+and using that @ https://github.com/chloepasturel/AnticipatorySPEM/blob/master/2020-03_video-abstract/2020-03-24_video-abstract.ipynb
+"""
+# Creating the movie using the (*excellent*) [MoviePy](http://zulko.github.io/moviepy/index.html) library:
+
+videoname = "Polychrony-review_2022_video-abstract"
+gifname = videoname + ".gif"
+fps = 30
+from moviepy.editor import VideoFileClip, ImageClip, TextClip, CompositeVideoClip
+
+H, W = 500, 800
+H_fig, W_fig = int(H-H/(1.618*3)), int(W-W/(1.618*3))
+
+
+opt_t = dict(font="Arial", size=(W,H), method='caption')
+opt_st = dict(font="Arial", size=(W,H), method='caption')
+opt_t = dict(size=(W,H), method='caption')
+opt_st = dict(size=(W,H), method='caption')
+
+clip = []
+t = 0 
+
+# TITRE
+texts = ["""
+Precise spiking motifs in neurobiological and neuromorphic data
+
+
+
+
+
+""", """
+Precise spiking motifs in neurobiological and neuromorphic data
+
+by Grimaldi, Gruel, Besnainou, Jérémie, 
+Martinet, and Perrinet
+
+
+""", """
+Precise spiking motifs in neurobiological and neuromorphic data
+
+by Grimaldi, Gruel, Besnainou, Jérémie, 
+Martinet, and Perrinet
+
+Brain Sciences (2022)
+"""]
+txt_opts = dict(align='center', color='white', **opt_t) #stroke_color='gray', stroke_width=.5
+duration = 2
+for text in texts:
+    txt = TextClip(text, fontsize=35, **txt_opts).set_start(t).set_duration(duration)
+    t += duration
+    clip.append(txt)
+
+# INTRO
+sub_opts = dict(fontsize=28, align='center', color='white', **opt_t)
+sub_duration = 1.5
+intro_subs = ["""
+The majority of information passing in the brain is
+mediated by all-or-none events, spikes.
+
+
+""", """
+The majority of information passing in the brain is
+mediated by all-or-none events, spikes.
+We present here novel evidence for the role of their 
+
+""", """
+The majority of information passing in the brain is
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -1,4 +1,5 @@
 # 2022-12-23_video-abstract
+
 """
  Using the template @ https://laurentperrinet.github.io/sciblog/posts/2019-09-11_video-abstract-vision.html
 and using that @ https://github.com/chloepasturel/AnticipatorySPEM/blob/master/2020-03_video-abstract/2020-03-24_video-abstract.ipynb
@@ -245,9 +246,7 @@ for text, fontsize in zip(texts, [30, 24]):
 
 video = CompositeVideoClip(clip)
 video.write_videofile(videoname + '.mp4', fps=fps)
-<BR>
-<center><video controls autoplay loop src="PasturelMontagniniPerrinet2020_video-abstract.mp4" width=61.8%/></a> </center>
-<BR>
+
 video.write_gif(gifname, fps=fps)
 from pygifsicle import optimize
 optimize(gifname)

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -6,8 +6,9 @@ and using that @ https://github.com/chloepasturel/AnticipatorySPEM/blob/master/2
 """
 # Creating the movie using the (*excellent*) [MoviePy](http://zulko.github.io/moviepy/index.html) library:
 
-videoname = "Polychrony-review_2022_video-abstract"
+videoname = "2022-12-23_polychrony-review_video-abstract"
 gifname = videoname + ".gif"
+gifname = None
 fps = 30
 from moviepy.editor import VideoFileClip, ImageClip, TextClip, CompositeVideoClip
 
@@ -24,25 +25,25 @@ clip = []
 t = 0 
 
 # TITRE
-texts = ["""
-Precise spiking motifs in neurobiological and neuromorphic data
+texts = ["""Precise spiking motifs in neurobiological and neuromorphic data
 
 
 
 
 
-""", """
-Precise spiking motifs in neurobiological and neuromorphic data
 
-by Grimaldi, Gruel, Besnainou, Jérémie, 
-Martinet, and Perrinet
+""", """Precise spiking motifs in neurobiological and neuromorphic data
+
+by Grimaldi, Gruel, 
+   Besnainou, Jérémie, 
+   Martinet, and Perrinet
 
 
-""", """
-Precise spiking motifs in neurobiological and neuromorphic data
+""", """Precise spiking motifs in neurobiological and neuromorphic data
 
-by Grimaldi, Gruel, Besnainou, Jérémie, 
-Martinet, and Perrinet
+by Grimaldi, Gruel, 
+   Besnainou, Jérémie, 
+   Martinet, and Perrinet
 
 Brain Sciences (2022)
 """]
@@ -57,20 +58,26 @@ for text in texts:
 sub_opts = dict(fontsize=28, align='center', color='white', **opt_t)
 sub_duration = 1.5
 intro_subs = ["""
-The majority of information passing in the brain is
-mediated by all-or-none events, spikes.
+The majority of information passing 
+in the brain is mediated by all-or-none 
+events, 
+
+
+
+""", """
+The majority of information passing 
+in the brain is mediated by all-or-none 
+events, *spikes*.
+
 
 
 """, """
-The majority of information passing in the brain is
-mediated by all-or-none events, spikes.
-We present here novel evidence for the role of their 
-
-""", """
-The majority of information passing in the brain is
-mediated by all-or-none events, spikes.
-We present here novel evidence for the role of their 
-precise timing from neurobiological and neuromorphic data.
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -88,8 +88,6 @@ for i_sub, subtitle in enumerate(intro_subs):
 
 
 
-# visual-latency.jpg
-
 # fig_name = {"BSM":['BSM.mp4'], # 30''
 #             "Bet_eyeMvt":['eyeMvt.mp4',  # 10''
 #                       # '1_B_Trace_moyenne.png',
@@ -175,38 +173,67 @@ for i_sub, subtitle in enumerate(intro_subs):
 #                "...a hallmark of their inter-individual differences."]
 
 
-# # http://zulko.github.io/moviepy/ref/VideoClip/VideoClip.html?highlight=compositevideoclip#textclip
-# txt_opts = dict(fontsize=65, bg_color='white', align='center', **opt_st)
-# sub_opts = dict(fontsize=28, align='South', color='white', **opt_st)
 
-# for text, color in zip(texts, colors):
-#     # duration = 1
-#     # txt = TextClip(text, color=color, **txt_opts).set_start(t).set_duration(duration)
-#     # t += duration
-#     # clip.append(txt)
+chapters = {'VS': dict(title="Visual system", color='black'), 
+            'poly': dict(title="Polychrony", color='orange'), 
+            'neuro': dict(title="Neurobiology", color='red'), 
+            'DVS': dict(title="Neuromorphic", color='blue'), 
+            }
 
-#     for fig in fig_name[text] :
-    
-#         if fig[-4:]=='.mp4' :
-#             img = VideoFileClip('%s/%s'%(text, fig), audio=False)
-#             print(fig, '--> duration:', img.duration, ', fps:', img.fps)
-#             duration = img.duration
-#         else :
-#             duration = duration_dict[fig]
-#             img = ImageClip('%s/%s'%(text, fig)).set_duration(duration)
+chapters['VS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
+                "This shows that different individuals have different...", 
+                "...compromise between exploration and the exploitation...", 
+                "...a hallmark of their inter-individual differences."
+)]
+chapters['poly']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
+                "This shows that different individuals have different...", 
+                "...compromise between exploration and the exploitation...", 
+                "...a hallmark of their inter-individual differences."
+)]
+chapters['neuro']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
+                "This shows that different individuals have different...", 
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -284,6 +284,7 @@ for text, fontsize in zip(texts, [30, 24]):
 video = CompositeVideoClip(clip)
 video.write_videofile(videoname + '.mp4', fps=fps)
 
-video.write_gif(gifname, fps=fps)
-from pygifsicle import optimize
-optimize(gifname)
+if not(gifname is None):
+    video.write_gif(gifname, fps=fps)
+    from pygifsicle import optimize
+    optimize(gifname)

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -183,23 +183,19 @@ chapters = {'VS': dict(title="Visual system", color='black'),
 chapters['VS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."
-)]
+                "...a hallmark of their inter-individual differences.")]
 chapters['poly']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."
-)]
+                "...a hallmark of their inter-individual differences.")]
 chapters['neuro']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."
-)]
+                "...a hallmark of their inter-individual differences.")]
 chapters['DVS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."
-)]
+                "...a hallmark of their inter-individual differences.")]
 
 
 # http://zulko.github.io/moviepy/ref/VideoClip/VideoClip.html?highlight=compositevideoclip#textclip

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -180,22 +180,22 @@ chapters = {'VS': dict(title="Visual system", color='black'),
             'DVS': dict(title="Neuromorphic", color='blue'), 
             }
 
-chapters['VS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
+chapters['VS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences.")]
-chapters['poly']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
+                "...a hallmark of their inter-individual differences."])]
+chapters['poly']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences.")]
-chapters['neuro']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
+                "...a hallmark of their inter-individual differences."])]
+chapters['neuro']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences.")]
-chapters['DVS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=
+                "...a hallmark of their inter-individual differences."])]
+chapters['DVS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences.")]
+                "...a hallmark of their inter-individual differences."])]
 
 
 # http://zulko.github.io/moviepy/ref/VideoClip/VideoClip.html?highlight=compositevideoclip#textclip

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -87,93 +87,6 @@ for i_sub, subtitle in enumerate(intro_subs):
     clip.append(sub)
 
 
-
-# fig_name = {"BSM":['BSM.mp4'], # 30''
-#             "Bet_eyeMvt":['eyeMvt.mp4',  # 10''
-#                       # '1_B_Trace_moyenne.png',
-#                       # 'demo_fit.png'
-#                          'Bet.mp4'],  # 15''
-#             "BBCP":['3_BCP_model.png', 
-#                     '3_BCP_readouts.png',
-#                     '2_results_enregistrement.mp4', # 30''
-#                     '4_A_result_psycho_aSPEM.png',
-#                     '4_B_result_psycho_bet.png',
-#                     '5B_inter-individual_differences_fit.png',
-#                     '5A_inter-individual_differences_fit.png']}
-
-# duration_dict = {'3_BCP_model.png':12, 
-#                  '3_BCP_readouts.png':9,
-#                  '4_A_result_psycho_aSPEM.png':7,
-#                  '4_B_result_psycho_bet.png':7,
-#                  '5A_inter-individual_differences_fit.png':6,
-#                  '5B_inter-individual_differences_fit.png':6}
-
-# # LES CHOSES SERIEUSES !
-# texts = ["BSM", "Bet_eyeMvt", "BBCP"]
-# colors = ['black', 'orange', 'blue', 'red']
-# subtitles = {}
-# subtitles['BSM.mp4'] = ["Take for instance the case of a doctor who tests...", # 
-#                "...her patients to check if they have a flu or not.", 
-#                "The underlying cause for these observations (Top graph)...", 
-#                # "...might be caused by a switch in probability rate (Bottom)...",
-#                "...might be a sudden increase in the occurrence of flu...",
-#                "...from a normal situation (low p) to an outbreak (high p).", 
-#                "In the longer-term, we may understand volatility as...", 
-#                "...different epochs, each with a stationary rate of cases and...", 
-#                "...separated by switches which occur at random times.", 
-#                # "...but with a given hazard rate (= rate of switches)."
-#                        ]
-
-# subtitles['eyeMvt.mp4'] = ["How do humans behave to anticipate future outcomes?", 
-#                            "We tested this by showing trial sequences...", 
-#                            "...drawn from such a switching process.", 
-#                            "Each trial consists of a dot on the screen...", 
-#                            "... which moves either left or right.", 
-#                            "First, we recorded the eye movements of observers....", 
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -122,7 +122,9 @@ for chapter_key in chapters.keys():
     t += duration
     clip.append(txt)
 
-    for content in chapter['content'] :
+    for content_key in chapter['content'].keys():
+        content = chapter[content_key]
+
         # set the figure
         figname = content['figure']
         if figname[-4:]=='.mp4' :
@@ -130,7 +132,8 @@ for chapter_key in chapters.keys():
             print(figname, '--> duration:', img.duration, ', fps:', img.fps)
             duration = img.duration
         else :
-            img = ImageClip(figname).set_duration(content[duration])
+            duration = content['duration']
+            img = ImageClip(figname).set_duration(duration)
 
         img = img.set_start(t).set_pos('center').resize(height=H_fig, width=W_fig)
 
@@ -139,8 +142,8 @@ for chapter_key in chapters.keys():
 
         # write the subtitles
         t_sub = t - duration
-        sub_duration = duration / len(content[subtitle])
-        for subtitle in content[subtitle]:
+        sub_duration = duration / len(content['subtitle'])
+        for subtitle in content['subtitle']:
             sub = TextClip(subtitle, **sub_opts).set_start(t_sub).set_duration(sub_duration)
             t_sub += sub_duration
             clip.append(sub)

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -118,7 +118,7 @@ sub_opts = dict(fontsize=28, align='South', color='white', **opt_st)
 for chapter_key in chapters.keys():
     chapter = chapters[chapter_key]
     duration = 1
-    txt = TextClip(chapter[title], color=chapter[color], **txt_opts).set_start(t).set_duration(duration)
+    txt = TextClip(chapter['title'], color=chapter['color'], **txt_opts).set_start(t).set_duration(duration)
     t += duration
     clip.append(txt)
 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -122,9 +122,7 @@ for chapter_key in chapters.keys():
     t += duration
     clip.append(txt)
 
-    for content_key in chapter['content'].keys():
-        content = chapter[content_key]
-
+    for content in chapter['content']:
         # set the figure
         figname = content['figure']
         if figname[-4:]=='.mp4' :

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -18,8 +18,8 @@ H_fig, W_fig = int(H-H/(1.618*3)), int(W-W/(1.618*3))
 
 opt_t = dict(font="Arial", size=(W,H), method='caption')
 opt_st = dict(font="Arial", size=(W,H), method='caption')
-opt_t = dict(size=(W,H), method='caption')
-opt_st = dict(size=(W,H), method='caption')
+# opt_t = dict(size=(W,H), method='caption')
+# opt_st = dict(size=(W,H), method='caption')
 
 clip = []
 t = 0 
@@ -31,19 +31,16 @@ texts = ["""Precise spiking motifs in neurobiological and neuromorphic data
 
 
 
+""", """Precise spiking motifs in neurobiological and neuromorphic data
+
+   Grimaldi, Gruel, Besnainou, 
+   Jérémie, Martinet, and Perrinet
+
 
 """, """Precise spiking motifs in neurobiological and neuromorphic data
 
-by Grimaldi, Gruel, 
-   Besnainou, Jérémie, 
-   Martinet, and Perrinet
-
-
-""", """Precise spiking motifs in neurobiological and neuromorphic data
-
-by Grimaldi, Gruel, 
-   Besnainou, Jérémie, 
-   Martinet, and Perrinet
+   Grimaldi, Gruel, Besnainou, 
+   Jérémie, Martinet, and Perrinet
 
 Brain Sciences (2022)
 """]
@@ -55,8 +52,8 @@ for text in texts:
     clip.append(txt)
 
 # INTRO
-sub_opts = dict(fontsize=28, align='center', color='white', **opt_t)
-sub_duration = 1.5
+sub_opts = dict(fontsize=28, align='left', color='white', **opt_t)
+sub_duration = 2.5
 intro_subs = ["""
 The majority of information passing 
 in the brain is mediated by all-or-none 
@@ -97,15 +94,19 @@ chapters['VS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
                 "...a hallmark of their inter-individual differences."])]
-chapters['poly']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
-                "This shows that different individuals have different...", 
+chapters['poly']['content'] = [dict(figure='figures/replicating_MainenSejnowski1995.png', duration=5, subtitle=["This shows that different individuals have different...", 
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -52,7 +52,7 @@ for text in texts:
     clip.append(txt)
 
 # INTRO
-sub_opts = dict(fontsize=28, align='left', color='white', **opt_t)
+sub_opts = dict(fontsize=28, align='West', color='white', **opt_t)
 sub_duration = 2.5
 intro_subs = ["""
 The majority of information passing 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -106,7 +106,7 @@ chapters['neuro']['content'] = [
                 dict(figure='figures/Ikegaya2004zse0150424620001.jpeg', duration=5, subtitle=["This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
                 "...a hallmark of their inter-individual differences."])]
-chapters['DVS']['content'] = [dict(figure='event_driven_computations.png', duration=5, subtitle=[
+chapters['DVS']['content'] = [dict(figure='figures/event_driven_computations.png', duration=5, subtitle=[
                 "This shows that different individuals have different...", 
                 "...compromise between exploration and the exploitation...", 
                 "...a hallmark of their inter-individual differences."])]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -84,40 +84,38 @@ for i_sub, subtitle in enumerate(intro_subs):
     clip.append(sub)
 
 
-chapters = {'VS': dict(title="Visual system", color='black'), 
-            'poly': dict(title="Polychrony", color='orange'), 
-            'neuro': dict(title="Neurobiology", color='red'), 
-            'DVS': dict(title="Neuromorphic", color='blue'), 
-            }
-
-chapters['VS']['content'] = [dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
-                "This shows that different individuals have different...", 
-                "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."])]
-chapters['poly']['content'] = [dict(figure='figures/replicating_MainenSejnowski1995.png', duration=5, subtitle=["This shows that different individuals have different...", 
-                "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."])]
-chapters['neuro']['content'] = [
-                dict(figure='figures/Ikegaya2004zse0150424620001.jpeg', 
-                     duration=5, subtitle=[
-                        "This shows that different individuals have different...", 
-                        "...compromise between exploration and the exploitation...", 
-                        "...a hallmark of their inter-individual differences."]),
-                dict(figure='figures/Ikegaya2004zse0150424620001.jpeg', duration=5, subtitle=["This shows that different individuals have different...", 
-                "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."])]
-chapters['DVS']['content'] = [dict(figure='figures/event_driven_computations.png', duration=5, subtitle=[
-                "This shows that different individuals have different...", 
-                "...compromise between exploration and the exploitation...", 
-                "...a hallmark of their inter-individual differences."])]
+chapters = [dict(title="Visual system", color='black',
+            content=[dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
+                            "This shows that different individuals have different...", 
+                            "...compromise between exploration and the exploitation...", 
+                            "...a hallmark of their inter-individual differences."])]), 
+           dict(title="Polychrony", color='orange',
+            content=[dict(figure='figures/replicating_MainenSejnowski1995.png', duration=5, subtitle=[
+                            "This shows that different individuals have different...", 
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -130,7 +130,7 @@ for chapter in chapters:
             duration = img.duration
         else :
             duration = content['duration']
-            img = ImageClip(figname).set_duration(duration)
+            img = ImageClip(figname, size=(W_fig, H_fig)).set_duration(duration)
 
         img = img.set_start(t).set_pos('center').resize(height=H_fig, width=W_fig)
 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -45,7 +45,7 @@ texts = ["""Precise spiking motifs in neurobiological and neuromorphic data
 Brain Sciences (2022)
 """]
 txt_opts = dict(align='center', color='white', **opt_t) #stroke_color='gray', stroke_width=.5
-duration = 2
+duration = 1.5
 for text in texts:
     txt = TextClip(text, fontsize=35, **txt_opts).set_start(t).set_duration(duration)
     t += duration
@@ -53,7 +53,7 @@ for text in texts:
 
 # INTRO
 sub_opts = dict(fontsize=28, align='West', color='white', **opt_t)
-sub_duration = 2.5
+sub_duration = 1.5
 intro_subs = ["""
 The majority of information passing 
 in the brain is mediated by all-or-none 
@@ -82,7 +82,7 @@ for i_sub, subtitle in enumerate(intro_subs):
     sub = TextClip(subtitle, **sub_opts).set_start(t).set_duration(sub_duration)
     t += sub_duration
     clip.append(sub)
-
+clip.append(sub) # another bit on the last bit
 
 chapters = [dict(title="Visual system", color='black',
             content=[dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
@@ -196,12 +196,14 @@ https://laurentperrinet.github.io/
 ]
 
 txt_opts = dict(align='center', **opt_t)
-duration = 3
-for text, fontsize in zip(texts, [30, 24]):
+duration = 2.5
+for text, fontsize in zip(texts, [28, 28]):
     txt = TextClip(text, color='orange', fontsize=fontsize, **txt_opts).set_start(t).set_duration(duration)
     t += duration
     clip.append(txt)    
 
+
+# COMPOSITING
 video = CompositeVideoClip(clip)
 video.write_videofile(videoname + '.mp4', fps=fps)
 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -94,7 +94,7 @@ chapters = [dict(title="Visual system", color='black',
                             "This shows that different individuals have different...", 
                             "...compromise between exploration and the exploitation...", 
                             "...a hallmark of their inter-individual differences."]),
-                    dict(figure='figures/event_driven_computations.png', duration=5, subtitle=[
+                    dict(figure='figures/THC_1a.pdf', duration=5, subtitle=[
                             "This shows that different individuals have different...", 
                             "...compromise between exploration and the exploitation...", 
                             "...a hallmark of their inter-individual differences."])]), 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -132,7 +132,13 @@ for chapter in chapters:
             duration = content['duration']
             img = ImageClip(figname, size=(W_fig, H_fig)).set_duration(duration)
 
-        img = img.set_start(t).set_pos('center').resize(height=H_fig, width=W_fig)
+        H_clip, W_clip, three = clip.get_frame(0).shape
+        if H_clip/W_clip > H_fig/W_fig: # portrait-like
+            img = img.resize(height=H_fig)
+        else: # landscape-like
+            img = img.resize(width=W_fig)
+
+        img = img.set_start(t).set_pos('center')#.resize(height=H_fig, width=W_fig)
 
         t += duration
         clip.append(img)

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -130,7 +130,7 @@ for chapter in chapters:
             duration = img.duration
         else :
             duration = content['duration']
-            img = ImageClip(figname, size=(W_fig, H_fig)).set_duration(duration)
+            img = ImageClip(figname).set_duration(duration)
 
         H_clip, W_clip, three = clip.get_frame(0).shape
         if H_clip/W_clip > H_fig/W_fig: # portrait-like

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -132,7 +132,7 @@ for chapter in chapters:
             duration = content['duration']
             img = ImageClip(figname).set_duration(duration)
 
-        H_clip, W_clip, three = clip.get_frame(0).shape
+        H_clip, W_clip, three = img.get_frame(0).shape
         if H_clip/W_clip > H_fig/W_fig: # portrait-like
             img = img.resize(height=H_fig)
         else: # landscape-like

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -94,7 +94,11 @@ chapters = [dict(title="Visual system", color='black',
                             "This shows that different individuals have different...", 
                             "...compromise between exploration and the exploitation...", 
                             "...a hallmark of their inter-individual differences."]),
-                    dict(figure='figures/THC_1a.pdf', duration=5, subtitle=[
+                    dict(figure='figures/THC_1a_k.png', duration=5, subtitle=[
+                            "This shows that different individuals have different...", 
+                            "...compromise between exploration and the exploitation...", 
+                            "...a hallmark of their inter-individual differences."]),
+                    dict(figure='figures/THC_1a.png', duration=5, subtitle=[
                             "This shows that different individuals have different...", 
                             "...compromise between exploration and the exploitation...", 
                             "...a hallmark of their inter-individual differences."])]), 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -86,9 +86,15 @@ clip.append(sub) # another bit on the last bit
 
 chapters = [dict(title="Visual system", color='black',
             content=[dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
-                            "This shows that different individuals have different...", 
-                            "...compromise between exploration and the exploitation...", 
-                            "...a hallmark of their inter-individual differences."])]), 
+                            "The visual system is very efficient to generate a...", 
+                            "...decision from the raw image to the different ...", 
+                            "...stages of the visual pathways here a reaction...", 
+                            "...of finger muscles in about 300 milliseconds."]),
+                     dict(figure='../AG_figures/animated_neurons/LIF.mp4', duration=5, subtitle=[
+                            "This efficiency is thought to be mediated by spikes...", 
+                            "...that is brief all-or-none events which are passed...", 
+                            "...from assemblies of neurons to others, in the...", 
+                            "...very large network which forms the brain."])]), 
            dict(title="Polychrony", color='orange',
             content=[dict(figure='figures/replicating_MainenSejnowski1995.png', duration=5, subtitle=[
                             "This shows that different individuals have different...", 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -52,7 +52,7 @@ for text in texts:
     clip.append(txt)
 
 # INTRO
-sub_opts = dict(fontsize=28, align='West', color='white', **opt_t)
+sub_opts = dict(fontsize=32, align='center', color='white', **opt_t)
 sub_duration = 1.5
 intro_subs = ["""
 The majority of information passing 
@@ -85,34 +85,39 @@ for i_sub, subtitle in enumerate(intro_subs):
 clip.append(sub) # another bit on the last bit
 
 chapters = [dict(title="Visual system", color='black',
-            content=[dict(figure='figures/visual-latency.jpg', duration=5, subtitle=[
+            content=[dict(figure='figures/visual-latency_bg.jpg', duration=3, subtitle=[
                             "The visual system is very efficient to generate a...", 
-                            "...decision from the raw image to the different ...", 
+                            "...decision from the raw image to the different ..."]),
+                    dict(figure='figures/visual-latency.jpg', duration=3, subtitle=[
                             "...stages of the visual pathways here a reaction...", 
                             "...of finger muscles in about 300 milliseconds."]),
                      dict(figure='../AG_figures/animated_neurons/LIF.mp4', duration=5, subtitle=[
                             "This efficiency is thought to be mediated by spikes...", 
-                            "...that is brief all-or-none events which are passed...", 
-                            "...from assemblies of neurons to others, in the...", 
-                            "...very large network which forms the brain."])]), 
+                            "...that is, brief all-or-none events which are passed...", 
+                            "...in the very large network which forms the brain...", 
+                            "...from assemblies of neurons to others."])]), 
            dict(title="Polychrony", color='orange',
-            content=[dict(figure='figures/replicating_MainenSejnowski1995.png', duration=5, subtitle=[
-                            "This shows that different individuals have different...", 
-                            "...compromise between exploration and the exploitation...", 
-                            "...a hallmark of their inter-individual differences."]),
+            content=[dict(figure='figures/izhikevich.png', duration=5, subtitle=[
+                            "We review here an hypothesis in which, rather than the...", 
+                            "...firing frequency of spikes, it is their precise timing...", 
+                            "...which would enable or not this message passing."]),
                     dict(figure='figures/THC_1a_k.png', duration=5, subtitle=[
-                            "This shows that different individuals have different...", 
-                            "...compromise between exploration and the exploitation...", 
+                            "We present different mathematical models to extract...", 
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -111,18 +111,23 @@ chapters = [dict(title="Visual system", color='black',
                             "...or neuromorphic engineering."])]), 
            dict(title="Neurobiology", color='red',
             content=[dict(figure='figures/haimerl2019.jpg', duration=5, subtitle=[
-                            "This shows that different individuals have different...", 
-                            "...compromise between exploration and the exploitation...", 
-                            "...a hallmark of their inter-individual haimerl2019.jpg."]),
+                            "This hypothesis is reviewed with respect to our ...", 
+                            "...knowledge of the neuobilogy, for instance in the...", 
+                            "...hippocampus of rodents. But we review also ."]),
                     dict(figure='figures/Ikegaya2004zse0150424620001.jpeg', duration=5, subtitle=[
-                            "This shows that different individuals have different...", 
-                            "...compromise between exploration and the exploitation...", 
-                            "...a hallmark of their inter-individual ."])]), 
+                            "...numerous and extensive work on the mechanisms...", 
+                            "... which may allow the neural system to learn...", 
+                            "...to actually use that precise spiking motifs."])]), 
            dict(title="Neuromorphic", color='blue',
             content=[dict(figure='figures/event_driven_computations.png', duration=5, subtitle=[
-                            "This shows that different individuals have different...", 
-                            "...compromise between exploration and the exploitation...", 
-                            "...a hallmark of their inter-individual differences."])]), 
+                            "We also review evidence from computational neuroscience...", 
+                            "...and neuromorphic engineering, in particular for novel...", 
+                            "...device such as event-based cameras which directly...", 
+                            "...transform lumnious information into spikes."]),
+                    dict(figure='../pyTERtorch/2022-11-30_BiolCybernetics/figures/2022-11-10_MotionDetection_input.mp4', duration=5, subtitle=[
+                            "For instance, we show how precise spike times may be...", 
+                            "...used to detect the direction of motion in a stream...", 
+                            "...of events in an ultra fast fashion."])]), 
            ]
 
 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -24,7 +24,10 @@ opt_st = dict(font="Arial", size=(W,H), method='caption')
 clip = []
 t = 0 
 
+#################################################################################
 # TITRE
+#################################################################################
+# 
 texts = ["""Precise spiking motifs in neurobiological and neuromorphic data
 
 
@@ -51,7 +54,10 @@ for text in texts:
     t += duration
     clip.append(txt)
 
+#################################################################################
 # INTRO
+#################################################################################
+
 sub_opts = dict(fontsize=32, align='center', color='white', **opt_t)
 sub_duration = 1.5
 intro_subs = ["""
@@ -84,7 +90,9 @@ for i_sub, subtitle in enumerate(intro_subs):
     clip.append(sub)
 clip.append(sub) # another bit on the last bit
 
-chapters = [dict(title="Visual system", color='black',
+#################################################################################
+#################################################################################
+chapters = [dict(title="Visual system", color='green',
             content=[dict(figure='figures/visual-latency_bg.jpg', duration=3, subtitle=[
                             "The visual system is very efficient to generate a...", 
                             "...decision from the raw image to the different ..."]),
@@ -171,7 +179,13 @@ for chapter in chapters:
             t_sub += sub_duration
             clip.append(sub)
 
+#################################################################################
+#################################################################################
+
+
+#################################################################################
 # OUTRO
+#################################################################################
 texts = ["""
 Overall, this manuscript reviews the potential
 of using the precise spiking motifs for a
@@ -232,7 +246,10 @@ img = ImageClip('figures/qrcode.png').set_duration(duration)
 img = img.resize(width=W_fig).set_start(t).set_pos('center')
 clip.append(img)
 
+#################################################################################
 # COMPOSITING
+#################################################################################
+
 video = CompositeVideoClip(clip)
 video.write_videofile(videoname + '.mp4', fps=fps)
 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -62,22 +62,22 @@ sub_opts = dict(fontsize=32, align='center', color='white', **opt_t)
 sub_duration = 1.5
 intro_subs = ["""
 The majority of information passing 
-in the brain is mediated by all-or-none 
-events, 
+in the brain is mediated by all-or-none events,
 
-
-
-""", """
-The majority of information passing 
-in the brain is mediated by all-or-none 
-events, *spikes*.
 
 
 
 """, """
 The majority of information passing 
-in the brain is mediated by all-or-none 
-events, *spikes*.
+in the brain is mediated by all-or-none events,
+*spikes*.
+
+
+
+""", """
+The majority of information passing 
+in the brain is mediated by all-or-none events,
+*spikes*.
 We present here novel evidence for the 
 role of their precise timing from 
 neurobiological and neuromorphic data.
@@ -111,10 +111,9 @@ chapters = [dict(title="Visual system", color='green',
                             "...which would enable or not this message passing."]),
                     dict(figure='figures/THC_1a_k.png', duration=5, subtitle=[
                             "We present different mathematical models to extract...", 
-                            "...such precise spiking motifs and...", 
-                            "...a hallmark of their inter-individual differences."]),
+                            "...such precise spiking motifs and...""]),
                     dict(figure='figures/THC_1a.png', duration=5, subtitle=[
-                            "... also some novel tools to extract these motifs...", 
+                            "...also some novel tools to extract these motifs...", 
                             "...in arbitrary raster plots, from neurobiology...", 
                             "...or neuromorphic engineering."])]), 
            dict(title="Neurobiology", color='red',
@@ -124,7 +123,7 @@ chapters = [dict(title="Visual system", color='green',
                             "...hippocampus of rodents. But we review also ."]),
                     dict(figure='figures/Ikegaya2004zse0150424620001.jpeg', duration=5, subtitle=[
                             "...numerous and extensive work on the mechanisms...", 
-                            "... which may allow the neural system to learn...", 
+                            "...which may allow the neural system to learn...", 
                             "...to actually use that precise spiking motifs."])]), 
            dict(title="Neuromorphic", color='blue',
             content=[dict(figure='figures/event_driven_computations.png', duration=5, subtitle=[
@@ -134,8 +133,8 @@ chapters = [dict(title="Visual system", color='green',
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -111,7 +111,7 @@ chapters = [dict(title="Visual system", color='green',
                             "...which would enable or not this message passing."]),
                     dict(figure='figures/THC_1a_k.png', duration=5, subtitle=[
                             "We present different mathematical models to extract...", 
-                            "...such precise spiking motifs and...""]),
+                            "...such precise spiking motifs and..."]),
                     dict(figure='figures/THC_1a.png', duration=5, subtitle=[
                             "...also some novel tools to extract these motifs...", 
                             "...in arbitrary raster plots, from neurobiology...", 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -119,12 +119,13 @@ chapters = [dict(title="Visual system", color='green',
            dict(title="Neurobiology", color='red',
             content=[dict(figure='figures/haimerl2019.jpg', duration=5, subtitle=[
                             "This hypothesis is reviewed with respect to our ...", 
-                            "...knowledge of the neuobilogy, for instance in the...", 
-                            "...hippocampus of rodents. But we review also ."]),
+                            "...knowledge of the neurobiology, for instance in the...", 
+                            "...hippocampus of rodents. Also, we review..."]),
                     dict(figure='figures/Ikegaya2004zse0150424620001.jpeg', duration=5, subtitle=[
                             "...numerous and extensive work on the mechanisms...", 
                             "...which may allow the neural system to learn...", 
-                            "...to actually use that precise spiking motifs."])]), 
+                            "...to actually use that precise spiking motifs", 
+                            "...by atuning the delay between pairs of neurons."])]), 
            dict(title="Neuromorphic", color='blue',
             content=[dict(figure='figures/event_driven_computations.png', duration=5, subtitle=[
                             "We also review evidence from computational neuroscience...", 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -242,6 +242,9 @@ for text in texts:
     t += duration
     clip.append(txt)    
 
+# QRCODE
+#
+# qrencode -o figures/qrcode.png -d 200 https://laurentperrinet.github.io/publication/grimaldi-22-polychronies
 img = ImageClip('figures/qrcode.png').set_duration(duration)
 img = img.resize(width=(W_fig*2)//3).set_start(t).set_pos('center')
 clip.append(img)

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -1,8 +1,12 @@
+"""
 # 2022-12-23_video-abstract
 
-"""
+
  Using the template @ https://laurentperrinet.github.io/sciblog/posts/2019-09-11_video-abstract-vision.html
 and using that @ https://github.com/chloepasturel/AnticipatorySPEM/blob/master/2020-03_video-abstract/2020-03-24_video-abstract.ipynb
+
+cp 2022-12-23_polychrony-review_video-abstract.mp4  /Users/laurentperrinet/metagit/blog/hugo_academic/content/publication/grimaldi-22-polychronies/
+
 """
 # Creating the movie using the (*excellent*) [MoviePy](http://zulko.github.io/moviepy/index.html) library:
 

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -62,8 +62,6 @@ for text in texts:
 # INTRO
 #################################################################################
 
-sub_opts = dict(fontsize=32, align='center', color='white', **opt_t)
-sub_duration = 1.5
 intro_subs = ["""
 The majority of information passing 
 in the brain is mediated by all-or-none events,
@@ -82,12 +80,14 @@ in the brain is mediated by all-or-none events,
 The majority of information passing 
 in the brain is mediated by all-or-none events,
 *spikes*.
-We present here novel evidence for the 
+We review here novel evidence for the 
 role of their precise timing from 
 neurobiological and neuromorphic data.
 """,
                ]
 
+sub_opts = dict(fontsize=32, align='center', color='white', **opt_t)
+sub_duration = 1.5
 for i_sub, subtitle in enumerate(intro_subs):
     sub = TextClip(subtitle, **sub_opts).set_start(t).set_duration(sub_duration)
     t += sub_duration
@@ -99,12 +99,12 @@ clip.append(sub) # another bit on the last bit
 chapters = [dict(title="Visual system", color='green',
             content=[dict(figure='figures/visual-latency_bg.jpg', duration=3, subtitle=[
                             "The visual system is very efficient to generate a...", 
-                            "...decision from the raw image to the different ..."]),
+                            "...decision from the retinal image to the different ..."]),
                     dict(figure='figures/visual-latency.jpg', duration=3, subtitle=[
-                            "...stages of the visual pathways here a reaction...", 
+                            "...stages of the visual pathways, here a reaction...", 
                             "...of finger muscles in about 300 milliseconds."]),
                      dict(figure='../AG_figures/animated_neurons/LIF.mp4', duration=5, subtitle=[
-                            "This efficiency is thought to be mediated by spikes...", 
+                            "This efficiency is thought to be allowed by spikes...", 
                             "...that is, brief all-or-none events which are passed...", 
                             "...in the very large network which forms the brain...", 
                             "...from assemblies of neurons to others."])]), 
@@ -113,7 +113,7 @@ chapters = [dict(title="Visual system", color='green',
                             "We review here an hypothesis in which, rather than the...", 
                             "...firing frequency of spikes, it is their precise timing...", 
                             "...which would enable or not this message passing."]),
-                    dict(figure='figures/THC_1a_k.png', duration=5, subtitle=[
+                    dict(figure='figures/THC_1a_k.png', duration=3, subtitle=[
[…]

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -83,12 +83,11 @@ in the brain is mediated by all-or-none events,
 We review here novel evidence for the 
 role of their precise timing from 
 neurobiological and neuromorphic data.
-""",
-               ]
+"""]
 
 sub_opts = dict(fontsize=32, align='center', color='white', **opt_t)
-sub_duration = 1.5
-for i_sub, subtitle in enumerate(intro_subs):
+sub_durations = [1.5, 1.5, 2.5]
+for subtitle, sub_duration in zip(intro_subs, sub_durations):
     sub = TextClip(subtitle, **sub_opts).set_start(t).set_duration(sub_duration)
     t += sub_duration
     clip.append(sub)
@@ -194,7 +193,7 @@ for chapter in chapters:
 texts = ["""
 Overall, this manuscript reviews the potential
 of using the precise spiking motifs for a
-building more efficient 
+building more efficient machine learning 
 
 
 
@@ -229,7 +228,6 @@ for text in texts:
 # FIN
 texts = ["""
 For more info, and the full, open-sourced code... visit
-
 
 
 """, """

### 2022-12-23 2022-12-23_video-abstract.py
Brain Sciences : 2022-12-23_video-abstract.py

@@ -192,8 +192,7 @@ for chapter in chapters:
 #################################################################################
 texts = ["""
 Overall, this manuscript reviews the potential
-of using the precise spiking motifs for a
-building more efficient machine learning 
+of using precise spiking motifs to design
 
 
 
@@ -201,21 +200,19 @@ building more efficient machine learning
 """,
 """
 Overall, this manuscript reviews the potential
-of using the precise spiking motifs for a
-building more efficient machine learning 
-algorithms, faster and more frugal,
-but also to better understand 
+of using precise spiking motifs to design
+more efficient machine learning algorithms, 
+that is, faster and more energy-frugal,
 
 
 """,
 """
 Overall, this manuscript reviews the potential
-of using the precise spiking motifs for a
-building more efficient machine learning 
-algorithms, faster and more frugal,
+of using precise spiking motifs to design
+more efficient machine learning algorithms, 
+that is, faster and more energy-frugal,
 but also to better understand 
-neurobiological functioning and their 
-disorders...
+neurobiological processes and their disorders...
 """]
 
 txt_opts = dict(fontsize=30, align='center', **opt_t)

### 2022-12-25 manuscript.tex
Update on Overleaf.
Update manuscript.tex
2022-10-04 conversion from manubot

--- 2022-12-19 manuscript.tex
@@
 % https://www.zotero.org/groups/4562620/polychronies
-%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\ResponseReviewerTwo}{
-We would like to thank reviewer 2 for his very detailed response which were essential in providing this new revision.
-
-> The authors present a relatively thorough review on neural coding based on spike patterns in populations of neurons, from biological, theoretical and neuromorphic engineering perspective.
-I find the review mostly fine, with some omissions that, when corrected, should enhance it.
-
-Thanks for the encouraging comment, we hope that in its present form, the result is even more useful to the community.
-
-> My main concern is the overall reliance on spike motifs as a coding mechanism.
-The authors should list alternatives and compare to spike motifs, at least cursory.
-A minor concern is the focus on mylenation-based delay learning, ignoring the obvious length-based delay adjustments.
-
-We have more progressively commented different alternatives for coding mechanisms in the beginning of section 2, highlighting their relative benefits.
-
-> I have commented heavily within the manuscript, and am attaching it for the authors to use.
-I tried to delineate wording suggestions from more significant critiques.
-I am also listing the bigger concerns below.
-
-Many thanks for providing the PDF of the manuscript with annotations, this was immensely useful to annotate and improve our revision.
-To respond to all these points, we have constructed a tracked changes manuscript in which we respond (by a change, removal or addition) to the reviewer's comments.
-
-> Specific issues:
-> - not referencing much of insect work, with a few exceptions that are put in mammalian context.
-OK to do just timing in vertebrate systems (as birds are used), but should be stated clearly, and cited work should comport to that constraint (e.g., cite rat olfactory, not locust olfactory).
-Should the authors wish to include invertebrate work, there is a LOT more to review.
-
-Indeed, our focus was implicitly on vertebrates but we cited some works about insect.
-As mentioned by the reviewer, adding researches done on insects would increase too much the size of the article.
-We decide to focus on vertebrates and mention this choice in the new manuscript.
-
-> - some references to original research are missing, while often much older ones are used (and not reviews).
-> - missing important lines of work on spike timing. In particular, Aurel Lazar's TEM work. 
-> - missing important lines of work on dendritic processing: Mel, Spruston.
-
-These critical references were indeed missing and we have added them now to the manuscript. Thanks for pointing out these omissions.
-
-> - discounting axonal length as delay mechanism, focus on myelin effects too high.
-
-This was corrected in section 4, providing a more balanced review.
-
-> - singular focus on synfire chain/polychronicity-like mechanisms.
-Alternative options, like Bayesian spiking (Deneve), probabilistic ensemble motifs (Ballard) and probabilistic population coding (Pouget and Latham), which do not rely on exact motifs of spikes, are not discussed.
-
-Indeed, we had made that choice to study precise temporal motifs.
-In the new manuscript, we mention these alternative options briefly even if we keep the focus on deterministic spiking motifs for this review.
-
-> - there are also some discrepancies - are SNN-s differentiable or not.
-
-We have corrected these discrepancies. Thanks again for all these useful comments.
-
-}
-
-% cite or die
-% ❯ ipython analyze_bibs.py
-
-% In recognition of your well-respected work, we are inviting you to submit one feature contribution (i.e., a long review or research paper) to an MDPI journal free of charge.
-If you wish to take advantage of this opportunity, please provide your information and the name of the journal you would like to submit to: https://www.surveymonkey.com/r/XQZ5FP9 (If you cannot find the journal through the link, please feel free to reply to us about your choice).
-
-% The submission deadline for this special offer is 31 October 2022.
-The journal editorial office will contact you soon once you choose the journal.
-Please contact the revelant journal’s Editorial Office if you require an extension.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %  LaTeX support: latex@mdpi.com 
@@
 %=================================================================
-\documentclass[brainsci, %arts,
-               review,submit,pdftex,moreauthors
-               %,  draft
-               ]{Definitions/mdpi} 
+\documentclass[brainsci,review,submit,pdftex,moreauthors]{Definitions/mdpi} 
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
+
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
+% supfile = supplementary materials
 
@@
 Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
 %=================================================================
@@
 
-%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
-
-\newcommand{\ms}{\si{\milli\second}}%
-
-% a simpler annotation system:
-\usepackage{soulutf8}
-\usepackage{color}
-\newcommand{\mynote}[1]{{\sethlcolor{yellow}\hl{#1}}}
-% \usepackage{trackchanges}
-\usepackage[finalnew]{trackchanges}
-\addeditor{Antoine}
-\addeditor{Amélie}
-\addeditor{Camille}
-\addeditor{Jean}
-\addeditor{JN}
-\addeditor{Laurent}
-
-%\newcommand{\note}[1]{}
-%
 %=================================================================
@@
 % Full title of the paper (Capitalized)
-\Title{Precise spiking motifs in neurobiological and neuromorphic data}
-%
+\Title{Review on precise spiking motifs in neurobiological and neuromorphic data}
+
 % MDPI internal command: Title for citation in the left column
 \TitleCitation{Precise spiking motifs}
-%
+
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0002-3107-4788} % AnG % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0002-9536-010X} % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0003-3916-0514} % AmG
-\newcommand{\orcidauthorC}{0000-0002-0054-2729} % CB
-\newcommand{\orcidauthorD}{0000-0002-9238-6654} % JN
-\newcommand{\orcidauthorE}{0000-0001-8821-5556} % JM
-\newcommand{\orcidauthorF}{0000-0002-9536-010X} % LP
-%
+\newcommand{\orcidauthorD}{0000-0002-3107-4788}
+
 % Authors, for the paper (add full first names)
 \Author{
-Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
-\orcidA{}, 
-Amélie Gruel $^{2}$\orcidB{}, % $^{2,\dagger,\ddagger}$,
-Camille Besnainou $^{1}$\orcidC{},
-Jean-Nicolas Jérémie $^{1}$\orcidD{},
-Jean Martinet $^{2}$\orcidE{},
-and Laurent Perrinet $^{1,}$\orcidF{}*}
-%
+Antoine Grimaldi $^{1,\dagger,\ddagger}$\orcidA{}, 
+Amélie Gruel $^{2,\dagger,\ddagger}$, 
+Jean Martinet $^{2,\ddagger}$ 
+and Laurent Perrinet $^{1,}$\orcidD{}*}
+
 %\longauthorlist{yes}
-%
+
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent Perrinet}
-%
+\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Grimaldi, A.; Gruel, A.; Besnainou, C.; Jérémie, J.-N.;Martinet, J.; and Perrinet, L.}
+\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
-%
+
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
 \address{%
-$^{1}$ \quad INT UMR 7289, Aix Marseille Univ, CNRS; 27 Bd Jean Moulin, 13005 Marseille, France\\
+$^{1}$ \quad INT UMR 7289, Université Aix Marseille, CNRS; 27 Bd Jean Moulin, 13005 Marseille, France\\
 $^{2}$ \quad SPARKS,  Côte d'Azur, CNRS, I3S; 2000 Rte des Lucioles, 06900 Sophia-Antipolis, France
 }
-%
+
 % Contact information of the corresponding author
-\corres{Correspondence: laurent.perrinet@univ-amu.fr}
-%
+\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
+
 % Current address and/or shared authorship
-% \firstnote{Current address: Affiliation 3.} 
-% \secondnote{These authors contributed equally to this work.}
+\firstnote{Current address: Affiliation 3.} 
+\secondnote{These authors contributed equally to this work.}
 % The commands \thirdnote{} till \eighthnote{} are available for further notes
-%
+
 %\simplesumm{} % Simple summary
-%
+
 %\conference{} % An extended version of a conference paper
-%
+
 % Abstract (Do not insert blank lines, i.e. \\) 
 \abstract{
+%A single paragraph of about 200 words maximum.
+For research articles, abstracts should give a pertinent overview of the work.
+We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations.
+The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.
 Why do neurons communicate through spikes?
-By definition, spikes are all-or-none neural events which occur at continuous times.
-In other words, spikes are on one side binary, existing or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock.
-This stands in stark contrast to the analog representation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks.
-As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the profusion of recorded data.
-With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption --- a major societal issue in the era of the digital economy and global warming.
-In this review paper, we provide evidence from biology, theory and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
+By definition, a spike, or action potential, is a all-or-none event ---it can occur or not without further details--- at asynchronous timings, i.e.~it can occur at any continuous time - differentially to the discretized timing classically used in digital processing.
+In the living world, neural systems almost systematically use this so-called event-based representation, though we do not yet have a clear idea why.
+A better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the masses of recorded data.
+It is also an emerging challenge in computer science to allow the efficient exploitation of a new class of sensors and event-based computers, called neuromorphic, which could allow significant gains in computing time and energy consumption ---a major societal challenge in the age of the digital economy and of global warming.
+%
+The response of a biological neuron depends largely on the precise timing of the sequence of presynaptic spikes as they reach the basal dendritic tree.
+This \emph{event-based representation} present in the neuronal code is essential in understanding information processing and yet, most neuronal models do not take advantage of this minute temporal dimension.
+Our goal here is to bring an interdisciplinary perspective on the computational advantage of time series representations for the brain and for information processing machines.
+In particular, we will focus on the hypothesis that there exists in an assembly of neurons a representation based on a set of motifs of different relative spike times.
+Here, we will review current litterature on the detection of such motifs in generic raster plots.
+It is work in progress, where anybody interested can \emph{openly} join.
+%
+This hypothesis is directly inspired by neurobiological observations in the hippocampus, and it expands the capabilities of analog representations based on the firing rate by considering a representation based on repetitions of these motifs at precise times of occurrence.
+A mathematical formalization would be particularly well suited to neuromorphic computing, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+We will first review some biological and theoretical evidence in neural information processing.
+We will then present some models for the detection of such motifs in arbitrary raster plots, synthetic, biological or artificial (notably from event-based cameras).
+In particular, we will discuss models which exploit the variety of synaptic delays on the dendritic tree.
+Then, we will try to outline some possible strategies for learning these patterns and finally discuss possible perspectives.
 }
-%
+
 % Keywords
-\keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; spiking motifs; polychronization.)} 
-%
+\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+
 % The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
@@
 %\JEL{}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction: Importance of precise spike timings in the brain}\label{sec:time}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Is there a neural code?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% CONTEXT
-Neural activity is directly influenced by our immediate environment and by internal states and is structured in order to generate motor actions.
-The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection.
-A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings observed in neurobiological experiments.
-One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
-an explicit representation \add[Antoine]{of cognitive processes} within the neural activity.
-Nevertheless, we will use this terminology in all generality to denote the existence of a structure in neural activity.
-In this respect, it is \change[Antoine]{known}{reasonable to declare} that neural activity may be related to specific measurable variables\change[Antoine]{, and since Galvani's}{.
-Since Galvani's} experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
-A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered \change[Antoine]{by Lord Adrian}{at the beginning of the $XX^{th}$ century}~\citep{adrian_impulses_1926}.
-In this study, the frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb.
-In the scope of this article, we focus mainly on these spikes\change[Antoine]{which are brief (about one millisecond) impulses that propagate along the axons of neurons.
-First, these have the particular property of being essentially binary in their amplitude, i.e.
-they are prototypical, ``all-or-none''}{in vertebrate systems.
-They can be described as brief (about one millisecond) and prototypical, i.e.
-``all-or-none'', impulses that propagate along the axons of neurons.}
- \change[Antoine]{Typically, action potentials are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
-The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except notably in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
-\remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
-This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
-
-% GAP
-However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role and that neurons \change[Antoine]{may be as well integrators than synchrony detectors}{may be synchrony detectors as well as integrators}~\citep{abeles_role_1982}.
-\add[Laurent]{In particular, it is possible that the minute arrangement of temporal delays between neurons may provide a computational advantage.} We will investigate this very hypothesis in this review.
-%Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
-In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
-For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{of the rate coding hypothesis that relies only on the firing rate} by considering \change[Antoine]{a representation based on repetitions of spiking motifs at precise times of occurrence.}{representations based on the precise timing of single spikes.} \remove[Antoine]{This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.} Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004,dilorenzo_spike_2013}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons.
-A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
-Crucially, validating this hypothesis would also be crucial in our understanding of neural processes.
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Dynamics of vision and consequences on the neural code}%Precise timing in the neural code enables efficient vision}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Let us start with a focus on the state-of-the-art of the role of dynamics in vision.
-Broadly speaking, vision is the set of processes that allow us \change[Antoine]{to make sense of the luminous world}{to make sense of the world through luminous signals}, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
-In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between the topology of external visual space and its representation on the retina, called retinotopy.
-The origins of this neuroscientific question can be found in the first experiments from Pierre Flourens which, using lesions in animals, demonstrated the relationship between visual sensations and activity in the cerebral cortex~\citep{flourens_recherches_1842,pearce_marie-jean-pierre_2009}.
-This was also observed \remove[Antoine]{by Lord Adrian} when recording the activity of the frog's visual system~\citep{adrian_impulses_1926}.
-In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as the visual orientation or motion of elementary contours.
-For a large proportion of neurons, there is \change[Antoine]{remarkably a monotonic}{a remarkably monotonic} relationship between the contrast of visual features and the firing frequency of neurons.
-However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitute essential pieces to this puzzle~\citep{carandini_normalization_2012}.
-
-%----------------------------%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+
+\section{Introduction: Importance of precise spike timings in the brain}\label{introduction-importance-of-precise-spike-timings-in-the-brain}
+\subsection{Ultra fast neural codes for ultra fast vision}\label{ultra-fast-neural-codes-for-ultra-fast-vision}
+
+Let us start our review of the state of the art on the role of dynamics in vision by presenting some surprising results that have been obtained in neuroscience.
+Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
+Its first experiment consisted in asking subjects to categorize images that do or do not contain animals~\citep{ref-G3GlupPg}.
+The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neuronal activity.
+These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade~\citep{ref-B9iZWoq3}.
+
+This fast processing also explains the surprising experiments of fast serial detection which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal.
+The performances decrease progressively as the frequency of presentation of the images increases.
+However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image~\citep{ref-EqCCVwl2}.
+
+This speed of the visual cortex, although surprising, is quite compatible with the latencies that are recorded at the neuro-physiological level.
+Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque~\citep{ref-Vu87UU6A}, see Figure {[}\ref{fig:thorpe}{]}.
+This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas~\citep{ref-15aMjRHVK}.
+
 \begin{figure}
 \centering
-\includegraphics[width=.75\textwidth]{figures/visual-latency.pdf}
-% thorpe
-% \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
-\caption{{\bf Latency of the different processing steps along the human visual pathway.} Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
-Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
-In order, the retina is first activated ($20-40~\ms$), then the thalamus and the primary visual cortex (V1, $60-90~\ms$).
-This visual information  projects to the temporal lobe to reach the infero-temporal area (IT, $150~\ms$) for object recognition.
-It then reaches the prefrontal cortex (PFC, $180~\ms$) which modulates decision making to evoke the Motor Cortex (MC, $220~\ms$) which may mediate an action.
-This is eventually relayed through the spinal cord to trigger finger muscles, with latencies about $280-400~\ms$.}\label{fig:thorpe}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/visual-latency-estimate.jpg}
+\caption{From~\citep{ref-12dWDpY61}}\label{fig:thorpe}
 \end{figure}
-%----------------------------%
-In particular, there have been some remarkable findings when studying the dynamics of vision.
-For instance, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second.
-\change[Antoine]{His}{Their} experiments consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
-The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
-These results have been extended to several species, including primates.
-Different experimental protocols have shown for example that the motor response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
-This fast processing correlates with the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and decoding via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
-As expected, the performances decrease progressively as the frequency of presentation of the images increases.
-However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only $14~\ms$ per image.
-
-Although surprising, this speed of the visual cortex in primates is compatible with the latencies that are recorded at the neuro-physiological level.
-The rapid propagation of the visual information in the thalamus, then in the primary visual cortex takes about $45~\ms$ in the macaque~\citep{schmolesky_signal_1998} and about $60~\ms$ in humans~\citep{vanni_coinciding_2001}.
-This functioning of visual processing as a forward pass is most prominent in fast processing (see Figure~\ref{fig:thorpe}), and can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}.
-An important consequence of this speed of processing of vision is that it implies that processing is carried out using only very few spikes per layer.
-As a comparison, the latencies in macaque monkeys are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.\add[Laurent]{Note that, since maximal conduction speeds are roughly constant, theses latencies are comparable to that found in humans, with a ratio given by the physical size of the whole system.} It follows that if we consider that a behavioral response occurra in only $200~\ms$, it would involve about ten processing stages along the ``forward'' pathways of the visual system.
-Such processes were indeed efficiently reproduced in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
-At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
-This figure is inspired by similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
-As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{How precise spike timing may encode vectors of real values}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\add[Laurent]{Let's now focus on one processing step along a cortical pathway. }
-% reliability of spike timings
-Sensory data is most often represented by continuous values, such as the energy produced by a flow of photons that hits the different photoreceptors of the retina.
-How may such information be encoded in neuronal activity?
-The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
-For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it at first sight inconsistent with the temporal integration of firing rate.
-Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: while the first spike is aligned to stimulus' time, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{bryant_spike_1976,mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
-However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see~\citep{ermentrout_reliability_2008}).
-This is consistent with the differential role of different stimulus frequencies (for instance the gamma range around $80~\si{\Hz}$) on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
-However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
-
-% retina
-\remove[Antoine]{How may information about the sensory world be encoded in the spiking pattern?} At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
-In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
-These results have been extended to natural images and show a qualitatively similar behavior.
-The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
-Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
-This evidence found in the retina can be extended to other areas such as the visual cortex.
-%----------------------------%
-% C_MainenSejnowski1995_Perrinet.ipynb
+
+\subsection{How timing encodes analogous profile}\label{how-timing-encodes-analogous-profile}
+An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief impulses that propagate along the axons of neurons.
+These have the particularity of being essentially binary in their amplitude (that is to say, they are prototypical, all or nothing).
+An important consequence of the speed of processing is that it implies that it is carried out using only very few spikes.
+Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
+Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{ref-zZUVkTfZ} but also the dependance across neurons~\citep{ref-13T2Gne9z}.
+
+Numerous studies demonstrated the importance of precise timing in neural population activity~\citep{ref-BNK6T5Uk}, efficient encoding thanks to the use of spike latencies~\citep{ref-1DpSM1kpQ,ref-IjepoonT} or precise timing in the auditory system~\citep{ref-mxPMgOmW,ref-1ECJhg2GM}.
+All these findings, and more~\citep{ref-WZsEyOlK}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
+Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{ref-XsHCA4IZ}.
+
 \begin{figure}
 \centering
-\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
-\caption{\textbf{Reproducibility of the spiking response of a neuron.
-}The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
-The stimulus current value over time for a step stimulus (top left) and for a noisy one (top right).
-Trial repetitions of a Leaky Integrate-and-Fire neuron stimulated by the stimulus on the upper row (middle row).
-Membrane potential is represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row).
-While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{frozen} noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
-See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of the results from~\citep{mainen_reliability_1995} using a simple LIF model.}\label{fig:mainen}
+\includegraphics{images/roc.jpg}
+\caption{Latency coding.
+An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency.
+In this example, one generates at most one spike per neuron.
+From~\citep{ref-YX5wqWH4}}\label{fig:roc}
 \end{figure}
-%----------------------------%
-
-% %----------------------------%
-% \begin{figure}
-% \centering
-% \mynote{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
-% \caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
-% }\label{fig:roc}
-% \end{figure}
-% %----------------------------%
-In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
-First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered, for instance for sound localization~\citep{chase_first-spike_2007}.
-Note also that timing is not entirely sensory or internal, but can be used as a general neural coding principle.
-In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
-Timing, at least in animals, may thus be fundamentally embodied and situated.''  Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}.
-These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}.
-One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron by using the rank of neurons' activation~\citep{delorme_spikenet_1999,bonilla_analyzing_2022}.
-Another class of artificial SNNs uses precise spike timing as a metric in order to determine the structure of the network in order to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods.
-The subsequent surrogate gradient method is now widely used in methods that attempt to transfer performance from analog (CNNs) to spike-based (SNNs) architectures~\citep{zenke_remarkable_2021}.
-This type of modeling often uses the classical task of categorizing images developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}.
-For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition or~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-However, the performance of SNNs is still lagging compared to that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
-\change[Antoine]{Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of event-based cameras}{Improvements in this new generation of Artificial Neural networks (ANN) would bring major advances in terms of efficient computations in machine learning.
-They would benefit in particular to a new generation of cameras called Silicon Retinas}~\citep{gallego_event-based_2022} (see Section~\ref{sec:neuromorphic}).
-\remove{These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation}%~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developed in Section~\ref{sec:neuromorphic}).This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.
-
-\change[Antoine]{To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.}{Even if technology lags far behind biology, this introduction demonstrates the importance of timing in neural processes and we will further review the role of precise spike timing in neural assemblies.
-We start by reviewing the different hypotheses that aim at deciphering the neural code with spatio-temporal spiking motifs.
-After listing some biological evidence for the use of precise spike timing we review some computational models and neuromorphic technics that add this temporal dimension to their computations.}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Role of precise spike timing in neural assemblies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
-There are multiple experimental evidences that neurons use different strategies to communicate.}~\cite{maunsell_functional_1983} \add[Antoine]{demonstrate that neurons of visual area MT in macaque monkeys encode motion direction through population activity, offering robustness to single neuron activity fluctuations.
-For rapid image processing, it has been observed that neurons in the salamander retina represent the spatial structure of an image with the relative timing of their first spikes}~\citep{gollisch_rapid_2008}\add[Antoine]{.
-In the primary visual area of macaque monkeys, the phase of firing convey information about naturalistic stimuli that cannot be described by firing rates alone}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{.
-Among a variety of strategies to communicate, so far we have also evidence that neurons can use correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}.
-\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
-Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\citep{maunsell_functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase--firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
-In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space.
-In all generality, this representation can encompass all the previous ones except for rate coding which is not defined locally in time.
-In this section, we choose to described different hypotheses making use of spatio-temporal patterns of spikes that can be propagated among neural assemblies.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{One first hypothesis: synchronous firing in cell assemblies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
-The book gradually leads the reader from the macroscopic cortical anatomy and standard electro-physiological properties of single neurons to neural network models.
-While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-The book then demonstrates that neural assemblies could form so-called ``synfire chains,'' that is, showing the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
-More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}: \add[Antoine]{"cells that fire together wire together"}.
-Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
-One possible function of this synchronization may serve the binding of information distributed in the brain~\citep{singer_visual_1995, roelfsema_visuomotor_1997}.
-
-Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}.
-Interestingly, these authors showed that ``accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
-Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
-These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
-\change[Antoine]{Note also that synchronicity may explain some unintuitive results.
-Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity.
-However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons.
-Among different explanations including travelling waves}~\citep{ben-yishai_traveling_1997}\add[Antoine]{, a synchronous activity at the synaptic level may be sufficient to elicit activity in the cortex.}~\citep{bruno_cortex_2006}.
-
-Theoretically\add[Antoine]{, even if vertebrate's neural system is not likely to be modeled only by such deterministic connectivity}~\citep{deneve_bayesian_2004, ballard_dual_2011}, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
+
+At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{ref-IjepoonT}.
+Notably one can appreciate in figure 1 that the response of ganglion cells to visual gratings that are flashed onto the retina.
+The authors showed that the neuronal response could be encoded in the latency of the response and not only in the frequency of discharge as is often assumed.
+In figure 4 of the same article, these results are extended to natural images and show a qualitatively similar behavior.
+The conclusion of the authors is that the discharge latency of the neurons allows to encode spatial characteristics of the image.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{ref-1H1Omh73g}.
+Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{ref-UlbjMU0W}.
+
+Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{ref-513xqzmL}.
+Many models have used these properties in temporal coding to build fast image categorization networks.
+These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{ref-1DpSM1kpQ}.
+This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{ref-7K80jcx9,ref-fRudGmR1}.
+This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events~\citep{ref-Oet1bgdb}.
+This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{ref-15cR83gqd}.
+Note also that timing is not entirely sensorial or internal but in~\citep{ref-1E8gkynCm}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{ref-QK8RfvTZ}.
+
+\subsection{From synfire chains\ldots{}}\label{from-synfire-chains}
+The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
+For instance, the firing rate is highly irregular~\citep{ref-sYHyiGsP}, which makes it inconsistent with temporal integration of random EPSPs.
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{ref-XsHCA4IZ} (see Figure {[}\ref{fig:mainen}{]}).
+When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
+This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in~\citep{ref-1AbvCTG03} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
+However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
+
+\begin{figure}
+\centering
+\includegraphics{figures/MainenSejnowski1995.png} % TODO replace by the simulation
+\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus.
+While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} nois at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
+\end{figure}
+
+In his book ``Corticonics''~\citep{ref-19MS6Bbwp}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
+The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models.
+While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{ref-lT1Ato7P}.
+The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion.
+Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{ref-R2E2wRFg}.
+
+It was shown that a simple model may allow the propagation of such synfire chains~\citep{ref-xbqYi2PD}.
+This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size.
 Each neuron of one group is connected by an excitatory synapse to the next.
 When a pulse is elicited in the first group, this may generate a spike in the next group.
-Depending on the concentration of synaptic weight values, this new activity may become more or less synchronized with respect to that of the previous pulse (as measured by the standard deviation of spike times within the pulse).
-Recursively applying this to a sequence of groups within a chain generates either a synfire propagation or not.
-A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
-A crucial aspect of this emergence is explained by the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
-This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling so-called push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
-Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
-In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013}.
-% (see Figure~\ref{fig:pynn}).
-%
-%----------------------------%
-%\begin{figure}
-%\centering
-%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
-%\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
-The background noise  is implemented as random Gaussian current.
-The state space gives %the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
-%\end{figure}
-%----------------------------%
-%----------------------------%
+Depending on the weight value, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse).
+Recursively applying this to a sequence of groups generates either a synfire propagation or not.
+A simple simulation is shown in Figure {[}\ref{fig:diesman}{]}.
+A crucial aspect of this emergence is in particular the balance between excitation and inhibition~\citep{ref-DL2NpHxr}.
+This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{ref-mszJ8wdv}.
+
 \begin{figure}
 \centering
-\includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
-\caption{\textbf{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.} The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
-Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreases from $5$ to $1$~\ms).
-This generates a pulse after a certain processing delay in the second group with a different jitter.
-While the first two pulses progressively vanish in the following groups, starting from the third input pulse (with a jitter of $3~\ms$), it is propagated to the following groups.
+\includegraphics{figures/Diesmann_et_al_1999.png}
+\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
+The model consists of 10 groups of 100 neurons each.
+A pulse with a given jitter is generated in the first group which here generates a pulse after a certain processing delay in the second group with a lower jitter.
 This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
 \end{figure}
-%----------------------------%
-
-Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Indeed, simultaneously recorded activities of neurons in the primary motor cortex of monkeys exhibited context-dependent, rapid changes in the patterns of coincident action potentials~\citep{riehle_spike_1997,grammont_precise_1999}.
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas yet it is less clear \remove[Antoine]{, though,} how these populations organize dynamically to cope with the momentary computational demands.
-In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical ANNs.
-It relies on the simple remark that a serie of repeated coincidences is in itself an invariant.
-Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, or the pitch of a musical note.
-
-This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}.
-The synchronicity will generate rhythms at different ranges of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down.
-Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion, i.e.
-the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
-More generally, it can be shown that the phase of alpha oscillations (about $10$ Hz) is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
-The question still remains open as to know if this is an epiphenomenon or a working mechanism of the neural code.
-%
-\subsection{A further hypothesis: travelling waves}
-%
-To further investigate the role of precise timing, let us also focus on the role of differential timings in an assembly of neurons.
-As we have seen, a visual feature will induce the firing of different cells at different latencies~\citep{celebrini_dynamics_1993}.
-Using intracellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is, a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
-In particular, it was shown that the network of the so-called horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within that area.
-For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions.
-Such interactions may underlay anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}.
-The underlying process could be the emergence of propagating waves on the surface of the cortex.
-
-Propagating waves in the neuronal response occur in many excitable media and were found in neural systems, for instance in the retina\note[Antoine]{added ref to Feller}~\citep{feller_dynamic_1997} or the neocortex~\citep{bienenstock_model_1995}.
-While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries.
-One possibility is that these waves were systematically missed in trial-averaged data, due to variability.
-A recent work~\citep{muller_stimulus-evoked_2014} presents a method for detecting propagating waves in noisy multichannel recordings.
-Applying this method to single-trial voltage-sensitive dye imaging data\note{The reviewer is right to point that calcium imaging may not be the best tool to capture precise spiking motif as both spatial and temporal resolution are not that great.
-We argue that few imaging techniques can capture the precise spike location, even good spike sorting algorithms.
-Also, here we study a propagation of activity defining a spatio-temporal pattern that may not have the same precision as some spiking motifs we describe later.}, the authors show that the stimulus-evoked population response in the primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
-A network model suggests that this reliability is the hallmark of the horizontal fiber network of superficial cortical layers.
-Propagating waves with similar properties occur independently in the secondary visual cortex, but maintain precise phase relations with the waves in the primary visual cortex.
-These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatio-temporal frame for further neural interactions.
-
-% \begin{figure}
-% \centering
-% %\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
-% \caption{a Spike raster for repeated presentations (%N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
+
+Attempts have been made to detect such synfire chains in neurobiological data~\citep{ref-qVu1OkoS}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies~\citep{ref-xBZqHOuN} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{ref-p7xk5w4f}.
+In particular, this was implemented using the pyNN language~\citep{ref-33SRV9MB} both in CPU-based and neuromorphic hardware~\citep{ref-DiIieN8s} (see Figure {[}\ref{fig:pynn}{]}).
+
+\begin{figure}
+\centering
+\includegraphics{figures/fnins-07-00011-g004.jpg}
+\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
+The background is only utilized in the original model, where it is implemented as random Gaussian current.
+For the presented hardware implementation it has been omitted due to network size constraints.
+As compensation for missing background stimuli, the resting potential was increased to ensure a comparable excitability of the neurons.
+The state space gives the duration of synfire chains as a function of the protocol's parameters and compares (C) CPU to (D) neuromorphic implementations.}\label{fig:pynn}
+\end{figure}
+
+Note also that synchronicity may explain some unintuitive results.
+Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
+However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{ref-18ZU9zP9l}.
+
+Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{ref-wVZpQvSk}.
+Interestingly, authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{ref-1HJjmmSpm} and showed some tuning to to movement direction and reaction time~\citep{ref-FZ9X31Qa}.
+It is important to note that synchronous events tended to lock to LFP beta waves~\citep{ref-Ejy1sAyh}, and was extended to larger assemblies~\citep{ref-Glu4NlLC} using statistical methods (see~\citep{ref-sec:detection}{\textbf{sec:detection?}}{]}).
+
+In~\citep{ref-3GlNNxH5}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
+It relies on the simple remark that a series of repeated coincidences is in itself an invariant.
+Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, the pitch of a musical note.
+
+\subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}
+TODO :~\citep{ref-zNDCo9Hy}
+
+Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
+This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{ref-Piao2Wzf}.
+This show precise spatio-temporal differences in this recurrently connected assembly.
+The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
+A similar dataset used in~\citep{ref-2AsQOOsE} is available from M.
+Berry's lab~\citep{ref-1FU0Y7Erc}.
+
+Propagating waves occur in many excitable media and were recently found in neural systems from retina to neocortex.
+While propagating waves are clearly present under anaesthesia, whether they also appear during awake and conscious states remains unclear.
+One possibility is that these waves are systematically missed in trial-averaged data, due to variability.
+A recent work~\citep{ref-hGYLURVS} presents a method for detecting propagating waves in noisy multichannel recordings.
+Applying this method to single-trial voltage-sensitive dye imaging data, they show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials.
+A network model suggests that this reliability is the hallmark of the horizontal fibre network of superficial cortical layers.
+Propagating waves with similar properties occur independently in secondary visual cortex, but maintain precise phase relations with the waves in primary visual cortex.
+These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatiotemporal frame for further neuronal interactions.
+
+\begin{figure}
+\centering
+%\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
+\caption{a Spike raster for repeated presentations (%N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line).
 A single-trial LFP trace is plotted in gray, and the average LFP response is plotted in black.
@@
 The relative power between baseline and evoked LFP fluctuations was not statistically different from parity (median = 1.23 dB, p = 0.087 two-tailed Wilcoxon's ranked-sum test).
-c An example of spontaneous LFP fluctuations structured as a travelling wave recorded from a spatially distributed multielectrode array in marmoset area MT.
+c An example of spontaneous LFP fluctuations structured as a traveling wave recorded from a spatially distributed multielectrode array in marmoset area MT.
 d Histogram of spontaneous spike probability as a function of the generalized phase of the LFP during fixation.
-e The average evoked response to low contrast stimuli was stronger when a more excitable phase (+ / - \pi rad) of a spontaneous travelling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test).
+e The average evoked response to low contrast stimuli was stronger when a more excitable phase (+ / - \pi rad) of a spontaneous traveling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test).
 Data for panels c--e modified from Davis et al.17 with permission.
-% }\label{fig:davis}
-% \end{figure}
-
-More recently, novel multi-unit recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
-Authors review these findings by considering the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function.
-In particular, spontaneous travelling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
+}\label{fig:davis}
+\end{figure}
+
+More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex~\citep{ref-vsKvZJ82}.
+Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
+In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{ref-BNK6T5Uk}.
 Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise.
-However, fluctuations of spontaneous activity, often organized as travelling waves, shape stimulus-evoked responses and perceptual sensitivity.
+However, fluctuations of spontaneous activity, often organized as traveling waves, shape stimulus-evoked responses and perceptual sensitivity.
 The mechanisms underlying these waves are unknown.
 Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings.
-In that paper, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
-They find that spontaneous travelling waves are a general property of these networks.
-The travelling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave.
+Here, authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
+They find that spontaneous traveling waves are a general property of these networks.
+The traveling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave.
 Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
-Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \textit{in vivo}.
-Such waves also occur in motor areas and~\citet{linden_movement_2022} have recently presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
-
-Interestingly, it can be shown that these travelling waves could have a measurable impact on the activity of the visual cortex.
-This has been illustrated in a recent study investigating the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
-The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in relatively close visual proximity.
-The lrAM is the core building block underlying the use of sequences of images to induce the perception of smooth, realistic visual scenes, which is at the base of movies seen in cinema theaters.
-In this study, the authors used voltage-sensitive dye imaging to record the activity of the primary visual cortex of macaque monkeys to the presentation of the pair of dots presented independently or in conjunction.
-A probabilistic modelling has shown that the activity of the joint presentation induced a suppressive wave in the direction opposed to the perceived direction, shaping the formation of a wave of propagation travelling at a speed compatible with the perceived motion.
-A computational model validated the hypothesis that this process could be mediated by diffusion in the horizontal layers connecting the different locations within this cortical area.
-In summary, the study by~\citet{chemla_suppressive_2019} gave a multi-disciplinary account to demonstrate the effect of travelling waves in the visual cortex.
-%
-% Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~ https://dx.doi.org/10.1073/pnas.1902653116
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A rediscovered hypothesis: precise spiking motifs in cell assemblies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Travelling waves indicate that spatio-temporal correlations could play an important role in shaping neural activity.
-For instance, statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-In particular, this was put in evidence by analyzing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}, which shows precise spatio-temporal differences in this recurrently connected assembly.
-The different aspects of information in the data are evaluated by a decoding strategy, highlighting the role of correlations.
-Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from Michael Berry's lab~\citep{berry_spike_2022} and allows \change[Antoine]{performing and validate}{testing in order to validate or falsify} these hypotheses.
-However, \remove[Antoine]{a limit of explaining such data with travelling waves is that} in theory, a \add[Antoine]{cortical} travelling wave would be stationary, which is incompatible with the limits in space and time of a neural system.
-Recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
-~\citet{miller_visual_2014} show that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
-Experiments are performed using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity.
-The study proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
-Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions.
-% (Fig.
-7).
-
-From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Explictly decoding spiking motifs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
-Going further, researchers have found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \textit{in vivo} and \textit{in vitro}.
-These patterns repeat after minutes, maintaining millisecond accuracy.
-Indeed,~\citet{ikegaya_synfire_2004} demonstrated that in cortical activity, one can find a repetition of several motifs in spike activity (duration around 1s +/- .5 s, some events in motifs are of similar size but sometimes absent).
-These sequences can be specific of a particular layer or column, can be synchronized with network activity oscillation, and can involve several cells.
-They also demonstrated that these sequences can form super sequences, so-called \emph{cortical songs}.
-It consists of the assembly of several sequences which repeat in a specific order with a compressed timing.
-\emph{In vivo} spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3-10 cells out (of about 900), and lasting up to 3 seconds.
-Sequences have specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and are associated with a structured spatial organization of the neurons that formed them.
-%In cortical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
+Further, by modulating local network state, they can shape responses to incoming inputs as observed in vivo.
+
+~\citep{ref-17EU8fPBv} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
+
+\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{ref-yLVM9IXi}.
+It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
+Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials.
+Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
+Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.
+
+~\citep{ref-OAorhMRa}
+
+\subsubsection{TODO \ldots{} and to synfire braids and polychronous groups}\label{todo-and-to-synfire-braids-and-polychronous-groups}
+\begin{itemize}
+ \item
+ ~\citep{ref-HjDb1gnq} : from synfire chains to Synfire braids
+\end{itemize}
+
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{ref-KS4oxnst,ref-ga3ZTFC7}.
+Notably, Izhikevich~\citep{ref-SM9G0xBK} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
+This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.
+
+sparse in time and space {[}2{]} AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. {[}3{]} CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
+
+\begin{itemize}
+\item
+  recent theories of binding by synchrony : Fries 2005 trends cog neuro with spikes arriving at peak susceptibility (top of a cycle) or down, van Rullen, Laura Dugué
+\item
+  cell assembly hypothesis: neurons coordinate their activity through the formation and repetitive co-activation of groups.
+https://link-springer-com.insb.bib.cnrs.fr/article/10.1007/s10827-021-00801-9 Hebb DO.
+The organisation of behaviour: a neuropsychological theory.
+New York: Science Editions; 1949.
+\item
+  A notable exception is the polychronization model of Izhikevich~\citep{ref-SM9G0xBK}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule.
+In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
+\end{itemize}
+
+\subsection{Are there precise heterosynaptic spiking motifs in the brain?}\label{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}
+Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
+Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ref-Mf6h1KNY}, packets~\citep{ref-2GovoaYw} or hippocampal sequences~\citep{ref-SVfasyCp,ref-PUNGVou} and on the role of such precise spike timing in downstream information transfer and coding~\citep{ref-19fAM6hcv,ref-13dQ0R4UM,ref-MfmcMsqA}.
+This is for instance relevant in sensory pathways in vision~\citep{ref-1EHZgcKyL}, audition~\citep{ref-FlKJw2VL}, olfaction~\citep{ref-AhSRXBJ2}{]} or touch~\citep{ref-gDinei4a}.
+In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
+In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs (Izhikevich, 2006) which he defined as polychronous groups.
+
+Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space~\citep{ref-13dQ0R4UM}.
+These sequences can be internally generated~\citep{ref-SVfasyCp,ref-PUNGVou} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016).
+Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from~\citep{ref-qI2HUEyy} below).
+As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
+
+\begin{figure}
+\centering
+%\includegraphics{https://www.pnas.org/cms/10.1073/pnas.1718518116/asset/ab9a48f6-b28f-40c1-978e-6f6dbec39b98/assets/graphic/pnas.1718518116fig01.jpeg}
+\caption{Mixed distance and duration representation in CA1.
+(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
+Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
+The black line represents speed of the mouse.}\label{fig:haimerl}
+\end{figure}
+
+TODO: read~\citep{ref-MfmcMsqA} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
+
+In~\citep{ref-4D8Jnhkz}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
+They develop a HMM model for quantifying the transitions.
+``In this study, van Kempen et al.~show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
+Top-down attention drives interareal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.''
+
+\subsection{Cortical songs}\label{cortical-songs}
+\begin{itemize}
+ \item
+  Ikegaya and colleagues~\citep{ref-Mf6h1KNY} worked on spontaneous activity in vitro and in vivo.
+They demonstrated that in cortical activity, we can find a repetition of several motifs.
+In PSCs, but also in spike activity.
+These sequences repeat after minutes and have a precise spatio temporal structure with a ms precision.
+They can be specific of a particular layer or colomn, are synchronized with network activity oscillation and can involve several cells.
+They also demonstated that these sequences can form supersequences : the cortical songs.
+It consist of the assembly of several sequences which repeat in a specific order with a compressed timing.
+\item
+  ``We find precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons in vivo and in vitro. These patterns repeat after minutes, maintaining millisecond accuracy.''
+\item
+  Precisely repeating motifs of spontaneous synaptic activity in slices: duration around 1s +/- .5 s.
+Some events in motifs are of similar size but sometimes absent - better described by Bernouilli than SE (and covariance)
+\item
+  \emph{in vivo} spontaneous activity also reveals repeating sequences. About 3000 sequences, each involving 3-10 cells out of about 900, and last up to 3 seconds
+\item
+  topography: ``Sequences had specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster (Fig.
+4, A and B, and fig.~S3B).
+(\ldots) Therefore, repeating temporal patterns of activation (\ldots) were associated with a structured spatial organization of the neurons that formed them.''
+\item
+  ``Cortical songs: modular assemblies of repeated sequences'': hierarchical detection.
+\item
+  in cotical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time.
 Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
-% %----------------------------%
-% \begin{figure}
-% \centering
-% \includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
-% \caption{Fig.
+\end{itemize}
+
+\begin{figure}
+\centering
+\includegraphics{images/Ikegaya2004zse0150424620001.jpeg}
+\caption{Fig.
 1.
-from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity \textit{in vitro} and \textit{in vivo}.
+from~\citep{ref-Mf6h1KNY} Repeated motifs of spontaneous synaptic activity in vitro and in vivo.
 (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
-Panels show segments (red) of the same voltage-clamp recording from the same cell, repeating seconds or minutes after the initial occurrence (blue).
+Panels show segments (red) of the same voltage-clamp recording from the same cell repeating seconds or minutes after the initial occurrence (blue).
 Arrows indicate timings of repeated PSCs.
@@
 The top traces show the motifs superimposed on each other (blue, green, and red), the middle traces show these same traces individually, and the bottom traces show temporally magnified regions of the motifs (a to c).
-(C) Repeated sequences of intracellular current-clamp recordings \textit{in vivo}.
+(C) Repeated sequences of intracellular current-clamp recordings in vivo.
 Two (i) and three (ii) repetitions of motifs are shown.
@@
 In (i), the blue trace is shifted --2.75 mV; in (ii), the blue trace is shifted --1.58 mV, and the green +0.79 mV.}\label{fig:Ikegaya2004}
-% \end{figure}
-% %----------------------------%
-
-Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
-This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{cleland_construction_2014, kashiwadani_synchronized_1999, rinberg_speed-accuracy_2006} or touch~\citep{johansson_first_2004}.
-In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
-These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and may be formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}.
-In that protocol, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}).
-As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
-%
-%----------------------------%
+\end{figure}
+
+It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{ref-Jzr1gwsr} which uses noise patterns.
+They " used random waveforms to probe the formation of new memories for arbitrary complex sounds.
+A behavioral measure was designed, based on the detection of repetitions embedded in noises up to 4 s long." The task is to detect the repetition of the same (frozen) noise within a trial.
+" Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block." they showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case.
+Note that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``Time reversal had no significant effect on the RefRN advantage'' (quite surprising).
+The Learning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).
+
+Such observations suggest that neuronal groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity~\citep{ref-UGHHptIH}.
+This work shows, using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity, that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+It proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes.
+Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions (Fig.
+7).
+
+\subsection{outline}\label{outline}
+A natural candidate to use these precise temporal patterns in the brain is to use Spiking Neural Networks (SNNs)~\citep{ref-DUpFdm8C}.
+The approach which is at present most prominent in the SNNs community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures.
+One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, for instance with a surrogate gradient.
+This approach is quite successful, and SNNs approach in some case the performance of Deep Learning algorithms, for instance on the N-MNIST dataset for categorizing digits in a stream of events.
+However, most biological neural systems use spikes and are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.
+There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+Our approach will be to focus on the temporal representation of information directly.
+In particular, our objective is to fully exploit the capacity of spiking neurons to detect synchronous patterns.
+
+{[}
+\textbf{TODO: a paragraph on our existing work}
+Our approach would be distinct than these approaches from us and colleagues as we will directly deal with delays in the system at the presynaptic level.
+
+there are temporal delays in the nervous system, both at the neural~\citep{ref-f7gkBktf} and behavioral~\citep{ref-1BSbXPPDF} levels.
+Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
+
+may be used for motion detection and interpolation~\citep{ref-WGEaBYgy}~\citep{ref-7GEdzh5o}
+
+Remarkably, novel neuromorphic chips use a representation similar to that of real neurons~\citep{ref-Oet1bgdb}.
+For example, event-based cameras provide a stream of binary asynchronous events signaling detectable changes in luminance, and information is represented by these spike-based temporal motifs, hence their name ``silicon retinas'' (see Figure \ref{fig:silicon_retina}).
+For such devices, it is crucial to better understand the potential of using such event-based representations in order to devise novel algorithms.
+
 \begin{figure}
 \centering
-\includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
-\caption{\textbf{An example of a precise temporal motifs seen in cortical slices.} In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
-(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
-Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}
+\includegraphics{figures/event_driven_computations.png}
+\caption{A miniature, event-based ATIS sensor.
+Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).}\label{fig:silicon_retina}
 \end{figure}
-%----------------------------%
-
-It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some frozen noise patterns.
-In particular, they used random waveforms to probe the formation of new memories for arbitrary complex sounds.
-A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \si{\second}$ long.
-%The task is to detect the repetition of the same (frozen) noise within a trial.
-Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block.
-They showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case.
-Note also that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``time reversal had no significant effect on the detection accuracy (which is quite surprising).
-The learning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).
-
-That results suggest that precise spiking motifs are not necessarily grouped on a topography and can be forming apparently randomly arranged connections.
-Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
-The rest of this review will be devoted to present evidence for the use of such precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
-As a summary, it seems that such precise structural information is essential to the neural code and that it seems imperative to include this information in decoding algorithm for a better understanding of neural activity.
-%
-% TODO: re-read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
-% \subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
-% \begin{itemize}
-%  \item
-%   Stochastic sampling links perceptual uncertainty to neural response variability
-% \item
-%   Model accounts for independent changes in strength and variability of responses
-% \item
-%   Model predicts relationship between noise, signal, and spontaneous correlations
-% \item
-%   Stimulus statistics dependence of response statistics is explained
-% \end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Understanding precise spiking motifs in neurobiology}\label{sec:detection}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding neural activity from firing rates}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we will review current evidence on how we may take advantage of spiking motifs in neurobiology, that is, in an effort to understand actual recordings from biological neural tissues.
-In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
-For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process\note[Laurent]{adding reference}~\citep{simoncelli_characterization_2003}.
-As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves, that is, its selectivity to a range of features~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.
-This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression.
-An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
-In this particular paper, authors fit summary statistics of neural data with a (differentiable) spiking network model.
-The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995}) and embedded with recurrent dynamics.
-In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the \note[Laurent]{now citing the URL}publicly available V1 dataset~\citep{kohn_utah_2016} which allows supervising the model with the input being the movie and the output the spikes recorded.
-These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
-\change[Antoine]{Using the right metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.}{Carefully picking the more appropriate metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.} Importantly, these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
-%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding neural activity using spike distances}
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-There are different solutions to provide with a distance between two given spike trains.
-A known measure is the Victor-Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
-\change[JN]{In a later study, Van Rossum (2001) have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citep{van_rossum_novel_2001}.
-\change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.
-Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure \change[JN]{which}{.
-An interesting conclusion since cross-correlation measure is} which is adapted to the event-based nature of spiking signals.
-These observations lead to the intuition that each distance may be as well  as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
-% ~\citep{}
-
-%Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
-Concerning spike timings,~\citet{levakova_review_2015} \change[Antoine]{has}{have} \change[JN]{reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning}{reviewed existing methods for estimating the latency of neural responses that include Bayesian binning.}.
-\change[JN]{More broadly }{Alternatively }, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
-This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}.
-A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}.
-By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}.
-This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[Laurent]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
-
-Another important algorithm, called SPOTDisClust, is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}.
-They introduced an unsupervised method based for their detection from high-dimensional neural ensembles.
-The algorithm measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other.
-Many approaches to this problem are supervised.
-In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for determining hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}.
-In SPOTDisClust, the learning is unsupervised.
-It uses the prior that there is only one spike per pattern.
-Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data.
-% (see Figure~\ref{fig:SPOTDisClust}).
-The limits of this method are that it is computationally complex, block-based and strictly specialized for the task at hand.
-To overcome these difficulties, a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
-%   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
-\add[JN]{Whether it is the distance between two given spike trains or a comparison of the spike timings, the complexity and the diversity of the methods used to measure them are witnesses of the growing interest of the integration of these measures in the understanding of the neural code.
-One of the steps to test their potential usefulness is then to scale these methods to larger amounts of data.}
-
-%----------------------------%
+
+This section has provided evidence that polychronous groups are an important apsect of information representation in biology with important application in data analysis and neuromorphic engineering.
+The rest of this review paper is organized as follows.
+
+First, we will review models for the detection of polychronous groups:
+
+\begin{itemize}
+\item
+  A crucial advantage of Spiking Neural Networks (SNNs) architectures lies in its processing of temporal information.
+Yet, most SNNs encode the temporal signal as an analog signal and try to ``cross-compile'' classical Neural Network to a spiking architecture.
+To go beyond the state-of-the-art, we will review here on one core computation of a spiking neuron, that is, is its ability to switch from the classical integrator mode (summing analog currents on its synapses) to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized.
+To overcome the diversity of input presynaptic patterns, we will explore different existing architectures to learn to detect stable ``polychronous`` events, that is, volleys of spikes which are stable up to certain synaptic delays.
+These models will be compared in light of neuroscientific and computational perspectives.
+We review theoretical and computational foundations of PG detection in models.
+\item
+  Application to Image processing using sparse spiking representations: Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals.
+design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
+\end{itemize}
+
+Finally we will discuss future avenues for effective PG detection and learning in event streams.
+
+\section{Theoretical models for the detection of precise heterosynaptic spiking motifs}\label{theoretical-models-for-the-detection-of-precise-heterosynaptic-spiking-motifs}
+\#sec:detection
+
+A popular model for the detection of latency patterns is the tempotron~\citep{ref-1GazUBEdT}.
+This model is in particular reviewed in~\citep{ref-QCyjy8wE}.
+The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
+This aims at extending the perceptron which does not incorporate a spike timing framework.
+The tempotron learning rule is derived by an optimization process and takes the form of a STDP rule.
+It is general consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{ref-EF6Larp8}.
+The limits of this model is that it's output is only binary and that it's storage capacity are limited.
+
+Some other models of latency motifs detection using STDP learning rules.
+For instance,~\citep{ref-Dd3kSDRW} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{ref-vePjgmRh} develops a form of spike-based, competitive learning applied for unsupervised learning.
+In~\citep{ref-RrgzWOSk}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
+
+In the following of this section, we will review theoretical models which directly tackle the problem of using precise heterosynaptic spiking motifs.
+
+\subsection{Izhikevitch's polychronization model}\label{izhikevitchs-polychronization-model}
+Considering the number of spikes required to exceed a voltage threshold, asynchronous signals are less efficient than synchronous signals.
+However, taking axonal propagation times into account, synchronous signals may not coincide significantly at the post-synaptic neuron contrary to asynchronous or polychronous signals.
+The term polychronous was first introduced in 2006, by E.
+Izhichevitch in~\citep{ref-SM9G0xBK}.
+He defines this term after pointing out a particular organization of the neurons of his spiking artificial neural network.
+The network is characterized by a timing-dependant learning rule for weights (STDP) and by fixed conduction delays between neurons.
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns.
+The neurons composing a group discharge at different times, but due to delays, the spikes reach the postsynaptic neuron at the same time.
+This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike and thus to the crossing of the voltage threshold and to the discharge of a spike.
+According to the STDP rule, the neurons involved in this activity will see their weight of synaptic connection increase and thus, constitute a polychronous group.
+Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system.
+
+{[}
+\textbf{TODO}
+
+-\textgreater{} insert fig 2 of~\citep{ref-doi}{\textbf{doi?}} 10.1162/089976606775093882{]}
+
+Thus, the learning of delays allowing this polychronous group organization may be useful to detect temporal sequences of interest.
+
+\begin{itemize}
+\item
+  Reproducing Polychronization: A Guide to Maximizing the Reproducibility of Spiking Network Models \url{https://www.frontiersin.org/articles/10.3389/fninf.2018.00046/full}
+
+  \begin{itemize}
+     \item
+    comes with python code
+  \end{itemize}
+\item
+  blog post by Paxon Frady \url{https://epaxon.blogspot.com/2012/07/izhikevich-2006-polychronization.html}
+\item
+  dynamic networks \& learning delays:
+
+  \begin{itemize}
+  \item
+    Huning H., Glunder H., and Palm G.
+(1998) Synaptic delay learning in pulse-coupled neurons.
+Neural Computation, 10:555--565.
+\url{https://www.deepdyve.com/lp/mit-press/synaptic-delay-learning-in-pulse-coupled-neurons-DGMiNHxp0A}
+  \item
+    \href{https://www.researchgate.net/publication/37921636_Dynamics_of_Self-Organized_Delay_Adaptation}{Eurich C., Pawelzik K., Ernst U., Cowan J., and Milton J.
+(1999) Dynamics of self-organazed delay adaptation.
+Phys.
+Rev.~Lett., 82:1594--1597.}
+  \item
+    The recent ``multi-neuronal spike sequence detector'' (MNSD) architecture integrates the weight- and delay-adjustment methods by combining heterosynaptic plasticity with the neurocomputational feature spike latency : \url{https://pubmed.ncbi.nlm.nih.gov/33679293/}
+  \item
+    an extensive (graph-centric) review on \href{https://arxiv.org/abs/2109.07618}{Synchronization in time-varying networks}
+  \end{itemize}
+\end{itemize}
+
+A Bayesian account: ``Previous methods for studying the PNG activation response to stimuli have been limited by the template-based methods used to identify PNG activation.
+In this letter, we outline a new method that overcomes these difficulties by establishing for the first time a probabilistic interpretation of PNG activation.
+We then demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system.''~\citep{ref-KS4oxnst}.
+Stimulation of a trained network produces the activation of a PNG, ie the propagation of firing activity through multiple layers due to convergent patterns of firing.
+
+Memory traces in dynamical systems~\citep{ref-o7rANZux} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
+This raises important questions about the fundamental limits of such memory traces and the properties required of dynamical systems to achieve these limits.
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+We introduce the Fisher Memory Curve (FMC) as a measure of the signal-to-noise ratio (SNR) embedded in the dynamical state relative to the input SNR.
+The integrated FMC indicates the total memory capacity.
+We apply this theory to linear neuronal networks and show that the capacity of networks with normal connectivity matrices is exactly 1 and that of any network of N neurons is, at most, N.
+A nonnormal network achieving this bound is subject to stringent design constraints: It must have a hidden feedforward architecture that superlinearly amplifies its input for a time of order N, and the input connectivity must optimally match this architecture.
+The memory capacity of networks subject to saturating nonlinearities is further limited, and cannot exceed $\sqrt N$.
+This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.
+We illustrate the generality of the theory by showing that memory in fluid systems can be sustained by transient nonnormal amplification due to convective instability or the onset of turbulence.''
+
+We address these issues by applying Fisher information theory to dynamical systems driven by time-dependent signals corrupted by noise.
+Memory capacity is constrained by architecture: ``This limit can be realized by feedforward structures with divergent fan out that distributes the signal across neurons, thereby avoiding saturation.''
+
+\section{Detecting precise spiking motifs in biological raster plots}\label{detecting-precise-spiking-motifs-in-biological-raster-plots}
+\{sec:detection\}
+
+\subsection{spike distances}\label{spike-distances}
+J. D. Victor and K. P. Purpura, ``Nature and precision of temporal coding in visual cortex: a metric-space analysis,'' J. Neurophysiol., vol.~76, pp.~1310--1326, Aug.~1996.
+
+M.
+C.
+W.
+van Rossum, ``A novel spike distance,'' Neural Comput., vol.~13, no.
+4, pp.~751--763, 2001.
+{[}21{]} D.
+Aronov and J.
+D.
+Victor, ``Non-Euclidean properties of spike train metric spaces,'' Physical Rev.~E (Statist., Nonlinear, Soft Matter Phys.), vol.~69, no.
+6, 2004.
+
+T. Kreuz, J. S. Haas, A. Morelli, H. D. I. Abarbanel, and A. Politi, ``Measuring spike train synchrony,'' J. Neurosci. Methods, vol.~165, no. 1, pp.~151--161, 2007. {[}23{]} H.
+
+Paper by~\citep{ref-LjkTSK3h} On Stability of Distance Measures for Event Sequences Induced by Level-Crossing Sampling
+
+Weyl's discrepency measure~\citep{ref-RYzP72lj} which may lead to the definition of a cross-correlation.
+
+Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~https://dx.doi.org/10.1073/pnas.1902653116
+
+A review of the methods for neuronal response latency estimation including bayesian binning~\citep{ref-Z7LAUup3}.
+
+\subsection{decoding neural activity}\label{decoding-neural-activity}
+In generic linear non linear lnl models, the output is assumed to be poisson.
+As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{ref-MgINa6bU}.
+This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
+
+Unitary event analysis is performed by a statistical model of coincidence detection~\citep{ref-13iJohyPk}.
+This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{ref-wVZpQvSk} for instance).
+
+Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{ref-1ErNTONy0}.
+By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range~\citep{ref-V9qkJFZU}.
+
+This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{ref-pExXxdMb} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{ref-d2m7wQpH}.
+
+~\citep{ref-doi}{\textbf{doi?}}{]}
+
+TODO: summarize ``Understanding Auditory Spectro-Temporal Receptive Fields and Their Changes with Input Statistics by Efficient Coding Principles'' - Spectro-temporal receptive fields (STRFs) have been widely used as linear approximations to the signal transform from sound spectrograms to neural responses along the auditory pathway.
+Their dependence on statistical attributes of the stimuli, such as sound intensity, is usually explained by nonlinear mechanisms and models.
+Here, we apply an efficient coding principle which has been successfully used to understand receptive fields in early stages of visual processing, in order to provide a computational understanding of the STRFs.
+According to this principle, STRFs result from an optimal tradeoff between maximizing the sensory information the brain receives, and minimizing the cost of the neural activities required to represent and transmit this information.
+Both terms depend on the statistical properties of the sensory inputs and the noise that corrupts them.
+The STRFs should therefore depend on the input power spectrum and the signal-to-noise ratio, which is assumed to increase with input intensity.
+We analytically derive the optimal STRFs when signal and noise are approximated as Gaussians.
+Under the constraint that they should be spectro-temporally local, the STRFs are predicted to adapt from being band-pass to low-pass filters as the input intensity reduces, or the input correlation becomes longer range in sound frequency or time.
+These predictions qualitatively match physiological observations.
+Our prediction as to how the STRFs should be determined by the input power spectrum could readily be tested, since this spectrum depends on the stimulus ensemble.
+The potentials and limitations of the efficient coding principle are discussed.
+~\citep{ref-s63GcujC}
+
+\subsection{Rastermap : decoding large-scale data}\label{rastermap-decoding-large-scale-data}
+\href{https://nbviewer.org/github/MouseLand/rastermap/blob/master/tutorial/tutorial.ipynb}{Rastermap} re-arranges neurons in the raster plot based on similarity of activity
+
+\begin{itemize}
+ \item
+  https://www.janelia.org/lab/stringer-lab
+\item
+  described in~\citep{ref-zc16plUu}
+\item
+  \href{https://github.com/MouseLand/rastermap}{rastermap}
+\item
+  deconvolution strategy
+\item
+  based on a linear model
+\end{itemize}
+
+\subsubsection{Stringer et al 2019, Nature~\citep{ref-brg2FFsb}}\label{stringer-et-al-2019-nature-stringer2019nature}
+\begin{itemize}
+ \item
+  ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
+Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
+\item
+  Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
+\item
+  Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
+\end{itemize}
+
+\subsubsection{Stringer et al 2019, Science~\citep{ref-1BPHUQLTf}}\label{stringer-et-al-2019-science-stringer2019science}
+\begin{itemize}
+\item
+  Stringer et al.~analyzed spontaneous neuronal firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
+The variability of neuronal responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
+\item
+  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{ref-1EMs3GS0B}
+\end{itemize}
+
+\subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
+\subsubsection{Paper by Grossberger, 2018~\citep{ref-9AK1cwY1}}\label{paper-by-grossberger-2018-grossberger2018}
+\begin{itemize}
+ \item
+  Temporally ordered multi-neuron patterns likely encode information in the brain.
+We introduce an unsupervised method, SPOTDisClust (Spike Pattern Optimal Transport Dissimilarity Clustering), for their detection from high-dimensional neural ensembles.
+SPOTDisClust measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other (SPOTDis).
+\item
+  Detecting these temporal patterns represents a major methodological challenge.
+\item
+  Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
+\item
+  Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
+\item
+  Lee AK, Wilson MA.
+A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order.
+J Neurophysiol.
+2004;92(4):2555--73.
+pmid:15212425
+\item
+  Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
+\item
+  only one spike per neuron: fig 1A = ``For each pattern and each neuron, a random position was chosen for the activation pulse.''
+\item
+  t-SNE projection with HDBSCAN labels shows that our clustering method can retrieve all patterns from the data.
+\item
+  data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
+\end{itemize}
+
 \begin{figure}
 \centering
-\includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
-\caption{\textbf{Detecting motifs using SPADE.} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}.
-This allowed to label different precise spike motifs which are denoted by different colors.
-Spikes belonging to the same motif have the same color.
-%As the synfire chain gets more precise, the detected patterns grow in size.
-}\label{fig:Diesmann_et_al_1999_spade}
+\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
+\caption{Fig 1 of~\citep{ref-9AK1cwY1}: ``Simulated example illustrating the steps in SPOTDisClust.
+A) Structure of five ``ground-truth'' patterns (\ldots).
+For each pattern and each neuron, a random position was chosen for the activation pulse.
+B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
 \end{figure}
-%----------------------------%
-
-% %%----------------------------%
-% \remove[LP]{
-% \begin{figure}
-% \centering
-% \includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
-% \caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm.
-\textbf{(A)} Structure of five ``ground-truth'' patterns (\ldots).
-For each pattern and each neuron, a random position was chosen for the activation pulse.
-\textbf{(B)} Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A).
-(© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
-% \end{figure}
-% %----------------------------%
-% }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Scaling up to very large scale data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
-Bridging across different spatial and temporal scales, combination of \textit{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
-The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurobiological knowledge to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
-As a consequence, there is a growing need to scale these methods to larger amounts of data.
-
-There are multiple \change[Antoine]{solutions}{approaches} which aim at tackling this problem.
-One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}.
-Basically, it rearranges neurons in the raster plot based on the similarity of their activity and applies a deconvolution strategy based on a linear model.
-\add[Laurent]{Yet this method was mainly tested on calcium imaging data which is known to add some imprecision to the timing of the original spiking activity.} The model is \note[Laurent]{now citing the URL}openly accessible~\citep{stringer_mouselandrastermap_2020} and has led to important discoveries.
-In~\citep{stringer_spontaneous_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
-Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements.
-In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
-%   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
-
-% \subsection{Extending the decoding at multiple time scales}
-In the paper by~\citet{russo_cell_2017}, the authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.
-It uses  sliding window as in~\citep{grun_unitary_2002} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
-They extend the measure to multiple lags~\citep{torre_synchronous_2016}.
-The core measure is based on a non-stationarity corrected parametric statistical test for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
-To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
-\change[Laurent]{They also introduce}{As originally introduced by}~\citep{kass_statistical_2005}, they use learnable time-warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
-At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
-They develop a Hidden Markov Model for quantifying the transitions.
-In particular, they show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
-Top-down attention drives inter-areal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
-Building such models for predicting changes of timings based on context, such as using a so-called change point model for blocked-based experimental protocols~\citep{pasturel_humans_2020}, could therefore provide useful prior information to enhance the decoding from neural activity.
-% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{What biological mechanism could allow learning spiking motifs?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Despite the evidence for the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject.
-The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviors allow us to put these doubts to rest.
-In the following paragraphs we will successively describe the first biological observations of delay learning, identify myelinization as an important actor and finally study this phenomenon at the molecular level.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Biological observations of delay adaptation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% First biological observations
-
-One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_functioning_1938}.
-Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{madadi_asl_dendritic_2018}.
-Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_effects_1992}.
-
-However, it was not until the study of the interaural time difference (ITD) by \remove[Amélie]{gerstnerneuronal1996}~\citet{carr_circuit_1990} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function.
-This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-\add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\citep{jeffress_place_1948}.
-\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
-According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.
-\add[Amélie]{This assertion confused the mean interspike interval, i.e.
-how often a neuron can fire, and the specific spike time, i.e.
-how precisely a spike can be generated.
-However, it has nevertheless contributed to the recognition of the importance of time in various biological mechanisms.} \remove[Amélie]{penacochlear2001 conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
-In the mouse somatosensory cortex, camontiming2019 found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
-the temporal disparity between left and right cochlear loci in the owl's brain.
-However, it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
-The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
-\add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.}
-
-\add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
-However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\citep{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.
-This mechanism adds to the axonal length delay, which was previously thought sole influence on the conduction velocity due to its anatomical soundness --- as}~\citet{seidl_mechanisms_2010}\add[Amélie]{ experimentally demonstrated, this mechanism by itself is not sufficient to explain the biological functionality but should be added to the one of axonal length delay.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The importance of myelination}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% The importance of myelination
-\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to \change[Amélie]{both the axon diameter by itself and the diameter of the axon wrapped in myelin}{the axon diameter, i.e.
-the amount of myelin wrapped around the axon}.
-Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
-Myelinization consists in ``two motions: the wrapping of the leading edge of the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}.
-Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the approx $1~\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_neuron-oligodendrocyte_2021}.
-\change[Antoine]{In}{On} average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
-This demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
-Thus, in several occasions, myelin has been identified as \change[Antoine]{a main}{an important} actor in the regulation of conduction velocity in neurons, i.e.
-axonal delay regulation.
-
-% role of myelination and neuronal delay
-
-\citet{fields_new_2015, fields_myelin_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments.
-This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations.
-Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{madadi_asl_dendritic_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_disruption_2020}, contextual fear conditioning~\citep{pan_preservation_2020} or oligodendrocyte precursor cells (OPCs) knock-out~\citep{wan_impaired_2020}.
-Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_new_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_neuron-oligodendrocyte_2021}.
-
-It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_new_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{madadi_asl_dendritic_2018}.
-Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_demyelination_2021}.
-~\citet{duncan_neuron-oligodendrocyte_2021} states that ``the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration''.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interplay of delay adaptation and neural activity}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% How is the neuronal activity detected ?
-However, one question remains: how do oligodendrocytes detect neuronal activity and regulate the myelinization accordingly?
-To answer this, we must first study the myelinization process.
-The OPCs first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
-Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
-Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as $Ca^{2+}$ activity~\citep{baraban_ca2_2018} of the neuregulin 1 growth factor~\citep{nave_axonal_2006}.
-The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity.
-Indeed,~\citet{cullen_periaxonal_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain.
-They also confirmed that the delay correlates with the level of skill acquisition.~\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
-
-A potential scenario for selective myelination on electrically active axons using non-synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
-This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_new_2015}.
-
-% myelination for axonal delay, but what about dendritic delay ? 
-The previous paragraphs present the biological mechanisms behind the axonal delay, regulated by myelination.
-However, the dendritic delay \add[Amélie]{as well as the axonal delay (see }~\citep{seidl_mechanisms_2010}) seem to play an equally important role in the precise timing within a sequence of spikes.
-Dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
-Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}.
-\citet{mel_synaptic_2017} \add[Amélie]{highlights the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
-Dendritic spiking involvement in synaptic potentiation following active backpropagation into dendrites was experimentally uncovered using calcium imaging to highlight dendritic calcium entry allowing for long-tem potentiation}~\citep{golding_dendritic_2002}.
-~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
-The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree and relies on the dendritic calcium influx moderation by NMDA receptors.
-
-All in all, learning spike motifs requires significantly complex pathways and biological mechanisms, whether in the dendrites or the axon of the neuron.
-More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop ANNs, whose learning rules would be more neuromorphic.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Modeling precise spiking motifs in theoretical \& computational neuroscience}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Now that we have reviewed biological foundations for the role of delays in neural computations, we review, in the following section, theoretical models which directly take advantage of using precise spiking motifs.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Spiking Neural Networks (SNNs)~\citep{maass_networks_1997} are natural candidates to use these precise temporal patterns in the brain.
-The approach which is currently most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021} One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.
-In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error's gradient.
-\change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the \remove[Antoine]{N-}MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
-\change[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
-However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{So far, this approach does not outperform classical architectures both in term of training efficiency and performances}~\citep{davies_loihi_2018}\add[Antoine]{.
-However, they remain the best candidates to reproduce biological neural systems and their capacities both in terms of accuracy, speed and energy consumption.} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
-To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) \change[Antoine]{to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized}{to a detector of precise spiking motifs}~\citep{paugam-moisy_computing_2012}.
-In particular, we will explore different existing architectures which are able to overcome the diversity of input presynaptic patterns and learn to detect stable spiking motifs, that is, volleys of spikes which are stable up to a certain onset time (see Figure~\ref{fig:THC}).
-These models will be compared in light of neuroscientific and computational perspectives.
-%We review in this section theoretical and computational foundations of spiking motif detection.
-%In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Izhikevich's polychronization model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% polychronization
-As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time encoding machine by}~\citet{lazar_time_2004} and \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
-\add[Antoine]{In this section we focus on the polychronization model} based on a random recurrent model of spiking neurons including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
-\change[Antoine]{It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons}{It was shown that spike timing (STDP) has an impact on the development of synaptic efficacy for many kinds of neurons}~\citep{caporale_spike_2008}.
-Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma.
-It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that delays are then ``frozen'' for the rest of the simulation.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs).
-One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
-This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevich}).
-According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group.
-%----------------------------%
-%
-\begin{figure}%[t!]
-  \centering
-  \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
-    \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
-    %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
-     }
-  \label{fig:izhikevich}
-\end{figure}
-% 
-%----------------------------%
-
-% follow-ups
-Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper.
-Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs.
-A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}.
-The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron, which does not incorporate a spike timing framework.
-The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a supervised STDP rule.
-The limits of this model are that its output is only binary and that its storage capacities are limited.
-An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}.
-Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete such as to allow for the reproduction of all results presented in that manuscript.
-This more recent work details how this code could be slightly corrected.
-It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
-Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}.
-In that work, based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation.
-They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system.
-Stimulation of the trained network produces the activation of a PG, i.e.
-the propagation of firing activity through multiple layers due to convergent patterns of firing.
-While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
-
-Strikingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:THC}).
-In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}.
-The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the modulation of spike latency emission~\citep{susi_nmnsd-spiking_2021}.
-Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
-This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
-Recently, by using a logistic regression model coupled with a temporal convolution, a model with heterogeneous delays was implemented to test the detection of the spiking motifs embedded in an event stream~\citep{grimaldi_learning_2022}.
-This allowed to detect a high number of superposed polychronous motifs in synthetic data, illustrating the computational benefit of such representations compared to that with a unique delay (see Figure~\ref{fig:THC}).
-As such, these models use the neural dynamics to handle input signals with different delays but do not explicitly take full advantage of the representation capacity offered by heterogeneous delays.
-%
-%----------------------------%
-%
-\begin{figure}%[t!]
-  \centering
-  \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
-  \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
-  \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
-  \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-    \caption{\textbf{Detecting event-based motifs using spiking neurons with heterogeneous delays.} 
-    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
-    {\bf (b)}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay.
-    {\bf (c)}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}.
-% such that the addresses are the synapses of these neurons (which we therefore call here synapses).
-    %Note that each afferent may be connected with multiple weights at different delays.
-The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
-%These PG motifs were efficiently learned by a logistic regression.
-    %{\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
-    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
-     }
-  \label{fig:THC} % TODO: citer biol cyb fig:THC
-\end{figure}
-% 
-%----------------------------%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Learning synaptic delays}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Training delays ~\citep{state_training_2019}
-First, the original model by Izhikevich use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
-For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
-Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
-They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
-In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
-More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
-Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
-
-At the implementation level, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}.
-The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns.
-Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task, but were prone to convergence issues.
-Another model of synaptic delay-weight plasticity integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes~\citep{zhang_supervised_2020}.
-This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
-In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron.
+
+\begin{itemize}
+ \item
+  ``However, SPOTDis has two principal weaknesses that we address here: (1) Its computational complexity, for comparing two time epochs, is O(N2), where N is the number of neurons.
+This becomes a major problem for computing an M dissimilarity matrix (for M time epochs) for thousands of neurons.
+(2) It relies exclusively on pairwise spike-timing relationships (i.e 2nd order correlations), because it does not solve the optimal transport problem for the entire spike pattern, but only for neuron pairs separately.
+Hence, it may not be sensitive to higher-order correlations in spiking
+  patterns.
+  Here, we develop a novel dissimilarity measure for multi-neuron spiking patterns called SpikeShip, which has linear computational complexity of O(N).
+We achieve this by (1) computing the minimum transport cost of spikes for each spike train separately, and (2) discounting a global translation term in the transport flow across neurons.''
+  https://doi.org/10.1101/2020.06.03.131573;
+\end{itemize}
+
+\subsubsection{Paper by Russo et al 2017~\citep{ref-RrgzWOSk}}\label{paper-by-russo-et-al-2017-russo2017}
+\begin{itemize}
+ \item
+  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{ref-RrgzWOSk}
+\item
+  sliding window as in~\citep{ref-1XpTamcl} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{ref-bc9czTxi}
+\item
+  based on a ``non-stationarity-corrected parametric test statistic for assessing the independence of pairs'' and ``an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies''
+\end{itemize}
+
+\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{ref-1300cS9RL}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
+\begin{itemize}
+ \item
+  Stochastic sampling links perceptual uncertainty to neural response variability
+\item
+  Model accounts for independent changes in strength and variability of responses
+\item
+  Model predicts relationship between noise, signal, and spontaneous correlations
+\item
+  Stimulus statistics dependence of response statistics is explained
+\end{itemize}
+
+To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{ref-1365b3Wwl} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time.
+They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
+
+\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}\label{dynamics-of-delay-coupled-excitable-neural-systems}
+uses a FPGA
+~\citep{ref-rYLmAESf}
+
+\begin{verbatim}
+February 2009International Journal of Bifurcation and Chaos 19(02):745-753
+\end{verbatim}
+
+V.
+Thanasoulis, B.
+Vogginger, J.
+Partzsch and C.
+Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
+~\citep{ref-1DadetlqZ}
+
+\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
+~\citep{ref-RrgzWOSk} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
+It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}.
+Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
+
+\section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}
+\subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
+spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{ref-19NyVq80B} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays.
+This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach.
+This will decouple the active synapses (similarly to a logistic regression) from the values of possible synaptic delays.
+
+\begin{itemize}
+ \item
+ ~\citep{ref-11p3NXmkg} : coherence detection
+\item
+ ~\citep{ref-j5xS3owo} : STDP
+\item
+  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
+PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+\end{itemize}
+
+Our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
+Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision.
+This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{ref-s09KZ8Kh}
+
+\textbf{TODO amelie} section sur l'apprentissage des délais dans la biologie
+
+~\citep{ref-UeTLXQCp}
+
+Bio-plausible Unsupervised Delay Learning for Extracting Temporal Features in Spiking Neural Networks
+Alireza Nadafian, Mohammad Ganjtabesh
+
+\begin{verbatim}
+The plasticity of the conduction delay between neurons plays a fundamental role in learning.
+However, the exact underlying mechanisms in the brain for this modulation is still an open problem.
+Understanding the precise adjustment of synaptic delays could help us in developing effective brain-inspired computational models in providing aligned insights with the experimental evidence.
+In this paper, we propose an unsupervised biologically plausible learning rule for adjusting the synaptic delays in spiking neural networks.
+Then, we provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns.
+Furthermore, the experimental results of applying an STDP-based spiking neural network equipped with our proposed delay learning rule on Random Dot Kinematogram indicate the efficacy of the proposed delay learning rule in extracting temporal features.
+\end{verbatim}
+
+Integrating synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes, that is referred to as synaptic delay-weight plasticity~\citep{ref-ga3ZTFC7}.
+Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
+
+In a recent paper~\citep{ref-B493Fl4K}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron.
 In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
-In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.
-
-%In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in birdsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+
+In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{ref-pPRQwZb}.
+This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.
+
+\subsection{Learning sequences}\label{learning-sequences}
+In~\citep{ref-WFwUKZb5}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje,ref-Dry1ANMD,ref-SAIh1pga} and this model shows how one could vary HVC activity using something similar to dropout in ML.
 Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
-In~\citep{agus_rapid_2010}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
-
-% Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
-% Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021
-
-% Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
-% Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020
-
-% Unsupervised Learning of Persistent and Sequential Activity.
-% Pereira U, Brunel N.Front Comput Neurosci. 2020
-
-% From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
-% Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019
-
-% Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
-% Pang R, Fairhall AL.Elife. 2019 May~\citep{pang_fast_2019}
-
-% Emergence of spontaneous assembly activity in developing neural networks without afferent input.
-% Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018
-
-% Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
-% Ocker GK, Doiron B.Cereb Cortex. 2019.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Real-world applications}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%----------------------------%
-% TODO
-\begin{figure}%[t!]
-  \centering
-  \includegraphics[width=0.980\linewidth]{figures/2022-09-27_MotionDetection_kernel.png}
-  %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
-    \caption{\textbf{Detecting visual motion in an event stream with heterogeneous delays.} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
-The models parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
-When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push-pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}.
-Global weight is globally decreasing from the lowest delay (right) to less recent information (left).
-     }
-  \label{fig:icip}
-\end{figure}
-% 
-%----------------------------%
-
-A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios.
-Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility.
-For instance, in a recent work,~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
-The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern.
-The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity.
+
+\begin{itemize}
+\item
+  in~\citep{ref-Jzr1gwsr}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
+\item
+  In Bellec~\citep{ref-aj6TgorR}, authors fit summary statistics of neural data with a differentiable spiking network simulator.
+
+  \begin{itemize}
+     \item
+    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{ref-14e3BCBii} with recurrent dynamics)
+  \item
+    sample and measure method to include latent / hidden neurons
+  \item
+    comes with code https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure
+  \item
+    V1-dataset : The dataset we used was collected by Smith and Kohn {[}49{]} and is publicly available at:
+    http://crcns.org/data-sets/vc/pvc-11 - it is in a sense supervised with the input being the movie and the output the spikes recorded.
+  \end{itemize}
+\end{itemize}
+
+A recent work~\citep{ref-10EU8MfxC} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream.
+The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern.
+The second stage takes the output of the spatiotemporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatiotemporal activity.
 For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
-Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.
-
-There are more applications to image processing using spiking neural networks.
-For instance, a set of models are based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections.
-Using this core computational unit, and extending it to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation.
-This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks.
-In essence, two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
-This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
-This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash-lag illusion~\citep{khoei_flash-lag_2017}.
-However, these latter models used parametric rules for defining the weights.
-Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore some exemplar results from neuromorphic engineering to get better insights on that aspect.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Applications of precise spiking motifs in neuromorphic engineering}
-\label{sec:neuromorphic}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Artificial intelligence has made huge advances in the past decades and deep learning algorithms, nowadays, outperform humans at complex tasks like natural image recognition or abstract strategy board games.
-Yet, machine learning algorithms suffer from adversarial attacks or a lack of generalization \change[Antoine]{capacity but}{capacity.
-But} their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
-Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019, javanshir_advancements_2022}.
-Because the reduced energy consumption of biological networks can be explained \add{in part} by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to \change{make}{perform} energy efficient computations.
-Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips.
-It can be materialized by the capacity of the physical connections between the processing units to store information~\citep{markovic_physics_2020}.
-An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted.
-Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2003, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
-
-This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by artificially reproducing such mechanisms, it is interesting to study neural circuitry.
-Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
-In this section, we make a description of the different neuromorphic hardware that have been developed and see how they can be used to deal with precise temporal motifs.
-%
-\subsection{The emergence of novel computational architectures}
-%
-To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_analog_1989}.
-It was implemented with transistors operating in the sub-threshold domain and responded to asynchronous events, but was not capable of discriminating two different spiking sequences with the same firing rate.
-Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, as it can be observed in biological neurons.
-Then, electronics circuits became more and more bio-realistic and, two decades later, ~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons.
-The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector''.
-Today, many devices are good candidates for implementing event-based algorithms and use the \change[Antoine]{AER}{address event representation (AER)}.
-They can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
-For a more complete review the reader can refer to~\citep{schuman_survey_2017}, here we site the most popular ones.
-SpiNNaker~\citep{furber_overview_2013, furber_spinnaker_2020}, Loihi~\citep{davies_loihi_2018} and TrueNorth~\citep{merolla_million_2014} chips are widely used examples of digital hardware implementations.
-Compared to TrueNorth which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning.
-This flexibility in the implementation comes at the cost of an increased energy consumption.
-Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}.
-They are mostly used by computational neuroscientists to model brain activity with different levels of abstraction.
-BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}.
-It is a wafer-scale neuromorphic hardware with analog components.
-Analog arrays (i.e.
-field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices.
-The pulsed current source synapse and the DPI are examples of such devices, we can also mention the field programmable neural array~\citep{farquhar_field_2006} and the NeuroFPAA~\citep{cheng_fpaa_2009} specifically designed for neuromorphic systems.
-Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.
-%
-%----------------------------%
-\begin{figure}
-\centering
-\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
-\caption{\textbf{A miniature, event-based ATIS sensor.} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
-These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
-Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
-\end{figure}
-%----------------------------%
-%
-Neuromorphic sensors are also developed with the idea to capture external stimuli more efficiently and closer to biological systems.
-A widely used example is the Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
-These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene.
-We also report other event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but will focus on DVS for the next subsection about the use of dynamics embedded in event-based signals.
-
-%Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by reproducing such mechanisms, it is interesting to study neural circuitry.
-Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
-Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
-In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
-There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
-Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.
-%
-\subsection{On the importance of spatio-temporal information in silicon retinas}
-%
-With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and processing of visual information.
-From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
-The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision tasks.
-
-In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
-An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
-It gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
-Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
-Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022} and show that this method could be efficiently applied to state-of-the-art benchmarks.
-
-More generally, three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}.
-With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window (see Figure~\ref{fig:icip}).
-This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
-Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
-This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of the precise temporal resolution of neuromorphic retina-like cameras.
-By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
-With this type of signal, the use of spatio-temporal motifs embedded in the event streams are essential to solve high level visual tasks.
-%
-\subsection{Computations with delays in neuromorphic hardware}
-%
-For the rest of this section, we report examples of implementations of event-based algorithms using precise spatio-temporal motifs on neuromorphic hardware. 
-\citep{hussain_deltron_2012} implemented a delay-learning algorithm on an analog chip.
-Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}.
-In addition to delay learning, a group at the University of West Sydney developed a neuromorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}.
-They show that STDP and Spike Timing Dependent Delay Plasticity rules could be implemented in both a digital and an analog chip.
-From the network parameters and the physical limitations to store it on-chip, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
-The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2014}.
-%This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-\citet{pfeil_neuromorphic_2013} implemented STDP on a mixed-analog-digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
-Coherence detection on input spikes coming from two sources was obtained with a 50 $ns$ precision.
-They claim that this unsupervised learning denoises the input and compensates for variations between neural components.
-The variability of response of the analog components can be compensated by population coding for responses robust to noise and this phenomenon is also observed in biological neural networks~\citep{boerlin_spike-based_2011}.
-A recent work makes implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}.
-They highlight the fact that using delays can be expensive \add[Laurent]{notably in memory bandwidth} because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay.
-Note that this problem is also due to the algorithm used in this paper and that the analog chips must not suffer from this type of limitation.
-
-Online on-chip learning and computations with delays are still emerging in neuromorphic engineering.
-The technical challenges linked to the development of this type of implementation and the growing interest in delay learning make advances in this field interesting for the future of computations with precise spatio-temporal motifs.
-While improvements are still to be made, neuromorphic chips seem to be a good candidate to efficiently make use of these particular features.
-%
-%they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is well-suited for areas such as pattern recognition and natural language processing.
-In this paper, we formally describe the STPU, implement the STPU on a field programmable gate array, and show measured performance data.\citep{hill_spike-timing_2017} not accessible
-%
-%review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)
-%
-%We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method.
-Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware~\citep{liang_neuromorphic_2022}
-%
-%Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160
-%
-% solutions usuellement utilisées - surrogate gradient
-%for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}\label{sec:discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Summary}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
-In this review paper, we have presented recent evidence for the role of precise spiking motifs in neuroscience.
-In particular, we have shown that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels and that they may have numerous applications in neuromorphic engineering.
-In particular, we have shown that:
-\begin{itemize}
-  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlights the importance of precise spike times,
-  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, travelling waves on arbitrary spiking motifs, \add[Laurent]{and notably that a an encoding based on precise spiking motifs may provide huge computational benefits},
-  \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
-  \item  using precise spiking motifs could ultimately be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
-\end{itemize}
- Overall, our reviewing effort has shown that a growing community is focusing on that aspect.
-This community is based on solid and validated evidence which is getting novel grounds thanks to current technical advances.
-Moreover, we have also shown that this community is highly diverse, operating on \emph{in vivo}, \emph{in computo} and \emph{in silico} systems.
-As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
-
-\subsection{Limits}
-Additionally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
-
-First, many models are based on a discretization of time.
-This assumption is important to allow for a useful representation of neuronal information in order to be processed in computers.
-This treatment amounts to transforming spike trains into a matrix form for processing in classical machine learning algorithms.
-This assumption therefore implies an ineffective use of the memory, as this representation transforms the sparse representation of a spike sequence into dense matrices.
-In addition, this representation can induce errors due to the discretization and the scale of temporal sampling.
-Finally, this representation encourages the use of classical methods, which are not adapted to disruptive applications such as event-driven representations.
-
-Moreover, the learning of patterns is often done in a supervised way.
-Indeed, the problem of detecting polychronous groups implies to infer both the address and the precise time of occurrence of these motifs.
-Most of the models we have presented are based on the assumption that at least one of its variables is known: either the pattern, its identity, or its time of occurrence.
-This constraint is to be put in parallel with the way a biological nervous system works in which learning is done autonomously, i.e.
-without supervision.
-However, we can note that some models can perform such learning, but only in the case of data for which the motifs are easily separable.
-More generally, to reproduce the efficiency of biological systems, one should account for the different temporal scales of adaptation, from seconds to years.
-For instance, the scaffolding of neural assemblies seems to follow critical periods during development~\citep{dard_rapid_2022}.
-
-A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output.
-However, we have seen that the neurobiological system uses processing loops within hierarchical graphs.
-In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}.
-The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
-Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts)~\citep{coull_distinction_2022}.
-%
-\subsection{Perspectives}
-%
-The limits that we have presented can be treated individually in each model, as evidenced by individual efforts which try to overcome them.
-However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.
-
-First, as we already noted, the effort is still largely scattered.
-This is in part due to the fact that interaction between neurobiology, theoretical and computational neuroscience and neuromorphic engineering are still scarce as of today.
-It was largely demonstrated that close, bidirectional interactions are essential to foster breakthroughs.
-For instance, the design of model-driven protocols has proven to be essential in modern neuroscience.
-Also, if neural networks were essential in shaping modern-day machine learning, e-g.
-computer vision using deep-learning, spiking neural networks should prove essential in future emerging technologies.
-
-In that perspective, it is essential that such models are tested on ecologically-relevant, real-world scenarios.
-Indeed, classic convolutional neural networks emerged as optimal solutions for instance to classify static images into categories, yet they are not well adapted for processing dynamic, multimodal sensory flows.
-The emerging necessity to be able to process more complex flows, such as the multiple flows of information in a car designed for autonomous driving, necessitates to modify such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs.
-Notably, the protocols used as well in neurobiology, theory or engineering should take into account this novel levels of complexity.
-
-Ultimately, the community should encourage the adoption of open, reproducible science.
-Indeed, the different models that we have displayed often come with the tools necessary to reproduce the results obtained.
-This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}.
-This aspect is essential to foster the emergence of interdisciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering.
-Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011} i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms.
-As such, this review aims at paving the way to openly share the variety of resources and to offer a unified view on the role of precise spiking motifs in neuroscience.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt} 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
-A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \authorcontributions{Conceptualization, LP; writing---original draft preparation, all authors; writing---review and editing, all authors; supervision, LP; funding acquisition, LP and JM.
-All authors have read and agreed to the published version of the manuscript.}
-
-\funding{
-This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
-% Agence Nationale de la Recherche %HL and
-LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}''
-% LP received funding 
-LP received support from the french government under the France 2030 investment plan, as part of the Initiative d’Excellence d’Aix-Marseille Université – A*MIDEX number AMX-21-RID-025 ``\href{https://laurentperrinet.github.io/grant/polychronies/}{Polychronies}''.}
-
-%\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
-You might choose to exclude this statement if the study did not require ethical approval.
-Please note that the Editorial Office might ask you for further information.
-Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
-OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
-OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
-
-%\informedconsent{Any research article describing a study involving humans should contain this statement.
-Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
-You might also choose to exclude this statement if the study did not involve humans.
-
-%Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
-Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
-
-\dataavailability{This works is made reproducible using the following tools.
-First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/XXX}{GitHub}.
-%~\citep{Grimaldi2023review}.
-The paper is available as an \href{https://arxiv.org/abs/YYYY.XXX}{arXiv preprint} with links to previous versions.
-Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{Zotero group} which was used to regroup relevant literature on the subject.}
-
-\acknowledgments{We would like to thank Hugo Ladret and Frdéric Chavane for valuable feedback while writing this review.
-Many thanks to Simon Thorpe for the useful discussions and references on the latency of the visual system in human and non-human primates.}
-
-\conflictsofinterest{The authors declare no conflict of interest.
-The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% %% Optional
-% \sampleavailability{Samples of the compounds ... are available from the authors.}
-
-% %% Only for journal Encyclopedia
-% %\entrylink{The Link to this entry published on the encyclopedia platform.}
-
-% \abbreviations{Abbreviations}{
-% The following abbreviations are used in this manuscript:\\
-
-% \noindent 
-% \begin{tabular}{@{}ll}
-% MDPI & Multidisciplinary Digital Publishing Institute\\
-% DOAJ & Directory of open access journals\\
-% TLA & Three letter acronym\\
-% LD & Linear dichroism
-% \end{tabular}
-% }
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% %% Optional
-% \appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
-If the appendix sections contain a heading then change the argument to "yes".
-% \appendixstart
-% \appendix
-% \section[\appendixname~\thesection]{}
-% \subsection[\appendixname~\thesubsection]{}
-% The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
-Mathematical proofs of results not central to the paper can be added as an appendix.
-
-% \begin{table}[H] 
-% \caption{This is a table caption.\label{tab5}}
-% \newcolumntype{C}{>{\centering\arraybackslash}X}
-% \begin{tabularx}{\textwidth}{CCC}
-% \toprule
-% \textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-% \midrule
-% Entry 1		& Data			& Data\\
-% Entry 2		& Data			& Data\\
-% \bottomrule
-% \end{tabularx}
-% \end{table}
-
-% \section[\appendixname~\thesection]{}
-% All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{adjustwidth}{-\extralength}{0cm}
-%\printendnotes[custom] % Un-comment to print a list of endnotes
-
-%%%%%%%%% REFERENCES
-\reftitle{References}
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
-%\bibliographystyle{Definitions/chicago2}
-\bibliography{references}
-
-%\begin{thebibliography}{00}
-%\printbibliography[heading=none]
-%\end{thebibliography}
-
-% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author:~\citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a},~\citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages:~\citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328;~\citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-\end{adjustwidth}
-\end{document}
+Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters.
+Showing the potential of such representation
+
+\subsection{TODO: more bib to read}\label{todo-more-bib-to-read}
+Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
+Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021
+
+Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
+Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020
+
+Unsupervised Learning of Persistent and Sequential Activity.
+Pereira U, Brunel N.Front Comput Neurosci. 2020
+
+From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
+Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019
+
+Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
+Pang R, Fairhall AL.Elife. 2019 May~\citep{ref-12zyUFdO4}
+
+Emergence of spontaneous assembly activity in developing neural networks without afferent input.
+Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018
+
+Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
+Ocker GK, Doiron B.Cereb Cortex. 2019.
+
+\textbf{TODO}
+
+\subsection{learning pattern detection on natural images / event-based cameras}\label{learning-pattern-detection-on-natural-images-event-based-cameras}
+\subsubsection{sparse coding on spatio-temporal data}\label{sparse-coding-on-spatio-temporal-data}
+\subsubsection{HOTS}\label{hots}
+\subsubsection{Grimaldi CBMI / PAMI}\label{grimaldi-cbmi-pami}
+\section{Discussion}\label{discussion}
+\textbf{TODO}
+
+\subsection{dynamical models}\label{dynamical-models}
+Dumas and colleagues~\citep{ref-WxQdcMB2} : three levels / fourth paradigm~\citep{ref-JrZPoxk9} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
+
+\subsection{our model}\label{our-model}
+Here, we develop a model for the efficient detection of such PGs based on the inversion of a probabilistic model defining the generation of the raster plot as a combination of such groups.
+We show that such an inference can be achieved by a neural-like computation that could itself be used as a spiking neuron, as can be implemented in a neuromorphic chip for instance.
+A first result is to show the efficiency of such a scheme in detecting different PGs occurring at specific times in synthetic data.
+The representational capacity of the PGs is particularly interesting compared to traditional models of neuronal encoding using spiking frequency.
+Our second result is to propose a novel learning method for learning PGs in raster plots in a self-supervised manner.
+Finally we demonstrate the use of this algorithm to the output of an event-based camera and how this may separate independent components from the stream of events.
+This end-to-end event-based computational brick could help improve the performance of current Spiking Neural Network solution currently used in neuromorphic chips.
+
+Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}.
+
+and oscillations? Journal Article • DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?
+
+\section{development}\label{development}
+scaffolding of neural assemblies / existence of critical periods :~\citep{ref-1HSyNOl48}
 
@@
 As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper~\citep{ref-journal}.
-Now citing a book reference~\citep{ref-book1,ref-book2} or other reference types~\citep{ref-unpublish,ref-communication,ref-proceeding}.
-Please use the command~\citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
+Citing a journal paper \cite{ref-journal}.
+Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}.
+Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``Conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work~reported.}
+
+\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+
+\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
+You might choose to exclude this statement if the study did not require ethical approval.
+Please note that the Editorial Office might ask you for further information.
+Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
+OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
+OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
+
+\informedconsent{Any research article describing a study involving humans should contain this statement.
+Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
+You might also choose to exclude this statement if the study did not involve humans.
+
+Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
+Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
+
+\dataavailability{
+%This works is made reproducible using the following tools.
+First the code reproducing all figures is available on \href{https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Readme.md}{GitHub}~\citep{Jeremie2022a}.
+
+%\noindent The paper is available a an \href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
+Find the associated \noindent \href{https://www.zotero.org/groups/4562620/polychronies}{zotero group} used to regroup relevant literature on the subject.
+
+}
+
+\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results''.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\sampleavailability{Samples of the compounds ... are available from the authors.}
+
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+\abbreviations{Abbreviations}{
+The following abbreviations are used in this manuscript:\\
+
+\noindent 
+\begin{tabular}{@{}ll}
+MDPI & Multidisciplinary Digital Publishing Institute\\
+DOAJ & Directory of open access journals\\
+TLA & Three letter acronym\\
+LD & Linear dichroism
+\end{tabular}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
+\appendix
+\section[\appendixname~\thesection]{}
+\subsection[\appendixname~\thesubsection]{}
+The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+\begin{table}[H] 
+\caption{This is a table caption.\label{tab5}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCC}
+\toprule
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+\midrule
+Entry 1		& Data			& Data\\
+Entry 2		& Data			& Data\\
+\bottomrule
+\end{tabularx}
+\end{table}
+
+\section[\appendixname~\thesection]{}
+All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{adjustwidth}{-\extralength}{0cm}
+%\printendnotes[custom] % Un-comment to print a list of endnotes
+
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+%\bibliography{your_external_BibTeX_file}
+
+%=====================================
+% References, variant B: internal bibliography
+%=====================================
+\begin{thebibliography}{999}
+% Reference 1
+\bibitem[Author1(year)]{ref-journal}
+Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+% Reference 2
+\bibitem[Author2(year)]{ref-book1}
+Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+% Reference 3
+\bibitem[Author3(year)]{ref-book2}
+Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
+% Reference 4
+\bibitem[Author4(year)]{ref-unpublish}
+Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
+% Reference 5
+\bibitem[Author5(year)]{ref-communication}
+Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
+% Reference 6
+\bibitem[Author6(year)]{ref-proceeding}
+Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
+Title of presentation.
+In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
+% Reference 7
+\bibitem[Author7(year)]{ref-thesis}
+Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
+% Reference 8
+\bibitem[Author8(year)]{ref-url}
+Title of Site. Available online: URL (accessed on Day Month Year).
+\end{thebibliography}
+
+% If authors have biography, please use the format below
+%\section*{Short Biography of Authors}
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
+%{\textbf{Firstname Lastname} Biography of first author}
+%
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
+%{\textbf{Firstname Lastname} Biography of second author}
+
+% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{adjustwidth}
+\end{document}

### 2022-12-25 manuscript_converted.tex
2022-10-04 conversion from manubot

--- 2022-12-25 manuscript.tex
@@
 The performances decrease progressively as the frequency of presentation of the images increases.
-However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image~\citep{ref-EqCCVwl2}.
+However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image~\citep{keysers_speed_2001}.
 
@@
 Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes.
-Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{ref-zZUVkTfZ} but also the dependance across neurons~\citep{ref-13T2Gne9z}.
-
-Numerous studies demonstrated the importance of precise timing in neural population activity~\citep{ref-BNK6T5Uk}, efficient encoding thanks to the use of spike latencies~\citep{ref-1DpSM1kpQ,ref-IjepoonT} or precise timing in the auditory system~\citep{ref-mxPMgOmW,ref-1ECJhg2GM}.
-All these findings, and more~\citep{ref-WZsEyOlK}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
-Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{ref-XsHCA4IZ}.
+Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing~\citep{perkel_neuronal_1967} but also the dependance across neurons~\citep{ref-13T2Gne9z}.
+
+Numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{ref-1DpSM1kpQ,gollisch_rapid_2008} or precise timing in the auditory system~\citep{ref-mxPMgOmW,carr_circuit_1990}.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train.
+Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons~\citep{mainen_reliability_1995}.
 
@@
 
-At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{ref-IjepoonT}.
+At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina~\citep{gollisch_rapid_2008}.
 Notably one can appreciate in figure 1 that the response of ganglion cells to visual gratings that are flashed onto the retina.
@@
 The conclusion of the authors is that the discharge latency of the neurons allows to encode spatial characteristics of the image.
-First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{ref-1H1Omh73g}.
-Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{ref-UlbjMU0W}.
+First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account~\citep{chase_first-spike_2007}.
+Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
 
@@
 This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation~\citep{ref-15cR83gqd}.
-Note also that timing is not entirely sensorial or internal but in~\citep{ref-1E8gkynCm}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
-Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{ref-QK8RfvTZ}.
+Note also that timing is not entirely sensorial or internal but in~\citep{safaie_turning_2020}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
 
@@
 For instance, the firing rate is highly irregular~\citep{ref-sYHyiGsP}, which makes it inconsistent with temporal integration of random EPSPs.
-It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{ref-XsHCA4IZ} (see Figure {[}\ref{fig:mainen}{]}).
+It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{mainen_reliability_1995} (see Figure {[}\ref{fig:mainen}{]}).
 When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible.
@@
 The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion.
-Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{ref-R2E2wRFg}.
+Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
 
@@
 Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity.
-However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{ref-18ZU9zP9l}.
-
-Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{ref-wVZpQvSk}.
+However, this pathway is sufficient to drive the cortex as this input is more often synchronously active~\citep{bruno_cortex_2006}.
+
+Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}.
 Interestingly, authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events.
 Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent.
-These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{ref-1HJjmmSpm} and showed some tuning to to movement direction and reaction time~\citep{ref-FZ9X31Qa}.
-It is important to note that synchronous events tended to lock to LFP beta waves~\citep{ref-Ejy1sAyh}, and was extended to larger assemblies~\citep{ref-Glu4NlLC} using statistical methods (see~\citep{ref-sec:detection}{\textbf{sec:detection?}}{]}).
-
-In~\citep{ref-3GlNNxH5}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
+These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{ref-1HJjmmSpm} and showed some tuning to to movement direction and reaction time~\citep{grammont_spike_2003}.
+It is important to note that synchronous events tended to lock to LFP beta waves~\citep{denker_lfp_2018}, and was extended to larger assemblies~\citep{ref-Glu4NlLC} using statistical methods (see~\citep{ref-sec:detection}{\textbf{sec:detection?}}{]}).
+
+In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks.
 It relies on the simple remark that a series of repeated coincidences is in itself an invariant.
@@
 Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{ref-Piao2Wzf}.
+This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}.
 This show precise spatio-temporal differences in this recurrently connected assembly.
 The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations.
-A similar dataset used in~\citep{ref-2AsQOOsE} is available from M.
+A similar dataset used in~\citep{schneidman_weak_2006} is available from M.
 Berry's lab~\citep{ref-1FU0Y7Erc}.
@@
 Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function.
-In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{ref-BNK6T5Uk}.
+In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
 Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise.
@@
 
-\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
-It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{ref-yLVM9IXi}.
+\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{riehle_spike_1997}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{grammont_precise_1999}.
 It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands.
@@
 
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{ref-KS4oxnst,ref-ga3ZTFC7}.
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{ref-KS4oxnst,zhang_supervised_2020}.
 Notably, Izhikevich~\citep{ref-SM9G0xBK} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays.
@@
 Currently, there is a consensus for rate-based coding models in computational and biological neuroscience.
-Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ref-Mf6h1KNY}, packets~\citep{ref-2GovoaYw} or hippocampal sequences~\citep{ref-SVfasyCp,ref-PUNGVou} and on the role of such precise spike timing in downstream information transfer and coding~\citep{ref-19fAM6hcv,ref-13dQ0R4UM,ref-MfmcMsqA}.
-This is for instance relevant in sensory pathways in vision~\citep{ref-1EHZgcKyL}, audition~\citep{ref-FlKJw2VL}, olfaction~\citep{ref-AhSRXBJ2}{]} or touch~\citep{ref-gDinei4a}.
+Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains~\citep{ref-Mf6h1KNY}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{ref-SVfasyCp,ref-PUNGVou} and on the role of such precise spike timing in downstream information transfer and coding~\citep{ref-19fAM6hcv,ref-13dQ0R4UM,ref-MfmcMsqA}.
+This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{wehr_odour_1996}{]} or touch~\citep{johansson_first_2004}.
 In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
@@
 Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs.
-They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from~\citep{ref-qI2HUEyy} below).
+They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from~\citep{haimerl_internal_2019} below).
 As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
@@
 
-In~\citep{ref-4D8Jnhkz}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
+In~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
 They develop a HMM model for quantifying the transitions.
@@
 
-there are temporal delays in the nervous system, both at the neural~\citep{ref-f7gkBktf} and behavioral~\citep{ref-1BSbXPPDF} levels.
+there are temporal delays in the nervous system, both at the neural~\citep{perrinet_active_2014} and behavioral~\citep{khoei_flash-lag_2017} levels.
 Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
@@
 A popular model for the detection of latency patterns is the tempotron~\citep{ref-1GazUBEdT}.
-This model is in particular reviewed in~\citep{ref-QCyjy8wE}.
+This model is in particular reviewed in~\citep{gutig_spike_2014}.
 The Tempotron is a supervised synaptic learning algorithm which classifies a distractor from a target motif.
@@
 Some other models of latency motifs detection using STDP learning rules.
-For instance,~\citep{ref-Dd3kSDRW} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{ref-vePjgmRh} develops a form of spike-based, competitive learning applied for unsupervised learning.
+For instance,~\citep{ref-Dd3kSDRW} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
 In~\citep{ref-RrgzWOSk}, authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which was applied to simultaneous electrophysiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
@@
 
-Memory traces in dynamical systems~\citep{ref-o7rANZux} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
+Memory traces in dynamical systems~\citep{ganguli_memory_2008} : ``To perform nontrivial, real-time computations on a sensory input stream, biological systems must retain a short-term memory trace of their recent inputs.
 It has been proposed that generic high-dimensional dynamical systems could retain a memory trace for past inputs in their current state.
@@
 In generic linear non linear lnl models, the output is assumed to be poisson.
-As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{ref-MgINa6bU}.
+As such a simple decoding strategy is to asscume it is to b inferned for a given tuning curves (Jayazeri) or simply by a simple regression~\citep{berens_fast_2012}.
 This latter model assumes a Bernoulli model for the generation of spikes such that the decoding amounts to a single-layer logistic regression.
@@
 Unitary event analysis is performed by a statistical model of coincidence detection~\citep{ref-13iJohyPk}.
-This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{ref-wVZpQvSk} for instance).
-
-Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{ref-1ErNTONy0}.
-By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range~\citep{ref-V9qkJFZU}.
-
-This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{ref-pExXxdMb} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
+This was extensively used in detecting above chance significant synchronous patterns, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance).
+
+Statistical evaluation of synchronous spike patterns extracted by frequent item set mining - a method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity~\citep{torre_statistical_2013}.
+By the same group the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the ms precision range~\citep{quaglio_methods_2018}.
+
+This was recently extended in ``3d-SPADE: Significance evaluation of spatio-temporal patterns of various temporal extents''~\citep{stella_3d-spade_2019} in order to find reoccurring spike patterns in parallel spike train data, and to determine their statistical significance.
 The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{ref-d2m7wQpH}.
@@
 ~\citep{ref-RrgzWOSk} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence.
-It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}.
+It is based on unitary events in multiple single-neuron spiking activity~\citep{grun_unitary_2002-1} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}.
 Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
@@
 \item
-  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
-PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
+  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{gilson_stdp_2010} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs.
+PLoS Comput Biol 9(2): e1002897.~\citep{kerr_delay_2013} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 \end{itemize}
@@
 \subsection{Learning sequences}\label{learning-sequences}
-In~\citep{ref-WFwUKZb5}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje,ref-Dry1ANMD,ref-SAIh1pga} and this model shows how one could vary HVC activity using something similar to dropout in ML.
+In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje,ref-Dry1ANMD,ref-SAIh1pga} and this model shows how one could vary HVC activity using something similar to dropout in ML.
 Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).
@@
 
-Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}.
+Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{di_mauro_alfio_sne_2022} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}.
 

### 2022-12-25 update_citekeys.py

@@ -1,0 +1,286 @@
+import calendar
+import os
+import re
+import subprocess
+import time
+from datetime import datetime
+import dateutil.parser
+from pathlib import Path
+
+import bibtexparser
+from bibtexparser.bibdatabase import BibDatabase
+from bibtexparser.bparser import BibTexParser
+from bibtexparser.bwriter import BibTexWriter
+from bibtexparser.customization import convert_to_unicode
+
+from academic import utils
+from academic.editFM import EditableFM
+from academic.publication_type import PUB_TYPES, PublicationType
+import os
+import glob
+
+
+def getDateTimeFromISO8601String(s, full=False):
+    d = dateutil.parser.parse(s)
+    if not full:
+        d = d.date()
+    return d
+
+
+def import_bibtex(
+    bibtex, pub_dir="publication", featured=False, overwrite=False, normalize=False, dry_run=False, verbose=False
+):
+
+    # 1- getting all citekeys
+    keys = []
+    # 1- Load BibTeX file for parsing.
+    with open(bibtex, 'r', encoding='utf-8') as bibtex_file:
+        parser = BibTexParser(common_strings=True)
+        parser.customization = convert_to_unicode
+        bib_database = bibtexparser.load(bibtex_file, parser=parser)
+        for entry in bib_database.entries:
+            #parse_bibtex_entry(entry, pub_dir=pub_dir, featured=featured, overwrite=overwrite, normalize=normalize)
+            keys.append(entry['ID'])
+
+    # 2- making a dictionary to slugify
+    from academic.import_bibtex import slugify, month2number
+
+    dico = {}
+    for key in keys:
+        dico[slugify(key)] = key
+
+    # 4- updating metadata with bibtex
+
+    import yaml
+
+    for file_path in glob.glob(f"{pub_dir}/**/index.md"):
+        new_key = file_path.split(f'{pub_dir}/')[-1].split('/index.md')[0]
+        if not new_key in ['___template___',  'person-re-id', 'template']:
+            with open(file_path, 'r', encoding='utf-8') as source_file:
+                page = source_file.read() + '\n'
+            metadata = page.split('---')
+            # print(metadata)
+            parsed_yaml = yaml.load(metadata[1], Loader=yaml.FullLoader)
+
+            #print('new_key =', new_key, 'dico[new_key]', dico[new_key])
+            old_key = dico[new_key]
+            #print('new_key =', new_key, 'dico[new_key]', dico[new_key])
[…]

### 2022-12-25 2022-12-19_coverletter.tex
Update on Overleaf.

@@
 We are very grateful to the reviewers for their very insightful remarks and we have now responded to all their remarks.
-Note that in agreement with the editorial office, we have added Jean-Nicolas Jérémie as an author, as his contributions were substantial during the process of writing this reviews.
+
+In addition to our reports to reviewers, there is a tracked changes PDF file which highlights all changes.
+
+\textbf{ Note that in agreement with the editorial office, we have added Jean-Nicolas Jérémie as an author, as his contributions were substantial for writing this review.}
 
@@
 
-This is an invited Feature Paper, with a full waiver on the APC. %In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source
+This is an invited Feature Paper, with a full waiver on the APC. In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source
 

### 2022-12-25 build/build.sh
big merge with overleaf

@@ -1,136 +1,0 @@
-#!/usr/bin/env bash
-
-## build.sh: compile manuscript outputs from content using Manubot and Pandoc
-
-set -o errexit \
-    -o nounset \
-    -o pipefail
-
-# Set timezone used by Python for setting the manuscript's date
-export TZ=Etc/UTC
-# Default Python to read/write text files using UTF-8 encoding
-export LC_ALL=en_US.UTF-8
-
-
-# Set DOCKER_RUNNING to true if docker is running, otherwise false.
-DOCKER_RUNNING="$(docker info &> /dev/null && echo "true" || (true && echo "false"))"
-
-# Set option defaults
-CI="${CI:-false}"
-BUILD_PDF="${BUILD_PDF:-true}"
-BUILD_DOCX="${BUILD_DOCX:-false}"
-BUILD_LATEX="${BUILD_LATEX:-true}"
-SPELLCHECK="${SPELLCHECK:-false}"
-MANUBOT_USE_DOCKER="${MANUBOT_USE_DOCKER:-$DOCKER_RUNNING}"
-# Pandoc's configuration is specified via files of option defaults
-# located in the $PANDOC_DATA_DIR/defaults directory.
-PANDOC_DATA_DIR="${PANDOC_DATA_DIR:-build/pandoc}"
-
-# Generate reference information
-echo >&2 "Retrieving and processing reference metadata"
-manubot process \
-  --content-directory=content \
-  --output-directory=output \
-  --cache-directory=ci/cache \
-  --skip-citations \
-  --log-level=INFO
-
-# Make output directory
-mkdir -p output
-
-# Create HTML output
-# https://pandoc.org/MANUAL.html
-echo >&2 "Exporting HTML manuscript"
-pandoc --verbose \
-  --data-dir="$PANDOC_DATA_DIR" \
-  --defaults=common.yaml \
-  --defaults=html.yaml
-
-# Create PDF output (unless BUILD_PDF environment variable equals "false")
-# If Docker is not available, use WeasyPrint to create PDF
-if [ "${BUILD_PDF}" != "false" ] && [ "${MANUBOT_USE_DOCKER}" != "true" ]; then
-  echo >&2 "Exporting PDF manuscript using WeasyPrint"
-  if [ -L images ]; then rm images; fi  # if images is a symlink, remove it
-  ln -s content/images
-  pandoc \
-    --data-dir="$PANDOC_DATA_DIR" \
[…]

### 2022-12-27 2022-12-23_video-abstract.py
Update 2022-12-23_video-abstract.py

@@ -97,13 +97,13 @@ clip.append(sub) # another bit on the last bit
 #################################################################################
 chapters = [dict(title="Visual system", color='green',
             content=[dict(figure='figures/visual-latency_bg.jpg', duration=3, subtitle=[
-                            "The visual system is very efficient to generate a...", 
+                            "The visual system is very efficient in generating a...", 
                             "...decision from the retinal image to the different ..."]),
                     dict(figure='figures/visual-latency.jpg', duration=3, subtitle=[
                             "...stages of the visual pathways, here a reaction...", 
                             "...of finger muscles in about 300 milliseconds."]),
-                     dict(figure='../AG_figures/animated_neurons/LIF.mp4', duration=5, subtitle=[
-                            "This efficiency is thought to be allowed by spikes...", 
+                     dict(figure='/Users/laurentperrinet/metagit/ABCD/AG_figures/animated_neurons/LIF.mp4', duration=5, subtitle=[
+                            "It is thought that this efficiency is achieved by spikes...", 
                             "...that is, brief all-or-none events which are passed...", 
                             "...in the very large network which forms the brain...", 
                             "...from assemblies of neurons to others."])]), 
@@ -111,31 +111,31 @@ chapters = [dict(title="Visual system", color='green',
             content=[dict(figure='figures/izhikevich.png', duration=5, subtitle=[
                             "We review here an hypothesis in which, rather than the...", 
                             "...firing frequency of spikes, it is their precise timing...", 
-                            "...which would enable or not this message passing."]),
+                            "...which would allow or not this message passing."]),
                     dict(figure='figures/THC_1a_k.png', duration=3, subtitle=[
-                            "We present different mathematical models to extract...", 
+                            "We present different mathematical models for extracting...", 
                             "...such precise spiking motifs and..."]),
                     dict(figure='figures/THC_1a.png', duration=5, subtitle=[
-                            "...also some novel tools to extract these motifs...", 
-                            "...in arbitrary raster plots, from neurobiology...", 
+                            "...also some novel tools for extracting these motifs...", 
+                            "...in arbitrary raster plots from neurobiology...", 
                             "...or neuromorphic engineering."])]), 
            dict(title="Neurobiology", color='red',
             content=[dict(figure='figures/haimerl2019.jpg', duration=5, subtitle=[
                             "This hypothesis is reviewed with respect to our ...", 
[…]

### 2022-12-28 2022-12-28_coverletter.tex
brainsci-2069565_accepted

--- 2022-12-25 2022-12-19_coverletter.tex
@@
 %----------------------------------------------------------------------------------------
-Many thanks for handling our manuscript entitled ``Precise spiking motifs in neurobiological and neuromorphic data'' for your consideration as a review paper in \emph{Brain Sciences}.
+Many thanks for accepting our manuscript entitled ``Precise spiking motifs in neurobiological and neuromorphic data'' for your consideration as a review paper in \emph{Brain Sciences}.
 We are very grateful to the reviewers for their very insightful remarks and we have now responded to all their remarks.
 
-In addition to our reports to reviewers, there is a tracked changes PDF file which highlights all changes.
+We have accepted all changes suggested at the proofreading stage and uploaded our files on the MDPI website.
 

### 2022-12-28 brainsci-2069565.tex
brainsci-2069565_accepted

% !TeX TS-program = pdfLaTeX
% !TeX encoding = UTF-8 Unicode
% !TeX spellcheck = en-US
% !BIB TS-program = bibtex
% -*- coding: UTF-8; -*-
% vim: set fenc=utf-8
%
% REV1
%
% https://www.overleaf.com/project/633bf90497b5b325a3ff692e
% https://framateam.org/spikeai/channels/review_polychrony
% https://www.zotero.org/groups/4562620/polychronies
%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\ResponseReviewerTwo}{
We would like to thank reviewer 2 for his very detailed response which were essential in providing this new revision.

> The authors present a relatively thorough review on neural coding based on spike patterns in populations of neurons, from biological, theoretical and neuromorphic engineering perspective. I find the review mostly fine, with some omissions that, when corrected, should enhance it.

Thanks for the encouraging comment, we hope that in its present form, the result is even more useful to the community.

> My main concern is the overall reliance on spike motifs as a coding mechanism. The authors should list alternatives and compare to spike motifs, at least cursory. A minor concern is the focus on mylenation-based delay learning, ignoring the obvious length-based delay adjustments.

We have more progressively commented different alternatives for coding mechanisms in the beginning of section 2, highlighting their relative benefits.

> I have commented heavily within the manuscript, and am attaching it for the authors to use. I tried to delineate wording suggestions from more significant critiques. I am also listing the bigger concerns below.

Many thanks for providing the PDF of the manuscript with annotations, this was immensely useful to annotate and improve our revision. To respond to all these points, we have constructed a tracked changes manuscript in which we respond (by a change, removal or addition) to the reviewer's comments.

> Specific issues:
> - not referencing much of insect work, with a few exceptions that are put in mammalian context. OK to do just timing in vertebrate systems (as birds are used), but should be stated clearly, and cited work should comport to that constraint (e.g., cite rat olfactory, not locust olfactory). Should the authors wish to include invertebrate work, there is a LOT more to review.

Indeed, our focus was implicitly on vertebrates but we cited some works about insect. As mentioned by the reviewer, adding researches done on insects would increase too much the size of the article. We decide to focus on vertebrates and mention this choice in the new manuscript.

> - some references to original research are missing, while often much older ones are used (and not reviews).
> - missing important lines of work on spike timing. In particular, Aurel Lazar's TEM work. 
> - missing important lines of work on dendritic processing: Mel, Spruston.

These critical references were indeed missing and we have added them now to the manuscript. Thanks for pointing out these omissions.

> - discounting axonal length as delay mechanism, focus on myelin effects too high.

This was corrected in section 4, providing a more balanced review.

> - singular focus on synfire chain/polychronicity-like mechanisms. Alternative options, like Bayesian spiking (Deneve), probabilistic ensemble motifs (Ballard) and probabilistic population coding (Pouget and Latham), which do not rely on exact motifs of spikes, are not discussed.

Indeed, we had made that choice to study precise temporal motifs. In the new manuscript, we mention these alternative options briefly even if we keep the focus on deterministic spiking motifs for this review.

> - there are also some discrepancies - are SNN-s differentiable or not.

We have corrected these discrepancies. Thanks again for all these useful comments.

}

% cite or die
% ❯ ipython analyze_bibs.py

% In recognition of your well-respected work, we are inviting you to submit one feature contribution (i.e., a long review or research paper) to an MDPI journal free of charge. If you wish to take advantage of this opportunity, please provide your information and the name of the journal you would like to submit to: https://www.surveymonkey.com/r/XQZ5FP9 (If you cannot find the journal through the link, please feel free to reply to us about your choice).

% The submission deadline for this special offer is 31 October 2022. The journal editorial office will contact you soon once you choose the journal. Please contact the revelant journal’s Editorial Office if you require an extension.

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[brainsci, %arts,
               review,accept,pdftex,moreauthors
               %,  draft
               ]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2022}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{16 November 2022} 
\daterevised{20 December 2022 } % Only for the journal Acoustics
\dateaccepted{23 December 2022 } 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.

\newcommand{\ms}{\si{\milli\second}}%

% a simpler annotation system:
\usepackage{soulutf8}
\usepackage{color}
\newcommand{\mynote}[1]{{\sethlcolor{yellow}\hl{#1}}}
% \usepackage{trackchanges}
\usepackage[finalnew]{trackchanges}
\addeditor{Antoine}
\addeditor{Amélie}
\addeditor{Camille}
\addeditor{Jean}
\addeditor{JN}
\addeditor{Laurent}

%\newcommand{\note}[1]{}
%
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Precise Spiking Motifs in Neurobiological and Neuromorphic~Data}
%
% MDPI internal command: Title for citation in the left column
\TitleCitation{Precise Spiking Motifs in Neurobiological and Neuromorphic Data}
%
% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0002-3107-4788} % AnG % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorB}{0000-0003-3916-0514} % AmG
\newcommand{\orcidauthorC}{0000-0002-0054-2729} % CB
\newcommand{\orcidauthorD}{0000-0002-9238-6654} % JN
\newcommand{\orcidauthorE}{0000-0001-8821-5556} % JM
\newcommand{\orcidauthorF}{0000-0002-9536-010X} % LP
%
% Authors, for the paper (add full first names)
\Author{\hl{Antoine Grimaldi} %MDPI: Please carefully check the accuracy of names and affiliations.
 $^{1}$% $^{1,\dagger,\ddagger}$
\orcidA{}, 
Amélie Gruel $^{2}$\orcidB{}, % $^{2,\dagger,\ddagger}$,
Camille Besnainou $^{1}$\orcidC{},
\hl{Jean-Nicolas Jérémie} %MDPI: The name of this author is different from the one submitted online at susy.mdpi.com. Please confirm which one is correct.
 $^{1}$\orcidD{},
Jean Martinet $^{2}$\orcidE{}
and~\hl{Laurent Perrinet} %MDPI: The name of this author is different from the one submitted online at susy.mdpi.com. Please confirm which one is correct.
 $^{1,}$*\orcidF{}}
%
%\longauthorlist{yes}
%
% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent Perrinet}
%
% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Grimaldi, A.; Gruel, A.; Besnainou, C.; Jérémie, J.-N.; Martinet, J.; Perrinet, L.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
%
% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad \hl{INT UMR 7289, Aix Marseille Univ, CNRS, 27 Bd Jean Moulin,} %Please confirm whether the address information is correct.
13005 Marseille, France; \hl{antoine.grimaldi@univ-amu.fr (A.G.); camille.besnainou@etu.univ-amu.fr (C.B.)} %MDPI: We added the email addresses here according to those submitted online at susy.mdpi.com. Please confirm.
 \\
$^{2}$ \quad \hl{SPARKS,  Côte d'Azur, CNRS, I3S, 2000 Rte des Lucioles,} %Please confirm whether the address information is correct.
06900 Sophia-Antipolis, France; \hl{amelie.gruel@i3s.unice.fr (A.G.); jean.martinet@univ-cotedazur.fr (J.M.)} %MDPI: We added the email addresses here according to those submitted online at susy.mdpi.com. Please confirm.

}
%
% Contact information of the corresponding author
\corres{Correspondence: laurent.perrinet@univ-amu.fr}
%
% Current address and/or shared authorship
% \firstnote{Current address: Affiliation 3.} 
% \secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes
%
%\simplesumm{} % Simple summary
%
%\conference{} % An extended version of a conference paper
%
% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Why do neurons communicate through spikes? By definition, spikes are all-or-none neural events which occur at continuous times. In other words, spikes are on one side binary, existing or not without further details, and on the other, can occur at any asynchronous time, without the need for a centralized clock. This stands in stark contrast to the analog representation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks. As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the profusion of recorded data. With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption---a major societal issue in the era of the digital economy and global warming. In this review paper, we provide evidence from biology, theory and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
}
%
% Keywords
\keyword{\textls[-15]{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic} engineering; heterogeneous delays; spiking motifs; polychronization} 
%
% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction: Importance of Precise Spike Timings in the Brain}\label{sec:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Is There a Neural Code?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CONTEXT
Neural activity is directly influenced by our immediate environment and by internal states and is structured in order to generate motor actions. The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection. A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings observed in neurobiological experiments. One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e., an explicit representation \add[Antoine]{of cognitive processes} within the neural activity. Nevertheless, we will use this terminology in all generality to denote the existence of a structure in neural activity. In this respect, it is \change[Antoine]{known}{reasonable to declare} that neural activity may be related to specific measurable variables\change[Antoine]{, and due to Galvani's}{. Since Galvani's} experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}). A central and well-studied way of communication between neurons is specific electrochemical events called action potentials, or spikes, which were first discovered \change[Antoine]{by Lord Adrian}{at the beginning of the $XX^{th}$ century}~\citep{adrian_impulses_1926}. In this study, the frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb.
In the scope of this article, we focus mainly on these spikes~\change[Antoine]{which are brief (about one millisecond) impulses that propagate along the axons of neurons. First, these have the particular property of being essentially binary in their amplitude, i.e., they are prototypical, ``all-or-none''}{in vertebrate systems. They can be described as brief (about one millisecond) and prototypical, i.e., ``all-or-none'', impulses that propagate along the axons of neurons.}
 \change[Antoine]{Typically, action potentials are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite. The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except notably in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
\remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} used the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons. This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.

% GAP
However, computational neuroscience models have suggested that the precise timing within a sequence of spikes may play a crucial role and that neurons \change[Antoine]{may be as well integrators than synchrony detectors}{may be synchrony detectors as well as integrators}~\citep{abeles_role_1982}. \add[Laurent]{In particular, it is possible that the minute arrangement of temporal delays between neurons may provide a computational advantage.} We will investigate this very hypothesis in this review. %Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}. 
In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this \emph{\hl{event-based representation} %MDPI: Please confirm if the italics is unnecessary and can be removed. The following highlights are the same.
} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}. For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{of the rate coding hypothesis that relies only on the firing rate} by considering \change[Antoine]{a representation based on repetitions of spiking motifs at precise times of occurrence.}{representations based on the precise timing of single spikes.} \remove[Antoine]{This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.} Additionally, numerous studies have demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}. All these findings, and more~\citep{bohte_evidence_2004,dilorenzo_spike_2013}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons. A mathematical formalization would be particularly well-suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.  Crucially, validating this hypothesis would also be crucial in our understanding of neural processes. %
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dynamics of Vision and Consequences on the Neural Code}%Precise timing in the neural code enables efficient vision}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Let us start with a focus on the state-of-the-art of the role of dynamics in vision. Broadly speaking, vision is the set of processes that allow us \change[Antoine]{to make sense of the luminous world}{to make sense of the world through luminous signals}, and is an intensively studied field in neuroscience, particularly with respect to deciphering the neural code. In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between the topology of external visual space and its representation on the retina, called retinotopy. The origins of this neuroscientific question can be found in the first experiments from Pierre Flourens which, using lesions in animals, demonstrated the relationship between visual sensations and activity in the cerebral cortex~\citep{flourens_recherches_1842,pearce_marie-jean-pierre_2009}. This was also observed \remove[Antoine]{by Lord Adrian} when recording the activity of the frog's visual system~\citep{adrian_impulses_1926}. In a series of seminal studies,~\citet{hubel_receptive_1968} showed that this activity could be selective to different features, such as the visual orientation or motion of elementary contours. For a large proportion of neurons, there is \change[Antoine]{remarkably a monotonic}{a remarkably monotonic} relationship between the contrast of visual features and the firing frequency of neurons. However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitute essential pieces to this puzzle~\citep{carandini_normalization_2012}.

%----------------------------%

%----------------------------%
In particular, there have been some remarkable findings when studying the dynamics of vision. For instance, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second. \change[Antoine]{His}{Their} experiments consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}. The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that, a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity. These results have been extended to several species, including primates. Different experimental protocols have shown, for example, that the motor response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}. This fast processing correlates with the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and decoding via the EEG if the observer can detect, for example, the presence of an animal~\citep{keysers_speed_2001}. As expected, the performances decrease progressively as the frequency of presentation of the images increases. However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only $14~\ms$ per image.

Although surprising, this speed of the visual cortex in primates is compatible with the latencies that are recorded at the neuro-physiological level. The rapid propagation of the visual information in the thalamus, then in the primary visual cortex takes about $45~\ms$ in the macaque~\citep{schmolesky_signal_1998} and about $60~\ms$ in humans~\citep{vanni_coinciding_2001}. This functioning of visual processing as a forward pass is most prominent in fast processing (see Figure~\ref{fig:thorpe}), and can be complemented with feedback loops from the higher areas to the sensory areas~\citep{lamme_distinct_2000}. An important consequence of this speed of processing of vision is that it implies that processing is carried out using only very few spikes per layer. As a comparison, the latencies in macaque monkeys are approximately as follows: Retina, 20--40~$\ms$; V1, 40--60~$\ms$; IT, 80--100~$\ms$; MC, 140--190~$\ms$; and to finger muscles, 180--260~$\ms$. \add[Laurent]{Note that, since maximal conduction speeds are roughly constant, theses latencies are comparable to that found in humans, with a ratio given by the physical size of the whole system.} It follows that if we consider that a behavioral response occurrs in only $200~\ms$, it would involve about ten processing stages along the ``forward'' pathways of the visual system. Such processes were indeed efficiently reproduced in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}. At the same time, it was demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.  This figure is inspired by similar schematics performed for monkeys in~\citep{thorpe_seeking_2001}. As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{How Precise Spike Timing May Encode Vectors of Real Values}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\add[Laurent]{Let us now focus on one processing step along a cortical pathway. }
% reliability of spike timings
Sensory data are most often represented by continuous values, such as the energy produced by a flow of photons that hits the different photoreceptors of the retina. How may such information be encoded in neuronal activity? The analysis of generic raster plots reveals particular traits that hint at the role of precise timing. For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it, at first sight, inconsistent with the temporal integration of firing rate. Remarkably, it was observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: while the first spike is aligned to stimulus' time, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{bryant_spike_1976,mainen_reliability_1995} (see Figure~\ref{fig:mainen}). However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see~\citep{ermentrout_reliability_2008}). This is consistent with the differential role of different stimulus frequencies (for instance, the gamma range around $80~\si{\Hz}$) on the reliability of the spike timing reported in~\citep{nowak_influence_1997}: ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies. However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''

\begin{figure}[H]
%\centering
\includegraphics[width=.75\textwidth]{figures/visual-latency.pdf}
% thorpe
% \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
\caption{{\bf \hl{Latency of the different processing steps along the human visual pathway.} %MDPI: Please confirm if the bold is unnecessary and can be removed. The following highlights are the same.
} Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes. Different areas are schematically represented by ellipses, and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$. In order, the retina is first activated (20--40~$\ms$), then the thalamus and the primary visual cortex (V1, 60--90~$\ms$). This visual information  projects to the temporal lobe to reach the infero-temporal area (IT, 150~$\ms$) for object recognition. It then reaches the prefrontal cortex (PFC, $180~\ms$), which modulates decision making to evoke the motor cortex (MC, $220~\ms$) which may mediate an action. This is eventually relayed through the spinal cord to trigger finger muscles, with latencies of about 280--400~$\ms$.}\label{fig:thorpe}
\end{figure}
\vspace{-6pt}
\begin{figure}[H]
%\centering
\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
\caption{\textbf{\hl{Reproducibility of the spiking response of a neuron.} }The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus. The stimulus current value over time for a step stimulus (top left) and for a noisy one (top right). Trial repetitions of a leaky integrate-and-fire neuron stimulated by the stimulus on the upper row (middle row). Membrane potential is represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row). While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{\hl{frozen}} noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input. See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{\hl{notebook}} %MDPI: Please confirm if the link is unnecessary and can be removed. The following highlights are the same.
for a replication of the results from~\citep{mainen_reliability_1995} using a simple LIF model.}\label{fig:mainen}
\end{figure}

% retina
\remove[Antoine]{How may information about the sensory world be encoded in the spiking pattern?} At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}. In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed. These results have been extended to natural images and show a qualitatively similar behavior. The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image. Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{\hl{Benham Top}} based on center--surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}. This evidence found in the retina can be extended to other areas, such as the visual cortex.
%----------------------------%
% C_MainenSejnowski1995_Perrinet.ipynb

%----------------------------%

% %----------------------------%
% \begin{figure}
% \centering
% \mynote{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
% \caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
% }\label{fig:roc}
% \end{figure}
% %----------------------------%
In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}. First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of the stimulus onset are considered, for instance, for sound localization~\citep{chase_first-spike_2007}.  Note also that timing is not entirely sensory or internal, but can be used as a general neural coding principle. In~\citep{safaie_turning_2020}, for instance, they found that the ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines. Timing, at least in animals, may thus be fundamentally embodied and situated.''  Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}. These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}. One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron by using the rank of neurons' activation~\citep{delorme_spikenet_1999,bonilla_analyzing_2022}. Another class of artificial SNNs uses precise spike timing as a metric in order to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods. The subsequent surrogate gradient method is now widely used in methods that attempt to transfer performance from analog (CNNs) to spike-based (SNNs) architectures~\citep{zenke_remarkable_2021}. This type of modeling often uses the classical task of categorizing images developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}. For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition or~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning. 
However, the performance of SNNs is still lagging compared to that of analog networks, and the question of the advantage of using spikes in machine learning and computer vision remains open. \change[Antoine]{Efficiently implementing this type of modeling is extremely important with respect to the development of a new generation of event-based cameras}{Improvements in this new generation of Artificial Neural networks (ANN) would bring major advances in terms of efficient computations in machine learning. They would benefit in particular to a new generation of cameras called Silicon Retinas}~\citep{gallego_event-based_2022} (see Section~\ref{sec:neuromorphic}). \remove{These represent visual information as events, similarly to the biologically-inspired representation described above and in contrast with a basic frame-based representation}%~\citep{lichtsteiner_128x128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developed in Section~\ref{sec:neuromorphic}).This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}.

\change[Antoine]{To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.}{~Even if technology lags far behind biology, this introduction demonstrates the importance of timing in neural processes, and we will further review the role of precise spike timing in neural assemblies. We start by reviewing the different hypotheses that aim at deciphering the neural code with spatio-temporal spiking motifs. After listing some biological evidence for the use of precise spike timing, we review some computational models and neuromorphic technics that add this temporal dimension to their computations.}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Role of Precise Spike Timing in Neural Assemblies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information. There are multiple experimental evidences that neurons use different strategies to communicate.}~\cite{maunsell_functional_1983} \add[Antoine]{demonstrate that neurons of visual area MT in macaque monkeys encode motion direction through population activity, offering robustness to single neuron activity fluctuations. For rapid image processing, it has been observed that neurons in the salamander retina represent the spatial structure of an image with the relative timing of their first spikes}~\citep{gollisch_rapid_2008}\add[Antoine]{. In the primary visual area of macaque monkeys, the phase of firing convey information about naturalistic stimuli that cannot be described by firing rates alone}~\citep{montemurro_phase-of-firing_2008}\add[Antoine]{. Among a variety of strategies to communicate, so far we have also evidence that neurons can use correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}. 
\add[Antoine]{In this first section, we introduced the notion of rate coding and demonstrated that spike timing can also carry information. Scientists found experimental evidence for various hypotheses of neural representations such as population coding}~\citep{maunsell_functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase--firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{. In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space. In all generality, this representation can encompass all the previous ones, except for rate coding, which is not defined locally in time. In this section, we choose to describe different hypotheses, making use of spatio-temporal patterns of spikes that can be propagated among neural assemblies.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{One First Hypothesis: Synchronous Firing in Cell Assemblies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
In his book \emph{Corticonics},~\citet{abeles_corticonics_1991} queried as to whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns. The book gradually leads the reader from the macroscopic cortical anatomy and standard electro-physiological properties of single neurons to neural network models. While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{abeles_role_1982,paugam-moisy_computing_2012}. The book then demonstrates that neural assemblies could form so-called ``synfire chains,'' that is, showing the emergence of synchronous activity on subsets of neurons, which could be propagated in a stable fashion. More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}: \add[Antoine]{``cells that fire together wire together''}. Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}. One possible function of this synchronization may serve the binding of information distributed in the brain~\citep{singer_visual_1995, roelfsema_visuomotor_1997}.

Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}. Interestingly, these authors showed that ``accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events. Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent. These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}. It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{\hl{LFP}} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details). \change[Antoine]{Note also that synchronicity may explain some unintuitive results. Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intracortical activity. However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons. Among different explanations, including travelling waves}~\citep{ben-yishai_traveling_1997}\add[Antoine]{, a synchronous activity at the synaptic level may be sufficient to elicit activity in the cortex}~\citep{bruno_cortex_2006}.

Theoretically\add[Antoine]{, even if the vertebrate's neural system is not likely to be modeled only by such deterministic connectivity}~\citep{deneve_bayesian_2004, ballard_dual_2011}, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of a similar size. Each neuron of one group is connected by an excitatory synapse to the next. When a pulse is elicited in the first group, this may generate a spike in the next group. Depending on the concentration of synaptic weight values, this new activity may become more or less synchronized with respect to that of the previous pulse (as measured by the standard deviation of spike times within the pulse). Recursively applying this to a sequence of groups within a chain generates either a synfire propagation or not. A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}. A crucial aspect of this emergence is explained by the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and, in particular, the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}. This balance was, for instance, modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling so-called push--pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}. Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}. In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013}. % (see Figure~\ref{fig:pynn}).
%
%----------------------------%
%\begin{figure}
%\centering
%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
%\caption{A pyNN implementation of a Synfire chain with feedforward inhibition. The background noise  is implemented as random Gaussian current. The state space gives %the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
%\end{figure}
%----------------------------%
%----------------------------%
\begin{figure}[H]
%\centering
\includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
\caption{\textbf{\hl{Simulation of a synfire propagation} \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{\hl{using Brian}}.} The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each. Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreases from $5$ to $1$~\ms). This generates a pulse after a certain processing delay in the second group with a different jitter. While the first two pulses progressively vanish in the following groups, starting from the third input pulse (with a jitter of $3~\ms$), it is propagated to the following groups. This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
\end{figure}
%----------------------------%

Attempts have been made to detect such synfire chains in neurobiological data. \citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Indeed, simultaneously recorded activities of neurons in the primary motor cortex of monkeys exhibited context-dependent, rapid changes in the patterns of coincident action potentials~\citep{riehle_spike_1997,grammont_precise_1999}. It is now commonly accepted that the planning and execution of movements are based on distributed processing by neural populations in motor cortical areas, yet it is less clear \remove[Antoine]{, though,} how these populations organize dynamically to cope with the momentary computational demands.  In~\citep{brette_computing_2012}, the author proposed a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical ANNs. It relies on the simple remark that a serie of repeated coincidences is in itself an invariant. Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, or the pitch of a musical note.

This is also expressed in the idea that different cortical areas could achieve binding by synchrony~\citep{fries_mechanism_2005}. The synchronicity will generate rhythms at different ranges of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down. Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion, i.e., the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}. More generally, it can be shown that the phase of alpha oscillations (about $10$ Hz) is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}. The question still remains open as to know if this is an epiphenomenon or a working mechanism of the neural code.
%
\subsection{A Further Hypothesis: Travelling Waves}
%
To further investigate the role of precise timing, let us also focus on the role of differential timings in an assembly of neurons. As we have seen, a visual feature will induce the firing of different cells at different latencies~\citep{celebrini_dynamics_1993}. Using intracellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is, a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}. In particular, it was shown that the network of the so-called horizontal connections within a cortical area is typically not myelinated, and that this latency basin would be determined by the propagation speed within that area. For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions. Such interactions may underlay anticipatory mechanisms in the primary visual cortex~\citep{benvenuti_anticipatory_2020,le_bec_horizontal_2022}. The underlying process could be the emergence of propagating waves on the surface of the cortex.

Propagating waves in the neuronal response occur in many excitable media and were found in neural systems, for instance, in the retina\note[Antoine]{added ref to Feller}~\citep{feller_dynamic_1997} or the neocortex~\citep{bienenstock_model_1995}. While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries. One possibility is that these waves were systematically missed in trial-averaged data, due to variability. A recent work~\citep{muller_stimulus-evoked_2014} presents a method for detecting propagating waves in noisy multichannel recordings. Applying this method to single-trial voltage-sensitive dye imaging data\note{The reviewer is right to point that calcium imaging may not be the best tool to capture precise spiking motif as both spatial and temporal resolution are not that great. We argue that few imaging techniques can capture the precise spike location, even good spike sorting algorithms. Also, here we study a propagation of activity defining a spatio-temporal pattern that may not have the same precision as some spiking motifs we describe later.}, the authors show that the stimulus-evoked population response in the primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials. A network model suggests that this reliability is the hallmark of the horizontal fiber network of superficial cortical layers. Propagating waves with similar properties occur independently in the secondary visual cortex, but maintain precise phase relations with the waves in the primary visual cortex. These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatio-temporal frame for further neural interactions.

% \begin{figure}
% \centering
% %\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
% \caption{a Spike raster for repeated presentations (%N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line). A single-trial LFP trace is plotted in gray, and the average LFP response is plotted in black. The relative power between baseline (−200 ms to stimulus-onset) and evoked fluctuations (stimulus-onset + 50--250 ms) significantly favored the evoked response (right panel; N = 110 trials; median = 1.89 dB, p = 0.000019, two-tailed Wilcoxon's ranked-sum test). b Same as in (a), but for a low contrast stimulus (2\% Michelson contrast). The relative power between baseline and evoked LFP fluctuations was not statistically different from parity (median = 1.23 dB, p = 0.087 two-tailed Wilcoxon's ranked-sum test). c An example of spontaneous LFP fluctuations structured as a travelling wave recorded from a spatially distributed multielectrode array in marmoset area MT. d Histogram of spontaneous spike probability as a function of the generalized phase of the LFP during fixation. e The average evoked response to low contrast stimuli was stronger when a more excitable phase (+ / - \pi rad) of a spontaneous travelling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test). Data for panels c--e modified from Davis et al.17 with permission.
% }\label{fig:davis}
% \end{figure}

More recently, novel multi-unit recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}. The authors reviewed these findings by considering the mechanisms by which travelling waves are generated, and evaluated their possible roles in cortical function. In particular, spontaneous travelling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}. Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally generated noise. However, fluctuations of spontaneous activity, often organized as travelling waves, shape stimulus-evoked responses and perceptual sensitivity. The mechanisms underlying these waves are unknown. Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings. In that paper, the authors describe a large-scale computational model with topographically organized connectivity and conduction delays relevant to biological scales. They find that spontaneous travelling waves are a general property of these networks. The travelling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave. Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states. Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed {\hl{in vivo}}. %MDPI: we remove italic, please confirm. The following highlights are the same.
Such waves also occur in motor areas and~\citet{linden_movement_2022} have recently presented ensemble recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic~behavior.

Interestingly, it can be shown that these travelling waves could have a measurable impact on the activity of the visual cortex. This was illustrated in a recent study investigating the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}. The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in relatively close visual proximity. The lrAM is the core building block underlying the use of sequences of images to induce the perception of smooth, realistic visual scenes, which is at the base of movies seen in cinema theaters. In this study, the authors used voltage-sensitive dye imaging to record the activity of the primary visual cortex of macaque monkeys to the presentation of the pair of dots presented independently or in conjunction. A probabilistic modelling showed that the activity of the joint presentation induced a suppressive wave in the direction opposed to the perceived direction, shaping the formation of a wave of propagation travelling at a speed compatible with the perceived motion. A computational model validated the hypothesis that this process could be mediated by diffusion in the horizontal layers connecting the different locations within this cortical area. In summary, the study by~\citet{chemla_suppressive_2019} gave a multi-disciplinary account to demonstrate the effect of travelling waves in the visual cortex.
%
% Robust computation with rhythmic spike patterns.~Proceedings of the National Academy of Sciences of the United States of America~116(36), 18050 - 18059.~ https://dx.doi.org/10.1073/pnas.1902653116
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Rediscovered Hypothesis: Precise Spiking Motifs in Cell Assemblies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Travelling waves indicate that spatio-temporal correlations could play an important role in shaping neural activity. For instance, statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it. In particular, this was put in evidence by analyzing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}, which shows the precise spatio-temporal differences in this recurrently connected assembly. The different aspects of information in the data are evaluated by a decoding strategy, highlighting the role of correlations. Note that a similar dataset used in~\citep{schneidman_weak_2006} is available from Michael Berry's lab~\citep{berry_spike_2022} and allows \change[Antoine]{performing and validate}{testing in order to validate or falsify} these hypotheses. However, \remove[Antoine]{a limit of explaining such data with travelling waves is that} in theory, a \add[Antoine]{cortical} travelling wave would be stationary, which is incompatible with the limits in space and time of a neural system. Recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity. ~\citet{miller_visual_2014} showed that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli. Experiments are performed using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity. The study proposes that visual stimuli recruit endogenously generated ensembles to represent visual attributes. Note that evoked ensembles in response to a natural movie played in a loop were precisely timed across repetitions. % (Fig. 7).

From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008}. %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Explictly decoding spiking motifs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}. 
Going further, researchers found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, {\hl{in vivo} %MDPI: We removed the italics. Please confirm this revision.
} and {\hl{in vitro}}. These patterns repeat after minutes, maintaining millisecond accuracy. Indeed,~\citet{ikegaya_synfire_2004} demonstrated that in cortical activity, one can find a repetition of several motifs in spike activity (duration around 1 s +/$-$ 0.5 s, some events in motifs are of similar size but sometimes absent). These sequences can be specific to a particular layer or column, can be synchronized with network activity oscillation, and can involve several cells. They also demonstrated that these sequences can form super sequences, so-called \emph{\hl{cortical songs}}. It consists of the assembly of several sequences which repeat in a specific order with a compressed timing. {\hl{In vivo}} spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3--10 cells out (of about 900), and lasting up to 3 s. Sequences have specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and are associated with a structured spatial organization of the neurons that formed them. %In cortical songs, there is a ``compressing timing'' which may be taken into account by a similar mechanism as maxpooling in CNNs for space, but in time. Or there may be a mechanism for controlling the replay speed (pulvinar, \ldots{} , ?)
% %----------------------------%
% \begin{figure}
% \centering
% \includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
% \caption{Fig. 1. from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity \textit{in vitro} and \textit{in vivo}. (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices. Panels show segments (red) of the same voltage-clamp recording from the same cell, repeating seconds or minutes after the initial occurrence (blue). Arrows indicate timings of repeated PSCs. (i) Upper trace: low--temporal resolution display of spontaneous activity of a neuron. Lower traces: higher resolution display of the repeated motif at indicated regions of the trace (a to c). (ii) Example of a longer motif. (B) Three repetitions of a motif. The top traces show the motifs superimposed on each other (blue, green, and red), the middle traces show these same traces individually, and the bottom traces show temporally magnified regions of the motifs (a to c). (C) Repeated sequences of intracellular current-clamp recordings \textit{in vivo}. Two (i) and three (ii) repetitions of motifs are shown. Shuffle tests were performed on traces (i), a to c, yielding significantly fewer repeats (fig.~S2, P  0.02). In (i), the blue trace is shifted --2.75 mV; in (ii), the blue trace is shifted --1.58 mV, and the green +0.79 mV.}\label{fig:Ikegaya2004}
% \end{figure}
% %----------------------------%

Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}. This is, for instance, relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{cleland_construction_2014, kashiwadani_synchronized_1999, rinberg_speed-accuracy_2006} or touch~\citep{johansson_first_2004}.  In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}. These sequences can be internally generated~\citep{pastalkova_internally_2008,villette_internally_2015} and may be formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets~\citep{malvache_awake_2016}. In that protocol, hippocampal sequences are formed by the ordered activation of smaller sequence motifs. They are stereotyped and robust since neurons can be activated in the same order across days (see Figure~\ref{fig:haimerl} from~\citep{haimerl_internal_2019}). As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.
%
%----------------------------%
\begin{figure}[H]
%\centering
\includegraphics[width=0.98\linewidth]{figures/haimerl2019.jpg}
\caption{\textbf{\hl{An example of a precise temporal motifs seen in cortical slices.}} In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds. (\textbf{A}) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions. Cells were selected and ordered with respect to their activity in the first imaging session (Top). The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{\hl{PNAS License}}).}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.

\end{figure}
%----------------------------%

It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010}, which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some frozen noise patterns. In particular, they used random waveforms to probe the formation of new memories for arbitrary complex sounds. A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4~ \si{\second}$ long. %The task is to detect the repetition of the same (frozen) noise within a trial.
Unbeknownst to listeners, some noise samples reoccurred randomly throughout an experimental block. They showed that the ``repeated exposure induced learning for otherwise totally unpredictable and meaningless sounds'' by showing that the sensitivity increases in that case. Note also that ``acoustical analyses failed to reveal any obvious differences between good and bad noises'' and that ``time reversal had no significant effect on the detection accuracy (which is quite surprising). The learning is unsupervised (statistical, automatic), fast-acting (phase transition, ``insight''), and long-lasting (memorization).

That results suggest that precise spiking motifs are not necessarily grouped on a topography and can be forming apparently randomly arranged connections. Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned. Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006}, which this author defined as \emph{\hl{polychronous groups}}. The rest of this review will be devoted to present evidence for the use of such precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering. As a summary, it seems that such precise structural information is essential to the neural code and that it seems imperative to include this information in the decoding algorithm for a better understanding of neural activity.
%
% TODO: re-read~\citep{luczak_packet-based_2015} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
% \subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{orban_neural_2016}}
% \begin{itemize}
%  \item
%   Stochastic sampling links perceptual uncertainty to neural response variability
% \item
%   Model accounts for independent changes in strength and variability of responses
% \item
%   Model predicts relationship between noise, signal, and spontaneous correlations
% \item
%   Stimulus statistics dependence of response statistics is explained
% \end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Understanding Precise Spiking Motifs in Neurobiology}\label{sec:detection}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Decoding Neural Activity from Firing Rates}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section, we will review current evidence on how we may take advantage of spiking motifs in neurobiology, that is, in an effort to understand actual recordings from biological neural tissues. In most generic computational models, the neural activity is assumed to be encoded in the firing rate. For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point process\note[Laurent]{adding reference}~\citep{simoncelli_characterization_2003}. As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves, that is, its selectivity to a range of features~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}. This latter model assumes a Bernoulli model for the generation of spikes, such that the decoding amounts to a single-layer logistic regression. An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}. In this particular paper, the authors fit the summary statistics of neural data with a (differentiable) spiking network model. The loss function is the cross entropy (following the Bernouilli hypothesis with a GLM, where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995}) and embedded with recurrent dynamics. In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{\hl{code}} and uses the \note[Laurent]{now citing the URL}publicly available V1 dataset~\citep{kohn_utah_2016}, which allows supervising the model, with the input being the movie and the output the spikes recorded. These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
\change[Antoine]{Using the right metric, as implemented in that paper by the corresponding neural models, is essential to better understanding neural data.}{Carefully picking the more appropriate metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.} Importantly, these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Decoding Neural Activity Using Spike Distances}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
There are different solutions to provide with a distance between two given spike trains. A known measure is the Victor--Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}. \change[JN]{In a later study, Van Rossum (2001) defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citep{van_rossum_novel_2001}. \change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}. Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure\change[JN]{which}{, an interesting conclusion since the cross-correlation measure is} that which is adapted to the event-based nature of spiking signals. These observations lead to the intuition that each distance may be as good as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}. % ~\citep{}

%Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
Concerning spike timings,~\citet{levakova_review_2015} \change[JN]{reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning}{reviewed existing methods for estimating the latency of neural responses that include Bayesian binning.} \change[JN]{More broadly }{Alternatively }, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}. This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, particularly in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}. A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}. By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}. This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance. The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, \add[Laurent]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.

\begin{figure}[H]
%\centering
\includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
\caption{\textbf{\hl{Detecting motifs using SPADE.}} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{\hl{tutorial}} on the data generated in  Figure~\ref{fig:diesman}. This allowed to label different precise spike motifs which are denoted by different colors. Spikes belonging to the same motif have the same color. %As the synfire chain gets more precise, the detected patterns grow in size.
}\label{fig:Diesmann_et_al_1999_spade}
\end{figure}

Another important algorithm, called SPOTDisClust, is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}. They introduced an unsupervised method based on their detection from high-dimensional neural ensembles. The algorithm measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other. Many approaches to this problem are supervised. In other words, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for determining hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity~\citep{nadasdy_replay_1999,lee_combinatorial_2004}. In SPOTDisClust, the learning is unsupervised. It uses the prior that there is only one spike per pattern. Using a so-called t-SNE projection (that allows to project this high-dimensional representation to a lower-dimension map) validated that this clustering method can retrieve all patterns from the data. % (see Figure~\ref{fig:SPOTDisClust}). 
The limits of this method are that it is computationally complex, block-based and strictly specialized for the task at hand. To overcome these difficulties, a novel method was recently developed~\citep{sotomayor-gomez_spikeship_2021}.
%   Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
\add[JN]{Whether it is the distance between two given spike trains or a comparison of the spike timings, the complexity and the diversity of the methods used to measure them are witnesses of the growing interest of the integration of these measures in the understanding of the neural code. One of the steps to test their potential usefulness is then to scale these methods to larger amounts of data.}

%----------------------------%

%----------------------------%

% %%----------------------------%
% \remove[LP]{
% \begin{figure}
% \centering
% \includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
% \caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm. \textbf{(A)} Structure of five ``ground-truth'' patterns (\ldots). For each pattern and each neuron, a random position was chosen for the activation pulse. \textbf{(B)} Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A). (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:SPOTDisClust}
% \end{figure}
% %----------------------------%
% }
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scaling Up to Very Large Scale Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts. Bridging across different spatial and temporal scales, combination of {in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks at an unprecedented level of detail and precision. The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurobiological knowledge to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits. As a consequence, there is a growing need to scale these methods to larger amounts of data.

There are multiple \change[Antoine]{solutions}{approaches} which aim at tackling this problem. One algorithm capable of achieving such a daunting task is the Rastermap algorithm~\citep{pachitariu_robustness_2018}. Basically, it rearranges neurons in the raster plot based on the similarity of their activity and applies a deconvolution strategy based on a linear model. \add[Laurent]{Yet this method was mainly tested on calcium imaging data, which are known to add some imprecision to the timing of the original spiking activity.} The model is \note[Laurent]{now citing the URL}openly accessible~\citep{stringer_mouselandrastermap_2020} and has led to important discoveries. In~\citep{stringer_spontaneous_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated. Then, in~\citep{stringer_high-precision_2021}, the authors analyzed spontaneous neural firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements. In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity. Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in~neuroscience.
%   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.

% \subsection{Extending the decoding at multiple time scales}
In the paper by~\citet{russo_cell_2017}, the authors present a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization. It uses  sliding window as in~\citep{grun_unitary_2002} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{pipa_neuroxidence_2008}. They extend the measure to multiple lags~\citep{torre_synchronous_2016}. The core measure is based on a non-stationarity corrected parametric statistical test for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.  To overcome the limits of models which require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} addressed each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time. \change[Laurent]{They also introduce}{As originally introduced by}~\citep{kass_statistical_2005}, they used learnable time-warping parameters to model sequences of varying duration, which were experimentally observed in neural circuits, and demonstrated these advantages on experimental recordings from the songbird higher vocal center and rodent hippocampus.
At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1. They develop a hidden Markov model for quantifying the transitions. In particular, they showed that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision. Top-down attention drives inter-areal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination. Building such models for predicting changes of timings based on context, such as using a so-called change point model for blocked-based experimental protocols~\citep{pasturel_humans_2020}, could therefore provide useful prior information to enhance the decoding from neural activity.
% sparse in time and space : AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355.  CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What Biological Mechanism Could Allow Learning Spiking Motifs?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Despite the evidence for the effectiveness of precise spiking detection we presented above, doubts may remain as to the reliability of this learning mechanism and whether there is a real need for further research on this subject. The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviors allow us to put these doubts to rest. In the following paragraphs, we will successively describe the first biological observations of delay learning, identify myelinization as an important actor, and finally study this phenomenon at the molecular level.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Biological Observations of Delay Adaptation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% First biological observations

One of the first significant pieces of evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_functioning_1938}. Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{madadi_asl_dendritic_2018}. Extensive measures of nerve conduction velocities were performed in different animals, including humans, and significant variations related to age, sex and other factors were identified~\citep{stetson_effects_1992}.

However, it was not until the study of the interaural time difference (ITD) by \remove[Amélie]{gerstnerneuronal1996}~\citet{carr_circuit_1990} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function. This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears. \add[Amélie]{It was first theorized in the Jeffress ``ITD-versus-place model''}~\citep{jeffress_place_1948}. \add[Amélie]{As hypothesized by this model, the \textit{\hl{nucleus laminaris}} of the avian brain contains coincidence detectors and, associated with the \textit{\hl{nucleus magnocellularis}} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}. According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'' and therefore necessarily need to play on the response time to do so. \add[Amélie]{This assertion confused the mean interspike interval, i.e., how often a neuron can fire, and the specific spike time, i.e., how precisely a spike can be generated. However, it has nevertheless contributed to the recognition of the importance of time in various biological mechanisms.} \remove[Amélie]{penacochlear2001 conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris. In the mouse somatosensory cortex, camontiming2019 found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e., the temporal disparity between the left and right cochlear loci in the owl's brain. However, it was quickly set aside, as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears. The authors supported a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors concluded that regulations at different sites within individual axons of at least two parameters, namely, the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays, thus allowing the ITD detection. \add[Amélie]{The authors also noted that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.}

\add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localization behaviour. However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\citep{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{. This mechanism adds to the axonal length delay, which was previously thought to be the sole influence on the conduction velocity due to its anatomical soundness---as}~\citet{seidl_mechanisms_2010}\add[Amélie]{ experimentally demonstrated, this mechanism by itself is not sufficient to explain the biological functionality but should be added to the one of axonal length delay.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Importance of Myelination}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The importance of myelination
\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to \change[Amélie]{both the axon diameter by itself and the diameter of the axon wrapped in myelin}{the axon diameter, i.e., the amount of myelin wrapped around the axon}. Indeed, the oligodendrocytes, one of the many glial cells present in the vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produces thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}. Myelinization consists in ``two motions: the wrapping of the leading edge of the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}. Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the approx $1~\upmu$m gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_neuron-oligodendrocyte_2021}. \change[Antoine]{In}{On} average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\upmu$m long~\citep{simons_oligodendrocytes_2016}. This demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}. Thus, in several occasions, myelin has been identified as \change[Antoine]{a main}{an important} actor in the regulation of conduction velocity in neurons, i.e., axonal delay regulation.

% role of myelination and neuronal delay
\citet{fields_new_2015, fields_myelin_2020} state that myelin facilitates both the neural circuit function and the behavioral performance: experiments on mammals show that myelinization is activity-dependent and directly related to learning and memory consolidation, especially sensory or motor training and in enriched environments. This biological phenomenon takes place both at an early age, where the amount of oligodendrocites is particularly high in the central nervous system~\citep{reynolds_study_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations. Myelination helps memory consolidation by coupling the activity of distant neuron populations and generating nearly synchronous responses in postsynaptic neurons involved among others in path integration~\citep{madadi_asl_dendritic_2018}, as was experimentally demonstrated on mice using a Morris water maze~\citep{steadman_disruption_2020}, contextual fear conditioning~\citep{pan_preservation_2020} or oligodendrocyte precursor cells (OPCs) knock-out~\citep{wan_impaired_2020}. Myelin also inhibits axon sprouting and synapse formation, especially in pyramidal neurons~\citep{fields_new_2015}, and is involved in axonal energy saving through a reduced axonal capacitance and a shift of the metabolic load from axons onto oligodendrocytes~\citep{duncan_neuron-oligodendrocyte_2021}.

It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_new_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{madadi_asl_dendritic_2018}. Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_demyelination_2021}. ~\citet{duncan_neuron-oligodendrocyte_2021} states that ``the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration''.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Interplay of Delay Adaptation and Neural Activity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% How is the neuronal activity detected ?
However, one question remains: how do oligodendrocytes detect neuronal activity and regulate the myelinization accordingly? To answer this, we must first study the myelinization process. The OPCs first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons. Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}. Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as Ca$^{2+}$ activity~\citep{baraban_ca2_2018} of the neuregulin 1 growth factor~\citep{nave_axonal_2006}. The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity. Indeed,~\citet{cullen_periaxonal_2021} experimentally demonstrated that learning and associated neuronal activity modify the Ranvier nodes' length and the periaxonal space width in the adult mouse brain. They also confirmed that the delay correlates with the level of skill acquisition.~\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis. Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or {\hl{via}} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.

A potential scenario for selective myelination on electrically active axons using non-synaptic junctions between an axon and an oligodendrocyte is as follows: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors. This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_new_2015}.

% myelination for axonal delay, but what about dendritic delay ? 
The previous paragraphs present the biological mechanisms behind the axonal delay, regulated by myelination. However, the dendritic delay \add[Amélie]{as well as the axonal delay (see}~\citep{seidl_mechanisms_2010}) seems to play an equally important role in the precise timing within a sequence of spikes. Dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}. Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}. \citet{mel_synaptic_2017} \add[Amélie]{highlighted the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc. Dendritic spiking involvement in synaptic potentiation following active backpropagation into dendrites was experimentally uncovered using calcium imaging to highlight dendritic calcium entry allowing for long-tem potentiation}~\citep{golding_dendritic_2002}. ~\citet{branco_dendritic_2010} demonstrated the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''. The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree and relies on the dendritic calcium influx moderation by NMDA receptors.

All in all, learning spike motifs requires significantly complex pathways and biological mechanisms, whether in the dendrites or the axon of the neuron. More and more is becoming known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop ANNs, whose learning rules would be more neuromorphic. 
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Modeling Precise Spiking Motifs in Theoretical and Computational Neuroscience}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Now that we have reviewed biological foundations for the role of delays in neural computations, we review, in the following section, theoretical models which directly take advantage of using precise spiking motifs. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Spiking neural networks (SNNs)~\citep{maass_networks_1997} are natural candidates to use these precise temporal patterns in the brain. The approach which is currently most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}. One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non-spiking) neural network. In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error's gradient. \change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, the gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical neural network to a spiking architecture~\citep{rueckauer_conversion_2017}. SNNs reach in some case a similar performance as their non-spiking equivalent, for instance on the \remove[Antoine]{N-}MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}. \change[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware. 
However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{So far, this approach does not outperform classical architectures both in term of training efficiency and performances}~\citep{davies_loihi_2018}\add[Antoine]{. However, they remain the best candidates to reproduce biological neural systems and their capacities in terms of accuracy, speed and energy consumption.} There is, therefore, an immense gap in the way we understand biology to translate it to the efficiency of SNNs. To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) \change[Antoine]{to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized}{to a detector of precise spiking motifs}~\citep{paugam-moisy_computing_2012}. In particular, we will explore different existing architectures which are able to overcome the diversity of input presynaptic patterns and learn to detect stable spiking motifs, that is, volleys of spikes which are stable up to a certain onset time (see Figure~\ref{fig:THC}). These models will be compared in light of neuroscientific and computational perspectives. %We review in this section theoretical and computational foundations of spiking motif detection. %In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Izhikevich's Polychronization Model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% polychronization
As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing rate. Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time-encoding machine by}~\citet{lazar_time_2004} and \textit{\hl{polychronization}} model of~\citet{izhikevich_polychronization_2006}. \add[Antoine]{In this section, we focus on the polychronization model} based on a random recurrent model of spiking neurons, including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a spike-time-dependent plasticity (STDP) learning rule~\citep{markram_regulation_1997}. \change[Antoine]{It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons}{It was shown that spike timing (STDP) has an impact on the development of synaptic efficacy for many kinds of neurons}~\citep{caporale_spike_2008}. Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma. It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that delays are then ``frozen'' for the rest of the simulation. Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.,~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs). One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time. This synchrony of arrival at the soma of the neuron leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:izhikevich}). According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group. 
%----------------------------%
%

% 
%----------------------------%

% follow-ups
Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper. Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999}, yet they had not shown potential applications to the detection of spiking motifs. A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}. The \emph{tempotron} is a supervised synaptic learning algorithm, which classifies a distractor from a target motif, in order to extend the perceptron, which does not incorporate a spike timing framework. The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a supervised STDP rule. The limits of this model are that its output is only binary and that its storage capacities are limited. An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}. Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete such as to allow for the reproduction of all results presented in that manuscript. This more recent work details how this code could be slightly corrected. It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process. Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}. In that work, based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation. They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system. Stimulation of the trained network produces the activation of a PG, i.e., the propagation of firing activity through multiple layers due to convergent patterns of firing. While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this~section.

Strikingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system (see Figure~\ref{fig:THC}). In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}. The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the modulation of spike latency emission~\citep{susi_nmnsd-spiking_2021}. Additional models for the detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}. This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}. Recently, by using a logistic regression model coupled with a temporal convolution, a model with heterogeneous delays was implemented to test the detection of the spiking motifs embedded in an event stream~\citep{grimaldi_learning_2022}. This allowed to detect a high number of superposed polychronous motifs in synthetic data, illustrating the computational benefit of such representations compared to that with a unique delay (see Figure~\ref{fig:THC}). As such, these models use the neural dynamics to handle input signals with different delays but do not explicitly take full advantage of the representation capacity offered by heterogeneous delays.

\begin{figure}[H]%[t!][]
%  \centering
  \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
  \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
  \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
  \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
    \caption{\textbf{\hl{Detecting event-based motifs using spiking neurons with heterogeneous delays.}} 
    { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs). 
    { (\textbf{b})}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay. 
    { (\textbf{c})}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}. % such that the addresses are the synapses of these neurons (which we therefore call here synapses).
    %Note that each afferent may be connected with multiple weights at different delays. The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses. %These PG motifs were efficiently learned by a logistic regression.
    %{\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
    { (\textbf{d})}~\hl{Knowing the} %MDPI: Please change the (a) into (d) in picture.
 results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic example, all spikes are now associated with a PG. 
     }
  \label{fig:THC} % TODO: citer biol cyb fig:THC
\end{figure}

\begin{figure}[H]%[t!]
%  \centering
  \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
    \caption{\textbf{\hl{Core mechanism of polychrony detection}~\citep{izhikevich_polychronization_2006}.} {(\textbf{Left})}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}. {(\textbf{Middle})}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons; therefore, no spike will be emitted, as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
    %at these different delays, and these may not be sufficient to generate a spike in either neuron.
    {(\textbf{Right})}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
     }
  \label{fig:izhikevich}
\end{figure}
%
%----------------------------%
%

% 
%----------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Learning Synaptic Delays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Training delays ~\citep{state_training_2019}
First, the original model by Izhikevich uses a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection. For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity--connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}. Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}. They were also extended to recurrent neuronal networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}. In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity. More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity, and these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.

At the implementation level, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}. The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns. Applying this STDP-based rule on delays to the spiking neural network, the experimental results were validated on a simple motion detection task, but were prone to convergence issues. Another model of synaptic delay-weight plasticity integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes~\citep{zhang_supervised_2020}. This was also presented by~\citep{wang_delay_2019}, who proposed a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance. Other models, such as that of~\citep{hazan_memory_2022}, propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST. In a recent paper~\citep{luo_supervised_2022}, the authors proposed a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron. In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
In yet another computational model,~\citet{sun_learning_2016} showed that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.

%In~\citep{duffy_variation_2019}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in birdsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{suthers_motor_2002,wild_descending_1993,yu_temporal_1996} and this model shows how one could vary HVC activity using something similar to dropout in ML. Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs). In~\citep{agus_rapid_2010}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.

% Learning compositional sequences with multiple time scales through a hierarchical network of spiking neurons.
% Maes A, Barahona M, Clopath C.PLoS Comput Biol. 2021

% Characteristics of sequential activity in networks with temporally asymmetric Hebbian learning.
% Gillett M, Pereira U, Brunel N.Proc Natl Acad Sci U S A. 2020

% Unsupervised Learning of Persistent and Sequential Activity.
% Pereira U, Brunel N.Front Comput Neurosci. 2020

% From space to time: Spatial inhomogeneities lead to the emergence of spatiotemporal sequences in spiking neuronal networks.
% Spreizer S, Aertsen A, Kumar A.PLoS Comput Biol. 2019

% Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
% Pang R, Fairhall AL.Elife. 2019 May~\citep{pang_fast_2019}

% Emergence of spontaneous assembly activity in developing neural networks without afferent input.
% Triplett MA, Avitan L, Goodhill GJ.PLoS Comput Biol. 2018

% Training and Spontaneous Reinforcement of Neuronal Assemblies by Spike Timing Plasticity.
% Ocker GK, Doiron B.Cereb Cortex. 2019.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Real-World Applications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%----------------------------%
% TODO

% 
%----------------------------%

A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios. Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods, such as image processing, would further prove their plausibility. For instance, in a recent work,~\citet{ghosh_spatiotemporal_2019} proposed a two-stage unsupervised--supervised system for the categorization of spatio-temporal actions from an event-based stream. The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal-related changes to a local spatio-temporal spike-event pattern. The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity. For testing the system, two datasets are considered: DVS gesture and a new action recognition dataset recorded for this work. Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.

There are more applications to image processing using spiking neural networks. For instance, a set of models are based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections. Using this core computational unit, and extending it to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation. This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks. In essence, two neurons, which were selective to specific motions, were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.  This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot, even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}. This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash--lag illusion~\citep{khoei_flash-lag_2017}. However, these latter models used parametric rules for defining the weights. Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore some exemplar results from neuromorphic engineering to obtain better insights on that aspect.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Applications of Precise Spiking Motifs in Neuromorphic Engineering}
\label{sec:neuromorphic}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Artificial intelligence has made huge advances in the past decades. Deep learning algorithms, nowadays, outperform humans at complex tasks such as natural image recognition or abstract strategy board games. Yet, machine learning algorithms suffer from adversarial attacks or a lack of generalization \change[Antoine]{capacity but}{capacity. However,} their main weakness, compared to biological neural networks, is their poor energy efficiency. Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019, javanshir_advancements_2022}. Because the reduced energy consumption of biological networks can be explained \add{in part} by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to \change{make}{perform} energy-efficient computations. Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips. It can be materialized by the capacity of the physical connections between the processing units to store information~\citep{markovic_physics_2020}. An example of such a connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted. Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems. Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2003, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing techniques show their advantages in terms of frugality and rapidity only on neuromorphic chips.

This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by artificially reproducing such mechanisms, it is interesting to study neural circuitry. Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}. In this section, we give a description of the different neuromorphic hardware that have been developed and see how they can be used to deal with precise temporal motifs. 
%
\subsection{The Emergence of Novel Computational Architectures}
%
To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_analog_1989}. It was implemented with transistors operating in the sub-threshold domain and responded to asynchronous events, but was not capable of discriminating two different spiking sequences with the same firing rate. Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, as it can be observed in biological neurons. Then, electronics circuits became more and more bio-realistic and, two decades later, ~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons. The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector''. Today, many devices are good candidates for implementing event-based algorithms and use the \change[Antoine]{AER}{address event representation (AER)}. They can be divided intro three major categories: digital, analog and mixed analog/digital platforms. For a more complete review, the reader can refer to~\citep{schuman_survey_2017}, where we site the most popular ones. SpiNNaker~\citep{furber_overview_2013, furber_spinnaker_2020}, Loihi~\citep{davies_loihi_2018} and TrueNorth~\citep{merolla_million_2014} chips are widely used examples of digital hardware implementations. Compared to TrueNorth, which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning. This flexibility in the implementation comes at the cost of an increased energy consumption. Mixed analog--digital systems were developed at Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}. They are mostly used by computational neuroscientists to model brain activity with different levels of abstraction. BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed analog--digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}. It is a wafer-scale neuromorphic hardware with analog components. Analog arrays (i.e., field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices. The pulsed current source synapse and the DPI are examples of such devices; we can also mention the field programmable neural array~\citep{farquhar_field_2006} and the NeuroFPAA~\citep{cheng_fpaa_2009} specifically designed for neuromorphic systems. Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing. 
%
%----------------------------%

%----------------------------%
%
Neuromorphic sensors have also been developed with the idea to capture external stimuli more efficiently and closer to biological systems. A widely used example is the dynamic vision sensors (DVSs) which provide a stream of binary asynchronous events signaling detectable changes in luminance. These devices, also named ``silicon retinas'' (see Figure~\ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene. We also report other event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but will focus on DVS for the next subsection about the use of dynamics embedded in event-based signals.

\begin{figure}[H]
\centering
\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
\caption{\textbf{\hl{A miniature, event-based ATIS sensor.}} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides events at the micro-second resolution. These are sparse, as they represent luminance increments or decrements (ON and OFF events, respectively). Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
\end{figure}

%Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems. This field of research is inspired by neuroscientific advances and a computational formalism to design innovative architectures and, by reproducing such mechanisms, it is interesting to study neural circuitry. Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}. Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER). In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field. There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams. Then, we dedicate one subsection to the implementation of delay-based algorithms on neuromorphic chips.  
%
\subsection{On the Importance of Spatio-Temporal Information in Silicon Retinas}
%
With the AER specification and their sub-millisecond temporal precision, DVSs bring a new approach to the storage and processing of visual information. From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor. The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision~tasks.

In~\citep{lagorce_hots_2017}, \textit{\hl{time surfaces}} were introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time. An exponential decay is applied on this delay to obtain the analog values of the time surfaces. It gives more precision to represent recent events and offers an analogy with the LIF spiking neuron. It is a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity, they can capture repeating motifs within the input signals. Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022} and show that this method could be efficiently applied to state-of-the-art benchmarks.

More generally, three-dimensional convolutions in both space and time are another representation of the spiking motifs embedded in the event stream~\citep{ghosh_spatiotemporal_2019, grimaldi_learning_2022, yu_stsc-snn_2022}. With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window (see Figure~\ref{fig:icip}). This representation is only limited by the time step used for the discretization of the signal; this factor defines the temporal precision of the representation. Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}. This non-exhaustive list of complex task solving is not directly linked to biological processes, but shows the potential of the precise temporal resolution of neuromorphic retina-like cameras. By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels. With this type of signal, the use of spatio-temporal motifs embedded in the event streams is essential to solve high-level visual tasks.

\begin{figure}[H]%[t!]
%  \centering
  \includegraphics[width=0.98\linewidth]{figures/2022-09-27_MotionDetection_kernel.png}
  %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
    \caption{\textbf{\hl{Detecting visual motion in an event stream with heterogeneous delays.}} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion. The models' parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected. When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push--pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}. Global weight is globally decreasing from the lowest delay (\textbf{right}) to less recent information (\textbf{left}).
     }
  \label{fig:icip}
\end{figure}

%
\subsection{Computations with Delays in Neuromorphic Hardware}
%
For the rest of this section, we report examples of implementations of event-based algorithms using precise spatio-temporal motifs on neuromorphic hardware. 
\citep{hussain_deltron_2012} implemented a delay-learning algorithm on an analog chip. Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was performed on the analog architecture; training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}. In addition to delay learning, a group at the University of West Sydney developed a neuromorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}. They showed that STDP and spike-timing-dependent delay plasticity rules could be implemented in both a digital and an analog chip. From the network parameters and the physical limitations to store it on-chip, they proved that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network. The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2014}. %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance. 
\citet{pfeil_neuromorphic_2013} implemented STDP on a mixed analog--digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}. Coherence detection on input spikes coming from two sources was obtained with a 50 \hl{$ns$} precision. They claimed that this unsupervised learning denoises the input and compensates for variations between neural components. The variability of response of the analog components can be compensated by population coding for responses robust to noise, and this phenomenon is also observed in biological neural networks~\citep{boerlin_spike-based_2011}. 
A recent work performed the implementation of a sparse vector symbol architecture binding operation on the Loihi chip, delay lines and coincidence detection, used to compute the binding operation~\citep{renner_sparse_2022}. They highlighted the fact that using delays can be expensive \add[Laurent]{notably in memory bandwidth} because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay. Note that this problem is also due to the algorithm used in this paper and that the analog chips must not suffer from this type of~limitation.

Online on-chip learning and computations with delays are still emerging in neuromorphic engineering. The technical challenges linked to the development of this type of implementation and the growing interest in delay learning make advances in this field interesting for the future of computations with precise spatio-temporal motifs. While improvements are still to be made, neuromorphic chips seem to be a good candidate to efficiently make use of these particular features.

%
%they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is well-suited for areas such as pattern recognition and natural language processing. In this paper, we formally describe the STPU, implement the STPU on a field programmable gate array, and show measured performance data.\citep{hill_spike-timing_2017} not accessible
%
%review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)
%
%We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method. Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware~\citep{liang_neuromorphic_2022} 
%
%Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160
%
% solutions usuellement utilisées - surrogate gradient
%for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Summary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
In this review paper, we presented recent evidence for the role of precise spiking motifs in neuroscience. In particular, we showed that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels, and that they may have numerous applications in neuromorphic engineering. In particular, we showed the following:
\begin{itemize}
  \item  The efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlights the importance of precise spike times;
  \item  Growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, travelling waves on arbitrary spiking motifs, \add[Laurent]{and notably that an encoding based on precise spiking motifs may provide huge computational benefits};
  \item  Many theoretical models already exist, taking into account the specificity of spiking motifs, notably by using heterogeneous delays;
  \item  Using precise spiking motifs could ultimately be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
\end{itemize}

 Overall, our reviewing effort has shown that a growing community is focusing on that aspect. This community is based on solid and validated evidence, which is breaking novel grounds thanks to the current technical advances. Moreover, we also showed that this community is highly diverse, operating on {\hl{in vivo}}, {\hl{in computo}} and {\hl{in silico}} systems. As a consequence, the global effort is still largely scattered, which limits its larger acceptance in~neuroscience.

\subsection{Limits}
Additionally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, and neuromorphic) individually present limitations that prevent their widespread application in neuroscience.

First, many models are based on a discretization of time. This assumption is important to allow for a useful representation of neuronal information in order to be processed in computers. This treatment amounts to transforming spike trains into a matrix form for processing in classical machine learning algorithms. This assumption therefore implies an ineffective use of the memory, as this representation transforms the sparse representation of a spike sequence into dense matrices. In addition, this representation can induce errors due to the discretization and the scale of temporal sampling. Finally, this representation encourages the use of classical methods, which are not adapted to disruptive applications, such as event-driven representations.

Moreover, the learning of patterns is often done in a supervised way. Indeed, the problem of detecting polychronous groups implies to infer both the address and the precise time of the occurrence of these motifs. Most of the models we have presented are based on the assumption that at least one of its variables is known: either the pattern, its identity, or its time of occurrence. This constraint is to be put in parallel with the way a biological nervous system works in which learning is performed autonomously, i.e., without supervision. However, we can note that some models can perform such learning, but only in the case of data for which the motifs are easily separable. More generally, to reproduce the efficiency of biological systems, one should account for the different temporal scales of adaptation, from seconds to years. For instance, the scaffolding of neural assemblies seems to follow critical periods during development~\citep{dard_rapid_2022}.

A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output. However, we saw that the neurobiological system uses processing loops within hierarchical graphs. In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}. The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance, the lack of a global clock, or the cost of fully connected topologies. Moreover,  these processes have to be distinguished from judgements on timing, such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts)~\citep{coull_distinction_2022}.
%
\subsection{Perspectives}
%
The limits that we have presented can be treated individually in each model, as evidenced by individual efforts, which try to overcome them. However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.

First, as we already noted, the effort is still largely scattered. This is in part due to the fact that interaction between neurobiology, theoretical and computational neuroscience and neuromorphic engineering are still scarce as of today. It was largely demonstrated that close, bidirectional interactions are essential to foster breakthroughs. For instance, the design of model-driven protocols has proven to be essential in modern neuroscience. Additionally, if neural networks were essential in shaping modern-day machine learning, e.g., computer vision using deep learning, spiking neural networks should prove essential in future emerging technologies.

In that perspective, it is essential that such models are tested on ecologically relevant, real-world scenarios. Indeed, classic convolutional neural networks have emerged as optimal solutions, for example, to classify static images into categories, yet they are not well-adapted for processing dynamic, multimodal sensory flows. The emerging necessity to be able to process more complex flows, such as the multiple flows of information in a car designed for autonomous driving, necessitates modifying such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs. Notably, the protocols used as well in neurobiology, theory or engineering should take into account these novel levels of complexity.

Ultimately, the community should encourage the adoption of open, reproducible science. Indeed, the different models that we have displayed often come with the tools necessary to reproduce the results obtained. This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}. This aspect is essential to foster the emergence of interdisciplinary projects, such as model-driven neurobiological experiments or biologically inspired neuromorphic engineering. Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011}, i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms. As such, this review aims at paving the way to openly share the variety of resources and to offer a unified view on the role of precise spiking motifs in neuroscience.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \authorcontributions{Conceptualization, LP; writing---original draft preparation, all authors; writing---review and editing, all authors; supervision, LP; funding acquisition, LP and JM. All authors have read and agreed to the published version of the manuscript.}

\funding{
This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{\hl{APROVIS3D}}''). % Agence Nationale de la Recherche %HL and 
LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{https://laurentperrinet.github.io/grant/anr-anr/}{\hl{AgileNeuroBot}}''.
% LP received funding 
LP received support from the french government under the France 2030 investment plan, as part of the Initiative d’Excellence d’Aix-Marseille Université – A*MIDEX number AMX-21-RID-025 ``\href{https://laurentperrinet.github.io/grant/polychronies/}{\hl{Polychronies}}''.}

%\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}

%\informedconsent{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.

%Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

\institutionalreview{\hl{ }
%In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.
}

\informedconsent{\hl{ }
%Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.
%
%Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.
}

\dataavailability{This works is made reproducible using the following tools. First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/XXX}{\hl{GitHub}}. %~\citep{Grimaldi2023review}. 
The paper is available as an \href{https://arxiv.org/abs/YYYY.XXX}{\hl{arXiv preprint}} with links to previous versions. Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{\hl{Zotero group}} which was used to regroup relevant literature on the subject.}

\acknowledgments{We would like to thank Hugo Ladret and Frdéric Chavane for valuable feedback while writing this review. Many thanks to Simon Thorpe for the useful discussions and references on the latency of the visual system in human and non-human primates.}

\conflictsofinterest{The authors declare no conflict of interest. The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% Optional
% \sampleavailability{Samples of the compounds ... are available from the authors.}

% %% Only for journal Encyclopedia
% %\entrylink{The Link to this entry published on the encyclopedia platform.}

% \abbreviations{Abbreviations}{
% The following abbreviations are used in this manuscript:\\

% \noindent 
% \begin{tabular}{@{}ll}
% MDPI & Multidisciplinary Digital Publishing Institute\\
% DOAJ & Directory of open access journals\\
% TLA & Three letter acronym\\
% LD & Linear dichroism
% \end{tabular}
% }

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% Optional
% \appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
% \appendixstart
% \appendix
% \section[\appendixname~\thesection]{}
% \subsection[\appendixname~\thesubsection]{}
% The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.

% \begin{table}[H] 
% \caption{This is a table caption.\label{tab5}}
% \newcolumntype{C}{>{\centering\arraybackslash}X}
% \begin{tabularx}{\textwidth}{CCC}
% \toprule
% \textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
% \midrule
% Entry 1		& Data			& Data\\
% Entry 2		& Data			& Data\\
% \bottomrule
% \end{tabularx}
% \end{table}

% \section[\appendixname~\thesection]{}
% All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

%%%%%%%%% REFERENCES
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
%\bibliographystyle{Definitions/chicago2}
\begin{thebibliography}{999}

\bibitem[Piccolino(1997)]{piccolino_luigi_1997}
Piccolino, M.
\newblock Luigi {Galvani} and animal electricity: two centuries after the
  foundation of electrophysiology.
\newblock {\em Trends Neurosci.} {\bf 1997}, {\em 20},~443--448.
\newblock {{https://doi.org/10.1016/S0166-2236(97)01101-6}}.

\bibitem[Adrian and Zotterman(1926)]{adrian_impulses_1926}
Adrian, E.D.; Zotterman, Y.
\newblock The impulses produced by sensory nerve endings.
\newblock {\em  J. Physiol.} {\bf 1926}, {\em 61},~465--483.

\bibitem[Gouras(1960)]{gouras_graded_1960}
Gouras, P.
\newblock Graded potentials of bream retina.
\newblock {\em  J. Physiol.} {\bf 1960}, {\em 152},~487--505.

\bibitem[Perkel {et~al.}(1967{\natexlab{a}})Perkel, Gerstein, and
  Moore]{perkel_neuronal_1967}
Perkel, D.H.; Gerstein, G.L.; Moore, G.P.
\newblock Neuronal {Spike} {Trains} and {Stochastic} {Point} {Processes}: {I}.
  {The} {Single} {Spike} {Train}.
\newblock {\em Biophys. J.} {\bf 1967}, {\em 7},~391--418.
\newblock {{https://doi.org/10.1016/S0006-3495(67)86596-2}}.

\bibitem[Perkel {et~al.}(1967{\natexlab{b}})Perkel, Gerstein, and
  Moore]{perkel_neuronal_1967-1}
Perkel, D.H.; Gerstein, G.L.; Moore, G.P.
\newblock Neuronal {Spike} {Trains} and {Stochastic} {Point} {Processes}: {II}.
  {Simultaneous} {Spike} {Trains}.
\newblock {\em Biophys. J.} {\bf 1967}, {\em 7},~419--440.
\newblock {{https://doi.org/10.1016/S0006-3495(67)86597-4}}.

\bibitem[Abeles(1982)]{abeles_role_1982}
Abeles, M.
\newblock Role of the cortical neuron: integrator or coincidence detector?
\newblock {\em Isr. J. Med. Sci.} {\bf 1982}, {\em
  18},~83--92.

\bibitem[Carr(1993)]{carr_processing_1993}
Carr, C.E.
\newblock Processing of {Temporal} {Information} in the {Brain}.
\newblock {\em Annu. Rev. Neurosci.} {\bf 1993}, {\em 16},~223--243.\linebreak
\newblock {{https://doi.org/10.1146/annurev.ne.16.030193.001255}}.

\bibitem[Davis {et~al.}(2021)Davis, Benigno, Fletterman, Desbordes, Steward,
  Sejnowski, H~Reynolds, and Muller]{davis_spontaneous_2021}
Davis, Z.W.; Benigno, G.B.; Fletterman, C.; Desbordes, T.; Steward, C.;
  Sejnowski, T.J.;~Reynolds, J.H.; Muller, L.
\newblock Spontaneous traveling waves naturally emerge from horizontal fiber
  time delays and travel through locally asynchronous-irregular states.
\newblock {\em Nat. Commun.} {\bf 2021}, {\em 12},~6057.
\newblock {{https://doi.org/10.1038/s41467-021-26175-1}}.

\bibitem[Perrinet {et~al.}(2004)Perrinet, Samuelides, and
  Thorpe]{perrinet_coding_2004}
Perrinet, L.; Samuelides, M.; Thorpe, S.
\newblock Coding static natural images using spiking event times: do neurons
  cooperate?
\newblock {\em IEEE Trans. Neural Netw.} {\bf 2004}, {\em
  15},~1164--1175.
 {{https://doi.org/10.1109/TNN.2004.833303}}.

\bibitem[Gollisch and Meister(2008)]{gollisch_rapid_2008}
Gollisch, T.; Meister, M.
\newblock Rapid {Neural} {Coding} in the {Retina} with {Relative} {Spike}
  {Latencies}.
\newblock {\em Science} {\bf 2008}, {\em 319},~1108--1111.
\newblock {{https://doi.org/10.1126/science.1149639}}.

\bibitem[DeWeese and Zador(2003)]{deweese_binary_2003}
DeWeese, M.R.; Zador, A.M.
\newblock {\em Binary Coding in Auditory Cortex}; \hl{Neural Information Processing
  Systems Foundation:} %MDPI: Please add the location of the publisher (City and Country).
   2003.

\bibitem[Carr and Konishi(1990)]{carr_circuit_1990}
Carr, C.E.; Konishi, M.
\newblock A circuit for detection of interaural time differences in the brain
  stem of the barn owl.
\newblock {\em J. Neurosci.} {\bf 1990}, {\em 10},~3227--3246.
%\newblock Publisher: Society for Neuroscience Section: Articles,
  {{https://doi.org/10.1523/JNEUROSCI.10-10-03227.1990}}.

\bibitem[Bohte(2004)]{bohte_evidence_2004}
Bohte, S.M.
\newblock The evidence for neural information processing with precise
  spike-times: {A} survey.
\newblock {\em Nat. Comput.} {\bf 2004}, {\em 3},~195--206.
\newblock {{https://doi.org/10.1023/B:NACO.0000027755.02868.60}}.

\bibitem[DiLorenzo and Victor(2013)]{dilorenzo_spike_2013}
DiLorenzo, P.M.; Victor, J.D.
\newblock {\em Spike {Timing}: {Mechanisms} and {Function}}; CRC Press:  \hl{Boca Raton, FL, USA,} %newly added information, please confirm
  2013.
%\newblock Google-Books-ID: KTHUIMUpQCUC.  %MDPI: We removed Google-Books-ID: KTHUIMUpQCUC. Please confirm this revision.

\bibitem[Roy {et~al.}(2019)Roy, Jaiswal, and Panda]{roy_towards_2019}
Roy, K.; Jaiswal, A.; Panda, P.
\newblock Towards spike-based machine intelligence with neuromorphic computing.
\newblock {\em Nature} {\bf 2019}, {\em 575},~607--617.
%\newblock Number: 7784 Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/s41586-019-1677-2}}.

\bibitem[Flourens(1842)]{flourens_recherches_1842}
Flourens, M.J.P.
\newblock {\em Recherches Expérimentales sur les Propriétés et les Fonctions
  du Système Nerveux, Dans les Animaux Vertébrés}; Balliere, J.-B., Ed.;  \hl{1842.} %MDPI: Please add the name of the publisher and their location.

\bibitem[Pearce(2009)]{pearce_marie-jean-pierre_2009}
Pearce, J.
\newblock Marie-{Jean}-{Pierre} {Flourens} (1794--1867) and {Cortical}
  {Localization}.
\newblock {\em Eur. Neurol.} {\bf 2009}, {\em 61},~311--314.\linebreak
\newblock {{https://doi.org/10.1159/000206858}}.

\bibitem[Hubel and Wiesel(1968)]{hubel_receptive_1968}
Hubel, D.H.; Wiesel, T.N.
\newblock Receptive fields and functional architecture of monkey striate
  cortex.
\newblock {\em  J. Physiol.} {\bf 1968}, {\em 195},~215--243.
\newblock {{https://doi.org/10.1113/jphysiol.1968.sp008455}}.

\bibitem[Carandini and Heeger(2012)]{carandini_normalization_2012}
Carandini, M.; Heeger, D.J.
\newblock Normalization as a canonical neural computation.
\newblock {\em Nat. Rev. Neurosci.} {\bf 2012}, {\em 13},~51--62.
\newblock {{https://doi.org/10.1038/nrn3136}}.

\bibitem[Thorpe {et~al.}(1996)Thorpe, Fize, and Marlot]{thorpe_speed_1996}
Thorpe, S.; Fize, D.; Marlot, C.
\newblock Speed of processing in the human visual system.
\newblock {\em Nature} {\bf 1996}, {\em 381},~520--522.\linebreak
\newblock {{https://doi.org/10.1038/381520a0}}.

\bibitem[Kirchner and Thorpe(2006)]{kirchner_ultra-rapid_2006}
Kirchner, H.; Thorpe, S.
\newblock Ultra-rapid object detection with saccadic eye movements: {Visual}
  processing speed revisited.
\newblock {\em Vis. Res.} {\bf 2006}, {\em 46},~1762--1776.
%\newblock Loaded from an external bibliography file by Manubot.
%  source\_bibliography: manual-references.bib standard\_id: Kirchner06,
  {{https://doi.org/10.1016/j.visres.2005.10.002}}.

\bibitem[Keysers {et~al.}(2001)Keysers, Xiao, Földiák, and
  Perrett]{keysers_speed_2001}
Keysers, C.; Xiao, D.K.; Földiák, P.; Perrett, D.I.
\newblock The {Speed} of {Sight}.
\newblock {\em J. Cogn. Neurosci.} {\bf 2001}, {\em
  13},~90--101.\linebreak
\newblock {{https://doi.org/10.1162/089892901564199}}.

\bibitem[Schmolesky {et~al.}(1998)Schmolesky, Wang, Hanes, Thompson,
  Leutgeb, Schall, and Leventhal]{schmolesky_signal_1998}
Schmolesky, M.T.; Wang, Y.; Hanes, D.P.; Thompson, K.G.; Leutgeb, S.; Schall,
  J.D.; Leventhal, A.G.
\newblock Signal timing across the macaque visual system.
\newblock {\em J. Neurophysiol.} {\bf 1998}, {\em 79},~3272--3278.
\newblock {{https://doi.org/10.1152/jn.1998.79.6.3272}}.

\bibitem[Vanni {et~al.}(2001)Vanni, Tanskanen, Seppä, Uutela, and
  Hari]{vanni_coinciding_2001}
Vanni, S.; Tanskanen, T.; Seppä, M.; Uutela, K.; Hari, R.
\newblock Coinciding early activation of the human primary visual cortex and
  anteromedial cuneus.
\newblock {\em Proc. Natl. Acad. Sci.  USA} {\bf 2001}, {\em 98},~2776--2780.
\newblock {{https://doi.org/10.1073/pnas.041600898}}.

\bibitem[Lamme and Roelfsema(2000)]{lamme_distinct_2000}
Lamme, V.A.; Roelfsema, P.R.
\newblock The distinct modes of vision offered by feedforward and recurrent
  processing.
\newblock {\em Trends Neurosci.} {\bf 2000}, {\em 23},~571--579.
\newblock {{https://doi.org/10.1016/s0166-2236(00)01657-x}}.

\bibitem[Serre {et~al.}(2007)Serre, Oliva, and
  Poggio]{serre_feedforward_2007}
Serre, T.; Oliva, A.; Poggio, T.
\newblock A feedforward architecture accounts for rapid categorization.
\newblock {\em Proc. Natl. Acad. Sci.  USA} {\bf 2007}, {\em 104},~6424--6429.
%\newblock tex.bdsk-url-2: https://doi.org/10.1073/pnas.0700622104
%  tex.date-added: 2022-05-05 19:08:15 +0200 tex.date-modified: 2022-05-05
%  19:08:15 +0200, 
  {{https://doi.org/10.1073/pnas.0700622104}}.

\bibitem[Jérémie and Perrinet(2022)]{jeremie_ultrafast_2022}
Jérémie, J.N.; Perrinet, L.U.
\newblock Ultrafast image categorization in vivo and in silico. \emph{arXiv}  \textbf{2022}, 
\newblock arXiv:2205.03635.

\bibitem[Nowak and Bullier(1997)]{nowak_timing_1997}
Nowak, L.G.; Bullier, J.
\newblock The {Timing} of {Information} {Transfer} in the {Visual} {System}. In
  {\em Extrastriate {Cortex} in {Primates}}; Springer: \hl{Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
  1997; pp. 205--241.

\bibitem[Thorpe and Fabre-Thorpe(2001)]{thorpe_seeking_2001}
Thorpe, S.J.; Fabre-Thorpe, M.
\newblock Seeking {Categories} in the {Brain}.
\newblock {\em Science} {\bf 2001}, {\em 291},~260--263.
\newblock {{https://doi.org/10.1126/science.1058249}}.

\bibitem[Rucci {et~al.}(2018)Rucci, Ahissar, and Burr]{rucci_temporal_2018}
Rucci, M.; Ahissar, E.; Burr, D.
\newblock Temporal {Coding} of {Visual} {Space}.
\newblock {\em Trends Cogn. Sci.} {\bf 2018}, {\em 22},~883--895.\linebreak
\newblock {{https://doi.org/10.1016/j.tics.2018.07.009}}.

\bibitem[Softky and Koch(1993)]{softky_highly_1993}
Softky, W.; Koch, C.
\newblock The highly irregular firing of cortical cells is inconsistent with
  temporal integration of random {EPSPs}.
\newblock {\em  J. Neurosci.} {\bf 1993}, {\em 13},~334--350.
\newblock {{https://doi.org/10.1523/jneurosci.13-01-00334.1993}}.

\bibitem[Bryant and Segundo(1976)]{bryant_spike_1976}
Bryant, H.L.; Segundo, J.P.
\newblock Spike initiation by transmembrane current: a white-noise analysis.
\newblock {\em  J. Physiol.} {\bf 1976}, {\em 260},~279--314.
%\newblock \_eprint:
%  https://onlinelibrary.wiley.com/doi/pdf/10.1113/jphysiol.1976.sp011516,
  {{https://doi.org/10.1113/jphysiol.1976.sp011516}}.

\bibitem[Mainen and Sejnowski(1995)]{mainen_reliability_1995}
Mainen, Z.F.; Sejnowski, T.J.
\newblock Reliability of {Spike} {Timing} in {Neocortical} {Neurons}.
\newblock {\em Science} {\bf 1995}, {\em 268},~1503--1506.\linebreak
\newblock {{https://doi.org/10.1126/science.7770778}}.

\bibitem[Ermentrout {et~al.}(2008)Ermentrout, Galán, and
  Urban]{ermentrout_reliability_2008}
Ermentrout, G.B.; Galán, R.F.; Urban, N.N.
\newblock Reliability, synchrony and noise.
\newblock {\em Trends Neurosci.} {\bf 2008}, {\em 31},~428--434.
\newblock {{https://doi.org/10.1016/j.tins.2008.06.002}}.

\bibitem[Nowak(1997)]{nowak_influence_1997}
Nowak, L.
\newblock Influence of low and high frequency inputs on spike timing in visual
  cortical neurons.
\newblock {\em Cereb. Cortex} {\bf 1997}, {\em 7},~487--501.
\newblock {{https://doi.org/10.1093/cercor/7.6.487}}.

\bibitem[Kenyon {et~al.}(2004)Kenyon, Hill, Theiler, George, and
  Marshak]{kenyon_theory_2004}
Kenyon, G.T.; Hill, D.; Theiler, J.; George, J.S.; Marshak, D.W.
\newblock A theory of the {Benham} {Top} based on center–surround
  interactions in the parvocellular pathway.
\newblock {\em Neural Netw.} {\bf 2004}, {\em 17},~773--786.
\newblock {{https://doi.org/10.1016/j.neunet.2004.05.005}}.

\bibitem[Celebrini {et~al.}(1993)Celebrini, Thorpe, Trotter, and
  Imbert]{celebrini_dynamics_1993}
Celebrini, S.; Thorpe, S.; Trotter, Y.; Imbert, M.
\newblock Dynamics of orientation coding in area {V1} of the awake primate.
\newblock {\em Vis. Neurosci.} {\bf 1993}, {\em 10},~811--825.
\newblock {{https://doi.org/10.1017/s0952523800006052}}.

\bibitem[Chase and Young(2007)]{chase_first-spike_2007}
Chase, S.M.; Young, E.D.
\newblock First-spike latency information in single neurons increases when
  referenced to population onset.
\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2007},
  {\em 104},~5175--5180.
\newblock {{https://doi.org/10.1073/pnas.0610368104}}.

\bibitem[Safaie {et~al.}(2020)Safaie, Jurado-Parras, Sarno, Louis,
  Karoutchi, Petit, Pasquet, Eloy, and Robbe]{safaie_turning_2020}
Safaie, M.; Jurado-Parras, M.T.; Sarno, S.; Louis, J.; Karoutchi, C.; Petit,
  L.F.; Pasquet, M.O.; Eloy, C.; Robbe, D.
\newblock Turning the body into a clock: {Accurate} timing is facilitated by
  simple stereotyped interactions with the environment.
\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2020},
  {\em 117},~13084--13093.
\newblock {{https://doi.org/10.1073/pnas.1921226117}}.

\bibitem[Gautrais and Thorpe(1998)]{gautrais_rate_1998}
Gautrais, J.; Thorpe, S.
\newblock Rate coding versus temporal order coding: a theoretical approach.
\newblock {\em Biosystems} {\bf 1998}, {\em 48},~57--65.
%\newblock Publisher: Elsevier,
  {{https://doi.org/10.1016/S0303-2647(98)00050-1}}.

\bibitem[Delorme {et~al.}(1999)Delorme, Gautrais, Van~Rullen, and
  Thorpe]{delorme_spikenet_1999}
Delorme, A.; Gautrais, J.; Van~Rullen, R.; Thorpe, S.
\newblock {SpikeNET}: {A} simulator for modeling large networks of integrate
  and fire neurons.
\newblock {\em Neurocomputing} {\bf 1999}, {\em 26},~989--996.
%\newblock Publisher: Elsevier,
  {{https://doi.org/10.1016/S0925-2312(99)00095-8}}.

\bibitem[Delorme {et~al.}(2000)Delorme, Richard, and
  Fabre-Thorpe]{delorme_ultra-rapid_2000}
Delorme, A.; Richard, G.; Fabre-Thorpe, M.
\newblock Ultra-rapid categorisation of natural scenes does not rely on colour
  cues: a study in monkeys and humans.
\newblock {\em Vis. Res.} {\bf 2000}, {\em 40},~2187--2200.
\newblock {{https://doi.org/10.1016/S0042-6989(00)00083-3}}.

\bibitem[Bonilla {et~al.}(2022)Bonilla, Gautrais, Thorpe, and
  Masquelier]{bonilla_analyzing_2022}
Bonilla, L.; Gautrais, J.; Thorpe, S.; Masquelier, T.
\newblock Analyzing time-to-first-spike coding schemes.
\newblock {\em Front. Neurosci.} {\bf 2022}, {\em 16}, 971937.
%\newblock Publisher: Frontiers,
  {{https://doi.org/10.3389/fnins.2022.971937}}.

\bibitem[Bohte {et~al.}(2002)Bohte, Kok, and
  La~Poutré]{bohte_error-backpropagation_2002}
Bohte, S.M.; Kok, J.N.; La~Poutré, H.
\newblock Error-backpropagation in temporally encoded networks of spiking
  neurons.
\newblock {\em Neurocomputing} {\bf 2002}, {\em 48},~17--37.
\newblock {{https://doi.org/10.1016/S0925-2312(01)00658-0}}.

\bibitem[Zenke and Vogels(2021)]{zenke_remarkable_2021}
Zenke, F.; Vogels, T.P.
\newblock The {Remarkable} {Robustness} of {Surrogate} {Gradient} {Learning}
  for {Instilling} {Complex} {Function} in {Spiking} {Neural} {Networks}.
\newblock {\em Neural Comput.} {\bf 2021}, {\em 33},~899--925.
\newblock {{https://doi.org/10.1162/neco\_a\_01367}}.

\bibitem[Göltz {et~al.}(2021)Göltz, Kriener, Baumbach, Billaudelle,
  Breitwieser, Cramer, Dold, Kungl, Senn, Schemmel, Meier, and
  Petrovici]{goltz_fast_2021}
Göltz, J.; Kriener, L.; Baumbach, A.; Billaudelle, S.; Breitwieser, O.;
  Cramer, B.; Dold, D.; Kungl, A.F.; Senn, W.; Schemmel, J.;  et~al.
\newblock Fast and energy-efficient neuromorphic deep learning with first-spike
  times. \emph{arXiv}  \textbf{2021}, arXiv:1912.11443.

\bibitem[Kheradpisheh {et~al.}(2018)Kheradpisheh, Ganjtabesh, Thorpe, and
  Masquelier]{kheradpisheh_stdp-based_2018}
Kheradpisheh, S.R.; Ganjtabesh, M.; Thorpe, S.J.; Masquelier, T.
\newblock {STDP}-based spiking deep convolutional neural networks for object
  recognition.
\newblock {\em Neural Netw.} {\bf 2018}, {\em 99},~56--67.
\newblock {{https://doi.org/10.1016/j.neunet.2017.12.005}}.

\bibitem[Tavanaei {et~al.}(2018)Tavanaei, Masquelier, and
  Maida]{tavanaei_representation_2018}
Tavanaei, A.; Masquelier, T.; Maida, A.
\newblock Representation learning using event-based {STDP}.
\newblock {\em Neural Netw.} {\bf 2018}, {\em 105},~294--303.
\newblock {{https://doi.org/10.1016/j.neunet.2018.05.018}}.

\bibitem[Gallego {et~al.}(2022)Gallego, Delbruck, Orchard, Bartolozzi, Taba,
  Censi, Leutenegger, Davison, Conradt, Daniilidis, and
  Scaramuzza]{gallego_event-based_2022}
Gallego, G.; Delbruck, T.; Orchard, G.; Bartolozzi, C.; Taba, B.; Censi, A.;
  Leutenegger, S.; Davison, A.J.; Conradt, J.; Daniilidis, K.;  et~al.
\newblock Event-{Based} {Vision}: {A} {Survey}.
\newblock {\em IEEE Trans. Pattern Anal. Mach. Intell.}
  {\bf 2022}, {\em 44},~154--180.
\newblock {{https://doi.org/10.1109/TPAMI.2020.3008413}}.

\bibitem[Maunsell and Van~Essen(1983)]{maunsell_functional_1983}
Maunsell, J.H.; Van~Essen, D.C.
\newblock Functional properties of neurons in middle temporal visual area of
  the macaque monkey. {I}. {Selectivity} for stimulus direction, speed, and
  orientation.
\newblock {\em J. Neurophysiol.} {\bf 1983}, {\em 49},~1127--1147.
\newblock {{https://doi.org/10.1152/jn.1983.49.5.1127}}.

\bibitem[Montemurro {et~al.}(2008)Montemurro, Rasch, Murayama, Logothetis,
  and Panzeri]{montemurro_phase--firing_2008}
Montemurro, M.A.; Rasch, M.J.; Murayama, Y.; Logothetis, N.K.; Panzeri, S.
\newblock Phase-of-{Firing} {Coding} of {Natural} {Visual} {Stimuli} in
  {Primary} {Visual} {Cortex}.
\newblock {\em Curr. Biol.} {\bf 2008}, {\em 18},~375--380.
%\newblock Publisher: Elsevier,
  {{https://doi.org/10.1016/j.cub.2008.02.023}}.

\bibitem[deCharms and Merzenich(1996)]{decharms_primary_1996}
deCharms, R.C.; Merzenich, M.M.
\newblock Primary cortical representation of sounds by the coordination of
  action-potential timing.
\newblock {\em Nature} {\bf 1996}, {\em 381},~610--613.
%\newblock Number: 6583 Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/381610a0}}.

\bibitem[Vinje and Gallant(2000)]{vinje_sparse_2000}
Vinje, W.E.; Gallant, J.L.
\newblock Sparse {Coding} and {Decorrelation} in {Primary} {Visual} {Cortex}
  {During} {Natural} {Vision}.
\newblock {\em Science} {\bf 2000}, {\em 287}, 1273--1276.
%\newblock Publisher: American Association for the Advancement of Science,
  {{https://doi.org/10.1126/science.287.5456.1273}}.

\bibitem[Abeles(1991)]{abeles_corticonics_1991}
Abeles, M.
\newblock {\em Corticonics: Neural Circuits of the Cerebral Cortex}; Cambridge
  University Press: Cambridge, MA, USA; New York, NY, USA, 1991.

\bibitem[Paugam-Moisy and Bohte(2012)]{paugam-moisy_computing_2012}
Paugam-Moisy, H.; Bohte, S.M.
\newblock Computing with spiking neuron networks. In {\em Handbook of Natural
  Computing}; Springer:  \hl{Berlin/Heidelberg, Germany,} %newly added information, please confirm
  2012.

\bibitem[Hebb(1949)]{hebb_organization_1949}
Hebb, D.O.
\newblock {\em The Organization of Behavior: {A} Neuropsychological Theory};
  Wiley: New York, NY, USA, 1949.

\bibitem[Harris {et~al.}(2003)Harris, Csicsvari, Hirase, Dragoi, and
  Buzsáki]{harris_organization_2003}
Harris, K.D.; Csicsvari, J.; Hirase, H.; Dragoi, G.; Buzsáki, G.
\newblock Organization of cell assemblies in the hippocampus.
\newblock {\em Nature} {\bf 2003}, {\em 424},~552--556.
\newblock {{https://doi.org/10.1038/nature01834}}.

\bibitem[Singer and Gray(1995)]{singer_visual_1995}
Singer, W.; Gray, C.M.
\newblock Visual {Feature} {Integration} and the {Temporal} {Correlation}
  {Hypothesis}.
\newblock {\em Annu. Rev. Neurosci.} {\bf 1995}, {\em 18},~555--586.
\newblock {{https://doi.org/10.1146/annurev.ne.18.030195.003011}}.

\bibitem[Roelfsema {et~al.}(1997)Roelfsema, Engel, König, and
  Singer]{roelfsema_visuomotor_1997}
Roelfsema, P.R.; Engel, A.K.; König, P.; Singer, W.
\newblock Visuomotor integration is associated with zero time-lag
  synchronization among cortical areas.
\newblock {\em Nature} {\bf 1997}, {\em 385},~157--161.
\newblock {{https://doi.org/10.1038/385157a0}}.

\bibitem[Riehle {et~al.}(1997)Riehle, Grun, Diesmann, and
  Aertsen]{riehle_spike_1997}
Riehle, A.; Grun, S.; Diesmann, M.; Aertsen, A.
\newblock Spike synchronization and rate modulation differentially involved in
  motor cortical function.
\newblock {\em Science} {\bf 1997}, {\em 278},~1950--1953.
%\newblock Publisher: American Association for the Advancement of Science,
  {{https://doi.org/10.1126/science.278.5345.1950}}.

\bibitem[Kilavik {et~al.}(2009)Kilavik, Roux, Ponce-Alvarez, Confais, Grun,
  and Riehle]{kilavik_long-term_2009}
Kilavik, B.E.; Roux, S.; Ponce-Alvarez, A.; Confais, J.; Grun, S.; Riehle, A.
\newblock Long-{Term} {Modifications} in {Motor} {Cortical} {Dynamics}
  {Induced} by {Intensive} {Practice}.
\newblock {\em J. Neurosci.} {\bf 2009}, {\em 29},~12653--12663.
\newblock {{https://doi.org/10.1523/jneurosci.1554-09.2009}}.

\bibitem[Grammont and Riehle(2003)]{grammont_spike_2003}
Grammont, F.; Riehle, A.
\newblock Spike synchronization and firing rate in a population of motor
  cortical neurons in relation to movement direction and reaction time.
\newblock {\em Biol. Cybern.} {\bf 2003}, {\em 88},~360--373.
\newblock {{https://doi.org/10.1007/s00422-002-0385-3}}.

\bibitem[Denker {et~al.}(2018)Denker, Zehl, Kilavik, Diesmann, Brochier,
  Riehle, and Grün]{denker_lfp_2018}
Denker, M.; Zehl, L.; Kilavik, B.E.; Diesmann, M.; Brochier, T.; Riehle, A.;
  Grün, S.
\newblock {LFP} beta amplitude is linked to mesoscopic spatio-temporal phase
  patterns.
\newblock {\em Sci. Rep.} {\bf 2018}, {\em 8}.
\newblock {{https://doi.org/10.1038/s41598-018-22990-7}}.

\bibitem[Torre {et~al.}(2016)Torre, Canova, Denker, Gerstein, Helias, and
  Grün]{torre_asset_2016}
Torre, E.; Canova, C.; Denker, M.; Gerstein, G.; Helias, M.; Grün, S.
\newblock {ASSET}: {Analysis} of {Sequences} of {Synchronous} {Events} in
  {Massively} {Parallel} {Spike} {Trains}.
\newblock {\em PLOS Comput. Biol.} {\bf 2016}, {\em 12},~e1004939.
\newblock {{https://doi.org/10.1371/journal.pcbi.1004939}}.

\bibitem[Ben-yishai and Hansel(1997)]{ben-yishai_traveling_1997}
Ben-yishai, R.; Hansel, D.
\newblock Traveling {Waves} and the {Processing} of {Weakly} {Tuned} {Inputs}
  in a {Cortical} {Network} {Module}.  \emph{J. Comput. Neurosci.} {\bf 1997}.
\newblock {\em 77},~57--77.

\bibitem[Bruno and Sakmann(2006)]{bruno_cortex_2006}
Bruno, R.M.; Sakmann, B.
\newblock Cortex {Is} {Driven} by {Weak} but {Synchronously} {Active}
  {Thalamocortical} {Synapses}.
\newblock {\em Science} {\bf 2006}, {\em 312},~1622--1627.
\newblock {{https://doi.org/10.1126/science.1124593}}.

\bibitem[Deneve(2004)]{deneve_bayesian_2004}
Deneve, S.
\newblock Bayesian inference in spiking neurons.
\newblock In \emph{Proceedings of the Advances in {Neural} {Information} {Processing}
  {Systems}}; MIT Press:  \hl{Cambridge, MA, USA,} %newly added information, please confirm
  2004, Volume~17.

\bibitem[Ballard and Jehee(2011)]{ballard_dual_2011}
Ballard, D.; Jehee, J.
\newblock Dual {Roles} for {Spike} {Signaling} in {Cortical} {Neural}
  {Populations}.
\newblock {\em Front. Comput. Neurosci.} {\bf 2011}, {\em 5}, 22.

\bibitem[Gewaltig {et~al.}(2001)Gewaltig, Diesmann, and
  Aertsen]{gewaltig_propagation_2001}
Gewaltig, M.O.; Diesmann, M.; Aertsen, A.
\newblock Propagation of cortical synfire activity: survival probability in
  single trials and stability in the mean.
\newblock {\em Neural Netw.} {\bf 2001}, {\em 14},~657--673.
\newblock {{https://doi.org/10.1016/S0893-6080(01)00070-3}}.

\bibitem[Gerstner(1995)]{gerstner_time_1995}
Gerstner, W.
\newblock Time structure of the activity in neural network models.
\newblock {\em Phys. Rev. E} {\bf 1995}, {\em 51},~738--758.\linebreak
\newblock {{https://doi.org/10.1103/physreve.51.738}}.

\bibitem[Azouz and Gray(2008)]{azouz_stimulus-selective_2008}
Azouz, R.; Gray, C.M.
\newblock Stimulus-selective spiking is driven by the relative timing of
  synchronous excitation and disinhibition in cat striate neurons {in
  vivo}.
\newblock {\em Eur. J. Neurosci.} {\bf 2008}, {\em
  28},~1286--1300.
\newblock {{https://doi.org/10.1111/j.1460-9568.2008.06434.x}}.

\bibitem[Kremkow {et~al.}(2016)Kremkow, Perrinet, Monier, Alonso, Aertsen,
  Frégnac, and Masson]{kremkow_push-pull_2016}
Kremkow, J.; Perrinet, L.U.; Monier, C.; Alonso, J.M.; Aertsen, A.; Frégnac,
  Y.; Masson, G.S.
\newblock Push-{Pull} {Receptive} {Field} {Organization} and {Synaptic}
  {Depression}: {Mechanisms} for {Reliably} {Encoding} {Naturalistic} {Stimuli}
  in {V1}.
\newblock {\em Front. Neural Circuits} {\bf 2016}, {\em 10}, 37.
\newblock {{https://doi.org/10.3389/fncir.2016.00037}}.

\bibitem[Aviel {et~al.}(2003)Aviel, Mehring, Abeles, and
  Horn]{aviel_embedding_2003}
Aviel, Y.; Mehring, C.; Abeles, M.; Horn, D.
\newblock On {Embedding} {Synfire} {Chains} in a {Balanced} {Network}.
\newblock {\em Neural Comput.} {\bf 2003}, {\em 15},~1321--1340.
\newblock {{https://doi.org/10.1162/089976603321780290}}.

\bibitem[Kremkow {et~al.}(2010)Kremkow, Perrinet, Masson, and
  Aertsen]{kremkow_functional_2010}
Kremkow, J.; Perrinet, L.U.; Masson, G.S.; Aertsen, A.
\newblock Functional consequences of correlated excitatory and inhibitory
  conductances in cortical networks.
\newblock {\em J. Comput. Neurosci.} {\bf 2010}, {\em
  28},~579--594.
\newblock {{https://doi.org/10.1007/s10827-010-0240-9}}.

\bibitem[Davison(2008)]{davison_pynn_2008}
Davison, A.P.
\newblock {PyNN}: A common interface for neuronal network simulators.
\newblock {\em Front. Neuroinformatics} {\bf 2008}, {\em 2}, 11.
\newblock {{https://doi.org/10.3389/neuro.11.011.2008}}.

\bibitem[Pfeil {et~al.}(2013)Pfeil, Grübl, Jeltsch, Müller, Müller,
  Petrovici, Schmuker, Brüderle, Schemmel, and Meier]{pfeil_six_2013}
Pfeil, T.; Grübl, A.; Jeltsch, S.; Müller, E.; Müller, P.; Petrovici, M.A.;
  Schmuker, M.; Brüderle, D.; Schemmel, J.; Meier, K.
\newblock Six {Networks} on a {Universal} {Neuromorphic} {Computing}
  {Substrate}.
\newblock {\em Front. Neurosci.} {\bf 2013}, {\em 7}, 11.
\newblock {{https://doi.org/10.3389/fnins.2013.00011}}.

\bibitem[Schrader {et~al.}(2008)Schrader, Grün, Diesmann, and
  Gerstein]{schrader_detecting_2008}
Schrader, S.; Grün, S.; Diesmann, M.; Gerstein, G.L.
\newblock Detecting {Synfire} {Chain} {Activity} {Using} {Massively} {Parallel}
  {Spike} {Train} {Recording}.
\newblock {\em J. Neurophysiol.} {\bf 2008}, {\em 100},~2165--2176.
\newblock {{https://doi.org/10.1152/jn.01245.2007}}.

\bibitem[Grammont and Riehle(1999)]{grammont_precise_1999}
Grammont, F.; Riehle, A.
\newblock Precise spike synchronization in monkey motor cortex involved in
  preparation for movement.
\newblock {\em Exp. Brain Res.} {\bf 1999}, {\em 128},~118--122.
\newblock {{https://doi.org/10.1007/s002210050826}}.

\bibitem[Brette(2012)]{brette_computing_2012}
Brette, R.
\newblock Computing with {Neural} {Synchrony}.
\newblock {\em PLoS Comput. Biol.} {\bf 2012}, {\em 8},~e1002561.
\newblock {{https://doi.org/10.1371/journal.pcbi.1002561}}.

\bibitem[Fries(2005)]{fries_mechanism_2005}
Fries, P.
\newblock A mechanism for cognitive dynamics: neuronal communication through
  neuronal coherence.
\newblock {\em Trends Cogn. Sci.} {\bf 2005}, {\em 9},~474--480.
\newblock {{https://doi.org/10.1016/j.tics.2005.08.011}}.

\bibitem[VanRullen {et~al.}(2006)VanRullen, Reddy, and
  Koch]{vanrullen_continuous_2006}
VanRullen, R.; Reddy, L.; Koch, C.
\newblock The {Continuous} {Wagon} {Wheel} {Illusion} {Is} {Associated} with
  {Changes} in {Electroencephalogram} {Power} at 13 {Hz}.
\newblock {\em J. Neurosci.} {\bf 2006}, {\em 26},~502--507.
 {{https://doi.org/10.1523/jneurosci.4654-05.2006}}.

\bibitem[Dugue {et~al.}(2011)Dugue, Marque, and VanRullen]{dugue_phase_2011}
Dugue, L.; Marque, P.; VanRullen, R.
\newblock The {Phase} of {Ongoing} {Oscillations} {Mediates} the {Causal}
  {Relation} between {Brain} {Excitation} and {Visual} {Perception}.
\newblock {\em J. Neurosci.} {\bf 2011}, {\em 31},~11889--11893.
\newblock {{https://doi.org/10.1523/JNEUROSCI.1161-11.2011}}.

\bibitem[Bringuier {et~al.}(1999)Bringuier, Chavane, Glaeser, and
  Frégnac]{bringuier_horizontal_1999}
Bringuier, V.; Chavane, F.; Glaeser, L.; Frégnac, Y.
\newblock Horizontal {Propagation} of {Visual} {Activity} in the {Synaptic}
  {Integration} {Field} of {Area} 17 {Neurons}.
\newblock {\em Science} {\bf 1999}, {\em 283},~695--699.
%\newblock 00535 Loaded from an external bibliography file by Manubot.
%  source\_bibliography: manual-references.bib standard\_id: Bringuier99,
  {{https://doi.org/10.1126/science.283.5402.695}}.

\bibitem[Benvenuti {et~al.}(2020)Benvenuti, Chemla, Boonman, Perrinet,
  Masson, and Chavane]{benvenuti_anticipatory_2020}
Benvenuti, G.; Chemla, S.; Boonman, A.; Perrinet, L.U.; Masson, G.S.; Chavane,
  F.
\newblock Anticipatory responses along motion trajectories in awake monkey area
  {V1}.
\newblock {\em bioRxiv  Prepr. Serv. Biol.} {\bf 2020}.
\newblock {{https://doi.org/10.1101/2020.03.26.010017}}.

\bibitem[Le~Bec {et~al.}(2022)Le~Bec, Troncoso, Desbois, Passarelli, Baudot,
  Monier, Pananceau, and Frégnac]{le_bec_horizontal_2022}
Le~Bec, B.; Troncoso, X.G.; Desbois, C.; Passarelli, Y.; Baudot, P.; Monier,
  C.; Pananceau, M.; Frégnac, Y.
\newblock Horizontal connectivity in {V1}: {Prediction} of coherence in contour
  and motion integration.
\newblock {\em PLoS ONE} {\bf 2022}, {\em 17},~e0268351.
\newblock {{https://doi.org/10.1371/journal.pone.0268351}}.

\bibitem[Feller {et~al.}(1997)Feller, Butts, Aaron, Rokhsar, and
  Shatz]{feller_dynamic_1997}
Feller, M.B.; Butts, D.A.; Aaron, H.L.; Rokhsar, D.S.; Shatz, C.J.
\newblock Dynamic {Processes} {Shape} {Spatiotemporal} {Properties} of
  {Retinal} {Waves}.
\newblock {\em Neuron} {\bf 1997}, {\em 19},~293--306.
\newblock {{https://doi.org/10.1016/S0896-6273(00)80940-X}}.

\bibitem[Bienenstock(1995)]{bienenstock_model_1995}
Bienenstock, E.
\newblock A model of neocortex.
\newblock {\em Netw. Comput. Neural Syst.} {\bf 1995}, {\em
  6},~179--224.
%\newblock Publisher: Taylor \& Francis \_eprint:
%  https://doi.org/10.1088/0954-898X\_6\_2\_004,
  {{https://doi.org/10.1088/0954-898X\_6\_2\_004}}.

\bibitem[Muller {et~al.}(2014)Muller, Reynaud, Chavane, and
  Destexhe]{muller_stimulus-evoked_2014}
Muller, L.; Reynaud, A.; Chavane, F.; Destexhe, A.
\newblock The stimulus-evoked population response in visual cortex of awake
  monkey is a propagating wave.
\newblock {\em Nat. Commun.} {\bf 2014}, {\em 5},~3675.
%\newblock 00068 Loaded from an external bibliography file by Manubot.
%  source\_bibliography: manual-references.bib standard\_id: Muller14,
  {{https://doi.org/10.1038/ncomms4675}}.

\bibitem[Muller {et~al.}(2018)Muller, Chavane, Reynolds, and
  Sejnowski]{muller_cortical_2018}
Muller, L.; Chavane, F.; Reynolds, J.; Sejnowski, T.J.
\newblock Cortical travelling waves: {Mechanisms} and computational principles.
\newblock {\em Nat. Rev. Neurosci.} {\bf 2018}, \emph{19}, 255--268.
%\newblock Loaded from an external bibliography file by Manubot.
%  source\_bibliography: manual-references.bib standard\_id: Muller18,
  {{https://doi.org/10.1038/nrn.2018.20}}.

\bibitem[Lindén {et~al.}(2022)Lindén, Petersen, Vestergaard, and
  Berg]{linden_movement_2022}
Lindén, H.; Petersen, P.C.; Vestergaard, M.; Berg, R.W.
\newblock Movement is governed by rotational neural dynamics in spinal motor
  networks.
\newblock {\em Nature} {\bf 2022}, {\em 610},~526--531.
\newblock {{https://doi.org/10.1038/s41586-022-05293-w}}.

\bibitem[Chemla {et~al.}(2019)Chemla, Reynaud, diVolo, Zerlaut, Perrinet,
  Destexhe, and Chavane]{chemla_suppressive_2019}
Chemla, S.; Reynaud, A.; diVolo, M.; Zerlaut, Y.; Perrinet, L.U.; Destexhe, A.;
  Chavane, F.Y.
\newblock Suppressive waves disambiguate the representation of long-range
  apparent motion in awake monkey {V1}.
\newblock {\em J. Neurosci.} {\bf 2019}, {\em 2792},~18.
%\newblock Loaded from an external bibliography file by Manubot.
%  source\_bibliography: manual-references.bib standard\_id: Chemla19,
  {{https://doi.org/10.1523/JNEUROSCI.2792-18.2019}}.

\bibitem[Pillow {et~al.}(2008)Pillow, Shlens, Paninski, Sher, Litke,
  Chichilnisky, and Simoncelli]{pillow_spatio-temporal_2008}
Pillow, J.W.; Shlens, J.; Paninski, L.; Sher, A.; Litke, A.M.; Chichilnisky,
  E.J.; Simoncelli, E.P.
\newblock Spatio-temporal correlations and visual signalling in a complete
  neuronal population.
\newblock {\em Nature} {\bf 2008}, {\em 454},~995--999.
\newblock {{https://doi.org/10.1038/nature07140}}.

\bibitem[Schneidman {et~al.}(2006)Schneidman, Berry, Segev, and
  Bialek]{schneidman_weak_2006}
Schneidman, E.; Berry, M.J.; Segev, R.; Bialek, W.
\newblock Weak pairwise correlations imply strongly correlated network states
  in a neural population.
\newblock {\em Nature} {\bf 2006}, {\em 440},~1007--1012.
\newblock {{https://doi.org/10.1038/nature04701}}.

\bibitem[Berry(2022)]{berry_spike_2022}
Puchalla, J.; Berry, M.J.
\newblock \emph{Spike {Trains} of {Retinal} {Ganglion} {Cells} {Viewing} a {Repeated}
  {Natural} {Movie}}; \hl{Princeton University: Princeton, NJ, USA,} %MDPI: We added the name of the publisher and location. Please confirm.
  {2022}.
\newblock {{https://doi.org/10.34770/V0V4-3H52}}.

\bibitem[Miller {et~al.}(2014)Miller, Ayzenshtat, Carrillo-Reid, and
  Yuste]{miller_visual_2014}
Miller, J.e.K.; Ayzenshtat, I.; Carrillo-Reid, L.; Yuste, R.
\newblock Visual stimuli recruit intrinsically generated cortical ensembles.
\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2014},
  {\em 111}, E4053--E4061.
\newblock {{https://doi.org/10.1073/pnas.1406077111}}.

\bibitem[Ikegaya {et~al.}(2004)Ikegaya, Aaron, Cossart, Aronov, Lampl,
  Ferster, and Yuste]{ikegaya_synfire_2004}
Ikegaya, Y.; Aaron, G.; Cossart, R.; Aronov, D.; Lampl, I.; Ferster, D.; Yuste,
  R.
\newblock Synfire {Chains} and {Cortical} {Songs}: {Temporal} {Modules} of
  {Cortical} {Activity}.
\newblock {\em Science} {\bf 2004}, {\em 304},~559--564.
\newblock {{https://doi.org/10.1126/science.1093173}}.

\bibitem[Luczak {et~al.}(2007)Luczak, Barthó, Marguet, Buzsáki, and
  Harris]{luczak_sequential_2007}
Luczak, A.; Barthó, P.; Marguet, S.L.; Buzsáki, G.; Harris, K.D.
\newblock Sequential structure of neocortical spontaneous activity in vivo.
\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2007},
  {\em 104},~347--352.
\newblock {{https://doi.org/10.1073/pnas.0605643104}}.

\bibitem[Pastalkova {et~al.}(2008)Pastalkova, Itskov, Amarasingham, and
  Buzsáki]{pastalkova_internally_2008}
Pastalkova, E.; Itskov, V.; Amarasingham, A.; Buzsáki, G.
\newblock Internally {Generated} {Cell} {Assembly} {Sequences} in the {Rat}
  {Hippocampus}.
\newblock {\em Science} {\bf 2008}, {\em 321},~1322--1327.
\newblock {{https://doi.org/10.1126/science.1159775}}.

\bibitem[Villette {et~al.}(2015)Villette, Malvache, Tressard, Dupuy, and
  Cossart]{villette_internally_2015}
Villette, V.; Malvache, A.; Tressard, T.; Dupuy, N.; Cossart, R.
\newblock Internally {Recurring} {Hippocampal} {Sequences} as a {Population}
  {Template} of {Spatiotemporal} {Information}.
\newblock {\em Neuron} {\bf 2015}, {\em 88},~357--366.
 {{https://doi.org/10.1016/j.neuron.2015.09.052}}.

\bibitem[Branco {et~al.}(2010)Branco, Clark, and
  Häusser]{branco_dendritic_2010}
Branco, T.; Clark, B.A.; Häusser, M.
\newblock Dendritic {Discrimination} of {Temporal} {Input} {Sequences} in
  {Cortical} {Neurons}.
\newblock {\em Science} {\bf 2010}, {\em 329},~1671--1675.
\newblock {{https://doi.org/10.1126/science.1189664}}.

\bibitem[Luczak {et~al.}(2015)Luczak, McNaughton, and
  Harris]{luczak_packet-based_2015}
Luczak, A.; McNaughton, B.L.; Harris, K.D.
\newblock Packet-based communication in the cortex.
\newblock {\em Nat. Rev. Neurosci.} {\bf 2015}, {\em 16},~745--755.
\newblock {{https://doi.org/10.1038/nrn4026}}.

\bibitem[Meister {et~al.}(1995)Meister, Lagnado, and
  Baylor]{meister_concerted_1995}
Meister, M.; Lagnado, L.; Baylor, D.A.
\newblock Concerted {Signaling} by {Retinal} {Ganglion} {Cells}.
\newblock {\em Science} {\bf 1995}, {\em 270},~1207--1210.
\newblock {{https://doi.org/10.1126/science.270.5239.1207}}.

\bibitem[Cleland(2014)]{cleland_construction_2014}
Cleland, T.A.
\newblock Construction of {Odor} {Representations} by {Olfactory} {Bulb}
  {Microcircuits}. In {\em Progress in {Brain} {Research}}; Elsevier:  \hl{Amsterdam, The Netherlands,} %newly added information, please confirm
  2014;
  Volume 208, pp. 177--203.
\newblock {{https://doi.org/10.1016/B978-0-444-63350-7.00007-3}}.

\bibitem[Kashiwadani {et~al.}(1999)Kashiwadani, Sasaki, Uchida, and
  Mori]{kashiwadani_synchronized_1999}
Kashiwadani, H.; Sasaki, Y.F.; Uchida, N.; Mori, K.
\newblock Synchronized {Oscillatory} {Discharges} of {Mitral}/{Tufted} {Cells}
  {With} {Different} {Molecular} {Receptive} {Ranges} in the {Rabbit}
  {Olfactory} {Bulb}.
\newblock {\em J. Neurophysiol.} {\bf 1999}, {\em 82},~1786--1792.\linebreak
%\newblock Publisher: American Physiological Society,
  {{https://doi.org/10.1152/jn.1999.82.4.1786}}.

\bibitem[Rinberg {et~al.}(2006)Rinberg, Koulakov, and
  Gelperin]{rinberg_speed-accuracy_2006}
Rinberg, D.; Koulakov, A.; Gelperin, A.
\newblock Speed-{Accuracy} {Tradeoff} in {Olfaction}.
\newblock {\em Neuron} {\bf 2006}, {\em 51},~351--358.\linebreak
\newblock {{https://doi.org/10.1016/j.neuron.2006.07.013}}.

\bibitem[Johansson and Birznieks(2004)]{johansson_first_2004}
Johansson, R.S.; Birznieks, I.
\newblock First spikes in ensembles of human tactile afferents code complex
  spatial fingertip events.
\newblock {\em Nat. Neurosci.} {\bf 2004}, {\em 7},~170--177.
\newblock {{https://doi.org/10.1038/nn1177}}.

\bibitem[Buzsáki and Tingley(2018)]{buzsaki_space_2018}
Buzsáki, G.; Tingley, D.
\newblock Space and {Time}: {The} {Hippocampus} as a {Sequence} {Generator}.
\newblock {\em Trends Cogn. Sci.} {\bf 2018}, {\em 22},~853--869.
\newblock {{https://doi.org/10.1016/j.tics.2018.07.006}}.

\bibitem[Malvache {et~al.}(2016)Malvache, Reichinnek, Villette, Haimerl, and
  Cossart]{malvache_awake_2016}
Malvache, A.; Reichinnek, S.; Villette, V.; Haimerl, C.; Cossart, R.
\newblock Awake hippocampal reactivations project onto orthogonal neuronal
  assemblies.
\newblock {\em Science} {\bf 2016}, {\em 353},~1280--1283.
 {{https://doi.org/10.1126/science.aaf3319}}.

\bibitem[Haimerl {et~al.}(2019)Haimerl, Angulo-Garcia, Villette, Reichinnek,
  Torcini, Cossart, and Malvache]{haimerl_internal_2019}
Haimerl, C.; Angulo-Garcia, D.; Villette, V.; Reichinnek, S.; Torcini, A.;
  Cossart, R.; Malvache, A.
\newblock Internal representation of hippocampal neuronal population spans a
  time-distance continuum.
\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2019},
  {\em 116},~7477--7482.
\newblock {{https://doi.org/10.1073/pnas.1718518116}}.

\bibitem[Agus {et~al.}(2010)Agus, Thorpe, and Pressnitzer]{agus_rapid_2010}
Agus, T.R.; Thorpe, S.J.; Pressnitzer, D.
\newblock Rapid {Formation} of {Robust} {Auditory} {Memories}: {Insights} from
  {Noise}.
\newblock {\em Neuron} {\bf 2010}, {\em 66},~610--618.
\newblock {{https://doi.org/10.1016/j.neuron.2010.04.014}}.

\bibitem[Izhikevich(2006)]{izhikevich_polychronization_2006}
Izhikevich, E.M.
\newblock Polychronization: {Computation} with {Spikes}.
\newblock {\em Neural Comput.} {\bf 2006}, {\em 18},~245--282.\linebreak
 {{https://doi.org/10.1162/089976606775093882}}.

\bibitem[Simoncelli {et~al.}(2003)Simoncelli, Paninski, Pillow, and
  Schwartz]{simoncelli_characterization_2003}
Simoncelli, E.P.; Paninski, L.; Pillow, J.; Schwartz, O.
\newblock Characterization of {Neural} {Responses} with {Stochastic} {Stimuli}. In \emph{\hl{The New Cognitive Neurosciences}}, \hl{3rd ed.; Gazzaniga, M., Ed.; MIT Press: Cambridge, MA, USA,} %MDPI: Newly added  information. Please confirm.
  {2003}.

\bibitem[Jazayeri and Movshon(2006)]{jazayeri_optimal_2006}
Jazayeri, M.; Movshon, J.A.
\newblock Optimal representation of sensory information by neural populations.
\newblock {\em Nat. Neurosci.} {\bf 2006}, {\em 9},~690--696.
%\newblock Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/nn1691}}.

\bibitem[Berens {et~al.}(2012)Berens, Ecker, Cotton, Ma, Bethge, and
  Tolias]{berens_fast_2012}
Berens, P.; Ecker, A.S.; Cotton, R.J.; Ma, W.J.; Bethge, M.; Tolias, A.S.
\newblock A {Fast} and {Simple} {Population} {Code} for {Orientation} in
  {Primate} {V1}.
\newblock {\em J. Neurosci.} {\bf 2012}, {\em 32},~10618--10626.
\newblock {{https://doi.org/10.1523/jneurosci.1335-12.2012}}.

\bibitem[Bellec {et~al.}(2021)Bellec, Wang, Modirshanechi, Brea, and
  Gerstner]{bellec_fitting_2021}
Bellec, G.; Wang, S.; Modirshanechi, A.; Brea, J.; Gerstner, W.
\newblock Fitting summary statistics of neural data with a differentiable
  spiking network simulator. \emph{arXiv} \textbf{2021},  	arXiv:2106.10064.
%\newblock Technical Report 2106.10064, arXiv,  .

\bibitem[Kohn and Smith(2016)]{kohn_utah_2016}
Kohn, A.; Smith, M.
\newblock Utah Array Extracellular Recordings of Spontaneous and Visually
  Evoked Activity from Anesthetized Macaque Primary Visual Cortex ({V1}). 
  2016.   
Available online: {\url{https://doi.org/10.6080/K0NC5Z4X}}  \hl{(accessed on date month year).} %MDPI: Please add the access date (format: Date Month Year), e.g., accessed on 1 January 2020.

\bibitem[Warner {et~al.}(2022)Warner, Ruda, and
  Sommer]{warner_probabilistic_2022}
Warner, C.; Ruda, K.; Sommer, F.T.
\newblock A probabilistic latent variable model for detecting structure in
  binary data. \emph{arXiv} \textbf{2022}, arXiv:2201.11108.
%\newblock {\em arXiv:2201.11108 [cs, q-bio, stat]} {\bf }.
%\newblock arXiv: 2201.11108.

\bibitem[Victor and Purpura(1996)]{victor_nature_1996}
Victor, J.D.; Purpura, K.P.
\newblock Nature and precision of temporal coding in visual cortex: A
  metric-space analysis.
\newblock {\em J. Neurophysiol.} {\bf 1996}, {\em 76},~1310--1326.
\newblock {{https://doi.org/10.1152/jn.1996.76.2.1310}}.

\bibitem[van Rossum(2001)]{van_rossum_novel_2001}
van Rossum, M.C.
\newblock A novel spike distance.
\newblock {\em Neural Comput.} {\bf 2001}, {\em 13},~751--763.
\newblock {{https://doi.org/10.1162/089976601300014321}}.

\bibitem[Kreuz {et~al.}(2007)Kreuz, Haas, Morelli, Abarbanel, and
  Politi]{kreuz_measuring_2007}
Kreuz, T.; Haas, J.S.; Morelli, A.; Abarbanel, H.D.I.; Politi, A.
\newblock Measuring spike train synchrony.
\newblock {\em J. Neurosci. Methods} {\bf 2007}, {\em
  165},~151--161.
\newblock {{https://doi.org/10.1016/j.jneumeth.2007.05.031}}.

\bibitem[Moser and Natschlager(2014)]{moser_stability_2014}
Moser, B.A.; Natschlager, T.
\newblock On {Stability} of {Distance} {Measures} for {Event} {Sequences}
  {Induced} by {Level}-{Crossing} {Sampling}.
\newblock {\em IEEE Trans. Signal Process.} {\bf 2014}, {\em
  62},~1987--1999.
\newblock {{https://doi.org/10.1109/tsp.2014.2305642}}.

\bibitem[Weyl(1916)]{weyl_ueber_1916}
Weyl, H.
\newblock Ueber die {Gleichverteilung} von {Zahlen} mod. {Eins}.
\newblock {\em Math. Ann.} {\bf 1916}, {\em 77},~313--352.
\newblock {{https://doi.org/10.1007/bf01475864}}.

\bibitem[Aronov and Victor(2004)]{aronov_non-euclidean_2004}
Aronov, D.; Victor, J.D.
\newblock Non-{Euclidean} properties of spike train metric spaces.
\newblock {\em Phys. Rev. E} {\bf 2004}, {\em 69},~061905.
%\newblock Publisher: American Physical Society,
  {{https://doi.org/10.1103/PhysRevE.69.061905}}.

\bibitem[Levakova {et~al.}(2015)Levakova, Tamborrino, Ditlevsen, and
  Lansky]{levakova_review_2015}
Levakova, M.; Tamborrino, M.; Ditlevsen, S.; Lansky, P.
\newblock A review of the methods for neuronal response latency estimation.
\newblock {\em Biosystems} {\bf 2015}, {\em 136},~23--34.
\newblock {{https://doi.org/10.1016/j.biosystems.2015.04.008}}.

\bibitem[Grün {et~al.}(2002)Grün, Diesmann, and
  Aertsen]{grun_unitary_2002-1}
Grün, S.; Diesmann, M.; Aertsen, A.
\newblock Unitary {Events} in {Multiple} {Single}-{Neuron} {Spiking}
  {Activity}: {II}. {Nonstationary} {Data}.
\newblock {\em Neural Comput.} {\bf 2002}, {\em 14},~81--119.
\newblock {{https://doi.org/10.1162/089976602753284464}}.

\bibitem[Grün {et~al.}(2010)Grün, Diesmann, and
  Aertsen]{grun_unitary_2010}
Grün, S.; Diesmann, M.; Aertsen, A.
\newblock Unitary {Event} {Analysis}. In {\em Analysis of {Parallel} {Spike}
  {Trains}}; Springer: \hl{Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
  2010; pp. 191--220.

\bibitem[Grün {et~al.}(2002)Grün, Diesmann, and
  Aertsen]{grun_unitary_2002}
Grün, S.; Diesmann, M.; Aertsen, A.
\newblock Unitary {Events} in {Multiple} {Single}-{Neuron} {Spiking}
  {Activity}: {I}. {Detection} and {Significance}.
\newblock {\em Neural Comput.} {\bf 2002}, {\em 14},~43--80.
\newblock {{https://doi.org/10.1162/089976602753284455}}.

\bibitem[Torre {et~al.}(2013)Torre, Picado-Muiño, Denker, Borgelt, and
  Grün]{torre_statistical_2013}
Torre, E.; Picado-Muiño, D.; Denker, M.; Borgelt, C.; Grün, S.
\newblock Statistical evaluation of synchronous spike patterns extracted by
  frequent item set mining.
\newblock {\em Front. Comput. Neurosci.} {\bf 2013}, {\em 7}, 132.
\newblock {{https://doi.org/10.3389/fncom.2013.00132}}.

\bibitem[Quaglio {et~al.}(2018)Quaglio, Rostami, Torre, and
  Grün]{quaglio_methods_2018}
Quaglio, P.; Rostami, V.; Torre, E.; Grün, S.
\newblock Methods for identification of spike patterns in massively parallel
  spike trains.
\newblock {\em Biol. Cybern.} {\bf 2018}, {\em 112},~57--80.
\newblock {{https://doi.org/10.1007/s00422-018-0755-0}}.

\bibitem[Stella {et~al.}(2019)Stella, Quaglio, Torre, and
  Grün]{stella_3d-spade_2019}
Stella, A.; Quaglio, P.; Torre, E.; Grün, S.
\newblock 3d-{SPADE}: {Significance} evaluation of spatio-temporal patterns of
  various temporal extents.
\newblock {\em Biosystems} {\bf 2019}, {\em 185},~104022.
\newblock {{https://doi.org/10.1016/j.biosystems.2019.104022}}.

\bibitem[Stella {et~al.}(2022)Stella, Bouss, Palm, and
  Grün]{stella_comparing_2022}
Stella, A.; Bouss, P.; Palm, G.; Grün, S.
\newblock Comparing {Surrogates} to {Evaluate} {Precisely} {Timed}
  {Higher}-{Order} {Spike} {Correlations}.
\newblock {\em eNeuro} {\bf 2022}, {\em 9},~ENEURO.0505--21.2022.
\newblock {{https://doi.org/10.1523/eneuro.0505-21.2022}}.

\bibitem[Grossberger {et~al.}(2018)Grossberger, Battaglia, and
  Vinck]{grossberger_unsupervised_2018}
Grossberger, L.; Battaglia, F.P.; Vinck, M.
\newblock Unsupervised clustering of temporal patterns in high-dimensional
  neuronal ensembles using a novel dissimilarity measure.
\newblock {\em PLoS Comput. Biol.} {\bf 2018}, {\em 14},~e1006283.
\newblock {{https://doi.org/10.1371/journal.pcbi.1006283}}.

\bibitem[Nádasdy {et~al.}(1999)Nádasdy, Hirase, Czurkó, Csicsvari, and
  Buzsáki]{nadasdy_replay_1999}
Nádasdy, Z.; Hirase, H.; Czurkó, A.; Csicsvari, J.; Buzsáki, G.
\newblock Replay and {Time} {Compression} of {Recurring} {Spike} {Sequences} in
  the {Hippocampus}.
\newblock {\em J. Neurosci.} {\bf 1999}, {\em 19},~9497--9507.
%\newblock Publisher: Society for Neuroscience Section: ARTICLE,
  {{https://doi.org/10.1523/JNEUROSCI.19-21-09497.1999}}.

\bibitem[Lee and Wilson(2004)]{lee_combinatorial_2004}
Lee, A.K.; Wilson, M.A.
\newblock A {Combinatorial} {Method} for {Analyzing} {Sequential} {Firing}
  {Patterns} {Involving} an {Arbitrary} {Number} of {Neurons} {Based} on
  {Relative} {Time} {Order}.
\newblock {\em J. Neurophysiol.} {\bf 2004}, {\em 92},~2555--2573.
\newblock {{https://doi.org/10.1152/jn.01030.2003}}.

\bibitem[Sotomayor-Gómez {et~al.}(2021)Sotomayor-Gómez, Battaglia, and
  Vinck]{sotomayor-gomez_spikeship_2021}
Sotomayor-Gómez, B.; Battaglia, F.P.; Vinck, M.
\newblock {SpikeShip}: {A} method for fast, unsupervised discovery of
  high-dimensional neural spiking patterns.
\newblock {\em bioRxiv  Prepr. Serv. Biol.} {\bf 2021}.
%\newblock Publisher: Cold Spring Harbor Laboratory,
  {{https://doi.org/10.1101/2020.06.03.131573}}.

\bibitem[Pachitariu {et~al.}(2018)Pachitariu, Stringer, and
  Harris]{pachitariu_robustness_2018}
Pachitariu, M.; Stringer, C.; Harris, K.D.
\newblock Robustness of {Spike} {Deconvolution} for {Neuronal} {Calcium}
  {Imaging}.
\newblock {\em  J. Neurosci.} {\bf 2018}, {\em 38},~7976--7985.
\newblock {{https://doi.org/10.1523/jneurosci.3339-17.2018}}.

\bibitem[Stringer(2020)]{stringer_mouselandrastermap_2020}
Stringer, C.
\newblock {MouseLand}/Rastermap: {A} Multi-Dimensional Embedding Algorithm.
  \hl{2020}. %MDPI: Please add website and access date. Please check.

\bibitem[Stringer {et~al.}(2019)Stringer, Pachitariu, Steinmetz, Reddy,
  Carandini, and Harris]{stringer_spontaneous_2019}
Stringer, C.; Pachitariu, M.; Steinmetz, N.; Reddy, C.B.; Carandini, M.;
  Harris, K.D.
\newblock Spontaneous behaviors drive multidimensional, brainwide activity.
\newblock {\em Science} {\bf 2019}, {\em 364}, 255.
\newblock {{https://doi.org/10.1126/science.aav7893}}.

\bibitem[Stringer {et~al.}(2021)Stringer, Michaelos, Tsyboulski, Lindo, and
  Pachitariu]{stringer_high-precision_2021}
Stringer, C.; Michaelos, M.; Tsyboulski, D.; Lindo, S.E.; Pachitariu, M.
\newblock High-precision coding in visual cortex.
\newblock {\em Cell} {\bf 2021}, {\em 184},~2767--2778.e15.
\newblock {{https://doi.org/10.1016/j.cell.2021.03.042}}.

\bibitem[Russo and Durstewitz(2017)]{russo_cell_2017}
Russo, E.; Durstewitz, D.
\newblock Cell assemblies at multiple time scales with arbitrary lag
  constellations.
\newblock {\em eLife} {\bf 2017}, {\em 6}, e19428.
\newblock {{https://doi.org/10.7554/elife.19428}}.

\bibitem[Pipa {et~al.}(2008)Pipa, Wheeler, Singer, and
  Nikolić]{pipa_neuroxidence_2008}
Pipa, G.; Wheeler, D.W.; Singer, W.; Nikolić, D.
\newblock {NeuroXidence}: Reliable and efficient analysis of an excess or
  deficiency of joint-spike events.
\newblock {\em J. Comput. Neurosci.} {\bf 2008}, {\em
  25},~64--88.
\newblock {{https://doi.org/10.1007/s10827-007-0065-3}}.

\bibitem[Torre {et~al.}(2016)Torre, Quaglio, Denker, Brochier, Riehle, and
  Grun]{torre_synchronous_2016}
Torre, E.; Quaglio, P.; Denker, M.; Brochier, T.; Riehle, A.; Grun, S.
\newblock Synchronous {Spike} {Patterns} in {Macaque} {Motor} {Cortex} during
  an {Instructed}-{Delay} {Reach}-to-{Grasp} {Task}.
\newblock {\em J. Neurosci.} {\bf 2016}, {\em 36},~8329--8340.
\newblock {{https://doi.org/10.1523/jneurosci.4375-15.2016}}.

\bibitem[Williams {et~al.}(2020)Williams, Degleris, Wang, and
  Linderman]{williams_point_2020}
Williams, A.H.; Degleris, A.; Wang, Y.; Linderman, S.W.
\newblock Point process models for sequence detection in high-dimensional
  neural spike trains. \emph{arXiv}  \textbf{2020},  	arXiv:2010.04875
%\newblock Technical Report 2010.04875, .

\bibitem[Kass {et~al.}(2005)Kass, Ventura, and Brown]{kass_statistical_2005}
Kass, R.E.; Ventura, V.; Brown, E.N.
\newblock Statistical issues in the analysis of neuronal data.
\newblock {\em J. Neurophysiol.} {\bf 2005}, {\em 94},~8--25.
%\newblock Publisher: American Physiological Society,
  {{https://doi.org/10.1152/jn.00648.2004}}.

\bibitem[van Kempen {et~al.}(2021)van Kempen, Gieselmann, Boyd, Steinmetz,
  Moore, Engel, and Thiele]{van_kempen_top-down_2021}
van Kempen, J.; Gieselmann, M.A.; Boyd, M.; Steinmetz, N.A.; Moore, T.; Engel,
  T.A.; Thiele, A.
\newblock Top-down coordination of local cortical state during selective
  attention.
\newblock {\em Neuron} {\bf 2021}, {\em 109},~894--904.e8.
\newblock {{https://doi.org/10.1016/j.neuron.2020.12.013}}.

\bibitem[Pasturel {et~al.}(2020)Pasturel, Montagnini, and
  Perrinet]{pasturel_humans_2020}
Pasturel, C.; Montagnini, A.; Perrinet, L.U.
\newblock Humans adapt their anticipatory eye movements to the volatility of
  visual motion properties.
\newblock {\em PLoS Comput. Biol.} {\bf 2020}, \emph{16}, e1007438.
\newblock {{https://doi.org/10.1371/journal.pcbi.1007438}}.

\bibitem[Von~Helmholz(1850)]{von_helmholz_messungen_1850}
Von~Helmholz, H.
\newblock Messungen über den zeitlichen {Verlauf} der {Zuckung} animalischer
  {Muskeln} und die {Fortpflanzungsgeschwindigkeit} der {Reizung} in den
  {Nerven}.
\newblock {\em Arch. Anat. Physiol. Wiss. Med.}
  {\bf 1850}, {\em 17},~176--364.

\bibitem[Peyrard(2020)]{peyrard_how_2020}
Peyrard, M.
\newblock How is information transmitted in a nerve?
\newblock {\em J. Biol. Phys.} {\bf 2020}, {\em 46},~327--341.
\newblock {{https://doi.org/10.1007/s10867-020-09557-2}}.

\bibitem[Young(1938)]{young_functioning_1938}
Young, J.Z.
\newblock The {Functioning} of the {Giant} {Nerve} {Fibres} of the {Squid}.
\newblock {\em J. Exp. Biol.} {\bf 1938}, {\em 15},~170--185.\linebreak
\newblock {{https://doi.org/10.1242/jeb.15.2.170}}.

\bibitem[Madadi~Asl {et~al.}(2018)Madadi~Asl, Valizadeh, and
  Tass]{madadi_asl_dendritic_2018}
Madadi~Asl, M.; Valizadeh, A.; Tass, P.A.
\newblock Dendritic and {Axonal} {Propagation} {Delays} {May} {Shape}
  {Neuronal} {Networks} {With} {Plastic} {Synapses}.
\newblock {\em Front. Physiol.} {\bf 2018}, {\em 9},~1849.
\newblock {{https://doi.org/10.3389/fphys.2018.01849}}.

\bibitem[Stetson {et~al.}(1992)Stetson, Albers, Silverstein, and
  Wolfe]{stetson_effects_1992}
Stetson, D.S.; Albers, J.W.; Silverstein, B.A.; Wolfe, R.A.
\newblock Effects of age, sex, and anthropometric factors on nerve conduction
  measures.
\newblock {\em Muscle Nerve} {\bf 1992}, {\em 15},~1095--1104.
%\newblock \_eprint:
%  https://onlinelibrary.wiley.com/doi/pdf/10.1002/mus.880151007,
  {{https://doi.org/10.1002/mus.880151007}}.

\bibitem[Jeffress(1948)]{jeffress_place_1948}
Jeffress, L.A.
\newblock A place theory of sound localization.
\newblock {\em J. Comp. Physiol. Psychol.} {\bf 1948},
  {\em 41},~35--39.
%\newblock Place: US Publisher: American Psychological Association,
  {{https://doi.org/10.1037/h0061495}}.

\bibitem[Konishi(2003)]{konishi_coding_2003}
Konishi, M.
\newblock Coding of auditory space.
\newblock {\em Annu. Rev. Neurosci.} {\bf 2003}, {\em 26},~31--55.\linebreak
\newblock {{https://doi.org/10.1146/annurev.neuro.26.041002.131123}}.

\bibitem[Gerstner {et~al.}(1996)Gerstner, Kempter, van Hemmen, and
  Wagner]{gerstner_neuronal_1996}
Gerstner, W.; Kempter, R.; van Hemmen, J.L.; Wagner, H.
\newblock A neuronal learning rule for sub-millisecond temporal coding.
\newblock {\em Nature} {\bf 1996}, {\em 383},~76--78.
%\newblock Number: 6595 Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/383076a0}}.

\bibitem[Seidl {et~al.}(2010)Seidl, Rubel, and
  Harris]{seidl_mechanisms_2010}
Seidl, A.H.; Rubel, E.W.; Harris, D.M.
\newblock Mechanisms for adjusting interaural time differences to achieve
  binaural coincidence detection.
\newblock {\em  J. Neurosci. Off. J. Soc. Neurosci.} {\bf 2010}, {\em 30},~70--80.
\newblock {{https://doi.org/10.1523/JNEUROSCI.3464-09.2010}}.

\bibitem[Camon {et~al.}(2019)Camon, Hugues, Erlandson, Robbe, Lagoun,
  Marouane, and Bureau]{camon_timing_2019}
Camon, J.; Hugues, S.; Erlandson, M.A.; Robbe, D.; Lagoun, S.; Marouane, E.;
  Bureau, I.
\newblock The {Timing} of {Sensory}-{Guided} {Behavioral} {Response} is
  {Represented} in the {Mouse} {Primary} {Somatosensory} {Cortex}.
\newblock {\em Cereb. Cortex} {\bf 2019}, {\em 29},~3034--3047.
\newblock {{https://doi.org/10.1093/cercor/bhy169}}.

\bibitem[Gasser and Grundfest(1939)]{gasser_axon_1939}
Gasser, H.S.; Grundfest, H.
\newblock {AXON} {DIAMETERS} {IN} {RELATION} {TO} {THE} {SPIKE} {DIMENSIONS}
  {AND} {THE} {CONDUCTION} {VELOCITY} {IN} {MAMMALIAN} {A} {FIBERS}.
\newblock {\em Am. J.-Physiol.-Leg. Content} {\bf 1939}. %MDPI: Please add the volume and page number.
%\newblock Publisher: American Physiological Society,
  {{https://doi.org/10.1152/ajplegacy.1939.127.2.393}}.

\bibitem[Brill {et~al.}(1977)Brill, Waxman, Moore, and
  Joyner]{brill_conduction_1977}
Brill, M.H.; Waxman, S.G.; Moore, J.W.; Joyner, R.W.
\newblock Conduction velocity and spike configuration in myelinated fibres:
  computed dependence on internode distance.
\newblock {\em J. Neurol. Neurosurg. Psychiatry} {\bf 1977},
  {\em 40},~769--774.

\bibitem[Pérez-Cerdá {et~al.}(2015)Pérez-Cerdá, Sánchez-Gómez, and
  Matute]{perez-cerda_pio_2015}
Pérez-Cerdá, F.; Sánchez-Gómez, M.V.; Matute, C.
\newblock Pío del {Río} {Hortega} and the discovery of the oligodendrocytes.
\newblock {\em Front. Neuroanat.} {\bf 2015}, {\em 9},~92.
\newblock {{https://doi.org/10.3389/fnana.2015.00092}}.

\bibitem[Schmitt and Bear(1939)]{schmitt_ultrastructure_1939}
Schmitt, F.O.; Bear, R.S.
\newblock The {Ultrastructure} of the {Nerve} {Axon} {Sheath}.
\newblock {\em Biol. Rev.} {\bf 1939}, {\em 14},~27--50.
%\newblock \_eprint:
%  https://onlinelibrary.wiley.com/doi/pdf/10.1111/j.1469-185X.1939.tb00922.x,
  {{https://doi.org/10.1111/j.1469-185X.1939.tb00922.x}}.

\bibitem[Simons and Nave(2016)]{simons_oligodendrocytes_2016}
Simons, M.; Nave, K.A.
\newblock Oligodendrocytes: {Myelination} and {Axonal} {Support}.
\newblock {\em Cold Spring Harb. Perspect. Biol.} {\bf 2016}, {\em
  8},~a020479.
\newblock {{https://doi.org/10.1101/cshperspect.a020479}}.

\bibitem[Duncan {et~al.}(2021)Duncan, Simkins, and
  Emery]{duncan_neuron-oligodendrocyte_2021}
Duncan, G.J.; Simkins, T.J.; Emery, B.
\newblock Neuron-{Oligodendrocyte} {Interactions} in the {Structure} and
  {Integrity} of {Axons}.
\newblock {\em Front. Cell Dev. Biol.} {\bf 2021}, {\em
  9}, 653101.
\newblock {{https://doi.org/10.3389/fcell.2021.653101}}.

\bibitem[Fields(2015)]{fields_new_2015}
Fields, R.D.
\newblock A new mechanism of nervous system plasticity: activity-dependent
  myelination.
\newblock {\em Nat. Rev. Neurosci.} {\bf 2015}, {\em 16},~756--767.
%\newblock Number: 12 Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/nrn4023}}.

\bibitem[Fields and Bukalo(2020)]{fields_myelin_2020}
Fields, R.D.; Bukalo, O.
\newblock Myelin makes memories.
\newblock {\em Nat. Neurosci.} {\bf 2020}, {\em 23},~469--470.
\newblock {{https://doi.org/10.1038/s41593-020-0606-x}}.

\bibitem[Reynolds and Slater(1928)]{reynolds_study_1928}
Reynolds, F.E.; Slater, J.K.
\newblock A {Study} of the {Structure} and {Function} of the {Interstitial}
  {Tissue} of the {Central} {Nervous} {System}.
\newblock {\em Edinb. Med. J.} {\bf 1928}, {\em 35},~49--57.

\bibitem[Steadman {et~al.}(2020)Steadman, Xia, Ahmed, Mocle, Penning,
  Geraghty, Steenland, Monje, Josselyn, and
  Frankland]{steadman_disruption_2020}
Steadman, P.E.; Xia, F.; Ahmed, M.; Mocle, A.J.; Penning, A.R.; Geraghty, A.C.;
  Steenland, H.W.; Monje, M.; Josselyn, S.A.; Frankland, P.W.
\newblock Disruption of {Oligodendrogenesis} {Impairs} {Memory} {Consolidation}
  in {Adult} {Mice}.
\newblock {\em Neuron} {\bf 2020}, {\em 105},~150--164.e6.
\newblock {{https://doi.org/10.1016/j.neuron.2019.10.013}}.

\bibitem[Pan {et~al.}(2020)Pan, Mayoral, Choi, Chan, and
  Kheirbek]{pan_preservation_2020}
Pan, S.; Mayoral, S.R.; Choi, H.S.; Chan, J.R.; Kheirbek, M.A.
\newblock Preservation of a remote fear memory requires new myelin formation.
\newblock {\em Nat. Neurosci.} {\bf 2020}, {\em 23},~487--499.
\newblock {{https://doi.org/10.1038/s41593-019-0582-1}}.

\bibitem[Wan {et~al.}(2020)Wan, Cheli, Santiago-González, Rosenblum, Wan,
  and Paez]{wan_impaired_2020}
Wan, R.; Cheli, V.T.; Santiago-González, D.A.; Rosenblum, S.L.; Wan, Q.; Paez,
  P.M.
\newblock Impaired {Postnatal} {Myelination} in a {Conditional} {Knockout}
  {Mouse} for the {Ferritin} {Heavy} {Chain} in {Oligodendroglial} {Cells}.
\newblock {\em  J. Neurosci.} {\bf 2020}, {\em 40},~7609--7624.
\newblock {{https://doi.org/10.1523/JNEUROSCI.1281-20.2020}}.

\bibitem[Xue {et~al.}(2021)Xue, Zhu, Liu, Lin, Li, Li, and
  Zhuo]{xue_demyelination_2021}
Xue, J.; Zhu, Y.; Liu, Z.; Lin, J.; Li, Y.; Li, Y.; Zhuo, Y.
\newblock Demyelination of the {Optic} {Nerve}: {An} {Underlying} {Factor} in
  {Glaucoma}?
\newblock {\em Front. Aging Neurosci.} {\bf 2021}, {\em 13},~701322.
\newblock {{https://doi.org/10.3389/fnagi.2021.701322}}.

\bibitem[Kuhn {et~al.}(2019)Kuhn, Gritti, Crooks, and
  Dombrowski]{kuhn_oligodendrocytes_2019}
Kuhn, S.; Gritti, L.; Crooks, D.; Dombrowski, Y.
\newblock Oligodendrocytes in {Development}, {Myelin} {Generation} and
  {Beyond}.
\newblock {\em Cells} {\bf 2019}, {\em 8},~1424.
\newblock {{https://doi.org/10.3390/cells8111424}}.

\bibitem[Baraban {et~al.}(2018)Baraban, Koudelka, and
  Lyons]{baraban_ca2_2018}
Baraban, M.; Koudelka, S.; Lyons, D.A.
\newblock Ca$^{2+}$ activity signatures of myelin sheath formation and growth in
  vivo.
\newblock {\em Nat. Neurosci.} {\bf 2018}, {\em 21},~19--23.
\newblock {{https://doi.org/10.1038/s41593-017-0040-x}}.

\bibitem[Nave and Salzer(2006)]{nave_axonal_2006}
Nave, K.A.; Salzer, J.L.
\newblock Axonal regulation of myelination by neuregulin 1.
\newblock {\em Curr. Opin. Neurobiol.} {\bf 2006}, {\em 16},~492--500.
\newblock {{https://doi.org/10.1016/j.conb.2006.08.008}}.

\bibitem[Cullen {et~al.}(2021)Cullen, Pepper, Clutterbuck, Pitman, Oorschot,
  Auderset, Tang, Ramm, Emery, Rodger, Jolivet, and
  Young]{cullen_periaxonal_2021}
Cullen, C.L.; Pepper, R.E.; Clutterbuck, M.T.; Pitman, K.A.; Oorschot, V.;
  Auderset, L.; Tang, A.D.; Ramm, G.; Emery, B.; Rodger, J.;  et~al.
\newblock Periaxonal and nodal plasticities modulate action potential
  conduction in the adult mouse brain.
\newblock {\em Cell Rep.} {\bf 2021}, {\em 34}, 108641.
%\newblock Publisher: Elsevier,
  {{https://doi.org/10.1016/j.celrep.2020.108641}}.

\bibitem[Gibson {et~al.}(2014)Gibson, Purger, Mount, Goldstein, Lin, Wood,
  Inema, Miller, Bieri, Zuchero, Barres, Woo, Vogel, and
  Monje]{gibson_neuronal_2014}
Gibson, E.M.; Purger, D.; Mount, C.W.; Goldstein, A.K.; Lin, G.L.; Wood, L.S.;
  Inema, I.; Miller, S.E.; Bieri, G.; Zuchero, J.B.;  et~al.
\newblock Neuronal {Activity} {Promotes} {Oligodendrogenesis} and {Adaptive}
  {Myelination} in the {Mammalian} {Brain}.
\newblock {\em Science} {\bf 2014}, {\em 344},~1252304.
\newblock {{https://doi.org/10.1126/science.1252304}}.

\bibitem[Spencer {et~al.}(2018)Spencer, Meffin, Burkitt, and
  Grayden]{spencer_compensation_2018}
Spencer, M.J.; Meffin, H.; Burkitt, A.N.; Grayden, D.B.
\newblock Compensation for {Traveling} {Wave} {Delay} {Through} {Selection} of
  {Dendritic} {Delays} {Using} {Spike}-{Timing}-{Dependent} {Plasticity} in a
  {Model} of the {Auditory} {Brainstem}.
\newblock {\em Front. Comput. Neurosci.} {\bf 2018}, {\em
  12},~36.
\newblock {{https://doi.org/10.3389/fncom.2018.00036}}.

\bibitem[Mel {et~al.}(2017)Mel, Schiller, and Poirazi]{mel_synaptic_2017}
Mel, B.W.; Schiller, J.; Poirazi, P.
\newblock Synaptic plasticity in dendrites: complications and coping
  strategies.
\newblock {\em Curr. Opin. Neurobiol.} {\bf 2017}, {\em 43},~177--186.
\newblock {{https://doi.org/10.1016/j.conb.2017.03.012}}.

\bibitem[Golding {et~al.}(2002)Golding, Staff, and
  Spruston]{golding_dendritic_2002}
Golding, N.L.; Staff, N.P.; Spruston, N.
\newblock Dendritic spikes as a mechanism for cooperative long-term
  potentiation.
\newblock {\em Nature} {\bf 2002}, {\em 418},~326--331.
%\newblock Number: 6895 Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/nature00854}}.

\bibitem[Maass(1997)]{maass_networks_1997}
Maass, W.
\newblock Networks of spiking neurons: {The} third generation of neural network
  models.
\newblock {\em Neural Netw.} {\bf 1997}, {\em 10},~1659--1671.

\bibitem[Neftci {et~al.}(2019)Neftci, Mostafa, and
  Zenke]{neftci_surrogate_2019}
Neftci, E.O.; Mostafa, H.; Zenke, F.
\newblock Surrogate {Gradient} {Learning} in {Spiking} {Neural} {Networks}:
  {Bringing} the {Power} of {Gradient}-{Based} {Optimization} to {Spiking}
  {Neural} {Networks}.
\newblock {\em IEEE Signal Process. Mag.} {\bf 2019}, {\em 36},~51--63.\linebreak
\newblock {{https://doi.org/10.1109/MSP.2019.2931595}}.

\bibitem[Rueckauer {et~al.}(2017)Rueckauer, Lungu, Hu, Pfeiffer, and
  Liu]{rueckauer_conversion_2017}
Rueckauer, B.; Lungu, I.A.; Hu, Y.; Pfeiffer, M.; Liu, S.C.
\newblock Conversion of {Continuous}-{Valued} {Deep} {Networks} to {Efficient}
  {Event}-{Driven} {Networks} for {Image} {Classification}.
\newblock {\em Front. Neurosci.} {\bf 2017}, {\em 11}.
\newblock {{https://doi.org/10.3389/fnins.2017.00682}}.

\bibitem[Susi {et~al.}(2021)Susi, Antón-Toro, Maestú, Pereda, and
  Mirasso]{susi_nmnsd-spiking_2021}
Susi, G.; Antón-Toro, L.F.; Maestú, F.; Pereda, E.; Mirasso, C.
\newblock {nMNSD}-{A} {Spiking} {Neuron}-{Based} {Classifier} {That} {Combines}
  {Weight}-{Adjustment} and {Delay}-{Shift}.
\newblock {\em Front. Neurosci.} {\bf 2021}, {\em 15},~582608.
\newblock {{https://doi.org/10.3389/fnins.2021.582608}}.

\bibitem[Davies {et~al.}(2018)Davies, Srinivasa, Lin, Chinya, Cao, Choday,
  Dimou, Joshi, Imam, Jain, Liao, Lin, Lines, Liu, Mathaikutty, McCoy, Paul,
  Tse, Venkataramanan, Weng, Wild, Yang, and Wang]{davies_loihi_2018}
Davies, M.; Srinivasa, N.; Lin, T.H.; Chinya, G.; Cao, Y.; Choday, S.H.; Dimou,
  G.; Joshi, P.; Imam, N.; Jain, S.;  et~al.
\newblock Loihi: {A} {Neuromorphic} {Manycore} {Processor} with {On}-{Chip}
  {Learning}.
\newblock {\em IEEE Micro} {\bf 2018}, {\em 38},~82--99.
\newblock {{https://doi.org/10.1109/MM.2018.112130359}}.

\bibitem[Lazar(2004)]{lazar_time_2004}
Lazar, A.A.
\newblock Time encoding with an integrate-and-fire neuron with a refractory
  period.
\newblock {\em Neurocomputing} {\bf 2004}, {\em 58--60},~53--58.
\newblock {{https://doi.org/10.1016/j.neucom.2004.01.022}}.

\bibitem[Markram {et~al.}(1997)Markram, Lübke, Frotscher, and
  Sakmann]{markram_regulation_1997}
Markram, H.; Lübke, J.; Frotscher, M.; Sakmann, B.
\newblock Regulation of {Synaptic} {Efficacy} by {Coincidence} of
  {Postsynaptic} {APs} and {EPSPs}.
\newblock {\em Science} {\bf 1997}, {\em 275},~213--215.
\newblock {{https://doi.org/10.1126/science.275.5297.213}}.

\bibitem[Caporale and Dan(2008)]{caporale_spike_2008}
Caporale, N.; Dan, Y.
\newblock Spike {Timing}–{Dependent} {Plasticity}: {A} {Hebbian} {Learning}
  {Rule}.
\newblock {\em Annu. Rev. Neurosci.} {\bf 2008}, {\em 31},~25--46.
\newblock {{https://doi.org/10.1146/annurev.neuro.31.060407.125639}}.

\bibitem[Hüning {et~al.}(1998)Hüning, Glünder, and
  Palm]{huning_synaptic_1998}
Hüning, H.; Glünder, H.; Palm, G.
\newblock Synaptic {Delay} {Learning} in {Pulse}-{Coupled} {Neurons}.
\newblock {\em Neural Comput.} {\bf 1998}, {\em 10},~555--565.
\newblock {{https://doi.org/10.1162/089976698300017665}}.

\bibitem[Eurich {et~al.}(1999)Eurich, Pawelzik, Ernst, Cowan, and
  Milton]{eurich_dynamics_1999}
Eurich, C.W.; Pawelzik, K.; Ernst, U.; Cowan, J.D.; Milton, J.G.
\newblock Dynamics of {Self}-{Organized} {Delay} {Adaptation}.
\newblock {\em Phys. Rev. Lett.} {\bf 1999}, {\em 82},~1594--1597.
%\newblock 00082 Publisher: American Physical Society,
  {{https://doi.org/10.1103/PhysRevLett.82.1594}}.

\bibitem[Gütig and Sompolinsky(2006)]{gutig_tempotron_2006}
Gütig, R.; Sompolinsky, H.
\newblock The tempotron: A neuron that learns spike timing–based decisions.
\newblock {\em Nat. Neurosci.} {\bf 2006}, {\em 9},~420--428.
\newblock {{https://doi.org/10.1038/nn1643}}.

\bibitem[Gütig(2014)]{gutig_spike_2014}
Gütig, R.
\newblock To spike, or when to spike?
\newblock {\em Curr. Opin. Neurobiol.} {\bf 2014}, {\em 25},~134--139.
\newblock {{https://doi.org/10.1016/j.conb.2014.01.004}}.

\bibitem[Pauli {et~al.}(2018)Pauli, Weidel, Kunkel, and
  Morrison]{pauli_reproducing_2018}
Pauli, R.; Weidel, P.; Kunkel, S.; Morrison, A.
\newblock Reproducing {Polychronization}: {A} {Guide} to {Maximizing} the
  {Reproducibility} of {Spiking} {Network} {Models}.
\newblock {\em Front. Neuroinformatics} {\bf 2018}, {\em 12},~46.
 {{https://doi.org/10.3389/fninf.2018.00046}}.

\bibitem[Guise {et~al.}(2014)Guise, Knott, and
  Benuskova]{guise_bayesian_2014}
Guise, M.; Knott, A.; Benuskova, L.
\newblock A {Bayesian} {Model} of {Polychronicity}.
\newblock {\em Neural Comput.} {\bf 2014}, {\em 26},~2052--2073.
\newblock {{https://doi.org/10.1162/neco\_a\_00620}}.

\bibitem[Zhang {et~al.}(2020)Zhang, Wu, Belatreche, Pan, Xie, Chua, Li, Qu,
  and Li]{zhang_supervised_2020}
Zhang, M.; Wu, J.; Belatreche, A.; Pan, Z.; Xie, X.; Chua, Y.; Li, G.; Qu, H.;
  Li, H.
\newblock Supervised learning in spiking neural networks with synaptic
  delay-weight plasticity.
\newblock {\em Neurocomputing} {\bf 2020}, {\em 409},~103--118.
\newblock {{https://doi.org/10.1016/j.neucom.2020.03.079}}.

\bibitem[Ghosh {et~al.}(2021)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
  Alfaro-Bittner, and Boccaletti]{ghosh_synchronization_2021}
Ghosh, D.; Frasca, M.; Rizzo, A.; Majhi, S.; Rakshit, S.; Alfaro-Bittner, K.;
  Boccaletti, S.
\newblock Synchronization in time-varying networks. \emph{arXiv} {\bf 2021}, arXiv:2109.07618
%\newblock {\em arXiv:2109.07618 [physics]} .
%\newblock 00000 arXiv: 2109.07618.

\bibitem[Ghosh {et~al.}(2022)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
  Alfaro-Bittner, and Boccaletti]{ghosh_synchronized_2022}
Ghosh, D.; Frasca, M.; Rizzo, A.; Majhi, S.; Rakshit, S.; Alfaro-Bittner, K.;
  Boccaletti, S.
\newblock The synchronized dynamics of time-varying networks.
\newblock {\em Phys. Rep.} {\bf 2022}, {\em 949},~1--63.
%\newblock arXiv:2109.07618 [physics],
  {{https://doi.org/10.1016/j.physrep.2021.10.006}}.

\bibitem[Izhikevich and Hoppensteadt(2009)]{izhikevich_polychronous_2009}
Izhikevich, E.M.; Hoppensteadt, F.C.
\newblock Polychronous {Wavefront} {Computations}.
\newblock {\em Int. J. Bifurc. Chaos} {\bf 2009}, {\em
  19},~1733--1739.
\newblock {{https://doi.org/10.1142/S0218127409023809}}.

\bibitem[Grimaldi and Perrinet(2022)]{grimaldi_learning_2022}
Grimaldi, A.; Perrinet, L.U.
\newblock Learning hetero-synaptic delays for motion detection in a single
  layer of spiking neurons.
\newblock In Proceedings of the 2022 {IEEE} {International} {Conference} on
  {Image} {Processing} ({ICIP}), \hl{Bordeaux, France, 16--19 October} %MDPI: We added the location and date of the conference. Please confirm.
 2022; pp.~3591--3595.
%\newblock ISSN: 2381-8549,
  {{https://doi.org/10.1109/ICIP46576.2022.9897394}}.

\bibitem[Madadi~Asl and
  Ramezani~Akbarabadi(2022)]{madadi_asl_delay-dependent_2022}
Madadi~Asl, M.; Ramezani~Akbarabadi, S.
\newblock Delay-dependent transitions of phase synchronization and coupling
  symmetry between neurons shaped by spike-timing-dependent plasticity.
\newblock {\em Cogn. Neurodynamics} {\bf \hl{2022} %MDPI: Please add the volume.
}, 1--14.
\newblock {{https://doi.org/10.1007/s11571-022-09850-x}}.

\bibitem[Perrinet and Samuelides(2002)]{perrinet_coherence_2002}
Perrinet, L.; Samuelides, M.
\newblock Coherence detection in a spiking neuron via {Hebbian} learning.
\newblock {\em Neurocomputing} {\bf 2002}, {\em 44-46},~133--139.
\newblock {{https://doi.org/10.1016/s0925-2312(02)00374-0}}.

\bibitem[Perrinet {et~al.}(2001)Perrinet, Delorme, Samuelides, and
  Thorpe]{perrinet_networks_2001}
Perrinet, L.; Delorme, A.; Samuelides, M.; Thorpe, S.
\newblock Networks of integrate-and-fire neuron using rank order coding {A}:
  {How} to implement spike time dependent {Hebbian} plasticity.
\newblock {\em Neurocomputing} {\bf 2001}, {\em 38--40},~817--822.
\newblock {{https://doi.org/10.1016/s0925-2312(01)00460-x}}.

\bibitem[Gilson(2010)]{gilson_stdp_2010}
Gilson, M.
\newblock {STDP} in recurrent neuronal networks.
\newblock {\em Front. Comput. Neurosci.} {\bf 2010}, {\em 4}, 23.
\newblock {{https://doi.org/10.3389/fncom.2010.00023}}.

\bibitem[Datadien {et~al.}(2011)Datadien, Haselager, and
  Sprinkhuizen-Kuyper]{datadien_right_2011}
Datadien, A.; Haselager, P.; Sprinkhuizen-Kuyper, I.
\newblock {\em The {Right} {Delay}---{Detecting} {Specific} {Spike} {Patterns}
  with {STDP} and {Axonal} {Conduction} {Delays}}; \hl{Springer:  Berlin/Heidelberg, Germany,} %MDPI: We added the name of the publisher and location. Please confirm.
 2011;
\newblock pp. 90--99.

\bibitem[Kerr {et~al.}(2013)Kerr, Burkitt, Thomas, Gilson, and
  Grayden]{kerr_delay_2013}
Kerr, R.R.; Burkitt, A.N.; Thomas, D.A.; Gilson, M.; Grayden, D.B.
\newblock Delay {Selection} by {Spike}-{Timing}-{Dependent} {Plasticity} in
  {Recurrent} {Networks} of {Spiking} {Neurons} {Receiving} {Oscillatory}
  {Inputs}.
\newblock {\em PLoS Comput. Biol.} {\bf 2013}, {\em 9},~e1002897.
\newblock {{https://doi.org/10.1371/journal.pcbi.1002897}}.

\bibitem[Burkitt and Hogendoorn(2021)]{burkitt_predictive_2021}
Burkitt, A.N.; Hogendoorn, H.
\newblock Predictive {Visual} {Motion} {Extrapolation} {Emerges}
  {Spontaneously} and without {Supervision} at {Each} {Layer} of a
  {Hierarchical} {Neural} {Network} with {Spike}-{Timing}-{Dependent}
  {Plasticity}.
\newblock {\em  J. Neurosci.} {\bf 2021}, {\em 41},~4428--4438.
\newblock {{https://doi.org/10.1523/jneurosci.2017-20.2021}}.

\bibitem[Nadafian and Ganjtabesh(2020)]{nadafian_bio-plausible_2020}
Nadafian, A.; Ganjtabesh, M.
\newblock Bio-plausible {Unsupervised} {Delay} {Learning} for {Extracting}
  {Temporal} {Features} in {Spiking} {Neural} {Networks}. \emph{arXiv}  \textbf{2020},  	arXiv:2011.09380.

\bibitem[Wang {et~al.}(2019)Wang, Lin, and Dang]{wang_delay_2019}
Wang, X.; Lin, X.; Dang, X.
\newblock A {Delay} {Learning} {Algorithm} {Based} on {Spike} {Train} {Kernels}
  for {Spiking} {Neurons}.
\newblock {\em Front. Neurosci.} {\bf 2019}, {\em 13}, 252.

\bibitem[Hazan {et~al.}(2022)Hazan, Caby, Earl, Siegelmann, and
  Levin]{hazan_memory_2022}
Hazan, H.; Caby, S.; Earl, C.; Siegelmann, H.; Levin, M.
\newblock Memory via {Temporal} {Delays} in weightless {Spiking} {Neural}
  {Network}.  \emph{arXiv} \textbf{2022}, arXiv:2202.07132.
%\newblock arXiv:2202.07132 [cs, q-bio, stat].

\bibitem[Luo {et~al.}(2022)Luo, Qu, Wang, Yi, Zhang, and
  Zhang]{luo_supervised_2022}
Luo, X.; Qu, H.; Wang, Y.; Yi, Z.; Zhang, J.; Zhang, M.
\newblock Supervised {Learning} in {Multilayer} {Spiking} {Neural} {Networks}
  {With} {Spike} {Temporal} {Error} {Backpropagation}.
\newblock {\em IEEE Trans. Neural Netw. Learn. Syst.} {\bf
  \hl{2022} %MDPI: Please add the volume.
}, 1--13.
\newblock {{https://doi.org/10.1109/tnnls.2022.3164930}}.

\bibitem[Sun {et~al.}(2016)Sun, Sourina, and Huang]{sun_learning_2016}
Sun, H.; Sourina, O.; Huang, G.B.
\newblock Learning polychronous neuronal groups using joint weight-delay
  spike-timing-dependent plasticity.
\newblock {\em Neural Comput.} {\bf 2016}, {\em 28},~2181--2212.
%\newblock tex.eprint:
%  https://direct.mit.edu/neco/article-pdf/28/10/2181/972099/neco{\textbackslash}\_a{\textbackslash}\_00879.pdf,
  {{https://doi.org/10.1162/NECO\_a\_00879}}.

\bibitem[Ghosh {et~al.}(2019)Ghosh, Gupta, Silva, Soares, and
  Thakor]{ghosh_spatiotemporal_2019}
Ghosh, R.; Gupta, A.; Silva, A.N.; Soares, A.; Thakor, N.V.
\newblock Spatiotemporal filtering for event-based action recognition. \emph{arXiv}  \textbf{2019},  	arXiv:1903.07067.

\bibitem[Perrinet {et~al.}(2014)Perrinet, Adams, and
  Friston]{perrinet_active_2014}
Perrinet, L.U.; Adams, R.A.; Friston, K.J.
\newblock Active inference, eye movements and oculomotor delays.
\newblock {\em Biol. Cybern.} {\bf 2014}, {\em 108},~777--801.
\newblock {{https://doi.org/10.1007/s00422-014-0620-8}}.

\bibitem[Hogendoorn and Burkitt(2019)]{hogendoorn_predictive_2019}
Hogendoorn, H.; Burkitt, A.N.
\newblock Predictive {Coding} with {Neural} {Transmission} {Delays}: {A}
  {Real}-{Time} {Temporal} {Alignment} {Hypothesis}.
\newblock {\em eNeuro} {\bf 2019}, {\em 6},~ENEURO.0412--18.2019.
\newblock {{https://doi.org/10.1523/eneuro.0412-18.2019}}.

\bibitem[Khoei {et~al.}(2013)Khoei, Masson, and
  Perrinet]{khoei_motion-based_2013}
Khoei, M.A.; Masson, G.S.; Perrinet, L.U.
\newblock Motion-based prediction explains the role of tracking in motion
  extrapolation.
\newblock {\em J. Physiol.-Paris} {\bf 2013}, {\em 107},~409--420.
\newblock {{https://doi.org/10.1016/j.jphysparis.2013.08.001}}.

\bibitem[Kaplan {et~al.}(2013)Kaplan, Lansner, Masson, and
  Perrinet]{kaplan_anisotropic_2013}
Kaplan, B.A.; Lansner, A.; Masson, G.S.; Perrinet, L.U.
\newblock Anisotropic connectivity implements motion-based prediction in a
  spiking neural network.
\newblock {\em Front. Comput. Neurosci.} {\bf 2013}, {\em 7}, 112.
%\newblock Loaded from an external bibliography file by Manubot.
%  source\_bibliography: manual-references.bib standard\_id: Kaplan13,
  {{https://doi.org/10.3389/fncom.2013.00112}}.

\bibitem[Khoei {et~al.}(2017)Khoei, Masson, and
  Perrinet]{khoei_flash-lag_2017}
Khoei, M.A.; Masson, G.S.; Perrinet, L.U.
\newblock The {Flash}-{Lag} {Effect} as a {Motion}-{Based} {Predictive}
  {Shift}.
\newblock {\em PLoS Comput. Biol.} {\bf 2017}, {\em 13},~e1005068.
%\newblock Publisher: Public Library of Science,
  {{https://doi.org/10.1371/journal.pcbi.1005068}}.

\bibitem[Javanshir {et~al.}(2022)Javanshir, Nguyen, Mahmud, and
  Kouzani]{javanshir_advancements_2022}
Javanshir, A.; Nguyen, T.T.; Mahmud, M.A.P.; Kouzani, A.Z.
\newblock Advancements in {Algorithms} and {Neuromorphic} {Hardware} for
  {Spiking} {Neural} {Networks}.
\newblock {\em Neural Comput.} {\bf 2022}, {\em 34},~1289--1328.
\newblock {{https://doi.org/10.1162/neco\_a\_01499}}.

\bibitem[Marković {et~al.}(2020)Marković, Mizrahi, Querlioz, and
  Grollier]{markovic_physics_2020}
Marković, D.; Mizrahi, A.; Querlioz, D.; Grollier, J.
\newblock Physics for neuromorphic computing.
\newblock {\em Nat. Rev. Phys.} {\bf 2020}, {\em 2},~499--510.
%\newblock Number: 9 Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/s42254-020-0208-2}}.

\bibitem[Rasetto {et~al.}(2022)Rasetto, Wan, Akolkar, Shi, Xiong, and
  Benosman]{rasetto_challenges_2022}
Rasetto, M.; Wan, Q.; Akolkar, H.; Shi, B.; Xiong, F.; Benosman, R.
\newblock The {Challenges} {Ahead} for {Bio}-inspired {Neuromorphic} {Event}
  {Processors}: {How} {Memristors} {Dynamic} {Properties} {Could}
  {Revolutionize} {Machine} {Learning}. \emph{arXiv}  \textbf{2022},  	arXiv:2201.12673.
%\newblock Technical Report 2201.12673, .

\bibitem[Diesmann and Gewaltig(2003)]{diesmann_nest_2003}
Diesmann, M.; Gewaltig, M.O.
\newblock {NEST}: {An} {Environment} for {Neural} {Systems} {Simulations}.
\newblock {\em GWDG-Bericht Nr. 58 Theo Plesser, Volker Macho (Hrsg.)}; {
  \hl{2003}}; %MDPI: Please add the name of the publisher and their location.
 p.~29.

\bibitem[Hazan {et~al.}(2018)Hazan, Saunders, Khan, Patel, Sanghavi,
  Siegelmann, and Kozma]{hazan_bindsnet_2018}
Hazan, H.; Saunders, D.J.; Khan, H.; Patel, D.; Sanghavi, D.T.; Siegelmann,
  H.T.; Kozma, R.
\newblock \textls[-15]{{BindsNET}: {A} {Machine} {Learning}-{Oriented} {Spiking} {Neural}
  {Networks} {Library} in {Python}.}
\newblock {\em Front. Neuroinformatics} {\bf 2018}, {\em 12}, 89.
\newblock {{https://doi.org/10.3389/fninf.2018.00089}}.

\bibitem[Stimberg {et~al.}(2019)Stimberg, Brette, and
  Goodman]{stimberg_brian_2019}
Stimberg, M.; Brette, R.; Goodman, D.F.
\newblock Brian 2, an intuitive and efficient neural simulator.
\newblock {\em eLife} {\bf 2019}, {\em 8},~e47314.
\newblock {{https://doi.org/10.7554/eLife.47314}}.

\bibitem[Zenke {et~al.}(2021)Zenke, Bohté, Clopath, Comşa, Göltz, Maass,
  Masquelier, Naud, Neftci, Petrovici, Scherr, and
  Goodman]{zenke_visualizing_2021}
Zenke, F.; Bohté, S.M.; Clopath, C.; Comşa, I.M.; Göltz, J.; Maass, W.;
  Masquelier, T.; Naud, R.; Neftci, E.O.; Petrovici, M.A.;  et~al.
\newblock Visualizing a joint future of neuroscience and neuromorphic
  engineering.
\newblock {\em Neuron} {\bf 2021}, {\em 109},~571--575.
\newblock {{https://doi.org/10.1016/j.neuron.2021.01.009}}.

\bibitem[Mead and Ismail(1989)]{mead_analog_1989}
Mead, C.; Ismail, M.
\newblock {\em Analog {VLSI} {Implementation} of {Neural} {Systems}}; Springer
  Science \& Business Media:  \hl{Berlin/Heidelberg, Germany,} %newly added information, please confirm
  1989.
%\newblock Google-Books-ID: 9e29dOiXeiMC.

\bibitem[Bartolozzi and Indiveri(2007)]{bartolozzi_synaptic_2007}
Bartolozzi, C.; Indiveri, G.
\newblock Synaptic {Dynamics} in {Analog} {VLSI}.
\newblock {\em Neural Comput.} {\bf 2007}, {\em 19},~2581--2603.\linebreak
\newblock {{https://doi.org/10.1162/neco.2007.19.10.2581}}.

\bibitem[Schuman {et~al.}(2017)Schuman, Potok, Patton, Birdwell, Dean, Rose,
  and Plank]{schuman_survey_2017}
Schuman, C.D.; Potok, T.E.; Patton, R.M.; Birdwell, J.D.; Dean, M.E.; Rose,
  G.S.; Plank, J.S.
\newblock A {Survey} of {Neuromorphic} {Computing} and {Neural} {Networks} in
  {Hardware}. \emph{arXiv} {\bf 2017}, arXiv:1705.06963
%\newblock {\em  [cs]} .
%\newblock arXiv: 1705.06963.

\bibitem[Furber {et~al.}(2013)Furber, Lester, Plana, Garside, Painkras,
  Temple, and Brown]{furber_overview_2013}
Furber, S.B.; Lester, D.R.; Plana, L.A.; Garside, J.D.; Painkras, E.; Temple,
  S.; Brown, A.D.
\newblock Overview of the {SpiNNaker} {System} {Architecture}.
\newblock {\em IEEE Trans. Comput.} {\bf 2013}, {\em
  62},~2454--2467.
\newblock {{https://doi.org/10.1109/TC.2012.142}}.

\bibitem[Furber and Bogdan(2020)]{furber_spinnaker_2020}
Furber, S.; Bogdan, P., Eds.
\newblock {\em {SpiNNaker}: {A} {Spiking} {Neural} {Network} {Architecture}};
  Now Publishers: \hl{Norwell, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
  2020.
\newblock {{https://doi.org/10.1561/9781680836523}}.

\bibitem[Merolla {et~al.}(2014)Merolla, Arthur, Alvarez-Icaza, Cassidy,
  Sawada, Akopyan, Jackson, Imam, Guo, Nakamura, Brezzo, Vo, Esser, Appuswamy,
  Taba, Amir, Flickner, Risk, Manohar, and Modha]{merolla_million_2014}
Merolla, P.A.; Arthur, J.V.; Alvarez-Icaza, R.; Cassidy, A.S.; Sawada, J.;
  Akopyan, F.; Jackson, B.L.; Imam, N.; Guo, C.; Nakamura, Y.;  et~al.
\newblock A million spiking-neuron integrated circuit with a scalable
  communication network and interface.
\newblock {\em Science} {\bf 2014}, {\em 345},~668--673.
\newblock {{https://doi.org/10.1126/science.1254642}}.

\bibitem[Benjamin {et~al.}(2014)Benjamin, Gao, McQuinn, Choudhary,
  Chandrasekaran, Bussat, Alvarez-Icaza, Arthur, Merolla, and
  Boahen]{benjamin_neurogrid_2014}
Benjamin, B.V.; Gao, P.; McQuinn, E.; Choudhary, S.; Chandrasekaran, A.R.;
  Bussat, J.M.; Alvarez-Icaza, R.; Arthur, J.V.; Merolla, P.A.; Boahen, K.
\newblock Neurogrid: {A} {Mixed}-{Analog}-{Digital} {Multichip} {System} for
  {Large}-{Scale} {Neural} {Simulations}.
\newblock {\em Proc.  IEEE} {\bf 2014}, {\em 102},~699--716.
%\newblock Conference Name: Proceedings of the IEEE,
  {{https://doi.org/10.1109/JPROC.2014.2313565}}.

\bibitem[Neckar {et~al.}(2019)Neckar, Fok, Benjamin, Stewart, Oza, Voelker,
  Eliasmith, Manohar, and Boahen]{neckar_braindrop_2019}
Neckar, A.; Fok, S.; Benjamin, B.V.; Stewart, T.C.; Oza, N.N.; Voelker, A.R.;
  Eliasmith, C.; Manohar, R.; Boahen, K.
\newblock Braindrop: {A} mixed-signal neuromorphic architecture with a
  dynamical systems-based programming model.
\newblock {\em Proc.  IEEE} {\bf 2019}, {\em 107},~144--164.
\newblock {{https://doi.org/10.1109/JPROC.2018.2881432}}.

\bibitem[Schemmel {et~al.}(2010)Schemmel, Brüderle, Grübl, Hock, Meier,
  and Millner]{schemmel_wafer-scale_2010}
Schemmel, J.; Brüderle, D.; Grübl, A.; Hock, M.; Meier, K.; Millner, S.
\newblock A wafer-scale neuromorphic hardware system for large-scale neural
  modeling.
\newblock In Proceedings of the 2010 {IEEE} {International} {Symposium} on
  {Circuits} and {Systems} ({ISCAS}), \hl{Paris, France, 30 May--2 June} %MDPI: We added the location and date of the conference. Please confirm.
 2010; pp. 1947--1950.
{{https://doi.org/10.1109/ISCAS.2010.5536970}}.

\bibitem[Markram {et~al.}(2011)Markram, Meier, Lippert, Grillner,
  Frackowiak, Dehaene, Knoll, Sompolinsky, Verstreken, DeFelipe, Grant,
  Changeux, and Saria]{markram_introducing_2011}
Markram, H.; Meier, K.; Lippert, T.; Grillner, S.; Frackowiak, R.; Dehaene, S.;
  Knoll, A.; Sompolinsky, H.; Verstreken, K.; DeFelipe, J.;  et~al.
\newblock Introducing the {Human} {Brain} {Project}.
\newblock {\em Procedia Comput. Sci.} {\bf 2011}, {\em 7},~39--42.
\newblock {{https://doi.org/10.1016/j.procs.2011.12.015}}.

\bibitem[Farquhar {et~al.}(2006)Farquhar, Gordon, and
  Hasler]{farquhar_field_2006}
Farquhar, E.; Gordon, C.; Hasler, P.
\newblock A {Field} {Programmable} {Neural} {Array}.
\newblock In Proceedings of the 2006 {IEEE} {International} {Symposium} on
  {Circuits} and {Systems}, \hl{Kos, Greece, 21--24 May} %MDPI: We added the location and date of the conference. Please confirm.
 2006; pp. 4114--4117.
\newblock {{https://doi.org/10.1109/ISCAS.2006.1693534}}.

\bibitem[Liu {et~al.}(2009)Liu, Yu, and Wang]{cheng_fpaa_2009}
Liu, M.; Yu, H.; Wang, W.
\newblock {FPAA} {Based} on {Integration} of {CMOS} and {Nanojunction}
  {Devices} for {Neuromorphic} {Applications}. In {\em Nano-{Net}}; Cheng, M.,
  Ed.; Springer: Berlin/Heidelberg, Germany,  2009; Volume~3, pp.
  44--48.

\bibitem[Chan {et~al.}(2007)Chan, Liu, and van Schaik]{chan_aer_2007}
Chan, V.; Liu, S.C.; van Schaik, A.
\newblock {AER} {EAR}: {A} {Matched} {Silicon} {Cochlea} {Pair} {With}
  {Address} {Event} {Representation} {Interface}.
\newblock {\em IEEE Trans. Circuits Syst. I Regul. Pap.}
  {\bf 2007}, {\em 54},~48--59.
%\newblock Conference Name: IEEE Transactions on Circuits and Systems I: Regular
%  Papers,
  {{https://doi.org/10.1109/TCSI.2006.887979}}.

\bibitem[Haessig {et~al.}(2020)Haessig, Milde, Aceituno, Oubari, Knight, van
  Schaik, Benosman, and Indiveri]{haessig_event-based_2020}
Haessig, G.; Milde, M.B.; Aceituno, P.V.; Oubari, O.; Knight, J.C.; van Schaik,
  A.; Benosman, R.B.; Indiveri, G.
\newblock Event-{Based} {Computation} for {Touch} {Localization} {Based} on
  {Precise} {Spike} {Timing}.
\newblock {\em Front. Neurosci.} {\bf 2020}, {\em 14},~420.
\newblock {{https://doi.org/10.3389/fnins.2020.00420}}.

\bibitem[Lagorce {et~al.}(2017)Lagorce, Orchard, Galluppi, Shi, and
  Benosman]{lagorce_hots_2017}
Lagorce, X.; Orchard, G.; Galluppi, F.; Shi, B.E.; Benosman, R.B.
\newblock {HOTS}: {A} {Hierarchy} of {Event}-{Based} {Time}-{Surfaces} for
  {Pattern} {Recognition}.
\newblock {\em IEEE Trans. Pattern Anal. Mach. Intell.}
  {\bf 2017}, {\em 39},~1346--1359.
\newblock {{https://doi.org/10.1109/TPAMI.2016.2574707}}.

\bibitem[Sironi {et~al.}(2018)Sironi, Brambilla, Bourdis, Lagorce, and
  Benosman]{sironi_hats_2018}
Sironi, A.; Brambilla, M.; Bourdis, N.; Lagorce, X.; Benosman, R.
\newblock {HATS}: {Histograms} of {Averaged} {Time} {Surfaces} for {Robust}
  {Event}-{Based} {Object} {Classification}.
\newblock In Proceedings of the 2018 {IEEE}/{CVF} {Conference} on {Computer}
  {Vision} and {Pattern} {Recognition}, Salt Lake City, UT, USA, \hl{18--23 June} %MDPI: We added the  date of the conference. Please confirm.
  2018;
  pp. 1731--1740.
\newblock {{https://doi.org/10.1109/CVPR.2018.00186}}.

\bibitem[Maro {et~al.}(2020)Maro, Ieng, and Benosman]{maro_event-based_2020}
Maro, J.M.; Ieng, S.H.; Benosman, R.
\newblock Event-{Based} {Gesture} {Recognition} {With} {Dynamic} {Background}
  {Suppression} {Using} {Smartphone} {Computational} {Capabilities}.
\newblock {\em Front. Neurosci.} {\bf 2020}, {\em 14}, 275.
%\newblock Publisher: Frontiers,
  {{https://doi.org/10.3389/fnins.2020.00275}}.

\bibitem[Grimaldi {et~al.}(2021)Grimaldi, Boutin, Perrinet, Ieng, and
  Benosman]{grimaldi_homeostatic_2021}
Grimaldi, A.; Boutin, V.; Perrinet, L.; Ieng, S.H.; Benosman, R.
\newblock A homeostatic gain control mechanism to improve event-driven object
  recognition.
\newblock In Proceedings of the 2021 {International} {Conference} on
  {Content}-{Based} {Multimedia} {Indexing} ({CBMI}),  \hl{Lille, France,  28--30 June} %MDPI: We added the location and date of the conference. Please confirm.
  2021.
\newblock {{https://doi.org/10.1109/cbmi50038.2021.9461901}}.

\bibitem[Grimaldi {et~al.}(2022)Grimaldi, Boutin, Ieng, Benosman, and
  Perrinet]{grimaldi_robust_2022}
Grimaldi, A.; Boutin, V.; Ieng, S.H.; Benosman, R.; Perrinet, L.U.
\newblock A robust event-driven approach to always-on object recognition.
\newblock {\em TechRxiv} {\bf 2022}.
\newblock {{https://doi.org/10.36227/techrxiv.18003077.v1}}.

\bibitem[Yu {et~al.}(2022)Yu, Gu, Li, Wang, Wang, and Li]{yu_stsc-snn_2022}
Yu, C.; Gu, Z.; Li, D.; Wang, G.; Wang, A.; Li, E.
\newblock {STSC}-{SNN}: {Spatio}-{Temporal} {Synaptic} {Connection} with
  {Temporal} {Convolution} and {Attention} for {Spiking} {Neural} {Networks}.  	\emph{arXiv}
  2022.
\newblock arXiv:2210.05241.

\bibitem[Benosman {et~al.}(2014)Benosman, Clercq, Lagorce, {Sio-Hoi Ieng},
  and Bartolozzi]{benosman_event-based_2014}
Benosman, R.; Clercq, C.; Lagorce, X.; Ieng, {S.-H}.; Bartolozzi, C.
\newblock Event-{Based} {Visual} {Flow}.
\newblock {\em IEEE Trans. Neural Netw. Learn. Syst.} {\bf
  2014}, {\em 25},~407--417.
\newblock {{https://doi.org/10.1109/tnnls.2013.2273537}}.

\bibitem[Clady {et~al.}(2014)Clady, Clercq, Ieng, Houseini, Randazzo,
  Natale, Bartolozzi, and Benosman]{clady_asynchronous_2014}
Clady, X.; Clercq, C.; Ieng, S.H.; Houseini, F.; Randazzo, M.; Natale, L.;
  Bartolozzi, C.; Benosman, R.B.
\newblock Asynchronous visual event-based time-to-contact.
\newblock {\em Front. Neurosci.} {\bf 2014}, {\em 8}, 9.
%\newblock 00000 Publisher: Frontiers,
  {{https://doi.org/10.3389/fnins.2014.00009}}.

\bibitem[Tschechne {et~al.}(2014)Tschechne, Sailer, and
  Neumann]{tschechne_bio-inspired_2014}
Tschechne, S.; Sailer, R.; Neumann, H.
\newblock Bio-{Inspired} {Optic} {Flow} from {Event}-{Based} {Neuromorphic}
  {Sensor} {Input}.
\newblock In \emph{Proceedings of the Artificial {Neural} {Networks} in {Pattern}
  {Recognition}}; El~Gayar, N.; Schwenker, F.; Suen, C., Eds.; Springer
  International Publishing: Cham, Switzerland, 2014; pp. 171--182.
\newblock {{https://doi.org/10.1007/978-3-319-11656-3\_16}}.

\bibitem[Hidalgo-Carrió {et~al.}(2020)Hidalgo-Carrió, Gehrig, and
  Scaramuzza]{hidalgo-carrio_learning_2020}
Hidalgo-Carrió, J.; Gehrig, D.; Scaramuzza, D.
\newblock Learning {Monocular} {Dense} {Depth} from {Events}.  \emph{arXiv}, {\bf 2020}, arXiv:2010.08350.
%\newblock {\em arXiv:2010.08350 [cs]} .
%\newblock arXiv: 2010.08350.

\bibitem[Dardelet {et~al.}(2021)Dardelet, Benosman, and
  Ieng]{dardelet_event-by-event_2021}
Dardelet, L.; Benosman, R.; Ieng, S.H.
\newblock An {Event}-by-{Event} {Feature} {Detection} and {Tracking}
  {Invariant} to {Motion} {Direction} and {Velocity}. \emph{TechRxiv} {\bf 2021}.
\newblock {{https://doi.org/10.36227/techrxiv.17013824.v1}}.

\bibitem[Stoffregen {et~al.}(2019)Stoffregen, Gallego, Drummond, Kleeman,
  and Scaramuzza]{stoffregen_event-based_2019}
Stoffregen, T.; Gallego, G.; Drummond, T.; Kleeman, L.; Scaramuzza, D.
\newblock Event-{Based} {Motion} {Segmentation} by {Motion} {Compensation}.
\newblock In Proceedings of the 2019 {IEEE}/{CVF} {International} {Conference}
  on {Computer} {Vision} ({ICCV}), \hl{Seoul, Republic of Korea, 27 February} %MDPI: We added the location and date of the conference. Please confirm.
 2019; pp. 7243--7252.
 {{https://doi.org/10.1109/ICCV.2019.00734}}.

\bibitem[Kim {et~al.}(2016)Kim, Leutenegger, and
  Davison]{kim_real-time_2016}
Kim, H.; Leutenegger, S.; Davison, A.J.
\newblock Real-{Time} {3D} {Reconstruction} and 6-{DoF} {Tracking} with an
  {Event} {Camera}.
\newblock In \emph{Proceedings of the Computer {Vision}---{ECCV} 2016}; Leibe, B.;
  Matas, J.; Sebe, N.; Welling, M., Eds.; Springer International Publishing:
  Cham,  Switzerland, 2016;  pp. 349--364.
\newblock {{https://doi.org/10.1007/978-3-319-46466-4\_21}}.

\bibitem[Hussain {et~al.}(2012)Hussain, Basu, Wang, and
  Hamilton]{hussain_deltron_2012}
Hussain, S.; Basu, A.; Wang, M.; Hamilton, T.J.
\newblock {DELTRON}: {Neuromorphic} architectures for delay based learning.
\newblock In Proceedings of the 2012 {IEEE} {Asia} {Pacific} {Conference} on
  {Circuits} and {Systems},  \hl{Kaohsiung, Taiwan, 2--5 December} %MDPI: We added the location and date of the conference. Please confirm.
 2012; pp. 304--307.
\newblock {{https://doi.org/10.1109/APCCAS.2012.6419032}}.

\bibitem[Wang {et~al.}(2015)Wang, Hamilton, Tapson, and van
  Schaik]{wang_neuromorphic_2015}
Wang, R.M.; Hamilton, T.J.; Tapson, J.C.; van Schaik, A.
\newblock A neuromorphic implementation of multiple spike-timing synaptic
  plasticity rules for large-scale neural networks.
\newblock {\em Front. Neurosci.} {\bf 2015}, {\em 9}, 180.
\newblock {{https://doi.org/10.3389/fnins.2015.00180}}.

\bibitem[Wang {et~al.}(2014)Wang, Hamilton, Tapson, and van
  Schaik]{wang_fpga_2014}
Wang, R.; Hamilton, T.J.; Tapson, J.; van Schaik, A.
\newblock An {FPGA} design framework for large-scale spiking neural networks.
\newblock In Proceedings of the 2014 {IEEE} {International} {Symposium} on
  {Circuits} and {Systems} ({ISCAS}), \hl{Melbourne, VIC, Australia, 1--5 June} %MDPI: We added the location and date of the conference. Please confirm.
 2014; pp. 457--460.
 {{https://doi.org/10.1109/ISCAS.2014.6865169}}.

\bibitem[Pfeil {et~al.}(2013)Pfeil, Scherzer, Schemmel, and
  Meier]{pfeil_neuromorphic_2013}
Pfeil, T.; Scherzer, A.C.; Schemmel, J.; Meier, K.
\newblock Neuromorphic learning towards nano second precision.
\newblock In Proceedings of the The 2013 {International} {Joint} {Conference}
  on {Neural} {Networks} ({IJCNN}), \hl{Dallas, TX, USA, 4--9 August} %MDPI: We added the location and date of the conference. Please confirm.
 2013; pp. 1--5.
 {{https://doi.org/10.1109/IJCNN.2013.6706828}}.

\bibitem[Boerlin and Denève(2011)]{boerlin_spike-based_2011}
Boerlin, M.; Denève, S.
\newblock Spike-{Based} {Population} {Coding} and {Working} {Memory}.
\newblock {\em PLoS Comput. Biol.} {\bf 2011}, {\em 7},~e1001080.
\newblock {{https://doi.org/10.1371/journal.pcbi.1001080}}.

\bibitem[Renner {et~al.}(2022)Renner, Sandamirskaya, Sommer, and
  Frady]{renner_sparse_2022}
Renner, A.; Sandamirskaya, Y.; Sommer, F.T.; Frady, E.P.
\newblock Sparse {Vector} {Binding} on {Spiking} {Neuromorphic} {Hardware}
  {Using} {Synaptic} {Delays}.
\newblock {\em Proceedings of the International Conference on Neuromorphic
  Systems};
%\newblock Conference Name: ICONS 2022: International Conference on Neuromorphic
%  Systems 2022 Meeting Name: ICONS 2022: International Conference on
%  Neuromorphic Systems 2022 Place:  
  Publisher: ACM Digital
  library: Knoxville, TN, USA, 2022. {{https://doi.org/10.1145/3546790.3546820}}.

\bibitem[Dard {et~al.}(2022)Dard, Leprince, Denis, Rao~Balappa, Suchkov,
  Boyce, Lopez, Giorgi-Kurz, Szwagier, Dumont, Rouault, Minlebaev, Baude,
  Cossart, and Picardo]{dard_rapid_2022}
Dard, R.F.; Leprince, E.; Denis, J.; Rao~Balappa, S.; Suchkov, D.; Boyce, R.;
  Lopez, C.; Giorgi-Kurz, M.; Szwagier, T.; Dumont, T.;  et~al.
\newblock The rapid developmental rise of somatic inhibition disengages
  hippocampal dynamics from self-motion.
\newblock {\em eLife} {\bf 2022}, {\em 11},~e78116.
\newblock {{https://doi.org/10.7554/eLife.78116}}.

\bibitem[Coull and Giersch(2022)]{coull_distinction_2022}
Coull, J.T.; Giersch, A.
\newblock The distinction between temporal order and duration processing, and
  implications for schizophrenia.
\newblock {\em Nat. Rev. Psychol.} {\bf 2022}, {\em 1},~257--271.
%\newblock Publisher: Nature Publishing Group,
  {{https://doi.org/10.1038/s44159-022-00038-y}}.

\bibitem[Panahi {et~al.}(2021)Panahi, Abrevaya, Gagnon-Audet, Voleti, Rish,
  and Dumas]{panahi_generative_2021}
Panahi, M.R.; Abrevaya, G.; Gagnon-Audet, J.C.; Voleti, V.; Rish, I.; Dumas, G.
\newblock Generative {Models} of {Brain} {Dynamics}---{A} review. \emph{arXiv} \textbf{2021},
\newblock arXiv:2112.12147.

\bibitem[Tolle {et~al.}(2011)Tolle, Tansley, and Hey]{tolle_fourth_2011}
Tolle, K.M.; Tansley, D.S.W.; Hey, A.J.G.
\newblock The {Fourth} {Paradigm}: {Data}-{Intensive} {Scientific} {Discovery}
  [{Point} of {View}].
\newblock {\em Proc.  IEEE} {\bf 2011}, {\em 99},~1334--1337.
\newblock {{https://doi.org/10.1109/jproc.2011.2155130}}.

\end{thebibliography}

%\begin{thebibliography}{00}
%\printbibliography[heading=none]
%\end{thebibliography}

% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author:~\citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a},~\citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages:~\citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328;~\citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)
\PublishersNote{}
\end{adjustwidth}
\end{document}

%
%The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper~\citep{ref-journal}. Now citing a book reference~\citep{ref-book1,ref-book2} or other reference types~\citep{ref-unpublish,ref-communication,ref-proceeding}. Please use the command~\citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Materials and Methods}
%
%Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
%
%Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.
%
%Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
%\begin{quote}
%This is an example of a quote.
%\end{quote}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Results}
%
%This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
%\section{Subsection}
%\subsection{Subsubsection}
%
%Bulleted lists look like this:
%\begin{itemize}
%\item	First bullet;
%\item	Second bullet;
%\item	Third bullet.
%\end{itemize}
%
%Numbered lists can be added as follows:
%\begin{enumerate}
%\item	First item; 
%\item	Second item;
%\item	Third item.
%\end{enumerate}
%
%The text continues here. 
%
%\section{Figures, Tables and Schemes}
%
%All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.
%
%\begin{figure}[H]
%\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
%\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
%\end{figure}   
%\unskip
%
%\begin{table}[H] 
%\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
%\newcolumntype{C}{>{\centering\arraybackslash}X}
%\begin{tabularx}{\textwidth}{CCC}
%\toprule
%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
%\midrule
%Entry 1		& Data			& Data\\
%Entry 2		& Data			& Data\\
%\bottomrule
%\end{tabularx}
%\end{table}
%\unskip
%
%\begin{table}[H]
%\caption{This is a wide table.\label{tab2}}
%	\begin{adjustwidth}{-\extralength}{0cm}
%		\newcolumntype{C}{>{\centering\arraybackslash}X}
%		\begin{tabularx}{\fulllength}{CCCC}
%			\toprule
%			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
%			\midrule
%			Entry 1		& Data			& Data			& Data\\
%			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
%			\bottomrule
%		\end{tabularx}
%	\end{adjustwidth}
%	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
%\end{table}
%
%%\begin{listing}[H]
%%\caption{Title of the listing}
%%\rule{\columnwidth}{1pt}
%%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%%\rule{\columnwidth}{1pt}
%%\end{listing}
%
%Text.
%
%Text.
%
%\section{Formatting of Mathematical Components}
%
%This is the example 1 of equation:
%\begin{linenomath}
%\begin{equation}
%a = 1,
%\end{equation}
%\end{linenomath}
%the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
%
%This is the example 2 of equation:
%\begin{adjustwidth}{-\extralength}{0cm}
%\begin{equation}
%a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
%\end{equation}
%\end{adjustwidth}
%
%% Example of a page in landscape format (with table and table footnote).
%%\startlandscape
%%\begin{table}[H] %% Table in wide page
%%\caption{This is a very wide table.\label{tab3}}
%%	\begin{tabularx}{\textwidth}{CCCC}
%%		\toprule
%%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
%%		\midrule
%%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
%%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
%%		\bottomrule
%%	\end{tabularx}
%%	\begin{adjustwidth}{+\extralength}{0cm}
%%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
%%	\end{adjustwidth}
%%\end{table}
%%\finishlandscape
%
%% Example of a figure that spans the whole page width. The same concept works for tables, too.
%\begin{figure}[H]
%\begin{adjustwidth}{-\extralength}{0cm}
%\centering
%\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
%\end{adjustwidth}
%\caption{This is a wide figure.\label{fig2}}
%\end{figure}  
%
%Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%%% Example of a theorem:
%\begin{Theorem}
%Example text of a theorem.
%\end{Theorem}
%
%The text continues here. Proofs must be formatted as follows:
%
%%% Example of a proof:
%\begin{proof}[Proof of Theorem 1]
%Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
%\end{proof}
%The text continues here.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Discussion}
%
%Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Conclusions}
%
%This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

### 2022-12-28 brainsci-2069565_accepted.tex
brainsci-2069565_accepted

--- 2022-12-28 brainsci-2069565.tex
@@
 \usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
-
 \newcommand{\ms}{\si{\milli\second}}%
@@
 % Authors, for the paper (add full first names)
-\Author{\hl{Antoine Grimaldi} %MDPI: Please carefully check the accuracy of names and affiliations.
- $^{1}$% $^{1,\dagger,\ddagger}$
+\Author{Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
 \orcidA{}, 
@@
 Camille Besnainou $^{1}$\orcidC{},
-\hl{Jean-Nicolas Jérémie} %MDPI: The name of this author is different from the one submitted online at susy.mdpi.com. Please confirm which one is correct.
- $^{1}$\orcidD{},
+Jean-Nicolas Jérémie $^{1}$\orcidD{},
 Jean Martinet $^{2}$\orcidE{}
-and~\hl{Laurent Perrinet} %MDPI: The name of this author is different from the one submitted online at susy.mdpi.com. Please confirm which one is correct.
- $^{1,}$*\orcidF{}}
+and~Laurent U Perrinet $^{1,}$*\orcidF{}}
 %
@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent Perrinet}
+\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent U Perrinet}
 %
@@
 \address{%
-$^{1}$ \quad \hl{INT UMR 7289, Aix Marseille Univ, CNRS, 27 Bd Jean Moulin,} %Please confirm whether the address information is correct.
-13005 Marseille, France; \hl{antoine.grimaldi@univ-amu.fr (A.G.); camille.besnainou@etu.univ-amu.fr (C.B.)} %MDPI: We added the email addresses here according to those submitted online at susy.mdpi.com.
-Please confirm.
+$^{1}$ \quad INT UMR 7289, Aix Marseille Univ, CNRS, 27 Bd Jean Moulin, 13005 Marseille, France; antoine.grimaldi@univ-amu.fr (A.G.); camille.besnainou@etu.univ-amu.fr (C.B.); jean-nicolas.jeremie@univ-amu.fr (J.-N.J.), laurent.perrinet@univ-amu.fr (L.U.P.)
  \\
-$^{2}$ \quad \hl{SPARKS,  Côte d'Azur, CNRS, I3S, 2000 Rte des Lucioles,} %Please confirm whether the address information is correct.
-06900 Sophia-Antipolis, France; \hl{amelie.gruel@i3s.unice.fr (A.G.); jean.martinet@univ-cotedazur.fr (J.M.)} %MDPI: We added the email addresses here according to those submitted online at susy.mdpi.com.
-Please confirm.
-
+$^{2}$ \quad SPARKS,  Côte d'Azur, CNRS, I3S, 2000 Rte des Lucioles,
+06900 Sophia-Antipolis, France; amelie.gruel@i3s.unice.fr (A.G.); jean.martinet@univ-cotedazur.fr (J.M.)
 }

### 2022-12-28 brainsci-2069565_accepted.tex
MDPI: Please confirm if the bold is unnecessary and can be removed. The following highlights are the same.

@@
 \usepackage[finalnew]{trackchanges}
-\addeditor{Antoine}
-\addeditor{Amélie}
-\addeditor{Camille}
-\addeditor{Jean}
-\addeditor{JN}
-\addeditor{Laurent}
+\addeditor{MDPI}
 
@@
 %Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
-In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this \emph{\hl{event-based representation} %MDPI: Please confirm if the italics is unnecessary and can be removed.
-The following highlights are the same.
-} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
+In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this event-based representation observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
 For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{of the rate coding hypothesis that relies only on the firing rate} by considering \change[Antoine]{a representation based on repetitions of spiking motifs at precise times of occurrence.}{representations based on the precise timing of single spikes.} \remove[Antoine]{This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.} Additionally, numerous studies have demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
@@
 % \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
-\caption{{\bf \hl{Latency of the different processing steps along the human visual pathway.} %MDPI: Please confirm if the bold is unnecessary and can be removed. The following highlights are the same.
-} Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
+\caption{
+  Latency of the different processing steps along the human visual pathway.Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
 Different areas are schematically represented by ellipses, and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
@@
 \includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
-\caption{\textbf{\hl{Reproducibility of the spiking response of a neuron.} }The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
+\caption{
+  Reproducibility of the spiking response of a neuron. 
+The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
 The stimulus current value over time for a step stimulus (top left) and for a noisy one (top right).
@@
 \includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
-\caption{\textbf{\hl{Simulation of a synfire propagation} \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{\hl{using Brian}}.} The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
+\caption{
+  Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{\hl{using Brian}}.} 
+  The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
 Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreases from $5$ to $1$~\ms).
@@
 \includegraphics[width=0.98\linewidth]{figures/haimerl2019.jpg}
-\caption{\textbf{\hl{An example of a precise temporal motifs seen in cortical slices.}} In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+\caption{
+        An example of a precise temporal motifs seen in cortical slices. 
+        In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
 (\textbf{A}) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
@@
 \includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
-\caption{\textbf{\hl{Detecting motifs using SPADE.}} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{\hl{tutorial}} on the data generated in  Figure~\ref{fig:diesman}.
+\caption{
+  Detecting motifs using SPADE. 
+  We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{\hl{tutorial}} on the data generated in  Figure~\ref{fig:diesman}.
 This allowed to label different precise spike motifs which are denoted by different colors.
@@
   \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-    \caption{\textbf{\hl{Detecting event-based motifs using spiking neurons with heterogeneous delays.}} 
+    \caption{
+      Detecting event-based motifs using spiking neurons with heterogeneous delays.
     { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
@@
   \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
-    \caption{\textbf{\hl{Core mechanism of polychrony detection}~\citep{izhikevich_polychronization_2006}.} {(\textbf{Left})}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
+    \caption{
+      Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.
+       {(\textbf{Left})}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
 {(\textbf{Middle})}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons; therefore, no spike will be emitted, as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
@@
 \includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
-\caption{\textbf{\hl{A miniature, event-based ATIS sensor.}} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides events at the micro-second resolution.
+\caption{
+ A miniature, event-based ATIS sensor.
+ Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides events at the micro-second resolution.
 These are sparse, as they represent luminance increments or decrements (ON and OFF events, respectively).
@@
   %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
-    \caption{\textbf{\hl{Detecting visual motion in an event stream with heterogeneous delays.}} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
+    \caption{
+      Detecting visual motion in an event stream with heterogeneous delays.
+      Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
 The models' parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
@@
 
-\institutionalreview{\hl{ }
+\institutionalreview{Not applicable.
 %In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
@@
 
-\informedconsent{\hl{ }
+\informedconsent{Not applicable.
 %Any research article describing a study involving humans should contain this statement.
@@
 
-\dataavailability{This works is made reproducible using the following tools.
-First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/XXX}{\hl{GitHub}}.
+\dataavailability{This work is made reproducible using the following tools.
+First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/2022_polychronies-review}{GitHub}.
 %~\citep{Grimaldi2023review}.
-The paper is available as an \href{https://arxiv.org/abs/YYYY.XXX}{\hl{arXiv preprint}} with links to previous versions.
-Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{\hl{Zotero group}} which was used to regroup relevant literature on the subject.}
-
-\acknowledgments{We would like to thank Hugo Ladret and Frdéric Chavane for valuable feedback while writing this review.
+This list also links to preprints versions of this review as well as links to previous versions.
+Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{Zotero group} which was used to regroup relevant literature on the subject.}
+
+\acknowledgments{We would like to thank Hugo Ladret and Frédéric Chavane for valuable feedback while writing this review.
 Many thanks to Simon Thorpe for the useful discussions and references on the latency of the visual system in human and non-human primates.}

### 2022-12-28 brainsci-2069565_accepted.tex
Update brainsci-2069565_accepted.tex

@@
 Membrane potential is represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row).
-While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{\hl{frozen}} noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
-See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{\hl{notebook}} %MDPI: Please confirm if the link is unnecessary and can be removed.
+While this seems paradoxical at first sight, it highlights the consequence of using the same frozen noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} %MDPI: Please confirm if the link is unnecessary and can be removed.
 The following highlights are the same.
@@
 The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
-Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{\hl{Benham Top}} based on center--surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center--surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
 This evidence found in the retina can be extended to other areas, such as the visual cortex.
@@
 \caption{
-  Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{\hl{using Brian}}.} 
+  Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{\hl{using Brian}}.
   The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.

### 2022-12-28 brainsci-2069565_accepted.tex
%MDPI: We removed the italics. Please confirm this revision. These were kept as they correspond to latin locutions

@@
 %
-% REV1
-%
-% https://www.overleaf.com/project/633bf90497b5b325a3ff692e
+% ACCEPTED VERSION
+%
+% https://github.com/SpikeAI/2022_polychronies-review
 % https://framateam.org/spikeai/channels/review_polychrony
@@
 These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}.
-It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{\hl{LFP}} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
+It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018} and were also extended to larger assemblies using statistical methods~\citep{torre_asset_2016} (see Section~\ref{sec:detection} for further details).
 \change[Antoine]{Note also that synchronicity may explain some unintuitive results.
@@
 \caption{
-  Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{\hl{using Brian}}.
+  Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
   The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
@@
 Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
-Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed {\hl{in vivo}}.
-%MDPI: we remove italic, please confirm.
-The following highlights are the same.
+Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed in biology.
 Such waves also occur in motor areas and~\citet{linden_movement_2022} have recently presented ensemble recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic~behavior.
@@
 %We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
-Going further, researchers found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, {\hl{in vivo} %MDPI: We removed the italics. Please confirm this revision.
-} and {\hl{in vitro}}.
+Going further, researchers found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \emph{in vivo} %MDPI: We removed the italics.
+Please confirm this revision.
+These were kept as they correspond to latin locutions
+ and \emph{in vitro}.
 These patterns repeat after minutes, maintaining millisecond accuracy.

### 2022-12-28 brainsci-2069565_accepted.tex
MDPI: Please add the explanation for B--D in the figure. The figure which should have been submitted was a crop on panel A : we have fixed that in the revision

@@
 These sequences can be specific to a particular layer or column, can be synchronized with network activity oscillation, and can involve several cells.
-They also demonstrated that these sequences can form super sequences, so-called \emph{\hl{cortical songs}}.
+They also demonstrated that these sequences can form super sequences, so-called ``cortical songs''.
 It consists of the assembly of several sequences which repeat in a specific order with a compressed timing.
-{\hl{In vivo}} spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3--10 cells out (of about 900), and lasting up to 3 s.
+This spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3--10 cells out (of about 900), and lasting up to 3 s.
 Sequences have specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and are associated with a structured spatial organization of the neurons that formed them.
@@
         In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
-(\textbf{A}) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
+Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
 Cells were selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{\hl{PNAS License}}).}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.
+The black line represents the speed of the mouse (modified from Figure 1-A from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.
+The figure which should have been submitted was a crop on panel A : we have fixed that in the revision
 

### 2022-12-28 brainsci-2069565_accepted.tex
MDPI: Please change the (a) into (d) in picture. Fixed !

@@
 Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006}, which this author defined as \emph{\hl{polychronous groups}}.
+Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006}, which this author defined as polychronous groups.
 The rest of this review will be devoted to present evidence for the use of such precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
@@
 The loss function is the cross entropy (following the Bernouilli hypothesis with a GLM, where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995}) and embedded with recurrent dynamics.
-In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{\hl{code}} and uses the \note[Laurent]{now citing the URL}publicly available V1 dataset~\citep{kohn_utah_2016}, which allows supervising the model, with the input being the movie and the output the spikes recorded.
+In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the \note[Laurent]{now citing the URL}publicly available V1 dataset~\citep{kohn_utah_2016}, which allows supervising the model, with the input being the movie and the output the spikes recorded.
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
@@
   Detecting motifs using SPADE. 
-  We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{\hl{tutorial}} on the data generated in  Figure~\ref{fig:diesman}.
+  We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}.
 This allowed to label different precise spike motifs which are denoted by different colors.
@@
 \add[Amélie]{It was first theorized in the Jeffress ``ITD-versus-place model''}~\citep{jeffress_place_1948}.
-\add[Amélie]{As hypothesized by this model, the \textit{\hl{nucleus laminaris}} of the avian brain contains coincidence detectors and, associated with the \textit{\hl{nucleus magnocellularis}} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
+\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
 According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'' and therefore necessarily need to play on the response time to do so.
@@
 They also confirmed that the delay correlates with the level of skill acquisition.~\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or {\hl{via}} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or {via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
 
@@
 As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing rate.
-Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time-encoding machine by}~\citet{lazar_time_2004} and \textit{\hl{polychronization}} model of~\citet{izhikevich_polychronization_2006}.
+Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time-encoding machine by}~\citet{lazar_time_2004} and polychronization model of~\citet{izhikevich_polychronization_2006}.
 \add[Antoine]{In this section, we focus on the polychronization model} based on a random recurrent model of spiking neurons, including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a spike-time-dependent plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
@@
     %{\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
-    { (\textbf{d})}~\hl{Knowing the} %MDPI: Please change the (a) into (d) in picture.
+    { (\textbf{d})}~Knowing the %MDPI: Please change the (a) into (d) in picture. Fixed !
  results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic example, all spikes are now associated with a PG. 

### 2022-12-28 brainsci-2069565_accepted.tex
moving figures 6 <-> 7 to respect apparition order
fixing caption

@@
 
+\begin{figure}[H]%[t!]
+%  \centering
+  \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
+    \caption{
+      Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.
+       {(\textbf{Left})}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
+{(\textbf{Middle})}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons; therefore, no spike will be emitted, as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+    %at these different delays, and these may not be sufficient to generate a spike in either neuron.
+    {(\textbf{Right})}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
+     }
+  \label{fig:izhikevich}
+\end{figure}
+%
+%----------------------------%
+%
+
 \begin{figure}[H]%[t!][]
@@
         Detecting event-based motifs using spiking neurons with heterogeneous delays.
-    { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
+      { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific neuron addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups''.
       { (\textbf{b})}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay.
@@
   \end{figure}
-
-\begin{figure}[H]%[t!]
-%  \centering
-  \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
-    \caption{
-      Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.
-       {(\textbf{Left})}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-{(\textbf{Middle})}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons; therefore, no spike will be emitted, as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
-    %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    {(\textbf{Right})}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
-     }
-  \label{fig:izhikevich}
-\end{figure}
-%
-%----------------------------%
-%
-
 % 

### 2022-12-28 brainsci-2069565_accepted.tex
%MDPI: Please add the explanation for B--D in the figure. The figure which should have been submitted was a crop on panel A : we have fixed that in the revision

@@
 \caption{
-        An example of a precise temporal motifs seen in cortical slices. 
-        In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
-Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
-Cells were selected and ordered with respect to their activity in the first imaging session (Top).
-The black line represents the speed of the mouse (modified from Figure 1-A from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.
+        An example of a precise temporal motifs seen in cortical slices in two subsequent days. 
+        In this study by~\citep{haimerl_internal_2019}, an analysis of calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+Cells were selected and ordered with respect to their activity in the first imaging session.
+The black line on top represents the speed of the mouse.
+Futher analysis showe that more than the majority of the cells participating in run sequences on the first day were recruited again in run sequences on the next day.
+(modified from Figure 1-A from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.
 The figure which should have been submitted was a crop on panel A : we have fixed that in the revision

### 2022-12-28 brainsci-2069565_accepted.tex
THC finally

@@
         Detecting event-based motifs using spiking neurons with heterogeneous delays.
-      { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific neuron addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups''.
-      { (\textbf{b})}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay.
-      { (\textbf{c})}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}.
+      { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific neuron addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise spiking motifs.
+      { (\textbf{b})}~The generative model is defined by this set of motifs (here $4$ of them) each defined by different weights at heterogeneous delays (red for excitatory, blue for inhibitory). 
+      { (\textbf{c})}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these spiking motifs~\citep{grimaldi_learning_2022}.
 % such that the addresses are the synapses of these neurons (which we therefore call here synapses).
@@
       { (\textbf{d})}~Knowing the %MDPI: Please change the (a) into (d) in picture. Fixed !
-   results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic example, all spikes are now associated with a PG.
+   results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic example, all spikes are now associated with a motif.
        }

### 2022-12-28 brainsci-2069565_accepted.tex
finished going through all \hl

@@
 
-In~\citep{lagorce_hots_2017}, \textit{\hl{time surfaces}} were introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+In~\citep{lagorce_hots_2017}, time surfaces were introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
 An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
@@
 \citet{pfeil_neuromorphic_2013} implemented STDP on a mixed analog--digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
-Coherence detection on input spikes coming from two sources was obtained with a 50 \hl{$ns$} precision.
+Coherence detection on input spikes coming from two sources was obtained with a 50 $ns$ precision.
 They claimed that this unsupervised learning denoises the input and compensates for variations between neural components.
@@
 This community is based on solid and validated evidence, which is breaking novel grounds thanks to the current technical advances.
-Moreover, we also showed that this community is highly diverse, operating on {\hl{in vivo}}, {\hl{in computo}} and {\hl{in silico}} systems.
+Moreover, we also showed that this community is highly diverse, operating in biology, computational neuroscience or neuromorphic engineering.
 As a consequence, the global effort is still largely scattered, which limits its larger acceptance in~neuroscience.
@@
 \funding{
-This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{\hl{APROVIS3D}}'').
+This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
 % Agence Nationale de la Recherche %HL and
-LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{https://laurentperrinet.github.io/grant/anr-anr/}{\hl{AgileNeuroBot}}''.
+LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}''.
 % LP received funding  
-LP received support from the french government under the France 2030 investment plan, as part of the Initiative d’Excellence d’Aix-Marseille Université – A*MIDEX number AMX-21-RID-025 ``\href{https://laurentperrinet.github.io/grant/polychronies/}{\hl{Polychronies}}''.}
+LP received support from the french government under the France 2030 investment plan, as part of the Initiative d’Excellence d’Aix-Marseille Université – A*MIDEX number AMX-21-RID-025 ``\href{https://laurentperrinet.github.io/grant/polychronies/}{Polychronies}''.}
 
@@
 DeWeese, M.R.; Zador, A.M.
-\newblock {\em Binary Coding in Auditory Cortex}; \hl{Neural Information Processing
-  Systems Foundation:} %MDPI: Please add the location of the publisher (City and Country).
+\newblock {\em Binary Coding in Auditory Cortex}; {Neural Information Processing
+  Systems Foundation:} 
+  \newblock MIT Press: Cambridge, MA, USA,
+  %MDPI: Please add the location of the publisher (City and Country).
    2003.
@@
 DiLorenzo, P.M.; Victor, J.D.
-\newblock {\em Spike {Timing}: {Mechanisms} and {Function}}; CRC Press:  \hl{Boca Raton, FL, USA,} %newly added information, please confirm
+\newblock {\em Spike {Timing}: {Mechanisms} and {Function}}; CRC Press:  {Boca Raton, FL, USA,} %newly added information, please confirm
   2013.
@@
 \newblock {\em Recherches Expérimentales sur les Propriétés et les Fonctions
-  du Système Nerveux, Dans les Animaux Vertébrés}; Balliere, J.-B., Ed.;  \hl{1842.} %MDPI: Please add the name of the publisher and their location.
+  du Système Nerveux, Dans les Animaux Vertébrés}; Hachette {BNF}: {Paris, France,} {1842.} %MDPI: Please add the name of the publisher and their location.
 
@@
 \newblock The {Timing} of {Information} {Transfer} in the {Visual} {System}. In
-  {\em Extrastriate {Cortex} in {Primates}}; Springer: \hl{Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
+  {\em Extrastriate {Cortex} in {Primates}}; Springer: {Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
   1997; pp. 205--241.
@@
 \newblock Computing with spiking neuron networks. In {\em Handbook of Natural
-  Computing}; Springer:  \hl{Berlin/Heidelberg, Germany,} %newly added information, please confirm
+  Computing}; Springer:  {Berlin/Heidelberg, Germany,} %newly added information, please confirm
   2012.
@@
 \newblock In \emph{Proceedings of the Advances in {Neural} {Information} {Processing}
-  {Systems}}; MIT Press:  \hl{Cambridge, MA, USA,} %newly added information, please confirm
+  {Systems}}; MIT Press: Cambridge, MA, USA, %newly added information, please confirm
   2004, Volume~17.
@@
 \newblock \emph{Spike {Trains} of {Retinal} {Ganglion} {Cells} {Viewing} a {Repeated}
-  {Natural} {Movie}}; \hl{Princeton University: Princeton, NJ, USA,} %MDPI: We added the name of the publisher and location. Please confirm.
+  {Natural} {Movie}}; {Princeton University: Princeton, NJ, USA,} %MDPI: We added the name of the publisher and location. Please confirm.
   {2022}.
@@
 \newblock Construction of {Odor} {Representations} by {Olfactory} {Bulb}
-  {Microcircuits}. In {\em Progress in {Brain} {Research}}; Elsevier:  \hl{Amsterdam, The Netherlands,} %newly added information, please confirm
+  {Microcircuits}. In {\em Progress in {Brain} {Research}}; Elsevier:  {Amsterdam, The Netherlands,} %newly added information, please confirm
   2014;
@@
 \newblock Characterization of {Neural} {Responses} with {Stochastic} {Stimuli}.
-In \emph{\hl{The New Cognitive Neurosciences}}, \hl{3rd ed.; Gazzaniga, M., Ed.; MIT Press: Cambridge, MA, USA,} %MDPI: Newly added  information.
+In \emph{{The New Cognitive Neurosciences}}, {3rd ed.; Gazzaniga, M., Ed.; MIT Press: Cambridge, MA, USA,} %MDPI: Newly added  information.
 Please confirm.
@@
   2016.   
-Available online: {\url{https://doi.org/10.6080/K0NC5Z4X}}  \hl{(accessed on date month year).} %MDPI: Please add the access date (format: Date Month Year), e.g., accessed on 1 January 2020.
+Available online: {\url{https://doi.org/10.6080/K0NC5Z4X}}  {(accessed on 25 December 2022).} %MDPI: Please add the access date (format: Date Month Year), e.g., accessed on 1 January 2020.
 
@@
 \newblock Unitary {Event} {Analysis}. In {\em Analysis of {Parallel} {Spike}
-  {Trains}}; Springer: \hl{Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
+  {Trains}}; Springer: {Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
   2010; pp. 191--220.
@@
 \newblock {MouseLand}/Rastermap: {A} Multi-Dimensional Embedding Algorithm.
-  \hl{2020}. %MDPI: Please add website and access date. Please check.
+  {2020}. %MDPI: Please add website and access date. Please check.
+  Available online: {\url{https://github.com/MouseLand/rastermap}}  {(accessed on 25 December 2022).} %MDPI: Please add the access date (format: Date Month Year), e.g., accessed on 1 January 2020.
 
@@
 \newblock In Proceedings of the 2022 {IEEE} {International} {Conference} on
-  {Image} {Processing} ({ICIP}), \hl{Bordeaux, France, 16--19 October} %MDPI: We added the location and date of the conference. Please confirm.
+  {Image} {Processing} ({ICIP}), {Bordeaux, France, 16--19 October} %MDPI: We added the location and date of the conference. Please confirm.
  2022; pp.~3591--3595.
@@
   symmetry between neurons shaped by spike-timing-dependent plasticity.
-\newblock {\em Cogn. Neurodynamics} {\bf \hl{2022} %MDPI: Please add the volume.
+\newblock {\em Cogn. Neurodynamics} {\bf {2022} %MDPI: Please add the volume. There is none for this citation
 }, 1--14.
@@
 \newblock {\em The {Right} {Delay}---{Detecting} {Specific} {Spike} {Patterns}
-  with {STDP} and {Axonal} {Conduction} {Delays}}; \hl{Springer:  Berlin/Heidelberg, Germany,} %MDPI: We added the name of the publisher and location. Please confirm.
+  with {STDP} and {Axonal} {Conduction} {Delays}}; {Springer:  Berlin/Heidelberg, Germany,} %MDPI: We added the name of the publisher and location. Please confirm.
  2011;
@@
 \newblock {\em IEEE Trans. Neural Netw. Learn. Syst.} {\bf
-  \hl{2022} %MDPI: Please add the volume.
+  {2022} %MDPI: Please add the volume. No volume for this citation
 }, 1--13.
@@
 \newblock {\em GWDG-Bericht Nr. 58 Theo Plesser, Volker Macho (Hrsg.)}; {
-  \hl{2003}}; %MDPI: Please add the name of the publisher and their location.
+  {2003}}; %MDPI: Please add the name of the publisher and their location. This is a tech report
  p.~29.
@@
 \newblock {\em Analog {VLSI} {Implementation} of {Neural} {Systems}}; Springer
-  Science \& Business Media:  \hl{Berlin/Heidelberg, Germany,} %newly added information, please confirm
+  Science \& Business Media:  {Berlin/Heidelberg, Germany,} %newly added information, please confirm
   1989.
@@
 \newblock {\em {SpiNNaker}: {A} {Spiking} {Neural} {Network} {Architecture}};
-  Now Publishers: \hl{Norwell, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
+  Now Publishers: {Norwell, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
   2020.
@@
 \newblock In Proceedings of the 2010 {IEEE} {International} {Symposium} on
-  {Circuits} and {Systems} ({ISCAS}), \hl{Paris, France, 30 May--2 June} %MDPI: We added the location and date of the conference. Please confirm.
+  {Circuits} and {Systems} ({ISCAS}), {Paris, France, 30 May--2 June} %MDPI: We added the location and date of the conference. Please confirm.
  2010; pp. 1947--1950.
@@
 \newblock In Proceedings of the 2006 {IEEE} {International} {Symposium} on
-  {Circuits} and {Systems}, \hl{Kos, Greece, 21--24 May} %MDPI: We added the location and date of the conference. Please confirm.
+  {Circuits} and {Systems}, {Kos, Greece, 21--24 May} %MDPI: We added the location and date of the conference. Please confirm.
  2006; pp. 4114--4117.
@@
 \newblock In Proceedings of the 2018 {IEEE}/{CVF} {Conference} on {Computer}
-  {Vision} and {Pattern} {Recognition}, Salt Lake City, UT, USA, \hl{18--23 June} %MDPI: We added the  date of the conference. Please confirm.
+  {Vision} and {Pattern} {Recognition}, Salt Lake City, UT, USA, {18--23 June} %MDPI: We added the  date of the conference. Please confirm.
   2018;
@@
 \newblock In Proceedings of the 2021 {International} {Conference} on
-  {Content}-{Based} {Multimedia} {Indexing} ({CBMI}),  \hl{Lille, France,  28--30 June} %MDPI: We added the location and date of the conference. Please confirm.
+  {Content}-{Based} {Multimedia} {Indexing} ({CBMI}), {Lille, France,  28--30 June} %MDPI: We added the location and date of the conference. Please confirm.
   2021.
@@
 \newblock In Proceedings of the 2019 {IEEE}/{CVF} {International} {Conference}
-  on {Computer} {Vision} ({ICCV}), \hl{Seoul, Republic of Korea, 27 February} %MDPI: We added the location and date of the conference. Please confirm.
+  on {Computer} {Vision} ({ICCV}), {Seoul, Republic of Korea, 27 February} %MDPI: We added the location and date of the conference. Please confirm.
  2019; pp. 7243--7252.
@@
 \newblock In Proceedings of the 2012 {IEEE} {Asia} {Pacific} {Conference} on
-  {Circuits} and {Systems},  \hl{Kaohsiung, Taiwan, 2--5 December} %MDPI: We added the location and date of the conference. Please confirm.
+  {Circuits} and {Systems},  {Kaohsiung, Taiwan, 2--5 December} %MDPI: We added the location and date of the conference. Please confirm.
  2012; pp. 304--307.
@@
 \newblock In Proceedings of the 2014 {IEEE} {International} {Symposium} on
-  {Circuits} and {Systems} ({ISCAS}), \hl{Melbourne, VIC, Australia, 1--5 June} %MDPI: We added the location and date of the conference. Please confirm.
+  {Circuits} and {Systems} ({ISCAS}), {Melbourne, VIC, Australia, 1--5 June} %MDPI: We added the location and date of the conference. Please confirm.
  2014; pp. 457--460.
@@
 \newblock In Proceedings of the The 2013 {International} {Joint} {Conference}
-  on {Neural} {Networks} ({IJCNN}), \hl{Dallas, TX, USA, 4--9 August} %MDPI: We added the location and date of the conference. Please confirm.
+  on {Neural} {Networks} ({IJCNN}), {Dallas, TX, USA, 4--9 August} %MDPI: We added the location and date of the conference. Please confirm.
  2013; pp. 1--5.

### 2022-12-28 2022-12-28_coverletter.tex
Update 2022-12-28_coverletter.tex

@@
 %----------------------------------------------------------------------------------------
-Many thanks for accepting our manuscript entitled ``Precise spiking motifs in neurobiological and neuromorphic data'' for your consideration as a review paper in \emph{Brain Sciences}.
-We are very grateful to the reviewers for their very insightful remarks and we have now responded to all their remarks.
-
+Many thanks for accepting our manuscript entitled ``Precise spiking motifs in neurobiological and neuromorphic data'' for publication as a review paper in \emph{Brain Sciences}.
 We have accepted all changes suggested at the proofreading stage and uploaded our files on the MDPI website.
-
-\textbf{ Note that in agreement with the editorial office, we have added Jean-Nicolas Jérémie as an author, as his contributions were substantial for writing this review.}
 
@@
 
-This is an invited Feature Paper, with a full waiver on the APC. In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source
+This is an invited Feature Paper, with a full waiver on the APC. In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source.
 

### 2022-12-28 brainsci-2069565_accepted.tex
PROOFRED VERSION

@@
 %
-% ACCEPTED VERSION
+% ACCEPTED VERSION - PROOFRED
 %
@@
 \caption{
-        An example of a precise temporal motifs seen in cortical slices in two subsequent days. 
-        In this study by~\citep{haimerl_internal_2019}, an analysis of calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+        An example of a precise temporal motif observed in two subsequent days. 
+        In this study by~\citep{haimerl_internal_2019}, an analysis of calcium fluorescence (heatmap) of hippocampal CA1 neurons participating to run sequences in consecutive imaging sessions shows repetitions of precise spiking motifs with a time scale of the order of seconds.
 Cells were selected and ordered with respect to their activity in the first imaging session.
 The black line on top represents the speed of the mouse.
-Futher analysis showe that more than the majority of the cells participating in run sequences on the first day were recruited again in run sequences on the next day.
-(modified from Figure 1-A from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.
+Futher analysis showed that more than the majority of the cells participating in run sequences on the first day were recruited again in run sequences on the next day.
+Modified from Figure 1-A from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}.}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.
 The figure which should have been submitted was a crop on panel A : we have fixed that in the revision
@@
     \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
+    \includegraphics[width=0.480\linewidth]{figures/THC_1c.pdf}
     \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
@@
 \funding{
-This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
+This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant number ANR-19-CHR3-0008-03 ``\href{http://aprovis3d.eu/}{APROVIS3D}''.
 % Agence Nationale de la Recherche %HL and

### 2022-12-28 manuscript_rev1.tex
pushed PROOFRED

--- 2022-12-28 brainsci-2069565_accepted.tex
@@
 %
-% ACCEPTED VERSION
-%
-% https://github.com/SpikeAI/2022_polychronies-review
+% REV1
+%
+% https://www.overleaf.com/project/633bf90497b5b325a3ff692e
 % https://framateam.org/spikeai/channels/review_polychrony
@@
 \documentclass[brainsci, %arts,
-               review,accept,pdftex,moreauthors
+               review,submit,pdftex,moreauthors
                %,  draft
@@
 \articlenumber{0}
-\pubyear{2023}
+\pubyear{2022}
 \copyrightyear{2022}
 %\externaleditor{Academic Editor: Firstname Lastname}
-\datereceived{16 November 2022} 
-\daterevised{20 December 2022 } % Only for the journal Acoustics
-\dateaccepted{23 December 2022 } 
+\datereceived{} 
+%\daterevised{} % Only for the journal Acoustics
+\dateaccepted{} 
 \datepublished{} 
@@
 \usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
+
 \newcommand{\ms}{\si{\milli\second}}%
@@
 \usepackage[finalnew]{trackchanges}
-\addeditor{MDPI}
+\addeditor{Antoine}
+\addeditor{Amélie}
+\addeditor{Camille}
+\addeditor{Jean}
+\addeditor{JN}
+\addeditor{Laurent}
 
@@
 % Full title of the paper (Capitalized)
-\Title{Precise Spiking Motifs in Neurobiological and Neuromorphic~Data}
+\Title{Precise spiking motifs in neurobiological and neuromorphic data}
 %
 % MDPI internal command: Title for citation in the left column
-\TitleCitation{Precise Spiking Motifs in Neurobiological and Neuromorphic Data}
+\TitleCitation{Precise spiking motifs}
 %
@@
 % Authors, for the paper (add full first names)
-\Author{Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
+\Author{
+Antoine Grimaldi $^{1}$% $^{1,\dagger,\ddagger}$
 \orcidA{}, 
@@
 Jean-Nicolas Jérémie $^{1}$\orcidD{},
-Jean Martinet $^{2}$\orcidE{}
-and~Laurent U Perrinet $^{1,}$*\orcidF{}}
+Jean Martinet $^{2}$\orcidE{},
+and Laurent Perrinet $^{1,}$\orcidF{}*}
 %
@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent U Perrinet}
+\AuthorNames{Antoine Grimaldi, Amélie Gruel, Camille Besnainou, Jean-Nicolas Jérémie, Jean Martinet and Laurent Perrinet}
 %
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Grimaldi, A.; Gruel, A.; Besnainou, C.; Jérémie, J.-N.; Martinet, J.; Perrinet, L.}
+\AuthorCitation{Grimaldi, A.; Gruel, A.; Besnainou, C.; Jérémie, J.-N.;Martinet, J.; and Perrinet, L.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 \address{%
-$^{1}$ \quad INT UMR 7289, Aix Marseille Univ, CNRS, 27 Bd Jean Moulin, 13005 Marseille, France; antoine.grimaldi@univ-amu.fr (A.G.); camille.besnainou@etu.univ-amu.fr (C.B.); jean-nicolas.jeremie@univ-amu.fr (J.-N.J.), laurent.perrinet@univ-amu.fr (L.U.P.)
- \\
-$^{2}$ \quad SPARKS,  Côte d'Azur, CNRS, I3S, 2000 Rte des Lucioles,
-06900 Sophia-Antipolis, France; amelie.gruel@i3s.unice.fr (A.G.); jean.martinet@univ-cotedazur.fr (J.M.)
+$^{1}$ \quad INT UMR 7289, Aix Marseille Univ, CNRS; 27 Bd Jean Moulin, 13005 Marseille, France\\
+$^{2}$ \quad SPARKS,  Côte d'Azur, CNRS, I3S; 2000 Rte des Lucioles, 06900 Sophia-Antipolis, France
 }
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{Why do neurons communicate through spikes?
+\abstract{
+Why do neurons communicate through spikes?
 By definition, spikes are all-or-none neural events which occur at continuous times.
-In other words, spikes are on one side binary, existing or not without further details, and on the other, can occur at any asynchronous time, without the need for a centralized clock.
+In other words, spikes are on one side binary, existing or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock.
 This stands in stark contrast to the analog representation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks.
 As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the profusion of recorded data.
-With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption---a major societal issue in the era of the digital economy and global warming.
+With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption --- a major societal issue in the era of the digital economy and global warming.
 In this review paper, we provide evidence from biology, theory and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
@@
 % Keywords
-\keyword{\textls[-15]{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic} engineering; heterogeneous delays; spiking motifs; polychronization} 
+\keyword{spikes; asynchronous computing; neurobiology; computational neuroscience; neuromorphic engineering; heterogeneous delays; spiking motifs; polychronization.)} 
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction: Importance of Precise Spike Timings in the Brain}\label{sec:time}
+\section{Introduction: Importance of precise spike timings in the brain}\label{sec:time}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Is There a Neural Code?}
+\subsection{Is there a neural code?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings observed in neurobiological experiments.
-One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e., an explicit representation \add[Antoine]{of cognitive processes} within the neural activity.
+One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e.
+an explicit representation \add[Antoine]{of cognitive processes} within the neural activity.
 Nevertheless, we will use this terminology in all generality to denote the existence of a structure in neural activity.
-In this respect, it is \change[Antoine]{known}{reasonable to declare} that neural activity may be related to specific measurable variables\change[Antoine]{, and due to Galvani's}{.
+In this respect, it is \change[Antoine]{known}{reasonable to declare} that neural activity may be related to specific measurable variables\change[Antoine]{, and since Galvani's}{.
 Since Galvani's} experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}).
@@
 In this study, the frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb.
-In the scope of this article, we focus mainly on these spikes~\change[Antoine]{which are brief (about one millisecond) impulses that propagate along the axons of neurons.
-First, these have the particular property of being essentially binary in their amplitude, i.e., they are prototypical, ``all-or-none''}{in vertebrate systems.
-They can be described as brief (about one millisecond) and prototypical, i.e., ``all-or-none'', impulses that propagate along the axons of neurons.}
+In the scope of this article, we focus mainly on these spikes\change[Antoine]{which are brief (about one millisecond) impulses that propagate along the axons of neurons.
+First, these have the particular property of being essentially binary in their amplitude, i.e.
+they are prototypical, ``all-or-none''}{in vertebrate systems.
+They can be described as brief (about one millisecond) and prototypical, i.e.
+``all-or-none'', impulses that propagate along the axons of neurons.}
  \change[Antoine]{Typically, action potentials are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons.}{Typically, a postsynaptic neuron receives incoming spikes from other afferent neurons on the arborized ``input'' dendrite.
 The integration of these spikes by the dendritic tree and the soma of the postsynaptic neuron results in the modification of its membrane potential that possibly leads to the emission of an action potential along its ``output'' axon to reach efferent neurons.} \add[Antoine]{Except notably in the retina where neurons communicate with graded potentials}~\citep{gouras_graded_1960}\add[Antoine]{, it is assumed that spike trains are the main component of the neural code.}
-\remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} used the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
+\remove[Antoine]{The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb(Adrian citation), and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code.} \change[Antoine]{Indded, neurophysiologists typically}{Until recently, most neurophysiologists} use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons.
 This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.
@@
 %Indeed, the response of one biological neuron largely depends on the precise timing of the sequence of incoming spikes that reach its soma~\citep{paugam-moisy_computing_2012}.
-In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this event-based representation observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
-For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{of the rate coding hypothesis that relies only on the firing rate} by considering \change[Antoine]{a representation based on repetitions of spiking motifs at precise times of occurrence.}{representations based on the precise timing of single spikes.} \remove[Antoine]{This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.} Additionally, numerous studies have demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
+In comparison to a classical \change[Antoine]{analogical}{analog} vector of inputs, this \emph{event-based representation} observed in the neural code is essential in understanding information processing~\citep{carr_processing_1993}.
+For instance, it expands the capabilities of representations \change[Antoine]{(which rely on the firing rate)}{of the rate coding hypothesis that relies only on the firing rate} by considering \change[Antoine]{a representation based on repetitions of spiking motifs at precise times of occurrence.}{representations based on the precise timing of single spikes.} \remove[Antoine]{This hypothesis is directly inspired by neurobiological observations and expands the capabilities of analog representations (which are based on spiking firing rates) by considering a representation based on instances of spiking motifs at precise times of occurrence.} Additionally, numerous studies demonstrated the importance of precise timing in neural population activity~\citep{davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004,gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2003,carr_circuit_1990}.
 All these findings, and more~\citep{bohte_evidence_2004,dilorenzo_spike_2013}, highlight the importance of the temporal aspect of the neural code and further suggest the existence of precise spatio-temporal spiking motifs in the input which excites neurons.
-A mathematical formalization would be particularly well-suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
+A mathematical formalization would be particularly well suited to neuromorphic computing~\citep{roy_towards_2019}, and would allow for the supervised or self-supervised learning of such motifs in any event-driven data.
 Crucially, validating this hypothesis would also be crucial in our understanding of neural processes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Dynamics of Vision and Consequences on the Neural Code}%Precise timing in the neural code enables efficient vision}
+\subsection{Dynamics of vision and consequences on the neural code}%Precise timing in the neural code enables efficient vision}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Let us start with a focus on the state-of-the-art of the role of dynamics in vision.
-Broadly speaking, vision is the set of processes that allow us \change[Antoine]{to make sense of the luminous world}{to make sense of the world through luminous signals}, and is an intensively studied field in neuroscience, particularly with respect to deciphering the neural code.
+Broadly speaking, vision is the set of processes that allow us \change[Antoine]{to make sense of the luminous world}{to make sense of the world through luminous signals}, and is an intensively studied field in neuroscience, in particular with respect to deciphering the neural code.
 In most mammals, light enters the eye to induce neural activity on the retina, which maintains a certain similitude between the topology of external visual space and its representation on the retina, called retinotopy.
@@
 
-%----------------------------%
-
-%----------------------------%
 In particular, there have been some remarkable findings when studying the dynamics of vision.
@@
 \change[Antoine]{His}{Their} experiments consisted in asking subjects to categorize images that do or do not contain animals~\citep{thorpe_speed_1996}.
-The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that, a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
+The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges with a very short latency in neural activity.
 These results have been extended to several species, including primates.
-Different experimental protocols have shown, for example, that the motor response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
-This fast processing correlates with the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and decoding via the EEG if the observer can detect, for example, the presence of an animal~\citep{keysers_speed_2001}.
+Different experimental protocols have shown for example that the motor response could be extremely fast (of the order of $120~\ms$) when the task was to perform a saccade~\citep{kirchner_ultra-rapid_2006}.
+This fast processing correlates with the surprising experiments of fast serial detection, which consists in presenting a fast succession of different images and decoding via the EEG if the observer can detect for example the presence of an animal~\citep{keysers_speed_2001}.
 As expected, the performances decrease progressively as the frequency of presentation of the images increases.
@@
 An important consequence of this speed of processing of vision is that it implies that processing is carried out using only very few spikes per layer.
-As a comparison, the latencies in macaque monkeys are approximately as follows: Retina, 20--40~$\ms$; V1, 40--60~$\ms$; IT, 80--100~$\ms$; MC, 140--190~$\ms$; and to finger muscles, 180--260~$\ms$.
-\add[Laurent]{Note that, since maximal conduction speeds are roughly constant, theses latencies are comparable to that found in humans, with a ratio given by the physical size of the whole system.} It follows that if we consider that a behavioral response occurrs in only $200~\ms$, it would involve about ten processing stages along the ``forward'' pathways of the visual system.
+As a comparison, the latencies in macaque monkeys are about: Retina, $20-40~\ms$; V1, $40-60~\ms$; IT, $80-100~\ms$; MC, $140-190~\ms$; to finger muscles, $180-260~\ms$.\add[Laurent]{Note that, since maximal conduction speeds are roughly constant, theses latencies are comparable to that found in humans, with a ratio given by the physical size of the whole system.} It follows that if we consider that a behavioral response occurra in only $200~\ms$, it would involve about ten processing stages along the ``forward'' pathways of the visual system.
 Such processes were indeed efficiently reproduced in feed-forward models trained with back-propagation~\citep{serre_feedforward_2007,jeremie_ultrafast_2022}.
-At the same time, it was demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
-This figure is inspired by similar schematics performed for monkeys in~\citep{thorpe_seeking_2001}.
+At the same time, it has been demonstrated that one spike requires a significant amount of time (about $10~\ms$) to be conducted from one layer to the next~\citep{nowak_timing_1997,thorpe_seeking_2001}.
+This figure is inspired by similar schematics done for monkeys in~\citep{thorpe_seeking_2001}.
 As a consequence, these results suggest that ``like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information''~\citep{rucci_temporal_2018}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{How Precise Spike Timing May Encode Vectors of Real Values}
+\subsection{How precise spike timing may encode vectors of real values}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\add[Laurent]{Let us now focus on one processing step along a cortical pathway. }
+\add[Laurent]{Let's now focus on one processing step along a cortical pathway. }
 % reliability of spike timings
-Sensory data are most often represented by continuous values, such as the energy produced by a flow of photons that hits the different photoreceptors of the retina.
+Sensory data is most often represented by continuous values, such as the energy produced by a flow of photons that hits the different photoreceptors of the retina.
 How may such information be encoded in neuronal activity?
-The analysis of generic raster plots reveals particular traits that hint at the role of precise timing.
-For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it, at first sight, inconsistent with the temporal integration of firing rate.
-Remarkably, it was observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: while the first spike is aligned to stimulus' time, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{bryant_spike_1976,mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
+The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing.
+For instance, the firing rate of cortical cells in awake monkeys is highly irregular~\citep{softky_highly_1993}, which makes it at first sight inconsistent with the temporal integration of firing rate.
+Remarkably, it has been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible: while the first spike is aligned to stimulus' time, the subsequent spike times tend to diffuse for independent repetitions of the stimulation~\citep{bryant_spike_1976,mainen_reliability_1995} (see Figure~\ref{fig:mainen}).
 However, if that same neuron is now driven by a \emph{frozen} noise, that is, a highly dynamic signal which is repeatedly replayed from trial to trial, then the output spikes are highly reproducible (for a review, see~\citep{ermentrout_reliability_2008}).
-This is consistent with the differential role of different stimulus frequencies (for instance, the gamma range around $80~\si{\Hz}$) on the reliability of the spike timing reported in~\citep{nowak_influence_1997}: ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
+This is consistent with the differential role of different stimulus frequencies (for instance the gamma range around $80~\si{\Hz}$) on the reliability of spike timing reported in~\citep{nowak_influence_1997} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies.
 However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''
 
+%----------------------------%
 \begin{figure}[H]
-%\centering
-\includegraphics[width=.75\textwidth]{figures/visual-latency.pdf}
+  \centering
+  \includegraphics[width=.7\textwidth]{../2022-12-20_polychrony_review_REV1/figures/visual-latency.pdf}
   % thorpe
-% \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
-\caption{
-  Latency of the different processing steps along the human visual pathway.Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
-Different areas are schematically represented by ellipses, and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
-In order, the retina is first activated (20--40~$\ms$), then the thalamus and the primary visual cortex (V1, 60--90~$\ms$).
-This visual information  projects to the temporal lobe to reach the infero-temporal area (IT, 150~$\ms$) for object recognition.
-It then reaches the prefrontal cortex (PFC, $180~\ms$), which modulates decision making to evoke the motor cortex (MC, $220~\ms$) which may mediate an action.
-This is eventually relayed through the spinal cord to trigger finger muscles, with latencies of about 280--400~$\ms$.}\label{fig:thorpe}
+  % \includegraphics[width=.7\textwidth]{../2022-12-20_polychrony_review_REV1/figures/visual-latency-estimate.jpg}%-estimate.jpg}
+  \caption{{\bf Latency of the different processing steps along the human visual pathway.} Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
+Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
+In order, the retina is first activated ($20-40~\ms$), then the thalamus and the primary visual cortex (V1, $60-90~\ms$).
+This visual information  projects to the temporal lobe to reach the infero-temporal area (IT, $150~\ms$) for object recognition.
+It then reaches the prefrontal cortex (PFC, $180~\ms$) which modulates decision making to evoke the Motor Cortex (MC, $220~\ms$) which may mediate an action.
+This is eventually relayed through the spinal cord to trigger finger muscles, with latencies about $280-400~\ms$.}\label{fig:thorpe}
   \end{figure}
+  %----------------------------%
   \vspace{-6pt}
+
+%----------------------------%
+% C_MainenSejnowski1995_Perrinet.ipynb
 \begin{figure}[H]
-%\centering
-\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
-\caption{
-  Reproducibility of the spiking response of a neuron. 
-The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
+  \centering
+  \includegraphics[width=\textwidth]{../2022-12-20_polychrony_review_REV1/figures/replicating_MainenSejnowski1995.pdf} 
+  \caption{\textbf{Reproducibility of the spiking response of a neuron.
+}The timing of the spikes produced following the repetition of a step stimulus is less reproducible than that to a noisy stimulus.
 The stimulus current value over time for a step stimulus (top left) and for a noisy one (top right).
-Trial repetitions of a leaky integrate-and-fire neuron stimulated by the stimulus on the upper row (middle row).
+Trial repetitions of a Leaky Integrate-and-Fire neuron stimulated by the stimulus on the upper row (middle row).
 Membrane potential is represented by dark blue color when light yellow colors when depolarized) and quantified by the average firing rate across trials (lower row).
-While this seems paradoxical at first sight, it highlights the consequence of using the same frozen noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
-See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} %MDPI: Please confirm if the link is unnecessary and can be removed.
-The following highlights are the same.
-for a replication of the results from~\citep{mainen_reliability_1995} using a simple LIF model.}\label{fig:mainen}
+While this seems paradoxical at first sight, it highlights the consequence of using the same \emph{frozen} noise at each repetition and highlights the highly reproducible pattern of spikes when it is driven by a highly dynamic input.
+See this~\href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995_Perrinet.ipynb}{notebook} for a replication of the results from~\citep{mainen_reliability_1995} using a simple LIF model.}\label{fig:mainen}
   \end{figure}
+  %----------------------------%
 
@@
 The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
-Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center--surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
-This evidence found in the retina can be extended to other areas, such as the visual cortex.
-%----------------------------%
-% C_MainenSejnowski1995_Perrinet.ipynb
-
-%----------------------------%
+Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+This evidence found in the retina can be extended to other areas such as the visual cortex.
 
@@
 In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}.
-First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of the stimulus onset are considered, for instance, for sound localization~\citep{chase_first-spike_2007}.
+First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered, for instance for sound localization~\citep{chase_first-spike_2007}.
 Note also that timing is not entirely sensory or internal, but can be used as a general neural coding principle.
-In~\citep{safaie_turning_2020}, for instance, they found that the ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
+In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines.
 Timing, at least in animals, may thus be fundamentally embodied and situated.''  Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}.
@@
 For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition or~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.
-However, the performance of SNNs is still lagging compared to that of analog networks, and the question of the advantage of using spikes in machine learning and computer vision remains open.
-\change[Antoine]{Efficiently implementing this type of modeling is extremely important with respect to the development of a new generation of event-based cameras}{Improvements in this new generation of Artificial Neural networks (ANN) would bring major advances in terms of efficient computations in machine learning.
+However, the performance of SNNs is still lagging compared to that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
+\change[Antoine]{Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of event-based cameras}{Improvements in this new generation of Artificial Neural networks (ANN) would bring major advances in terms of efficient computations in machine learning.
 They would benefit in particular to a new generation of cameras called Silicon Retinas}~\citep{gallego_event-based_2022} (see Section~\ref{sec:neuromorphic}).
@@
 
-\change[Antoine]{To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.}{~Even if technology lags far behind biology, this introduction demonstrates the importance of timing in neural processes, and we will further review the role of precise spike timing in neural assemblies.
+\change[Antoine]{To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.}{Even if technology lags far behind biology, this introduction demonstrates the importance of timing in neural processes and we will further review the role of precise spike timing in neural assemblies.
 We start by reviewing the different hypotheses that aim at deciphering the neural code with spatio-temporal spiking motifs.
-After listing some biological evidence for the use of precise spike timing, we review some computational models and neuromorphic technics that add this temporal dimension to their computations.}
+After listing some biological evidence for the use of precise spike timing we review some computational models and neuromorphic technics that add this temporal dimension to their computations.}
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Role of Precise Spike Timing in Neural Assemblies}
+\section{Role of precise spike timing in neural assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Among a variety of strategies to communicate, so far we have also evidence that neurons can use correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}.
-\add[Antoine]{In this first section, we introduced the notion of rate coding and demonstrated that spike timing can also carry information.
-Scientists found experimental evidence for various hypotheses of neural representations such as population coding}~\citep{maunsell_functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase--firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
+\add[Antoine]{In this first section, we have introduced the notion of rate coding and demonstrated that spike timing can also carry information.
+Scientists found experimental evidence for various hypotheses of neural representations like population coding}~\citep{maunsell_functional_1983}\add[Antoine]{, time-to-first-spike coding}~\citep{gollisch_rapid_2008}\add[Antoine]{, phase-of-firing coding}~\citep{montemurro_phase--firing_2008}\add[Antoine]{, correlation coding}~\citep{decharms_primary_1996}\add[Antoine]{ or sparse coding}~\citep{vinje_sparse_2000}\add[Antoine]{.
 In the scope of this review, we infer that spike trains are composed of repeating spiking motifs and we focus on precise spatio-temporal representations composed of a motif of spikes defined precisely in time and in the presynaptic address space.
-In all generality, this representation can encompass all the previous ones, except for rate coding, which is not defined locally in time.
-In this section, we choose to describe different hypotheses, making use of spatio-temporal patterns of spikes that can be propagated among neural assemblies.}
+In all generality, this representation can encompass all the previous ones except for rate coding which is not defined locally in time.
+In this section, we choose to described different hypotheses making use of spatio-temporal patterns of spikes that can be propagated among neural assemblies.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{One First Hypothesis: Synchronous Firing in Cell Assemblies}
+\subsection{One first hypothesis: synchronous firing in cell assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-In his book \emph{Corticonics},~\citet{abeles_corticonics_1991} queried as to whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
+In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns.
 The book gradually leads the reader from the macroscopic cortical anatomy and standard electro-physiological properties of single neurons to neural network models.
 While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code~\citep{abeles_role_1982,paugam-moisy_computing_2012}.
-The book then demonstrates that neural assemblies could form so-called ``synfire chains,'' that is, showing the emergence of synchronous activity on subsets of neurons, which could be propagated in a stable fashion.
-More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}: \add[Antoine]{``cells that fire together wire together''}.
+The book then demonstrates that neural assemblies could form so-called ``synfire chains,'' that is, showing the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion.
+More broadly, the idea of using the synchrony of co-activation in a cell assembly is reminiscent of the hypothesis that was formalized by~\citet{hebb_organization_1949}: \add[Antoine]{"cells that fire together wire together"}.
 Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}.
@@
 However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active, allowing it to more efficiently drive cortical neurons}{Synchronicity is also an interesting proposition to explain how the relatively weak thalamo-cortical synapses are able to drive cortical neurons.
-Among different explanations, including travelling waves}~\citep{ben-yishai_traveling_1997}\add[Antoine]{, a synchronous activity at the synaptic level may be sufficient to elicit activity in the cortex}~\citep{bruno_cortex_2006}.
-
-Theoretically\add[Antoine]{, even if the vertebrate's neural system is not likely to be modeled only by such deterministic connectivity}~\citep{deneve_bayesian_2004, ballard_dual_2011}, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of a similar size.
+Among different explanations including travelling waves}~\citep{ben-yishai_traveling_1997}\add[Antoine]{, a synchronous activity at the synaptic level may be sufficient to elicit activity in the cortex.}~\citep{bruno_cortex_2006}.
+
+Theoretically\add[Antoine]{, even if vertebrate's neural system is not likely to be modeled only by such deterministic connectivity}~\citep{deneve_bayesian_2004, ballard_dual_2011}, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size.
 Each neuron of one group is connected by an excitatory synapse to the next.
@@
 A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}.
-A crucial aspect of this emergence is explained by the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and, in particular, the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
-This balance was, for instance, modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling so-called push--pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
+A crucial aspect of this emergence is explained by the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}.
+This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling so-called push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}.
 Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}.
@@
 %\centering
-%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
+%\includegraphics[width=.7\textwidth]{../2022-12-20_polychrony_review_REV1/figures/fnins-07-00011-g004.jpg}
 %\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
@@
 %----------------------------%
-\begin{figure}[H]
-%\centering
-\includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
-\caption{
-  Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.
-  The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../2022-12-20_polychrony_review_REV1/figures/Diesmann_et_al_1999.pdf}
+\caption{\textbf{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.} The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
 Five pulses with decreasing jitter are generated in the first group around times $10$, $30$, $50$, $70$ and $90$~\ms~(with jitters given by a standard deviation which linearly decreases from $5$ to $1$~\ms).
@@
 
-Attempts have been made to detect such synfire chains in neurobiological data.
-\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Indeed, simultaneously recorded activities of neurons in the primary motor cortex of monkeys exhibited context-dependent, rapid changes in the patterns of coincident action potentials~\citep{riehle_spike_1997,grammont_precise_1999}.
-It is now commonly accepted that the planning and execution of movements are based on distributed processing by neural populations in motor cortical areas, yet it is less clear \remove[Antoine]{, though,} how these populations organize dynamically to cope with the momentary computational demands.
-In~\citep{brette_computing_2012}, the author proposed a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical ANNs.
+Attempts have been made to detect such synfire chains in neurobiological data.~\citet{schrader_detecting_2008} envisioned that ``sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Indeed, simultaneously recorded activities of neurons in the primary motor cortex of monkeys exhibited context-dependent, rapid changes in the patterns of coincident action potentials~\citep{riehle_spike_1997,grammont_precise_1999}.
+It is now commonly accepted that planning and execution of movements are based on distributed processing by neural populations in motor cortical areas yet it is less clear \remove[Antoine]{, though,} how these populations organize dynamically to cope with the momentary computational demands.
+In~\citep{brette_computing_2012}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical ANNs.
 It relies on the simple remark that a serie of repeated coincidences is in itself an invariant.
@@
 The synchronicity will generate rhythms at different ranges of frequencies, with spikes arriving at peak susceptibility (top of a cycle) or down.
-Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion, i.e., the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
+Such a theory has surprisingly been validated in EEG recordings to explain for example the continuous wagon-wheel illusion, i.e.
+the perceived reversal of the rotational movements of the spikes in a rotating wheel~\citep{vanrullen_continuous_2006}.
 More generally, it can be shown that the phase of alpha oscillations (about $10$ Hz) is causally linked with modulations of cortical excitability and with visual perception~\citep{dugue_phase_2011}.
@@
 %
-\subsection{A Further Hypothesis: Travelling Waves}
+\subsection{A further hypothesis: travelling waves}
 %
@@
 Using intracellular recordings, it was shown that the response to a focal visual activation would elicit a latency basin, that is, a graded onset of the neural response from the most activated to neighboring neurons~\citep{bringuier_horizontal_1999}.
-In particular, it was shown that the network of the so-called horizontal connections within a cortical area is typically not myelinated, and that this latency basin would be determined by the propagation speed within that area.
+In particular, it was shown that the network of the so-called horizontal connections within a cortical area is typically not myelinated and that this latency basin would be determined by the propagation speed within that area.
 For generic visual scenes, these processes would generate a complex interplay between the dynamics of the sensory signal and the spatio-temporal determinants of these interactions.
@@
 
-Propagating waves in the neuronal response occur in many excitable media and were found in neural systems, for instance, in the retina\note[Antoine]{added ref to Feller}~\citep{feller_dynamic_1997} or the neocortex~\citep{bienenstock_model_1995}.
+Propagating waves in the neuronal response occur in many excitable media and were found in neural systems, for instance in the retina\note[Antoine]{added ref to Feller}~\citep{feller_dynamic_1997} or the neocortex~\citep{bienenstock_model_1995}.
 While propagating waves are clearly present under anesthesia, whether they also appear during awake and conscious states remained unclear until recent discoveries.
@@
 More recently, novel multi-unit recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}.
-The authors reviewed these findings by considering the mechanisms by which travelling waves are generated, and evaluated their possible roles in cortical function.
+Authors review these findings by considering the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function.
 In particular, spontaneous travelling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}.
-Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally generated noise.
+Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise.
 However, fluctuations of spontaneous activity, often organized as travelling waves, shape stimulus-evoked responses and perceptual sensitivity.
@@
 Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings.
-In that paper, the authors describe a large-scale computational model with topographically organized connectivity and conduction delays relevant to biological scales.
+In that paper, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales.
 They find that spontaneous travelling waves are a general property of these networks.
@@
 Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
-Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed in biology.
-Such waves also occur in motor areas and~\citet{linden_movement_2022} have recently presented ensemble recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic~behavior.
+Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \emph{in vivo}.
+Such waves also occur in motor areas and~\citet{linden_movement_2022} have recently presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
 
 Interestingly, it can be shown that these travelling waves could have a measurable impact on the activity of the visual cortex.
-This was illustrated in a recent study investigating the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
+This has been illustrated in a recent study investigating the long-range apparent motion effect (lrAM)~\citep{chemla_suppressive_2019}.
 The lrAM is the simple phenomenon of perceiving a smooth motion when showing two dots in a temporal sequence and in relatively close visual proximity.
@@
 In this study, the authors used voltage-sensitive dye imaging to record the activity of the primary visual cortex of macaque monkeys to the presentation of the pair of dots presented independently or in conjunction.
-A probabilistic modelling showed that the activity of the joint presentation induced a suppressive wave in the direction opposed to the perceived direction, shaping the formation of a wave of propagation travelling at a speed compatible with the perceived motion.
+A probabilistic modelling has shown that the activity of the joint presentation induced a suppressive wave in the direction opposed to the perceived direction, shaping the formation of a wave of propagation travelling at a speed compatible with the perceived motion.
 A computational model validated the hypothesis that this process could be mediated by diffusion in the horizontal layers connecting the different locations within this cortical area.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A Rediscovered Hypothesis: Precise Spiking Motifs in Cell Assemblies}
+\subsection{A rediscovered hypothesis: precise spiking motifs in cell assemblies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For instance, statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it.
-In particular, this was put in evidence by analyzing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}, which shows the precise spatio-temporal differences in this recurrently connected assembly.
+In particular, this was put in evidence by analyzing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses~\citep{pillow_spatio-temporal_2008}, which shows precise spatio-temporal differences in this recurrently connected assembly.
 The different aspects of information in the data are evaluated by a decoding strategy, highlighting the role of correlations.
@@
 Recent observations may suggest that neural groups or ensembles, rather than individual neurons, are emergent functional units of cortical activity.
-~\citet{miller_visual_2014} showed that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
+~\citet{miller_visual_2014} show that whereas intrinsic ensembles recur at random time intervals, visually evoked ensembles are time-locked to stimuli.
 Experiments are performed using two-photon calcium imaging of populations of neurons from the primary visual cortex of awake mice during visual stimulation and spontaneous activity.
@@
 
-From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008}.
+From another viewpoint, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007} or hippocampal sequences~\citep{pastalkova_internally_2008}.
 %
@@
 %We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
-Going further, researchers found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \emph{in vivo} %MDPI: We removed the italics.
-Please confirm this revision.
-These were kept as they correspond to latin locutions
- and \emph{in vitro}.
+Going further, researchers have found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \emph{in vivo} and \emph{in vitro}.
 These patterns repeat after minutes, maintaining millisecond accuracy.
-Indeed,~\citet{ikegaya_synfire_2004} demonstrated that in cortical activity, one can find a repetition of several motifs in spike activity (duration around 1 s +/$-$ 0.5 s, some events in motifs are of similar size but sometimes absent).
-These sequences can be specific to a particular layer or column, can be synchronized with network activity oscillation, and can involve several cells.
-They also demonstrated that these sequences can form super sequences, so-called ``cortical songs''.
+Indeed,~\citet{ikegaya_synfire_2004} demonstrated that in cortical activity, one can find a repetition of several motifs in spike activity (duration around 1s +/- .5 s, some events in motifs are of similar size but sometimes absent).
+These sequences can be specific of a particular layer or column, can be synchronized with network activity oscillation, and can involve several cells.
+They also demonstrated that these sequences can form super sequences, so-called \emph{cortical songs}.
 It consists of the assembly of several sequences which repeat in a specific order with a compressed timing.
-This spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3--10 cells out (of about 900), and lasting up to 3 s.
+\emph{In vivo} spontaneous activity also reveals repeating sequences: about 3000 sequences, each involving 3-10 cells out (of about 900), and lasting up to 3 seconds.
 Sequences have specific topographic structures, in some cases involving only a particular layer or a vertical column of cells or cells located in a cluster, and are associated with a structured spatial organization of the neurons that formed them.
@@
 % \centering
-% \includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
+% \includegraphics[width=.6\linewidth]{../2022-12-20_polychrony_review_REV1/figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
 % \caption{Fig.
 1.
-from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity \textit{in vitro} and \textit{in vivo}.
+from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity \emph{in vitro} and \emph{in vivo}.
 (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
@@
 The top traces show the motifs superimposed on each other (blue, green, and red), the middle traces show these same traces individually, and the bottom traces show temporally magnified regions of the motifs (a to c).
-(C) Repeated sequences of intracellular current-clamp recordings \textit{in vivo}.
+(C) Repeated sequences of intracellular current-clamp recordings \emph{in vivo}.
 Two (i) and three (ii) repetitions of motifs are shown.
@@
 Additional studies detail the role of such precise spike timing in downstream information transfer and coding~\citep{villette_internally_2015,branco_dendritic_2010,luczak_packet-based_2015}.
-This is, for instance, relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{cleland_construction_2014, kashiwadani_synchronized_1999, rinberg_speed-accuracy_2006} or touch~\citep{johansson_first_2004}.
+This is for instance relevant in sensory pathways in vision~\citep{meister_concerted_1995}, audition~\citep{decharms_primary_1996}, olfaction~\citep{cleland_construction_2014, kashiwadani_synchronized_1999, rinberg_speed-accuracy_2006} or touch~\citep{johansson_first_2004}.
 In particular, stereotyped sequences of neural activation have been described in the adult hippocampus and related to its function in mental travel in time and space~\citep{buzsaki_space_2018}.
@@
 %----------------------------%
-\begin{figure}[H]
-%\centering
-\includegraphics[width=0.98\linewidth]{figures/haimerl2019.jpg}
-\caption{
-        An example of a precise temporal motifs seen in cortical slices in two subsequent days. 
-        In this study by~\citep{haimerl_internal_2019}, an analysis of calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions shows repetitions of precise spiking motifs with a time scale of the order of seconds.
-Cells were selected and ordered with respect to their activity in the first imaging session.
-The black line on top represents the speed of the mouse.
-Futher analysis showe that more than the majority of the cells participating in run sequences on the first day were recruited again in run sequences on the next day.
-(modified from Figure 1-A from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}  %MDPI: Please add the explanation for B--D in the figure.
-The figure which should have been submitted was a crop on panel A : we have fixed that in the revision
-
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{../2022-12-20_polychrony_review_REV1/figures/haimerl2019.jpg}
+\caption{\textbf{An example of a precise temporal motifs seen in cortical slices.} In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
+(A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions.
+Cells have been selected and ordered with respect to their activity in the first imaging session (Top).
+The black line represents the speed of the mouse (modified from Figure 1 from~\citep{haimerl_internal_2019} under the CC-BY \href{https://www.pnas.org/doi/full/10.1073/pnas.1718518116}{PNAS License}).}\label{fig:haimerl}
 \end{figure}
@@
 
-It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010}, which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some frozen noise patterns.
+It is interesting to make a parallel with the ``Rapid Formation of Robust Auditory Memories'' reported in~\citep{agus_rapid_2010} which uses noise patterns to observe if listeners could learn to detect repeated occurrences of some frozen noise patterns.
 In particular, they used random waveforms to probe the formation of new memories for arbitrary complex sounds.
-A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4~ \si{\second}$ long.
+A behavioral measure was designed, based on the detection of repetitions embedded in noises up to $4 \si{\second}$ long.
 %The task is to detect the repetition of the same (frozen) noise within a trial.
@@
 Interestingly, one theoretical viewpoint considers synfire braids~\citep{bienenstock_model_1995}, where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned.
-Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006}, which this author defined as polychronous groups.
+Furthermore, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs~\citep{izhikevich_polychronization_2006} which this author defined as \emph{polychronous groups}.
 The rest of this review will be devoted to present evidence for the use of such precise spiking motifs in computational neuroscience, neurobiology and neuromorphic engineering.
-As a summary, it seems that such precise structural information is essential to the neural code and that it seems imperative to include this information in the decoding algorithm for a better understanding of neural activity.
+As a summary, it seems that such precise structural information is essential to the neural code and that it seems imperative to include this information in decoding algorithm for a better understanding of neural activity.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Understanding Precise Spiking Motifs in Neurobiology}\label{sec:detection}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding Neural Activity from Firing Rates}
+\section{Understanding precise spiking motifs in neurobiology}\label{sec:detection}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decoding neural activity from firing rates}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In most generic computational models, the neural activity is assumed to be encoded in the firing rate.
-For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point process\note[Laurent]{adding reference}~\citep{simoncelli_characterization_2003}.
+For instance, the output of the so-called linear non-linear (L-NL) models is assumed to model the response of a biological neuron as the sequence of a linear integration followed by a non-linear spiking response generating spikes according to a Poisson point-process\note[Laurent]{adding reference}~\citep{simoncelli_characterization_2003}.
 As such, a simple decoding strategy is to infer the input knowing the neuron's tuning curves, that is, its selectivity to a range of features~\citep{jazayeri_optimal_2006} or simply by a simple regression~\citep{berens_fast_2012}.
@@
 An important perspective of these methods used to decipher the recorded activity is that it could be ultimately used to fit neural network models to the recorded activity~\citep{bellec_fitting_2021}.
-In this particular paper, the authors fit the summary statistics of neural data with a (differentiable) spiking network model.
-The loss function is the cross entropy (following the Bernouilli hypothesis with a GLM, where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995}) and embedded with recurrent dynamics.
-In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the \note[Laurent]{now citing the URL}publicly available V1 dataset~\citep{kohn_utah_2016}, which allows supervising the model, with the input being the movie and the output the spikes recorded.
+In this particular paper, authors fit summary statistics of neural data with a (differentiable) spiking network model.
+The loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with an SRM neuron~\citep{gerstner_time_1995}) and embedded with recurrent dynamics.
+In particular, it comes with \href{https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure}{code} and uses the \note[Laurent]{now citing the URL}publicly available V1 dataset~\citep{kohn_utah_2016} which allows supervising the model with the input being the movie and the output the spikes recorded.
 These type of model may infer sparse activity in a set of binary latent variables, each describing the activity of a cell assembly~\citep{warner_probabilistic_2022}.
-\change[Antoine]{Using the right metric, as implemented in that paper by the corresponding neural models, is essential to better understanding neural data.}{Carefully picking the more appropriate metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.} Importantly, these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
-%
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Decoding Neural Activity Using Spike Distances}
+\change[Antoine]{Using the right metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.}{Carefully picking the more appropriate metric, as implemented in that paper by the corresponding neural models, is essential in better understanding neural data.} Importantly, these models are dependent on a core definition of spike measures, and we will review here how precise spiking motifs are taken into account by such spike distances.
+%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decoding neural activity using spike distances}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 There are different solutions to provide with a distance between two given spike trains.
-A known measure is the Victor--Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
-\change[JN]{In a later study, Van Rossum (2001) defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citep{van_rossum_novel_2001}.
+A known measure is the Victor-Purpura distance, which overcomes inconsistencies experienced with a \remove[JN]{simple} firing rate (Poisson model) of spike trains~\citep{victor_nature_1996}.
+\change[JN]{In a later study, Van Rossum (2001) have defined a distance that uses a time constant as a parameter}{Then a study tries to solve the problem by including a time constant as a parameter}~\citep{van_rossum_novel_2001}.
 \change[JN]{Depending on this parameter, the distance interpolates between a coincidence detector and a rate difference counter.}{This parameter will then be used to interpolate the distance between a coincidence detector and a rate difference counter.} Such distances were extended to non-Euclidean metrics and use morphological manipulations to compute spike train dissimilarity~\citep{kreuz_measuring_2007}.
-Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure\change[JN]{which}{, an interesting conclusion since the cross-correlation measure is} that which is adapted to the event-based nature of spiking signals.
-These observations lead to the intuition that each distance may be as good as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
+Mathematically, the stability of distance measures induced by level-crossing sampling can be evaluated~\citep{moser_stability_2014}, notably in light of the so-called Weyl's discrepancy measure~\citep{weyl_ueber_1916}, which may lead to the definition of a cross-correlation measure \change[JN]{which}{.
+An interesting conclusion since cross-correlation measure is} which is adapted to the event-based nature of spiking signals.
+These observations lead to the intuition that each distance may be as well  as the optimal solution of a generative model for these measures, possibly through non-linear relations~\citep{aronov_non-euclidean_2004}.
 % ~\citep{}
@@
 %Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
-Concerning spike timings,~\citet{levakova_review_2015} \change[JN]{reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning}{reviewed existing methods for estimating the latency of neural responses that include Bayesian binning.} \change[JN]{More broadly }{Alternatively }, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
-This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, particularly in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}.
+Concerning spike timings,~\citet{levakova_review_2015} \change[Antoine]{has}{have} \change[JN]{reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning}{reviewed existing methods for estimating the latency of neural responses that include Bayesian binning.}.
+\change[JN]{More broadly }{Alternatively }, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}.
+This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}.
 A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}.
@@
 This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance.
-The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data, \add[Laurent]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
-
+The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[Laurent]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
+
+%----------------------------%
 \begin{figure}[H]
-%\centering
-\includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
-\caption{
-  Detecting motifs using SPADE. 
-  We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}.
+  \centering
+  \includegraphics[width=\linewidth]{../2022-12-20_polychrony_review_REV1/figures/Diesmann_et_al_1999_spade.pdf}
+  \caption{\textbf{Detecting motifs using SPADE.} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}.
 This allowed to label different precise spike motifs which are denoted by different colors.
@@
   \end{figure}
+  %----------------------------%
 
 Another important algorithm, called SPOTDisClust, is based on the detection of structured temporal patterns~\citep{grossberger_unsupervised_2018}.
-They introduced an unsupervised method based on their detection from high-dimensional neural ensembles.
+They introduced an unsupervised method based for their detection from high-dimensional neural ensembles.
 The algorithm measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other.
@@
 
-%----------------------------%
-
-%----------------------------%
-
 % %%----------------------------%
@@
 % \centering
-% \includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
+% \includegraphics[width=\linewidth]{../2022-12-20_polychrony_review_REV1/figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
 % \caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Scaling Up to Very Large Scale Data}
+\subsection{Scaling up to very large scale data}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
-Bridging across different spatial and temporal scales, combination of {in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks at an unprecedented level of detail and precision.
+Bridging across different spatial and temporal scales, combination of \emph{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
 The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurobiological knowledge to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
@@
 Basically, it rearranges neurons in the raster plot based on the similarity of their activity and applies a deconvolution strategy based on a linear model.
-\add[Laurent]{Yet this method was mainly tested on calcium imaging data, which are known to add some imprecision to the timing of the original spiking activity.} The model is \note[Laurent]{now citing the URL}openly accessible~\citep{stringer_mouselandrastermap_2020} and has led to important discoveries.
+\add[Laurent]{Yet this method was mainly tested on calcium imaging data which is known to add some imprecision to the timing of the original spiking activity.} The model is \note[Laurent]{now citing the URL}openly accessible~\citep{stringer_mouselandrastermap_2020} and has led to important discoveries.
 In~\citep{stringer_spontaneous_2019} for instance, it was shown that a neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated.
@@
 In~\citep{russo_cell_2017}, the authors developed novel machine learning tools and statistical tests for unsupervised spatio-temporal pattern detection in non-stationary environments, which were applied to simultaneous electro-physiological recordings from tens to hundreds of neurons for decoding cognitive processes from neural activity.
-Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in~neuroscience.
+Altogether, this provides evidence for the importance of such machine-learning-based tools to provide with breakthroughs in neuroscience.
 %   Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
@@
 The core measure is based on a non-stationarity corrected parametric statistical test for assessing the independence of pairs and an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.
-To overcome the limits of models which require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} addressed each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
-\change[Laurent]{They also introduce}{As originally introduced by}~\citep{kass_statistical_2005}, they used learnable time-warping parameters to model sequences of varying duration, which were experimentally observed in neural circuits, and demonstrated these advantages on experimental recordings from the songbird higher vocal center and rodent hippocampus.
+To overcome the limits of models which either require spike times to be discretized, utilize a suboptimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{williams_point_2020} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a few marked events in continuous time.
+\change[Laurent]{They also introduce}{As originally introduced by}~\citep{kass_statistical_2005}, they use learnable time-warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits, and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.
 At a larger scale, in~\citep{van_kempen_top-down_2021}, it was shown that attentional information from V4 or arousal can change the timings of groups of events in V1.
-They develop a hidden Markov model for quantifying the transitions.
-In particular, they showed that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
+They develop a Hidden Markov Model for quantifying the transitions.
+In particular, they show that fluctuations in neural excitability are coordinated between visual areas with retinotopic precision.
 Top-down attention drives inter-areal coordination along the reverse cortical hierarchy, predicting better behavioral performance with increased coordination.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{What Biological Mechanism Could Allow Learning Spiking Motifs?}
+\section{What biological mechanism could allow learning spiking motifs?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The discovery of the existence of an equivalent biological mechanism in vertebrates as well as the demonstration of the importance of its role in various developed behaviors allow us to put these doubts to rest.
-In the following paragraphs, we will successively describe the first biological observations of delay learning, identify myelinization as an important actor, and finally study this phenomenon at the molecular level.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Biological Observations of Delay Adaptation}
+In the following paragraphs we will successively describe the first biological observations of delay learning, identify myelinization as an important actor and finally study this phenomenon at the molecular level.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Biological observations of delay adaptation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-One of the first significant pieces of evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_functioning_1938}.
+One of the first significant evidence of any neuronal delay in the information propagation within the animal neural system came from Hermann von Helmholtz's study of a frog's sciatic nerve in 1850~\citep{von_helmholz_messungen_1850,peyrard_how_2020}, and was later confirmed with Young's study of the squid giant axon~\citep{young_functioning_1938}.
 Dendritic propagation delays vary from sub-milliseconds to a few milliseconds, while axonal propagation delays range from a few milliseconds to tens of milliseconds, depending on the neuronal population studied~\citep{madadi_asl_dendritic_2018}.
-Extensive measures of nerve conduction velocities were performed in different animals, including humans, and significant variations related to age, sex and other factors were identified~\citep{stetson_effects_1992}.
+Extensive measures of nerve conduction velocities were done in different animals, including humans, and significant variations related to age, sex and other factors have been identified~\citep{stetson_effects_1992}.
 
@@
 This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears.
-\add[Amélie]{It was first theorized in the Jeffress ``ITD-versus-place model''}~\citep{jeffress_place_1948}.
-\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
-According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'' and therefore necessarily need to play on the response time to do so.
-\add[Amélie]{This assertion confused the mean interspike interval, i.e., how often a neuron can fire, and the specific spike time, i.e., how precisely a spike can be generated.
+\add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\citep{jeffress_place_1948}.
+\add[Amélie]{As hypothesized by this model, the \emph{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \emph{nucleus magnocellularis} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
+According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.
+\add[Amélie]{This assertion confused the mean interspike interval, i.e.
+how often a neuron can fire, and the specific spike time, i.e.
+how precisely a spike can be generated.
 However, it has nevertheless contributed to the recognition of the importance of time in various biological mechanisms.} \remove[Amélie]{penacochlear2001 conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris.
-In the mouse somatosensory cortex, camontiming2019 found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e., the temporal disparity between the left and right cochlear loci in the owl's brain.
-However, it was quickly set aside, as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
-The authors supported a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors concluded that regulations at different sites within individual axons of at least two parameters, namely, the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays, thus allowing the ITD detection.
-\add[Amélie]{The authors also noted that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.}
-
-\add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localization behaviour.
+In the mouse somatosensory cortex, camontiming2019 found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e.
+the temporal disparity between left and right cochlear loci in the owl's brain.
+However, it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears.
+The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection.
+\add[Amélie]{The authors also note that the barn owl's axons seem to change in length, thus implementing a ``pure delay line''.}
+
+\add[Amélie]{The experiments described above thus conclude on the important role of physiological delays in the avian sound localisation behaviour.
 However, the relevance of precise timing in spikes is not limited to birds; for example, in the mouse somatosensory cortex,}~\citep{camon_timing_2019} \add[Amélie]{found a strong correlation between the delay of the mouse behavioral response and the timing of multiunit activity evoked by a trained whisker.} These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977}\change[Amélie]{--- instead of solely the axon's length as was previously thought.}{.
-This mechanism adds to the axonal length delay, which was previously thought to be the sole influence on the conduction velocity due to its anatomical soundness---as}~\citet{seidl_mechanisms_2010}\add[Amélie]{ experimentally demonstrated, this mechanism by itself is not sufficient to explain the biological functionality but should be added to the one of axonal length delay.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Importance of Myelination}
+This mechanism adds to the axonal length delay, which was previously thought sole influence on the conduction velocity due to its anatomical soundness --- as}~\citet{seidl_mechanisms_2010}\add[Amélie]{ experimentally demonstrated, this mechanism by itself is not sufficient to explain the biological functionality but should be added to the one of axonal length delay.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The importance of myelination}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % The importance of myelination
-\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to \change[Amélie]{both the axon diameter by itself and the diameter of the axon wrapped in myelin}{the axon diameter, i.e., the amount of myelin wrapped around the axon}.
-Indeed, the oligodendrocytes, one of the many glial cells present in the vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produces thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
+\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to \change[Amélie]{both the axon diameter by itself and the diameter of the axon wrapped in myelin}{the axon diameter, i.e.
+the amount of myelin wrapped around the axon}.
+Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}.
 Myelinization consists in ``two motions: the wrapping of the leading edge of the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}.
-Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the approx $1~\upmu$m gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_neuron-oligodendrocyte_2021}.
-\change[Antoine]{In}{On} average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\upmu$m long~\citep{simons_oligodendrocytes_2016}.
+Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the approx $1~\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_neuron-oligodendrocyte_2021}.
+\change[Antoine]{In}{On} average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}.
 This demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}.
-Thus, in several occasions, myelin has been identified as \change[Antoine]{a main}{an important} actor in the regulation of conduction velocity in neurons, i.e., axonal delay regulation.
+Thus, in several occasions, myelin has been identified as \change[Antoine]{a main}{an important} actor in the regulation of conduction velocity in neurons, i.e.
+axonal delay regulation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interplay of Delay Adaptation and Neural Activity}
+\subsection{Interplay of delay adaptation and neural activity}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 To answer this, we must first study the myelinization process.
-The OPCs first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
+The OPCs first proliferate in the white matter \emph{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
 Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
-Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as Ca$^{2+}$ activity~\citep{baraban_ca2_2018} of the neuregulin 1 growth factor~\citep{nave_axonal_2006}.
+Not much is known about how oligodendrocytes select the axons to myelinate, but it seems that myelination only takes place on large enough axons and is strongly regulated by several factors~\citep{kuhn_oligodendrocytes_2019}, such as $Ca^{2+}$ activity~\citep{baraban_ca2_2018} of the neuregulin 1 growth factor~\citep{nave_axonal_2006}.
 The important role of myelination on delay learning and biological behaviors, as highlighted in the previous paragraph, suggests that the identification of target neurons as well as the myelin production is also regulated by neuronal activity.
@@
 They also confirmed that the delay correlates with the level of skill acquisition.~\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or {via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
-
-A potential scenario for selective myelination on electrically active axons using non-synaptic junctions between an axon and an oligodendrocyte is as follows: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \emph{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
+
+A potential scenario for selective myelination on electrically active axons using non-synaptic junctions between an axon and an oligodendrocyte is: the axon releases glutamate in the extracellular environment by vesicle fusion, which activate the oligodendrocyte's NMDA and metabotropic glutamate receptors.
 This triggers the axo-glial signalling complex, involving the phosphorylation of the SRC family kinase FYN followed by the translation of heterogeneous nuclear ribonucleoprotein A2 into local myelin basic proteins~\citep{fields_new_2015}.
@@
 The previous paragraphs present the biological mechanisms behind the axonal delay, regulated by myelination.
-However, the dendritic delay \add[Amélie]{as well as the axonal delay (see}~\citep{seidl_mechanisms_2010}) seems to play an equally important role in the precise timing within a sequence of spikes.
+However, the dendritic delay \add[Amélie]{as well as the axonal delay (see }~\citep{seidl_mechanisms_2010}) seem to play an equally important role in the precise timing within a sequence of spikes.
 Dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}.
 Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brain stem~\citep{spencer_compensation_2018}.
-\citet{mel_synaptic_2017} \add[Amélie]{highlighted the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
+\citet{mel_synaptic_2017} \add[Amélie]{highlights the dendrites' impact on neuronal plasticity, which is caused by the wide variation of numerous parameters: plasticity rules applied to different dendritic subtrees or dendritic subregions, local passive cable properties, distance travelled by remote dendritic inputs, branching structures, dendritic diameters, the relative timing of back-propagating somatic action potentials, etc.
 Dendritic spiking involvement in synaptic potentiation following active backpropagation into dendrites was experimentally uncovered using calcium imaging to highlight dendritic calcium entry allowing for long-tem potentiation}~\citep{golding_dendritic_2002}.
-~\citet{branco_dendritic_2010} demonstrated the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
+~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''.
 The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree and relies on the dendritic calcium influx moderation by NMDA receptors.
@@
 All in all, learning spike motifs requires significantly complex pathways and biological mechanisms, whether in the dendrites or the axon of the neuron.
-More and more is becoming known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop ANNs, whose learning rules would be more neuromorphic.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Modeling Precise Spiking Motifs in Theoretical and Computational Neuroscience}
+More and more is known about the non-trivial research topic that is biological delay learning, and extensive experimental data help develop ANNs, whose learning rules would be more neuromorphic.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Modeling precise spiking motifs in theoretical \& computational neuroscience}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Spiking neural networks (SNNs)~\citep{maass_networks_1997} are natural candidates to use these precise temporal patterns in the brain.
-The approach which is currently most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021}.
-One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non-spiking) neural network.
+Spiking Neural Networks (SNNs)~\citep{maass_networks_1997} are natural candidates to use these precise temporal patterns in the brain.
+The approach which is currently most prominent in the SNN community is to use existing algorithms from machine learning and to adapt them to the specificity of spiking architectures~\citep{goltz_fast_2021} One such example is to adapt the successes of deep learning algorithms and to transfer the back-propagation algorithm to SNNs, as it is the most widely used to tune the weights of a classical (non spiking) Neural Network.
 In a nutshell, it considers the system as implementing an input / output function and iteratively updates the weights according to the direction and magnitude of the error's gradient.
-\change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, the gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical neural network to a spiking architecture~\citep{rueckauer_conversion_2017}.
-SNNs reach in some case a similar performance as their non-spiking equivalent, for instance on the \remove[Antoine]{N-}MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
+\change[Antoine]{Since the activation function of a spiking neuron is not differentiable,}{In deep learning, gradient is computed on the activation function and since spikes are not differentiable,} a recent popular approach consists in using a surrogate gradient~\citep{neftci_surrogate_2019} to ``cross-compile'' a classical Neural Network to a spiking architecture~\citep{rueckauer_conversion_2017}.
+SNNs reach in some case a similar performance as their non spiking equivalent, for instance on the \remove[Antoine]{N-}MNIST dataset for categorizing digits in a stream of events~\citep{susi_nmnsd-spiking_2021}.
 \change[Antoine]{The main reason for adopting this approach instead of classical architectures is the possibility to benefit from dedicated neuromorphic low-energy hardware.
 However, most biological neural systems are obviously more efficient than current state-of-the-art vision systems, both in terms of efficiency (accuracy), in speed (latency), and energy consumption.}{So far, this approach does not outperform classical architectures both in term of training efficiency and performances}~\citep{davies_loihi_2018}\add[Antoine]{.
-However, they remain the best candidates to reproduce biological neural systems and their capacities in terms of accuracy, speed and energy consumption.} There is, therefore, an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
+However, they remain the best candidates to reproduce biological neural systems and their capacities both in terms of accuracy, speed and energy consumption.} There is therefore an immense gap in the way we understand biology to translate it to the efficiency of SNNs.
 To go beyond the state-of-the-art, we will focus here on one core computation of a spiking neuron, that is, its ability to switch from the classical integrator mode (summing analog currents on its \change[Antoine]{synapses}{dendrites}) \change[Antoine]{to a synchrony detector where it emits a spike whenever presynaptic spiking inputs are synchronized}{to a detector of precise spiking motifs}~\citep{paugam-moisy_computing_2012}.
@@
 %In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Izhikevich's Polychronization Model}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Izhikevich's polychronization model}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % polychronization
-As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing rate.
-Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time-encoding machine by}~\citet{lazar_time_2004} and polychronization model of~\citet{izhikevich_polychronization_2006}.
-\add[Antoine]{In this section, we focus on the polychronization model} based on a random recurrent model of spiking neurons, including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a spike-time-dependent plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
+As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
+Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time encoding machine by}~\citet{lazar_time_2004} and \emph{polychronization} model of~\citet{izhikevich_polychronization_2006}.
+\add[Antoine]{In this section we focus on the polychronization model} based on a random recurrent model of spiking neurons including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
 \change[Antoine]{It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons}{It was shown that spike timing (STDP) has an impact on the development of synaptic efficacy for many kinds of neurons}~\citep{caporale_spike_2008}.
@@
 It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that delays are then ``frozen'' for the rest of the simulation.
-Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.,~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs).
+Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked firing patterns which  the author defined as ``polychronous groups'' (PGs).
 One core ingredient of this model is the fact that the neurons composing a group fire at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time.
@@
 According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group.
-%----------------------------%
-%
-
-% 
-%----------------------------%
 
@@
 Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper.
-Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999}, yet they had not shown potential applications to the detection of spiking motifs.
+Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs.
 A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}.
-The \emph{tempotron} is a supervised synaptic learning algorithm, which classifies a distractor from a target motif, in order to extend the perceptron, which does not incorporate a spike timing framework.
+The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron, which does not incorporate a spike timing framework.
 The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a supervised STDP rule.
@@
 They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system.
-Stimulation of the trained network produces the activation of a PG, i.e., the propagation of firing activity through multiple layers due to convergent patterns of firing.
-While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this~section.
+Stimulation of the trained network produces the activation of a PG, i.e.
+the propagation of firing activity through multiple layers due to convergent patterns of firing.
+While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
 
@@
 The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the modulation of spike latency emission~\citep{susi_nmnsd-spiking_2021}.
-Additional models for the detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
+Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}.
 This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.
@@
 
+%----------------------------%
+%
 \begin{figure}[H]%[t!]
-%  \centering
-  \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
-    \caption{
-      Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.
-       {(\textbf{Left})}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-{(\textbf{Middle})}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons; therefore, no spike will be emitted, as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+  \centering
+  \includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
+    \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \emph{(Left)}~In this example, three presynaptic neurons denoted \emph{b}, \emph{c} and, \emph{d} are fully connected to two post-synaptic neurons \emph{a} and \emph{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \emph{a} and  $8$, $5$, and $1~\ms$ for \emph{e}.
+\emph{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \emph{a} and \emph{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
     %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    {(\textbf{Right})}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
+    \emph{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \emph{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \emph{e}.
      }
@@
 %----------------------------%
-%
-
-\begin{figure}[H]%[t!][]
-  %  \centering
-    \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-      \caption{
-        Detecting event-based motifs using spiking neurons with heterogeneous delays.
-      { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific neuron addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise spiking motifs.
-      { (\textbf{b})}~The generative model is defined by this set of motifs (here $4$ of them) each defined by different weights at heterogeneous delays (red for excitatory, blue for inhibitory). 
-      { (\textbf{c})}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these spiking motifs~\citep{grimaldi_learning_2022}.
+
+%
+%----------------------------%
+%
+\begin{figure}[H]%[t!]
+  \centering
+  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1a_k.pdf}
+  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1c.pdf}
+  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1b.pdf}
+  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1a.pdf}
+    \caption{\textbf{Detecting event-based motifs using spiking neurons with heterogeneous delays.} 
+    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs).
+    {\bf (b)}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay.
+    {\bf (c)}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}.
 % such that the addresses are the synapses of these neurons (which we therefore call here synapses).
@@
     %{\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
-      { (\textbf{d})}~Knowing the %MDPI: Please change the (a) into (d) in picture. Fixed !
-   results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic example, all spikes are now associated with a motif.
+    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
      }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Learning Synaptic Delays}
+\subsection{Learning synaptic delays}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Training delays ~\citep{state_training_2019}
-First, the original model by Izhikevich uses a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
-For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity--connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
+First, the original model by Izhikevich use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection.
+For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}.
 Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}.
-They were also extended to recurrent neuronal networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
+They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}.
 In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
 More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission.
-Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity, and these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
+Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
 
@@
 The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns.
-Applying this STDP-based rule on delays to the spiking neural network, the experimental results were validated on a simple motion detection task, but were prone to convergence issues.
+Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task, but were prone to convergence issues.
 Another model of synaptic delay-weight plasticity integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes~\citep{zhang_supervised_2020}.
-This was also presented by~\citep{wang_delay_2019}, who proposed a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-Other models, such as that of~\citep{hazan_memory_2022}, propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
-In a recent paper~\citep{luo_supervised_2022}, the authors proposed a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron.
+This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
+Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
+In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron.
 In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
-In yet another computational model,~\citet{sun_learning_2016} showed that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.
+In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Real-World Applications}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%----------------------------%
-% TODO
-
-% 
-%----------------------------%
+\subsection{Real-world applications}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios.
-Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods, such as image processing, would further prove their plausibility.
-For instance, in a recent work,~\citet{ghosh_spatiotemporal_2019} proposed a two-stage unsupervised--supervised system for the categorization of spatio-temporal actions from an event-based stream.
-The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal-related changes to a local spatio-temporal spike-event pattern.
+Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility.
+For instance, in a recent work,~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream.
+The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern.
 The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity.
-For testing the system, two datasets are considered: DVS gesture and a new action recognition dataset recorded for this work.
+For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work.
 Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.
@@
 This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks.
-In essence, two neurons, which were selective to specific motions, were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
-This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot, even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
-This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash--lag illusion~\citep{khoei_flash-lag_2017}.
+In essence, two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.
+This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}.
+This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash-lag illusion~\citep{khoei_flash-lag_2017}.
 However, these latter models used parametric rules for defining the weights.
-Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore some exemplar results from neuromorphic engineering to obtain better insights on that aspect.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Applications of Precise Spiking Motifs in Neuromorphic Engineering}
+Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore some exemplar results from neuromorphic engineering to get better insights on that aspect.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Applications of precise spiking motifs in neuromorphic engineering}
 \label{sec:neuromorphic}
@@
 %
-Artificial intelligence has made huge advances in the past decades.
-Deep learning algorithms, nowadays, outperform humans at complex tasks such as natural image recognition or abstract strategy board games.
+Artificial intelligence has made huge advances in the past decades and deep learning algorithms, nowadays, outperform humans at complex tasks like natural image recognition or abstract strategy board games.
 Yet, machine learning algorithms suffer from adversarial attacks or a lack of generalization \change[Antoine]{capacity but}{capacity.
-However,} their main weakness, compared to biological neural networks, is their poor energy efficiency.
+But} their main weakness, compared to biological neural networks, rely on their poor energy efficiency.
 Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence~\citep{roy_towards_2019, javanshir_advancements_2022}.
-Because the reduced energy consumption of biological networks can be explained \add{in part} by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to \change{make}{perform} energy-efficient computations.
+Because the reduced energy consumption of biological networks can be explained \add{in part} by the use of spikes and asynchronous responses to exchange information~\citep{maass_networks_1997}, neuromorphic chips use this parallel and event-based representation to \change{make}{perform} energy efficient computations.
 Another important distinction with classical von Neumann architectures is the localized memory of this new type of chips.
 It can be materialized by the capacity of the physical connections between the processing units to store information~\citep{markovic_physics_2020}.
-An example of such a connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted.
+An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{rasetto_challenges_2022} for which the resistance value can be dynamically adjusted.
 Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
-Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2003, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing techniques show their advantages in terms of frugality and rapidity only on neuromorphic chips.
+Even if some useful SNNs simulators run on GPUs~\citep{diesmann_nest_2003, hazan_bindsnet_2018, stimberg_brian_2019}, such event-based computing technics show their advantages in terms of frugality and rapidity only on neuromorphic chips.
 
@@
 Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges~\citep{zenke_visualizing_2021}.
-In this section, we give a description of the different neuromorphic hardware that have been developed and see how they can be used to deal with precise temporal motifs.
-%
-\subsection{The Emergence of Novel Computational Architectures}
+In this section, we make a description of the different neuromorphic hardware that have been developed and see how they can be used to deal with precise temporal motifs.
+%
+\subsection{The emergence of novel computational architectures}
 %
@@
 Today, many devices are good candidates for implementing event-based algorithms and use the \change[Antoine]{AER}{address event representation (AER)}.
-They can be divided intro three major categories: digital, analog and mixed analog/digital platforms.
-For a more complete review, the reader can refer to~\citep{schuman_survey_2017}, where we site the most popular ones.
+They can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms.
+For a more complete review the reader can refer to~\citep{schuman_survey_2017}, here we site the most popular ones.
 SpiNNaker~\citep{furber_overview_2013, furber_spinnaker_2020}, Loihi~\citep{davies_loihi_2018} and TrueNorth~\citep{merolla_million_2014} chips are widely used examples of digital hardware implementations.
-Compared to TrueNorth, which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning.
+Compared to TrueNorth which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning.
 This flexibility in the implementation comes at the cost of an increased energy consumption.
-Mixed analog--digital systems were developed at Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}.
+Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}.
 They are mostly used by computational neuroscientists to model brain activity with different levels of abstraction.
-BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed analog--digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}.
+BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}.
 It is a wafer-scale neuromorphic hardware with analog components.
-Analog arrays (i.e., field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices.
-The pulsed current source synapse and the DPI are examples of such devices; we can also mention the field programmable neural array~\citep{farquhar_field_2006} and the NeuroFPAA~\citep{cheng_fpaa_2009} specifically designed for neuromorphic systems.
+Analog arrays (i.e.
+field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices.
+The pulsed current source synapse and the DPI are examples of such devices, we can also mention the field programmable neural array~\citep{farquhar_field_2006} and the NeuroFPAA~\citep{cheng_fpaa_2009} specifically designed for neuromorphic systems.
 Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.
 %
+
+Neuromorphic sensors are also developed with the idea to capture external stimuli more efficiently and closer to biological systems.
+A widely used example is the Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
+These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene.
+We also report other event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but will focus on DVS for the next subsection about the use of dynamics embedded in event-based signals.
+
 %----------------------------%
-
-%----------------------------%
-%
-Neuromorphic sensors have also been developed with the idea to capture external stimuli more efficiently and closer to biological systems.
-A widely used example is the dynamic vision sensors (DVSs) which provide a stream of binary asynchronous events signaling detectable changes in luminance.
-These devices, also named ``silicon retinas'' (see Figure~\ref{fig:silicon_retina}), show great improvements in terms of memory allocation, or power consumption, for the recording of a visual scene.
-We also report other event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} but will focus on DVS for the next subsection about the use of dynamics embedded in event-based signals.
-
 \begin{figure}[H]
   \centering
-\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
-\caption{
- A miniature, event-based ATIS sensor.
- Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides events at the micro-second resolution.
-These are sparse, as they represent luminance increments or decrements (ON and OFF events, respectively).
+\includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/event_driven_computations.png}
+\caption{\textbf{A miniature, event-based ATIS sensor.} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
 Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
 \end{figure}
+%----------------------------%
+%
 
@@
 Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
-In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
+In this section, we will use indifferently \emph{events} or \emph{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
 There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
@@
 %
-\subsection{On the Importance of Spatio-Temporal Information in Silicon Retinas}
-%
-With the AER specification and their sub-millisecond temporal precision, DVSs bring a new approach to the storage and processing of visual information.
+\subsection{On the importance of spatio-temporal information in silicon retinas}
+%
+With the AER specification and their sub-millisecond temporal precision, DVS bring a new approach to the storage and processing of visual information.
 From their generative model, these sparse events are markers of the dynamics of the visual scene captured by the sensor.
-The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision~tasks.
-
-In~\citep{lagorce_hots_2017}, time surfaces were introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+The dynamics of the event streams have to be used to make sense of the recorded information, and new algorithms are needed to solve efficiently classical computer vision tasks.
+
+In~\citep{lagorce_hots_2017}, \emph{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
 An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
 It gives more precision to represent recent events and offers an analogy with the LIF spiking neuron.
-It is a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity, they can capture repeating motifs within the input signals.
+Its a way to represent the local dynamics embedded in the event-based recordings, and with unsupervised learning on the event stream based on the cosine similarity they can capture repeating motifs within the input signals.
 Learned time surfaces can be used for object recognition~\citep{lagorce_hots_2017, sironi_hats_2018, maro_event-based_2020, grimaldi_homeostatic_2021,grimaldi_robust_2022} and show that this method could be efficiently applied to state-of-the-art benchmarks.
@@
 With their additional temporal dimension, their kernels can capture multiple events on the same pixel address, as long as they belong to the local temporal window (see Figure~\ref{fig:icip}).
-This representation is only limited by the time step used for the discretization of the signal; this factor defines the temporal precision of the representation.
+This representation is only limited by the time step used for the discretization of the signal, this factor defines the temporal precision of the representation.
 Other methods make direct use of the precise timing of events captured by the DVS to solve optical flow and time-to-contact challenges~\citep{benosman_event-based_2014, clady_asynchronous_2014, tschechne_bio-inspired_2014}, inferring depth~\citep{hidalgo-carrio_learning_2020}, feature detection and tracking~\citep{dardelet_event-by-event_2021}, motion segmentation~\citep{stoffregen_event-based_2019} or the simultaneous localization and mapping problem~\citep{kim_real-time_2016}.
-This non-exhaustive list of complex task solving is not directly linked to biological processes, but shows the potential of the precise temporal resolution of neuromorphic retina-like cameras.
+This non-exhaustive list of complex task solving is not directly linked to biological processes, but show the potential of the precise temporal resolution of neuromorphic retina-like cameras.
 By essence, these sensors offer a novel view of visual information processing due to the asynchronous responses of the different pixels.
-With this type of signal, the use of spatio-temporal motifs embedded in the event streams is essential to solve high-level visual tasks.
-
+With this type of signal, the use of spatio-temporal motifs embedded in the event streams are essential to solve high level visual tasks.
+%
+
+%----------------------------%
+% TODO
 \begin{figure}[H]%[t!]
-%  \centering
-  \includegraphics[width=0.98\linewidth]{figures/2022-09-27_MotionDetection_kernel.png}
-  %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
-    \caption{
-      Detecting visual motion in an event stream with heterogeneous delays.
-      Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
-The models' parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
-When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push--pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}.
-Global weight is globally decreasing from the lowest delay (\textbf{right}) to less recent information (\textbf{left}).
+  \centering
+  \includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/2022-09-27_MotionDetection_kernel.png}
+  %\includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/2022-10-10_MotionDetection_kernel.png}
+    \caption{\textbf{Detecting visual motion in an event stream with heterogeneous delays.} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
+The models parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
+When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push-pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}.
+Global weight is globally decreasing from the lowest delay (right) to less recent information (left).
      }
@@
 \end{figure}
-
-%
-\subsection{Computations with Delays in Neuromorphic Hardware}
+% 
+%----------------------------%
+\subsection{Computations with delays in neuromorphic hardware}
 %
@@
 \citep{hussain_deltron_2012} implemented a delay-learning algorithm on an analog chip.
-Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was performed on the analog architecture; training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}.
+Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}.
 In addition to delay learning, a group at the University of West Sydney developed a neuromorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}.
-They showed that STDP and spike-timing-dependent delay plasticity rules could be implemented in both a digital and an analog chip.
-From the network parameters and the physical limitations to store it on-chip, they proved that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
+They show that STDP and Spike Timing Dependent Delay Plasticity rules could be implemented in both a digital and an analog chip.
+From the network parameters and the physical limitations to store it on-chip, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network.
 The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2014}.
 %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance.
-\citet{pfeil_neuromorphic_2013} implemented STDP on a mixed analog--digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
+\citet{pfeil_neuromorphic_2013} implemented STDP on a mixed-analog-digital chip to simulate the sound localization processes observed on the barn owl auditory system~\citep{gerstner_neuronal_1996}.
 Coherence detection on input spikes coming from two sources was obtained with a 50 $ns$ precision.
-They claimed that this unsupervised learning denoises the input and compensates for variations between neural components.
-The variability of response of the analog components can be compensated by population coding for responses robust to noise, and this phenomenon is also observed in biological neural networks~\citep{boerlin_spike-based_2011}.
-A recent work performed the implementation of a sparse vector symbol architecture binding operation on the Loihi chip, delay lines and coincidence detection, used to compute the binding operation~\citep{renner_sparse_2022}.
-They highlighted the fact that using delays can be expensive \add[Laurent]{notably in memory bandwidth} because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay.
-Note that this problem is also due to the algorithm used in this paper and that the analog chips must not suffer from this type of~limitation.
+They claim that this unsupervised learning denoises the input and compensates for variations between neural components.
+The variability of response of the analog components can be compensated by population coding for responses robust to noise and this phenomenon is also observed in biological neural networks~\citep{boerlin_spike-based_2011}.
+A recent work makes implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}.
+They highlight the fact that using delays can be expensive \add[Laurent]{notably in memory bandwidth} because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay.
+Note that this problem is also due to the algorithm used in this paper and that the analog chips must not suffer from this type of limitation.
 
@@
 While improvements are still to be made, neuromorphic chips seem to be a good candidate to efficiently make use of these particular features.
-
 %
@@
 %: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
-In this review paper, we presented recent evidence for the role of precise spiking motifs in neuroscience.
-In particular, we showed that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels, and that they may have numerous applications in neuromorphic engineering.
-In particular, we showed the following:
+In this review paper, we have presented recent evidence for the role of precise spiking motifs in neuroscience.
+In particular, we have shown that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels and that they may have numerous applications in neuromorphic engineering.
+In particular, we have shown that:
 \begin{itemize}
-  \item  The efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlights the importance of precise spike times;
-  \item  Growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, travelling waves on arbitrary spiking motifs, \add[Laurent]{and notably that an encoding based on precise spiking motifs may provide huge computational benefits};
-  \item  Many theoretical models already exist, taking into account the specificity of spiking motifs, notably by using heterogeneous delays;
-  \item  Using precise spiking motifs could ultimately be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
+  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlights the importance of precise spike times,
+  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, travelling waves on arbitrary spiking motifs, \add[Laurent]{and notably that a an encoding based on precise spiking motifs may provide huge computational benefits},
+  \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
+  \item  using precise spiking motifs could ultimately be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
 \end{itemize}
-
  Overall, our reviewing effort has shown that a growing community is focusing on that aspect.
-This community is based on solid and validated evidence, which is breaking novel grounds thanks to the current technical advances.
-Moreover, we also showed that this community is highly diverse, operating in biology, computational neuroscience or neuromorphic engineering.
-As a consequence, the global effort is still largely scattered, which limits its larger acceptance in~neuroscience.
+This community is based on solid and validated evidence which is getting novel grounds thanks to current technical advances.
+Moreover, we have also shown that this community is highly diverse, operating on \emph{in vivo}, \emph{in computo} and \emph{in silico} systems.
+As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
 
 \subsection{Limits}
-Additionally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, and neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
+Additionally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.
 
@@
 In addition, this representation can induce errors due to the discretization and the scale of temporal sampling.
-Finally, this representation encourages the use of classical methods, which are not adapted to disruptive applications, such as event-driven representations.
+Finally, this representation encourages the use of classical methods, which are not adapted to disruptive applications such as event-driven representations.
 
 Moreover, the learning of patterns is often done in a supervised way.
-Indeed, the problem of detecting polychronous groups implies to infer both the address and the precise time of the occurrence of these motifs.
+Indeed, the problem of detecting polychronous groups implies to infer both the address and the precise time of occurrence of these motifs.
 Most of the models we have presented are based on the assumption that at least one of its variables is known: either the pattern, its identity, or its time of occurrence.
-This constraint is to be put in parallel with the way a biological nervous system works in which learning is performed autonomously, i.e., without supervision.
+This constraint is to be put in parallel with the way a biological nervous system works in which learning is done autonomously, i.e.
+without supervision.
 However, we can note that some models can perform such learning, but only in the case of data for which the motifs are easily separable.
@@
 A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output.
-However, we saw that the neurobiological system uses processing loops within hierarchical graphs.
+However, we have seen that the neurobiological system uses processing loops within hierarchical graphs.
 In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}.
-The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance, the lack of a global clock, or the cost of fully connected topologies.
-Moreover,  these processes have to be distinguished from judgements on timing, such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts)~\citep{coull_distinction_2022}.
+The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
+Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts)~\citep{coull_distinction_2022}.
 %
@@
 %
-The limits that we have presented can be treated individually in each model, as evidenced by individual efforts, which try to overcome them.
+The limits that we have presented can be treated individually in each model, as evidenced by individual efforts which try to overcome them.
 However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.
@@
 For instance, the design of model-driven protocols has proven to be essential in modern neuroscience.
-Additionally, if neural networks were essential in shaping modern-day machine learning, e.g., computer vision using deep learning, spiking neural networks should prove essential in future emerging technologies.
-
-In that perspective, it is essential that such models are tested on ecologically relevant, real-world scenarios.
-Indeed, classic convolutional neural networks have emerged as optimal solutions, for example, to classify static images into categories, yet they are not well-adapted for processing dynamic, multimodal sensory flows.
-The emerging necessity to be able to process more complex flows, such as the multiple flows of information in a car designed for autonomous driving, necessitates modifying such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs.
-Notably, the protocols used as well in neurobiology, theory or engineering should take into account these novel levels of complexity.
+Also, if neural networks were essential in shaping modern-day machine learning, e-g.
+computer vision using deep-learning, spiking neural networks should prove essential in future emerging technologies.
+
+In that perspective, it is essential that such models are tested on ecologically-relevant, real-world scenarios.
+Indeed, classic convolutional neural networks emerged as optimal solutions for instance to classify static images into categories, yet they are not well adapted for processing dynamic, multimodal sensory flows.
+The emerging necessity to be able to process more complex flows, such as the multiple flows of information in a car designed for autonomous driving, necessitates to modify such modelling paradigms, and in particular, to take into account that the generated actions may modify the sensory inputs.
+Notably, the protocols used as well in neurobiology, theory or engineering should take into account this novel levels of complexity.
 
@@
 This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}.
-This aspect is essential to foster the emergence of interdisciplinary projects, such as model-driven neurobiological experiments or biologically inspired neuromorphic engineering.
-Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011}, i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms.
+This aspect is essential to foster the emergence of interdisciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering.
+Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011} i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms.
 As such, this review aims at paving the way to openly share the variety of resources and to offer a unified view on the role of precise spiking motifs in neuroscience.
@@
 % Agence Nationale de la Recherche %HL and
-LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}''.
+LP, CB and AG received funding from the ANR grant number ANR-20-CE23-0021 ``\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}''. %
 % LP received funding 
@@
 
-\institutionalreview{Not applicable.
-%In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
-You might choose to exclude this statement if the study did not require ethical approval.
-Please note that the Editorial Office might ask you for further information.
-Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
-OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
-OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.
-}
-
-\informedconsent{Not applicable.
-%Any research article describing a study involving humans should contain this statement.
-Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
-You might also choose to exclude this statement if the study did not involve humans.
-%
-%Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
-Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.
-}
-
 \dataavailability{This work is made reproducible using the following tools.
-First the manuscript's source code, and the code reproducing all figures is available on \href{https://github.com/SpikeAI/2022_polychronies-review}{GitHub}.
+First the manuscript's source code, and the code reproducing all figures are available on \href{https://github.com/SpikeAI/XXX}{GitHub}.
 %~\citep{Grimaldi2023review}.
-This list also links to preprints versions of this review as well as links to previous versions.
+The paper is available as an \href{https://arxiv.org/abs/YYYY.XXX}{arXiv preprint} with links to previous versions.
 Find the associated \href{https://www.zotero.org/groups/4562620/polychronies}{Zotero group} which was used to regroup relevant literature on the subject.}
@@
 %\bibliographystyle{Definitions/chicago2}
+% \bibliography{references}
 \begin{thebibliography}{999}
@@
     foundation of electrophysiology.
-\newblock {\em Trends Neurosci.} {\bf 1997}, {\em 20},~443--448.
-\newblock {{https://doi.org/10.1016/S0166-2236(97)01101-6}}.
+  \newblock {\em Trends in Neurosciences} {\bf 1997}, {\em 20},~443--448.
+  \newblock {\url{https://doi.org/10.1016/S0166-2236(97)01101-6}}.
 
@@
   \newblock The impulses produced by sensory nerve endings.
-\newblock {\em  J. Physiol.} {\bf 1926}, {\em 61},~465--483.
+  \newblock {\em The Journal of Physiology} {\bf 1926}, {\em 61},~465--483.
 
@@
   \newblock Graded potentials of bream retina.
-\newblock {\em  J. Physiol.} {\bf 1960}, {\em 152},~487--505.
-
-\bibitem[Perkel {et~al.}(1967{\natexlab{a}})Perkel, Gerstein, and
+  \newblock {\em The Journal of Physiology} {\bf 1960}, {\em 152},~487--505.
+
+  \bibitem[Perkel \em{et~al.}(1967{\natexlab{a}})Perkel, Gerstein, and
     Moore]{perkel_neuronal_1967}
@@
     {The} {Single} {Spike} {Train}.
-\newblock {\em Biophys. J.} {\bf 1967}, {\em 7},~391--418.
-\newblock {{https://doi.org/10.1016/S0006-3495(67)86596-2}}.
-
-\bibitem[Perkel {et~al.}(1967{\natexlab{b}})Perkel, Gerstein, and
+  \newblock {\em Biophysical Journal} {\bf 1967}, {\em 7},~391--418.
+  \newblock {\url{https://doi.org/10.1016/S0006-3495(67)86596-2}}.
+
+  \bibitem[Perkel \em{et~al.}(1967{\natexlab{b}})Perkel, Gerstein, and
     Moore]{perkel_neuronal_1967-1}
@@
     {Simultaneous} {Spike} {Trains}.
-\newblock {\em Biophys. J.} {\bf 1967}, {\em 7},~419--440.
-\newblock {{https://doi.org/10.1016/S0006-3495(67)86597-4}}.
+  \newblock {\em Biophysical Journal} {\bf 1967}, {\em 7},~419--440.
+  \newblock {\url{https://doi.org/10.1016/S0006-3495(67)86597-4}}.
 
@@
   \newblock Role of the cortical neuron: integrator or coincidence detector?
-\newblock {\em Isr. J. Med. Sci.} {\bf 1982}, {\em
+  \newblock {\em Israel journal of medical sciences} {\bf 1982}, {\em
     18},~83--92.
@@
   \newblock Processing of {Temporal} {Information} in the {Brain}.
-\newblock {\em Annu. Rev. Neurosci.} {\bf 1993}, {\em 16},~223--243.\linebreak
-\newblock {{https://doi.org/10.1146/annurev.ne.16.030193.001255}}.
-
-\bibitem[Davis {et~al.}(2021)Davis, Benigno, Fletterman, Desbordes, Steward,
+  \newblock {\em Annual Review of Neuroscience} {\bf 1993}, {\em 16},~223--243.
+  \newblock {\url{https://doi.org/10.1146/annurev.ne.16.030193.001255}}.
+
+  \bibitem[Davis \em{et~al.}(2021)Davis, Benigno, Fletterman, Desbordes, Steward,
     Sejnowski, H~Reynolds, and Muller]{davis_spontaneous_2021}
   Davis, Z.W.; Benigno, G.B.; Fletterman, C.; Desbordes, T.; Steward, C.;
-  Sejnowski, T.J.;~Reynolds, J.H.; Muller, L.
+    Sejnowski, T.J.; H~Reynolds, J.; Muller, L.
   \newblock Spontaneous traveling waves naturally emerge from horizontal fiber
     time delays and travel through locally asynchronous-irregular states.
-\newblock {\em Nat. Commun.} {\bf 2021}, {\em 12},~6057.
-\newblock {{https://doi.org/10.1038/s41467-021-26175-1}}.
-
-\bibitem[Perrinet {et~al.}(2004)Perrinet, Samuelides, and
+  \newblock {\em Nature Communications} {\bf 2021}, {\em 12},~1--16.
+  \newblock {\url{https://doi.org/10.1038/s41467-021-26175-1}}.
+
+  \bibitem[Perrinet \em{et~al.}(2004)Perrinet, Samuelides, and
     Thorpe]{perrinet_coding_2004}
@@
     cooperate?
-\newblock {\em IEEE Trans. Neural Netw.} {\bf 2004}, {\em
+  \newblock {\em IEEE Transactions on neural networks} {\bf 2004}, {\em
     15},~1164--1175.
- {{https://doi.org/10.1109/TNN.2004.833303}}.
+  \newblock Publisher: IEEE, {\url{https://doi.org/10.1109/TNN.2004.833303}}.
 
@@
   \newblock {\em Science} {\bf 2008}, {\em 319},~1108--1111.
-\newblock {{https://doi.org/10.1126/science.1149639}}.
+  \newblock {\url{https://doi.org/10.1126/science.1149639}}.
 
@@
   DeWeese, M.R.; Zador, A.M.
-\newblock {\em Binary Coding in Auditory Cortex}; {Neural Information Processing
-  Systems Foundation:} 
-  \newblock MIT Press: Cambridge, MA, USA,
-  %MDPI: Please add the location of the publisher (City and Country).
-   2003.
+  \newblock {\em Binary coding in auditory cortex}; Neural Information Processing
+    Systems Foundation,  2003.
 
@@
     stem of the barn owl.
-\newblock {\em J. Neurosci.} {\bf 1990}, {\em 10},~3227--3246.
-%\newblock Publisher: Society for Neuroscience Section: Articles,
-  {{https://doi.org/10.1523/JNEUROSCI.10-10-03227.1990}}.
+  \newblock {\em Journal of Neuroscience} {\bf 1990}, {\em 10},~3227--3246.
+  \newblock Publisher: Society for Neuroscience Section: Articles,
+    {\url{https://doi.org/10.1523/JNEUROSCI.10-10-03227.1990}}.
 
@@
     spike-times: {A} survey.
-\newblock {\em Nat. Comput.} {\bf 2004}, {\em 3},~195--206.
-\newblock {{https://doi.org/10.1023/B:NACO.0000027755.02868.60}}.
+  \newblock {\em Natural Computing} {\bf 2004}, {\em 3},~195--206.
+  \newblock {\url{https://doi.org/10.1023/B:NACO.0000027755.02868.60}}.
 
@@
   DiLorenzo, P.M.; Victor, J.D.
-\newblock {\em Spike {Timing}: {Mechanisms} and {Function}}; CRC Press:  {Boca Raton, FL, USA,} %newly added information, please confirm
-  2013.
-%\newblock Google-Books-ID: KTHUIMUpQCUC.  %MDPI: We removed Google-Books-ID: KTHUIMUpQCUC. Please confirm this revision.
-
-\bibitem[Roy {et~al.}(2019)Roy, Jaiswal, and Panda]{roy_towards_2019}
+  \newblock {\em Spike {Timing}: {Mechanisms} and {Function}}; CRC Press,  2013.
+  \newblock Google-Books-ID: KTHUIMUpQCUC.
+
+  \bibitem[Roy \em{et~al.}(2019)Roy, Jaiswal, and Panda]{roy_towards_2019}
   Roy, K.; Jaiswal, A.; Panda, P.
@@
   \newblock {\em Nature} {\bf 2019}, {\em 575},~607--617.
-%\newblock Number: 7784 Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/s41586-019-1677-2}}.
+  \newblock Number: 7784 Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/s41586-019-1677-2}}.
 
@@
   Flourens, M.J.P.
-\newblock {\em Recherches Expérimentales sur les Propriétés et les Fonctions
-  du Système Nerveux, Dans les Animaux Vertébrés}; Hachette {BNF}: {Paris, France,} {1842.} %MDPI: Please add the name of the publisher and their location.
+  \newblock {\em Recherches expérimentales sur les propriétés et les fonctions
+    du système nerveux, dans les animaux vertébrés}; J.-B. Balliere,  1842.
 
@@
   Pearce, J.
-\newblock Marie-{Jean}-{Pierre} {Flourens} (1794--1867) and {Cortical}
+  \newblock Marie-{Jean}-{Pierre} {Flourens} (1794–1867) and {Cortical}
     {Localization}.
-\newblock {\em Eur. Neurol.} {\bf 2009}, {\em 61},~311--314.\linebreak
-\newblock {{https://doi.org/10.1159/000206858}}.
+  \newblock {\em European Neurology} {\bf 2009}, {\em 61},~311--314.
+  \newblock {\url{https://doi.org/10.1159/000206858}}.
 
@@
     cortex.
-\newblock {\em  J. Physiol.} {\bf 1968}, {\em 195},~215--243.
-\newblock {{https://doi.org/10.1113/jphysiol.1968.sp008455}}.
+  \newblock {\em The Journal of Physiology} {\bf 1968}, {\em 195},~215--243.
+  \newblock {\url{https://doi.org/10.1113/jphysiol.1968.sp008455}}.
 
@@
   \newblock Normalization as a canonical neural computation.
-\newblock {\em Nat. Rev. Neurosci.} {\bf 2012}, {\em 13},~51--62.
-\newblock {{https://doi.org/10.1038/nrn3136}}.
-
-\bibitem[Thorpe {et~al.}(1996)Thorpe, Fize, and Marlot]{thorpe_speed_1996}
+  \newblock {\em Nature Reviews Neuroscience} {\bf 2012}, {\em 13},~51--62.
+  \newblock {\url{https://doi.org/10.1038/nrn3136}}.
+
+  \bibitem[Thorpe \em{et~al.}(1996)Thorpe, Fize, and Marlot]{thorpe_speed_1996}
   Thorpe, S.; Fize, D.; Marlot, C.
   \newblock Speed of processing in the human visual system.
-\newblock {\em Nature} {\bf 1996}, {\em 381},~520--522.\linebreak
-\newblock {{https://doi.org/10.1038/381520a0}}.
+  \newblock {\em Nature} {\bf 1996}, {\em 381},~520--522.
+  \newblock {\url{https://doi.org/10.1038/381520a0}}.
 
@@
     processing speed revisited.
-\newblock {\em Vis. Res.} {\bf 2006}, {\em 46},~1762--1776.
-%\newblock Loaded from an external bibliography file by Manubot.
-%  source\_bibliography: manual-references.bib standard\_id: Kirchner06,
-  {{https://doi.org/10.1016/j.visres.2005.10.002}}.
-
-\bibitem[Keysers {et~al.}(2001)Keysers, Xiao, Földiák, and
+  \newblock {\em Vision Research} {\bf 2006}, {\em 46},~1762--76.
+  \newblock Loaded from an external bibliography file by Manubot.
+    source\_bibliography: manual-references.bib standard\_id: Kirchner06,
+    {\url{https://doi.org/10.1016/j.visres.2005.10.002}}.
+
+  \bibitem[Keysers \em{et~al.}(2001)Keysers, Xiao, Földiák, and
     Perrett]{keysers_speed_2001}
@@
   \newblock The {Speed} of {Sight}.
-\newblock {\em J. Cogn. Neurosci.} {\bf 2001}, {\em
-  13},~90--101.\linebreak
-\newblock {{https://doi.org/10.1162/089892901564199}}.
-
-\bibitem[Schmolesky {et~al.}(1998)Schmolesky, Wang, Hanes, Thompson,
+  \newblock {\em Journal of Cognitive Neuroscience} {\bf 2001}, {\em
+    13},~90--101.
+  \newblock {\url{https://doi.org/10.1162/089892901564199}}.
+
+  \bibitem[Schmolesky \em{et~al.}(1998)Schmolesky, Wang, Hanes, Thompson,
     Leutgeb, Schall, and Leventhal]{schmolesky_signal_1998}
@@
   \newblock Signal timing across the macaque visual system.
-\newblock {\em J. Neurophysiol.} {\bf 1998}, {\em 79},~3272--3278.
-\newblock {{https://doi.org/10.1152/jn.1998.79.6.3272}}.
-
-\bibitem[Vanni {et~al.}(2001)Vanni, Tanskanen, Seppä, Uutela, and
+  \newblock {\em Journal of Neurophysiology} {\bf 1998}, {\em 79},~3272--3278.
+  \newblock {\url{https://doi.org/10.1152/jn.1998.79.6.3272}}.
+
+  \bibitem[Vanni \em{et~al.}(2001)Vanni, Tanskanen, Seppä, Uutela, and
     Hari]{vanni_coinciding_2001}
@@
     anteromedial cuneus.
-\newblock {\em Proc. Natl. Acad. Sci.  USA} {\bf 2001}, {\em 98},~2776--2780.
-\newblock {{https://doi.org/10.1073/pnas.041600898}}.
+  \newblock {\em Proceedings of the National Academy of Sciences of the United
+    States of America} {\bf 2001}, {\em 98},~2776--2780.
+  \newblock {\url{https://doi.org/10.1073/pnas.041600898}}.
 
@@
     processing.
-\newblock {\em Trends Neurosci.} {\bf 2000}, {\em 23},~571--579.
-\newblock {{https://doi.org/10.1016/s0166-2236(00)01657-x}}.
-
-\bibitem[Serre {et~al.}(2007)Serre, Oliva, and
+  \newblock {\em Trends in Neurosciences} {\bf 2000}, {\em 23},~571--579.
+  \newblock {\url{https://doi.org/10.1016/s0166-2236(00)01657-x}}.
+
+  \bibitem[Serre \em{et~al.}(2007)Serre, Oliva, and
     Poggio]{serre_feedforward_2007}
@@
   \newblock A feedforward architecture accounts for rapid categorization.
-\newblock {\em Proc. Natl. Acad. Sci.  USA} {\bf 2007}, {\em 104},~6424--6429.
-%\newblock tex.bdsk-url-2: https://doi.org/10.1073/pnas.0700622104
-%  tex.date-added: 2022-05-05 19:08:15 +0200 tex.date-modified: 2022-05-05
-%  19:08:15 +0200, 
-  {{https://doi.org/10.1073/pnas.0700622104}}.
+  \newblock {\em Proceedings of the National Academy of Sciences of the United
+    States of America} {\bf 2007}, {\em 104},~6424--6429.
+  \newblock tex.bdsk-url-2: https://doi.org/10.1073/pnas.0700622104
+    tex.date-added: 2022-05-05 19:08:15 +0200 tex.date-modified: 2022-05-05
+    19:08:15 +0200, {\url{https://doi.org/10.1073/pnas.0700622104}}.
 
@@
   Jérémie, J.N.; Perrinet, L.U.
-\newblock Ultrafast image categorization in vivo and in silico. \emph{arXiv}  \textbf{2022}, 
-\newblock arXiv:2205.03635.
+  \newblock Ultrafast image categorization in vivo and in silico,  2022.
+  \newblock arXiv:2205.03635 [cs, q-bio].
 
@@
   \newblock The {Timing} of {Information} {Transfer} in the {Visual} {System}. In
-  {\em Extrastriate {Cortex} in {Primates}}; Springer: {Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
-  1997; pp. 205--241.
+    {\em Extrastriate {Cortex} in {Primates}}; Springer US,  1997; pp. 205--241.
 
@@
   \newblock {\em Science} {\bf 2001}, {\em 291},~260--263.
-\newblock {{https://doi.org/10.1126/science.1058249}}.
-
-\bibitem[Rucci {et~al.}(2018)Rucci, Ahissar, and Burr]{rucci_temporal_2018}
+  \newblock {\url{https://doi.org/10.1126/science.1058249}}.
+
+  \bibitem[Rucci \em{et~al.}(2018)Rucci, Ahissar, and Burr]{rucci_temporal_2018}
   Rucci, M.; Ahissar, E.; Burr, D.
   \newblock Temporal {Coding} of {Visual} {Space}.
-\newblock {\em Trends Cogn. Sci.} {\bf 2018}, {\em 22},~883--895.\linebreak
-\newblock {{https://doi.org/10.1016/j.tics.2018.07.009}}.
+  \newblock {\em Trends in Cognitive Sciences} {\bf 2018}, {\em 22},~883--895.
+  \newblock {\url{https://doi.org/10.1016/j.tics.2018.07.009}}.
 
@@
     temporal integration of random {EPSPs}.
-\newblock {\em  J. Neurosci.} {\bf 1993}, {\em 13},~334--350.
-\newblock {{https://doi.org/10.1523/jneurosci.13-01-00334.1993}}.
+  \newblock {\em The Journal of Neuroscience} {\bf 1993}, {\em 13},~334--350.
+  \newblock {\url{https://doi.org/10.1523/jneurosci.13-01-00334.1993}}.
 
@@
   \newblock Spike initiation by transmembrane current: a white-noise analysis.
-\newblock {\em  J. Physiol.} {\bf 1976}, {\em 260},~279--314.
-%\newblock \_eprint:
-%  https://onlinelibrary.wiley.com/doi/pdf/10.1113/jphysiol.1976.sp011516,
-  {{https://doi.org/10.1113/jphysiol.1976.sp011516}}.
+  \newblock {\em The Journal of Physiology} {\bf 1976}, {\em 260},~279--314.
+  \newblock \_eprint:
+    https://onlinelibrary.wiley.com/doi/pdf/10.1113/jphysiol.1976.sp011516,
+    {\url{https://doi.org/10.1113/jphysiol.1976.sp011516}}.
 
@@
   \newblock Reliability of {Spike} {Timing} in {Neocortical} {Neurons}.
-\newblock {\em Science} {\bf 1995}, {\em 268},~1503--1506.\linebreak
-\newblock {{https://doi.org/10.1126/science.7770778}}.
-
-\bibitem[Ermentrout {et~al.}(2008)Ermentrout, Galán, and
+  \newblock {\em Science} {\bf 1995}, {\em 268},~1503--1506.
+  \newblock {\url{https://doi.org/10.1126/science.7770778}}.
+
+  \bibitem[Ermentrout \em{et~al.}(2008)Ermentrout, Galán, and
     Urban]{ermentrout_reliability_2008}
@@
   \newblock Reliability, synchrony and noise.
-\newblock {\em Trends Neurosci.} {\bf 2008}, {\em 31},~428--434.
-\newblock {{https://doi.org/10.1016/j.tins.2008.06.002}}.
+  \newblock {\em Trends in neurosciences} {\bf 2008}, {\em 31},~428--434.
+  \newblock {\url{https://doi.org/10.1016/j.tins.2008.06.002}}.
 
@@
     cortical neurons.
-\newblock {\em Cereb. Cortex} {\bf 1997}, {\em 7},~487--501.
-\newblock {{https://doi.org/10.1093/cercor/7.6.487}}.
-
-\bibitem[Kenyon {et~al.}(2004)Kenyon, Hill, Theiler, George, and
+  \newblock {\em Cerebral Cortex} {\bf 1997}, {\em 7},~487--501.
+  \newblock {\url{https://doi.org/10.1093/cercor/7.6.487}}.
+
+  \bibitem[Kenyon \em{et~al.}(2004)Kenyon, Hill, Theiler, George, and
     Marshak]{kenyon_theory_2004}
@@
     interactions in the parvocellular pathway.
-\newblock {\em Neural Netw.} {\bf 2004}, {\em 17},~773--786.
-\newblock {{https://doi.org/10.1016/j.neunet.2004.05.005}}.
-
-\bibitem[Celebrini {et~al.}(1993)Celebrini, Thorpe, Trotter, and
+  \newblock {\em Neural Networks} {\bf 2004}, {\em 17},~773--786.
+  \newblock {\url{https://doi.org/10.1016/j.neunet.2004.05.005}}.
+
+  \bibitem[Celebrini \em{et~al.}(1993)Celebrini, Thorpe, Trotter, and
     Imbert]{celebrini_dynamics_1993}
@@
   \newblock Dynamics of orientation coding in area {V1} of the awake primate.
-\newblock {\em Vis. Neurosci.} {\bf 1993}, {\em 10},~811--825.
-\newblock {{https://doi.org/10.1017/s0952523800006052}}.
+  \newblock {\em Visual Neuroscience} {\bf 1993}, {\em 10},~811--825.
+  \newblock {\url{https://doi.org/10.1017/s0952523800006052}}.
 
@@
     referenced to population onset.
-\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2007},
+  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2007},
     {\em 104},~5175--5180.
-\newblock {{https://doi.org/10.1073/pnas.0610368104}}.
-
-\bibitem[Safaie {et~al.}(2020)Safaie, Jurado-Parras, Sarno, Louis,
+  \newblock {\url{https://doi.org/10.1073/pnas.0610368104}}.
+
+  \bibitem[Safaie \em{et~al.}(2020)Safaie, Jurado-Parras, Sarno, Louis,
     Karoutchi, Petit, Pasquet, Eloy, and Robbe]{safaie_turning_2020}
@@
     simple stereotyped interactions with the environment.
-\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2020},
+  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2020},
     {\em 117},~13084--13093.
-\newblock {{https://doi.org/10.1073/pnas.1921226117}}.
+  \newblock {\url{https://doi.org/10.1073/pnas.1921226117}}.
 
@@
   \newblock {\em Biosystems} {\bf 1998}, {\em 48},~57--65.
-%\newblock Publisher: Elsevier,
-  {{https://doi.org/10.1016/S0303-2647(98)00050-1}}.
-
-\bibitem[Delorme {et~al.}(1999)Delorme, Gautrais, Van~Rullen, and
+  \newblock Publisher: Elsevier,
+    {\url{https://doi.org/10.1016/S0303-2647(98)00050-1}}.
+
+  \bibitem[Delorme \em{et~al.}(1999)Delorme, Gautrais, Van~Rullen, and
     Thorpe]{delorme_spikenet_1999}
@@
   \newblock {\em Neurocomputing} {\bf 1999}, {\em 26},~989--996.
-%\newblock Publisher: Elsevier,
-  {{https://doi.org/10.1016/S0925-2312(99)00095-8}}.
-
-\bibitem[Delorme {et~al.}(2000)Delorme, Richard, and
+  \newblock Publisher: Elsevier,
+    {\url{https://doi.org/10.1016/S0925-2312(99)00095-8}}.
+
+  \bibitem[Delorme \em{et~al.}(2000)Delorme, Richard, and
     Fabre-Thorpe]{delorme_ultra-rapid_2000}
@@
     cues: a study in monkeys and humans.
-\newblock {\em Vis. Res.} {\bf 2000}, {\em 40},~2187--2200.
-\newblock {{https://doi.org/10.1016/S0042-6989(00)00083-3}}.
-
-\bibitem[Bonilla {et~al.}(2022)Bonilla, Gautrais, Thorpe, and
+  \newblock {\em Vision Research} {\bf 2000}, {\em 40},~2187--2200.
+  \newblock {\url{https://doi.org/10.1016/S0042-6989(00)00083-3}}.
+
+  \bibitem[Bonilla \em{et~al.}(2022)Bonilla, Gautrais, Thorpe, and
     Masquelier]{bonilla_analyzing_2022}
@@
   \newblock Analyzing time-to-first-spike coding schemes.
-\newblock {\em Front. Neurosci.} {\bf 2022}, {\em 16}, 971937.
-%\newblock Publisher: Frontiers,
-  {{https://doi.org/10.3389/fnins.2022.971937}}.
-
-\bibitem[Bohte {et~al.}(2002)Bohte, Kok, and
+  \newblock {\em Frontiers in Neuroscience} {\bf 2022}, {\em 16}.
+  \newblock Publisher: Frontiers,
+    {\url{https://doi.org/10.3389/fnins.2022.971937}}.
+
+  \bibitem[Bohte \em{et~al.}(2002)Bohte, Kok, and
     La~Poutré]{bohte_error-backpropagation_2002}
@@
   \newblock {\em Neurocomputing} {\bf 2002}, {\em 48},~17--37.
-\newblock {{https://doi.org/10.1016/S0925-2312(01)00658-0}}.
+  \newblock {\url{https://doi.org/10.1016/S0925-2312(01)00658-0}}.
 
@@
     for {Instilling} {Complex} {Function} in {Spiking} {Neural} {Networks}.
-\newblock {\em Neural Comput.} {\bf 2021}, {\em 33},~899--925.
-\newblock {{https://doi.org/10.1162/neco\_a\_01367}}.
-
-\bibitem[Göltz {et~al.}(2021)Göltz, Kriener, Baumbach, Billaudelle,
+  \newblock {\em Neural Computation} {\bf 2021}, {\em 33},~899--925.
+  \newblock {\url{https://doi.org/10.1162/neco_a_01367}}.
+
+  \bibitem[Göltz \em{et~al.}(2021)Göltz, Kriener, Baumbach, Billaudelle,
     Breitwieser, Cramer, Dold, Kungl, Senn, Schemmel, Meier, and
@@
   \newblock Fast and energy-efficient neuromorphic deep learning with first-spike
-  times. \emph{arXiv}  \textbf{2021}, arXiv:1912.11443.
-
-\bibitem[Kheradpisheh {et~al.}(2018)Kheradpisheh, Ganjtabesh, Thorpe, and
+    times.
+  \newblock Technical Report 1912.11443, arXiv,  2021.
+
+  \bibitem[Kheradpisheh \em{et~al.}(2018)Kheradpisheh, Ganjtabesh, Thorpe, and
     Masquelier]{kheradpisheh_stdp-based_2018}
@@
     recognition.
-\newblock {\em Neural Netw.} {\bf 2018}, {\em 99},~56--67.
-\newblock {{https://doi.org/10.1016/j.neunet.2017.12.005}}.
-
-\bibitem[Tavanaei {et~al.}(2018)Tavanaei, Masquelier, and
+  \newblock {\em Neural Networks} {\bf 2018}, {\em 99},~56--67.
+  \newblock {\url{https://doi.org/10.1016/j.neunet.2017.12.005}}.
+
+  \bibitem[Tavanaei \em{et~al.}(2018)Tavanaei, Masquelier, and
     Maida]{tavanaei_representation_2018}
@@
   \newblock Representation learning using event-based {STDP}.
-\newblock {\em Neural Netw.} {\bf 2018}, {\em 105},~294--303.
-\newblock {{https://doi.org/10.1016/j.neunet.2018.05.018}}.
-
-\bibitem[Gallego {et~al.}(2022)Gallego, Delbruck, Orchard, Bartolozzi, Taba,
+  \newblock {\em Neural Networks} {\bf 2018}, {\em 105},~294--303.
+  \newblock {\url{https://doi.org/10.1016/j.neunet.2018.05.018}}.
+
+  \bibitem[Gallego \em{et~al.}(2022)Gallego, Delbruck, Orchard, Bartolozzi, Taba,
     Censi, Leutenegger, Davison, Conradt, Daniilidis, and
@@
   \newblock Event-{Based} {Vision}: {A} {Survey}.
-\newblock {\em IEEE Trans. Pattern Anal. Mach. Intell.}
+  \newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}
     {\bf 2022}, {\em 44},~154--180.
-\newblock {{https://doi.org/10.1109/TPAMI.2020.3008413}}.
+  \newblock {\url{https://doi.org/10.1109/TPAMI.2020.3008413}}.
 
@@
     orientation.
-\newblock {\em J. Neurophysiol.} {\bf 1983}, {\em 49},~1127--1147.
-\newblock {{https://doi.org/10.1152/jn.1983.49.5.1127}}.
-
-\bibitem[Montemurro {et~al.}(2008)Montemurro, Rasch, Murayama, Logothetis,
+  \newblock {\em Journal of Neurophysiology} {\bf 1983}, {\em 49},~1127--1147.
+  \newblock {\url{https://doi.org/10.1152/jn.1983.49.5.1127}}.
+
+  \bibitem[Montemurro \em{et~al.}(2008)Montemurro, Rasch, Murayama, Logothetis,
     and Panzeri]{montemurro_phase--firing_2008}
@@
     {Primary} {Visual} {Cortex}.
-\newblock {\em Curr. Biol.} {\bf 2008}, {\em 18},~375--380.
-%\newblock Publisher: Elsevier,
-  {{https://doi.org/10.1016/j.cub.2008.02.023}}.
+  \newblock {\em Current Biology} {\bf 2008}, {\em 18},~375--380.
+  \newblock Publisher: Elsevier,
+    {\url{https://doi.org/10.1016/j.cub.2008.02.023}}.
 
@@
   \newblock {\em Nature} {\bf 1996}, {\em 381},~610--613.
-%\newblock Number: 6583 Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/381610a0}}.
+  \newblock Number: 6583 Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/381610a0}}.
 
@@
     {During} {Natural} {Vision}.
-\newblock {\em Science} {\bf 2000}, {\em 287}, 1273--1276.
-%\newblock Publisher: American Association for the Advancement of Science,
-  {{https://doi.org/10.1126/science.287.5456.1273}}.
+  \newblock {\em Science} {\bf 2000}, {\em 287},~1273--1276.
+  \newblock Publisher: American Association for the Advancement of Science,
+    {\url{https://doi.org/10.1126/science.287.5456.1273}}.
 
@@
   Abeles, M.
-\newblock {\em Corticonics: Neural Circuits of the Cerebral Cortex}; Cambridge
-  University Press: Cambridge, MA, USA; New York, NY, USA, 1991.
+  \newblock {\em Corticonics: neural circuits of the cerebral cortex}; Cambridge
+    University Press: Cambridge ; New York,  1991.
 
@@
   Paugam-Moisy, H.; Bohte, S.M.
-\newblock Computing with spiking neuron networks. In {\em Handbook of Natural
-  Computing}; Springer:  {Berlin/Heidelberg, Germany,} %newly added information, please confirm
-  2012.
+  \newblock Computing with spiking neuron networks. In {\em Handbook of natural
+    computing}; Springer-Verlag,  2012.
 
@@
   Hebb, D.O.
-\newblock {\em The Organization of Behavior: {A} Neuropsychological Theory};
-  Wiley: New York, NY, USA, 1949.
-
-\bibitem[Harris {et~al.}(2003)Harris, Csicsvari, Hirase, Dragoi, and
+  \newblock {\em The organization of behavior: {A} neuropsychological theory};
+    Wiley: New York,  1949.
+
+  \bibitem[Harris \em{et~al.}(2003)Harris, Csicsvari, Hirase, Dragoi, and
     Buzsáki]{harris_organization_2003}
@@
   \newblock {\em Nature} {\bf 2003}, {\em 424},~552--556.
-\newblock {{https://doi.org/10.1038/nature01834}}.
+  \newblock {\url{https://doi.org/10.1038/nature01834}}.
 
@@
     {Hypothesis}.
-\newblock {\em Annu. Rev. Neurosci.} {\bf 1995}, {\em 18},~555--586.
-\newblock {{https://doi.org/10.1146/annurev.ne.18.030195.003011}}.
-
-\bibitem[Roelfsema {et~al.}(1997)Roelfsema, Engel, König, and
+  \newblock {\em Annual Review of Neuroscience} {\bf 1995}, {\em 18},~555--586.
+  \newblock {\url{https://doi.org/10.1146/annurev.ne.18.030195.003011}}.
+
+  \bibitem[Roelfsema \em{et~al.}(1997)Roelfsema, Engel, König, and
     Singer]{roelfsema_visuomotor_1997}
@@
   \newblock {\em Nature} {\bf 1997}, {\em 385},~157--161.
-\newblock {{https://doi.org/10.1038/385157a0}}.
-
-\bibitem[Riehle {et~al.}(1997)Riehle, Grun, Diesmann, and
+  \newblock {\url{https://doi.org/10.1038/385157a0}}.
+
+  \bibitem[Riehle \em{et~al.}(1997)Riehle, Grun, Diesmann, and
     Aertsen]{riehle_spike_1997}
@@
     motor cortical function.
-\newblock {\em Science} {\bf 1997}, {\em 278},~1950--1953.
-%\newblock Publisher: American Association for the Advancement of Science,
-  {{https://doi.org/10.1126/science.278.5345.1950}}.
-
-\bibitem[Kilavik {et~al.}(2009)Kilavik, Roux, Ponce-Alvarez, Confais, Grun,
+  \newblock {\em Science (New York, N.Y.)} {\bf 1997}, {\em 278},~1950--1953.
+  \newblock Publisher: American Association for the Advancement of Science,
+    {\url{https://doi.org/10.1126/science.278.5345.1950}}.
+
+  \bibitem[Kilavik \em{et~al.}(2009)Kilavik, Roux, Ponce-Alvarez, Confais, Grun,
     and Riehle]{kilavik_long-term_2009}
@@
     {Induced} by {Intensive} {Practice}.
-\newblock {\em J. Neurosci.} {\bf 2009}, {\em 29},~12653--12663.
-\newblock {{https://doi.org/10.1523/jneurosci.1554-09.2009}}.
+  \newblock {\em Journal of Neuroscience} {\bf 2009}, {\em 29},~12653--12663.
+  \newblock {\url{https://doi.org/10.1523/jneurosci.1554-09.2009}}.
 
@@
     cortical neurons in relation to movement direction and reaction time.
-\newblock {\em Biol. Cybern.} {\bf 2003}, {\em 88},~360--373.
-\newblock {{https://doi.org/10.1007/s00422-002-0385-3}}.
-
-\bibitem[Denker {et~al.}(2018)Denker, Zehl, Kilavik, Diesmann, Brochier,
+  \newblock {\em Biological Cybernetics} {\bf 2003}, {\em 88},~360--373.
+  \newblock {\url{https://doi.org/10.1007/s00422-002-0385-3}}.
+
+  \bibitem[Denker \em{et~al.}(2018)Denker, Zehl, Kilavik, Diesmann, Brochier,
     Riehle, and Grün]{denker_lfp_2018}
@@
     patterns.
-\newblock {\em Sci. Rep.} {\bf 2018}, {\em 8}.
-\newblock {{https://doi.org/10.1038/s41598-018-22990-7}}.
-
-\bibitem[Torre {et~al.}(2016)Torre, Canova, Denker, Gerstein, Helias, and
+  \newblock {\em Scientific Reports} {\bf 2018}, {\em 8}.
+  \newblock {\url{https://doi.org/10.1038/s41598-018-22990-7}}.
+
+  \bibitem[Torre \em{et~al.}(2016)Torre, Canova, Denker, Gerstein, Helias, and
     Grün]{torre_asset_2016}
@@
     {Massively} {Parallel} {Spike} {Trains}.
-\newblock {\em PLOS Comput. Biol.} {\bf 2016}, {\em 12},~e1004939.
-\newblock {{https://doi.org/10.1371/journal.pcbi.1004939}}.
+  \newblock {\em PLOS Computational Biology} {\bf 2016}, {\em 12},~e1004939.
+  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1004939}}.
 
@@
   \newblock Traveling {Waves} and the {Processing} of {Weakly} {Tuned} {Inputs}
-  in a {Cortical} {Network} {Module}.  \emph{J. Comput. Neurosci.} {\bf 1997}.
+    in a {Cortical} {Network} {Module} {\bf 1997}.
   \newblock {\em 77},~57--77.
+  \newblock 00000.
 
@@
   \newblock {\em Science} {\bf 2006}, {\em 312},~1622--1627.
-\newblock {{https://doi.org/10.1126/science.1124593}}.
+  \newblock {\url{https://doi.org/10.1126/science.1124593}}.
 
@@
   \newblock Bayesian inference in spiking neurons.
-\newblock In \emph{Proceedings of the Advances in {Neural} {Information} {Processing}
-  {Systems}}; MIT Press: Cambridge, MA, USA, %newly added information, please confirm
-  2004, Volume~17.
+  \newblock In Proceedings of the Advances in {Neural} {Information} {Processing}
+    {Systems}. MIT Press,  2004, Vol.~17.
 
@@
     {Populations}.
-\newblock {\em Front. Comput. Neurosci.} {\bf 2011}, {\em 5}, 22.
-
-\bibitem[Gewaltig {et~al.}(2001)Gewaltig, Diesmann, and
+  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2011}, {\em 5}.
+
+  \bibitem[Gewaltig \em{et~al.}(2001)Gewaltig, Diesmann, and
     Aertsen]{gewaltig_propagation_2001}
@@
     single trials and stability in the mean.
-\newblock {\em Neural Netw.} {\bf 2001}, {\em 14},~657--673.
-\newblock {{https://doi.org/10.1016/S0893-6080(01)00070-3}}.
+  \newblock {\em Neural Networks} {\bf 2001}, {\em 14},~657--673.
+  \newblock {\url{https://doi.org/10.1016/S0893-6080(01)00070-3}}.
 
@@
   \newblock Time structure of the activity in neural network models.
-\newblock {\em Phys. Rev. E} {\bf 1995}, {\em 51},~738--758.\linebreak
-\newblock {{https://doi.org/10.1103/physreve.51.738}}.
+  \newblock {\em Physical Review E} {\bf 1995}, {\em 51},~738--758.
+  \newblock {\url{https://doi.org/10.1103/physreve.51.738}}.
 
@@
   \newblock Stimulus-selective spiking is driven by the relative timing of
-  synchronous excitation and disinhibition in cat striate neurons {in
+    synchronous excitation and disinhibition in cat striate neurons\textit{in
     vivo}.
-\newblock {\em Eur. J. Neurosci.} {\bf 2008}, {\em
+  \newblock {\em European Journal of Neuroscience} {\bf 2008}, {\em
     28},~1286--1300.
-\newblock {{https://doi.org/10.1111/j.1460-9568.2008.06434.x}}.
-
-\bibitem[Kremkow {et~al.}(2016)Kremkow, Perrinet, Monier, Alonso, Aertsen,
+  \newblock {\url{https://doi.org/10.1111/j.1460-9568.2008.06434.x}}.
+
+  \bibitem[Kremkow \em{et~al.}(2016)Kremkow, Perrinet, Monier, Alonso, Aertsen,
     Frégnac, and Masson]{kremkow_push-pull_2016}
@@
     in {V1}.
-\newblock {\em Front. Neural Circuits} {\bf 2016}, {\em 10}, 37.
-\newblock {{https://doi.org/10.3389/fncir.2016.00037}}.
-
-\bibitem[Aviel {et~al.}(2003)Aviel, Mehring, Abeles, and
+  \newblock {\em Frontiers in Neural Circuits} {\bf 2016}, {\em 10}.
+  \newblock {\url{https://doi.org/10.3389/fncir.2016.00037}}.
+
+  \bibitem[Aviel \em{et~al.}(2003)Aviel, Mehring, Abeles, and
     Horn]{aviel_embedding_2003}
@@
   \newblock On {Embedding} {Synfire} {Chains} in a {Balanced} {Network}.
-\newblock {\em Neural Comput.} {\bf 2003}, {\em 15},~1321--1340.
-\newblock {{https://doi.org/10.1162/089976603321780290}}.
-
-\bibitem[Kremkow {et~al.}(2010)Kremkow, Perrinet, Masson, and
+  \newblock {\em Neural Computation} {\bf 2003}, {\em 15},~1321--1340.
+  \newblock {\url{https://doi.org/10.1162/089976603321780290}}.
+
+  \bibitem[Kremkow \em{et~al.}(2010)Kremkow, Perrinet, Masson, and
     Aertsen]{kremkow_functional_2010}
@@
     conductances in cortical networks.
-\newblock {\em J. Comput. Neurosci.} {\bf 2010}, {\em
+  \newblock {\em Journal of Computational Neuroscience} {\bf 2010}, {\em
     28},~579--594.
-\newblock {{https://doi.org/10.1007/s10827-010-0240-9}}.
+  \newblock {\url{https://doi.org/10.1007/s10827-010-0240-9}}.
 
@@
   Davison, A.P.
-\newblock {PyNN}: A common interface for neuronal network simulators.
-\newblock {\em Front. Neuroinformatics} {\bf 2008}, {\em 2}, 11.
-\newblock {{https://doi.org/10.3389/neuro.11.011.2008}}.
-
-\bibitem[Pfeil {et~al.}(2013)Pfeil, Grübl, Jeltsch, Müller, Müller,
+  \newblock {PyNN}: a common interface for neuronal network simulators.
+  \newblock {\em Frontiers in Neuroinformatics} {\bf 2008}, {\em 2}.
+  \newblock {\url{https://doi.org/10.3389/neuro.11.011.2008}}.
+
+  \bibitem[Pfeil \em{et~al.}(2013)Pfeil, Grübl, Jeltsch, Müller, Müller,
     Petrovici, Schmuker, Brüderle, Schemmel, and Meier]{pfeil_six_2013}
@@
     {Substrate}.
-\newblock {\em Front. Neurosci.} {\bf 2013}, {\em 7}, 11.
-\newblock {{https://doi.org/10.3389/fnins.2013.00011}}.
-
-\bibitem[Schrader {et~al.}(2008)Schrader, Grün, Diesmann, and
+  \newblock {\em Frontiers in Neuroscience} {\bf 2013}, {\em 7}.
+  \newblock {\url{https://doi.org/10.3389/fnins.2013.00011}}.
+
+  \bibitem[Schrader \em{et~al.}(2008)Schrader, Grün, Diesmann, and
     Gerstein]{schrader_detecting_2008}
@@
     {Spike} {Train} {Recording}.
-\newblock {\em J. Neurophysiol.} {\bf 2008}, {\em 100},~2165--2176.
-\newblock {{https://doi.org/10.1152/jn.01245.2007}}.
+  \newblock {\em Journal of Neurophysiology} {\bf 2008}, {\em 100},~2165--2176.
+  \newblock {\url{https://doi.org/10.1152/jn.01245.2007}}.
 
@@
     preparation for movement.
-\newblock {\em Exp. Brain Res.} {\bf 1999}, {\em 128},~118--122.
-\newblock {{https://doi.org/10.1007/s002210050826}}.
+  \newblock {\em Experimental Brain Research} {\bf 1999}, {\em 128},~118--122.
+  \newblock {\url{https://doi.org/10.1007/s002210050826}}.
 
@@
   \newblock Computing with {Neural} {Synchrony}.
-\newblock {\em PLoS Comput. Biol.} {\bf 2012}, {\em 8},~e1002561.
-\newblock {{https://doi.org/10.1371/journal.pcbi.1002561}}.
+  \newblock {\em PLoS Computational Biology} {\bf 2012}, {\em 8},~e1002561.
+  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1002561}}.
 
@@
     neuronal coherence.
-\newblock {\em Trends Cogn. Sci.} {\bf 2005}, {\em 9},~474--480.
-\newblock {{https://doi.org/10.1016/j.tics.2005.08.011}}.
-
-\bibitem[VanRullen {et~al.}(2006)VanRullen, Reddy, and
+  \newblock {\em Trends in Cognitive Sciences} {\bf 2005}, {\em 9},~474--480.
+  \newblock {\url{https://doi.org/10.1016/j.tics.2005.08.011}}.
+
+  \bibitem[VanRullen \em{et~al.}(2006)VanRullen, Reddy, and
     Koch]{vanrullen_continuous_2006}
@@
     {Changes} in {Electroencephalogram} {Power} at 13 {Hz}.
-\newblock {\em J. Neurosci.} {\bf 2006}, {\em 26},~502--507.
- {{https://doi.org/10.1523/jneurosci.4654-05.2006}}.
-
-\bibitem[Dugue {et~al.}(2011)Dugue, Marque, and VanRullen]{dugue_phase_2011}
+  \newblock {\em Journal of Neuroscience} {\bf 2006}, {\em 26},~502--507.
+  \newblock 00000, {\url{https://doi.org/10.1523/jneurosci.4654-05.2006}}.
+
+  \bibitem[Dugue \em{et~al.}(2011)Dugue, Marque, and VanRullen]{dugue_phase_2011}
   Dugue, L.; Marque, P.; VanRullen, R.
@@
     {Relation} between {Brain} {Excitation} and {Visual} {Perception}.
-\newblock {\em J. Neurosci.} {\bf 2011}, {\em 31},~11889--11893.
-\newblock {{https://doi.org/10.1523/JNEUROSCI.1161-11.2011}}.
-
-\bibitem[Bringuier {et~al.}(1999)Bringuier, Chavane, Glaeser, and
+  \newblock {\em Journal of Neuroscience} {\bf 2011}, {\em 31},~11889--11893.
+  \newblock {\url{https://doi.org/10.1523/JNEUROSCI.1161-11.2011}}.
+
+  \bibitem[Bringuier \em{et~al.}(1999)Bringuier, Chavane, Glaeser, and
     Frégnac]{bringuier_horizontal_1999}
@@
   \newblock {\em Science} {\bf 1999}, {\em 283},~695--699.
-%\newblock 00535 Loaded from an external bibliography file by Manubot.
-%  source\_bibliography: manual-references.bib standard\_id: Bringuier99,
-  {{https://doi.org/10.1126/science.283.5402.695}}.
-
-\bibitem[Benvenuti {et~al.}(2020)Benvenuti, Chemla, Boonman, Perrinet,
+  \newblock 00535 Loaded from an external bibliography file by Manubot.
+    source\_bibliography: manual-references.bib standard\_id: Bringuier99,
+    {\url{https://doi.org/10.1126/science.283.5402.695}}.
+
+  \bibitem[Benvenuti \em{et~al.}(2020)Benvenuti, Chemla, Boonman, Perrinet,
     Masson, and Chavane]{benvenuti_anticipatory_2020}
@@
     {V1}.
-\newblock {\em bioRxiv  Prepr. Serv. Biol.} {\bf 2020}.
-\newblock {{https://doi.org/10.1101/2020.03.26.010017}}.
-
-\bibitem[Le~Bec {et~al.}(2022)Le~Bec, Troncoso, Desbois, Passarelli, Baudot,
+  \newblock {\em bioRxiv : the preprint server for biology} {\bf 2020}, p.
+    2020.03.26.010017.
+  \newblock {\url{https://doi.org/10.1101/2020.03.26.010017}}.
+
+  \bibitem[Le~Bec \em{et~al.}(2022)Le~Bec, Troncoso, Desbois, Passarelli, Baudot,
     Monier, Pananceau, and Frégnac]{le_bec_horizontal_2022}
@@
     and motion integration.
-\newblock {\em PLoS ONE} {\bf 2022}, {\em 17},~e0268351.
-\newblock {{https://doi.org/10.1371/journal.pone.0268351}}.
-
-\bibitem[Feller {et~al.}(1997)Feller, Butts, Aaron, Rokhsar, and
+  \newblock {\em PLOS ONE} {\bf 2022}, {\em 17},~e0268351.
+  \newblock {\url{https://doi.org/10.1371/journal.pone.0268351}}.
+
+  \bibitem[Feller \em{et~al.}(1997)Feller, Butts, Aaron, Rokhsar, and
     Shatz]{feller_dynamic_1997}
@@
   \newblock {\em Neuron} {\bf 1997}, {\em 19},~293--306.
-\newblock {{https://doi.org/10.1016/S0896-6273(00)80940-X}}.
+  \newblock {\url{https://doi.org/10.1016/S0896-6273(00)80940-X}}.
 
@@
   \newblock A model of neocortex.
-\newblock {\em Netw. Comput. Neural Syst.} {\bf 1995}, {\em
+  \newblock {\em Network: Computation in Neural Systems} {\bf 1995}, {\em
     6},~179--224.
-%\newblock Publisher: Taylor \& Francis \_eprint:
-%  https://doi.org/10.1088/0954-898X\_6\_2\_004,
-  {{https://doi.org/10.1088/0954-898X\_6\_2\_004}}.
-
-\bibitem[Muller {et~al.}(2014)Muller, Reynaud, Chavane, and
+  \newblock Publisher: Taylor \& Francis \_eprint:
+    https://doi.org/10.1088/0954-898X\_6\_2\_004,
+    {\url{https://doi.org/10.1088/0954-898X_6_2_004}}.
+
+  \bibitem[Muller \em{et~al.}(2014)Muller, Reynaud, Chavane, and
     Destexhe]{muller_stimulus-evoked_2014}
@@
     monkey is a propagating wave.
-\newblock {\em Nat. Commun.} {\bf 2014}, {\em 5},~3675.
-%\newblock 00068 Loaded from an external bibliography file by Manubot.
-%  source\_bibliography: manual-references.bib standard\_id: Muller14,
-  {{https://doi.org/10.1038/ncomms4675}}.
-
-\bibitem[Muller {et~al.}(2018)Muller, Chavane, Reynolds, and
+  \newblock {\em Nature Communications} {\bf 2014}, {\em 5},~3675.
+  \newblock 00068 Loaded from an external bibliography file by Manubot.
+    source\_bibliography: manual-references.bib standard\_id: Muller14,
+    {\url{https://doi.org/10.1038/ncomms4675}}.
+
+  \bibitem[Muller \em{et~al.}(2018)Muller, Chavane, Reynolds, and
     Sejnowski]{muller_cortical_2018}
@@
   \newblock Cortical travelling waves: {Mechanisms} and computational principles.
-\newblock {\em Nat. Rev. Neurosci.} {\bf 2018}, \emph{19}, 255--268.
-%\newblock Loaded from an external bibliography file by Manubot.
-%  source\_bibliography: manual-references.bib standard\_id: Muller18,
-  {{https://doi.org/10.1038/nrn.2018.20}}.
-
-\bibitem[Lindén {et~al.}(2022)Lindén, Petersen, Vestergaard, and
+  \newblock {\em Nature Reviews Neuroscience} {\bf 2018}.
+  \newblock Loaded from an external bibliography file by Manubot.
+    source\_bibliography: manual-references.bib standard\_id: Muller18,
+    {\url{https://doi.org/10.1038/nrn.2018.20}}.
+
+  \bibitem[Lindén \em{et~al.}(2022)Lindén, Petersen, Vestergaard, and
     Berg]{linden_movement_2022}
@@
   \newblock {\em Nature} {\bf 2022}, {\em 610},~526--531.
-\newblock {{https://doi.org/10.1038/s41586-022-05293-w}}.
-
-\bibitem[Chemla {et~al.}(2019)Chemla, Reynaud, diVolo, Zerlaut, Perrinet,
+  \newblock {\url{https://doi.org/10.1038/s41586-022-05293-w}}.
+
+  \bibitem[Chemla \em{et~al.}(2019)Chemla, Reynaud, diVolo, Zerlaut, Perrinet,
     Destexhe, and Chavane]{chemla_suppressive_2019}
@@
     apparent motion in awake monkey {V1}.
-\newblock {\em J. Neurosci.} {\bf 2019}, {\em 2792},~18.
-%\newblock Loaded from an external bibliography file by Manubot.
-%  source\_bibliography: manual-references.bib standard\_id: Chemla19,
-  {{https://doi.org/10.1523/JNEUROSCI.2792-18.2019}}.
-
-\bibitem[Pillow {et~al.}(2008)Pillow, Shlens, Paninski, Sher, Litke,
+  \newblock {\em Journal of Neuroscience} {\bf 2019}, {\em 2792},~18.
+  \newblock Loaded from an external bibliography file by Manubot.
+    source\_bibliography: manual-references.bib standard\_id: Chemla19,
+    {\url{https://doi.org/10.1523/JNEUROSCI.2792-18.2019}}.
+
+  \bibitem[Pillow \em{et~al.}(2008)Pillow, Shlens, Paninski, Sher, Litke,
     Chichilnisky, and Simoncelli]{pillow_spatio-temporal_2008}
@@
   \newblock {\em Nature} {\bf 2008}, {\em 454},~995--999.
-\newblock {{https://doi.org/10.1038/nature07140}}.
-
-\bibitem[Schneidman {et~al.}(2006)Schneidman, Berry, Segev, and
+  \newblock {\url{https://doi.org/10.1038/nature07140}}.
+
+  \bibitem[Schneidman \em{et~al.}(2006)Schneidman, Berry, Segev, and
     Bialek]{schneidman_weak_2006}
@@
   \newblock {\em Nature} {\bf 2006}, {\em 440},~1007--1012.
-\newblock {{https://doi.org/10.1038/nature04701}}.
+  \newblock {\url{https://doi.org/10.1038/nature04701}}.
 
   \bibitem[Berry(2022)]{berry_spike_2022}
-Puchalla, J.; Berry, M.J.
-\newblock \emph{Spike {Trains} of {Retinal} {Ganglion} {Cells} {Viewing} a {Repeated}
-  {Natural} {Movie}}; {Princeton University: Princeton, NJ, USA,} %MDPI: We added the name of the publisher and location. Please confirm.
-  {2022}.
-\newblock {{https://doi.org/10.34770/V0V4-3H52}}.
-
-\bibitem[Miller {et~al.}(2014)Miller, Ayzenshtat, Carrillo-Reid, and
+  Berry, M.J.
+  \newblock Spike {Trains} of {Retinal} {Ganglion} {Cells} {Viewing} a {Repeated}
+    {Natural} {Movie} {\bf 2022}.
+  \newblock {\url{https://doi.org/10.34770/V0V4-3H52}}.
+
+  \bibitem[Miller \em{et~al.}(2014)Miller, Ayzenshtat, Carrillo-Reid, and
     Yuste]{miller_visual_2014}
@@
   \newblock Visual stimuli recruit intrinsically generated cortical ensembles.
-\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2014},
-  {\em 111}, E4053--E4061.
-\newblock {{https://doi.org/10.1073/pnas.1406077111}}.
-
-\bibitem[Ikegaya {et~al.}(2004)Ikegaya, Aaron, Cossart, Aronov, Lampl,
+  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2014},
+    {\em 111}.
+  \newblock {\url{https://doi.org/10.1073/pnas.1406077111}}.
+
+  \bibitem[Ikegaya \em{et~al.}(2004)Ikegaya, Aaron, Cossart, Aronov, Lampl,
     Ferster, and Yuste]{ikegaya_synfire_2004}
@@
   \newblock {\em Science} {\bf 2004}, {\em 304},~559--564.
-\newblock {{https://doi.org/10.1126/science.1093173}}.
-
-\bibitem[Luczak {et~al.}(2007)Luczak, Barthó, Marguet, Buzsáki, and
+  \newblock {\url{https://doi.org/10/djckcn}}.
+
+  \bibitem[Luczak \em{et~al.}(2007)Luczak, Barthó, Marguet, Buzsáki, and
     Harris]{luczak_sequential_2007}
@@
   \newblock Sequential structure of neocortical spontaneous activity in vivo.
-\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2007},
+  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2007},
     {\em 104},~347--352.
-\newblock {{https://doi.org/10.1073/pnas.0605643104}}.
-
-\bibitem[Pastalkova {et~al.}(2008)Pastalkova, Itskov, Amarasingham, and
+  \newblock {\url{https://doi.org/10.1073/pnas.0605643104}}.
+
+  \bibitem[Pastalkova \em{et~al.}(2008)Pastalkova, Itskov, Amarasingham, and
     Buzsáki]{pastalkova_internally_2008}
@@
     {Hippocampus}.
-\newblock {\em Science} {\bf 2008}, {\em 321},~1322--1327.
-\newblock {{https://doi.org/10.1126/science.1159775}}.
-
-\bibitem[Villette {et~al.}(2015)Villette, Malvache, Tressard, Dupuy, and
+  \newblock {\em Science (New York, N.Y.)} {\bf 2008}, {\em 321},~1322--1327.
+  \newblock {\url{https://doi.org/10.1126/science.1159775}}.
+
+  \bibitem[Villette \em{et~al.}(2015)Villette, Malvache, Tressard, Dupuy, and
     Cossart]{villette_internally_2015}
@@
   \newblock {\em Neuron} {\bf 2015}, {\em 88},~357--366.
- {{https://doi.org/10.1016/j.neuron.2015.09.052}}.
-
-\bibitem[Branco {et~al.}(2010)Branco, Clark, and
+  \newblock 00085, {\url{https://doi.org/10/f7whnn}}.
+
+  \bibitem[Branco \em{et~al.}(2010)Branco, Clark, and
     Häusser]{branco_dendritic_2010}
@@
   \newblock {\em Science} {\bf 2010}, {\em 329},~1671--1675.
-\newblock {{https://doi.org/10.1126/science.1189664}}.
-
-\bibitem[Luczak {et~al.}(2015)Luczak, McNaughton, and
+  \newblock {\url{https://doi.org/10.1126/science.1189664}}.
+
+  \bibitem[Luczak \em{et~al.}(2015)Luczak, McNaughton, and
     Harris]{luczak_packet-based_2015}
@@
   \newblock Packet-based communication in the cortex.
-\newblock {\em Nat. Rev. Neurosci.} {\bf 2015}, {\em 16},~745--755.
-\newblock {{https://doi.org/10.1038/nrn4026}}.
-
-\bibitem[Meister {et~al.}(1995)Meister, Lagnado, and
+  \newblock {\em Nature reviews. Neuroscience} {\bf 2015}, {\em 16},~745--55.
+  \newblock {\url{https://doi.org/10.1038/nrn4026}}.
+
+  \bibitem[Meister \em{et~al.}(1995)Meister, Lagnado, and
     Baylor]{meister_concerted_1995}
@@
   \newblock {\em Science} {\bf 1995}, {\em 270},~1207--1210.
-\newblock {{https://doi.org/10.1126/science.270.5239.1207}}.
+  \newblock {\url{https://doi.org/10.1126/science.270.5239.1207}}.
 
@@
   \newblock Construction of {Odor} {Representations} by {Olfactory} {Bulb}
-  {Microcircuits}. In {\em Progress in {Brain} {Research}}; Elsevier:  {Amsterdam, The Netherlands,} %newly added information, please confirm
-  2014;
-  Volume 208, pp. 177--203.
-\newblock {{https://doi.org/10.1016/B978-0-444-63350-7.00007-3}}.
-
-\bibitem[Kashiwadani {et~al.}(1999)Kashiwadani, Sasaki, Uchida, and
+    {Microcircuits}. In {\em Progress in {Brain} {Research}}; Elsevier,  2014;
+    Vol. 208, pp. 177--203.
+  \newblock {\url{https://doi.org/10.1016/B978-0-444-63350-7.00007-3}}.
+
+  \bibitem[Kashiwadani \em{et~al.}(1999)Kashiwadani, Sasaki, Uchida, and
     Mori]{kashiwadani_synchronized_1999}
@@
     {Olfactory} {Bulb}.
-\newblock {\em J. Neurophysiol.} {\bf 1999}, {\em 82},~1786--1792.\linebreak
-%\newblock Publisher: American Physiological Society,
-  {{https://doi.org/10.1152/jn.1999.82.4.1786}}.
-
-\bibitem[Rinberg {et~al.}(2006)Rinberg, Koulakov, and
+  \newblock {\em Journal of Neurophysiology} {\bf 1999}, {\em 82},~1786--1792.
+  \newblock Publisher: American Physiological Society,
+    {\url{https://doi.org/10.1152/jn.1999.82.4.1786}}.
+
+  \bibitem[Rinberg \em{et~al.}(2006)Rinberg, Koulakov, and
     Gelperin]{rinberg_speed-accuracy_2006}
@@
   \newblock Speed-{Accuracy} {Tradeoff} in {Olfaction}.
-\newblock {\em Neuron} {\bf 2006}, {\em 51},~351--358.\linebreak
-\newblock {{https://doi.org/10.1016/j.neuron.2006.07.013}}.
+  \newblock {\em Neuron} {\bf 2006}, {\em 51},~351--358.
+  \newblock {\url{https://doi.org/10.1016/j.neuron.2006.07.013}}.
 
@@
     spatial fingertip events.
-\newblock {\em Nat. Neurosci.} {\bf 2004}, {\em 7},~170--177.
-\newblock {{https://doi.org/10.1038/nn1177}}.
+  \newblock {\em Nature Neuroscience} {\bf 2004}, {\em 7},~170--177.
+  \newblock {\url{https://doi.org/10.1038/nn1177}}.
 
@@
   \newblock Space and {Time}: {The} {Hippocampus} as a {Sequence} {Generator}.
-\newblock {\em Trends Cogn. Sci.} {\bf 2018}, {\em 22},~853--869.
-\newblock {{https://doi.org/10.1016/j.tics.2018.07.006}}.
-
-\bibitem[Malvache {et~al.}(2016)Malvache, Reichinnek, Villette, Haimerl, and
+  \newblock {\em Trends in Cognitive Sciences} {\bf 2018}, {\em 22},~853--869.
+  \newblock {\url{https://doi.org/10.1016/j.tics.2018.07.006}}.
+
+  \bibitem[Malvache \em{et~al.}(2016)Malvache, Reichinnek, Villette, Haimerl, and
     Cossart]{malvache_awake_2016}
@@
     assemblies.
-\newblock {\em Science} {\bf 2016}, {\em 353},~1280--1283.
- {{https://doi.org/10.1126/science.aaf3319}}.
-
-\bibitem[Haimerl {et~al.}(2019)Haimerl, Angulo-Garcia, Villette, Reichinnek,
+  \newblock {\em Science (New York, N.Y.)} {\bf 2016}, {\em 353},~1280--1283.
+  \newblock 00105, {\url{https://doi.org/10/bqpq}}.
+
+  \bibitem[Haimerl \em{et~al.}(2019)Haimerl, Angulo-Garcia, Villette, Reichinnek,
     Torcini, Cossart, and Malvache]{haimerl_internal_2019}
@@
     time-distance continuum.
-\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2019},
+  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2019},
     {\em 116},~7477--7482.
-\newblock {{https://doi.org/10.1073/pnas.1718518116}}.
-
-\bibitem[Agus {et~al.}(2010)Agus, Thorpe, and Pressnitzer]{agus_rapid_2010}
+  \newblock {\url{https://doi.org/10.1073/pnas.1718518116}}.
+
+  \bibitem[Agus \em{et~al.}(2010)Agus, Thorpe, and Pressnitzer]{agus_rapid_2010}
   Agus, T.R.; Thorpe, S.J.; Pressnitzer, D.
@@
   \newblock {\em Neuron} {\bf 2010}, {\em 66},~610--618.
-\newblock {{https://doi.org/10.1016/j.neuron.2010.04.014}}.
+  \newblock {\url{https://doi.org/10.1016/j.neuron.2010.04.014}}.
 
@@
   \newblock Polychronization: {Computation} with {Spikes}.
-\newblock {\em Neural Comput.} {\bf 2006}, {\em 18},~245--282.\linebreak
- {{https://doi.org/10.1162/089976606775093882}}.
-
-\bibitem[Simoncelli {et~al.}(2003)Simoncelli, Paninski, Pillow, and
+  \newblock {\em Neural Computation} {\bf 2006}, {\em 18},~245--282.
+  \newblock 00000, {\url{https://doi.org/10/bgh4qv}}.
+
+  \bibitem[Simoncelli \em{et~al.}(2003)Simoncelli, Paninski, Pillow, and
     Schwartz]{simoncelli_characterization_2003}
   Simoncelli, E.P.; Paninski, L.; Pillow, J.; Schwartz, O.
-\newblock Characterization of {Neural} {Responses} with {Stochastic} {Stimuli}.
-In \emph{{The New Cognitive Neurosciences}}, {3rd ed.; Gazzaniga, M., Ed.; MIT Press: Cambridge, MA, USA,} %MDPI: Newly added  information.
-Please confirm.
-  {2003}.
+  \newblock Characterization of {Neural} {Responses} with {Stochastic} {Stimuli}
+    {\bf 2003}.
 
@@
   \newblock Optimal representation of sensory information by neural populations.
-\newblock {\em Nat. Neurosci.} {\bf 2006}, {\em 9},~690--696.
-%\newblock Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/nn1691}}.
-
-\bibitem[Berens {et~al.}(2012)Berens, Ecker, Cotton, Ma, Bethge, and
+  \newblock {\em Nature neuroscience} {\bf 2006}, {\em 9},~690--696.
+  \newblock Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/nn1691}}.
+
+  \bibitem[Berens \em{et~al.}(2012)Berens, Ecker, Cotton, Ma, Bethge, and
     Tolias]{berens_fast_2012}
@@
     {Primate} {V1}.
-\newblock {\em J. Neurosci.} {\bf 2012}, {\em 32},~10618--10626.
-\newblock {{https://doi.org/10.1523/jneurosci.1335-12.2012}}.
-
-\bibitem[Bellec {et~al.}(2021)Bellec, Wang, Modirshanechi, Brea, and
+  \newblock {\em Journal of Neuroscience} {\bf 2012}, {\em 32},~10618--10626.
+  \newblock {\url{https://doi.org/10.1523/jneurosci.1335-12.2012}}.
+
+  \bibitem[Bellec \em{et~al.}(2021)Bellec, Wang, Modirshanechi, Brea, and
     Gerstner]{bellec_fitting_2021}
@@
   \newblock Fitting summary statistics of neural data with a differentiable
-  spiking network simulator. \emph{arXiv} \textbf{2021},  	arXiv:2106.10064.
-%\newblock Technical Report 2106.10064, arXiv,  .
+    spiking network simulator.
+  \newblock Technical Report 2106.10064, arXiv,  2021.
 
@@
   Kohn, A.; Smith, M.
-\newblock Utah Array Extracellular Recordings of Spontaneous and Visually
-  Evoked Activity from Anesthetized Macaque Primary Visual Cortex ({V1}). 
+  \newblock Utah array extracellular recordings of spontaneous and visually
+    evoked activity from anesthetized macaque primary visual cortex ({V1}).,
     2016.
-Available online: {\url{https://doi.org/10.6080/K0NC5Z4X}}  {(accessed on 25 December 2022).} %MDPI: Please add the access date (format: Date Month Year), e.g., accessed on 1 January 2020.
-
-\bibitem[Warner {et~al.}(2022)Warner, Ruda, and
+  \newblock {\url{https://doi.org/10.6080/K0NC5Z4X}}.
+
+  \bibitem[Warner \em{et~al.}(2022)Warner, Ruda, and
     Sommer]{warner_probabilistic_2022}
@@
   \newblock A probabilistic latent variable model for detecting structure in
-  binary data. \emph{arXiv} \textbf{2022}, arXiv:2201.11108.
-%\newblock {\em arXiv:2201.11108 [cs, q-bio, stat]} {\bf }.
-%\newblock arXiv: 2201.11108.
+    binary data.
+  \newblock {\em arXiv:2201.11108 [cs, q-bio, stat]} {\bf 2022}.
+  \newblock arXiv: 2201.11108.
 
@@
   Victor, J.D.; Purpura, K.P.
-\newblock Nature and precision of temporal coding in visual cortex: A
+  \newblock Nature and precision of temporal coding in visual cortex: a
     metric-space analysis.
-\newblock {\em J. Neurophysiol.} {\bf 1996}, {\em 76},~1310--1326.
-\newblock {{https://doi.org/10.1152/jn.1996.76.2.1310}}.
+  \newblock {\em Journal of Neurophysiology} {\bf 1996}, {\em 76},~1310--1326.
+  \newblock {\url{https://doi.org/10.1152/jn.1996.76.2.1310}}.
 
@@
   \newblock A novel spike distance.
-\newblock {\em Neural Comput.} {\bf 2001}, {\em 13},~751--763.
-\newblock {{https://doi.org/10.1162/089976601300014321}}.
-
-\bibitem[Kreuz {et~al.}(2007)Kreuz, Haas, Morelli, Abarbanel, and
+  \newblock {\em Neural Computation} {\bf 2001}, {\em 13},~751--763.
+  \newblock {\url{https://doi.org/10.1162/089976601300014321}}.
+
+  \bibitem[Kreuz \em{et~al.}(2007)Kreuz, Haas, Morelli, Abarbanel, and
     Politi]{kreuz_measuring_2007}
@@
   \newblock Measuring spike train synchrony.
-\newblock {\em J. Neurosci. Methods} {\bf 2007}, {\em
+  \newblock {\em Journal of Neuroscience Methods} {\bf 2007}, {\em
     165},~151--161.
-\newblock {{https://doi.org/10.1016/j.jneumeth.2007.05.031}}.
+  \newblock {\url{https://doi.org/10.1016/j.jneumeth.2007.05.031}}.
 
@@
     {Induced} by {Level}-{Crossing} {Sampling}.
-\newblock {\em IEEE Trans. Signal Process.} {\bf 2014}, {\em
+  \newblock {\em IEEE Transactions on Signal Processing} {\bf 2014}, {\em
     62},~1987--1999.
-\newblock {{https://doi.org/10.1109/tsp.2014.2305642}}.
+  \newblock {\url{https://doi.org/10.1109/tsp.2014.2305642}}.
 
@@
   \newblock Ueber die {Gleichverteilung} von {Zahlen} mod. {Eins}.
-\newblock {\em Math. Ann.} {\bf 1916}, {\em 77},~313--352.
-\newblock {{https://doi.org/10.1007/bf01475864}}.
+  \newblock {\em Mathematische Annalen} {\bf 1916}, {\em 77},~313--352.
+  \newblock {\url{https://doi.org/10.1007/bf01475864}}.
 
@@
   \newblock Non-{Euclidean} properties of spike train metric spaces.
-\newblock {\em Phys. Rev. E} {\bf 2004}, {\em 69},~061905.
-%\newblock Publisher: American Physical Society,
-  {{https://doi.org/10.1103/PhysRevE.69.061905}}.
-
-\bibitem[Levakova {et~al.}(2015)Levakova, Tamborrino, Ditlevsen, and
+  \newblock {\em Physical Review E} {\bf 2004}, {\em 69},~061905.
+  \newblock Publisher: American Physical Society,
+    {\url{https://doi.org/10.1103/PhysRevE.69.061905}}.
+
+  \bibitem[Levakova \em{et~al.}(2015)Levakova, Tamborrino, Ditlevsen, and
     Lansky]{levakova_review_2015}
@@
   \newblock {\em Biosystems} {\bf 2015}, {\em 136},~23--34.
-\newblock {{https://doi.org/10.1016/j.biosystems.2015.04.008}}.
-
-\bibitem[Grün {et~al.}(2002)Grün, Diesmann, and
+  \newblock {\url{https://doi.org/10.1016/j.biosystems.2015.04.008}}.
+
+  \bibitem[Grün \em{et~al.}(2002)Grün, Diesmann, and
     Aertsen]{grun_unitary_2002-1}
@@
     {Activity}: {II}. {Nonstationary} {Data}.
-\newblock {\em Neural Comput.} {\bf 2002}, {\em 14},~81--119.
-\newblock {{https://doi.org/10.1162/089976602753284464}}.
-
-\bibitem[Grün {et~al.}(2010)Grün, Diesmann, and
+  \newblock {\em Neural Computation} {\bf 2002}, {\em 14},~81--119.
+  \newblock {\url{https://doi.org/10.1162/089976602753284464}}.
+
+  \bibitem[Grün \em{et~al.}(2010)Grün, Diesmann, and
     Aertsen]{grun_unitary_2010}
@@
   \newblock Unitary {Event} {Analysis}. In {\em Analysis of {Parallel} {Spike}
-  {Trains}}; Springer: {Boston, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
-  2010; pp. 191--220.
-
-\bibitem[Grün {et~al.}(2002)Grün, Diesmann, and
+    {Trains}}; Springer US,  2010; pp. 191--220.
+
+  \bibitem[Grün \em{et~al.}(2002)Grün, Diesmann, and
     Aertsen]{grun_unitary_2002}
@@
     {Activity}: {I}. {Detection} and {Significance}.
-\newblock {\em Neural Comput.} {\bf 2002}, {\em 14},~43--80.
-\newblock {{https://doi.org/10.1162/089976602753284455}}.
-
-\bibitem[Torre {et~al.}(2013)Torre, Picado-Muiño, Denker, Borgelt, and
+  \newblock {\em Neural Computation} {\bf 2002}, {\em 14},~43--80.
+  \newblock {\url{https://doi.org/10.1162/089976602753284455}}.
+
+  \bibitem[Torre \em{et~al.}(2013)Torre, Picado-Muiño, Denker, Borgelt, and
     Grün]{torre_statistical_2013}
@@
     frequent item set mining.
-\newblock {\em Front. Comput. Neurosci.} {\bf 2013}, {\em 7}, 132.
-\newblock {{https://doi.org/10.3389/fncom.2013.00132}}.
-
-\bibitem[Quaglio {et~al.}(2018)Quaglio, Rostami, Torre, and
+  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2013}, {\em 7}.
+  \newblock {\url{https://doi.org/10.3389/fncom.2013.00132}}.
+
+  \bibitem[Quaglio \em{et~al.}(2018)Quaglio, Rostami, Torre, and
     Grün]{quaglio_methods_2018}
@@
     spike trains.
-\newblock {\em Biol. Cybern.} {\bf 2018}, {\em 112},~57--80.
-\newblock {{https://doi.org/10.1007/s00422-018-0755-0}}.
-
-\bibitem[Stella {et~al.}(2019)Stella, Quaglio, Torre, and
+  \newblock {\em Biological Cybernetics} {\bf 2018}, {\em 112},~57--80.
+  \newblock {\url{https://doi.org/10.1007/s00422-018-0755-0}}.
+
+  \bibitem[Stella \em{et~al.}(2019)Stella, Quaglio, Torre, and
     Grün]{stella_3d-spade_2019}
@@
   \newblock {\em Biosystems} {\bf 2019}, {\em 185},~104022.
-\newblock {{https://doi.org/10.1016/j.biosystems.2019.104022}}.
-
-\bibitem[Stella {et~al.}(2022)Stella, Bouss, Palm, and
+  \newblock {\url{https://doi.org/10.1016/j.biosystems.2019.104022}}.
+
+  \bibitem[Stella \em{et~al.}(2022)Stella, Bouss, Palm, and
     Grün]{stella_comparing_2022}
@@
     {Higher}-{Order} {Spike} {Correlations}.
-\newblock {\em eNeuro} {\bf 2022}, {\em 9},~ENEURO.0505--21.2022.
-\newblock {{https://doi.org/10.1523/eneuro.0505-21.2022}}.
-
-\bibitem[Grossberger {et~al.}(2018)Grossberger, Battaglia, and
+  \newblock {\em eneuro} {\bf 2022}, {\em 9},~ENEURO.0505--21.2022.
+  \newblock {\url{https://doi.org/10.1523/eneuro.0505-21.2022}}.
+
+  \bibitem[Grossberger \em{et~al.}(2018)Grossberger, Battaglia, and
     Vinck]{grossberger_unsupervised_2018}
@@
     neuronal ensembles using a novel dissimilarity measure.
-\newblock {\em PLoS Comput. Biol.} {\bf 2018}, {\em 14},~e1006283.
-\newblock {{https://doi.org/10.1371/journal.pcbi.1006283}}.
-
-\bibitem[Nádasdy {et~al.}(1999)Nádasdy, Hirase, Czurkó, Csicsvari, and
+  \newblock {\em PLOS Computational Biology} {\bf 2018}, {\em 14},~e1006283.
+  \newblock {\url{https://doi.org/10/gdvbsx}}.
+
+  \bibitem[Nádasdy \em{et~al.}(1999)Nádasdy, Hirase, Czurkó, Csicsvari, and
     Buzsáki]{nadasdy_replay_1999}
@@
     the {Hippocampus}.
-\newblock {\em J. Neurosci.} {\bf 1999}, {\em 19},~9497--9507.
-%\newblock Publisher: Society for Neuroscience Section: ARTICLE,
-  {{https://doi.org/10.1523/JNEUROSCI.19-21-09497.1999}}.
+  \newblock {\em Journal of Neuroscience} {\bf 1999}, {\em 19},~9497--9507.
+  \newblock Publisher: Society for Neuroscience Section: ARTICLE,
+    {\url{https://doi.org/10.1523/JNEUROSCI.19-21-09497.1999}}.
 
@@
     {Relative} {Time} {Order}.
-\newblock {\em J. Neurophysiol.} {\bf 2004}, {\em 92},~2555--2573.
-\newblock {{https://doi.org/10.1152/jn.01030.2003}}.
-
-\bibitem[Sotomayor-Gómez {et~al.}(2021)Sotomayor-Gómez, Battaglia, and
+  \newblock {\em Journal of Neurophysiology} {\bf 2004}, {\em 92},~2555--2573.
+  \newblock {\url{https://doi.org/10.1152/jn.01030.2003}}.
+
+  \bibitem[Sotomayor-Gómez \em{et~al.}(2021)Sotomayor-Gómez, Battaglia, and
     Vinck]{sotomayor-gomez_spikeship_2021}
@@
     high-dimensional neural spiking patterns.
-\newblock {\em bioRxiv  Prepr. Serv. Biol.} {\bf 2021}.
-%\newblock Publisher: Cold Spring Harbor Laboratory,
-  {{https://doi.org/10.1101/2020.06.03.131573}}.
-
-\bibitem[Pachitariu {et~al.}(2018)Pachitariu, Stringer, and
+  \newblock {\em bioRxiv : the preprint server for biology} {\bf 2021}, pp.
+    2020--06.
+  \newblock Publisher: Cold Spring Harbor Laboratory,
+    {\url{https://doi.org/10.1101/2020.06.03.131573}}.
+
+  \bibitem[Pachitariu \em{et~al.}(2018)Pachitariu, Stringer, and
     Harris]{pachitariu_robustness_2018}
@@
     {Imaging}.
-\newblock {\em  J. Neurosci.} {\bf 2018}, {\em 38},~7976--7985.
-\newblock {{https://doi.org/10.1523/jneurosci.3339-17.2018}}.
+  \newblock {\em The Journal of Neuroscience} {\bf 2018}, {\em 38},~7976--7985.
+  \newblock {\url{https://doi.org/10.1523/jneurosci.3339-17.2018}}.
 
@@
   Stringer, C.
-\newblock {MouseLand}/Rastermap: {A} Multi-Dimensional Embedding Algorithm.
-  {2020}. %MDPI: Please add website and access date. Please check.
-  Available online: {\url{https://github.com/MouseLand/rastermap}}  {(accessed on 25 December 2022).} %MDPI: Please add the access date (format: Date Month Year), e.g., accessed on 1 January 2020.
-
-\bibitem[Stringer {et~al.}(2019)Stringer, Pachitariu, Steinmetz, Reddy,
+  \newblock {MouseLand}/rastermap: {A} multi-dimensional embedding algorithm,
+    2020.
+
+  \bibitem[Stringer \em{et~al.}(2019)Stringer, Pachitariu, Steinmetz, Reddy,
     Carandini, and Harris]{stringer_spontaneous_2019}
@@
   \newblock Spontaneous behaviors drive multidimensional, brainwide activity.
-\newblock {\em Science} {\bf 2019}, {\em 364}, 255.
-\newblock {{https://doi.org/10.1126/science.aav7893}}.
-
-\bibitem[Stringer {et~al.}(2021)Stringer, Michaelos, Tsyboulski, Lindo, and
+  \newblock {\em Science} {\bf 2019}, {\em 364}.
+  \newblock {\url{https://doi.org/10.1126/science.aav7893}}.
+
+  \bibitem[Stringer \em{et~al.}(2021)Stringer, Michaelos, Tsyboulski, Lindo, and
     Pachitariu]{stringer_high-precision_2021}
@@
   \newblock {\em Cell} {\bf 2021}, {\em 184},~2767--2778.e15.
-\newblock {{https://doi.org/10.1016/j.cell.2021.03.042}}.
+  \newblock {\url{https://doi.org/10.1016/j.cell.2021.03.042}}.
 
@@
     constellations.
-\newblock {\em eLife} {\bf 2017}, {\em 6}, e19428.
-\newblock {{https://doi.org/10.7554/elife.19428}}.
-
-\bibitem[Pipa {et~al.}(2008)Pipa, Wheeler, Singer, and
+  \newblock {\em eLife} {\bf 2017}, {\em 6}.
+  \newblock {\url{https://doi.org/10.7554/elife.19428}}.
+
+  \bibitem[Pipa \em{et~al.}(2008)Pipa, Wheeler, Singer, and
     Nikolić]{pipa_neuroxidence_2008}
   Pipa, G.; Wheeler, D.W.; Singer, W.; Nikolić, D.
-\newblock {NeuroXidence}: Reliable and efficient analysis of an excess or
+  \newblock {NeuroXidence}: reliable and efficient analysis of an excess or
     deficiency of joint-spike events.
-\newblock {\em J. Comput. Neurosci.} {\bf 2008}, {\em
+  \newblock {\em Journal of Computational Neuroscience} {\bf 2008}, {\em
     25},~64--88.
-\newblock {{https://doi.org/10.1007/s10827-007-0065-3}}.
-
-\bibitem[Torre {et~al.}(2016)Torre, Quaglio, Denker, Brochier, Riehle, and
+  \newblock {\url{https://doi.org/10.1007/s10827-007-0065-3}}.
+
+  \bibitem[Torre \em{et~al.}(2016)Torre, Quaglio, Denker, Brochier, Riehle, and
     Grun]{torre_synchronous_2016}
@@
     an {Instructed}-{Delay} {Reach}-to-{Grasp} {Task}.
-\newblock {\em J. Neurosci.} {\bf 2016}, {\em 36},~8329--8340.
-\newblock {{https://doi.org/10.1523/jneurosci.4375-15.2016}}.
-
-\bibitem[Williams {et~al.}(2020)Williams, Degleris, Wang, and
+  \newblock {\em Journal of Neuroscience} {\bf 2016}, {\em 36},~8329--8340.
+  \newblock {\url{https://doi.org/10.1523/jneurosci.4375-15.2016}}.
+
+  \bibitem[Williams \em{et~al.}(2020)Williams, Degleris, Wang, and
     Linderman]{williams_point_2020}
@@
   \newblock Point process models for sequence detection in high-dimensional
-  neural spike trains. \emph{arXiv}  \textbf{2020},  	arXiv:2010.04875
-%\newblock Technical Report 2010.04875, .
-
-\bibitem[Kass {et~al.}(2005)Kass, Ventura, and Brown]{kass_statistical_2005}
+    neural spike trains.
+  \newblock Technical Report 2010.04875, arXiv,  2020.
+
+  \bibitem[Kass \em{et~al.}(2005)Kass, Ventura, and Brown]{kass_statistical_2005}
   Kass, R.E.; Ventura, V.; Brown, E.N.
   \newblock Statistical issues in the analysis of neuronal data.
-\newblock {\em J. Neurophysiol.} {\bf 2005}, {\em 94},~8--25.
-%\newblock Publisher: American Physiological Society,
-  {{https://doi.org/10.1152/jn.00648.2004}}.
-
-\bibitem[van Kempen {et~al.}(2021)van Kempen, Gieselmann, Boyd, Steinmetz,
+  \newblock {\em Journal of neurophysiology} {\bf 2005}, {\em 94},~8--25.
+  \newblock Publisher: American Physiological Society,
+    {\url{https://doi.org/10.1152/jn.00648.2004}}.
+
+  \bibitem[van Kempen \em{et~al.}(2021)van Kempen, Gieselmann, Boyd, Steinmetz,
     Moore, Engel, and Thiele]{van_kempen_top-down_2021}
@@
   \newblock {\em Neuron} {\bf 2021}, {\em 109},~894--904.e8.
-\newblock {{https://doi.org/10.1016/j.neuron.2020.12.013}}.
-
-\bibitem[Pasturel {et~al.}(2020)Pasturel, Montagnini, and
+  \newblock {\url{https://doi.org/10.1016/j.neuron.2020.12.013}}.
+
+  \bibitem[Pasturel \em{et~al.}(2020)Pasturel, Montagnini, and
     Perrinet]{pasturel_humans_2020}
@@
     visual motion properties.
-\newblock {\em PLoS Comput. Biol.} {\bf 2020}, \emph{16}, e1007438.
-\newblock {{https://doi.org/10.1371/journal.pcbi.1007438}}.
+  \newblock {\em PLoS Computational Biology} {\bf 2020}.
+  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1007438}}.
 
@@
     {Nerven}.
-\newblock {\em Arch. Anat. Physiol. Wiss. Med.}
+  \newblock {\em Archiv für Anatomie, Physiologie und wissenschaftliche Medicin}
     {\bf 1850}, {\em 17},~176--364.
@@
   \newblock How is information transmitted in a nerve?
-\newblock {\em J. Biol. Phys.} {\bf 2020}, {\em 46},~327--341.
-\newblock {{https://doi.org/10.1007/s10867-020-09557-2}}.
+  \newblock {\em Journal of Biological Physics} {\bf 2020}, {\em 46},~327--341.
+  \newblock {\url{https://doi.org/10.1007/s10867-020-09557-2}}.
 
@@
   \newblock The {Functioning} of the {Giant} {Nerve} {Fibres} of the {Squid}.
-\newblock {\em J. Exp. Biol.} {\bf 1938}, {\em 15},~170--185.\linebreak
-\newblock {{https://doi.org/10.1242/jeb.15.2.170}}.
-
-\bibitem[Madadi~Asl {et~al.}(2018)Madadi~Asl, Valizadeh, and
+  \newblock {\em Journal of Experimental Biology} {\bf 1938}, {\em 15},~170--185.
+  \newblock {\url{https://doi.org/10.1242/jeb.15.2.170}}.
+
+  \bibitem[Madadi~Asl \em{et~al.}(2018)Madadi~Asl, Valizadeh, and
     Tass]{madadi_asl_dendritic_2018}
@@
     {Neuronal} {Networks} {With} {Plastic} {Synapses}.
-\newblock {\em Front. Physiol.} {\bf 2018}, {\em 9},~1849.
-\newblock {{https://doi.org/10.3389/fphys.2018.01849}}.
-
-\bibitem[Stetson {et~al.}(1992)Stetson, Albers, Silverstein, and
+  \newblock {\em Frontiers in Physiology} {\bf 2018}, {\em 9},~1849.
+  \newblock {\url{https://doi.org/10.3389/fphys.2018.01849}}.
+
+  \bibitem[Stetson \em{et~al.}(1992)Stetson, Albers, Silverstein, and
     Wolfe]{stetson_effects_1992}
@@
     measures.
-\newblock {\em Muscle Nerve} {\bf 1992}, {\em 15},~1095--1104.
-%\newblock \_eprint:
-%  https://onlinelibrary.wiley.com/doi/pdf/10.1002/mus.880151007,
-  {{https://doi.org/10.1002/mus.880151007}}.
+  \newblock {\em Muscle \& Nerve} {\bf 1992}, {\em 15},~1095--1104.
+  \newblock \_eprint:
+    https://onlinelibrary.wiley.com/doi/pdf/10.1002/mus.880151007,
+    {\url{https://doi.org/10.1002/mus.880151007}}.
 
@@
   \newblock A place theory of sound localization.
-\newblock {\em J. Comp. Physiol. Psychol.} {\bf 1948},
+  \newblock {\em Journal of Comparative and Physiological Psychology} {\bf 1948},
     {\em 41},~35--39.
-%\newblock Place: US Publisher: American Psychological Association,
-  {{https://doi.org/10.1037/h0061495}}.
+  \newblock Place: US Publisher: American Psychological Association,
+    {\url{https://doi.org/10.1037/h0061495}}.
 
@@
   \newblock Coding of auditory space.
-\newblock {\em Annu. Rev. Neurosci.} {\bf 2003}, {\em 26},~31--55.\linebreak
-\newblock {{https://doi.org/10.1146/annurev.neuro.26.041002.131123}}.
-
-\bibitem[Gerstner {et~al.}(1996)Gerstner, Kempter, van Hemmen, and
+  \newblock {\em Annual Review of Neuroscience} {\bf 2003}, {\em 26},~31--55.
+  \newblock {\url{https://doi.org/10.1146/annurev.neuro.26.041002.131123}}.
+
+  \bibitem[Gerstner \em{et~al.}(1996)Gerstner, Kempter, van Hemmen, and
     Wagner]{gerstner_neuronal_1996}
@@
   \newblock {\em Nature} {\bf 1996}, {\em 383},~76--78.
-%\newblock Number: 6595 Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/383076a0}}.
-
-\bibitem[Seidl {et~al.}(2010)Seidl, Rubel, and
+  \newblock Number: 6595 Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/383076a0}}.
+
+  \bibitem[Seidl \em{et~al.}(2010)Seidl, Rubel, and
     Harris]{seidl_mechanisms_2010}
@@
     binaural coincidence detection.
-\newblock {\em  J. Neurosci. Off. J. Soc. Neurosci.} {\bf 2010}, {\em 30},~70--80.
-\newblock {{https://doi.org/10.1523/JNEUROSCI.3464-09.2010}}.
-
-\bibitem[Camon {et~al.}(2019)Camon, Hugues, Erlandson, Robbe, Lagoun,
+  \newblock {\em The Journal of Neuroscience: The Official Journal of the Society
+    for Neuroscience} {\bf 2010}, {\em 30},~70--80.
+  \newblock {\url{https://doi.org/10.1523/JNEUROSCI.3464-09.2010}}.
+
+  \bibitem[Camon \em{et~al.}(2019)Camon, Hugues, Erlandson, Robbe, Lagoun,
     Marouane, and Bureau]{camon_timing_2019}
@@
     {Represented} in the {Mouse} {Primary} {Somatosensory} {Cortex}.
-\newblock {\em Cereb. Cortex} {\bf 2019}, {\em 29},~3034--3047.
-\newblock {{https://doi.org/10.1093/cercor/bhy169}}.
+  \newblock {\em Cerebral Cortex} {\bf 2019}, {\em 29},~3034--3047.
+  \newblock {\url{https://doi.org/10.1093/cercor/bhy169}}.
 
@@
     {AND} {THE} {CONDUCTION} {VELOCITY} {IN} {MAMMALIAN} {A} {FIBERS}.
-\newblock {\em Am. J.-Physiol.-Leg. Content} {\bf 1939}. %MDPI: Please add the volume and page number.
-%\newblock Publisher: American Physiological Society,
-  {{https://doi.org/10.1152/ajplegacy.1939.127.2.393}}.
-
-\bibitem[Brill {et~al.}(1977)Brill, Waxman, Moore, and
+  \newblock {\em American Journal of Physiology-Legacy Content} {\bf 1939}.
+  \newblock Publisher: American Physiological Society,
+    {\url{https://doi.org/10.1152/ajplegacy.1939.127.2.393}}.
+
+  \bibitem[Brill \em{et~al.}(1977)Brill, Waxman, Moore, and
     Joyner]{brill_conduction_1977}
@@
     computed dependence on internode distance.
-\newblock {\em J. Neurol. Neurosurg. Psychiatry} {\bf 1977},
+  \newblock {\em Journal of Neurology, Neurosurgery, and Psychiatry} {\bf 1977},
     {\em 40},~769--774.
 
-\bibitem[Pérez-Cerdá {et~al.}(2015)Pérez-Cerdá, Sánchez-Gómez, and
+  \bibitem[Pérez-Cerdá \em{et~al.}(2015)Pérez-Cerdá, Sánchez-Gómez, and
     Matute]{perez-cerda_pio_2015}
@@
   \newblock Pío del {Río} {Hortega} and the discovery of the oligodendrocytes.
-\newblock {\em Front. Neuroanat.} {\bf 2015}, {\em 9},~92.
-\newblock {{https://doi.org/10.3389/fnana.2015.00092}}.
+  \newblock {\em Frontiers in Neuroanatomy} {\bf 2015}, {\em 9},~92.
+  \newblock {\url{https://doi.org/10.3389/fnana.2015.00092}}.
 
@@
   \newblock The {Ultrastructure} of the {Nerve} {Axon} {Sheath}.
-\newblock {\em Biol. Rev.} {\bf 1939}, {\em 14},~27--50.
-%\newblock \_eprint:
-%  https://onlinelibrary.wiley.com/doi/pdf/10.1111/j.1469-185X.1939.tb00922.x,
-  {{https://doi.org/10.1111/j.1469-185X.1939.tb00922.x}}.
+  \newblock {\em Biological Reviews} {\bf 1939}, {\em 14},~27--50.
+  \newblock \_eprint:
+    https://onlinelibrary.wiley.com/doi/pdf/10.1111/j.1469-185X.1939.tb00922.x,
+    {\url{https://doi.org/10.1111/j.1469-185X.1939.tb00922.x}}.
 
@@
   \newblock Oligodendrocytes: {Myelination} and {Axonal} {Support}.
-\newblock {\em Cold Spring Harb. Perspect. Biol.} {\bf 2016}, {\em
+  \newblock {\em Cold Spring Harbor Perspectives in Biology} {\bf 2016}, {\em
     8},~a020479.
-\newblock {{https://doi.org/10.1101/cshperspect.a020479}}.
-
-\bibitem[Duncan {et~al.}(2021)Duncan, Simkins, and
+  \newblock {\url{https://doi.org/10.1101/cshperspect.a020479}}.
+
+  \bibitem[Duncan \em{et~al.}(2021)Duncan, Simkins, and
     Emery]{duncan_neuron-oligodendrocyte_2021}
@@
     {Integrity} of {Axons}.
-\newblock {\em Front. Cell Dev. Biol.} {\bf 2021}, {\em
-  9}, 653101.
-\newblock {{https://doi.org/10.3389/fcell.2021.653101}}.
+  \newblock {\em Frontiers in Cell and Developmental Biology} {\bf 2021}, {\em
+    9}.
+  \newblock {\url{https://doi.org/10.3389/fcell.2021.653101}}.
 
@@
     myelination.
-\newblock {\em Nat. Rev. Neurosci.} {\bf 2015}, {\em 16},~756--767.
-%\newblock Number: 12 Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/nrn4023}}.
+  \newblock {\em Nature Reviews Neuroscience} {\bf 2015}, {\em 16},~756--767.
+  \newblock Number: 12 Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/nrn4023}}.
 
@@
   \newblock Myelin makes memories.
-\newblock {\em Nat. Neurosci.} {\bf 2020}, {\em 23},~469--470.
-\newblock {{https://doi.org/10.1038/s41593-020-0606-x}}.
+  \newblock {\em Nature Neuroscience} {\bf 2020}, {\em 23},~469--470.
+  \newblock {\url{https://doi.org/10.1038/s41593-020-0606-x}}.
 
@@
     {Tissue} of the {Central} {Nervous} {System}.
-\newblock {\em Edinb. Med. J.} {\bf 1928}, {\em 35},~49--57.
-
-\bibitem[Steadman {et~al.}(2020)Steadman, Xia, Ahmed, Mocle, Penning,
+  \newblock {\em Edinburgh Medical Journal} {\bf 1928}, {\em 35},~49--57.
+
+  \bibitem[Steadman \em{et~al.}(2020)Steadman, Xia, Ahmed, Mocle, Penning,
     Geraghty, Steenland, Monje, Josselyn, and
@@
   \newblock {\em Neuron} {\bf 2020}, {\em 105},~150--164.e6.
-\newblock {{https://doi.org/10.1016/j.neuron.2019.10.013}}.
-
-\bibitem[Pan {et~al.}(2020)Pan, Mayoral, Choi, Chan, and
+  \newblock {\url{https://doi.org/10.1016/j.neuron.2019.10.013}}.
+
+  \bibitem[Pan \em{et~al.}(2020)Pan, Mayoral, Choi, Chan, and
     Kheirbek]{pan_preservation_2020}
@@
   \newblock Preservation of a remote fear memory requires new myelin formation.
-\newblock {\em Nat. Neurosci.} {\bf 2020}, {\em 23},~487--499.
-\newblock {{https://doi.org/10.1038/s41593-019-0582-1}}.
-
-\bibitem[Wan {et~al.}(2020)Wan, Cheli, Santiago-González, Rosenblum, Wan,
+  \newblock {\em Nature Neuroscience} {\bf 2020}, {\em 23},~487--499.
+  \newblock {\url{https://doi.org/10.1038/s41593-019-0582-1}}.
+
+  \bibitem[Wan \em{et~al.}(2020)Wan, Cheli, Santiago-González, Rosenblum, Wan,
     and Paez]{wan_impaired_2020}
@@
     {Mouse} for the {Ferritin} {Heavy} {Chain} in {Oligodendroglial} {Cells}.
-\newblock {\em  J. Neurosci.} {\bf 2020}, {\em 40},~7609--7624.
-\newblock {{https://doi.org/10.1523/JNEUROSCI.1281-20.2020}}.
-
-\bibitem[Xue {et~al.}(2021)Xue, Zhu, Liu, Lin, Li, Li, and
+  \newblock {\em The Journal of Neuroscience} {\bf 2020}, {\em 40},~7609--7624.
+  \newblock {\url{https://doi.org/10.1523/JNEUROSCI.1281-20.2020}}.
+
+  \bibitem[Xue \em{et~al.}(2021)Xue, Zhu, Liu, Lin, Li, Li, and
     Zhuo]{xue_demyelination_2021}
@@
     {Glaucoma}?
-\newblock {\em Front. Aging Neurosci.} {\bf 2021}, {\em 13},~701322.
-\newblock {{https://doi.org/10.3389/fnagi.2021.701322}}.
-
-\bibitem[Kuhn {et~al.}(2019)Kuhn, Gritti, Crooks, and
+  \newblock {\em Frontiers in Aging Neuroscience} {\bf 2021}, {\em 13},~701322.
+  \newblock {\url{https://doi.org/10.3389/fnagi.2021.701322}}.
+
+  \bibitem[Kuhn \em{et~al.}(2019)Kuhn, Gritti, Crooks, and
     Dombrowski]{kuhn_oligodendrocytes_2019}
@@
   \newblock {\em Cells} {\bf 2019}, {\em 8},~1424.
-\newblock {{https://doi.org/10.3390/cells8111424}}.
-
-\bibitem[Baraban {et~al.}(2018)Baraban, Koudelka, and
+  \newblock {\url{https://doi.org/10.3390/cells8111424}}.
+
+  \bibitem[Baraban \em{et~al.}(2018)Baraban, Koudelka, and
     Lyons]{baraban_ca2_2018}
   Baraban, M.; Koudelka, S.; Lyons, D.A.
-\newblock Ca$^{2+}$ activity signatures of myelin sheath formation and growth in
+  \newblock Ca2+ activity signatures of myelin sheath formation and growth in
     vivo.
-\newblock {\em Nat. Neurosci.} {\bf 2018}, {\em 21},~19--23.
-\newblock {{https://doi.org/10.1038/s41593-017-0040-x}}.
+  \newblock {\em Nature neuroscience} {\bf 2018}, {\em 21},~19--23.
+  \newblock {\url{https://doi.org/10.1038/s41593-017-0040-x}}.
 
@@
   \newblock Axonal regulation of myelination by neuregulin 1.
-\newblock {\em Curr. Opin. Neurobiol.} {\bf 2006}, {\em 16},~492--500.
-\newblock {{https://doi.org/10.1016/j.conb.2006.08.008}}.
-
-\bibitem[Cullen {et~al.}(2021)Cullen, Pepper, Clutterbuck, Pitman, Oorschot,
+  \newblock {\em Current Opinion in Neurobiology} {\bf 2006}, {\em 16},~492--500.
+  \newblock {\url{https://doi.org/10.1016/j.conb.2006.08.008}}.
+
+  \bibitem[Cullen \em{et~al.}(2021)Cullen, Pepper, Clutterbuck, Pitman, Oorschot,
     Auderset, Tang, Ramm, Emery, Rodger, Jolivet, and
@@
     conduction in the adult mouse brain.
-\newblock {\em Cell Rep.} {\bf 2021}, {\em 34}, 108641.
-%\newblock Publisher: Elsevier,
-  {{https://doi.org/10.1016/j.celrep.2020.108641}}.
-
-\bibitem[Gibson {et~al.}(2014)Gibson, Purger, Mount, Goldstein, Lin, Wood,
+  \newblock {\em Cell Reports} {\bf 2021}, {\em 34}.
+  \newblock Publisher: Elsevier,
+    {\url{https://doi.org/10.1016/j.celrep.2020.108641}}.
+
+  \bibitem[Gibson \em{et~al.}(2014)Gibson, Purger, Mount, Goldstein, Lin, Wood,
     Inema, Miller, Bieri, Zuchero, Barres, Woo, Vogel, and
@@
     {Myelination} in the {Mammalian} {Brain}.
-\newblock {\em Science} {\bf 2014}, {\em 344},~1252304.
-\newblock {{https://doi.org/10.1126/science.1252304}}.
-
-\bibitem[Spencer {et~al.}(2018)Spencer, Meffin, Burkitt, and
+  \newblock {\em Science (New York, N.Y.)} {\bf 2014}, {\em 344},~1252304.
+  \newblock {\url{https://doi.org/10.1126/science.1252304}}.
+
+  \bibitem[Spencer \em{et~al.}(2018)Spencer, Meffin, Burkitt, and
     Grayden]{spencer_compensation_2018}
@@
     {Model} of the {Auditory} {Brainstem}.
-\newblock {\em Front. Comput. Neurosci.} {\bf 2018}, {\em
+  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2018}, {\em
     12},~36.
-\newblock {{https://doi.org/10.3389/fncom.2018.00036}}.
-
-\bibitem[Mel {et~al.}(2017)Mel, Schiller, and Poirazi]{mel_synaptic_2017}
+  \newblock {\url{https://doi.org/10.3389/fncom.2018.00036}}.
+
+  \bibitem[Mel \em{et~al.}(2017)Mel, Schiller, and Poirazi]{mel_synaptic_2017}
   Mel, B.W.; Schiller, J.; Poirazi, P.
@@
     strategies.
-\newblock {\em Curr. Opin. Neurobiol.} {\bf 2017}, {\em 43},~177--186.
-\newblock {{https://doi.org/10.1016/j.conb.2017.03.012}}.
-
-\bibitem[Golding {et~al.}(2002)Golding, Staff, and
+  \newblock {\em Current Opinion in Neurobiology} {\bf 2017}, {\em 43},~177--186.
+  \newblock {\url{https://doi.org/10.1016/j.conb.2017.03.012}}.
+
+  \bibitem[Golding \em{et~al.}(2002)Golding, Staff, and
     Spruston]{golding_dendritic_2002}
@@
   \newblock {\em Nature} {\bf 2002}, {\em 418},~326--331.
-%\newblock Number: 6895 Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/nature00854}}.
+  \newblock Number: 6895 Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/nature00854}}.
 
@@
     models.
-\newblock {\em Neural Netw.} {\bf 1997}, {\em 10},~1659--1671.
-
-\bibitem[Neftci {et~al.}(2019)Neftci, Mostafa, and
+  \newblock {\em Neural Networks} {\bf 1997}, {\em 10},~1659--1671.
+
+  \bibitem[Neftci \em{et~al.}(2019)Neftci, Mostafa, and
     Zenke]{neftci_surrogate_2019}
@@
     {Neural} {Networks}.
-\newblock {\em IEEE Signal Process. Mag.} {\bf 2019}, {\em 36},~51--63.\linebreak
-\newblock {{https://doi.org/10.1109/MSP.2019.2931595}}.
-
-\bibitem[Rueckauer {et~al.}(2017)Rueckauer, Lungu, Hu, Pfeiffer, and
+  \newblock {\em IEEE Signal Processing Magazine} {\bf 2019}, {\em 36},~51--63.
+  \newblock {\url{https://doi.org/10.1109/MSP.2019.2931595}}.
+
+  \bibitem[Rueckauer \em{et~al.}(2017)Rueckauer, Lungu, Hu, Pfeiffer, and
     Liu]{rueckauer_conversion_2017}
@@
     {Event}-{Driven} {Networks} for {Image} {Classification}.
-\newblock {\em Front. Neurosci.} {\bf 2017}, {\em 11}.
-\newblock {{https://doi.org/10.3389/fnins.2017.00682}}.
-
-\bibitem[Susi {et~al.}(2021)Susi, Antón-Toro, Maestú, Pereda, and
+  \newblock {\em Frontiers in Neuroscience} {\bf 2017}, {\em 11}.
+  \newblock {\url{https://doi.org/10.3389/fnins.2017.00682}}.
+
+  \bibitem[Susi \em{et~al.}(2021)Susi, Antón-Toro, Maestú, Pereda, and
     Mirasso]{susi_nmnsd-spiking_2021}
@@
     {Weight}-{Adjustment} and {Delay}-{Shift}.
-\newblock {\em Front. Neurosci.} {\bf 2021}, {\em 15},~582608.
-\newblock {{https://doi.org/10.3389/fnins.2021.582608}}.
-
-\bibitem[Davies {et~al.}(2018)Davies, Srinivasa, Lin, Chinya, Cao, Choday,
+  \newblock {\em Frontiers in Neuroscience} {\bf 2021}, {\em 15},~582608.
+  \newblock {\url{https://doi.org/10.3389/fnins.2021.582608}}.
+
+  \bibitem[Davies \em{et~al.}(2018)Davies, Srinivasa, Lin, Chinya, Cao, Choday,
     Dimou, Joshi, Imam, Jain, Liao, Lin, Lines, Liu, Mathaikutty, McCoy, Paul,
@@
   \newblock {\em IEEE Micro} {\bf 2018}, {\em 38},~82--99.
-\newblock {{https://doi.org/10.1109/MM.2018.112130359}}.
+  \newblock {\url{https://doi.org/10.1109/MM.2018.112130359}}.
 
@@
     period.
-\newblock {\em Neurocomputing} {\bf 2004}, {\em 58--60},~53--58.
-\newblock {{https://doi.org/10.1016/j.neucom.2004.01.022}}.
-
-\bibitem[Markram {et~al.}(1997)Markram, Lübke, Frotscher, and
+  \newblock {\em Neurocomputing} {\bf 2004}, {\em 58-60},~53--58.
+  \newblock {\url{https://doi.org/10.1016/j.neucom.2004.01.022}}.
+
+  \bibitem[Markram \em{et~al.}(1997)Markram, Lübke, Frotscher, and
     Sakmann]{markram_regulation_1997}
@@
   \newblock {\em Science} {\bf 1997}, {\em 275},~213--215.
-\newblock {{https://doi.org/10.1126/science.275.5297.213}}.
+  \newblock {\url{https://doi.org/10.1126/science.275.5297.213}}.
 
@@
     {Rule}.
-\newblock {\em Annu. Rev. Neurosci.} {\bf 2008}, {\em 31},~25--46.
-\newblock {{https://doi.org/10.1146/annurev.neuro.31.060407.125639}}.
-
-\bibitem[Hüning {et~al.}(1998)Hüning, Glünder, and
+  \newblock {\em Annual Review of Neuroscience} {\bf 2008}, {\em 31},~25--46.
+  \newblock {\url{https://doi.org/10.1146/annurev.neuro.31.060407.125639}}.
+
+  \bibitem[Hüning \em{et~al.}(1998)Hüning, Glünder, and
     Palm]{huning_synaptic_1998}
@@
   \newblock Synaptic {Delay} {Learning} in {Pulse}-{Coupled} {Neurons}.
-\newblock {\em Neural Comput.} {\bf 1998}, {\em 10},~555--565.
-\newblock {{https://doi.org/10.1162/089976698300017665}}.
-
-\bibitem[Eurich {et~al.}(1999)Eurich, Pawelzik, Ernst, Cowan, and
+  \newblock {\em Neural Computation} {\bf 1998}, {\em 10},~555--565.
+  \newblock {\url{https://doi.org/10/cthfps}}.
+
+  \bibitem[Eurich \em{et~al.}(1999)Eurich, Pawelzik, Ernst, Cowan, and
     Milton]{eurich_dynamics_1999}
@@
   \newblock Dynamics of {Self}-{Organized} {Delay} {Adaptation}.
-\newblock {\em Phys. Rev. Lett.} {\bf 1999}, {\em 82},~1594--1597.
-%\newblock 00082 Publisher: American Physical Society,
-  {{https://doi.org/10.1103/PhysRevLett.82.1594}}.
+  \newblock {\em Physical Review Letters} {\bf 1999}, {\em 82},~1594--1597.
+  \newblock 00082 Publisher: American Physical Society,
+    {\url{https://doi.org/10/dj9c7q}}.
 
@@
   Gütig, R.; Sompolinsky, H.
-\newblock The tempotron: A neuron that learns spike timing–based decisions.
-\newblock {\em Nat. Neurosci.} {\bf 2006}, {\em 9},~420--428.
-\newblock {{https://doi.org/10.1038/nn1643}}.
+  \newblock The tempotron: a neuron that learns spike timing–based decisions.
+  \newblock {\em Nature Neuroscience} {\bf 2006}, {\em 9},~420--428.
+  \newblock {\url{https://doi.org/10.1038/nn1643}}.
 
@@
   \newblock To spike, or when to spike?
-\newblock {\em Curr. Opin. Neurobiol.} {\bf 2014}, {\em 25},~134--139.
-\newblock {{https://doi.org/10.1016/j.conb.2014.01.004}}.
-
-\bibitem[Pauli {et~al.}(2018)Pauli, Weidel, Kunkel, and
+  \newblock {\em Current Opinion in Neurobiology} {\bf 2014}, {\em 25},~134--139.
+  \newblock {\url{https://doi.org/10.1016/j.conb.2014.01.004}}.
+
+  \bibitem[Pauli \em{et~al.}(2018)Pauli, Weidel, Kunkel, and
     Morrison]{pauli_reproducing_2018}
@@
     {Reproducibility} of {Spiking} {Network} {Models}.
-\newblock {\em Front. Neuroinformatics} {\bf 2018}, {\em 12},~46.
- {{https://doi.org/10.3389/fninf.2018.00046}}.
-
-\bibitem[Guise {et~al.}(2014)Guise, Knott, and
+  \newblock {\em Frontiers in Neuroinformatics} {\bf 2018}, {\em 12},~46.
+  \newblock 00000, {\url{https://doi.org/10/gd8zj5}}.
+
+  \bibitem[Guise \em{et~al.}(2014)Guise, Knott, and
     Benuskova]{guise_bayesian_2014}
@@
   \newblock A {Bayesian} {Model} of {Polychronicity}.
-\newblock {\em Neural Comput.} {\bf 2014}, {\em 26},~2052--2073.
-\newblock {{https://doi.org/10.1162/neco\_a\_00620}}.
-
-\bibitem[Zhang {et~al.}(2020)Zhang, Wu, Belatreche, Pan, Xie, Chua, Li, Qu,
+  \newblock {\em Neural Computation} {\bf 2014}, {\em 26},~2052--2073.
+  \newblock {\url{https://doi.org/10.1162/neco_a_00620}}.
+
+  \bibitem[Zhang \em{et~al.}(2020)Zhang, Wu, Belatreche, Pan, Xie, Chua, Li, Qu,
     and Li]{zhang_supervised_2020}
@@
   \newblock {\em Neurocomputing} {\bf 2020}, {\em 409},~103--118.
-\newblock {{https://doi.org/10.1016/j.neucom.2020.03.079}}.
-
-\bibitem[Ghosh {et~al.}(2021)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
+  \newblock {\url{https://doi.org/10.1016/j.neucom.2020.03.079}}.
+
+  \bibitem[Ghosh \em{et~al.}(2021)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
     Alfaro-Bittner, and Boccaletti]{ghosh_synchronization_2021}
@@
     Boccaletti, S.
-\newblock Synchronization in time-varying networks. \emph{arXiv} {\bf 2021}, arXiv:2109.07618
-%\newblock {\em arXiv:2109.07618 [physics]} .
-%\newblock 00000 arXiv: 2109.07618.
-
-\bibitem[Ghosh {et~al.}(2022)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
+  \newblock Synchronization in time-varying networks.
+  \newblock {\em arXiv:2109.07618 [physics]} {\bf 2021}.
+  \newblock 00000 arXiv: 2109.07618.
+
+  \bibitem[Ghosh \em{et~al.}(2022)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
     Alfaro-Bittner, and Boccaletti]{ghosh_synchronized_2022}
@@
   \newblock The synchronized dynamics of time-varying networks.
-\newblock {\em Phys. Rep.} {\bf 2022}, {\em 949},~1--63.
-%\newblock arXiv:2109.07618 [physics],
-  {{https://doi.org/10.1016/j.physrep.2021.10.006}}.
+  \newblock {\em Physics Reports} {\bf 2022}, {\em 949},~1--63.
+  \newblock arXiv:2109.07618 [physics],
+    {\url{https://doi.org/10.1016/j.physrep.2021.10.006}}.
 
@@
   \newblock Polychronous {Wavefront} {Computations}.
-\newblock {\em Int. J. Bifurc. Chaos} {\bf 2009}, {\em
+  \newblock {\em International Journal of Bifurcation and Chaos} {\bf 2009}, {\em
     19},~1733--1739.
-\newblock {{https://doi.org/10.1142/S0218127409023809}}.
+  \newblock {\url{https://doi.org/10/db98d7}}.
 
@@
   \newblock In Proceedings of the 2022 {IEEE} {International} {Conference} on
-  {Image} {Processing} ({ICIP}), {Bordeaux, France, 16--19 October} %MDPI: We added the location and date of the conference. Please confirm.
- 2022; pp.~3591--3595.
-%\newblock ISSN: 2381-8549,
-  {{https://doi.org/10.1109/ICIP46576.2022.9897394}}.
+    {Image} {Processing} ({ICIP}),  2022, pp. 3591--3595.
+  \newblock ISSN: 2381-8549,
+    {\url{https://doi.org/10.1109/ICIP46576.2022.9897394}}.
 
@@
     symmetry between neurons shaped by spike-timing-dependent plasticity.
-\newblock {\em Cogn. Neurodynamics} {\bf {2022} %MDPI: Please add the volume. There is none for this citation
-}, 1--14.
-\newblock {{https://doi.org/10.1007/s11571-022-09850-x}}.
+  \newblock {\em Cognitive Neurodynamics} {\bf 2022}.
+  \newblock {\url{https://doi.org/10.1007/s11571-022-09850-x}}.
 
@@
   \newblock {\em Neurocomputing} {\bf 2002}, {\em 44-46},~133--139.
-\newblock {{https://doi.org/10.1016/s0925-2312(02)00374-0}}.
-
-\bibitem[Perrinet {et~al.}(2001)Perrinet, Delorme, Samuelides, and
+  \newblock {\url{https://doi.org/10.1016/s0925-2312(02)00374-0}}.
+
+  \bibitem[Perrinet \em{et~al.}(2001)Perrinet, Delorme, Samuelides, and
     Thorpe]{perrinet_networks_2001}
@@
     {How} to implement spike time dependent {Hebbian} plasticity.
-\newblock {\em Neurocomputing} {\bf 2001}, {\em 38--40},~817--822.
-\newblock {{https://doi.org/10.1016/s0925-2312(01)00460-x}}.
+  \newblock {\em Neurocomputing} {\bf 2001}, {\em 38-40},~817--822.
+  \newblock {\url{https://doi.org/10.1016/s0925-2312(01)00460-x}}.
 
@@
   \newblock {STDP} in recurrent neuronal networks.
-\newblock {\em Front. Comput. Neurosci.} {\bf 2010}, {\em 4}, 23.
-\newblock {{https://doi.org/10.3389/fncom.2010.00023}}.
-
-\bibitem[Datadien {et~al.}(2011)Datadien, Haselager, and
+  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2010}, {\em 4}.
+  \newblock {\url{https://doi.org/10.3389/fncom.2010.00023}}.
+
+  \bibitem[Datadien \em{et~al.}(2011)Datadien, Haselager, and
     Sprinkhuizen-Kuyper]{datadien_right_2011}
   Datadien, A.; Haselager, P.; Sprinkhuizen-Kuyper, I.
-\newblock {\em The {Right} {Delay}---{Detecting} {Specific} {Spike} {Patterns}
-  with {STDP} and {Axonal} {Conduction} {Delays}}; {Springer:  Berlin/Heidelberg, Germany,} %MDPI: We added the name of the publisher and location. Please confirm.
- 2011;
-\newblock pp. 90--99.
-
-\bibitem[Kerr {et~al.}(2013)Kerr, Burkitt, Thomas, Gilson, and
+  \newblock {\em The {Right} {Delay} - {Detecting} {Specific} {Spike} {Patterns}
+    with {STDP} and {Axonal} {Conduction} {Delays}.};  2011.
+  \newblock Pages: 99.
+
+  \bibitem[Kerr \em{et~al.}(2013)Kerr, Burkitt, Thomas, Gilson, and
     Grayden]{kerr_delay_2013}
@@
     {Inputs}.
-\newblock {\em PLoS Comput. Biol.} {\bf 2013}, {\em 9},~e1002897.
-\newblock {{https://doi.org/10.1371/journal.pcbi.1002897}}.
+  \newblock {\em PLoS Computational Biology} {\bf 2013}, {\em 9},~e1002897.
+  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1002897}}.
 
@@
     {Plasticity}.
-\newblock {\em  J. Neurosci.} {\bf 2021}, {\em 41},~4428--4438.
-\newblock {{https://doi.org/10.1523/jneurosci.2017-20.2021}}.
+  \newblock {\em The Journal of Neuroscience} {\bf 2021}, {\em 41},~4428--4438.
+  \newblock {\url{https://doi.org/10.1523/jneurosci.2017-20.2021}}.
 
@@
   \newblock Bio-plausible {Unsupervised} {Delay} {Learning} for {Extracting}
-  {Temporal} {Features} in {Spiking} {Neural} {Networks}. \emph{arXiv}  \textbf{2020},  	arXiv:2011.09380.
-
-\bibitem[Wang {et~al.}(2019)Wang, Lin, and Dang]{wang_delay_2019}
+    {Temporal} {Features} in {Spiking} {Neural} {Networks},  2020.
+
+  \bibitem[Wang \em{et~al.}(2019)Wang, Lin, and Dang]{wang_delay_2019}
   Wang, X.; Lin, X.; Dang, X.
@@
     for {Spiking} {Neurons}.
-\newblock {\em Front. Neurosci.} {\bf 2019}, {\em 13}, 252.
-
-\bibitem[Hazan {et~al.}(2022)Hazan, Caby, Earl, Siegelmann, and
+  \newblock {\em Frontiers in Neuroscience} {\bf 2019}, {\em 13}.
+
+  \bibitem[Hazan \em{et~al.}(2022)Hazan, Caby, Earl, Siegelmann, and
     Levin]{hazan_memory_2022}
@@
   \newblock Memory via {Temporal} {Delays} in weightless {Spiking} {Neural}
-  {Network}.  \emph{arXiv} \textbf{2022}, arXiv:2202.07132.
-%\newblock arXiv:2202.07132 [cs, q-bio, stat].
-
-\bibitem[Luo {et~al.}(2022)Luo, Qu, Wang, Yi, Zhang, and
+    {Network},  2022.
+  \newblock arXiv:2202.07132 [cs, q-bio, stat].
+
+  \bibitem[Luo \em{et~al.}(2022)Luo, Qu, Wang, Yi, Zhang, and
     Zhang]{luo_supervised_2022}
@@
     {With} {Spike} {Temporal} {Error} {Backpropagation}.
-\newblock {\em IEEE Trans. Neural Netw. Learn. Syst.} {\bf
-  {2022} %MDPI: Please add the volume. No volume for this citation
-}, 1--13.
-\newblock {{https://doi.org/10.1109/tnnls.2022.3164930}}.
-
-\bibitem[Sun {et~al.}(2016)Sun, Sourina, and Huang]{sun_learning_2016}
+  \newblock {\em IEEE Transactions on Neural Networks and Learning Systems} {\bf
+    2022}, pp. 1--13.
+  \newblock {\url{https://doi.org/10.1109/tnnls.2022.3164930}}.
+
+  \bibitem[Sun \em{et~al.}(2016)Sun, Sourina, and Huang]{sun_learning_2016}
   Sun, H.; Sourina, O.; Huang, G.B.
@@
     spike-timing-dependent plasticity.
-\newblock {\em Neural Comput.} {\bf 2016}, {\em 28},~2181--2212.
-%\newblock tex.eprint:
-%  https://direct.mit.edu/neco/article-pdf/28/10/2181/972099/neco{\textbackslash}\_a{\textbackslash}\_00879.pdf,
-  {{https://doi.org/10.1162/NECO\_a\_00879}}.
-
-\bibitem[Ghosh {et~al.}(2019)Ghosh, Gupta, Silva, Soares, and
+  \newblock {\em Neural Computation} {\bf 2016}, {\em 28},~2181--2212.
+  \newblock tex.eprint:
+    https://direct.mit.edu/neco/article-pdf/28/10/2181/972099/neco{\textbackslash}\_a{\textbackslash}\_00879.pdf,
+    {\url{https://doi.org/10.1162/NECO_a_00879}}.
+
+  \bibitem[Ghosh \em{et~al.}(2019)Ghosh, Gupta, Silva, Soares, and
     Thakor]{ghosh_spatiotemporal_2019}
   Ghosh, R.; Gupta, A.; Silva, A.N.; Soares, A.; Thakor, N.V.
-\newblock Spatiotemporal filtering for event-based action recognition. \emph{arXiv}  \textbf{2019},  	arXiv:1903.07067.
-
-\bibitem[Perrinet {et~al.}(2014)Perrinet, Adams, and
+  \newblock Spatiotemporal filtering for event-based action recognition,  2019.
+
+  \bibitem[Perrinet \em{et~al.}(2014)Perrinet, Adams, and
     Friston]{perrinet_active_2014}
@@
   \newblock Active inference, eye movements and oculomotor delays.
-\newblock {\em Biol. Cybern.} {\bf 2014}, {\em 108},~777--801.
-\newblock {{https://doi.org/10.1007/s00422-014-0620-8}}.
+  \newblock {\em Biological Cybernetics} {\bf 2014}, {\em 108},~777--801.
+  \newblock {\url{https://doi.org/10.1007/s00422-014-0620-8}}.
 
@@
     {Real}-{Time} {Temporal} {Alignment} {Hypothesis}.
-\newblock {\em eNeuro} {\bf 2019}, {\em 6},~ENEURO.0412--18.2019.
-\newblock {{https://doi.org/10.1523/eneuro.0412-18.2019}}.
-
-\bibitem[Khoei {et~al.}(2013)Khoei, Masson, and
+  \newblock {\em eneuro} {\bf 2019}, {\em 6},~ENEURO.0412--18.2019.
+  \newblock {\url{https://doi.org/10.1523/eneuro.0412-18.2019}}.
+
+  \bibitem[Khoei \em{et~al.}(2013)Khoei, Masson, and
     Perrinet]{khoei_motion-based_2013}
@@
     extrapolation.
-\newblock {\em J. Physiol.-Paris} {\bf 2013}, {\em 107},~409--420.
-\newblock {{https://doi.org/10.1016/j.jphysparis.2013.08.001}}.
-
-\bibitem[Kaplan {et~al.}(2013)Kaplan, Lansner, Masson, and
+  \newblock {\em Journal of Physiology-Paris} {\bf 2013}, {\em 107},~409--420.
+  \newblock {\url{https://doi.org/10.1016/j.jphysparis.2013.08.001}}.
+
+  \bibitem[Kaplan \em{et~al.}(2013)Kaplan, Lansner, Masson, and
     Perrinet]{kaplan_anisotropic_2013}
@@
     spiking neural network.
-\newblock {\em Front. Comput. Neurosci.} {\bf 2013}, {\em 7}, 112.
-%\newblock Loaded from an external bibliography file by Manubot.
-%  source\_bibliography: manual-references.bib standard\_id: Kaplan13,
-  {{https://doi.org/10.3389/fncom.2013.00112}}.
-
-\bibitem[Khoei {et~al.}(2017)Khoei, Masson, and
+  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2013}, {\em 7}.
+  \newblock Loaded from an external bibliography file by Manubot.
+    source\_bibliography: manual-references.bib standard\_id: Kaplan13,
+    {\url{https://doi.org/10.3389/fncom.2013.00112}}.
+
+  \bibitem[Khoei \em{et~al.}(2017)Khoei, Masson, and
     Perrinet]{khoei_flash-lag_2017}
@@
     {Shift}.
-\newblock {\em PLoS Comput. Biol.} {\bf 2017}, {\em 13},~e1005068.
-%\newblock Publisher: Public Library of Science,
-  {{https://doi.org/10.1371/journal.pcbi.1005068}}.
-
-\bibitem[Javanshir {et~al.}(2022)Javanshir, Nguyen, Mahmud, and
+  \newblock {\em PLOS Computational Biology} {\bf 2017}, {\em 13},~e1005068.
+  \newblock Publisher: Public Library of Science,
+    {\url{https://doi.org/10.1371/journal.pcbi.1005068}}.
+
+  \bibitem[Javanshir \em{et~al.}(2022)Javanshir, Nguyen, Mahmud, and
     Kouzani]{javanshir_advancements_2022}
@@
     {Spiking} {Neural} {Networks}.
-\newblock {\em Neural Comput.} {\bf 2022}, {\em 34},~1289--1328.
-\newblock {{https://doi.org/10.1162/neco\_a\_01499}}.
-
-\bibitem[Marković {et~al.}(2020)Marković, Mizrahi, Querlioz, and
+  \newblock {\em Neural Computation} {\bf 2022}, {\em 34},~1289--1328.
+  \newblock {\url{https://doi.org/10.1162/neco_a_01499}}.
+
+  \bibitem[Marković \em{et~al.}(2020)Marković, Mizrahi, Querlioz, and
     Grollier]{markovic_physics_2020}
@@
   \newblock Physics for neuromorphic computing.
-\newblock {\em Nat. Rev. Phys.} {\bf 2020}, {\em 2},~499--510.
-%\newblock Number: 9 Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/s42254-020-0208-2}}.
-
-\bibitem[Rasetto {et~al.}(2022)Rasetto, Wan, Akolkar, Shi, Xiong, and
+  \newblock {\em Nature Reviews Physics} {\bf 2020}, {\em 2},~499--510.
+  \newblock Number: 9 Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/s42254-020-0208-2}}.
+
+  \bibitem[Rasetto \em{et~al.}(2022)Rasetto, Wan, Akolkar, Shi, Xiong, and
     Benosman]{rasetto_challenges_2022}
@@
     {Processors}: {How} {Memristors} {Dynamic} {Properties} {Could}
-  {Revolutionize} {Machine} {Learning}. \emph{arXiv}  \textbf{2022},  	arXiv:2201.12673.
-%\newblock Technical Report 2201.12673, .
+    {Revolutionize} {Machine} {Learning}.
+  \newblock Technical Report 2201.12673, arXiv,  2022.
 
@@
   \newblock {NEST}: {An} {Environment} for {Neural} {Systems} {Simulations}.
-\newblock {\em GWDG-Bericht Nr. 58 Theo Plesser, Volker Macho (Hrsg.)}; {
-  {2003}}; %MDPI: Please add the name of the publisher and their location. This is a tech report
- p.~29.
-
-\bibitem[Hazan {et~al.}(2018)Hazan, Saunders, Khan, Patel, Sanghavi,
+  \newblock {\em GWDG-Bericht Nr. 58 Theo Plesser, Volker Macho (Hrsg.)} {\bf
+    2003}, p.~29.
+
+  \bibitem[Hazan \em{et~al.}(2018)Hazan, Saunders, Khan, Patel, Sanghavi,
     Siegelmann, and Kozma]{hazan_bindsnet_2018}
@@
     H.T.; Kozma, R.
-\newblock \textls[-15]{{BindsNET}: {A} {Machine} {Learning}-{Oriented} {Spiking} {Neural}
-  {Networks} {Library} in {Python}.}
-\newblock {\em Front. Neuroinformatics} {\bf 2018}, {\em 12}, 89.
-\newblock {{https://doi.org/10.3389/fninf.2018.00089}}.
-
-\bibitem[Stimberg {et~al.}(2019)Stimberg, Brette, and
+  \newblock {BindsNET}: {A} {Machine} {Learning}-{Oriented} {Spiking} {Neural}
+    {Networks} {Library} in {Python}.
+  \newblock {\em Frontiers in Neuroinformatics} {\bf 2018}, {\em 12}.
+  \newblock {\url{https://doi.org/10.3389/fninf.2018.00089}}.
+
+  \bibitem[Stimberg \em{et~al.}(2019)Stimberg, Brette, and
     Goodman]{stimberg_brian_2019}
@@
   \newblock {\em eLife} {\bf 2019}, {\em 8},~e47314.
-\newblock {{https://doi.org/10.7554/eLife.47314}}.
-
-\bibitem[Zenke {et~al.}(2021)Zenke, Bohté, Clopath, Comşa, Göltz, Maass,
+  \newblock {\url{https://doi.org/10.7554/eLife.47314}}.
+
+  \bibitem[Zenke \em{et~al.}(2021)Zenke, Bohté, Clopath, Comşa, Göltz, Maass,
     Masquelier, Naud, Neftci, Petrovici, Scherr, and
@@
   \newblock {\em Neuron} {\bf 2021}, {\em 109},~571--575.
-\newblock {{https://doi.org/10.1016/j.neuron.2021.01.009}}.
+  \newblock {\url{https://doi.org/10.1016/j.neuron.2021.01.009}}.
 
@@
   \newblock {\em Analog {VLSI} {Implementation} of {Neural} {Systems}}; Springer
-  Science \& Business Media:  {Berlin/Heidelberg, Germany,} %newly added information, please confirm
-  1989.
-%\newblock Google-Books-ID: 9e29dOiXeiMC.
+    Science \& Business Media,  1989.
+  \newblock Google-Books-ID: 9e29dOiXeiMC.
 
@@
   \newblock Synaptic {Dynamics} in {Analog} {VLSI}.
-\newblock {\em Neural Comput.} {\bf 2007}, {\em 19},~2581--2603.\linebreak
-\newblock {{https://doi.org/10.1162/neco.2007.19.10.2581}}.
-
-\bibitem[Schuman {et~al.}(2017)Schuman, Potok, Patton, Birdwell, Dean, Rose,
+  \newblock {\em Neural Computation} {\bf 2007}, {\em 19},~2581--2603.
+  \newblock {\url{https://doi.org/10.1162/neco.2007.19.10.2581}}.
+
+  \bibitem[Schuman \em{et~al.}(2017)Schuman, Potok, Patton, Birdwell, Dean, Rose,
     and Plank]{schuman_survey_2017}
@@
   \newblock A {Survey} of {Neuromorphic} {Computing} and {Neural} {Networks} in
-  {Hardware}. \emph{arXiv} {\bf 2017}, arXiv:1705.06963
-%\newblock {\em  [cs]} .
-%\newblock arXiv: 1705.06963.
-
-\bibitem[Furber {et~al.}(2013)Furber, Lester, Plana, Garside, Painkras,
+    {Hardware}.
+  \newblock {\em arXiv:1705.06963 [cs]} {\bf 2017}.
+  \newblock arXiv: 1705.06963.
+
+  \bibitem[Furber \em{et~al.}(2013)Furber, Lester, Plana, Garside, Painkras,
     Temple, and Brown]{furber_overview_2013}
@@
   \newblock Overview of the {SpiNNaker} {System} {Architecture}.
-\newblock {\em IEEE Trans. Comput.} {\bf 2013}, {\em
+  \newblock {\em IEEE Transactions on Computers} {\bf 2013}, {\em
     62},~2454--2467.
-\newblock {{https://doi.org/10.1109/TC.2012.142}}.
+  \newblock {\url{https://doi.org/10.1109/TC.2012.142}}.
 
@@
   \newblock {\em {SpiNNaker}: {A} {Spiking} {Neural} {Network} {Architecture}};
-  Now Publishers: {Norwell, MA, USA,} %MDPI: We added the location of the publisher. Please confirm.
-  2020.
-\newblock {{https://doi.org/10.1561/9781680836523}}.
-
-\bibitem[Merolla {et~al.}(2014)Merolla, Arthur, Alvarez-Icaza, Cassidy,
+    Now Publishers,  2020.
+  \newblock {\url{https://doi.org/10.1561/9781680836523}}.
+
+  \bibitem[Merolla \em{et~al.}(2014)Merolla, Arthur, Alvarez-Icaza, Cassidy,
     Sawada, Akopyan, Jackson, Imam, Guo, Nakamura, Brezzo, Vo, Esser, Appuswamy,
@@
   \newblock {\em Science} {\bf 2014}, {\em 345},~668--673.
-\newblock {{https://doi.org/10.1126/science.1254642}}.
-
-\bibitem[Benjamin {et~al.}(2014)Benjamin, Gao, McQuinn, Choudhary,
+  \newblock {\url{https://doi.org/10.1126/science.1254642}}.
+
+  \bibitem[Benjamin \em{et~al.}(2014)Benjamin, Gao, McQuinn, Choudhary,
     Chandrasekaran, Bussat, Alvarez-Icaza, Arthur, Merolla, and
@@
     {Large}-{Scale} {Neural} {Simulations}.
-\newblock {\em Proc.  IEEE} {\bf 2014}, {\em 102},~699--716.
-%\newblock Conference Name: Proceedings of the IEEE,
-  {{https://doi.org/10.1109/JPROC.2014.2313565}}.
-
-\bibitem[Neckar {et~al.}(2019)Neckar, Fok, Benjamin, Stewart, Oza, Voelker,
+  \newblock {\em Proceedings of the IEEE} {\bf 2014}, {\em 102},~699--716.
+  \newblock Conference Name: Proceedings of the IEEE,
+    {\url{https://doi.org/10.1109/JPROC.2014.2313565}}.
+
+  \bibitem[Neckar \em{et~al.}(2019)Neckar, Fok, Benjamin, Stewart, Oza, Voelker,
     Eliasmith, Manohar, and Boahen]{neckar_braindrop_2019}
@@
     dynamical systems-based programming model.
-\newblock {\em Proc.  IEEE} {\bf 2019}, {\em 107},~144--164.
-\newblock {{https://doi.org/10.1109/JPROC.2018.2881432}}.
-
-\bibitem[Schemmel {et~al.}(2010)Schemmel, Brüderle, Grübl, Hock, Meier,
+  \newblock {\em Proceedings of the IEEE} {\bf 2019}, {\em 107},~144--164.
+  \newblock {\url{https://doi.org/10.1109/JPROC.2018.2881432}}.
+
+  \bibitem[Schemmel \em{et~al.}(2010)Schemmel, Brüderle, Grübl, Hock, Meier,
     and Millner]{schemmel_wafer-scale_2010}
@@
   \newblock In Proceedings of the 2010 {IEEE} {International} {Symposium} on
-  {Circuits} and {Systems} ({ISCAS}), {Paris, France, 30 May--2 June} %MDPI: We added the location and date of the conference. Please confirm.
- 2010; pp. 1947--1950.
-{{https://doi.org/10.1109/ISCAS.2010.5536970}}.
-
-\bibitem[Markram {et~al.}(2011)Markram, Meier, Lippert, Grillner,
+    {Circuits} and {Systems} ({ISCAS}),  2010, pp. 1947--1950.
+  \newblock ISSN: 2158-1525, {\url{https://doi.org/10.1109/ISCAS.2010.5536970}}.
+
+  \bibitem[Markram \em{et~al.}(2011)Markram, Meier, Lippert, Grillner,
     Frackowiak, Dehaene, Knoll, Sompolinsky, Verstreken, DeFelipe, Grant,
@@
   \newblock Introducing the {Human} {Brain} {Project}.
-\newblock {\em Procedia Comput. Sci.} {\bf 2011}, {\em 7},~39--42.
-\newblock {{https://doi.org/10.1016/j.procs.2011.12.015}}.
-
-\bibitem[Farquhar {et~al.}(2006)Farquhar, Gordon, and
+  \newblock {\em Procedia Computer Science} {\bf 2011}, {\em 7},~39--42.
+  \newblock {\url{https://doi.org/10.1016/j.procs.2011.12.015}}.
+
+  \bibitem[Farquhar \em{et~al.}(2006)Farquhar, Gordon, and
     Hasler]{farquhar_field_2006}
@@
   \newblock In Proceedings of the 2006 {IEEE} {International} {Symposium} on
-  {Circuits} and {Systems}, {Kos, Greece, 21--24 May} %MDPI: We added the location and date of the conference. Please confirm.
- 2006; pp. 4114--4117.
-\newblock {{https://doi.org/10.1109/ISCAS.2006.1693534}}.
-
-\bibitem[Liu {et~al.}(2009)Liu, Yu, and Wang]{cheng_fpaa_2009}
+    {Circuits} and {Systems}; IEEE: Island of Kos, Greece,  2006; pp. 4114--4117.
+  \newblock {\url{https://doi.org/10.1109/ISCAS.2006.1693534}}.
+
+  \bibitem[Liu \em{et~al.}(2009)Liu, Yu, and Wang]{cheng_fpaa_2009}
   Liu, M.; Yu, H.; Wang, W.
@@
     {Devices} for {Neuromorphic} {Applications}. In {\em Nano-{Net}}; Cheng, M.,
-  Ed.; Springer: Berlin/Heidelberg, Germany,  2009; Volume~3, pp.
+    Ed.; Springer Berlin Heidelberg: Berlin, Heidelberg,  2009; Vol.~3, pp.
     44--48.
 
-\bibitem[Chan {et~al.}(2007)Chan, Liu, and van Schaik]{chan_aer_2007}
+  \bibitem[Chan \em{et~al.}(2007)Chan, Liu, and van Schaik]{chan_aer_2007}
   Chan, V.; Liu, S.C.; van Schaik, A.
@@
     {Address} {Event} {Representation} {Interface}.
-\newblock {\em IEEE Trans. Circuits Syst. I Regul. Pap.}
+  \newblock {\em IEEE Transactions on Circuits and Systems I: Regular Papers}
     {\bf 2007}, {\em 54},~48--59.
-%\newblock Conference Name: IEEE Transactions on Circuits and Systems I: Regular
-%  Papers,
-  {{https://doi.org/10.1109/TCSI.2006.887979}}.
-
-\bibitem[Haessig {et~al.}(2020)Haessig, Milde, Aceituno, Oubari, Knight, van
+  \newblock Conference Name: IEEE Transactions on Circuits and Systems I: Regular
+    Papers, {\url{https://doi.org/10.1109/TCSI.2006.887979}}.
+
+  \bibitem[Haessig \em{et~al.}(2020)Haessig, Milde, Aceituno, Oubari, Knight, van
     Schaik, Benosman, and Indiveri]{haessig_event-based_2020}
@@
     {Precise} {Spike} {Timing}.
-\newblock {\em Front. Neurosci.} {\bf 2020}, {\em 14},~420.
-\newblock {{https://doi.org/10.3389/fnins.2020.00420}}.
-
-\bibitem[Lagorce {et~al.}(2017)Lagorce, Orchard, Galluppi, Shi, and
+  \newblock {\em Frontiers in Neuroscience} {\bf 2020}, {\em 14},~420.
+  \newblock {\url{https://doi.org/10/ghvnxj}}.
+
+  \bibitem[Lagorce \em{et~al.}(2017)Lagorce, Orchard, Galluppi, Shi, and
     Benosman]{lagorce_hots_2017}
@@
     {Pattern} {Recognition}.
-\newblock {\em IEEE Trans. Pattern Anal. Mach. Intell.}
+  \newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}
     {\bf 2017}, {\em 39},~1346--1359.
-\newblock {{https://doi.org/10.1109/TPAMI.2016.2574707}}.
-
-\bibitem[Sironi {et~al.}(2018)Sironi, Brambilla, Bourdis, Lagorce, and
+  \newblock {\url{https://doi.org/10.1109/TPAMI.2016.2574707}}.
+
+  \bibitem[Sironi \em{et~al.}(2018)Sironi, Brambilla, Bourdis, Lagorce, and
     Benosman]{sironi_hats_2018}
@@
   \newblock In Proceedings of the 2018 {IEEE}/{CVF} {Conference} on {Computer}
-  {Vision} and {Pattern} {Recognition}, Salt Lake City, UT, USA, {18--23 June} %MDPI: We added the  date of the conference. Please confirm.
-  2018;
+    {Vision} and {Pattern} {Recognition}; IEEE: Salt Lake City, UT, USA,  2018;
     pp. 1731--1740.
-\newblock {{https://doi.org/10.1109/CVPR.2018.00186}}.
-
-\bibitem[Maro {et~al.}(2020)Maro, Ieng, and Benosman]{maro_event-based_2020}
+  \newblock {\url{https://doi.org/10.1109/CVPR.2018.00186}}.
+
+  \bibitem[Maro \em{et~al.}(2020)Maro, Ieng, and Benosman]{maro_event-based_2020}
   Maro, J.M.; Ieng, S.H.; Benosman, R.
@@
     {Suppression} {Using} {Smartphone} {Computational} {Capabilities}.
-\newblock {\em Front. Neurosci.} {\bf 2020}, {\em 14}, 275.
-%\newblock Publisher: Frontiers,
-  {{https://doi.org/10.3389/fnins.2020.00275}}.
-
-\bibitem[Grimaldi {et~al.}(2021)Grimaldi, Boutin, Perrinet, Ieng, and
+  \newblock {\em Frontiers in Neuroscience} {\bf 2020}, {\em 14}.
+  \newblock Publisher: Frontiers,
+    {\url{https://doi.org/10.3389/fnins.2020.00275}}.
+
+  \bibitem[Grimaldi \em{et~al.}(2021)Grimaldi, Boutin, Perrinet, Ieng, and
     Benosman]{grimaldi_homeostatic_2021}
@@
   \newblock In Proceedings of the 2021 {International} {Conference} on
-  {Content}-{Based} {Multimedia} {Indexing} ({CBMI}), {Lille, France,  28--30 June} %MDPI: We added the location and date of the conference. Please confirm.
-  2021.
-\newblock {{https://doi.org/10.1109/cbmi50038.2021.9461901}}.
-
-\bibitem[Grimaldi {et~al.}(2022)Grimaldi, Boutin, Ieng, Benosman, and
+    {Content}-{Based} {Multimedia} {Indexing} ({CBMI}). IEEE,  2021.
+  \newblock {\url{https://doi.org/10.1109/cbmi50038.2021.9461901}}.
+
+  \bibitem[Grimaldi \em{et~al.}(2022)Grimaldi, Boutin, Ieng, Benosman, and
     Perrinet]{grimaldi_robust_2022}
@@
   \newblock A robust event-driven approach to always-on object recognition.
-\newblock {\em TechRxiv} {\bf 2022}.
-\newblock {{https://doi.org/10.36227/techrxiv.18003077.v1}}.
-
-\bibitem[Yu {et~al.}(2022)Yu, Gu, Li, Wang, Wang, and Li]{yu_stsc-snn_2022}
+  \newblock {\em TechRxiv preprint} {\bf 2022}.
+  \newblock {\url{https://doi.org/10.36227/techrxiv.18003077.v1}}.
+
+  \bibitem[Yu \em{et~al.}(2022)Yu, Gu, Li, Wang, Wang, and Li]{yu_stsc-snn_2022}
   Yu, C.; Gu, Z.; Li, D.; Wang, G.; Wang, A.; Li, E.
   \newblock {STSC}-{SNN}: {Spatio}-{Temporal} {Synaptic} {Connection} with
-  {Temporal} {Convolution} and {Attention} for {Spiking} {Neural} {Networks}.  	\emph{arXiv}
+    {Temporal} {Convolution} and {Attention} for {Spiking} {Neural} {Networks},
     2022.
-\newblock arXiv:2210.05241.
-
-\bibitem[Benosman {et~al.}(2014)Benosman, Clercq, Lagorce, {Sio-Hoi Ieng},
+  \newblock arXiv:2210.05241 [cs, q-bio, stat].
+
+  \bibitem[Benosman \em{et~al.}(2014)Benosman, Clercq, Lagorce, {Sio-Hoi Ieng},
     and Bartolozzi]{benosman_event-based_2014}
-Benosman, R.; Clercq, C.; Lagorce, X.; Ieng, {S.-H}.; Bartolozzi, C.
+  Benosman, R.; Clercq, C.; Lagorce, X.; {Sio-Hoi Ieng}.; Bartolozzi, C.
   \newblock Event-{Based} {Visual} {Flow}.
-\newblock {\em IEEE Trans. Neural Netw. Learn. Syst.} {\bf
+  \newblock {\em IEEE Transactions on Neural Networks and Learning Systems} {\bf
     2014}, {\em 25},~407--417.
-\newblock {{https://doi.org/10.1109/tnnls.2013.2273537}}.
-
-\bibitem[Clady {et~al.}(2014)Clady, Clercq, Ieng, Houseini, Randazzo,
+  \newblock {\url{https://doi.org/10.1109/tnnls.2013.2273537}}.
+
+  \bibitem[Clady \em{et~al.}(2014)Clady, Clercq, Ieng, Houseini, Randazzo,
     Natale, Bartolozzi, and Benosman]{clady_asynchronous_2014}
@@
   \newblock Asynchronous visual event-based time-to-contact.
-\newblock {\em Front. Neurosci.} {\bf 2014}, {\em 8}, 9.
-%\newblock 00000 Publisher: Frontiers,
-  {{https://doi.org/10.3389/fnins.2014.00009}}.
-
-\bibitem[Tschechne {et~al.}(2014)Tschechne, Sailer, and
+  \newblock {\em Frontiers in Neuroscience} {\bf 2014}, {\em 8}.
+  \newblock 00000 Publisher: Frontiers,
+    {\url{https://doi.org/10.3389/fnins.2014.00009}}.
+
+  \bibitem[Tschechne \em{et~al.}(2014)Tschechne, Sailer, and
     Neumann]{tschechne_bio-inspired_2014}
@@
     {Sensor} {Input}.
-\newblock In \emph{Proceedings of the Artificial {Neural} {Networks} in {Pattern}
-  {Recognition}}; El~Gayar, N.; Schwenker, F.; Suen, C., Eds.; Springer
-  International Publishing: Cham, Switzerland, 2014; pp. 171--182.
-\newblock {{https://doi.org/10.1007/978-3-319-11656-3\_16}}.
-
-\bibitem[Hidalgo-Carrió {et~al.}(2020)Hidalgo-Carrió, Gehrig, and
+  \newblock In Proceedings of the Artificial {Neural} {Networks} in {Pattern}
+    {Recognition}; El~Gayar, N.; Schwenker, F.; Suen, C., Eds.; Springer
+    International Publishing: Cham,  2014; Lecture {Notes} in {Computer}
+    {Science}, pp. 171--182.
+  \newblock {\url{https://doi.org/10.1007/978-3-319-11656-3_16}}.
+
+  \bibitem[Hidalgo-Carrió \em{et~al.}(2020)Hidalgo-Carrió, Gehrig, and
     Scaramuzza]{hidalgo-carrio_learning_2020}
   Hidalgo-Carrió, J.; Gehrig, D.; Scaramuzza, D.
-\newblock Learning {Monocular} {Dense} {Depth} from {Events}.  \emph{arXiv}, {\bf 2020}, arXiv:2010.08350.
-%\newblock {\em arXiv:2010.08350 [cs]} .
-%\newblock arXiv: 2010.08350.
-
-\bibitem[Dardelet {et~al.}(2021)Dardelet, Benosman, and
+  \newblock Learning {Monocular} {Dense} {Depth} from {Events}.
+  \newblock {\em arXiv:2010.08350 [cs]} {\bf 2020}.
+  \newblock arXiv: 2010.08350.
+
+  \bibitem[Dardelet \em{et~al.}(2021)Dardelet, Benosman, and
     Ieng]{dardelet_event-by-event_2021}
@@
   \newblock An {Event}-by-{Event} {Feature} {Detection} and {Tracking}
-  {Invariant} to {Motion} {Direction} and {Velocity}. \emph{TechRxiv} {\bf 2021}.
-\newblock {{https://doi.org/10.36227/techrxiv.17013824.v1}}.
-
-\bibitem[Stoffregen {et~al.}(2019)Stoffregen, Gallego, Drummond, Kleeman,
+    {Invariant} to {Motion} {Direction} and {Velocity} {\bf 2021}.
+  \newblock {\url{https://doi.org/10.36227/techrxiv.17013824.v1}}.
+
+  \bibitem[Stoffregen \em{et~al.}(2019)Stoffregen, Gallego, Drummond, Kleeman,
     and Scaramuzza]{stoffregen_event-based_2019}
@@
   \newblock In Proceedings of the 2019 {IEEE}/{CVF} {International} {Conference}
-  on {Computer} {Vision} ({ICCV}), {Seoul, Republic of Korea, 27 February} %MDPI: We added the location and date of the conference. Please confirm.
- 2019; pp. 7243--7252.
- {{https://doi.org/10.1109/ICCV.2019.00734}}.
-
-\bibitem[Kim {et~al.}(2016)Kim, Leutenegger, and
+    on {Computer} {Vision} ({ICCV}),  2019, pp. 7243--7252.
+  \newblock ISSN: 2380-7504, {\url{https://doi.org/10.1109/ICCV.2019.00734}}.
+
+  \bibitem[Kim \em{et~al.}(2016)Kim, Leutenegger, and
     Davison]{kim_real-time_2016}
@@
     {Event} {Camera}.
-\newblock In \emph{Proceedings of the Computer {Vision}---{ECCV} 2016}; Leibe, B.;
+  \newblock In Proceedings of the Computer {Vision} – {ECCV} 2016; Leibe, B.;
     Matas, J.; Sebe, N.; Welling, M., Eds.; Springer International Publishing:
-  Cham,  Switzerland, 2016;  pp. 349--364.
-\newblock {{https://doi.org/10.1007/978-3-319-46466-4\_21}}.
-
-\bibitem[Hussain {et~al.}(2012)Hussain, Basu, Wang, and
+    Cham,  2016; Lecture {Notes} in {Computer} {Science}, pp. 349--364.
+  \newblock {\url{https://doi.org/10.1007/978-3-319-46466-4_21}}.
+
+  \bibitem[Hussain \em{et~al.}(2012)Hussain, Basu, Wang, and
     Hamilton]{hussain_deltron_2012}
@@
   \newblock In Proceedings of the 2012 {IEEE} {Asia} {Pacific} {Conference} on
-  {Circuits} and {Systems},  {Kaohsiung, Taiwan, 2--5 December} %MDPI: We added the location and date of the conference. Please confirm.
- 2012; pp. 304--307.
-\newblock {{https://doi.org/10.1109/APCCAS.2012.6419032}}.
-
-\bibitem[Wang {et~al.}(2015)Wang, Hamilton, Tapson, and van
+    {Circuits} and {Systems},  2012, pp. 304--307.
+  \newblock {\url{https://doi.org/10.1109/APCCAS.2012.6419032}}.
+
+  \bibitem[Wang \em{et~al.}(2015)Wang, Hamilton, Tapson, and van
     Schaik]{wang_neuromorphic_2015}
@@
     plasticity rules for large-scale neural networks.
-\newblock {\em Front. Neurosci.} {\bf 2015}, {\em 9}, 180.
-\newblock {{https://doi.org/10.3389/fnins.2015.00180}}.
-
-\bibitem[Wang {et~al.}(2014)Wang, Hamilton, Tapson, and van
+  \newblock {\em Frontiers in Neuroscience} {\bf 2015}, {\em 9}.
+  \newblock {\url{https://doi.org/10.3389/fnins.2015.00180}}.
+
+  \bibitem[Wang \em{et~al.}(2014)Wang, Hamilton, Tapson, and van
     Schaik]{wang_fpga_2014}
@@
   \newblock In Proceedings of the 2014 {IEEE} {International} {Symposium} on
-  {Circuits} and {Systems} ({ISCAS}), {Melbourne, VIC, Australia, 1--5 June} %MDPI: We added the location and date of the conference. Please confirm.
- 2014; pp. 457--460.
- {{https://doi.org/10.1109/ISCAS.2014.6865169}}.
-
-\bibitem[Pfeil {et~al.}(2013)Pfeil, Scherzer, Schemmel, and
+    {Circuits} and {Systems} ({ISCAS}),  2014, pp. 457--460.
+  \newblock ISSN: 2158-1525, {\url{https://doi.org/10.1109/ISCAS.2014.6865169}}.
+
+  \bibitem[Pfeil \em{et~al.}(2013)Pfeil, Scherzer, Schemmel, and
     Meier]{pfeil_neuromorphic_2013}
@@
   \newblock In Proceedings of the The 2013 {International} {Joint} {Conference}
-  on {Neural} {Networks} ({IJCNN}), {Dallas, TX, USA, 4--9 August} %MDPI: We added the location and date of the conference. Please confirm.
- 2013; pp. 1--5.
- {{https://doi.org/10.1109/IJCNN.2013.6706828}}.
+    on {Neural} {Networks} ({IJCNN}),  2013, pp. 1--5.
+  \newblock ISSN: 2161-4407, {\url{https://doi.org/10.1109/IJCNN.2013.6706828}}.
 
@@
   \newblock Spike-{Based} {Population} {Coding} and {Working} {Memory}.
-\newblock {\em PLoS Comput. Biol.} {\bf 2011}, {\em 7},~e1001080.
-\newblock {{https://doi.org/10.1371/journal.pcbi.1001080}}.
-
-\bibitem[Renner {et~al.}(2022)Renner, Sandamirskaya, Sommer, and
+  \newblock {\em PLOS Computational Biology} {\bf 2011}, {\em 7},~e1001080.
+  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1001080}}.
+
+  \bibitem[Renner \em{et~al.}(2022)Renner, Sandamirskaya, Sommer, and
     Frady]{renner_sparse_2022}
@@
   \newblock {\em Proceedings of the International Conference on Neuromorphic
-  Systems};
-%\newblock Conference Name: ICONS 2022: International Conference on Neuromorphic
-%  Systems 2022 Meeting Name: ICONS 2022: International Conference on
-%  Neuromorphic Systems 2022 Place:  
-  Publisher: ACM Digital
-  library: Knoxville, TN, USA, 2022. {{https://doi.org/10.1145/3546790.3546820}}.
-
-\bibitem[Dard {et~al.}(2022)Dard, Leprince, Denis, Rao~Balappa, Suchkov,
+    Systems} {\bf 2022}.
+  \newblock Conference Name: ICONS 2022: International Conference on Neuromorphic
+    Systems 2022 Meeting Name: ICONS 2022: International Conference on
+    Neuromorphic Systems 2022 Place: Knoxville, TN, USA Publisher: ACM Digital
+    library, {\url{https://doi.org/10.1145/3546790.3546820}}.
+
+  \bibitem[Dard \em{et~al.}(2022)Dard, Leprince, Denis, Rao~Balappa, Suchkov,
     Boyce, Lopez, Giorgi-Kurz, Szwagier, Dumont, Rouault, Minlebaev, Baude,
@@
   \newblock {\em eLife} {\bf 2022}, {\em 11},~e78116.
-\newblock {{https://doi.org/10.7554/eLife.78116}}.
+  \newblock {\url{https://doi.org/10.7554/eLife.78116}}.
 
@@
     implications for schizophrenia.
-\newblock {\em Nat. Rev. Psychol.} {\bf 2022}, {\em 1},~257--271.
-%\newblock Publisher: Nature Publishing Group,
-  {{https://doi.org/10.1038/s44159-022-00038-y}}.
-
-\bibitem[Panahi {et~al.}(2021)Panahi, Abrevaya, Gagnon-Audet, Voleti, Rish,
+  \newblock {\em Nature Reviews Psychology} {\bf 2022}, {\em 1},~257--271.
+  \newblock Publisher: Nature Publishing Group,
+    {\url{https://doi.org/10.1038/s44159-022-00038-y}}.
+
+  \bibitem[Panahi \em{et~al.}(2021)Panahi, Abrevaya, Gagnon-Audet, Voleti, Rish,
     and Dumas]{panahi_generative_2021}
   Panahi, M.R.; Abrevaya, G.; Gagnon-Audet, J.C.; Voleti, V.; Rish, I.; Dumas, G.
-\newblock Generative {Models} of {Brain} {Dynamics}---{A} review. \emph{arXiv} \textbf{2021},
-\newblock arXiv:2112.12147.
-
-\bibitem[Tolle {et~al.}(2011)Tolle, Tansley, and Hey]{tolle_fourth_2011}
+  \newblock Generative {Models} of {Brain} {Dynamics} -- {A} review.
+  \newblock Technical Report 2112.12147, arXiv,  2021.
+
+  \bibitem[Tolle \em{et~al.}(2011)Tolle, Tansley, and Hey]{tolle_fourth_2011}
   Tolle, K.M.; Tansley, D.S.W.; Hey, A.J.G.
@@
     [{Point} of {View}].
-\newblock {\em Proc.  IEEE} {\bf 2011}, {\em 99},~1334--1337.
-\newblock {{https://doi.org/10.1109/jproc.2011.2155130}}.
+  \newblock {\em Proceedings of the IEEE} {\bf 2011}, {\em 99},~1334--1337.
+  \newblock {\url{https://doi.org/10.1109/jproc.2011.2155130}}.
 
@@
 475)
-\PublishersNote{}
+
 \end{adjustwidth}

### 2024-04-11 manuscript-arxiv.tex
Create manuscript-arxiv.tex

--- 2022-12-28 manuscript_rev1.tex
@@
 However, there is no consensus to explain the multiple nonlinear mechanisms that transform the visual scene into retinotopic neural activity maps, even though these processes seem to constitute essential pieces to this puzzle~\citep{carandini_normalization_2012}.
-
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=.7\textwidth]{figures/visual-latency.pdf}
+% thorpe
+% \includegraphics[width=.7\textwidth]{figures/visual-latency-estimate.jpg}%-estimate.jpg}
+\caption{{\bf Latency of the different processing steps along the human visual pathway.} Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
+Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
+In order, the retina is first activated ($20-40~\ms$), then the thalamus and the primary visual cortex (V1, $60-90~\ms$).
+This visual information  projects to the temporal lobe to reach the infero-temporal area (IT, $150~\ms$) for object recognition.
+It then reaches the prefrontal cortex (PFC, $180~\ms$) which modulates decision making to evoke the Motor Cortex (MC, $220~\ms$) which may mediate an action.
+This is eventually relayed through the spinal cord to trigger finger muscles, with latencies about $280-400~\ms$.}\label{fig:thorpe}
+\end{figure}
+%----------------------------%
 In particular, there have been some remarkable findings when studying the dynamics of vision.
@@
 
-%----------------------------%
-\begin{figure}[H]
+% retina
+\remove[Antoine]{How may information about the sensory world be encoded in the spiking pattern?} At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
+In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
+These results have been extended to natural images and show a qualitatively similar behavior.
+The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
+Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
+This evidence found in the retina can be extended to other areas such as the visual cortex.
+%----------------------------%
+% C_MainenSejnowski1995_Perrinet.ipynb
+\begin{figure}
 \centering
-  \includegraphics[width=.7\textwidth]{../2022-12-20_polychrony_review_REV1/figures/visual-latency.pdf}
-  % thorpe
-  % \includegraphics[width=.7\textwidth]{../2022-12-20_polychrony_review_REV1/figures/visual-latency-estimate.jpg}%-estimate.jpg}
-  \caption{{\bf Latency of the different processing steps along the human visual pathway.} Though the visual system is highly inter-connected, one can follow the sequence of activations whenever an image (here a yellow star) is flashed in front of the eyes.
-Different areas are schematically represented by ellipses and arrows denote the fastest feed-forward activation, ordered with respect to their activation latency in $\ms$.
-In order, the retina is first activated ($20-40~\ms$), then the thalamus and the primary visual cortex (V1, $60-90~\ms$).
-This visual information  projects to the temporal lobe to reach the infero-temporal area (IT, $150~\ms$) for object recognition.
-It then reaches the prefrontal cortex (PFC, $180~\ms$) which modulates decision making to evoke the Motor Cortex (MC, $220~\ms$) which may mediate an action.
-This is eventually relayed through the spinal cord to trigger finger muscles, with latencies about $280-400~\ms$.}\label{fig:thorpe}
-  \end{figure}
-  %----------------------------%
-  \vspace{-6pt}
-
-%----------------------------%
-% C_MainenSejnowski1995_Perrinet.ipynb
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{../2022-12-20_polychrony_review_REV1/figures/replicating_MainenSejnowski1995.pdf} 
+\includegraphics[width=\textwidth]{figures/replicating_MainenSejnowski1995.pdf} 
 \caption{\textbf{Reproducibility of the spiking response of a neuron.
@@
 %----------------------------%
-
-% retina
-\remove[Antoine]{How may information about the sensory world be encoded in the spiking pattern?} At the level of the retina, it has been shown that a coding of luminance values in the image using the timing of the spikes may be at work~\citep{gollisch_rapid_2008}.
-In particular, these results show that the response of ganglion cells to the visual gratings projected on the retina could be encoded in the latency of the response and not only in the frequency of the discharge, as it is often assumed.
-These results have been extended to natural images and show a qualitatively similar behavior.
-The authors' conclusion is that the precise spiking latency of the neurons encodes the spatial features of the image.
-Interestingly, such a precise latency mechanism may underlie some visual illusions, e.g., the false color illusion in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based on center-surround interactions in the parvocellular pathway~\citep{kenyon_theory_2004}.
-This evidence found in the retina can be extended to other areas such as the visual cortex.
 
@@
 %\centering
-%\includegraphics[width=.7\textwidth]{../2022-12-20_polychrony_review_REV1/figures/fnins-07-00011-g004.jpg}
+%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
 %\caption{A pyNN implementation of a Synfire chain with feedforward inhibition.
@@
 \centering
-\includegraphics[width=\textwidth]{../2022-12-20_polychrony_review_REV1/figures/Diesmann_et_al_1999.pdf}
+\includegraphics[width=\textwidth]{figures/Diesmann_et_al_1999.pdf}
 \caption{\textbf{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}.} The model consists of $10$ groups (arranged with the first group represented in the lowest row) of $100$ neurons each.
@@
 Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states.
-Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \emph{in vivo}.
+Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \textit{in vivo}.
 Such waves also occur in motor areas and~\citet{linden_movement_2022} have recently presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
@@
 %We have previously seen that neurobiologists have described stereotyped sequences of neural activation, notably in the hippocampus~\citep{pastalkova_internally_2008,villette_internally_2015,malvache_awake_2016}, and that neurons can be activated in the same order across days~\citep{haimerl_internal_2019}.
-Going further, researchers have found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \emph{in vivo} and \emph{in vitro}.
+Going further, researchers have found precise repetitions of spontaneous patterns of synaptic inputs in neocortical neurons, \textit{in vivo} and \textit{in vitro}.
 These patterns repeat after minutes, maintaining millisecond accuracy.
@@
 % \centering
-% \includegraphics[width=.6\linewidth]{../2022-12-20_polychrony_review_REV1/figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
+% \includegraphics[width=.6\linewidth]{figures/Ikegaya2004zse0150424620001.jpeg}%pcbi.1006283.g001.PNG_L.png}
 % \caption{Fig.
 1.
-from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity \emph{in vitro} and \emph{in vivo}.
+from~\citep{ikegaya_synfire_2004} repeated motifs of spontaneous synaptic activity \textit{in vitro} and \textit{in vivo}.
 (A) Repeated motifs of intracellular activity from layer 5 pyramidal neurons in slices.
@@
 The top traces show the motifs superimposed on each other (blue, green, and red), the middle traces show these same traces individually, and the bottom traces show temporally magnified regions of the motifs (a to c).
-(C) Repeated sequences of intracellular current-clamp recordings \emph{in vivo}.
+(C) Repeated sequences of intracellular current-clamp recordings \textit{in vivo}.
 Two (i) and three (ii) repetitions of motifs are shown.
@@
 \centering
-\includegraphics[width=\linewidth]{../2022-12-20_polychrony_review_REV1/figures/haimerl2019.jpg}
+\includegraphics[width=\linewidth]{figures/haimerl2019.jpg}
 \caption{\textbf{An example of a precise temporal motifs seen in cortical slices.} In this study by~\citep{haimerl_internal_2019}, an analysis of the raster plot shows repetitions of precise spiking motifs with a time scale of the order of seconds.
@@
 The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[Laurent]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.
-
-%----------------------------%
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../2022-12-20_polychrony_review_REV1/figures/Diesmann_et_al_1999_spade.pdf}
-  \caption{\textbf{Detecting motifs using SPADE.} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}.
-This allowed to label different precise spike motifs which are denoted by different colors.
-Spikes belonging to the same motif have the same color.
-%As the synfire chain gets more precise, the detected patterns grow in size.
-  }\label{fig:Diesmann_et_al_1999_spade}
-  \end{figure}
-  %----------------------------%
 
@@
 
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
+\caption{\textbf{Detecting motifs using SPADE.} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}.
+This allowed to label different precise spike motifs which are denoted by different colors.
+Spikes belonging to the same motif have the same color.
+%As the synfire chain gets more precise, the detected patterns grow in size.
+}\label{fig:Diesmann_et_al_1999_spade}
+\end{figure}
+%----------------------------%
+
 % %%----------------------------%
@@
 % \centering
-% \includegraphics[width=\linewidth]{../2022-12-20_polychrony_review_REV1/figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
+% \includegraphics[width=\linewidth]{figures/pcbi.1006283.g001.png}%pcbi.1006283.g001.PNG_L.png}
 % \caption{\textbf{The SPOTDisClust algorithm.} Fig 1 of~\citep{grossberger_unsupervised_2018}, showing simulated example illustrating the steps in the algorihtm.
@@
 Over the past decade, tremendous technological advances across several disciplines have dramatically expanded the frontiers of experimentally accessible neuroscientific facts.
-Bridging across different spatial and temporal scales, combination of \emph{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
+Bridging across different spatial and temporal scales, combination of \textit{in vivo} two photon imaging, large population recording-array technologies, optogenetic circuit control tools, transgenic manipulations as well as large volume circuit reconstructions are now used to examine the function, structure and dynamics of neural networks on an unprecedented level of detail and precision.
 The daunting complexity of the biological reality revealed by these technologies highlights the importance of neurobiological knowledge to provide a conceptual bridge between abstract principles of brain function and their biological implementations within neural circuits.
@@
 \add[Amélie]{It was first theorized in the Jeffress "ITD-versus-place model"}~\citep{jeffress_place_1948}.
-\add[Amélie]{As hypothesized by this model, the \emph{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \emph{nucleus magnocellularis} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
+\add[Amélie]{As hypothesized by this model, the \textit{nucleus laminaris} of the avian brain contains coincidence detectors and, associated with the \textit{nucleus magnocellularis} axons, forms circuits for processing ITD}~\citep{konishi_coding_2003}.
 According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since ``neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower'', therefore necessarily need to play on the response time to do so.
@@
 To answer this, we must first study the myelinization process.
-The OPCs first proliferate in the white matter \emph{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
+The OPCs first proliferate in the white matter \textit{via} a self-repulsive process, thus allowing for an evenly spaced network, and identify target axons.
 Most OPCs then differentiate into oligodendrocytes and immediately initiate myelinization, with no further migration~\citep{simons_oligodendrocytes_2016}.
@@
 They also confirmed that the delay correlates with the level of skill acquisition.~\citet{gibson_neuronal_2014} suggests that neuronal activity does not solely promote adaptive myelination in the mammalian brain, but also OPC differentiation and oligodendrogenesis.
-Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \emph{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
+Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or \textit{via} exocytosis, but does not require any axo-glial synaptic communication~\citep{fields_new_2015}.
 
@@
 %In particular, our objective is to describe methods which fully exploit the capacity of spiking neurons to detect synchronous patterns.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As we saw, most SNN, and in particular those adapted from analogous deep-learning-like architectures, rely on an encoding of information based on a continuously varying firing-rate.
-Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time encoding machine by}~\citet{lazar_time_2004} and \emph{polychronization} model of~\citet{izhikevich_polychronization_2006}.
+Notable exceptions of SNNs using precise spike timings are the \add[Antoine]{time encoding machine by}~\citet{lazar_time_2004} and \textit{polychronization} model of~\citet{izhikevich_polychronization_2006}.
 \add[Antoine]{In this section we focus on the polychronization model} based on a random recurrent model of spiking neurons including synaptic delays chosen from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}.
@@
 According to the STDP rule, the group of neurons involved in this polychronous activity will see their synaptic weight increase and, thus, may consolidate the formation of a polychronous group.
+%----------------------------%
+%
+\begin{figure}%[t!]
+  \centering
+  \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
+    \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \textit{(Left)}~In this example, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+    %at these different delays, and these may not be sufficient to generate a spike in either neuron.
+    \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
+     }
+  \label{fig:izhikevich}
+\end{figure}
+% 
+%----------------------------%
 
@@
 As such, these models use the neural dynamics to handle input signals with different delays but do not explicitly take full advantage of the representation capacity offered by heterogeneous delays.
-
-%----------------------------%
-%
-\begin{figure}[H]%[t!]
+%
+%----------------------------%
+%
+\begin{figure}%[t!]
   \centering
-  \includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
-    \caption{\textbf{Core mechanism of polychrony detection~\citep{izhikevich_polychronization_2006}.} \emph{(Left)}~In this example, three presynaptic neurons denoted \emph{b}, \emph{c} and, \emph{d} are fully connected to two post-synaptic neurons \emph{a} and \emph{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \emph{a} and  $8$, $5$, and $1~\ms$ for \emph{e}.
-\emph{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \emph{a} and \emph{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
-    %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-    \emph{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \emph{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \emph{e}.
-     }
-  \label{fig:izhikevich}
-\end{figure}
-% 
-%----------------------------%
-
-%
-%----------------------------%
-%
-\begin{figure}[H]%[t!]
-  \centering
-  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1a_k.pdf}
-  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1c.pdf}
-  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1b.pdf}
-  \includegraphics[width=0.490\linewidth]{../2022-12-20_polychrony_review_REV1/figures/THC_1a.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
+  \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
     \caption{\textbf{Detecting event-based motifs using spiking neurons with heterogeneous delays.} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------%
+% TODO
+\begin{figure}%[t!]
+  \centering
+  \includegraphics[width=0.980\linewidth]{figures/2022-09-27_MotionDetection_kernel.png}
+  %\includegraphics[width=0.980\linewidth]{figures/2022-10-10_MotionDetection_kernel.png}
+    \caption{\textbf{Detecting visual motion in an event stream with heterogeneous delays.} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
+The models parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
+When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push-pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}.
+Global weight is globally decreasing from the lowest delay (right) to less recent information (left).
+     }
+  \label{fig:icip}
+\end{figure}
+% 
+%----------------------------%
 
@@
 %
-
+%----------------------------%
+\begin{figure}
+\centering
+\includegraphics[width=0.980\linewidth]{figures/event_driven_computations.png}
+\caption{\textbf{A miniature, event-based ATIS sensor.} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
+These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
+Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
+\end{figure}
+%----------------------------%
+%
 Neuromorphic sensors are also developed with the idea to capture external stimuli more efficiently and closer to biological systems.
@@
 
-%----------------------------%
-\begin{figure}[H]
-  \centering
-\includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/event_driven_computations.png}
-\caption{\textbf{A miniature, event-based ATIS sensor.} Contrary to a classical frame-based camera for which a full dense image representation is given at discrete, regularly spaced timings, the event-based camera provides with events at the micro-second resolution.
-These are sparse as they represent luminance increments or decrements (ON and OFF events, respectively).
-Figure courtesy of Sio-Hoi Ieng (Sorbonne Université/UPMC, Institut de la Vision).}\label{fig:silicon_retina}
-\end{figure}
-%----------------------------%
-%
-
 %Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems.
@@
 Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER).
-In this section, we will use indifferently \emph{events} or \emph{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
+In this section, we will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and we will extend the review of the use of precise spike timing to this emerging field.
 There exist different types of event-based sensing devices for sound~\citep{chan_aer_2007} and touch~\citep{haessig_event-based_2020} and here we focus only on neuromorphic cameras to review different methods using efficiently the dynamics of event streams.
@@
 
-In~\citep{lagorce_hots_2017}, \emph{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
+In~\citep{lagorce_hots_2017}, \textit{time surfaces} are introduced as an event-driven 2D image of the delay between the last event recorded at the address of a pixel and the current time.
 An exponential decay is applied on this delay to obtain the analog values of the time surfaces.
@@
 %
-
-%----------------------------%
-% TODO
-\begin{figure}[H]%[t!]
-  \centering
-  \includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/2022-09-27_MotionDetection_kernel.png}
-  %\includegraphics[width=0.980\linewidth]{../2022-12-20_polychrony_review_REV1/figures/2022-10-10_MotionDetection_kernel.png}
-    \caption{\textbf{Detecting visual motion in an event stream with heterogeneous delays.} Extending the polychrony detection model to the spatial domain,~\citet{grimaldi_learning_2022} have applied a supervised learning scheme to the detection of motion.
-The models parameters are represented by different spatio-temporal kernels, and we show three examples as pairs of rows, one targeting ON spikes, the other OFF spikes, the first column representing the corresponding motion detected.
-When trained on a set of natural images, it shows the emergence of localized, oriented kernels organized in a so-called push-pull organization for which weights to an ON spike are negatively proportional to that to an OFF cell~\citep{kremkow_push-pull_2016}.
-Global weight is globally decreasing from the lowest delay (right) to less recent information (left).
-     }
-  \label{fig:icip}
-\end{figure}
-% 
-%----------------------------%
 \subsection{Computations with delays in neuromorphic hardware}
@@
 This community is based on solid and validated evidence which is getting novel grounds thanks to current technical advances.
-Moreover, we have also shown that this community is highly diverse, operating on \emph{in vivo}, \emph{in computo} and \emph{in silico} systems.
+Moreover, we have also shown that this community is highly diverse, operating in biology, computational neuroscience or neuromorphic engineering.
 As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
@@
 %\bibliographystyle{Definitions/chicago2}
-% \bibliography{references}
-\begin{thebibliography}{999}
-
-  \bibitem[Piccolino(1997)]{piccolino_luigi_1997}
-  Piccolino, M.
-  \newblock Luigi {Galvani} and animal electricity: two centuries after the
-    foundation of electrophysiology.
-  \newblock {\em Trends in Neurosciences} {\bf 1997}, {\em 20},~443--448.
-  \newblock {\url{https://doi.org/10.1016/S0166-2236(97)01101-6}}.
-
-  \bibitem[Adrian and Zotterman(1926)]{adrian_impulses_1926}
-  Adrian, E.D.; Zotterman, Y.
-  \newblock The impulses produced by sensory nerve endings.
-  \newblock {\em The Journal of Physiology} {\bf 1926}, {\em 61},~465--483.
-
-  \bibitem[Gouras(1960)]{gouras_graded_1960}
-  Gouras, P.
-  \newblock Graded potentials of bream retina.
-  \newblock {\em The Journal of Physiology} {\bf 1960}, {\em 152},~487--505.
-
-  \bibitem[Perkel \em{et~al.}(1967{\natexlab{a}})Perkel, Gerstein, and
-    Moore]{perkel_neuronal_1967}
-  Perkel, D.H.; Gerstein, G.L.; Moore, G.P.
-  \newblock Neuronal {Spike} {Trains} and {Stochastic} {Point} {Processes}: {I}.
-    {The} {Single} {Spike} {Train}.
-  \newblock {\em Biophysical Journal} {\bf 1967}, {\em 7},~391--418.
-  \newblock {\url{https://doi.org/10.1016/S0006-3495(67)86596-2}}.
-
-  \bibitem[Perkel \em{et~al.}(1967{\natexlab{b}})Perkel, Gerstein, and
-    Moore]{perkel_neuronal_1967-1}
-  Perkel, D.H.; Gerstein, G.L.; Moore, G.P.
-  \newblock Neuronal {Spike} {Trains} and {Stochastic} {Point} {Processes}: {II}.
-    {Simultaneous} {Spike} {Trains}.
-  \newblock {\em Biophysical Journal} {\bf 1967}, {\em 7},~419--440.
-  \newblock {\url{https://doi.org/10.1016/S0006-3495(67)86597-4}}.
-
-  \bibitem[Abeles(1982)]{abeles_role_1982}
-  Abeles, M.
-  \newblock Role of the cortical neuron: integrator or coincidence detector?
-  \newblock {\em Israel journal of medical sciences} {\bf 1982}, {\em
-    18},~83--92.
-
-  \bibitem[Carr(1993)]{carr_processing_1993}
-  Carr, C.E.
-  \newblock Processing of {Temporal} {Information} in the {Brain}.
-  \newblock {\em Annual Review of Neuroscience} {\bf 1993}, {\em 16},~223--243.
-  \newblock {\url{https://doi.org/10.1146/annurev.ne.16.030193.001255}}.
-
-  \bibitem[Davis \em{et~al.}(2021)Davis, Benigno, Fletterman, Desbordes, Steward,
-    Sejnowski, H~Reynolds, and Muller]{davis_spontaneous_2021}
-  Davis, Z.W.; Benigno, G.B.; Fletterman, C.; Desbordes, T.; Steward, C.;
-    Sejnowski, T.J.; H~Reynolds, J.; Muller, L.
-  \newblock Spontaneous traveling waves naturally emerge from horizontal fiber
-    time delays and travel through locally asynchronous-irregular states.
-  \newblock {\em Nature Communications} {\bf 2021}, {\em 12},~1--16.
-  \newblock {\url{https://doi.org/10.1038/s41467-021-26175-1}}.
-
-  \bibitem[Perrinet \em{et~al.}(2004)Perrinet, Samuelides, and
-    Thorpe]{perrinet_coding_2004}
-  Perrinet, L.; Samuelides, M.; Thorpe, S.
-  \newblock Coding static natural images using spiking event times: do neurons
-    cooperate?
-  \newblock {\em IEEE Transactions on neural networks} {\bf 2004}, {\em
-    15},~1164--1175.
-  \newblock Publisher: IEEE, {\url{https://doi.org/10.1109/TNN.2004.833303}}.
-
-  \bibitem[Gollisch and Meister(2008)]{gollisch_rapid_2008}
-  Gollisch, T.; Meister, M.
-  \newblock Rapid {Neural} {Coding} in the {Retina} with {Relative} {Spike}
-    {Latencies}.
-  \newblock {\em Science} {\bf 2008}, {\em 319},~1108--1111.
-  \newblock {\url{https://doi.org/10.1126/science.1149639}}.
-
-  \bibitem[DeWeese and Zador(2003)]{deweese_binary_2003}
-  DeWeese, M.R.; Zador, A.M.
-  \newblock {\em Binary coding in auditory cortex}; Neural Information Processing
-    Systems Foundation,  2003.
-
-  \bibitem[Carr and Konishi(1990)]{carr_circuit_1990}
-  Carr, C.E.; Konishi, M.
-  \newblock A circuit for detection of interaural time differences in the brain
-    stem of the barn owl.
-  \newblock {\em Journal of Neuroscience} {\bf 1990}, {\em 10},~3227--3246.
-  \newblock Publisher: Society for Neuroscience Section: Articles,
-    {\url{https://doi.org/10.1523/JNEUROSCI.10-10-03227.1990}}.
-
-  \bibitem[Bohte(2004)]{bohte_evidence_2004}
-  Bohte, S.M.
-  \newblock The evidence for neural information processing with precise
-    spike-times: {A} survey.
-  \newblock {\em Natural Computing} {\bf 2004}, {\em 3},~195--206.
-  \newblock {\url{https://doi.org/10.1023/B:NACO.0000027755.02868.60}}.
-
-  \bibitem[DiLorenzo and Victor(2013)]{dilorenzo_spike_2013}
-  DiLorenzo, P.M.; Victor, J.D.
-  \newblock {\em Spike {Timing}: {Mechanisms} and {Function}}; CRC Press,  2013.
-  \newblock Google-Books-ID: KTHUIMUpQCUC.
-
-  \bibitem[Roy \em{et~al.}(2019)Roy, Jaiswal, and Panda]{roy_towards_2019}
-  Roy, K.; Jaiswal, A.; Panda, P.
-  \newblock Towards spike-based machine intelligence with neuromorphic computing.
-  \newblock {\em Nature} {\bf 2019}, {\em 575},~607--617.
-  \newblock Number: 7784 Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/s41586-019-1677-2}}.
-
-  \bibitem[Flourens(1842)]{flourens_recherches_1842}
-  Flourens, M.J.P.
-  \newblock {\em Recherches expérimentales sur les propriétés et les fonctions
-    du système nerveux, dans les animaux vertébrés}; J.-B. Balliere,  1842.
-
-  \bibitem[Pearce(2009)]{pearce_marie-jean-pierre_2009}
-  Pearce, J.
-  \newblock Marie-{Jean}-{Pierre} {Flourens} (1794–1867) and {Cortical}
-    {Localization}.
-  \newblock {\em European Neurology} {\bf 2009}, {\em 61},~311--314.
-  \newblock {\url{https://doi.org/10.1159/000206858}}.
-
-  \bibitem[Hubel and Wiesel(1968)]{hubel_receptive_1968}
-  Hubel, D.H.; Wiesel, T.N.
-  \newblock Receptive fields and functional architecture of monkey striate
-    cortex.
-  \newblock {\em The Journal of Physiology} {\bf 1968}, {\em 195},~215--243.
-  \newblock {\url{https://doi.org/10.1113/jphysiol.1968.sp008455}}.
-
-  \bibitem[Carandini and Heeger(2012)]{carandini_normalization_2012}
-  Carandini, M.; Heeger, D.J.
-  \newblock Normalization as a canonical neural computation.
-  \newblock {\em Nature Reviews Neuroscience} {\bf 2012}, {\em 13},~51--62.
-  \newblock {\url{https://doi.org/10.1038/nrn3136}}.
-
-  \bibitem[Thorpe \em{et~al.}(1996)Thorpe, Fize, and Marlot]{thorpe_speed_1996}
-  Thorpe, S.; Fize, D.; Marlot, C.
-  \newblock Speed of processing in the human visual system.
-  \newblock {\em Nature} {\bf 1996}, {\em 381},~520--522.
-  \newblock {\url{https://doi.org/10.1038/381520a0}}.
-
-  \bibitem[Kirchner and Thorpe(2006)]{kirchner_ultra-rapid_2006}
-  Kirchner, H.; Thorpe, S.
-  \newblock Ultra-rapid object detection with saccadic eye movements: {Visual}
-    processing speed revisited.
-  \newblock {\em Vision Research} {\bf 2006}, {\em 46},~1762--76.
-  \newblock Loaded from an external bibliography file by Manubot.
-    source\_bibliography: manual-references.bib standard\_id: Kirchner06,
-    {\url{https://doi.org/10.1016/j.visres.2005.10.002}}.
-
-  \bibitem[Keysers \em{et~al.}(2001)Keysers, Xiao, Földiák, and
-    Perrett]{keysers_speed_2001}
-  Keysers, C.; Xiao, D.K.; Földiák, P.; Perrett, D.I.
-  \newblock The {Speed} of {Sight}.
-  \newblock {\em Journal of Cognitive Neuroscience} {\bf 2001}, {\em
-    13},~90--101.
-  \newblock {\url{https://doi.org/10.1162/089892901564199}}.
-
-  \bibitem[Schmolesky \em{et~al.}(1998)Schmolesky, Wang, Hanes, Thompson,
-    Leutgeb, Schall, and Leventhal]{schmolesky_signal_1998}
-  Schmolesky, M.T.; Wang, Y.; Hanes, D.P.; Thompson, K.G.; Leutgeb, S.; Schall,
-    J.D.; Leventhal, A.G.
-  \newblock Signal timing across the macaque visual system.
-  \newblock {\em Journal of Neurophysiology} {\bf 1998}, {\em 79},~3272--3278.
-  \newblock {\url{https://doi.org/10.1152/jn.1998.79.6.3272}}.
-
-  \bibitem[Vanni \em{et~al.}(2001)Vanni, Tanskanen, Seppä, Uutela, and
-    Hari]{vanni_coinciding_2001}
-  Vanni, S.; Tanskanen, T.; Seppä, M.; Uutela, K.; Hari, R.
-  \newblock Coinciding early activation of the human primary visual cortex and
-    anteromedial cuneus.
-  \newblock {\em Proceedings of the National Academy of Sciences of the United
-    States of America} {\bf 2001}, {\em 98},~2776--2780.
-  \newblock {\url{https://doi.org/10.1073/pnas.041600898}}.
-
-  \bibitem[Lamme and Roelfsema(2000)]{lamme_distinct_2000}
-  Lamme, V.A.; Roelfsema, P.R.
-  \newblock The distinct modes of vision offered by feedforward and recurrent
-    processing.
-  \newblock {\em Trends in Neurosciences} {\bf 2000}, {\em 23},~571--579.
-  \newblock {\url{https://doi.org/10.1016/s0166-2236(00)01657-x}}.
-
-  \bibitem[Serre \em{et~al.}(2007)Serre, Oliva, and
-    Poggio]{serre_feedforward_2007}
-  Serre, T.; Oliva, A.; Poggio, T.
-  \newblock A feedforward architecture accounts for rapid categorization.
-  \newblock {\em Proceedings of the National Academy of Sciences of the United
-    States of America} {\bf 2007}, {\em 104},~6424--6429.
-  \newblock tex.bdsk-url-2: https://doi.org/10.1073/pnas.0700622104
-    tex.date-added: 2022-05-05 19:08:15 +0200 tex.date-modified: 2022-05-05
-    19:08:15 +0200, {\url{https://doi.org/10.1073/pnas.0700622104}}.
-
-  \bibitem[Jérémie and Perrinet(2022)]{jeremie_ultrafast_2022}
-  Jérémie, J.N.; Perrinet, L.U.
-  \newblock Ultrafast image categorization in vivo and in silico,  2022.
-  \newblock arXiv:2205.03635 [cs, q-bio].
-
-  \bibitem[Nowak and Bullier(1997)]{nowak_timing_1997}
-  Nowak, L.G.; Bullier, J.
-  \newblock The {Timing} of {Information} {Transfer} in the {Visual} {System}. In
-    {\em Extrastriate {Cortex} in {Primates}}; Springer US,  1997; pp. 205--241.
-
-  \bibitem[Thorpe and Fabre-Thorpe(2001)]{thorpe_seeking_2001}
-  Thorpe, S.J.; Fabre-Thorpe, M.
-  \newblock Seeking {Categories} in the {Brain}.
-  \newblock {\em Science} {\bf 2001}, {\em 291},~260--263.
-  \newblock {\url{https://doi.org/10.1126/science.1058249}}.
-
-  \bibitem[Rucci \em{et~al.}(2018)Rucci, Ahissar, and Burr]{rucci_temporal_2018}
-  Rucci, M.; Ahissar, E.; Burr, D.
-  \newblock Temporal {Coding} of {Visual} {Space}.
-  \newblock {\em Trends in Cognitive Sciences} {\bf 2018}, {\em 22},~883--895.
-  \newblock {\url{https://doi.org/10.1016/j.tics.2018.07.009}}.
-
-  \bibitem[Softky and Koch(1993)]{softky_highly_1993}
-  Softky, W.; Koch, C.
-  \newblock The highly irregular firing of cortical cells is inconsistent with
-    temporal integration of random {EPSPs}.
-  \newblock {\em The Journal of Neuroscience} {\bf 1993}, {\em 13},~334--350.
-  \newblock {\url{https://doi.org/10.1523/jneurosci.13-01-00334.1993}}.
-
-  \bibitem[Bryant and Segundo(1976)]{bryant_spike_1976}
-  Bryant, H.L.; Segundo, J.P.
-  \newblock Spike initiation by transmembrane current: a white-noise analysis.
-  \newblock {\em The Journal of Physiology} {\bf 1976}, {\em 260},~279--314.
-  \newblock \_eprint:
-    https://onlinelibrary.wiley.com/doi/pdf/10.1113/jphysiol.1976.sp011516,
-    {\url{https://doi.org/10.1113/jphysiol.1976.sp011516}}.
-
-  \bibitem[Mainen and Sejnowski(1995)]{mainen_reliability_1995}
-  Mainen, Z.F.; Sejnowski, T.J.
-  \newblock Reliability of {Spike} {Timing} in {Neocortical} {Neurons}.
-  \newblock {\em Science} {\bf 1995}, {\em 268},~1503--1506.
-  \newblock {\url{https://doi.org/10.1126/science.7770778}}.
-
-  \bibitem[Ermentrout \em{et~al.}(2008)Ermentrout, Galán, and
-    Urban]{ermentrout_reliability_2008}
-  Ermentrout, G.B.; Galán, R.F.; Urban, N.N.
-  \newblock Reliability, synchrony and noise.
-  \newblock {\em Trends in neurosciences} {\bf 2008}, {\em 31},~428--434.
-  \newblock {\url{https://doi.org/10.1016/j.tins.2008.06.002}}.
-
-  \bibitem[Nowak(1997)]{nowak_influence_1997}
-  Nowak, L.
-  \newblock Influence of low and high frequency inputs on spike timing in visual
-    cortical neurons.
-  \newblock {\em Cerebral Cortex} {\bf 1997}, {\em 7},~487--501.
-  \newblock {\url{https://doi.org/10.1093/cercor/7.6.487}}.
-
-  \bibitem[Kenyon \em{et~al.}(2004)Kenyon, Hill, Theiler, George, and
-    Marshak]{kenyon_theory_2004}
-  Kenyon, G.T.; Hill, D.; Theiler, J.; George, J.S.; Marshak, D.W.
-  \newblock A theory of the {Benham} {Top} based on center–surround
-    interactions in the parvocellular pathway.
-  \newblock {\em Neural Networks} {\bf 2004}, {\em 17},~773--786.
-  \newblock {\url{https://doi.org/10.1016/j.neunet.2004.05.005}}.
-
-  \bibitem[Celebrini \em{et~al.}(1993)Celebrini, Thorpe, Trotter, and
-    Imbert]{celebrini_dynamics_1993}
-  Celebrini, S.; Thorpe, S.; Trotter, Y.; Imbert, M.
-  \newblock Dynamics of orientation coding in area {V1} of the awake primate.
-  \newblock {\em Visual Neuroscience} {\bf 1993}, {\em 10},~811--825.
-  \newblock {\url{https://doi.org/10.1017/s0952523800006052}}.
-
-  \bibitem[Chase and Young(2007)]{chase_first-spike_2007}
-  Chase, S.M.; Young, E.D.
-  \newblock First-spike latency information in single neurons increases when
-    referenced to population onset.
-  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2007},
-    {\em 104},~5175--5180.
-  \newblock {\url{https://doi.org/10.1073/pnas.0610368104}}.
-
-  \bibitem[Safaie \em{et~al.}(2020)Safaie, Jurado-Parras, Sarno, Louis,
-    Karoutchi, Petit, Pasquet, Eloy, and Robbe]{safaie_turning_2020}
-  Safaie, M.; Jurado-Parras, M.T.; Sarno, S.; Louis, J.; Karoutchi, C.; Petit,
-    L.F.; Pasquet, M.O.; Eloy, C.; Robbe, D.
-  \newblock Turning the body into a clock: {Accurate} timing is facilitated by
-    simple stereotyped interactions with the environment.
-  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2020},
-    {\em 117},~13084--13093.
-  \newblock {\url{https://doi.org/10.1073/pnas.1921226117}}.
-
-  \bibitem[Gautrais and Thorpe(1998)]{gautrais_rate_1998}
-  Gautrais, J.; Thorpe, S.
-  \newblock Rate coding versus temporal order coding: a theoretical approach.
-  \newblock {\em Biosystems} {\bf 1998}, {\em 48},~57--65.
-  \newblock Publisher: Elsevier,
-    {\url{https://doi.org/10.1016/S0303-2647(98)00050-1}}.
-
-  \bibitem[Delorme \em{et~al.}(1999)Delorme, Gautrais, Van~Rullen, and
-    Thorpe]{delorme_spikenet_1999}
-  Delorme, A.; Gautrais, J.; Van~Rullen, R.; Thorpe, S.
-  \newblock {SpikeNET}: {A} simulator for modeling large networks of integrate
-    and fire neurons.
-  \newblock {\em Neurocomputing} {\bf 1999}, {\em 26},~989--996.
-  \newblock Publisher: Elsevier,
-    {\url{https://doi.org/10.1016/S0925-2312(99)00095-8}}.
-
-  \bibitem[Delorme \em{et~al.}(2000)Delorme, Richard, and
-    Fabre-Thorpe]{delorme_ultra-rapid_2000}
-  Delorme, A.; Richard, G.; Fabre-Thorpe, M.
-  \newblock Ultra-rapid categorisation of natural scenes does not rely on colour
-    cues: a study in monkeys and humans.
-  \newblock {\em Vision Research} {\bf 2000}, {\em 40},~2187--2200.
-  \newblock {\url{https://doi.org/10.1016/S0042-6989(00)00083-3}}.
-
-  \bibitem[Bonilla \em{et~al.}(2022)Bonilla, Gautrais, Thorpe, and
-    Masquelier]{bonilla_analyzing_2022}
-  Bonilla, L.; Gautrais, J.; Thorpe, S.; Masquelier, T.
-  \newblock Analyzing time-to-first-spike coding schemes.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2022}, {\em 16}.
-  \newblock Publisher: Frontiers,
-    {\url{https://doi.org/10.3389/fnins.2022.971937}}.
-
-  \bibitem[Bohte \em{et~al.}(2002)Bohte, Kok, and
-    La~Poutré]{bohte_error-backpropagation_2002}
-  Bohte, S.M.; Kok, J.N.; La~Poutré, H.
-  \newblock Error-backpropagation in temporally encoded networks of spiking
-    neurons.
-  \newblock {\em Neurocomputing} {\bf 2002}, {\em 48},~17--37.
-  \newblock {\url{https://doi.org/10.1016/S0925-2312(01)00658-0}}.
-
-  \bibitem[Zenke and Vogels(2021)]{zenke_remarkable_2021}
-  Zenke, F.; Vogels, T.P.
-  \newblock The {Remarkable} {Robustness} of {Surrogate} {Gradient} {Learning}
-    for {Instilling} {Complex} {Function} in {Spiking} {Neural} {Networks}.
-  \newblock {\em Neural Computation} {\bf 2021}, {\em 33},~899--925.
-  \newblock {\url{https://doi.org/10.1162/neco_a_01367}}.
-
-  \bibitem[Göltz \em{et~al.}(2021)Göltz, Kriener, Baumbach, Billaudelle,
-    Breitwieser, Cramer, Dold, Kungl, Senn, Schemmel, Meier, and
-    Petrovici]{goltz_fast_2021}
-  Göltz, J.; Kriener, L.; Baumbach, A.; Billaudelle, S.; Breitwieser, O.;
-    Cramer, B.; Dold, D.; Kungl, A.F.; Senn, W.; Schemmel, J.;  et~al.
-  \newblock Fast and energy-efficient neuromorphic deep learning with first-spike
-    times.
-  \newblock Technical Report 1912.11443, arXiv,  2021.
-
-  \bibitem[Kheradpisheh \em{et~al.}(2018)Kheradpisheh, Ganjtabesh, Thorpe, and
-    Masquelier]{kheradpisheh_stdp-based_2018}
-  Kheradpisheh, S.R.; Ganjtabesh, M.; Thorpe, S.J.; Masquelier, T.
-  \newblock {STDP}-based spiking deep convolutional neural networks for object
-    recognition.
-  \newblock {\em Neural Networks} {\bf 2018}, {\em 99},~56--67.
-  \newblock {\url{https://doi.org/10.1016/j.neunet.2017.12.005}}.
-
-  \bibitem[Tavanaei \em{et~al.}(2018)Tavanaei, Masquelier, and
-    Maida]{tavanaei_representation_2018}
-  Tavanaei, A.; Masquelier, T.; Maida, A.
-  \newblock Representation learning using event-based {STDP}.
-  \newblock {\em Neural Networks} {\bf 2018}, {\em 105},~294--303.
-  \newblock {\url{https://doi.org/10.1016/j.neunet.2018.05.018}}.
-
-  \bibitem[Gallego \em{et~al.}(2022)Gallego, Delbruck, Orchard, Bartolozzi, Taba,
-    Censi, Leutenegger, Davison, Conradt, Daniilidis, and
-    Scaramuzza]{gallego_event-based_2022}
-  Gallego, G.; Delbruck, T.; Orchard, G.; Bartolozzi, C.; Taba, B.; Censi, A.;
-    Leutenegger, S.; Davison, A.J.; Conradt, J.; Daniilidis, K.;  et~al.
-  \newblock Event-{Based} {Vision}: {A} {Survey}.
-  \newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}
-    {\bf 2022}, {\em 44},~154--180.
-  \newblock {\url{https://doi.org/10.1109/TPAMI.2020.3008413}}.
-
-  \bibitem[Maunsell and Van~Essen(1983)]{maunsell_functional_1983}
-  Maunsell, J.H.; Van~Essen, D.C.
-  \newblock Functional properties of neurons in middle temporal visual area of
-    the macaque monkey. {I}. {Selectivity} for stimulus direction, speed, and
-    orientation.
-  \newblock {\em Journal of Neurophysiology} {\bf 1983}, {\em 49},~1127--1147.
-  \newblock {\url{https://doi.org/10.1152/jn.1983.49.5.1127}}.
-
-  \bibitem[Montemurro \em{et~al.}(2008)Montemurro, Rasch, Murayama, Logothetis,
-    and Panzeri]{montemurro_phase--firing_2008}
-  Montemurro, M.A.; Rasch, M.J.; Murayama, Y.; Logothetis, N.K.; Panzeri, S.
-  \newblock Phase-of-{Firing} {Coding} of {Natural} {Visual} {Stimuli} in
-    {Primary} {Visual} {Cortex}.
-  \newblock {\em Current Biology} {\bf 2008}, {\em 18},~375--380.
-  \newblock Publisher: Elsevier,
-    {\url{https://doi.org/10.1016/j.cub.2008.02.023}}.
-
-  \bibitem[deCharms and Merzenich(1996)]{decharms_primary_1996}
-  deCharms, R.C.; Merzenich, M.M.
-  \newblock Primary cortical representation of sounds by the coordination of
-    action-potential timing.
-  \newblock {\em Nature} {\bf 1996}, {\em 381},~610--613.
-  \newblock Number: 6583 Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/381610a0}}.
-
-  \bibitem[Vinje and Gallant(2000)]{vinje_sparse_2000}
-  Vinje, W.E.; Gallant, J.L.
-  \newblock Sparse {Coding} and {Decorrelation} in {Primary} {Visual} {Cortex}
-    {During} {Natural} {Vision}.
-  \newblock {\em Science} {\bf 2000}, {\em 287},~1273--1276.
-  \newblock Publisher: American Association for the Advancement of Science,
-    {\url{https://doi.org/10.1126/science.287.5456.1273}}.
-
-  \bibitem[Abeles(1991)]{abeles_corticonics_1991}
-  Abeles, M.
-  \newblock {\em Corticonics: neural circuits of the cerebral cortex}; Cambridge
-    University Press: Cambridge ; New York,  1991.
-
-  \bibitem[Paugam-Moisy and Bohte(2012)]{paugam-moisy_computing_2012}
-  Paugam-Moisy, H.; Bohte, S.M.
-  \newblock Computing with spiking neuron networks. In {\em Handbook of natural
-    computing}; Springer-Verlag,  2012.
-
-  \bibitem[Hebb(1949)]{hebb_organization_1949}
-  Hebb, D.O.
-  \newblock {\em The organization of behavior: {A} neuropsychological theory};
-    Wiley: New York,  1949.
-
-  \bibitem[Harris \em{et~al.}(2003)Harris, Csicsvari, Hirase, Dragoi, and
-    Buzsáki]{harris_organization_2003}
-  Harris, K.D.; Csicsvari, J.; Hirase, H.; Dragoi, G.; Buzsáki, G.
-  \newblock Organization of cell assemblies in the hippocampus.
-  \newblock {\em Nature} {\bf 2003}, {\em 424},~552--556.
-  \newblock {\url{https://doi.org/10.1038/nature01834}}.
-
-  \bibitem[Singer and Gray(1995)]{singer_visual_1995}
-  Singer, W.; Gray, C.M.
-  \newblock Visual {Feature} {Integration} and the {Temporal} {Correlation}
-    {Hypothesis}.
-  \newblock {\em Annual Review of Neuroscience} {\bf 1995}, {\em 18},~555--586.
-  \newblock {\url{https://doi.org/10.1146/annurev.ne.18.030195.003011}}.
-
-  \bibitem[Roelfsema \em{et~al.}(1997)Roelfsema, Engel, König, and
-    Singer]{roelfsema_visuomotor_1997}
-  Roelfsema, P.R.; Engel, A.K.; König, P.; Singer, W.
-  \newblock Visuomotor integration is associated with zero time-lag
-    synchronization among cortical areas.
-  \newblock {\em Nature} {\bf 1997}, {\em 385},~157--161.
-  \newblock {\url{https://doi.org/10.1038/385157a0}}.
-
-  \bibitem[Riehle \em{et~al.}(1997)Riehle, Grun, Diesmann, and
-    Aertsen]{riehle_spike_1997}
-  Riehle, A.; Grun, S.; Diesmann, M.; Aertsen, A.
-  \newblock Spike synchronization and rate modulation differentially involved in
-    motor cortical function.
-  \newblock {\em Science (New York, N.Y.)} {\bf 1997}, {\em 278},~1950--1953.
-  \newblock Publisher: American Association for the Advancement of Science,
-    {\url{https://doi.org/10.1126/science.278.5345.1950}}.
-
-  \bibitem[Kilavik \em{et~al.}(2009)Kilavik, Roux, Ponce-Alvarez, Confais, Grun,
-    and Riehle]{kilavik_long-term_2009}
-  Kilavik, B.E.; Roux, S.; Ponce-Alvarez, A.; Confais, J.; Grun, S.; Riehle, A.
-  \newblock Long-{Term} {Modifications} in {Motor} {Cortical} {Dynamics}
-    {Induced} by {Intensive} {Practice}.
-  \newblock {\em Journal of Neuroscience} {\bf 2009}, {\em 29},~12653--12663.
-  \newblock {\url{https://doi.org/10.1523/jneurosci.1554-09.2009}}.
-
-  \bibitem[Grammont and Riehle(2003)]{grammont_spike_2003}
-  Grammont, F.; Riehle, A.
-  \newblock Spike synchronization and firing rate in a population of motor
-    cortical neurons in relation to movement direction and reaction time.
-  \newblock {\em Biological Cybernetics} {\bf 2003}, {\em 88},~360--373.
-  \newblock {\url{https://doi.org/10.1007/s00422-002-0385-3}}.
-
-  \bibitem[Denker \em{et~al.}(2018)Denker, Zehl, Kilavik, Diesmann, Brochier,
-    Riehle, and Grün]{denker_lfp_2018}
-  Denker, M.; Zehl, L.; Kilavik, B.E.; Diesmann, M.; Brochier, T.; Riehle, A.;
-    Grün, S.
-  \newblock {LFP} beta amplitude is linked to mesoscopic spatio-temporal phase
-    patterns.
-  \newblock {\em Scientific Reports} {\bf 2018}, {\em 8}.
-  \newblock {\url{https://doi.org/10.1038/s41598-018-22990-7}}.
-
-  \bibitem[Torre \em{et~al.}(2016)Torre, Canova, Denker, Gerstein, Helias, and
-    Grün]{torre_asset_2016}
-  Torre, E.; Canova, C.; Denker, M.; Gerstein, G.; Helias, M.; Grün, S.
-  \newblock {ASSET}: {Analysis} of {Sequences} of {Synchronous} {Events} in
-    {Massively} {Parallel} {Spike} {Trains}.
-  \newblock {\em PLOS Computational Biology} {\bf 2016}, {\em 12},~e1004939.
-  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1004939}}.
-
-  \bibitem[Ben-yishai and Hansel(1997)]{ben-yishai_traveling_1997}
-  Ben-yishai, R.; Hansel, D.
-  \newblock Traveling {Waves} and the {Processing} of {Weakly} {Tuned} {Inputs}
-    in a {Cortical} {Network} {Module} {\bf 1997}.
-  \newblock {\em 77},~57--77.
-  \newblock 00000.
-
-  \bibitem[Bruno and Sakmann(2006)]{bruno_cortex_2006}
-  Bruno, R.M.; Sakmann, B.
-  \newblock Cortex {Is} {Driven} by {Weak} but {Synchronously} {Active}
-    {Thalamocortical} {Synapses}.
-  \newblock {\em Science} {\bf 2006}, {\em 312},~1622--1627.
-  \newblock {\url{https://doi.org/10.1126/science.1124593}}.
-
-  \bibitem[Deneve(2004)]{deneve_bayesian_2004}
-  Deneve, S.
-  \newblock Bayesian inference in spiking neurons.
-  \newblock In Proceedings of the Advances in {Neural} {Information} {Processing}
-    {Systems}. MIT Press,  2004, Vol.~17.
-
-  \bibitem[Ballard and Jehee(2011)]{ballard_dual_2011}
-  Ballard, D.; Jehee, J.
-  \newblock Dual {Roles} for {Spike} {Signaling} in {Cortical} {Neural}
-    {Populations}.
-  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2011}, {\em 5}.
-
-  \bibitem[Gewaltig \em{et~al.}(2001)Gewaltig, Diesmann, and
-    Aertsen]{gewaltig_propagation_2001}
-  Gewaltig, M.O.; Diesmann, M.; Aertsen, A.
-  \newblock Propagation of cortical synfire activity: survival probability in
-    single trials and stability in the mean.
-  \newblock {\em Neural Networks} {\bf 2001}, {\em 14},~657--673.
-  \newblock {\url{https://doi.org/10.1016/S0893-6080(01)00070-3}}.
-
-  \bibitem[Gerstner(1995)]{gerstner_time_1995}
-  Gerstner, W.
-  \newblock Time structure of the activity in neural network models.
-  \newblock {\em Physical Review E} {\bf 1995}, {\em 51},~738--758.
-  \newblock {\url{https://doi.org/10.1103/physreve.51.738}}.
-
-  \bibitem[Azouz and Gray(2008)]{azouz_stimulus-selective_2008}
-  Azouz, R.; Gray, C.M.
-  \newblock Stimulus-selective spiking is driven by the relative timing of
-    synchronous excitation and disinhibition in cat striate neurons\textit{in
-    vivo}.
-  \newblock {\em European Journal of Neuroscience} {\bf 2008}, {\em
-    28},~1286--1300.
-  \newblock {\url{https://doi.org/10.1111/j.1460-9568.2008.06434.x}}.
-
-  \bibitem[Kremkow \em{et~al.}(2016)Kremkow, Perrinet, Monier, Alonso, Aertsen,
-    Frégnac, and Masson]{kremkow_push-pull_2016}
-  Kremkow, J.; Perrinet, L.U.; Monier, C.; Alonso, J.M.; Aertsen, A.; Frégnac,
-    Y.; Masson, G.S.
-  \newblock Push-{Pull} {Receptive} {Field} {Organization} and {Synaptic}
-    {Depression}: {Mechanisms} for {Reliably} {Encoding} {Naturalistic} {Stimuli}
-    in {V1}.
-  \newblock {\em Frontiers in Neural Circuits} {\bf 2016}, {\em 10}.
-  \newblock {\url{https://doi.org/10.3389/fncir.2016.00037}}.
-
-  \bibitem[Aviel \em{et~al.}(2003)Aviel, Mehring, Abeles, and
-    Horn]{aviel_embedding_2003}
-  Aviel, Y.; Mehring, C.; Abeles, M.; Horn, D.
-  \newblock On {Embedding} {Synfire} {Chains} in a {Balanced} {Network}.
-  \newblock {\em Neural Computation} {\bf 2003}, {\em 15},~1321--1340.
-  \newblock {\url{https://doi.org/10.1162/089976603321780290}}.
-
-  \bibitem[Kremkow \em{et~al.}(2010)Kremkow, Perrinet, Masson, and
-    Aertsen]{kremkow_functional_2010}
-  Kremkow, J.; Perrinet, L.U.; Masson, G.S.; Aertsen, A.
-  \newblock Functional consequences of correlated excitatory and inhibitory
-    conductances in cortical networks.
-  \newblock {\em Journal of Computational Neuroscience} {\bf 2010}, {\em
-    28},~579--594.
-  \newblock {\url{https://doi.org/10.1007/s10827-010-0240-9}}.
-
-  \bibitem[Davison(2008)]{davison_pynn_2008}
-  Davison, A.P.
-  \newblock {PyNN}: a common interface for neuronal network simulators.
-  \newblock {\em Frontiers in Neuroinformatics} {\bf 2008}, {\em 2}.
-  \newblock {\url{https://doi.org/10.3389/neuro.11.011.2008}}.
-
-  \bibitem[Pfeil \em{et~al.}(2013)Pfeil, Grübl, Jeltsch, Müller, Müller,
-    Petrovici, Schmuker, Brüderle, Schemmel, and Meier]{pfeil_six_2013}
-  Pfeil, T.; Grübl, A.; Jeltsch, S.; Müller, E.; Müller, P.; Petrovici, M.A.;
-    Schmuker, M.; Brüderle, D.; Schemmel, J.; Meier, K.
-  \newblock Six {Networks} on a {Universal} {Neuromorphic} {Computing}
-    {Substrate}.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2013}, {\em 7}.
-  \newblock {\url{https://doi.org/10.3389/fnins.2013.00011}}.
-
-  \bibitem[Schrader \em{et~al.}(2008)Schrader, Grün, Diesmann, and
-    Gerstein]{schrader_detecting_2008}
-  Schrader, S.; Grün, S.; Diesmann, M.; Gerstein, G.L.
-  \newblock Detecting {Synfire} {Chain} {Activity} {Using} {Massively} {Parallel}
-    {Spike} {Train} {Recording}.
-  \newblock {\em Journal of Neurophysiology} {\bf 2008}, {\em 100},~2165--2176.
-  \newblock {\url{https://doi.org/10.1152/jn.01245.2007}}.
-
-  \bibitem[Grammont and Riehle(1999)]{grammont_precise_1999}
-  Grammont, F.; Riehle, A.
-  \newblock Precise spike synchronization in monkey motor cortex involved in
-    preparation for movement.
-  \newblock {\em Experimental Brain Research} {\bf 1999}, {\em 128},~118--122.
-  \newblock {\url{https://doi.org/10.1007/s002210050826}}.
-
-  \bibitem[Brette(2012)]{brette_computing_2012}
-  Brette, R.
-  \newblock Computing with {Neural} {Synchrony}.
-  \newblock {\em PLoS Computational Biology} {\bf 2012}, {\em 8},~e1002561.
-  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1002561}}.
-
-  \bibitem[Fries(2005)]{fries_mechanism_2005}
-  Fries, P.
-  \newblock A mechanism for cognitive dynamics: neuronal communication through
-    neuronal coherence.
-  \newblock {\em Trends in Cognitive Sciences} {\bf 2005}, {\em 9},~474--480.
-  \newblock {\url{https://doi.org/10.1016/j.tics.2005.08.011}}.
-
-  \bibitem[VanRullen \em{et~al.}(2006)VanRullen, Reddy, and
-    Koch]{vanrullen_continuous_2006}
-  VanRullen, R.; Reddy, L.; Koch, C.
-  \newblock The {Continuous} {Wagon} {Wheel} {Illusion} {Is} {Associated} with
-    {Changes} in {Electroencephalogram} {Power} at 13 {Hz}.
-  \newblock {\em Journal of Neuroscience} {\bf 2006}, {\em 26},~502--507.
-  \newblock 00000, {\url{https://doi.org/10.1523/jneurosci.4654-05.2006}}.
-
-  \bibitem[Dugue \em{et~al.}(2011)Dugue, Marque, and VanRullen]{dugue_phase_2011}
-  Dugue, L.; Marque, P.; VanRullen, R.
-  \newblock The {Phase} of {Ongoing} {Oscillations} {Mediates} the {Causal}
-    {Relation} between {Brain} {Excitation} and {Visual} {Perception}.
-  \newblock {\em Journal of Neuroscience} {\bf 2011}, {\em 31},~11889--11893.
-  \newblock {\url{https://doi.org/10.1523/JNEUROSCI.1161-11.2011}}.
-
-  \bibitem[Bringuier \em{et~al.}(1999)Bringuier, Chavane, Glaeser, and
-    Frégnac]{bringuier_horizontal_1999}
-  Bringuier, V.; Chavane, F.; Glaeser, L.; Frégnac, Y.
-  \newblock Horizontal {Propagation} of {Visual} {Activity} in the {Synaptic}
-    {Integration} {Field} of {Area} 17 {Neurons}.
-  \newblock {\em Science} {\bf 1999}, {\em 283},~695--699.
-  \newblock 00535 Loaded from an external bibliography file by Manubot.
-    source\_bibliography: manual-references.bib standard\_id: Bringuier99,
-    {\url{https://doi.org/10.1126/science.283.5402.695}}.
-
-  \bibitem[Benvenuti \em{et~al.}(2020)Benvenuti, Chemla, Boonman, Perrinet,
-    Masson, and Chavane]{benvenuti_anticipatory_2020}
-  Benvenuti, G.; Chemla, S.; Boonman, A.; Perrinet, L.U.; Masson, G.S.; Chavane,
-    F.
-  \newblock Anticipatory responses along motion trajectories in awake monkey area
-    {V1}.
-  \newblock {\em bioRxiv : the preprint server for biology} {\bf 2020}, p.
-    2020.03.26.010017.
-  \newblock {\url{https://doi.org/10.1101/2020.03.26.010017}}.
-
-  \bibitem[Le~Bec \em{et~al.}(2022)Le~Bec, Troncoso, Desbois, Passarelli, Baudot,
-    Monier, Pananceau, and Frégnac]{le_bec_horizontal_2022}
-  Le~Bec, B.; Troncoso, X.G.; Desbois, C.; Passarelli, Y.; Baudot, P.; Monier,
-    C.; Pananceau, M.; Frégnac, Y.
-  \newblock Horizontal connectivity in {V1}: {Prediction} of coherence in contour
-    and motion integration.
-  \newblock {\em PLOS ONE} {\bf 2022}, {\em 17},~e0268351.
-  \newblock {\url{https://doi.org/10.1371/journal.pone.0268351}}.
-
-  \bibitem[Feller \em{et~al.}(1997)Feller, Butts, Aaron, Rokhsar, and
-    Shatz]{feller_dynamic_1997}
-  Feller, M.B.; Butts, D.A.; Aaron, H.L.; Rokhsar, D.S.; Shatz, C.J.
-  \newblock Dynamic {Processes} {Shape} {Spatiotemporal} {Properties} of
-    {Retinal} {Waves}.
-  \newblock {\em Neuron} {\bf 1997}, {\em 19},~293--306.
-  \newblock {\url{https://doi.org/10.1016/S0896-6273(00)80940-X}}.
-
-  \bibitem[Bienenstock(1995)]{bienenstock_model_1995}
-  Bienenstock, E.
-  \newblock A model of neocortex.
-  \newblock {\em Network: Computation in Neural Systems} {\bf 1995}, {\em
-    6},~179--224.
-  \newblock Publisher: Taylor \& Francis \_eprint:
-    https://doi.org/10.1088/0954-898X\_6\_2\_004,
-    {\url{https://doi.org/10.1088/0954-898X_6_2_004}}.
-
-  \bibitem[Muller \em{et~al.}(2014)Muller, Reynaud, Chavane, and
-    Destexhe]{muller_stimulus-evoked_2014}
-  Muller, L.; Reynaud, A.; Chavane, F.; Destexhe, A.
-  \newblock The stimulus-evoked population response in visual cortex of awake
-    monkey is a propagating wave.
-  \newblock {\em Nature Communications} {\bf 2014}, {\em 5},~3675.
-  \newblock 00068 Loaded from an external bibliography file by Manubot.
-    source\_bibliography: manual-references.bib standard\_id: Muller14,
-    {\url{https://doi.org/10.1038/ncomms4675}}.
-
-  \bibitem[Muller \em{et~al.}(2018)Muller, Chavane, Reynolds, and
-    Sejnowski]{muller_cortical_2018}
-  Muller, L.; Chavane, F.; Reynolds, J.; Sejnowski, T.J.
-  \newblock Cortical travelling waves: {Mechanisms} and computational principles.
-  \newblock {\em Nature Reviews Neuroscience} {\bf 2018}.
-  \newblock Loaded from an external bibliography file by Manubot.
-    source\_bibliography: manual-references.bib standard\_id: Muller18,
-    {\url{https://doi.org/10.1038/nrn.2018.20}}.
-
-  \bibitem[Lindén \em{et~al.}(2022)Lindén, Petersen, Vestergaard, and
-    Berg]{linden_movement_2022}
-  Lindén, H.; Petersen, P.C.; Vestergaard, M.; Berg, R.W.
-  \newblock Movement is governed by rotational neural dynamics in spinal motor
-    networks.
-  \newblock {\em Nature} {\bf 2022}, {\em 610},~526--531.
-  \newblock {\url{https://doi.org/10.1038/s41586-022-05293-w}}.
-
-  \bibitem[Chemla \em{et~al.}(2019)Chemla, Reynaud, diVolo, Zerlaut, Perrinet,
-    Destexhe, and Chavane]{chemla_suppressive_2019}
-  Chemla, S.; Reynaud, A.; diVolo, M.; Zerlaut, Y.; Perrinet, L.U.; Destexhe, A.;
-    Chavane, F.Y.
-  \newblock Suppressive waves disambiguate the representation of long-range
-    apparent motion in awake monkey {V1}.
-  \newblock {\em Journal of Neuroscience} {\bf 2019}, {\em 2792},~18.
-  \newblock Loaded from an external bibliography file by Manubot.
-    source\_bibliography: manual-references.bib standard\_id: Chemla19,
-    {\url{https://doi.org/10.1523/JNEUROSCI.2792-18.2019}}.
-
-  \bibitem[Pillow \em{et~al.}(2008)Pillow, Shlens, Paninski, Sher, Litke,
-    Chichilnisky, and Simoncelli]{pillow_spatio-temporal_2008}
-  Pillow, J.W.; Shlens, J.; Paninski, L.; Sher, A.; Litke, A.M.; Chichilnisky,
-    E.J.; Simoncelli, E.P.
-  \newblock Spatio-temporal correlations and visual signalling in a complete
-    neuronal population.
-  \newblock {\em Nature} {\bf 2008}, {\em 454},~995--999.
-  \newblock {\url{https://doi.org/10.1038/nature07140}}.
-
-  \bibitem[Schneidman \em{et~al.}(2006)Schneidman, Berry, Segev, and
-    Bialek]{schneidman_weak_2006}
-  Schneidman, E.; Berry, M.J.; Segev, R.; Bialek, W.
-  \newblock Weak pairwise correlations imply strongly correlated network states
-    in a neural population.
-  \newblock {\em Nature} {\bf 2006}, {\em 440},~1007--1012.
-  \newblock {\url{https://doi.org/10.1038/nature04701}}.
-
-  \bibitem[Berry(2022)]{berry_spike_2022}
-  Berry, M.J.
-  \newblock Spike {Trains} of {Retinal} {Ganglion} {Cells} {Viewing} a {Repeated}
-    {Natural} {Movie} {\bf 2022}.
-  \newblock {\url{https://doi.org/10.34770/V0V4-3H52}}.
-
-  \bibitem[Miller \em{et~al.}(2014)Miller, Ayzenshtat, Carrillo-Reid, and
-    Yuste]{miller_visual_2014}
-  Miller, J.e.K.; Ayzenshtat, I.; Carrillo-Reid, L.; Yuste, R.
-  \newblock Visual stimuli recruit intrinsically generated cortical ensembles.
-  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2014},
-    {\em 111}.
-  \newblock {\url{https://doi.org/10.1073/pnas.1406077111}}.
-
-  \bibitem[Ikegaya \em{et~al.}(2004)Ikegaya, Aaron, Cossart, Aronov, Lampl,
-    Ferster, and Yuste]{ikegaya_synfire_2004}
-  Ikegaya, Y.; Aaron, G.; Cossart, R.; Aronov, D.; Lampl, I.; Ferster, D.; Yuste,
-    R.
-  \newblock Synfire {Chains} and {Cortical} {Songs}: {Temporal} {Modules} of
-    {Cortical} {Activity}.
-  \newblock {\em Science} {\bf 2004}, {\em 304},~559--564.
-  \newblock {\url{https://doi.org/10/djckcn}}.
-
-  \bibitem[Luczak \em{et~al.}(2007)Luczak, Barthó, Marguet, Buzsáki, and
-    Harris]{luczak_sequential_2007}
-  Luczak, A.; Barthó, P.; Marguet, S.L.; Buzsáki, G.; Harris, K.D.
-  \newblock Sequential structure of neocortical spontaneous activity in vivo.
-  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2007},
-    {\em 104},~347--352.
-  \newblock {\url{https://doi.org/10.1073/pnas.0605643104}}.
-
-  \bibitem[Pastalkova \em{et~al.}(2008)Pastalkova, Itskov, Amarasingham, and
-    Buzsáki]{pastalkova_internally_2008}
-  Pastalkova, E.; Itskov, V.; Amarasingham, A.; Buzsáki, G.
-  \newblock Internally {Generated} {Cell} {Assembly} {Sequences} in the {Rat}
-    {Hippocampus}.
-  \newblock {\em Science (New York, N.Y.)} {\bf 2008}, {\em 321},~1322--1327.
-  \newblock {\url{https://doi.org/10.1126/science.1159775}}.
-
-  \bibitem[Villette \em{et~al.}(2015)Villette, Malvache, Tressard, Dupuy, and
-    Cossart]{villette_internally_2015}
-  Villette, V.; Malvache, A.; Tressard, T.; Dupuy, N.; Cossart, R.
-  \newblock Internally {Recurring} {Hippocampal} {Sequences} as a {Population}
-    {Template} of {Spatiotemporal} {Information}.
-  \newblock {\em Neuron} {\bf 2015}, {\em 88},~357--366.
-  \newblock 00085, {\url{https://doi.org/10/f7whnn}}.
-
-  \bibitem[Branco \em{et~al.}(2010)Branco, Clark, and
-    Häusser]{branco_dendritic_2010}
-  Branco, T.; Clark, B.A.; Häusser, M.
-  \newblock Dendritic {Discrimination} of {Temporal} {Input} {Sequences} in
-    {Cortical} {Neurons}.
-  \newblock {\em Science} {\bf 2010}, {\em 329},~1671--1675.
-  \newblock {\url{https://doi.org/10.1126/science.1189664}}.
-
-  \bibitem[Luczak \em{et~al.}(2015)Luczak, McNaughton, and
-    Harris]{luczak_packet-based_2015}
-  Luczak, A.; McNaughton, B.L.; Harris, K.D.
-  \newblock Packet-based communication in the cortex.
-  \newblock {\em Nature reviews. Neuroscience} {\bf 2015}, {\em 16},~745--55.
-  \newblock {\url{https://doi.org/10.1038/nrn4026}}.
-
-  \bibitem[Meister \em{et~al.}(1995)Meister, Lagnado, and
-    Baylor]{meister_concerted_1995}
-  Meister, M.; Lagnado, L.; Baylor, D.A.
-  \newblock Concerted {Signaling} by {Retinal} {Ganglion} {Cells}.
-  \newblock {\em Science} {\bf 1995}, {\em 270},~1207--1210.
-  \newblock {\url{https://doi.org/10.1126/science.270.5239.1207}}.
-
-  \bibitem[Cleland(2014)]{cleland_construction_2014}
-  Cleland, T.A.
-  \newblock Construction of {Odor} {Representations} by {Olfactory} {Bulb}
-    {Microcircuits}. In {\em Progress in {Brain} {Research}}; Elsevier,  2014;
-    Vol. 208, pp. 177--203.
-  \newblock {\url{https://doi.org/10.1016/B978-0-444-63350-7.00007-3}}.
-
-  \bibitem[Kashiwadani \em{et~al.}(1999)Kashiwadani, Sasaki, Uchida, and
-    Mori]{kashiwadani_synchronized_1999}
-  Kashiwadani, H.; Sasaki, Y.F.; Uchida, N.; Mori, K.
-  \newblock Synchronized {Oscillatory} {Discharges} of {Mitral}/{Tufted} {Cells}
-    {With} {Different} {Molecular} {Receptive} {Ranges} in the {Rabbit}
-    {Olfactory} {Bulb}.
-  \newblock {\em Journal of Neurophysiology} {\bf 1999}, {\em 82},~1786--1792.
-  \newblock Publisher: American Physiological Society,
-    {\url{https://doi.org/10.1152/jn.1999.82.4.1786}}.
-
-  \bibitem[Rinberg \em{et~al.}(2006)Rinberg, Koulakov, and
-    Gelperin]{rinberg_speed-accuracy_2006}
-  Rinberg, D.; Koulakov, A.; Gelperin, A.
-  \newblock Speed-{Accuracy} {Tradeoff} in {Olfaction}.
-  \newblock {\em Neuron} {\bf 2006}, {\em 51},~351--358.
-  \newblock {\url{https://doi.org/10.1016/j.neuron.2006.07.013}}.
-
-  \bibitem[Johansson and Birznieks(2004)]{johansson_first_2004}
-  Johansson, R.S.; Birznieks, I.
-  \newblock First spikes in ensembles of human tactile afferents code complex
-    spatial fingertip events.
-  \newblock {\em Nature Neuroscience} {\bf 2004}, {\em 7},~170--177.
-  \newblock {\url{https://doi.org/10.1038/nn1177}}.
-
-  \bibitem[Buzsáki and Tingley(2018)]{buzsaki_space_2018}
-  Buzsáki, G.; Tingley, D.
-  \newblock Space and {Time}: {The} {Hippocampus} as a {Sequence} {Generator}.
-  \newblock {\em Trends in Cognitive Sciences} {\bf 2018}, {\em 22},~853--869.
-  \newblock {\url{https://doi.org/10.1016/j.tics.2018.07.006}}.
-
-  \bibitem[Malvache \em{et~al.}(2016)Malvache, Reichinnek, Villette, Haimerl, and
-    Cossart]{malvache_awake_2016}
-  Malvache, A.; Reichinnek, S.; Villette, V.; Haimerl, C.; Cossart, R.
-  \newblock Awake hippocampal reactivations project onto orthogonal neuronal
-    assemblies.
-  \newblock {\em Science (New York, N.Y.)} {\bf 2016}, {\em 353},~1280--1283.
-  \newblock 00105, {\url{https://doi.org/10/bqpq}}.
-
-  \bibitem[Haimerl \em{et~al.}(2019)Haimerl, Angulo-Garcia, Villette, Reichinnek,
-    Torcini, Cossart, and Malvache]{haimerl_internal_2019}
-  Haimerl, C.; Angulo-Garcia, D.; Villette, V.; Reichinnek, S.; Torcini, A.;
-    Cossart, R.; Malvache, A.
-  \newblock Internal representation of hippocampal neuronal population spans a
-    time-distance continuum.
-  \newblock {\em Proceedings of the National Academy of Sciences} {\bf 2019},
-    {\em 116},~7477--7482.
-  \newblock {\url{https://doi.org/10.1073/pnas.1718518116}}.
-
-  \bibitem[Agus \em{et~al.}(2010)Agus, Thorpe, and Pressnitzer]{agus_rapid_2010}
-  Agus, T.R.; Thorpe, S.J.; Pressnitzer, D.
-  \newblock Rapid {Formation} of {Robust} {Auditory} {Memories}: {Insights} from
-    {Noise}.
-  \newblock {\em Neuron} {\bf 2010}, {\em 66},~610--618.
-  \newblock {\url{https://doi.org/10.1016/j.neuron.2010.04.014}}.
-
-  \bibitem[Izhikevich(2006)]{izhikevich_polychronization_2006}
-  Izhikevich, E.M.
-  \newblock Polychronization: {Computation} with {Spikes}.
-  \newblock {\em Neural Computation} {\bf 2006}, {\em 18},~245--282.
-  \newblock 00000, {\url{https://doi.org/10/bgh4qv}}.
-
-  \bibitem[Simoncelli \em{et~al.}(2003)Simoncelli, Paninski, Pillow, and
-    Schwartz]{simoncelli_characterization_2003}
-  Simoncelli, E.P.; Paninski, L.; Pillow, J.; Schwartz, O.
-  \newblock Characterization of {Neural} {Responses} with {Stochastic} {Stimuli}
-    {\bf 2003}.
-
-  \bibitem[Jazayeri and Movshon(2006)]{jazayeri_optimal_2006}
-  Jazayeri, M.; Movshon, J.A.
-  \newblock Optimal representation of sensory information by neural populations.
-  \newblock {\em Nature neuroscience} {\bf 2006}, {\em 9},~690--696.
-  \newblock Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/nn1691}}.
-
-  \bibitem[Berens \em{et~al.}(2012)Berens, Ecker, Cotton, Ma, Bethge, and
-    Tolias]{berens_fast_2012}
-  Berens, P.; Ecker, A.S.; Cotton, R.J.; Ma, W.J.; Bethge, M.; Tolias, A.S.
-  \newblock A {Fast} and {Simple} {Population} {Code} for {Orientation} in
-    {Primate} {V1}.
-  \newblock {\em Journal of Neuroscience} {\bf 2012}, {\em 32},~10618--10626.
-  \newblock {\url{https://doi.org/10.1523/jneurosci.1335-12.2012}}.
-
-  \bibitem[Bellec \em{et~al.}(2021)Bellec, Wang, Modirshanechi, Brea, and
-    Gerstner]{bellec_fitting_2021}
-  Bellec, G.; Wang, S.; Modirshanechi, A.; Brea, J.; Gerstner, W.
-  \newblock Fitting summary statistics of neural data with a differentiable
-    spiking network simulator.
-  \newblock Technical Report 2106.10064, arXiv,  2021.
-
-  \bibitem[Kohn and Smith(2016)]{kohn_utah_2016}
-  Kohn, A.; Smith, M.
-  \newblock Utah array extracellular recordings of spontaneous and visually
-    evoked activity from anesthetized macaque primary visual cortex ({V1}).,
-    2016.
-  \newblock {\url{https://doi.org/10.6080/K0NC5Z4X}}.
-
-  \bibitem[Warner \em{et~al.}(2022)Warner, Ruda, and
-    Sommer]{warner_probabilistic_2022}
-  Warner, C.; Ruda, K.; Sommer, F.T.
-  \newblock A probabilistic latent variable model for detecting structure in
-    binary data.
-  \newblock {\em arXiv:2201.11108 [cs, q-bio, stat]} {\bf 2022}.
-  \newblock arXiv: 2201.11108.
-
-  \bibitem[Victor and Purpura(1996)]{victor_nature_1996}
-  Victor, J.D.; Purpura, K.P.
-  \newblock Nature and precision of temporal coding in visual cortex: a
-    metric-space analysis.
-  \newblock {\em Journal of Neurophysiology} {\bf 1996}, {\em 76},~1310--1326.
-  \newblock {\url{https://doi.org/10.1152/jn.1996.76.2.1310}}.
-
-  \bibitem[van Rossum(2001)]{van_rossum_novel_2001}
-  van Rossum, M.C.
-  \newblock A novel spike distance.
-  \newblock {\em Neural Computation} {\bf 2001}, {\em 13},~751--763.
-  \newblock {\url{https://doi.org/10.1162/089976601300014321}}.
-
-  \bibitem[Kreuz \em{et~al.}(2007)Kreuz, Haas, Morelli, Abarbanel, and
-    Politi]{kreuz_measuring_2007}
-  Kreuz, T.; Haas, J.S.; Morelli, A.; Abarbanel, H.D.I.; Politi, A.
-  \newblock Measuring spike train synchrony.
-  \newblock {\em Journal of Neuroscience Methods} {\bf 2007}, {\em
-    165},~151--161.
-  \newblock {\url{https://doi.org/10.1016/j.jneumeth.2007.05.031}}.
-
-  \bibitem[Moser and Natschlager(2014)]{moser_stability_2014}
-  Moser, B.A.; Natschlager, T.
-  \newblock On {Stability} of {Distance} {Measures} for {Event} {Sequences}
-    {Induced} by {Level}-{Crossing} {Sampling}.
-  \newblock {\em IEEE Transactions on Signal Processing} {\bf 2014}, {\em
-    62},~1987--1999.
-  \newblock {\url{https://doi.org/10.1109/tsp.2014.2305642}}.
-
-  \bibitem[Weyl(1916)]{weyl_ueber_1916}
-  Weyl, H.
-  \newblock Ueber die {Gleichverteilung} von {Zahlen} mod. {Eins}.
-  \newblock {\em Mathematische Annalen} {\bf 1916}, {\em 77},~313--352.
-  \newblock {\url{https://doi.org/10.1007/bf01475864}}.
-
-  \bibitem[Aronov and Victor(2004)]{aronov_non-euclidean_2004}
-  Aronov, D.; Victor, J.D.
-  \newblock Non-{Euclidean} properties of spike train metric spaces.
-  \newblock {\em Physical Review E} {\bf 2004}, {\em 69},~061905.
-  \newblock Publisher: American Physical Society,
-    {\url{https://doi.org/10.1103/PhysRevE.69.061905}}.
-
-  \bibitem[Levakova \em{et~al.}(2015)Levakova, Tamborrino, Ditlevsen, and
-    Lansky]{levakova_review_2015}
-  Levakova, M.; Tamborrino, M.; Ditlevsen, S.; Lansky, P.
-  \newblock A review of the methods for neuronal response latency estimation.
-  \newblock {\em Biosystems} {\bf 2015}, {\em 136},~23--34.
-  \newblock {\url{https://doi.org/10.1016/j.biosystems.2015.04.008}}.
-
-  \bibitem[Grün \em{et~al.}(2002)Grün, Diesmann, and
-    Aertsen]{grun_unitary_2002-1}
-  Grün, S.; Diesmann, M.; Aertsen, A.
-  \newblock Unitary {Events} in {Multiple} {Single}-{Neuron} {Spiking}
-    {Activity}: {II}. {Nonstationary} {Data}.
-  \newblock {\em Neural Computation} {\bf 2002}, {\em 14},~81--119.
-  \newblock {\url{https://doi.org/10.1162/089976602753284464}}.
-
-  \bibitem[Grün \em{et~al.}(2010)Grün, Diesmann, and
-    Aertsen]{grun_unitary_2010}
-  Grün, S.; Diesmann, M.; Aertsen, A.
-  \newblock Unitary {Event} {Analysis}. In {\em Analysis of {Parallel} {Spike}
-    {Trains}}; Springer US,  2010; pp. 191--220.
-
-  \bibitem[Grün \em{et~al.}(2002)Grün, Diesmann, and
-    Aertsen]{grun_unitary_2002}
-  Grün, S.; Diesmann, M.; Aertsen, A.
-  \newblock Unitary {Events} in {Multiple} {Single}-{Neuron} {Spiking}
-    {Activity}: {I}. {Detection} and {Significance}.
-  \newblock {\em Neural Computation} {\bf 2002}, {\em 14},~43--80.
-  \newblock {\url{https://doi.org/10.1162/089976602753284455}}.
-
-  \bibitem[Torre \em{et~al.}(2013)Torre, Picado-Muiño, Denker, Borgelt, and
-    Grün]{torre_statistical_2013}
-  Torre, E.; Picado-Muiño, D.; Denker, M.; Borgelt, C.; Grün, S.
-  \newblock Statistical evaluation of synchronous spike patterns extracted by
-    frequent item set mining.
-  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2013}, {\em 7}.
-  \newblock {\url{https://doi.org/10.3389/fncom.2013.00132}}.
-
-  \bibitem[Quaglio \em{et~al.}(2018)Quaglio, Rostami, Torre, and
-    Grün]{quaglio_methods_2018}
-  Quaglio, P.; Rostami, V.; Torre, E.; Grün, S.
-  \newblock Methods for identification of spike patterns in massively parallel
-    spike trains.
-  \newblock {\em Biological Cybernetics} {\bf 2018}, {\em 112},~57--80.
-  \newblock {\url{https://doi.org/10.1007/s00422-018-0755-0}}.
-
-  \bibitem[Stella \em{et~al.}(2019)Stella, Quaglio, Torre, and
-    Grün]{stella_3d-spade_2019}
-  Stella, A.; Quaglio, P.; Torre, E.; Grün, S.
-  \newblock 3d-{SPADE}: {Significance} evaluation of spatio-temporal patterns of
-    various temporal extents.
-  \newblock {\em Biosystems} {\bf 2019}, {\em 185},~104022.
-  \newblock {\url{https://doi.org/10.1016/j.biosystems.2019.104022}}.
-
-  \bibitem[Stella \em{et~al.}(2022)Stella, Bouss, Palm, and
-    Grün]{stella_comparing_2022}
-  Stella, A.; Bouss, P.; Palm, G.; Grün, S.
-  \newblock Comparing {Surrogates} to {Evaluate} {Precisely} {Timed}
-    {Higher}-{Order} {Spike} {Correlations}.
-  \newblock {\em eneuro} {\bf 2022}, {\em 9},~ENEURO.0505--21.2022.
-  \newblock {\url{https://doi.org/10.1523/eneuro.0505-21.2022}}.
-
-  \bibitem[Grossberger \em{et~al.}(2018)Grossberger, Battaglia, and
-    Vinck]{grossberger_unsupervised_2018}
-  Grossberger, L.; Battaglia, F.P.; Vinck, M.
-  \newblock Unsupervised clustering of temporal patterns in high-dimensional
-    neuronal ensembles using a novel dissimilarity measure.
-  \newblock {\em PLOS Computational Biology} {\bf 2018}, {\em 14},~e1006283.
-  \newblock {\url{https://doi.org/10/gdvbsx}}.
-
-  \bibitem[Nádasdy \em{et~al.}(1999)Nádasdy, Hirase, Czurkó, Csicsvari, and
-    Buzsáki]{nadasdy_replay_1999}
-  Nádasdy, Z.; Hirase, H.; Czurkó, A.; Csicsvari, J.; Buzsáki, G.
-  \newblock Replay and {Time} {Compression} of {Recurring} {Spike} {Sequences} in
-    the {Hippocampus}.
-  \newblock {\em Journal of Neuroscience} {\bf 1999}, {\em 19},~9497--9507.
-  \newblock Publisher: Society for Neuroscience Section: ARTICLE,
-    {\url{https://doi.org/10.1523/JNEUROSCI.19-21-09497.1999}}.
-
-  \bibitem[Lee and Wilson(2004)]{lee_combinatorial_2004}
-  Lee, A.K.; Wilson, M.A.
-  \newblock A {Combinatorial} {Method} for {Analyzing} {Sequential} {Firing}
-    {Patterns} {Involving} an {Arbitrary} {Number} of {Neurons} {Based} on
-    {Relative} {Time} {Order}.
-  \newblock {\em Journal of Neurophysiology} {\bf 2004}, {\em 92},~2555--2573.
-  \newblock {\url{https://doi.org/10.1152/jn.01030.2003}}.
-
-  \bibitem[Sotomayor-Gómez \em{et~al.}(2021)Sotomayor-Gómez, Battaglia, and
-    Vinck]{sotomayor-gomez_spikeship_2021}
-  Sotomayor-Gómez, B.; Battaglia, F.P.; Vinck, M.
-  \newblock {SpikeShip}: {A} method for fast, unsupervised discovery of
-    high-dimensional neural spiking patterns.
-  \newblock {\em bioRxiv : the preprint server for biology} {\bf 2021}, pp.
-    2020--06.
-  \newblock Publisher: Cold Spring Harbor Laboratory,
-    {\url{https://doi.org/10.1101/2020.06.03.131573}}.
-
-  \bibitem[Pachitariu \em{et~al.}(2018)Pachitariu, Stringer, and
-    Harris]{pachitariu_robustness_2018}
-  Pachitariu, M.; Stringer, C.; Harris, K.D.
-  \newblock Robustness of {Spike} {Deconvolution} for {Neuronal} {Calcium}
-    {Imaging}.
-  \newblock {\em The Journal of Neuroscience} {\bf 2018}, {\em 38},~7976--7985.
-  \newblock {\url{https://doi.org/10.1523/jneurosci.3339-17.2018}}.
-
-  \bibitem[Stringer(2020)]{stringer_mouselandrastermap_2020}
-  Stringer, C.
-  \newblock {MouseLand}/rastermap: {A} multi-dimensional embedding algorithm,
-    2020.
-
-  \bibitem[Stringer \em{et~al.}(2019)Stringer, Pachitariu, Steinmetz, Reddy,
-    Carandini, and Harris]{stringer_spontaneous_2019}
-  Stringer, C.; Pachitariu, M.; Steinmetz, N.; Reddy, C.B.; Carandini, M.;
-    Harris, K.D.
-  \newblock Spontaneous behaviors drive multidimensional, brainwide activity.
-  \newblock {\em Science} {\bf 2019}, {\em 364}.
-  \newblock {\url{https://doi.org/10.1126/science.aav7893}}.
-
-  \bibitem[Stringer \em{et~al.}(2021)Stringer, Michaelos, Tsyboulski, Lindo, and
-    Pachitariu]{stringer_high-precision_2021}
-  Stringer, C.; Michaelos, M.; Tsyboulski, D.; Lindo, S.E.; Pachitariu, M.
-  \newblock High-precision coding in visual cortex.
-  \newblock {\em Cell} {\bf 2021}, {\em 184},~2767--2778.e15.
-  \newblock {\url{https://doi.org/10.1016/j.cell.2021.03.042}}.
-
-  \bibitem[Russo and Durstewitz(2017)]{russo_cell_2017}
-  Russo, E.; Durstewitz, D.
-  \newblock Cell assemblies at multiple time scales with arbitrary lag
-    constellations.
-  \newblock {\em eLife} {\bf 2017}, {\em 6}.
-  \newblock {\url{https://doi.org/10.7554/elife.19428}}.
-
-  \bibitem[Pipa \em{et~al.}(2008)Pipa, Wheeler, Singer, and
-    Nikolić]{pipa_neuroxidence_2008}
-  Pipa, G.; Wheeler, D.W.; Singer, W.; Nikolić, D.
-  \newblock {NeuroXidence}: reliable and efficient analysis of an excess or
-    deficiency of joint-spike events.
-  \newblock {\em Journal of Computational Neuroscience} {\bf 2008}, {\em
-    25},~64--88.
-  \newblock {\url{https://doi.org/10.1007/s10827-007-0065-3}}.
-
-  \bibitem[Torre \em{et~al.}(2016)Torre, Quaglio, Denker, Brochier, Riehle, and
-    Grun]{torre_synchronous_2016}
-  Torre, E.; Quaglio, P.; Denker, M.; Brochier, T.; Riehle, A.; Grun, S.
-  \newblock Synchronous {Spike} {Patterns} in {Macaque} {Motor} {Cortex} during
-    an {Instructed}-{Delay} {Reach}-to-{Grasp} {Task}.
-  \newblock {\em Journal of Neuroscience} {\bf 2016}, {\em 36},~8329--8340.
-  \newblock {\url{https://doi.org/10.1523/jneurosci.4375-15.2016}}.
-
-  \bibitem[Williams \em{et~al.}(2020)Williams, Degleris, Wang, and
-    Linderman]{williams_point_2020}
-  Williams, A.H.; Degleris, A.; Wang, Y.; Linderman, S.W.
-  \newblock Point process models for sequence detection in high-dimensional
-    neural spike trains.
-  \newblock Technical Report 2010.04875, arXiv,  2020.
-
-  \bibitem[Kass \em{et~al.}(2005)Kass, Ventura, and Brown]{kass_statistical_2005}
-  Kass, R.E.; Ventura, V.; Brown, E.N.
-  \newblock Statistical issues in the analysis of neuronal data.
-  \newblock {\em Journal of neurophysiology} {\bf 2005}, {\em 94},~8--25.
-  \newblock Publisher: American Physiological Society,
-    {\url{https://doi.org/10.1152/jn.00648.2004}}.
-
-  \bibitem[van Kempen \em{et~al.}(2021)van Kempen, Gieselmann, Boyd, Steinmetz,
-    Moore, Engel, and Thiele]{van_kempen_top-down_2021}
-  van Kempen, J.; Gieselmann, M.A.; Boyd, M.; Steinmetz, N.A.; Moore, T.; Engel,
-    T.A.; Thiele, A.
-  \newblock Top-down coordination of local cortical state during selective
-    attention.
-  \newblock {\em Neuron} {\bf 2021}, {\em 109},~894--904.e8.
-  \newblock {\url{https://doi.org/10.1016/j.neuron.2020.12.013}}.
-
-  \bibitem[Pasturel \em{et~al.}(2020)Pasturel, Montagnini, and
-    Perrinet]{pasturel_humans_2020}
-  Pasturel, C.; Montagnini, A.; Perrinet, L.U.
-  \newblock Humans adapt their anticipatory eye movements to the volatility of
-    visual motion properties.
-  \newblock {\em PLoS Computational Biology} {\bf 2020}.
-  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1007438}}.
-
-  \bibitem[Von~Helmholz(1850)]{von_helmholz_messungen_1850}
-  Von~Helmholz, H.
-  \newblock Messungen über den zeitlichen {Verlauf} der {Zuckung} animalischer
-    {Muskeln} und die {Fortpflanzungsgeschwindigkeit} der {Reizung} in den
-    {Nerven}.
-  \newblock {\em Archiv für Anatomie, Physiologie und wissenschaftliche Medicin}
-    {\bf 1850}, {\em 17},~176--364.
-
-  \bibitem[Peyrard(2020)]{peyrard_how_2020}
-  Peyrard, M.
-  \newblock How is information transmitted in a nerve?
-  \newblock {\em Journal of Biological Physics} {\bf 2020}, {\em 46},~327--341.
-  \newblock {\url{https://doi.org/10.1007/s10867-020-09557-2}}.
-
-  \bibitem[Young(1938)]{young_functioning_1938}
-  Young, J.Z.
-  \newblock The {Functioning} of the {Giant} {Nerve} {Fibres} of the {Squid}.
-  \newblock {\em Journal of Experimental Biology} {\bf 1938}, {\em 15},~170--185.
-  \newblock {\url{https://doi.org/10.1242/jeb.15.2.170}}.
-
-  \bibitem[Madadi~Asl \em{et~al.}(2018)Madadi~Asl, Valizadeh, and
-    Tass]{madadi_asl_dendritic_2018}
-  Madadi~Asl, M.; Valizadeh, A.; Tass, P.A.
-  \newblock Dendritic and {Axonal} {Propagation} {Delays} {May} {Shape}
-    {Neuronal} {Networks} {With} {Plastic} {Synapses}.
-  \newblock {\em Frontiers in Physiology} {\bf 2018}, {\em 9},~1849.
-  \newblock {\url{https://doi.org/10.3389/fphys.2018.01849}}.
-
-  \bibitem[Stetson \em{et~al.}(1992)Stetson, Albers, Silverstein, and
-    Wolfe]{stetson_effects_1992}
-  Stetson, D.S.; Albers, J.W.; Silverstein, B.A.; Wolfe, R.A.
-  \newblock Effects of age, sex, and anthropometric factors on nerve conduction
-    measures.
-  \newblock {\em Muscle \& Nerve} {\bf 1992}, {\em 15},~1095--1104.
-  \newblock \_eprint:
-    https://onlinelibrary.wiley.com/doi/pdf/10.1002/mus.880151007,
-    {\url{https://doi.org/10.1002/mus.880151007}}.
-
-  \bibitem[Jeffress(1948)]{jeffress_place_1948}
-  Jeffress, L.A.
-  \newblock A place theory of sound localization.
-  \newblock {\em Journal of Comparative and Physiological Psychology} {\bf 1948},
-    {\em 41},~35--39.
-  \newblock Place: US Publisher: American Psychological Association,
-    {\url{https://doi.org/10.1037/h0061495}}.
-
-  \bibitem[Konishi(2003)]{konishi_coding_2003}
-  Konishi, M.
-  \newblock Coding of auditory space.
-  \newblock {\em Annual Review of Neuroscience} {\bf 2003}, {\em 26},~31--55.
-  \newblock {\url{https://doi.org/10.1146/annurev.neuro.26.041002.131123}}.
-
-  \bibitem[Gerstner \em{et~al.}(1996)Gerstner, Kempter, van Hemmen, and
-    Wagner]{gerstner_neuronal_1996}
-  Gerstner, W.; Kempter, R.; van Hemmen, J.L.; Wagner, H.
-  \newblock A neuronal learning rule for sub-millisecond temporal coding.
-  \newblock {\em Nature} {\bf 1996}, {\em 383},~76--78.
-  \newblock Number: 6595 Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/383076a0}}.
-
-  \bibitem[Seidl \em{et~al.}(2010)Seidl, Rubel, and
-    Harris]{seidl_mechanisms_2010}
-  Seidl, A.H.; Rubel, E.W.; Harris, D.M.
-  \newblock Mechanisms for adjusting interaural time differences to achieve
-    binaural coincidence detection.
-  \newblock {\em The Journal of Neuroscience: The Official Journal of the Society
-    for Neuroscience} {\bf 2010}, {\em 30},~70--80.
-  \newblock {\url{https://doi.org/10.1523/JNEUROSCI.3464-09.2010}}.
-
-  \bibitem[Camon \em{et~al.}(2019)Camon, Hugues, Erlandson, Robbe, Lagoun,
-    Marouane, and Bureau]{camon_timing_2019}
-  Camon, J.; Hugues, S.; Erlandson, M.A.; Robbe, D.; Lagoun, S.; Marouane, E.;
-    Bureau, I.
-  \newblock The {Timing} of {Sensory}-{Guided} {Behavioral} {Response} is
-    {Represented} in the {Mouse} {Primary} {Somatosensory} {Cortex}.
-  \newblock {\em Cerebral Cortex} {\bf 2019}, {\em 29},~3034--3047.
-  \newblock {\url{https://doi.org/10.1093/cercor/bhy169}}.
-
-  \bibitem[Gasser and Grundfest(1939)]{gasser_axon_1939}
-  Gasser, H.S.; Grundfest, H.
-  \newblock {AXON} {DIAMETERS} {IN} {RELATION} {TO} {THE} {SPIKE} {DIMENSIONS}
-    {AND} {THE} {CONDUCTION} {VELOCITY} {IN} {MAMMALIAN} {A} {FIBERS}.
-  \newblock {\em American Journal of Physiology-Legacy Content} {\bf 1939}.
-  \newblock Publisher: American Physiological Society,
-    {\url{https://doi.org/10.1152/ajplegacy.1939.127.2.393}}.
-
-  \bibitem[Brill \em{et~al.}(1977)Brill, Waxman, Moore, and
-    Joyner]{brill_conduction_1977}
-  Brill, M.H.; Waxman, S.G.; Moore, J.W.; Joyner, R.W.
-  \newblock Conduction velocity and spike configuration in myelinated fibres:
-    computed dependence on internode distance.
-  \newblock {\em Journal of Neurology, Neurosurgery, and Psychiatry} {\bf 1977},
-    {\em 40},~769--774.
-
-  \bibitem[Pérez-Cerdá \em{et~al.}(2015)Pérez-Cerdá, Sánchez-Gómez, and
-    Matute]{perez-cerda_pio_2015}
-  Pérez-Cerdá, F.; Sánchez-Gómez, M.V.; Matute, C.
-  \newblock Pío del {Río} {Hortega} and the discovery of the oligodendrocytes.
-  \newblock {\em Frontiers in Neuroanatomy} {\bf 2015}, {\em 9},~92.
-  \newblock {\url{https://doi.org/10.3389/fnana.2015.00092}}.
-
-  \bibitem[Schmitt and Bear(1939)]{schmitt_ultrastructure_1939}
-  Schmitt, F.O.; Bear, R.S.
-  \newblock The {Ultrastructure} of the {Nerve} {Axon} {Sheath}.
-  \newblock {\em Biological Reviews} {\bf 1939}, {\em 14},~27--50.
-  \newblock \_eprint:
-    https://onlinelibrary.wiley.com/doi/pdf/10.1111/j.1469-185X.1939.tb00922.x,
-    {\url{https://doi.org/10.1111/j.1469-185X.1939.tb00922.x}}.
-
-  \bibitem[Simons and Nave(2016)]{simons_oligodendrocytes_2016}
-  Simons, M.; Nave, K.A.
-  \newblock Oligodendrocytes: {Myelination} and {Axonal} {Support}.
-  \newblock {\em Cold Spring Harbor Perspectives in Biology} {\bf 2016}, {\em
-    8},~a020479.
-  \newblock {\url{https://doi.org/10.1101/cshperspect.a020479}}.
-
-  \bibitem[Duncan \em{et~al.}(2021)Duncan, Simkins, and
-    Emery]{duncan_neuron-oligodendrocyte_2021}
-  Duncan, G.J.; Simkins, T.J.; Emery, B.
-  \newblock Neuron-{Oligodendrocyte} {Interactions} in the {Structure} and
-    {Integrity} of {Axons}.
-  \newblock {\em Frontiers in Cell and Developmental Biology} {\bf 2021}, {\em
-    9}.
-  \newblock {\url{https://doi.org/10.3389/fcell.2021.653101}}.
-
-  \bibitem[Fields(2015)]{fields_new_2015}
-  Fields, R.D.
-  \newblock A new mechanism of nervous system plasticity: activity-dependent
-    myelination.
-  \newblock {\em Nature Reviews Neuroscience} {\bf 2015}, {\em 16},~756--767.
-  \newblock Number: 12 Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/nrn4023}}.
-
-  \bibitem[Fields and Bukalo(2020)]{fields_myelin_2020}
-  Fields, R.D.; Bukalo, O.
-  \newblock Myelin makes memories.
-  \newblock {\em Nature Neuroscience} {\bf 2020}, {\em 23},~469--470.
-  \newblock {\url{https://doi.org/10.1038/s41593-020-0606-x}}.
-
-  \bibitem[Reynolds and Slater(1928)]{reynolds_study_1928}
-  Reynolds, F.E.; Slater, J.K.
-  \newblock A {Study} of the {Structure} and {Function} of the {Interstitial}
-    {Tissue} of the {Central} {Nervous} {System}.
-  \newblock {\em Edinburgh Medical Journal} {\bf 1928}, {\em 35},~49--57.
-
-  \bibitem[Steadman \em{et~al.}(2020)Steadman, Xia, Ahmed, Mocle, Penning,
-    Geraghty, Steenland, Monje, Josselyn, and
-    Frankland]{steadman_disruption_2020}
-  Steadman, P.E.; Xia, F.; Ahmed, M.; Mocle, A.J.; Penning, A.R.; Geraghty, A.C.;
-    Steenland, H.W.; Monje, M.; Josselyn, S.A.; Frankland, P.W.
-  \newblock Disruption of {Oligodendrogenesis} {Impairs} {Memory} {Consolidation}
-    in {Adult} {Mice}.
-  \newblock {\em Neuron} {\bf 2020}, {\em 105},~150--164.e6.
-  \newblock {\url{https://doi.org/10.1016/j.neuron.2019.10.013}}.
-
-  \bibitem[Pan \em{et~al.}(2020)Pan, Mayoral, Choi, Chan, and
-    Kheirbek]{pan_preservation_2020}
-  Pan, S.; Mayoral, S.R.; Choi, H.S.; Chan, J.R.; Kheirbek, M.A.
-  \newblock Preservation of a remote fear memory requires new myelin formation.
-  \newblock {\em Nature Neuroscience} {\bf 2020}, {\em 23},~487--499.
-  \newblock {\url{https://doi.org/10.1038/s41593-019-0582-1}}.
-
-  \bibitem[Wan \em{et~al.}(2020)Wan, Cheli, Santiago-González, Rosenblum, Wan,
-    and Paez]{wan_impaired_2020}
-  Wan, R.; Cheli, V.T.; Santiago-González, D.A.; Rosenblum, S.L.; Wan, Q.; Paez,
-    P.M.
-  \newblock Impaired {Postnatal} {Myelination} in a {Conditional} {Knockout}
-    {Mouse} for the {Ferritin} {Heavy} {Chain} in {Oligodendroglial} {Cells}.
-  \newblock {\em The Journal of Neuroscience} {\bf 2020}, {\em 40},~7609--7624.
-  \newblock {\url{https://doi.org/10.1523/JNEUROSCI.1281-20.2020}}.
-
-  \bibitem[Xue \em{et~al.}(2021)Xue, Zhu, Liu, Lin, Li, Li, and
-    Zhuo]{xue_demyelination_2021}
-  Xue, J.; Zhu, Y.; Liu, Z.; Lin, J.; Li, Y.; Li, Y.; Zhuo, Y.
-  \newblock Demyelination of the {Optic} {Nerve}: {An} {Underlying} {Factor} in
-    {Glaucoma}?
-  \newblock {\em Frontiers in Aging Neuroscience} {\bf 2021}, {\em 13},~701322.
-  \newblock {\url{https://doi.org/10.3389/fnagi.2021.701322}}.
-
-  \bibitem[Kuhn \em{et~al.}(2019)Kuhn, Gritti, Crooks, and
-    Dombrowski]{kuhn_oligodendrocytes_2019}
-  Kuhn, S.; Gritti, L.; Crooks, D.; Dombrowski, Y.
-  \newblock Oligodendrocytes in {Development}, {Myelin} {Generation} and
-    {Beyond}.
-  \newblock {\em Cells} {\bf 2019}, {\em 8},~1424.
-  \newblock {\url{https://doi.org/10.3390/cells8111424}}.
-
-  \bibitem[Baraban \em{et~al.}(2018)Baraban, Koudelka, and
-    Lyons]{baraban_ca2_2018}
-  Baraban, M.; Koudelka, S.; Lyons, D.A.
-  \newblock Ca2+ activity signatures of myelin sheath formation and growth in
-    vivo.
-  \newblock {\em Nature neuroscience} {\bf 2018}, {\em 21},~19--23.
-  \newblock {\url{https://doi.org/10.1038/s41593-017-0040-x}}.
-
-  \bibitem[Nave and Salzer(2006)]{nave_axonal_2006}
-  Nave, K.A.; Salzer, J.L.
-  \newblock Axonal regulation of myelination by neuregulin 1.
-  \newblock {\em Current Opinion in Neurobiology} {\bf 2006}, {\em 16},~492--500.
-  \newblock {\url{https://doi.org/10.1016/j.conb.2006.08.008}}.
-
-  \bibitem[Cullen \em{et~al.}(2021)Cullen, Pepper, Clutterbuck, Pitman, Oorschot,
-    Auderset, Tang, Ramm, Emery, Rodger, Jolivet, and
-    Young]{cullen_periaxonal_2021}
-  Cullen, C.L.; Pepper, R.E.; Clutterbuck, M.T.; Pitman, K.A.; Oorschot, V.;
-    Auderset, L.; Tang, A.D.; Ramm, G.; Emery, B.; Rodger, J.;  et~al.
-  \newblock Periaxonal and nodal plasticities modulate action potential
-    conduction in the adult mouse brain.
-  \newblock {\em Cell Reports} {\bf 2021}, {\em 34}.
-  \newblock Publisher: Elsevier,
-    {\url{https://doi.org/10.1016/j.celrep.2020.108641}}.
-
-  \bibitem[Gibson \em{et~al.}(2014)Gibson, Purger, Mount, Goldstein, Lin, Wood,
-    Inema, Miller, Bieri, Zuchero, Barres, Woo, Vogel, and
-    Monje]{gibson_neuronal_2014}
-  Gibson, E.M.; Purger, D.; Mount, C.W.; Goldstein, A.K.; Lin, G.L.; Wood, L.S.;
-    Inema, I.; Miller, S.E.; Bieri, G.; Zuchero, J.B.;  et~al.
-  \newblock Neuronal {Activity} {Promotes} {Oligodendrogenesis} and {Adaptive}
-    {Myelination} in the {Mammalian} {Brain}.
-  \newblock {\em Science (New York, N.Y.)} {\bf 2014}, {\em 344},~1252304.
-  \newblock {\url{https://doi.org/10.1126/science.1252304}}.
-
-  \bibitem[Spencer \em{et~al.}(2018)Spencer, Meffin, Burkitt, and
-    Grayden]{spencer_compensation_2018}
-  Spencer, M.J.; Meffin, H.; Burkitt, A.N.; Grayden, D.B.
-  \newblock Compensation for {Traveling} {Wave} {Delay} {Through} {Selection} of
-    {Dendritic} {Delays} {Using} {Spike}-{Timing}-{Dependent} {Plasticity} in a
-    {Model} of the {Auditory} {Brainstem}.
-  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2018}, {\em
-    12},~36.
-  \newblock {\url{https://doi.org/10.3389/fncom.2018.00036}}.
-
-  \bibitem[Mel \em{et~al.}(2017)Mel, Schiller, and Poirazi]{mel_synaptic_2017}
-  Mel, B.W.; Schiller, J.; Poirazi, P.
-  \newblock Synaptic plasticity in dendrites: complications and coping
-    strategies.
-  \newblock {\em Current Opinion in Neurobiology} {\bf 2017}, {\em 43},~177--186.
-  \newblock {\url{https://doi.org/10.1016/j.conb.2017.03.012}}.
-
-  \bibitem[Golding \em{et~al.}(2002)Golding, Staff, and
-    Spruston]{golding_dendritic_2002}
-  Golding, N.L.; Staff, N.P.; Spruston, N.
-  \newblock Dendritic spikes as a mechanism for cooperative long-term
-    potentiation.
-  \newblock {\em Nature} {\bf 2002}, {\em 418},~326--331.
-  \newblock Number: 6895 Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/nature00854}}.
-
-  \bibitem[Maass(1997)]{maass_networks_1997}
-  Maass, W.
-  \newblock Networks of spiking neurons: {The} third generation of neural network
-    models.
-  \newblock {\em Neural Networks} {\bf 1997}, {\em 10},~1659--1671.
-
-  \bibitem[Neftci \em{et~al.}(2019)Neftci, Mostafa, and
-    Zenke]{neftci_surrogate_2019}
-  Neftci, E.O.; Mostafa, H.; Zenke, F.
-  \newblock Surrogate {Gradient} {Learning} in {Spiking} {Neural} {Networks}:
-    {Bringing} the {Power} of {Gradient}-{Based} {Optimization} to {Spiking}
-    {Neural} {Networks}.
-  \newblock {\em IEEE Signal Processing Magazine} {\bf 2019}, {\em 36},~51--63.
-  \newblock {\url{https://doi.org/10.1109/MSP.2019.2931595}}.
-
-  \bibitem[Rueckauer \em{et~al.}(2017)Rueckauer, Lungu, Hu, Pfeiffer, and
-    Liu]{rueckauer_conversion_2017}
-  Rueckauer, B.; Lungu, I.A.; Hu, Y.; Pfeiffer, M.; Liu, S.C.
-  \newblock Conversion of {Continuous}-{Valued} {Deep} {Networks} to {Efficient}
-    {Event}-{Driven} {Networks} for {Image} {Classification}.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2017}, {\em 11}.
-  \newblock {\url{https://doi.org/10.3389/fnins.2017.00682}}.
-
-  \bibitem[Susi \em{et~al.}(2021)Susi, Antón-Toro, Maestú, Pereda, and
-    Mirasso]{susi_nmnsd-spiking_2021}
-  Susi, G.; Antón-Toro, L.F.; Maestú, F.; Pereda, E.; Mirasso, C.
-  \newblock {nMNSD}-{A} {Spiking} {Neuron}-{Based} {Classifier} {That} {Combines}
-    {Weight}-{Adjustment} and {Delay}-{Shift}.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2021}, {\em 15},~582608.
-  \newblock {\url{https://doi.org/10.3389/fnins.2021.582608}}.
-
-  \bibitem[Davies \em{et~al.}(2018)Davies, Srinivasa, Lin, Chinya, Cao, Choday,
-    Dimou, Joshi, Imam, Jain, Liao, Lin, Lines, Liu, Mathaikutty, McCoy, Paul,
-    Tse, Venkataramanan, Weng, Wild, Yang, and Wang]{davies_loihi_2018}
-  Davies, M.; Srinivasa, N.; Lin, T.H.; Chinya, G.; Cao, Y.; Choday, S.H.; Dimou,
-    G.; Joshi, P.; Imam, N.; Jain, S.;  et~al.
-  \newblock Loihi: {A} {Neuromorphic} {Manycore} {Processor} with {On}-{Chip}
-    {Learning}.
-  \newblock {\em IEEE Micro} {\bf 2018}, {\em 38},~82--99.
-  \newblock {\url{https://doi.org/10.1109/MM.2018.112130359}}.
-
-  \bibitem[Lazar(2004)]{lazar_time_2004}
-  Lazar, A.A.
-  \newblock Time encoding with an integrate-and-fire neuron with a refractory
-    period.
-  \newblock {\em Neurocomputing} {\bf 2004}, {\em 58-60},~53--58.
-  \newblock {\url{https://doi.org/10.1016/j.neucom.2004.01.022}}.
-
-  \bibitem[Markram \em{et~al.}(1997)Markram, Lübke, Frotscher, and
-    Sakmann]{markram_regulation_1997}
-  Markram, H.; Lübke, J.; Frotscher, M.; Sakmann, B.
-  \newblock Regulation of {Synaptic} {Efficacy} by {Coincidence} of
-    {Postsynaptic} {APs} and {EPSPs}.
-  \newblock {\em Science} {\bf 1997}, {\em 275},~213--215.
-  \newblock {\url{https://doi.org/10.1126/science.275.5297.213}}.
-
-  \bibitem[Caporale and Dan(2008)]{caporale_spike_2008}
-  Caporale, N.; Dan, Y.
-  \newblock Spike {Timing}–{Dependent} {Plasticity}: {A} {Hebbian} {Learning}
-    {Rule}.
-  \newblock {\em Annual Review of Neuroscience} {\bf 2008}, {\em 31},~25--46.
-  \newblock {\url{https://doi.org/10.1146/annurev.neuro.31.060407.125639}}.
-
-  \bibitem[Hüning \em{et~al.}(1998)Hüning, Glünder, and
-    Palm]{huning_synaptic_1998}
-  Hüning, H.; Glünder, H.; Palm, G.
-  \newblock Synaptic {Delay} {Learning} in {Pulse}-{Coupled} {Neurons}.
-  \newblock {\em Neural Computation} {\bf 1998}, {\em 10},~555--565.
-  \newblock {\url{https://doi.org/10/cthfps}}.
-
-  \bibitem[Eurich \em{et~al.}(1999)Eurich, Pawelzik, Ernst, Cowan, and
-    Milton]{eurich_dynamics_1999}
-  Eurich, C.W.; Pawelzik, K.; Ernst, U.; Cowan, J.D.; Milton, J.G.
-  \newblock Dynamics of {Self}-{Organized} {Delay} {Adaptation}.
-  \newblock {\em Physical Review Letters} {\bf 1999}, {\em 82},~1594--1597.
-  \newblock 00082 Publisher: American Physical Society,
-    {\url{https://doi.org/10/dj9c7q}}.
-
-  \bibitem[Gütig and Sompolinsky(2006)]{gutig_tempotron_2006}
-  Gütig, R.; Sompolinsky, H.
-  \newblock The tempotron: a neuron that learns spike timing–based decisions.
-  \newblock {\em Nature Neuroscience} {\bf 2006}, {\em 9},~420--428.
-  \newblock {\url{https://doi.org/10.1038/nn1643}}.
-
-  \bibitem[Gütig(2014)]{gutig_spike_2014}
-  Gütig, R.
-  \newblock To spike, or when to spike?
-  \newblock {\em Current Opinion in Neurobiology} {\bf 2014}, {\em 25},~134--139.
-  \newblock {\url{https://doi.org/10.1016/j.conb.2014.01.004}}.
-
-  \bibitem[Pauli \em{et~al.}(2018)Pauli, Weidel, Kunkel, and
-    Morrison]{pauli_reproducing_2018}
-  Pauli, R.; Weidel, P.; Kunkel, S.; Morrison, A.
-  \newblock Reproducing {Polychronization}: {A} {Guide} to {Maximizing} the
-    {Reproducibility} of {Spiking} {Network} {Models}.
-  \newblock {\em Frontiers in Neuroinformatics} {\bf 2018}, {\em 12},~46.
-  \newblock 00000, {\url{https://doi.org/10/gd8zj5}}.
-
-  \bibitem[Guise \em{et~al.}(2014)Guise, Knott, and
-    Benuskova]{guise_bayesian_2014}
-  Guise, M.; Knott, A.; Benuskova, L.
-  \newblock A {Bayesian} {Model} of {Polychronicity}.
-  \newblock {\em Neural Computation} {\bf 2014}, {\em 26},~2052--2073.
-  \newblock {\url{https://doi.org/10.1162/neco_a_00620}}.
-
-  \bibitem[Zhang \em{et~al.}(2020)Zhang, Wu, Belatreche, Pan, Xie, Chua, Li, Qu,
-    and Li]{zhang_supervised_2020}
-  Zhang, M.; Wu, J.; Belatreche, A.; Pan, Z.; Xie, X.; Chua, Y.; Li, G.; Qu, H.;
-    Li, H.
-  \newblock Supervised learning in spiking neural networks with synaptic
-    delay-weight plasticity.
-  \newblock {\em Neurocomputing} {\bf 2020}, {\em 409},~103--118.
-  \newblock {\url{https://doi.org/10.1016/j.neucom.2020.03.079}}.
-
-  \bibitem[Ghosh \em{et~al.}(2021)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
-    Alfaro-Bittner, and Boccaletti]{ghosh_synchronization_2021}
-  Ghosh, D.; Frasca, M.; Rizzo, A.; Majhi, S.; Rakshit, S.; Alfaro-Bittner, K.;
-    Boccaletti, S.
-  \newblock Synchronization in time-varying networks.
-  \newblock {\em arXiv:2109.07618 [physics]} {\bf 2021}.
-  \newblock 00000 arXiv: 2109.07618.
-
-  \bibitem[Ghosh \em{et~al.}(2022)Ghosh, Frasca, Rizzo, Majhi, Rakshit,
-    Alfaro-Bittner, and Boccaletti]{ghosh_synchronized_2022}
-  Ghosh, D.; Frasca, M.; Rizzo, A.; Majhi, S.; Rakshit, S.; Alfaro-Bittner, K.;
-    Boccaletti, S.
-  \newblock The synchronized dynamics of time-varying networks.
-  \newblock {\em Physics Reports} {\bf 2022}, {\em 949},~1--63.
-  \newblock arXiv:2109.07618 [physics],
-    {\url{https://doi.org/10.1016/j.physrep.2021.10.006}}.
-
-  \bibitem[Izhikevich and Hoppensteadt(2009)]{izhikevich_polychronous_2009}
-  Izhikevich, E.M.; Hoppensteadt, F.C.
-  \newblock Polychronous {Wavefront} {Computations}.
-  \newblock {\em International Journal of Bifurcation and Chaos} {\bf 2009}, {\em
-    19},~1733--1739.
-  \newblock {\url{https://doi.org/10/db98d7}}.
-
-  \bibitem[Grimaldi and Perrinet(2022)]{grimaldi_learning_2022}
-  Grimaldi, A.; Perrinet, L.U.
-  \newblock Learning hetero-synaptic delays for motion detection in a single
-    layer of spiking neurons.
-  \newblock In Proceedings of the 2022 {IEEE} {International} {Conference} on
-    {Image} {Processing} ({ICIP}),  2022, pp. 3591--3595.
-  \newblock ISSN: 2381-8549,
-    {\url{https://doi.org/10.1109/ICIP46576.2022.9897394}}.
-
-  \bibitem[Madadi~Asl and
-    Ramezani~Akbarabadi(2022)]{madadi_asl_delay-dependent_2022}
-  Madadi~Asl, M.; Ramezani~Akbarabadi, S.
-  \newblock Delay-dependent transitions of phase synchronization and coupling
-    symmetry between neurons shaped by spike-timing-dependent plasticity.
-  \newblock {\em Cognitive Neurodynamics} {\bf 2022}.
-  \newblock {\url{https://doi.org/10.1007/s11571-022-09850-x}}.
-
-  \bibitem[Perrinet and Samuelides(2002)]{perrinet_coherence_2002}
-  Perrinet, L.; Samuelides, M.
-  \newblock Coherence detection in a spiking neuron via {Hebbian} learning.
-  \newblock {\em Neurocomputing} {\bf 2002}, {\em 44-46},~133--139.
-  \newblock {\url{https://doi.org/10.1016/s0925-2312(02)00374-0}}.
-
-  \bibitem[Perrinet \em{et~al.}(2001)Perrinet, Delorme, Samuelides, and
-    Thorpe]{perrinet_networks_2001}
-  Perrinet, L.; Delorme, A.; Samuelides, M.; Thorpe, S.
-  \newblock Networks of integrate-and-fire neuron using rank order coding {A}:
-    {How} to implement spike time dependent {Hebbian} plasticity.
-  \newblock {\em Neurocomputing} {\bf 2001}, {\em 38-40},~817--822.
-  \newblock {\url{https://doi.org/10.1016/s0925-2312(01)00460-x}}.
-
-  \bibitem[Gilson(2010)]{gilson_stdp_2010}
-  Gilson, M.
-  \newblock {STDP} in recurrent neuronal networks.
-  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2010}, {\em 4}.
-  \newblock {\url{https://doi.org/10.3389/fncom.2010.00023}}.
-
-  \bibitem[Datadien \em{et~al.}(2011)Datadien, Haselager, and
-    Sprinkhuizen-Kuyper]{datadien_right_2011}
-  Datadien, A.; Haselager, P.; Sprinkhuizen-Kuyper, I.
-  \newblock {\em The {Right} {Delay} - {Detecting} {Specific} {Spike} {Patterns}
-    with {STDP} and {Axonal} {Conduction} {Delays}.};  2011.
-  \newblock Pages: 99.
-
-  \bibitem[Kerr \em{et~al.}(2013)Kerr, Burkitt, Thomas, Gilson, and
-    Grayden]{kerr_delay_2013}
-  Kerr, R.R.; Burkitt, A.N.; Thomas, D.A.; Gilson, M.; Grayden, D.B.
-  \newblock Delay {Selection} by {Spike}-{Timing}-{Dependent} {Plasticity} in
-    {Recurrent} {Networks} of {Spiking} {Neurons} {Receiving} {Oscillatory}
-    {Inputs}.
-  \newblock {\em PLoS Computational Biology} {\bf 2013}, {\em 9},~e1002897.
-  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1002897}}.
-
-  \bibitem[Burkitt and Hogendoorn(2021)]{burkitt_predictive_2021}
-  Burkitt, A.N.; Hogendoorn, H.
-  \newblock Predictive {Visual} {Motion} {Extrapolation} {Emerges}
-    {Spontaneously} and without {Supervision} at {Each} {Layer} of a
-    {Hierarchical} {Neural} {Network} with {Spike}-{Timing}-{Dependent}
-    {Plasticity}.
-  \newblock {\em The Journal of Neuroscience} {\bf 2021}, {\em 41},~4428--4438.
-  \newblock {\url{https://doi.org/10.1523/jneurosci.2017-20.2021}}.
-
-  \bibitem[Nadafian and Ganjtabesh(2020)]{nadafian_bio-plausible_2020}
-  Nadafian, A.; Ganjtabesh, M.
-  \newblock Bio-plausible {Unsupervised} {Delay} {Learning} for {Extracting}
-    {Temporal} {Features} in {Spiking} {Neural} {Networks},  2020.
-
-  \bibitem[Wang \em{et~al.}(2019)Wang, Lin, and Dang]{wang_delay_2019}
-  Wang, X.; Lin, X.; Dang, X.
-  \newblock A {Delay} {Learning} {Algorithm} {Based} on {Spike} {Train} {Kernels}
-    for {Spiking} {Neurons}.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2019}, {\em 13}.
-
-  \bibitem[Hazan \em{et~al.}(2022)Hazan, Caby, Earl, Siegelmann, and
-    Levin]{hazan_memory_2022}
-  Hazan, H.; Caby, S.; Earl, C.; Siegelmann, H.; Levin, M.
-  \newblock Memory via {Temporal} {Delays} in weightless {Spiking} {Neural}
-    {Network},  2022.
-  \newblock arXiv:2202.07132 [cs, q-bio, stat].
-
-  \bibitem[Luo \em{et~al.}(2022)Luo, Qu, Wang, Yi, Zhang, and
-    Zhang]{luo_supervised_2022}
-  Luo, X.; Qu, H.; Wang, Y.; Yi, Z.; Zhang, J.; Zhang, M.
-  \newblock Supervised {Learning} in {Multilayer} {Spiking} {Neural} {Networks}
-    {With} {Spike} {Temporal} {Error} {Backpropagation}.
-  \newblock {\em IEEE Transactions on Neural Networks and Learning Systems} {\bf
-    2022}, pp. 1--13.
-  \newblock {\url{https://doi.org/10.1109/tnnls.2022.3164930}}.
-
-  \bibitem[Sun \em{et~al.}(2016)Sun, Sourina, and Huang]{sun_learning_2016}
-  Sun, H.; Sourina, O.; Huang, G.B.
-  \newblock Learning polychronous neuronal groups using joint weight-delay
-    spike-timing-dependent plasticity.
-  \newblock {\em Neural Computation} {\bf 2016}, {\em 28},~2181--2212.
-  \newblock tex.eprint:
-    https://direct.mit.edu/neco/article-pdf/28/10/2181/972099/neco{\textbackslash}\_a{\textbackslash}\_00879.pdf,
-    {\url{https://doi.org/10.1162/NECO_a_00879}}.
-
-  \bibitem[Ghosh \em{et~al.}(2019)Ghosh, Gupta, Silva, Soares, and
-    Thakor]{ghosh_spatiotemporal_2019}
-  Ghosh, R.; Gupta, A.; Silva, A.N.; Soares, A.; Thakor, N.V.
-  \newblock Spatiotemporal filtering for event-based action recognition,  2019.
-
-  \bibitem[Perrinet \em{et~al.}(2014)Perrinet, Adams, and
-    Friston]{perrinet_active_2014}
-  Perrinet, L.U.; Adams, R.A.; Friston, K.J.
-  \newblock Active inference, eye movements and oculomotor delays.
-  \newblock {\em Biological Cybernetics} {\bf 2014}, {\em 108},~777--801.
-  \newblock {\url{https://doi.org/10.1007/s00422-014-0620-8}}.
-
-  \bibitem[Hogendoorn and Burkitt(2019)]{hogendoorn_predictive_2019}
-  Hogendoorn, H.; Burkitt, A.N.
-  \newblock Predictive {Coding} with {Neural} {Transmission} {Delays}: {A}
-    {Real}-{Time} {Temporal} {Alignment} {Hypothesis}.
-  \newblock {\em eneuro} {\bf 2019}, {\em 6},~ENEURO.0412--18.2019.
-  \newblock {\url{https://doi.org/10.1523/eneuro.0412-18.2019}}.
-
-  \bibitem[Khoei \em{et~al.}(2013)Khoei, Masson, and
-    Perrinet]{khoei_motion-based_2013}
-  Khoei, M.A.; Masson, G.S.; Perrinet, L.U.
-  \newblock Motion-based prediction explains the role of tracking in motion
-    extrapolation.
-  \newblock {\em Journal of Physiology-Paris} {\bf 2013}, {\em 107},~409--420.
-  \newblock {\url{https://doi.org/10.1016/j.jphysparis.2013.08.001}}.
-
-  \bibitem[Kaplan \em{et~al.}(2013)Kaplan, Lansner, Masson, and
-    Perrinet]{kaplan_anisotropic_2013}
-  Kaplan, B.A.; Lansner, A.; Masson, G.S.; Perrinet, L.U.
-  \newblock Anisotropic connectivity implements motion-based prediction in a
-    spiking neural network.
-  \newblock {\em Frontiers in Computational Neuroscience} {\bf 2013}, {\em 7}.
-  \newblock Loaded from an external bibliography file by Manubot.
-    source\_bibliography: manual-references.bib standard\_id: Kaplan13,
-    {\url{https://doi.org/10.3389/fncom.2013.00112}}.
-
-  \bibitem[Khoei \em{et~al.}(2017)Khoei, Masson, and
-    Perrinet]{khoei_flash-lag_2017}
-  Khoei, M.A.; Masson, G.S.; Perrinet, L.U.
-  \newblock The {Flash}-{Lag} {Effect} as a {Motion}-{Based} {Predictive}
-    {Shift}.
-  \newblock {\em PLOS Computational Biology} {\bf 2017}, {\em 13},~e1005068.
-  \newblock Publisher: Public Library of Science,
-    {\url{https://doi.org/10.1371/journal.pcbi.1005068}}.
-
-  \bibitem[Javanshir \em{et~al.}(2022)Javanshir, Nguyen, Mahmud, and
-    Kouzani]{javanshir_advancements_2022}
-  Javanshir, A.; Nguyen, T.T.; Mahmud, M.A.P.; Kouzani, A.Z.
-  \newblock Advancements in {Algorithms} and {Neuromorphic} {Hardware} for
-    {Spiking} {Neural} {Networks}.
-  \newblock {\em Neural Computation} {\bf 2022}, {\em 34},~1289--1328.
-  \newblock {\url{https://doi.org/10.1162/neco_a_01499}}.
-
-  \bibitem[Marković \em{et~al.}(2020)Marković, Mizrahi, Querlioz, and
-    Grollier]{markovic_physics_2020}
-  Marković, D.; Mizrahi, A.; Querlioz, D.; Grollier, J.
-  \newblock Physics for neuromorphic computing.
-  \newblock {\em Nature Reviews Physics} {\bf 2020}, {\em 2},~499--510.
-  \newblock Number: 9 Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/s42254-020-0208-2}}.
-
-  \bibitem[Rasetto \em{et~al.}(2022)Rasetto, Wan, Akolkar, Shi, Xiong, and
-    Benosman]{rasetto_challenges_2022}
-  Rasetto, M.; Wan, Q.; Akolkar, H.; Shi, B.; Xiong, F.; Benosman, R.
-  \newblock The {Challenges} {Ahead} for {Bio}-inspired {Neuromorphic} {Event}
-    {Processors}: {How} {Memristors} {Dynamic} {Properties} {Could}
-    {Revolutionize} {Machine} {Learning}.
-  \newblock Technical Report 2201.12673, arXiv,  2022.
-
-  \bibitem[Diesmann and Gewaltig(2003)]{diesmann_nest_2003}
-  Diesmann, M.; Gewaltig, M.O.
-  \newblock {NEST}: {An} {Environment} for {Neural} {Systems} {Simulations}.
-  \newblock {\em GWDG-Bericht Nr. 58 Theo Plesser, Volker Macho (Hrsg.)} {\bf
-    2003}, p.~29.
-
-  \bibitem[Hazan \em{et~al.}(2018)Hazan, Saunders, Khan, Patel, Sanghavi,
-    Siegelmann, and Kozma]{hazan_bindsnet_2018}
-  Hazan, H.; Saunders, D.J.; Khan, H.; Patel, D.; Sanghavi, D.T.; Siegelmann,
-    H.T.; Kozma, R.
-  \newblock {BindsNET}: {A} {Machine} {Learning}-{Oriented} {Spiking} {Neural}
-    {Networks} {Library} in {Python}.
-  \newblock {\em Frontiers in Neuroinformatics} {\bf 2018}, {\em 12}.
-  \newblock {\url{https://doi.org/10.3389/fninf.2018.00089}}.
-
-  \bibitem[Stimberg \em{et~al.}(2019)Stimberg, Brette, and
-    Goodman]{stimberg_brian_2019}
-  Stimberg, M.; Brette, R.; Goodman, D.F.
-  \newblock Brian 2, an intuitive and efficient neural simulator.
-  \newblock {\em eLife} {\bf 2019}, {\em 8},~e47314.
-  \newblock {\url{https://doi.org/10.7554/eLife.47314}}.
-
-  \bibitem[Zenke \em{et~al.}(2021)Zenke, Bohté, Clopath, Comşa, Göltz, Maass,
-    Masquelier, Naud, Neftci, Petrovici, Scherr, and
-    Goodman]{zenke_visualizing_2021}
-  Zenke, F.; Bohté, S.M.; Clopath, C.; Comşa, I.M.; Göltz, J.; Maass, W.;
-    Masquelier, T.; Naud, R.; Neftci, E.O.; Petrovici, M.A.;  et~al.
-  \newblock Visualizing a joint future of neuroscience and neuromorphic
-    engineering.
-  \newblock {\em Neuron} {\bf 2021}, {\em 109},~571--575.
-  \newblock {\url{https://doi.org/10.1016/j.neuron.2021.01.009}}.
-
-  \bibitem[Mead and Ismail(1989)]{mead_analog_1989}
-  Mead, C.; Ismail, M.
-  \newblock {\em Analog {VLSI} {Implementation} of {Neural} {Systems}}; Springer
-    Science \& Business Media,  1989.
-  \newblock Google-Books-ID: 9e29dOiXeiMC.
-
-  \bibitem[Bartolozzi and Indiveri(2007)]{bartolozzi_synaptic_2007}
-  Bartolozzi, C.; Indiveri, G.
-  \newblock Synaptic {Dynamics} in {Analog} {VLSI}.
-  \newblock {\em Neural Computation} {\bf 2007}, {\em 19},~2581--2603.
-  \newblock {\url{https://doi.org/10.1162/neco.2007.19.10.2581}}.
-
-  \bibitem[Schuman \em{et~al.}(2017)Schuman, Potok, Patton, Birdwell, Dean, Rose,
-    and Plank]{schuman_survey_2017}
-  Schuman, C.D.; Potok, T.E.; Patton, R.M.; Birdwell, J.D.; Dean, M.E.; Rose,
-    G.S.; Plank, J.S.
-  \newblock A {Survey} of {Neuromorphic} {Computing} and {Neural} {Networks} in
-    {Hardware}.
-  \newblock {\em arXiv:1705.06963 [cs]} {\bf 2017}.
-  \newblock arXiv: 1705.06963.
-
-  \bibitem[Furber \em{et~al.}(2013)Furber, Lester, Plana, Garside, Painkras,
-    Temple, and Brown]{furber_overview_2013}
-  Furber, S.B.; Lester, D.R.; Plana, L.A.; Garside, J.D.; Painkras, E.; Temple,
-    S.; Brown, A.D.
-  \newblock Overview of the {SpiNNaker} {System} {Architecture}.
-  \newblock {\em IEEE Transactions on Computers} {\bf 2013}, {\em
-    62},~2454--2467.
-  \newblock {\url{https://doi.org/10.1109/TC.2012.142}}.
-
-  \bibitem[Furber and Bogdan(2020)]{furber_spinnaker_2020}
-  Furber, S.; Bogdan, P., Eds.
-  \newblock {\em {SpiNNaker}: {A} {Spiking} {Neural} {Network} {Architecture}};
-    Now Publishers,  2020.
-  \newblock {\url{https://doi.org/10.1561/9781680836523}}.
-
-  \bibitem[Merolla \em{et~al.}(2014)Merolla, Arthur, Alvarez-Icaza, Cassidy,
-    Sawada, Akopyan, Jackson, Imam, Guo, Nakamura, Brezzo, Vo, Esser, Appuswamy,
-    Taba, Amir, Flickner, Risk, Manohar, and Modha]{merolla_million_2014}
-  Merolla, P.A.; Arthur, J.V.; Alvarez-Icaza, R.; Cassidy, A.S.; Sawada, J.;
-    Akopyan, F.; Jackson, B.L.; Imam, N.; Guo, C.; Nakamura, Y.;  et~al.
-  \newblock A million spiking-neuron integrated circuit with a scalable
-    communication network and interface.
-  \newblock {\em Science} {\bf 2014}, {\em 345},~668--673.
-  \newblock {\url{https://doi.org/10.1126/science.1254642}}.
-
-  \bibitem[Benjamin \em{et~al.}(2014)Benjamin, Gao, McQuinn, Choudhary,
-    Chandrasekaran, Bussat, Alvarez-Icaza, Arthur, Merolla, and
-    Boahen]{benjamin_neurogrid_2014}
-  Benjamin, B.V.; Gao, P.; McQuinn, E.; Choudhary, S.; Chandrasekaran, A.R.;
-    Bussat, J.M.; Alvarez-Icaza, R.; Arthur, J.V.; Merolla, P.A.; Boahen, K.
-  \newblock Neurogrid: {A} {Mixed}-{Analog}-{Digital} {Multichip} {System} for
-    {Large}-{Scale} {Neural} {Simulations}.
-  \newblock {\em Proceedings of the IEEE} {\bf 2014}, {\em 102},~699--716.
-  \newblock Conference Name: Proceedings of the IEEE,
-    {\url{https://doi.org/10.1109/JPROC.2014.2313565}}.
-
-  \bibitem[Neckar \em{et~al.}(2019)Neckar, Fok, Benjamin, Stewart, Oza, Voelker,
-    Eliasmith, Manohar, and Boahen]{neckar_braindrop_2019}
-  Neckar, A.; Fok, S.; Benjamin, B.V.; Stewart, T.C.; Oza, N.N.; Voelker, A.R.;
-    Eliasmith, C.; Manohar, R.; Boahen, K.
-  \newblock Braindrop: {A} mixed-signal neuromorphic architecture with a
-    dynamical systems-based programming model.
-  \newblock {\em Proceedings of the IEEE} {\bf 2019}, {\em 107},~144--164.
-  \newblock {\url{https://doi.org/10.1109/JPROC.2018.2881432}}.
-
-  \bibitem[Schemmel \em{et~al.}(2010)Schemmel, Brüderle, Grübl, Hock, Meier,
-    and Millner]{schemmel_wafer-scale_2010}
-  Schemmel, J.; Brüderle, D.; Grübl, A.; Hock, M.; Meier, K.; Millner, S.
-  \newblock A wafer-scale neuromorphic hardware system for large-scale neural
-    modeling.
-  \newblock In Proceedings of the 2010 {IEEE} {International} {Symposium} on
-    {Circuits} and {Systems} ({ISCAS}),  2010, pp. 1947--1950.
-  \newblock ISSN: 2158-1525, {\url{https://doi.org/10.1109/ISCAS.2010.5536970}}.
-
-  \bibitem[Markram \em{et~al.}(2011)Markram, Meier, Lippert, Grillner,
-    Frackowiak, Dehaene, Knoll, Sompolinsky, Verstreken, DeFelipe, Grant,
-    Changeux, and Saria]{markram_introducing_2011}
-  Markram, H.; Meier, K.; Lippert, T.; Grillner, S.; Frackowiak, R.; Dehaene, S.;
-    Knoll, A.; Sompolinsky, H.; Verstreken, K.; DeFelipe, J.;  et~al.
-  \newblock Introducing the {Human} {Brain} {Project}.
-  \newblock {\em Procedia Computer Science} {\bf 2011}, {\em 7},~39--42.
-  \newblock {\url{https://doi.org/10.1016/j.procs.2011.12.015}}.
-
-  \bibitem[Farquhar \em{et~al.}(2006)Farquhar, Gordon, and
-    Hasler]{farquhar_field_2006}
-  Farquhar, E.; Gordon, C.; Hasler, P.
-  \newblock A {Field} {Programmable} {Neural} {Array}.
-  \newblock In Proceedings of the 2006 {IEEE} {International} {Symposium} on
-    {Circuits} and {Systems}; IEEE: Island of Kos, Greece,  2006; pp. 4114--4117.
-  \newblock {\url{https://doi.org/10.1109/ISCAS.2006.1693534}}.
-
-  \bibitem[Liu \em{et~al.}(2009)Liu, Yu, and Wang]{cheng_fpaa_2009}
-  Liu, M.; Yu, H.; Wang, W.
-  \newblock {FPAA} {Based} on {Integration} of {CMOS} and {Nanojunction}
-    {Devices} for {Neuromorphic} {Applications}. In {\em Nano-{Net}}; Cheng, M.,
-    Ed.; Springer Berlin Heidelberg: Berlin, Heidelberg,  2009; Vol.~3, pp.
-    44--48.
-
-  \bibitem[Chan \em{et~al.}(2007)Chan, Liu, and van Schaik]{chan_aer_2007}
-  Chan, V.; Liu, S.C.; van Schaik, A.
-  \newblock {AER} {EAR}: {A} {Matched} {Silicon} {Cochlea} {Pair} {With}
-    {Address} {Event} {Representation} {Interface}.
-  \newblock {\em IEEE Transactions on Circuits and Systems I: Regular Papers}
-    {\bf 2007}, {\em 54},~48--59.
-  \newblock Conference Name: IEEE Transactions on Circuits and Systems I: Regular
-    Papers, {\url{https://doi.org/10.1109/TCSI.2006.887979}}.
-
-  \bibitem[Haessig \em{et~al.}(2020)Haessig, Milde, Aceituno, Oubari, Knight, van
-    Schaik, Benosman, and Indiveri]{haessig_event-based_2020}
-  Haessig, G.; Milde, M.B.; Aceituno, P.V.; Oubari, O.; Knight, J.C.; van Schaik,
-    A.; Benosman, R.B.; Indiveri, G.
-  \newblock Event-{Based} {Computation} for {Touch} {Localization} {Based} on
-    {Precise} {Spike} {Timing}.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2020}, {\em 14},~420.
-  \newblock {\url{https://doi.org/10/ghvnxj}}.
-
-  \bibitem[Lagorce \em{et~al.}(2017)Lagorce, Orchard, Galluppi, Shi, and
-    Benosman]{lagorce_hots_2017}
-  Lagorce, X.; Orchard, G.; Galluppi, F.; Shi, B.E.; Benosman, R.B.
-  \newblock {HOTS}: {A} {Hierarchy} of {Event}-{Based} {Time}-{Surfaces} for
-    {Pattern} {Recognition}.
-  \newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}
-    {\bf 2017}, {\em 39},~1346--1359.
-  \newblock {\url{https://doi.org/10.1109/TPAMI.2016.2574707}}.
-
-  \bibitem[Sironi \em{et~al.}(2018)Sironi, Brambilla, Bourdis, Lagorce, and
-    Benosman]{sironi_hats_2018}
-  Sironi, A.; Brambilla, M.; Bourdis, N.; Lagorce, X.; Benosman, R.
-  \newblock {HATS}: {Histograms} of {Averaged} {Time} {Surfaces} for {Robust}
-    {Event}-{Based} {Object} {Classification}.
-  \newblock In Proceedings of the 2018 {IEEE}/{CVF} {Conference} on {Computer}
-    {Vision} and {Pattern} {Recognition}; IEEE: Salt Lake City, UT, USA,  2018;
-    pp. 1731--1740.
-  \newblock {\url{https://doi.org/10.1109/CVPR.2018.00186}}.
-
-  \bibitem[Maro \em{et~al.}(2020)Maro, Ieng, and Benosman]{maro_event-based_2020}
-  Maro, J.M.; Ieng, S.H.; Benosman, R.
-  \newblock Event-{Based} {Gesture} {Recognition} {With} {Dynamic} {Background}
-    {Suppression} {Using} {Smartphone} {Computational} {Capabilities}.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2020}, {\em 14}.
-  \newblock Publisher: Frontiers,
-    {\url{https://doi.org/10.3389/fnins.2020.00275}}.
-
-  \bibitem[Grimaldi \em{et~al.}(2021)Grimaldi, Boutin, Perrinet, Ieng, and
-    Benosman]{grimaldi_homeostatic_2021}
-  Grimaldi, A.; Boutin, V.; Perrinet, L.; Ieng, S.H.; Benosman, R.
-  \newblock A homeostatic gain control mechanism to improve event-driven object
-    recognition.
-  \newblock In Proceedings of the 2021 {International} {Conference} on
-    {Content}-{Based} {Multimedia} {Indexing} ({CBMI}). IEEE,  2021.
-  \newblock {\url{https://doi.org/10.1109/cbmi50038.2021.9461901}}.
-
-  \bibitem[Grimaldi \em{et~al.}(2022)Grimaldi, Boutin, Ieng, Benosman, and
-    Perrinet]{grimaldi_robust_2022}
-  Grimaldi, A.; Boutin, V.; Ieng, S.H.; Benosman, R.; Perrinet, L.U.
-  \newblock A robust event-driven approach to always-on object recognition.
-  \newblock {\em TechRxiv preprint} {\bf 2022}.
-  \newblock {\url{https://doi.org/10.36227/techrxiv.18003077.v1}}.
-
-  \bibitem[Yu \em{et~al.}(2022)Yu, Gu, Li, Wang, Wang, and Li]{yu_stsc-snn_2022}
-  Yu, C.; Gu, Z.; Li, D.; Wang, G.; Wang, A.; Li, E.
-  \newblock {STSC}-{SNN}: {Spatio}-{Temporal} {Synaptic} {Connection} with
-    {Temporal} {Convolution} and {Attention} for {Spiking} {Neural} {Networks},
-    2022.
-  \newblock arXiv:2210.05241 [cs, q-bio, stat].
-
-  \bibitem[Benosman \em{et~al.}(2014)Benosman, Clercq, Lagorce, {Sio-Hoi Ieng},
-    and Bartolozzi]{benosman_event-based_2014}
-  Benosman, R.; Clercq, C.; Lagorce, X.; {Sio-Hoi Ieng}.; Bartolozzi, C.
-  \newblock Event-{Based} {Visual} {Flow}.
-  \newblock {\em IEEE Transactions on Neural Networks and Learning Systems} {\bf
-    2014}, {\em 25},~407--417.
-  \newblock {\url{https://doi.org/10.1109/tnnls.2013.2273537}}.
-
-  \bibitem[Clady \em{et~al.}(2014)Clady, Clercq, Ieng, Houseini, Randazzo,
-    Natale, Bartolozzi, and Benosman]{clady_asynchronous_2014}
-  Clady, X.; Clercq, C.; Ieng, S.H.; Houseini, F.; Randazzo, M.; Natale, L.;
-    Bartolozzi, C.; Benosman, R.B.
-  \newblock Asynchronous visual event-based time-to-contact.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2014}, {\em 8}.
-  \newblock 00000 Publisher: Frontiers,
-    {\url{https://doi.org/10.3389/fnins.2014.00009}}.
-
-  \bibitem[Tschechne \em{et~al.}(2014)Tschechne, Sailer, and
-    Neumann]{tschechne_bio-inspired_2014}
-  Tschechne, S.; Sailer, R.; Neumann, H.
-  \newblock Bio-{Inspired} {Optic} {Flow} from {Event}-{Based} {Neuromorphic}
-    {Sensor} {Input}.
-  \newblock In Proceedings of the Artificial {Neural} {Networks} in {Pattern}
-    {Recognition}; El~Gayar, N.; Schwenker, F.; Suen, C., Eds.; Springer
-    International Publishing: Cham,  2014; Lecture {Notes} in {Computer}
-    {Science}, pp. 171--182.
-  \newblock {\url{https://doi.org/10.1007/978-3-319-11656-3_16}}.
-
-  \bibitem[Hidalgo-Carrió \em{et~al.}(2020)Hidalgo-Carrió, Gehrig, and
-    Scaramuzza]{hidalgo-carrio_learning_2020}
-  Hidalgo-Carrió, J.; Gehrig, D.; Scaramuzza, D.
-  \newblock Learning {Monocular} {Dense} {Depth} from {Events}.
-  \newblock {\em arXiv:2010.08350 [cs]} {\bf 2020}.
-  \newblock arXiv: 2010.08350.
-
-  \bibitem[Dardelet \em{et~al.}(2021)Dardelet, Benosman, and
-    Ieng]{dardelet_event-by-event_2021}
-  Dardelet, L.; Benosman, R.; Ieng, S.H.
-  \newblock An {Event}-by-{Event} {Feature} {Detection} and {Tracking}
-    {Invariant} to {Motion} {Direction} and {Velocity} {\bf 2021}.
-  \newblock {\url{https://doi.org/10.36227/techrxiv.17013824.v1}}.
-
-  \bibitem[Stoffregen \em{et~al.}(2019)Stoffregen, Gallego, Drummond, Kleeman,
-    and Scaramuzza]{stoffregen_event-based_2019}
-  Stoffregen, T.; Gallego, G.; Drummond, T.; Kleeman, L.; Scaramuzza, D.
-  \newblock Event-{Based} {Motion} {Segmentation} by {Motion} {Compensation}.
-  \newblock In Proceedings of the 2019 {IEEE}/{CVF} {International} {Conference}
-    on {Computer} {Vision} ({ICCV}),  2019, pp. 7243--7252.
-  \newblock ISSN: 2380-7504, {\url{https://doi.org/10.1109/ICCV.2019.00734}}.
-
-  \bibitem[Kim \em{et~al.}(2016)Kim, Leutenegger, and
-    Davison]{kim_real-time_2016}
-  Kim, H.; Leutenegger, S.; Davison, A.J.
-  \newblock Real-{Time} {3D} {Reconstruction} and 6-{DoF} {Tracking} with an
-    {Event} {Camera}.
-  \newblock In Proceedings of the Computer {Vision} – {ECCV} 2016; Leibe, B.;
-    Matas, J.; Sebe, N.; Welling, M., Eds.; Springer International Publishing:
-    Cham,  2016; Lecture {Notes} in {Computer} {Science}, pp. 349--364.
-  \newblock {\url{https://doi.org/10.1007/978-3-319-46466-4_21}}.
-
-  \bibitem[Hussain \em{et~al.}(2012)Hussain, Basu, Wang, and
-    Hamilton]{hussain_deltron_2012}
-  Hussain, S.; Basu, A.; Wang, M.; Hamilton, T.J.
-  \newblock {DELTRON}: {Neuromorphic} architectures for delay based learning.
-  \newblock In Proceedings of the 2012 {IEEE} {Asia} {Pacific} {Conference} on
-    {Circuits} and {Systems},  2012, pp. 304--307.
-  \newblock {\url{https://doi.org/10.1109/APCCAS.2012.6419032}}.
-
-  \bibitem[Wang \em{et~al.}(2015)Wang, Hamilton, Tapson, and van
-    Schaik]{wang_neuromorphic_2015}
-  Wang, R.M.; Hamilton, T.J.; Tapson, J.C.; van Schaik, A.
-  \newblock A neuromorphic implementation of multiple spike-timing synaptic
-    plasticity rules for large-scale neural networks.
-  \newblock {\em Frontiers in Neuroscience} {\bf 2015}, {\em 9}.
-  \newblock {\url{https://doi.org/10.3389/fnins.2015.00180}}.
-
-  \bibitem[Wang \em{et~al.}(2014)Wang, Hamilton, Tapson, and van
-    Schaik]{wang_fpga_2014}
-  Wang, R.; Hamilton, T.J.; Tapson, J.; van Schaik, A.
-  \newblock An {FPGA} design framework for large-scale spiking neural networks.
-  \newblock In Proceedings of the 2014 {IEEE} {International} {Symposium} on
-    {Circuits} and {Systems} ({ISCAS}),  2014, pp. 457--460.
-  \newblock ISSN: 2158-1525, {\url{https://doi.org/10.1109/ISCAS.2014.6865169}}.
-
-  \bibitem[Pfeil \em{et~al.}(2013)Pfeil, Scherzer, Schemmel, and
-    Meier]{pfeil_neuromorphic_2013}
-  Pfeil, T.; Scherzer, A.C.; Schemmel, J.; Meier, K.
-  \newblock Neuromorphic learning towards nano second precision.
-  \newblock In Proceedings of the The 2013 {International} {Joint} {Conference}
-    on {Neural} {Networks} ({IJCNN}),  2013, pp. 1--5.
-  \newblock ISSN: 2161-4407, {\url{https://doi.org/10.1109/IJCNN.2013.6706828}}.
-
-  \bibitem[Boerlin and Denève(2011)]{boerlin_spike-based_2011}
-  Boerlin, M.; Denève, S.
-  \newblock Spike-{Based} {Population} {Coding} and {Working} {Memory}.
-  \newblock {\em PLOS Computational Biology} {\bf 2011}, {\em 7},~e1001080.
-  \newblock {\url{https://doi.org/10.1371/journal.pcbi.1001080}}.
-
-  \bibitem[Renner \em{et~al.}(2022)Renner, Sandamirskaya, Sommer, and
-    Frady]{renner_sparse_2022}
-  Renner, A.; Sandamirskaya, Y.; Sommer, F.T.; Frady, E.P.
-  \newblock Sparse {Vector} {Binding} on {Spiking} {Neuromorphic} {Hardware}
-    {Using} {Synaptic} {Delays}.
-  \newblock {\em Proceedings of the International Conference on Neuromorphic
-    Systems} {\bf 2022}.
-  \newblock Conference Name: ICONS 2022: International Conference on Neuromorphic
-    Systems 2022 Meeting Name: ICONS 2022: International Conference on
-    Neuromorphic Systems 2022 Place: Knoxville, TN, USA Publisher: ACM Digital
-    library, {\url{https://doi.org/10.1145/3546790.3546820}}.
-
-  \bibitem[Dard \em{et~al.}(2022)Dard, Leprince, Denis, Rao~Balappa, Suchkov,
-    Boyce, Lopez, Giorgi-Kurz, Szwagier, Dumont, Rouault, Minlebaev, Baude,
-    Cossart, and Picardo]{dard_rapid_2022}
-  Dard, R.F.; Leprince, E.; Denis, J.; Rao~Balappa, S.; Suchkov, D.; Boyce, R.;
-    Lopez, C.; Giorgi-Kurz, M.; Szwagier, T.; Dumont, T.;  et~al.
-  \newblock The rapid developmental rise of somatic inhibition disengages
-    hippocampal dynamics from self-motion.
-  \newblock {\em eLife} {\bf 2022}, {\em 11},~e78116.
-  \newblock {\url{https://doi.org/10.7554/eLife.78116}}.
-
-  \bibitem[Coull and Giersch(2022)]{coull_distinction_2022}
-  Coull, J.T.; Giersch, A.
-  \newblock The distinction between temporal order and duration processing, and
-    implications for schizophrenia.
-  \newblock {\em Nature Reviews Psychology} {\bf 2022}, {\em 1},~257--271.
-  \newblock Publisher: Nature Publishing Group,
-    {\url{https://doi.org/10.1038/s44159-022-00038-y}}.
-
-  \bibitem[Panahi \em{et~al.}(2021)Panahi, Abrevaya, Gagnon-Audet, Voleti, Rish,
-    and Dumas]{panahi_generative_2021}
-  Panahi, M.R.; Abrevaya, G.; Gagnon-Audet, J.C.; Voleti, V.; Rish, I.; Dumas, G.
-  \newblock Generative {Models} of {Brain} {Dynamics} -- {A} review.
-  \newblock Technical Report 2112.12147, arXiv,  2021.
-
-  \bibitem[Tolle \em{et~al.}(2011)Tolle, Tansley, and Hey]{tolle_fourth_2011}
-  Tolle, K.M.; Tansley, D.S.W.; Hey, A.J.G.
-  \newblock The {Fourth} {Paradigm}: {Data}-{Intensive} {Scientific} {Discovery}
-    [{Point} of {View}].
-  \newblock {\em Proceedings of the IEEE} {\bf 2011}, {\em 99},~1334--1337.
-  \newblock {\url{https://doi.org/10.1109/jproc.2011.2155130}}.
-
-  \end{thebibliography}
+\bibliography{references}
 
